### 2018-06-21 Stickleback_Paper.tex
First Commit

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{comment}
\usepackage[backgroundcolor=pink,linecolor=red]{todonotes}

%\singlespacing
\onehalfspacing
%\doublespacing
%\setstretch{1.1}

\begin{document}

\linespread{1.5}

\section{Abstract}

Note: Finish after the whole paper has been written

%wider audience by more general questions about ... introduction maybe too fishy???

%a question about the genetic basis for adaptstion, how is it maintianed?
% theres alot of debate in the field

% more specifically - sticklback fish

% we want to examine this, This is why our approach is unique
% This is why the field could benefit from this.

%methods - we did this this and this specifically
%results - we found this and this specifically

% discussion get more broad again.

\section{Introduction}

In the mid 19th century,
Darwin theorized that the process of natural selection resulted in speciation, 
meaning all living species derived from a common ancestor. 
He also believed that this process took place on ecological timescales that would not allow for 
this process to be observed in a single lifetime. 
Today, we have found evidence of evolution
in tens of generations and are starting to grasp the complexity behind such a simple idea. 
Famously, Drs. Peter and Rosemary Grant went on to study Darwin's finches and
found that phenotypes could adapt to changes in the food source within mere generations.
%more specific?
We have have found many examples of rapid adaptation in many commonly studied models
but many questions still remain surrounding the genomics which allow for this.

Recently, rapid and parallel adaptation has brought about question of selection of
standing genetic variation in natural populations 
of Stickleback.
In 1964, The Great Alaskan Earthquake caused Middleton island to raise up and in turn, 
introduced a group of new freshwater ponds around the perimeter of the island. 
Quickly inhabited by the surrounding marine population of Ninespine Stickleback,
we've observed significant phenotypic changes in less than 50 years that appear to be 
parallel to freshwater stickleback that have been separated for over $13,000$ years. \cite{LescakE7204} 
In freshwater stickleback, 
the number of lateral plates are reduced
and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region (Cite Kristin, Cresko et. al)

The leading hypothesis is that rather than acting on new mutations, ecological speciation is sped up through selection on standing genetic variation found in marine populations. 
One clear example of this is the gene \textit{eda} 
which has been shown to regulate the number of lateral plates. 
While this gene arose millions of years ago, 
it is found in freshwater ponds which have formed much more recently.
Novel evidence from natural populations has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern. \cite{Nelson2017}

In 2009, Dolph Schluter and Gina L. Conte suggested the "transportation"-hypothesis. 
This outlined the flow of freshwater alleles into marine populations through offspring of hybridization events.
It suggests freshwater haplotypes are distributed though marine individuals and are continually selected upon in introduced freshwater populations.
The continued selection on freshwater favored alleles,
and introgression between the two sub-species,
allows the marine to maintain the freshwater haplotype dispersed in low frequency among its' individuals. 
Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection rebuilds the freshwater haplotype. \cite{Schluter2009} 
The alternative to this hypothesis is that individuals from other freshwater environments migrate directly to the new environments, and their haplotype is passed down directly. 
This hypothesis has been shown to be unlikely due to 
finding a high frequency of freshwater alleles in the ocean,
and almost no freshwater individuals.

Assuming they were correct about transportation, we might ask; 
how much migration and gene flow is to be expected for this relationship to be maintained?
(Other questions)
how does selection, at the genomic level, alter given a range of migration and recombination rates?
Here, we use forward moving simulations to model marine and freshwater populations of stickleback.
We then observe the effect (focusing on migration and recombination, for now) 
of different parameters on local adaptation of introduced freshwater environments. 
We compare to real data and make predictions about realistic parameters.

Simulations are a powerful tool when there is some map between the parameters we want to learn about
and the genetic data we observe in nature. 
We don't know what that map is, so we invert the relationship. 
This means going from genomic data that we see to the parameter values that must be driving the system, 
assuming the model is correct.

\section{Methods}

Our models are forward-moving evolutionary models which
emulate the geography, selective pressure, and genomic architecture 
of coastal marine and freshwater lake populations. 
Here we give a brief overview of the software used and the models created to observe the effects of selection.

\todo{Each set of simulations was run multiple times to ensure the data we observed was not stochastic.}

Note: Maybe talk about some things are stochastic: 
(1) Correlation of mutation frequencies at low migration rates
(2) Distribution of Effect Size? (run another set to make sure)

\subsection{SLiM}

To model the stickleback populations, we used a flexible evolutionary framework, SLiM. (cite SLiM 2) 
The model-type we used was based off of an extended wright-fisher. 
The simulation life cycle of each generation is as follows.
First, 
generate offspring by:
(1) choosing parents based on cached fitness values and migration rates, 
(2) performing recombination of parental genomes, 
(3) allow for mutation, 
(4) modify each child by a defined callback. 
Next, 
offspring become individuals before fitness is evaluated for the next generation.
Finally,
the generation counter is incremented. 
The population structure in SLiM can be arranged in any number of subpopulations, 
continuous or discrete, connected by any rate of migration.

\subsection{Model}

\subsubsection{Geography}

What we've modeled geographically is a coastal marine population 
connected by migration to multiple, smaller freshwater populations that are located along the coast.
The population structure for the marine is modeled as a one-dimensional, continuous population
ranging from 0.0 to 10.0.
There are two sets of lakes which represent freshwater populations; 
the first which evolves in parallel to the marine, 
and a set introduced later in the simulation as a subset of the marine. 
In SLiM, these freshwater sets are defined as just two continuous subpopulations. 
Using SLiM?s $modifyChild()$ callback, we?ve modeled both to act as 10 separate discrete subpopulations. 
Each of the $10$ lakes, $i$, is located along the marine at $i - 0.5$, 
and connected by migration only through the marine environment. 
The marine has $2000$ individuals while 
each of the lakes fluctuate around $200$.

\todo{Should I put my geography picture here?}

\subsubsection{Selective Pressure}

To emulate the freshwater and marine selective pressure, 
we set up a quantitative genetics model in which fitness is phenotypically based. 
Freshwater and Marine environments are distinguished by a single numeric value. 
This value is the optimum phenotype for each environment
and acts as a representation of lateral plate number and opercle shape in the stickleback populations.
The fitness of an individual is then determined by a probability density for a normal distribution 
at the quantile (?) of the difference between the optimum and the individual's phenotype

    \[
\textnormal{Individual.Fitness = } X \sim \mathcal{N}(\textnormal{Optimum - Individual.Phenotype},\,0.0\,15.0).
    \]

Where an individual?s phenotype is also defined by a single numeric value; 
determined as a summation of the selection coefficients of all mutations an individual possesses.

	\[
{\textnormal{Individual.Phenotype = } \sum_{m \in mutations} \textnormal{m.selectionCoeff}}
	\]

\todo{This isn't quite right because of how we calculate additive, recessive, and dominant}

We allow six mutation types which affect phenotype along with one neutral mutation type.
Among the effect mutations; 
there are additive, dominant, and recessive mutations which effect phenotype
in either the positive or negative direction. 
The selection coefficients are pulled from a beta distribution 
with a shape parameter of $1$ and a mean of either $0.5$, or ${-0.5}$.

	\[
\textnormal{Mutation.selectionCoeff = } X \sim \Gamma (1,\pm0.5)
	\]

\subsubsection{Genomic Architecture}

Although still uncertain about how much of the genome is directly associated with the distinguished phenotypes, 
GWAS has indicated clusters of loci (linkage groups? Operon?) along the genome to be causative. 
To mimic this and compare dynamics of neutral vs. effect under selection, 
we create regions of effect. These genomic elements are the only loci which allow for mutation that impact phenotype. 
This allows us to observe differences in selection at the genomic level. 
In SLiM, the genome is conceptually a linear array of loci for which we can define
different amount 
\todo{finish this}

\subsection{Sampling}

%how did we use slim 
%More question based, Why did we measure what we measued 
%are there ways to have two or three subheaders 
%decriptive statist. 
%genetic basis -- 
%process of plotting. 
%if it's something new

Here, we're interested in observing how different parameter values
(mainly migrations rates and recombination rates)
impact the sharing of alleles between populations. 
we define freshwater adapted alleles (FAA),
at any given generation,
to be a mutations with a frequency higher than $0.5$ in \textit{any} of the original lakes, 
while remaining lower than $0.5$ in the marine. 
This is because the transportation hypothesis does not specify where or when an advantageous mutation arises; 
but simply suggests that when one comes to high enough frequency, it too, could participate in the transportation process. \cite{Schluter2009}
To observe transportation of freshwater alleles in our simulations, we use a variety of metrics when sampling.
Throughout the simulation, we measure;
mutation frequencies, 
$F_{st}$ values, 
average phenotype, 
mean number of original lakes each FAA appears in,
and total number of unique FAA. 
to understand where, as well as how much, of these freshwater haplotypes are being transported,
we track the mean percentage of FAA per individual (MPFAI) across all populations throughout the simulation:
computed as the sum of the FAA mutation frequencies among each population.
we're particularly interested in using MPFAI to measure the marine population' capacity to carry FAA,
as well as observing the introduced populations' ability to select upon them.

Because we are interested in the local adaptation of the introduced freshwater populations,
we define time to adaptation ($T_{adapt}$) of the introduced population to be the generation at which
the difference between the average phenotype of the original and the introduced freshwater populations is less than 0.5. 
At this generation, we look at:
the correlation (B. Pearson) between the mutation frequencies of all effect loci and FAA,
and the number of shared high frequency alleles. 
These metrics tells us how much of the FAA  were used in the lakes once they have locally adapted.

\section{Results}

%We found this
%--justify it

% A LIST OF PLOTS
% GENERAL OBSERVATIONS

% [1]. Fst Neutral 
% [2]. Fst Effect 
% [3]. Fst Across The 
% [4]. Median Number of FAA Per individual / Total Number of FAA
% [5]. Phenotype Distribution
% [6]. Percentage of shared High Frequency Alleles Shared Between Original and Introduced
% [7]. Correlation of Allele Frequency 
% [8]. Time until Adaptation
% [9]. Total Num FWAA - Average number of lakes each allele appears in.
% [10]. Distribution of # of FWAA Per marine individual.

\subsection{Local Adaptation: differentiation with gene flow}

Local adaptation occurred across all simulated parameter values.
Starting from the same baseline freshwater and marine populations diverged phenotypically, 
until they reached an equilibrium,
where the population means were close to or at the optimal phenotype. 
phenotypic variation within each population was small compared to the difference between populations.
Polymorphic loci that effected the phenotype,
showed greater differentiation compared to neutral loci between marine and freshwater populations, with the exception 
of migration rate at a value of $5e-5$. 
$F_{st}$ plots along the genome, show peaks at loci underlying local adaptation.

\todo{ add more general observations and caveats about not all parameter values. }

\subsection{The Effect of Migration Rate}

\subsubsection{Migration is Mixing}

%FIGURES THAT SHOW THIS:
%- Fst Plots.
%- SGV Plots.
%- Haplotype plots

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=01.0\linewidth]{PhenotypeDist}
  		\caption{A boat.}
  		\label{fig:boat1}
	\end{center}
\end{figure}

Across increasing parameters of migration, we observed more gene flow across the populations. 
As can be seen in Figure (neutral fst figure)
$F_{st}$ values for neutral Alleles steadily decline between all sets of populations as 
migration increases, this illustrates less difference between sections of the genome which are not acted up 
upon by selection, in all populations.
Additionally, Standing Genetic Variance \todo{computed as?}
(Seen in SGV Figure)
values in the marine population steady increased with migration which suggests that more alleles from the 
lakes surrounding were exporting alleles into the marine individuals.
\todo{Something about SGV rates for the the other population}

%TIES TO NATURE?
%SPECIFICALLY (interesting stuff)

\subsubsection{migration affects the speed of adaptation}

%FIGURES THAT SHOW THIS:
%-Time to adaptation

We observed a dramatic shift in time until adaptation for the introduced populations
as migration rates increased. 
As can be seen in Figure (Time Until Adaptation) at the lowest migration rate of $5 * 10^{-5}$,
It took the introduced population over 30 thousand generations for the average phenotype of all the lakes to 
get to within 0.5 of the original lakes original phenotypes. 
This suggests that selection was acting primarily on new mutations, meaning the 
introduced lakes needed to wait for a beneficial mutation to arise before 
it was selected upon. 
There was a positive correlation between the amount of standing genetic variation in the 
marine and the rapid adaptation of the introduced populations.

%TIES TO NATURE?

\subsubsection{sharing of freshwater adapted alleles}

%FIGURES THAT SHOW THIS:
%-MPAA / IND
%-Total/Avg shared

To further investigate this correlation, we take a look at the FAA driving 
local adaptation of the introduced population. 
One common metric we investigate is the percentage of FAA per individual in each of the populations. 
This gives us a general look at the where these alleles are being distributed. 
Looking at the lowest migration rate for the original lakes population In Figure (MPAA / IND) 
We see that the average individual has almost exactly $1/10^{th}$ of the total defined FWAA throughout the simulation. 
Because FAA are defined among \textit{any} population, this suggests that each one of the $10$ 
lakes has created their own solution to adaptation of the the freshwater selective pressure.

As we can see in Figure (Total/Avg shared)
as migration rate parameter increases for the simulations, 
we see the distribution of the total number of FAA's decrease, and 
the average number of lakes each allele appears in at high frequency ($p > 0.5$), increase.
With the distribution of effect size remaining the same across all simulations, 
these are both suggestive of the original lakes sharing solutions to the same selective pressure.

%TIES TO NATURE?

%SPECIFICALLY (interesting stuff)

\subsubsection{We see Migration Load after a certain threshold of migration}

%FIGURES THAT SHOW THIS:
%- Phenotype Distribution

%GENERAL

In the distributions we have shown across migration rate parameter values, 
We have experienced the most dramatic shifts of the population dynamics at $M = 5x10^{-3}$.
After a threshold between this and $M = 5x10^{-2}$, we start to experience migration load. 
Significant gene flow constricts local adaptation
as a consequence of a large number of offspring through hybridization events between subpopulations.
In Figure (Phenotype distribution) at $M = 5x10^{-2}$, we can see the distribution of average phenotype throughout the simulation
pull towards the opposing selective pressure value and away from the local optimum in all subpopulations.

%TIES TO NATURE?

%SPECIFICALLY (interesting stuff)

\subsubsection {There is a Qualitative Threshold of migration rates}

%FIGURES THAT SHOW THIS:
%all
%TIES TO NATURE?
%SPECIFICALLY (interesting stuff)

%GENERAL
We have found that too little migration leads to selection upon new mutations in all subpopulations and lakes alike. 
In contrast, at high migration rates we have seen that migration load limits 
the ability for species to locally adapt to the selective pressure of their environment.
This leads us to consider a window (Goldilocks Zone) of introgression which allows for the transportation
of FAA's without migration load.

\subsubsection {Low Recombination Causes Clustering}

\begin{comment}
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
Across migration rates we observe a large difference in:
gene flow, 
$F_{st}$ values,
amount of time until adaptation,
sharing of solutions (haplotytes?),
$SGV$,
and, migration load. 
we run multiple simulations across a range of migration rates starting at $5e-5$ and increasing by
one order of magnitude up to $.05$. 
With the marine and freshwater both at a population size 2000,
this represents exactly $1/10_{th}$, $1$, $10$, and $100$ migrants per generation, respectively.

At low migration rates of $5e-5$ (1/10 migrants per generation), we observed little sharing of genetic solutions and 
almost no freshwater alleles existing in the marine population.
With an MPFAI of $0.0999$ in the original lakes;
each individual in the original lakes had almost exactly $1/10_{th}$ of the total FAA. 
In addition to each allele appearing in an average of $1.08$ lakes, 
this tells us that each of the 10 lakes has individually selected their own set of solutions (haplotypes) 
rather than sharing alleles through the marine. 
This is further supported by the low capacity of the marine to transport the alleles with an MPFAI of $0.003$.
$T_{adapt}$ remained on an ecological timescale of $18K$ generations, and the correlation of FAA frequencies was at $0.09$.
With the introduced population having an MPFAI of $0.02$, 
we see little evidence that the lakes were able the select on the alleles found in that marine. 
it may be important to note that while we are sampling the 10 introduced lakes as a whole population,
It is possible that only a couple of the lakes even received a copy some alleles during the introduction.
However, to see evidence of the haplotypes being carried throughout the marine we would expect to observe a pattern off all lakes
being able to select upon similar alleles. 
This observation gives a migration rate below the threshold needed to support the transportation hypothesis.

scaling up migration rates to $0.0005$ (1 migrants per generation)
we observe the first evidence of transportation.
The mean total number of FAA drops to 45 and the average number of lakes which each alleles appears in jumped to $1.33$. 
This shows us the gene flow between populations allowed for some freshwater alleles to travel through the marine and be shared by the lakes. 
Interestingly, with the marine showing an MPFAI of $0.02$, 
the introduced was much more able to select upon the alleles;
The introduced population showed a MPFAI of $0.078$, compared to the original lakes which was at $0.129$,
Looking at the FAA frequencies in the two freshwater populations at time of adaptation, we see a correlation of $.68$
While this is evidence of parallel local adaptation, the  $T_{adapt}$ was still at $8,365$ generations.

At $0.005$ (10 migrants per generation) we experience high correlation between the FAA frequencies
in the original and introduced populations, $0.96$, and rapid local adaptation of the introduced population, (81 generations). 
With the introduced population showing an MPFAI of  ~ $0.34$ which is strikingly close to the original freshwater population which is at $0.34$
This migration rate clearly shows the capacity to share alleles in order for us to see rapid and parallel adaptation.

Interestingly, is that if we look at the distribution and total amount of FAA in the marine population at time of introduction between 
$0.0005$ and $0.005$, they are virtually identical.
This comes with two questions
(1) why is there a similar amount of FAA in the marine with such a substantial difference in migration, and
(2) what is it about the higher migration rate of  $0.005$ that allows it perform so much better in terms of rapid adaptation? 
To answer the first question we can see that the median number of defined FAA is much higher, 45
While with the higher migration there is only a median of 10. 
This is a result of the solutions not being shared across the original lakes at lower migration rates.
This can be seen by the average number of lakes each FAA appears in being 
$1.54$ at 1 migrant per generation and 
$3.75$ at 10 migrants per generation.
consequently more alleles were defined as FAA and sampled as such in the marine population. 
This however does not explain why, at higher migration rates, 
the introduced freshwater population was able to rebuild the haplotype so much more quickly. 
For this, It is important to note that after introduction, there is continued migration. 
This means that for a migrate of $0.005$ there is roughly 10 times as 
much of the FAA being brought into the freshwater every generation compared to migrate $0.0005$.

This seams to suggest that continued migration is quite important 
for rapid adaptation at this migration rate as opposed to the simply the concentration of 
freshwater adapted alleles found in the marine.

Note: we should probably investigate the distribution of effect size more to clearly answer the above questions ...

(NOTE:PETER/BILL how does this compare to what we see in nature

\subsection{Qualitative Threshold}

\subsection{Migration Load}

%\begin{figure}
%	\begin{center}
%  		\includegraphics[width=1.2\linewidth]{MigrationLoad}
%  		\caption{A boat.}
%  		\label{fig:boat1}
%	\end{center}
%\end{figure}

While both of the higher migration rates of $0.005$ and $0.05$ showed evidence of sharing alleles and rapid adaptation,
we see substantial difference in the median phenotype compared to the optimum for each subpopulation. 
This is because migration tends to oppose the natural divergence of populations undergoing ecological speciation. \cite{Bolnick}
Migration load depreciation of fitness between two subpopulations with a high amount of introgression between them. 
This is often observed in nature when the sharing of genes constricts the adaptation of both species.

 In our simulations, at high migration rates of $0.05$, 
we consistently saw this effect in all subpopulations.

\subsection{Low Recombination in Effect Regions}

Goldilocks Zone?

at 10 migrants/generation (or 0.005 of the population) 
is the first parameter value at which we observe

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
\end{comment}

\section{Discussion}

% at low M we should be able to predict something ---
% at high M we should be able to predict something ----

These simulation we're

\bibliography{Citations}{}
\bibliographystyle{plain}

\end{document}

### 2018-06-22 Stickleback_Paper.tex
latest =
added matplotlib plots

@@
 
-\subsubsection{Migration is Mixing}
+\subsubsection*{Migration is Mixing}
 
@@
 
-\begin{figure}[h!tb]
+\begin{figure}[htb]
 	\begin{center}
-  		\includegraphics[width=01.0\linewidth]{PhenotypeDist}
+  		\includegraphics[width=01.0\linewidth]{matplotlib/Fst.png}
   		\caption{A boat.}
@@
 As can be seen in Figure (neutral fst figure)
-$F_{st}$ values for neutral Alleles steadily decline between all sets of populations as 
+$F_{st}$ values for neutral alleles steadily decline between all sets of populations as 
 migration increases, this illustrates less difference between sections of the genome which are not acted up 
@@
 
-\subsubsection{migration affects the speed of adaptation}
+\subsubsection*{migration affects the speed of adaptation}
 
@@
 %-Time to adaptation
+%\begin{figure}[h!tb]
+%	\begin{center}
+%  		\includegraphics[width=01.0\linewidth]{???}
+%  		\caption{A boat.}
+%  		\label{fig:boat1}
+%	\end{center}
+%\end{figure}
 
@@
 
-\subsubsection{sharing of freshwater adapted alleles}
+\subsubsection*{sharing of freshwater adapted alleles}
 
@@
 %-Total/Avg shared
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.8\linewidth]{matplotlib/MPFAI.png}
+  		\caption{A boat.}
+  		\label{fig:boat2}
+	\end{center}
+\end{figure}
 
@@
 
-\subsubsection{We see Migration Load after a certain threshold of migration}
+\subsubsection*{We see Migration Load after a certain threshold of migration}
 
@@
 
-\subsubsection {There is a Qualitative Threshold of migration rates}
+\subsubsection*{There is a Qualitative Threshold of migration rates}
 
@@
 
-\subsubsection {Low Recombination Causes Clustering}
+\subsubsection*{Low Recombination Causes Clustering}
 
@@
 %--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+%SHOULD PROBABLY JUST DELETE ALL THIS. V UNORGANIZED.
 Across migration rates we observe a large difference in:

### 2018-06-22 Stickleback_Paper.tex
latest plots

@@
 
-Note: Finish after the whole paper has been written
+\todo{Do last.}
 
@@
 Here we give a brief overview of the software used and the models created to observe the effects of selection. 
+Each set of simulations was run multiple times to ensure the data we observed was ???
 
-\todo{Each set of simulations was run multiple times to ensure the data we observed was not stochastic.}
-
-Note: Maybe talk about some things are stochastic: 
-(1) Correlation of mutation frequencies at low migration rates
-(2) Distribution of Effect Size? (run another set to make sure)
+%(1) Correlation of mutation frequencies at low migration rates
+%(2) Distribution of Effect Size? (run another set to make sure)
 
@@
 %-Time to adaptation
-%\begin{figure}[h!tb]
-%	\begin{center}
-%  		\includegraphics[width=01.0\linewidth]{???}
-%  		\caption{A boat.}
-%  		\label{fig:boat1}
-%	\end{center}
-%\end{figure}
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=01.0\linewidth]{matplotlib/TimeToAdaptation.png}
+  		\caption{A boat.}
+  		\label{fig:boat1}
+	\end{center}
+\end{figure}
 
@@
 
+\newpage
 \bibliography{Citations}{}

### 2018-06-25 Stickleback_Paper.tex
latest

@@
 
+\todo{maybe the order of these should be reversed?}
+
 \subsubsection{Genomic Architecture}
@@
   		\includegraphics[width=01.0\linewidth]{matplotlib/Fst.png}
-  		\caption{A boat.}
-  		\label{fig:boat1}
+  		\caption{Distributions of $F_{st}$  values between the three subpopulations throughout the simulations.
+		On the left, $F_{st}$ was calculated only on neutral mutation frequencies. On the right, $F_{st}$ was calculated
+		only on effect (Impact on phenotype) mutation frequencies. 
+		Effect mutations are acted upon by selection as they have an affect on fitness where neutral mutations  }
+  		\label{fig:Fst}
 	\end{center}
@@
 Across increasing parameters of migration, we observed more gene flow across the populations. 
-As can be seen in Figure (neutral fst figure)
+As can be seen in Figure \ref{fig:Fst},
 $F_{st}$ values for neutral alleles steadily decline between all sets of populations as 
@@
   		\includegraphics[width=01.0\linewidth]{matplotlib/TimeToAdaptation.png}
-  		\caption{A boat.}
-  		\label{fig:boat1}
+  		\caption{Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
+		it takes for the introduced population's mean phenotype to come within 0.5 of the original lakes average phenotype. 
+		Each point represents a simulation run at some value of $M$. 
+		The yellow dashed line is the average of all points at each respective parameter value}
+  		\label{fig:TimeToAdaptation}
 	\end{center}
@@
 as migration rates increased. 
-As can be seen in Figure (Time Until Adaptation) at the lowest migration rate of $5 * 10^{-5}$,
+As can be seen in Figure \ref{fig:TimeToAdaptation} at the lowest migration rate of $5 * 10^{-5}$,
 It took the introduced population over 30 thousand generations for the average phenotype of all the lakes to 
@@
   		\includegraphics[width=0.8\linewidth]{matplotlib/MPFAI.png}
-  		\caption{A boat.}
-  		\label{fig:boat2}
+  		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) throughout the simulation run, for each subpopulation.
+		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
+		freshwater adapted alleles.
+		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
+		while the percentage tells us the variance of the haplotype (note: I also have boxplots telling us total num FAA for the average individual that are interesting to look at "matplotlib/MFAI.png")}
+  		\label{fig:MPFAI}
 	\end{center}
@@
 %- Phenotype Distribution
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.8\linewidth]{matplotlib/MeanPhenotype1.png}
+  		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
+		The dashed pink line (Pheno = +10) is the optimum phenotype any individual in the marine environment.
+		In contrast the purple line at (Pheno = -10) represents the optimum for any individual in the freshwater environment. 
+		All individuals at generation 0 (beginning of the simulation) 
+		}
+  		\label{fig:MeanPhenotype}
+	\end{center}
+\end{figure}
 
@@
 as a consequence of a large number of offspring through hybridization events between subpopulations.
-In Figure (Phenotype distribution) at $M = 5x10^{-2}$, we can see the distribution of average phenotype throughout the simulation
+In Figure \ref{fig:MeanPhenotype} at $M = 5x10^{-2}$, we can see the distribution of average phenotype throughout the simulation
 pull towards the opposing selective pressure value and away from the local optimum in all subpopulations.
@@
 This leads us to consider a window (Goldilocks Zone) of introgression which allows for the transportation
-of FAA's without migration load.
+of FAA's without migration load. In a species that commonly is subjected to two general types of selective pressure 
+such as marine and
 
@@
 
-\section{Discussion}
+\section{Conclusion \& Discussion}
 
@@
 
-These simulation we're
+Here, we suggest a range of migration rate (introgression) parameter values which would allow for the rapid (and in turn, parallel) adaptation 
+of marine stickleback introduced intro a freshwater environment.
+While this range is heavily effected by all other selections of parameter values.
 

### 2018-06-26 Stickleback_Paper.tex
latest

@@
 \section{Introduction}
+
+\todo{Take out any "we's" and be more specific}
 
@@
 
-\todo{Should I put my geography picture here?}
+\begin{figure}[htb]
+	\begin{center}
+  		\includegraphics[width=0.8\linewidth]{GeographyDiagram}
+  		\caption{A representation of the geographic and evolutionary history of all populations throughout the simulation. 
+		The marine is a one-dimensional, continuous population with spatial positions ranging from [0.0 - 10.0]. Each 
+		lake$_{i}$ is a discrete population connected by migration only through the marine at position $i - 0.5$. The introduced lakes 
+		have the same location and selective pressure as the original lakes, but arise at 40K generations, 
+		initialized as a copy of all marine individuals.}
+  		\label{fig:Geo}
+	\end{center}
+\end{figure}
 

### 2018-06-26 Stickleback_Paper.tex
latest

@@
 
+In SLiM, the genome is conceptually a linear array of loci defining locations for 
+mutations that arise and breakpoints for recombination. 
+Genomic elements define a range depicting the types of mutations that can arise, the ratio at which they can arise. 
 Although still uncertain about how much of the genome is directly associated with the distinguished phenotypes, 
@@
 To mimic this and compare dynamics of neutral vs. effect under selection, 
-we create regions of effect. These genomic elements are the only loci which allow for mutation that impact phenotype. 
-This allows us to observe differences in selection at the genomic level. 
-In SLiM, the genome is conceptually a linear array of loci for which we can define
-different amount 
+we create regions of effect. 
+These genomic elements are the only loci which allow for mutations that impact phenotype.
+
+In our simulations, there are 10 effect regions of size 100 loci, uniformly placed between buffer regions of size 4950 loci. 
+This means the effect regions make up exactly $1\%$ of the entire genome for every individual throughout the simulation. 
+The entire genome has a mutation rate of $10^{-7}$ meaning approximately 1 in every 100 individuals, per generation, is subject to a new mutation. 
+The genome has a recombination rate of $10^{-5}$, which allows for approximately 1 breakpoint event, per individual, per generation.
+
+%This allows us to observe differences in selection at the genomic level.
+
 \todo{finish this}

### 2018-06-26 Stickleback_Paper.tex
latest

@@
 we create regions of effect. 
-These genomic elements are the only loci which allow for mutations that impact phenotype.
+In these genomic elements of effect, we uniformly allow additive, recessive, and dominant, mutation that can push phenotype in either the positive or negative direction to arise. 
+Every genomic position outside of these effect regions only allows for selectively neutral mutations to arise.
 
@@
 The entire genome has a mutation rate of $10^{-7}$ meaning approximately 1 in every 100 individuals, per generation, is subject to a new mutation. 
-The genome has a recombination rate of $10^{-5}$, which allows for approximately 1 breakpoint event, per individual, per generation.
+The genome has a recombination rate of $10^{-5}$, which is approximately 1 breakpoint event, per individual, per generation.
 

### 2018-07-04 Stickleback_Paper.tex
latest

@@
 \usepackage[backgroundcolor=pink,linecolor=red]{todonotes}
+\usepackage{fullpage}
 
@@
 \todo{Do last.}
-
-%wider audience by more general questions about ... introduction maybe too fishy???
-
-%a question about the genetic basis for adaptstion, how is it maintianed?
-% theres alot of debate in the field
-
-% more specifically - sticklback fish
-
-% we want to examine this, This is why our approach is unique
-% This is why the field could benefit from this.
-
-%methods - we did this this and this specifically
-%results - we found this and this specifically
-
-% discussion get more broad again.
 
@@
 
-In 2009, Dolph Schluter and Gina L. Conte suggested the "transportation"-hypothesis. 
+In 2009, Dolph Schluter and Gina L. Conte suggested the "transporter"-hypothesis. 
 This outlined the flow of freshwater alleles into marine populations through offspring of hybridization events.
@@
 how much migration and gene flow is to be expected for this relationship to be maintained?
-(Other questions)
-how does selection, at the genomic level, alter given a range of migration and recombination rates?
+%(Other questions)
+%how does selection, at the genomic level, alter given a range of migration and recombination rates?
 Here, we use forward moving simulations to model marine and freshwater populations of stickleback.
@@
 We compare to real data and make predictions about realistic parameters.
-
-Simulations are a powerful tool when there is some map between the parameters we want to learn about
-and the genetic data we observe in nature. 
-We don't know what that map is, so we invert the relationship. 
-This means going from genomic data that we see to the parameter values that must be driving the system, 
-assuming the model is correct.
 
@@
 
-\subsubsection{Geography}
+\subsubsection*{Geography}
 
@@
 
-\begin{figure}[htb]
+\begin{figure}
 	\begin{center}
-  		\includegraphics[width=0.8\linewidth]{GeographyDiagram}
+  		\includegraphics[width=0.6\linewidth]{GeographyDiagram}
   		\caption{A representation of the geographic and evolutionary history of all populations throughout the simulation. 
@@
 		lake$_{i}$ is a discrete population connected by migration only through the marine at position $i - 0.5$. The introduced lakes 
-		have the same location and selective pressure as the original lakes, but arise at 40K generations, 
-		initialized as a copy of all marine individuals.}
+		have the same location and selective pressure as the original lakes, but arise at 40K generations.
+		the introduced population is initialized as a copy of all marine individuals to model marine 
+		individuals inhabiting a newly created freshwater environment such as the ponds around Middleton Island.
+		We then observe the selection process of marine individuals and following generations 
+		with the new selective pressure across a range of parameter values. 
+		%Migration between the marine and original lakes is 
+		%set to $0$ the generation before introduction to avoid direct migrants from original $->$ introduced lakes.
+		}
   		\label{fig:Geo}
@@
 
-\subsubsection{Selective Pressure}
+\subsubsection*{Selective Pressure}
 
@@
 
-\subsubsection{Genomic Architecture}
+\subsubsection*{Genomic Architecture}
 
@@
 mutations that arise and breakpoints for recombination. 
-Genomic elements define a range depicting the types of mutations that can arise, the ratio at which they can arise. 
+Genomic elements define a range depicting the types of mutations that can arise, and the ratio at which different mutation types can arise. 
 Although still uncertain about how much of the genome is directly associated with the distinguished phenotypes, 
@@
 The genome has a recombination rate of $10^{-5}$, which is approximately 1 breakpoint event, per individual, per generation.
-
-%This allows us to observe differences in selection at the genomic level.
 
@@
 
-\begin{figure}[htb]
+\begin{figure}
 	\begin{center}
-  		\includegraphics[width=01.0\linewidth]{matplotlib/Fst.png}
+  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/FST_HRR.png}
   		\caption{Distributions of $F_{st}$  values between the three subpopulations throughout the simulations.
@@
 %-Time to adaptation
-\begin{figure}[h!tb]
+\begin{figure}
 	\begin{center}
-  		\includegraphics[width=01.0\linewidth]{matplotlib/TimeToAdaptation.png}
+  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/TimeToAdaptation.png}
   		\caption{Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
@@
 %-Total/Avg shared
-\begin{figure}[h!tb]
+\begin{figure}
 	\begin{center}
-  		\includegraphics[width=0.8\linewidth]{matplotlib/MPFAI.png}
-  		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) throughout the simulation run, for each subpopulation.
+  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MPFAI.png}
+  		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
 		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
@@
 		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
-		while the percentage tells us the variance of the haplotype (note: I also have boxplots telling us total num FAA for the average individual that are interesting to look at "matplotlib/MFAI.png")}
+		while the percentage tells us the variance of the haplotype}
 		\label{fig:MPFAI}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MFAI.png}
+  		\caption{Distributions of mean number of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
+		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
+		freshwater adapted alleles.
+		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
+		}
+  		\label{fig:MPFAI}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=0.6\linewidth]{plotlyPlots/NumFAA.png}
+  		\caption{(PLACEHOLDER)}
+		\label{fig:NumFAA}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=0.6\linewidth]{plotlyPlots/StandingGeneticVariation.png}
+  		\caption{(PLACEHOLDER)}
+		\label{fig:NumFAA}
 	\end{center}
@@
 	\begin{center}
-  		\includegraphics[width=0.8\linewidth]{matplotlib/MeanPhenotype1.png}
+  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MeanPhenotype1.png}
   		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
@@
 
-\subsubsection*{Low Recombination Causes Clustering}
+\subsection{Genomic Architecture}
 
-\begin{comment}
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-%SHOULD PROBABLY JUST DELETE ALL THIS. V UNORGANIZED.
-Across migration rates we observe a large difference in:
-gene flow, 
-$F_{st}$ values,
-amount of time until adaptation,
-sharing of solutions (haplotytes?),
-$SGV$,
-and, migration load. 
-we run multiple simulations across a range of migration rates starting at $5e-5$ and increasing by
-one order of magnitude up to $.05$. 
-With the marine and freshwater both at a population size 2000,
-this represents exactly $1/10_{th}$, $1$, $10$, and $100$ migrants per generation, respectively.
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{matplotlibPlots/Power_FP.png}
+  		\caption{ 
+		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
+		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
+		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
+		}
+  		\label{fig:Power_FP}
+	\end{center}
+\end{figure}
 
-At low migration rates of $5e-5$ (1/10 migrants per generation), we observed little sharing of genetic solutions and 
-almost no freshwater alleles existing in the marine population.
-With an MPFAI of $0.0999$ in the original lakes;
-each individual in the original lakes had almost exactly $1/10_{th}$ of the total FAA. 
-In addition to each allele appearing in an average of $1.08$ lakes, 
-this tells us that each of the 10 lakes has individually selected their own set of solutions (haplotypes) 
-rather than sharing alleles through the marine. 
-This is further supported by the low capacity of the marine to transport the alleles with an MPFAI of $0.003$.
-$T_{adapt}$ remained on an ecological timescale of $18K$ generations, and the correlation of FAA frequencies was at $0.09$.
-With the introduced population having an MPFAI of $0.02$, 
-we see little evidence that the lakes were able the select on the alleles found in that marine. 
-it may be important to note that while we are sampling the 10 introduced lakes as a whole population,
-It is possible that only a couple of the lakes even received a copy some alleles during the introduction.
-However, to see evidence of the haplotypes being carried throughout the marine we would expect to observe a pattern off all lakes
-being able to select upon similar alleles. 
-This observation gives a migration rate below the threshold needed to support the transportation hypothesis.
+Here, we take a closer look at the genomic architecture of local adaptation between freshwater and marine individuals. 
+Often when looking at real datas, biologists observe patterns of $F_{st}$ and clustering of alleles that suggest causative loci for certain traits.
+However, researchers would like to know how parameters such as migration and recombination impact the results from natural populations.
+To give an idea about the impact of these parameters in our model, we examine the distribution, location, and false positive rate on predicting effect loci
+given the $F_{st}$ per SNP across the genome.
 
-scaling up migration rates to $0.0005$ (1 migrants per generation)
-we observe the first evidence of transportation.
-The mean total number of FAA drops to 45 and the average number of lakes which each alleles appears in jumped to $1.33$. 
-This shows us the gene flow between populations allowed for some freshwater alleles to travel through the marine and be shared by the lakes. 
-Interestingly, with the marine showing an MPFAI of $0.02$, 
-the introduced was much more able to select upon the alleles;
-The introduced population showed a MPFAI of $0.078$, compared to the original lakes which was at $0.129$,
-Looking at the FAA frequencies in the two freshwater populations at time of adaptation, we see a correlation of $.68$
-While this is evidence of parallel local adaptation, the  $T_{adapt}$ was still at $8,365$ generations.
+\subsubsection*{False-Positives \& Statistical Power}
 
-At $0.005$ (10 migrants per generation) we experience high correlation between the FAA frequencies
-in the original and introduced populations, $0.96$, and rapid local adaptation of the introduced population, (81 generations). 
-With the introduced population showing an MPFAI of  ~ $0.34$ which is strikingly close to the original freshwater population which is at $0.34$
-This migration rate clearly shows the capacity to share alleles in order for us to see rapid and parallel adaptation.
+Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$. 
+Knowing exactly which mutations effect phenotype in our simulations, 
+we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome. 
+In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power. 
+This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.
 
-Interestingly, is that if we look at the distribution and total amount of FAA in the marine population at time of introduction between 
-$0.0005$ and $0.005$, they are virtually identical.
-This comes with two questions
-(1) why is there a similar amount of FAA in the marine with such a substantial difference in migration, and
-(2) what is it about the higher migration rate of  $0.005$ that allows it perform so much better in terms of rapid adaptation? 
-To answer the first question we can see that the median number of defined FAA is much higher, 45
-While with the higher migration there is only a median of 10. 
-This is a result of the solutions not being shared across the original lakes at lower migration rates.
-This can be seen by the average number of lakes each FAA appears in being 
-$1.54$ at 1 migrant per generation and 
-$3.75$ at 10 migrants per generation.
-consequently more alleles were defined as FAA and sampled as such in the marine population. 
-This however does not explain why, at higher migration rates, 
-the introduced freshwater population was able to rebuild the haplotype so much more quickly. 
-For this, It is important to note that after introduction, there is continued migration. 
-This means that for a migrate of $0.005$ there is roughly 10 times as 
-much of the FAA being brought into the freshwater every generation compared to migrate $0.0005$.
+%The lowest migration rate of $10^{-5}$ also has a very low false positive rate across all $F_{st}$ threshold values.
 
-This seams to suggest that continued migration is quite important 
-for rapid adaptation at this migration rate as opposed to the simply the concentration of 
-freshwater adapted alleles found in the marine.
+%\subsubsection*{Low Recombination Causes Clustering}
+%\begin{comment}
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-5.png}
+  		\caption{ (PLACEHOLDER)
+		}
+  		\label{fig:Fst1}
+	\end{center}
+\end{figure}
 
-Note: we should probably investigate the distribution of effect size more to clearly answer the above questions ...
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-4.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:Fst2}
+	\end{center}
+\end{figure}
 
-(NOTE:PETER/BILL how does this compare to what we see in nature
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:Fst3}
+	\end{center}
+\end{figure}
 
-\subsection{Qualitative Threshold}
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-2.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:Fst4}
+	\end{center}
+\end{figure}
 
-\subsection{Migration Load}
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-5.png}
+  		\caption{ (PLACEHOLDER)
+		}
+  		\label{fig:Haplo1}
+	\end{center}
+\end{figure}
 
-%\begin{figure}
-%	\begin{center}
-%  		\includegraphics[width=1.2\linewidth]{MigrationLoad}
-%  		\caption{A boat.}
-%  		\label{fig:boat1}
-%	\end{center}
-%\end{figure}
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-4.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:Haplo2}
+	\end{center}
+\end{figure}
 
-While both of the higher migration rates of $0.005$ and $0.05$ showed evidence of sharing alleles and rapid adaptation,
-we see substantial difference in the median phenotype compared to the optimum for each subpopulation. 
-This is because migration tends to oppose the natural divergence of populations undergoing ecological speciation. \cite{Bolnick}
-Migration load depreciation of fitness between two subpopulations with a high amount of introgression between them. 
-This is often observed in nature when the sharing of genes constricts the adaptation of both species.
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-3.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:Haplo3}
+	\end{center}
+\end{figure}
 
-In our simulations, at high migration rates of $0.05$, 
-we consistently saw this effect in all subpopulations.
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-2.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:Haplo4}
+	\end{center}
+\end{figure}
 
-\subsection{Low Recombination in Effect Regions}
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-5.png}
+  		\caption{ (PLACEHOLDER)
+		}
+  		\label{fig:Haplo1}
+	\end{center}
+\end{figure}
 
-Goldilocks Zone?
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:Haplo2}
+	\end{center}
+\end{figure}
 
-at 10 migrants/generation (or 0.005 of the population) 
-is the first parameter value at which we observe
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:Haplo3}
+	\end{center}
+\end{figure}
 
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-\end{comment}
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-2.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:Haplo4}
+	\end{center}
+\end{figure}
+%\end{comment}
 

### 2018-07-04 Stickleback_Paper.tex
plr comments

@@
 \usepackage[english]{babel}
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{setspace}
@@
 \usepackage[backgroundcolor=pink,linecolor=red]{todonotes}
-\usepackage{fullpage}
+% \usepackage{fullpage}
+\usepackage[hidelinks]{hyperref}
+\usepackage{xcolor}
 
@@
 %\setstretch{1.1}
+
+\newcommand{\plr}[1]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue]{#1}}
 
@@
 
+\plr{I like how this starts out, but going to Darwin is maybe a bit too ``big'';
+    better to say that we now have a number of examples of rapid evolution;
+    cite the Grants and a few others (Gough island house mice? insecticide resistance in Drosophila?),
+    then move on to the sticklebacks in more detail.}
+
 Recently, rapid and parallel adaptation has brought about question of selection of
@@
 we've observed significant phenotypic changes in less than 50 years that appear to be 
-parallel to freshwater stickleback that have been separated for over $13,000$ years. \cite{LescakE7204} 
+parallel to freshwater stickleback that have been separated for over $13,000$ years \citep{LescakE7204}.
+\plr{clarify what ``parallel'' means}
+\plr{``citep'' for ``cite in parentheses''. (and, citations come before the punctuation)}
 In freshwater stickleback, 
@@
 and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region (Cite Kristin, Cresko et. al)
+\plr{End with a question that the next two paragraphs propose an answer to.}
 
 The leading hypothesis is that rather than acting on new mutations, ecological speciation is sped up through selection on standing genetic variation found in marine populations. 
+\plr{Is this ``ecological speciation''? Speciation would imply the ecotypes are developing reproductive incompatibility;
+    I think that we think the pattern is stable. Maybe ``adaptation to freshwater habitats''?}
 One clear example of this is the gene \textit{eda} 
@@
 
-In 2009, Dolph Schluter and Gina L. Conte suggested the "transporter"-hypothesis. 
+% In 2009, Dolph Schluter and Gina L. Conte 
+\citet{schluter2009genetics} 
+\plr{"citet" for "cite, as part of the text"}
+suggested the ``transporter''-hypothesis.  
+\plr{in latex, use ` `words' ' to get the open- and closed-double quotes (not the double-quote character)}
 This outlined the flow of freshwater alleles into marine populations through offspring of hybridization events.
@@
 allows the marine to maintain the freshwater haplotype dispersed in low frequency among its' individuals. 
-Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection rebuilds the freshwater haplotype. \cite{Schluter2009} 
+Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection rebuilds the freshwater haplotype. \cite{schluter2009genetics} 
 The alternative to this hypothesis is that individuals from other freshwater environments migrate directly to the new environments, and their haplotype is passed down directly. 
@@
 and almost no freshwater individuals. 
+\plr{Bring up the possibility of more than one freshwater haplotype here.}
 
@@
 how much migration and gene flow is to be expected for this relationship to be maintained?
+\plr{motivating questions need more work}
 %(Other questions)
@@
 Here we give a brief overview of the software used and the models created to observe the effects of selection. 
+\plr{Should focus on the model, not the software -- describe how the simulation works}
 Each set of simulations was run multiple times to ensure the data we observed was ???
@@
 The simulation life cycle of each generation is as follows.
+\plr{This can be merged with the next section, perhaps as a section called ``Life cycle''.  
+    Here and below, avoid using SLiM jargon (don't say ``callback'').}
 First, 
@@
 Using SLiM?s $modifyChild()$ callback, we?ve modeled both to act as 10 separate discrete subpopulations. 
+\plr{no need to say this: we will provide the SLiM recipes as supplementary material
+    so that someone interested can see exactly how it works.}
 Each of the $10$ lakes, $i$, is located along the marine at $i - 0.5$, 
@@
 and acts as a representation of lateral plate number and opercle shape in the stickleback populations.
+\plr{More concisely, these could be called the ``optimal trait values''.}
 The fitness of an individual is then determined by a probability density for a normal distribution 
 at the quantile (?) of the difference between the optimum and the individual's phenotype
+\plr{instead, we can say that we use a ``Gaussian fitness function'', 
+    where the fitness of an individual with trait $x$ in a population with optimal trait value $x_0$ is:}
+\begin{align}
+    f(x; x_0) &= \exp\left(\frac{(x - x_0)^2 }{ 2 \sigma^2}\right),
+\end{align}
+with $\sigma = 15$.
 
-    \[
-\textnormal{Individual.Fitness = } X \sim \mathcal{N}(\textnormal{Optimum - Individual.Phenotype},\,0.0\,15.0).
-    \]
+%     \[
+% \textnormal{Individual.Fitness = } X \sim \mathcal{N}(\textnormal{Optimum - Individual.Phenotype},\,0.0\,15.0).
+%     \]
 
@@
 determined as a summation of the selection coefficients of all mutations an individual possesses. 
-
+\plr{Hm. Should we say ``phenotype'' or ``trait''? I've been saying ``trait'', which is less jargon-ey?}
+% no empty lines before equations, usually - they should read like part of the text itself.
 	\[
@@
 Among the effect mutations; 
+\plr{``effect mutation'' is not a standard term; should say ``mutations that affect the phenotype'' here,
+    and be careful about what we call them elsewhere.}
 there are additive, dominant, and recessive mutations which effect phenotype
@@
 with a shape parameter of $1$ and a mean of either $0.5$, or ${-0.5}$. 
-
+\plr{Can just say this in words, no need to have an equation.
+    Also, a Gamma with shape 1 and mean 0.5 is an Exponential with mean 0.5.}
 	\[
@@
 \todo{maybe the order of these should be reversed?}
+\plr{I agree: first explain how phenotypes are calculated, then say what fitness is.
+    Notation suggestion: phenotype for an individual that is heterozygous at loci $H$
+    and homozygous derived at loci $D$ is $x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
+    where $h_\ell$ is the dominance coefficient for locus $\ell$ (0 for recessive, 1 for dominant, 0.5 for additive),
+    and $s_\ell$ is the effect of the derived mutation on the phenotype}
 
@@
 
+\plr{This should maybe come before the phenotype.}
 In SLiM, the genome is conceptually a linear array of loci defining locations for 
@@
 \subsection{Sampling}
+\plr{I think the purpose of this section is to define terms and statistics used later;
+    so maybe a better title is ``Descriptive statistics''?}
 
@@
 impact the sharing of alleles between populations. 
+\plr{Below, I think we should make sure to use ``population'' to mean either ``lake'' or ``ocean'';
+    so when we compute freshwater-marine $F_{ST}$ this measures differentiation between habitats,
+    not between populations.}
 we define freshwater adapted alleles (FAA),
+\plr{I vote to remove the abbreviations. Also, should make the dependence on generation more clear: maybe ``freshwater adapted alleles at time $t$''?}
 at any given generation,
@@
 This is because the transportation hypothesis does not specify where or when an advantageous mutation arises; 
-but simply suggests that when one comes to high enough frequency, it too, could participate in the transportation process. \cite{Schluter2009}
+but simply suggests that when one comes to high enough frequency, it too, could participate in the transportation process. \cite{schluter2009genetics}
 To observe transportation of freshwater alleles in our simulations, we use a variety of metrics when sampling.
@@
 mutation frequencies, 
-$F_{st}$ values, 
+$F_{ST}$ values, \plr{how is $F_{ST}$ computed?}
 average phenotype, 
@@
 computed as the sum of the FAA mutation frequencies among each population.
+\plr{Define ``mean percentage''; no need to say that it is the sum of the frequencies, I think.}
+\plr{Here FAA are defined in terms of the original populations, explaining why the introduced populations can have MPFAI of 0.
+    This is counter-intuitive from the name (FAA); maybe we should call them ``pre-existing freshwater alleles''?
+    ``standing genetic variants''?}
 we're particularly interested in using MPFAI to measure the marine population' capacity to carry FAA,
 as well as observing the introduced populations' ability to select upon them. 
+\plr{omit this last sentence?}
 
@@
 and the number of shared high frequency alleles. 
+\plr{Here we need to define terms; we don't need to say exactly what was computed,
+    as that will be redundant with the Results.}
 These metrics tells us how much of the FAA  were used in the lakes once they have locally adapted.
@@
 Local adaptation occurred across all simulated parameter values.
-Starting from the same baseline freshwater and marine populations diverged phenotypically, 
+Starting from the same baseline, freshwater and marine populations diverged phenotypically, 
 until they reached an equilibrium,
@@
 phenotypic variation within each population was small compared to the difference between populations.
-Polymorphic loci that effected the phenotype,
+Polymorphic loci that affected the phenotype,
 showed greater differentiation compared to neutral loci between marine and freshwater populations, with the exception 
@@
 $F_{st}$ plots along the genome, show peaks at loci underlying local adaptation.
+\plr{Rather than referring to e.g., ``migration rate of $5 \times 10^{-5}$'',
+    it'd be easier to follow if we had names for the different migration rate situations.
+    Maybe ``low, medium-low, medium-high, and high''?
+    Or talk about the number of migrants per generation per lake, e.g.,
+    ``with the exception of the lowest migration rate of 0.01 migrants per generation per lake''?}
 
 \todo{ add more general observations and caveats about not all parameter values. }
+
+\plr{It would be nice to have a small table of parameter values,
+    so that when someone wonders ``What was the population size again?'' they could easily find it.}
+
+\plr{I vote to put figures \ref{fig:phenotype_ts2} and \ref{fig:phenotype_ts3} 
+    together as one figure and put them here.
+    There's a lot in those we could talk about: you can see the stepwise introduction
+    of individual alleles in figure \ref{fig:phenotype_ts2}, for instance.}
 
@@
 \subsubsection*{Migration is Mixing}
+\plr{Title: ``Migration homogenizes genetic variation''}
 
@@
 upon by selection, in all populations.
-Additionally, Standing Genetic Variance \todo{computed as?}
-(Seen in SGV Figure)
+Additionally, standing genetic variance \todo{computed as?}
+(seen in Figure \ref{fig:SGV})
 values in the marine population steady increased with migration which suggests that more alleles from the 
@@
 get to within 0.5 of the original lakes original phenotypes. 
+\plr{From the plots it looks more like adaptation happens (at least in most lakes) by like 10K or 15K.}
 This suggests that selection was acting primarily on new mutations, meaning the 
@@
 it was selected upon. 
+\plr{We should do the calculation to see if this makes sense: 
+    there's one new effect mutation in the right direction every 100 generations in the lakes,
+    one of effect size $s$ has probability of about $s/2$ of not being lost to drift...}
 There was a positive correlation between the amount of standing genetic variation in the 
@@
 	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MPFAI.png}
+  		\includegraphics[width=\linewidth]{matplotlibPlots/MPFAI.png}
   		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
@@
 	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MFAI.png}
+  		\includegraphics[width=\linewidth]{matplotlibPlots/MFAI.png}
   		\caption{Distributions of mean number of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
@@
 		}
-  		\label{fig:MPFAI}
+  		\label{fig:MNFAI}
 	\end{center}
@@
 	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{plotlyPlots/NumFAA.png}
+  		\includegraphics[width=\linewidth]{plotlyPlots/NumFAA.png}
   		\caption{(PLACEHOLDER)}
@@
 	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{plotlyPlots/StandingGeneticVariation.png}
+  		\includegraphics[width=\linewidth]{plotlyPlots/StandingGeneticVariation.png}
   		\caption{(PLACEHOLDER)}
-		\label{fig:NumFAA}
+		\label{fig:SGV}
 	\end{center}
@@
 such as marine and 
+\plr{Maybe this goes in the Discussion? Remaining observations could merge with the previous section.}
 
@@
 
+\plr{Start off by talking about Fst along the genome, then say ``if we identify QTL as the outliers, then...''}
+\plr{I vote for combining and including figures \ref{fig:Fst2} and \ref{fig:Fst3}.}
 Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$. 
@@
 		}
-  		\label{fig:Haplo1}
+  		\label{fig:phenotype_ts1}
 	\end{center}
@@
 		}
-  		\label{fig:Haplo2}
+  		\label{fig:phenotype_ts2}
 	\end{center}
@@
 		}
-  		\label{fig:Haplo3}
+  		\label{fig:phenotype_ts3}
 	\end{center}
@@
 		}
-  		\label{fig:Haplo4}
+  		\label{fig:phenotype_ts4}
 	\end{center}
@@
 \newpage
+\plr{some problems because you had bibliographystyle before bibliography}
+\bibliographystyle{plainnat}
 \bibliography{Citations}{}
-\bibliographystyle{plain}
 

### 2018-07-06 Stickleback_Paper.tex
Introduction updated

@@
 \usepackage[backgroundcolor=pink,linecolor=red]{todonotes}
-% \usepackage{fullpage}
+ \usepackage{fullpage}
 \usepackage[hidelinks]{hyperref}
@@
 \newcommand{\plr}[1]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue]{#1}}
+\newcommand{\jgg}[1]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=black]{#1}}
 
@@
 \linespread{1.5}
+
+JARGON WE SHOULD THINK ON ??
+
+Trait | Phenotype = Probably Trait
+
+effect loci | QTL | causal: = ??
+
+Original Lakes | $P_{1}$, $P_{3}$, $P_{3}$ = ?
+
+Migration Rate | Introgression = ?
 
@@
 
-In the mid 19th century,
-Darwin theorized that the process of natural selection resulted in speciation, 
-meaning all living species derived from a common ancestor. 
-He also believed that this process took place on ecological timescales that would not allow for 
-this process to be observed in a single lifetime. 
-Today, we have found evidence of evolution
-in tens of generations and are starting to grasp the complexity behind such a simple idea. 
-Famously, Drs. Peter and Rosemary Grant went on to study Darwin's finches and
-found that phenotypes could adapt to changes in the food source within mere generations.
-%more specific?
-We have have found many examples of rapid adaptation in many commonly studied models
-but many questions still remain surrounding the genomics which allow for this.
+% Jared's general notes for formatting
 
-\plr{I like how this starts out, but going to Darwin is maybe a bit too ``big'';
-    better to say that we now have a number of examples of rapid evolution;
-    cite the Grants and a few others (Gough island house mice? insecticide resistance in Drosophila?),
-    then move on to the sticklebacks in more detail.}
+%\plr{``citep'' for ``cite in parentheses''. (and, citations come before the punctuation)}
+%\plr{in latex, use ` `words' ' to get the open- and closed-double quotes (not the double-quote character)}
+%\plr{"citet" for "cite, as part of the text"}
 
-Recently, rapid and parallel adaptation has brought about question of selection of
-standing genetic variation in natural populations 
-of Stickleback.
-In 1964, The Great Alaskan Earthquake caused Middleton island to raise up and in turn, 
-introduced a group of new freshwater ponds around the perimeter of the island. 
-Quickly inhabited by the surrounding marine population of Ninespine Stickleback,
-we've observed significant phenotypic changes in less than 50 years that appear to be 
-parallel to freshwater stickleback that have been separated for over $13,000$ years \citep{LescakE7204}.
-\plr{clarify what ``parallel'' means}
-\plr{``citep'' for ``cite in parentheses''. (and, citations come before the punctuation)}
-In freshwater stickleback, 
-the number of lateral plates are reduced
-and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region (Cite Kristin, Cresko et. al)
-\plr{End with a question that the next two paragraphs propose an answer to.}
+\jgg{Starting a paper is awkward, help}
+The rate and tempo of evolution remains largely unknown.
+Today, we have found multiple examples as evidence of rapid evolution 
+in tens of generations or fewer.
+Biologists are now starting to address the complexity behind ecological speciation. 
+\jgg{Peter and Bill: Could use help with examples of rapid evolution and the story leading up to stickleback right here}
 
-The leading hypothesis is that rather than acting on new mutations, ecological speciation is sped up through selection on standing genetic variation found in marine populations. 
-\plr{Is this ``ecological speciation''? Speciation would imply the ecotypes are developing reproductive incompatibility;
-    I think that we think the pattern is stable. Maybe ``adaptation to freshwater habitats''?}
+Recently, multiple instances of similar (parallel) underlying genetic basis of rapid local adaptation has brought about 
+questions surrounding the origins of adaptive alleles.
+A empirical model is the alaskan populations of freshwater and marine Ninespine Stickleback fish.
+In 1964, The Great Alaskan Earthquake caused an uplift of Middleton island and in turn,
+introduced a group of freshwater ponds around the perimeter of the island. 
+Quickly inhabited by the surrounding marine population of Stickleback,
+\citet{LescakE7204} observed significant phenotypic changes in less than 50 years that appear to be 
+parallel to freshwater stickleback that have been separated for over $13,000$ years.
+In freshwater stickleback, the number of lateral plates are reduced
+and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region.
+\jgg{Bill - more specifics about Bill, Suzie, Kristin  \& or Thom's papers .. ? }
+These results leave us with questions surrounding the nature of rapid adaptation.
+Does convergent evolution breed it's own solution (haplotype) for every new selective pressure, 
+or can these solutions can be efficiently shared across multiple sub - populations facing similar selective pressures.
+
+The leading hypothesis for stickleback is that rather than acting on new mutations, 
+adaptation to freshwater environments is sped up through selection on standing genetic variation (SGV) found in marine populations. 
+\jgg{Bill: maybe you could make this section a little more specific}
 One clear example of this is the gene \textit{eda} 
 which has been shown to regulate the number of lateral plates. 
-While this gene arose millions of years ago, 
-it is found in freshwater ponds which have formed much more recently.
+While this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently.
 Novel evidence from natural populations has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern. \cite{Nelson2017} 
-
-% In 2009, Dolph Schluter and Gina L. Conte 
 \citet{schluter2009genetics} 
-\plr{"citet" for "cite, as part of the text"}
 suggested the ``transporter''-hypothesis.  
-\plr{in latex, use ` `words' ' to get the open- and closed-double quotes (not the double-quote character)}
 This outlined the flow of freshwater alleles into marine populations through offspring of hybridization events.
 It suggests freshwater haplotypes are distributed though marine individuals and are continually selected upon in introduced freshwater populations.
-The continued selection on freshwater favored alleles,
-and introgression between the two sub-species,
+The continued selection on freshwater favored alleles and introgression between the two sub-species,
 allows the marine to maintain the freshwater haplotype dispersed in low frequency among its' individuals. 
-Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection rebuilds the freshwater haplotype. \cite{schluter2009genetics} 
-The alternative to this hypothesis is that individuals from other freshwater environments migrate directly to the new environments, and their haplotype is passed down directly. 
-This hypothesis has been shown to be unlikely due to 
-finding a high frequency of freshwater alleles in the ocean,
+Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection reconstructs the freshwater haplotype. \cite{schluter2009genetics} 
+An alternative to this hypothesis is that individuals from other freshwater environments migrate directly to the new environments, 
+and their haplotype is passed down directly. 
+This hypothesis has been shown to be unlikely due to finding a high frequency of freshwater alleles in the ocean,
 and almost no freshwater individuals.
+
 \plr{Bring up the possibility of more than one freshwater haplotype here.}
+\jgg{Re: should I? Doesn't Thom's work and parallel above suggest they have the same alleles?}
 
-Assuming they were correct about transportation, we might ask; 
-how much migration and gene flow is to be expected for this relationship to be maintained?
-\plr{motivating questions need more work}
-%(Other questions)
-%how does selection, at the genomic level, alter given a range of migration and recombination rates?
-Here, we use forward moving simulations to model marine and freshwater populations of stickleback.
-We then observe the effect (focusing on migration and recombination, for now) 
-of different parameters on local adaptation of introduced freshwater environments. 
-We compare to real data and make predictions about realistic parameters.
+If selection on SGV is key in rapid adaptation, 
+many questions are brought to light concerning the surrounding population genetics parameters and underlying genomic architecture.
+How rapidly can selection act on standing genetic variants at a given value of migration, $M$?
+How many alleles underly a certain trait? (expand here)
+Furthermore, how can we infer causal loci for regions of the genome that must be driving the rapid adaptation, from real data.
+Many biologist today make use of genome wide association studies (GWAS) and 
+$F_{st}$ across the genome to estimate regions responsible for certain traits. 
+Unfortunately, the data can be heavily influenced by the level of introgression and population structure of the samples. 
+This being the case, what are the biases we can expect to see in real data for certain levels of introgression.
 
-\section{Methods}
+There are many different parameters that carry a large impact on the questions above.
+\citet{Ralph2015} 
+explored the dichotomy of selection on new alleles and those brought in by migrants from a similar selective pressure.
+This lead to an understanding of the process by which alleles move from an existing 'patch' to an introduced 'patch' with the 
+same selective pressure. 
+%read though Peter's paper more and connect it to what we're doing here.
 
-Our models are forward-moving evolutionary models which
+In this paper, we explore time to local adaptation as a function of migration rate through an environment (marine) with opposing selective pressure. 
+Next, we compare to real data and explore how $F_{st}$ data can be impacted by introgression, or lack of.
+%Explain what we do and what we find that is important
+
+%We use forward moving simulations to model marine and freshwater populations of stickleback.
+%We then observe the effect (focusing on migration and recombination, for now) 
+%of different parameters on local adaptation of introduced freshwater environments.
+
+\section{Simulation Model}
+
+Our simulations are forward-moving evolutionary models which
 emulate the geography, selective pressure, and genomic architecture 
 of coastal marine and freshwater lake populations. 
-Here we give a brief overview of the software used and the models created to observe the effects of selection. 
-\plr{Should focus on the model, not the software -- describe how the simulation works}
-Each set of simulations was run multiple times to ensure the data we observed was ???
+Here, we describe the specifics of out model.
 
-%(1) Correlation of mutation frequencies at low migration rates
-%(2) Distribution of Effect Size? (run another set to make sure)
+\subsection*{life cycle}
 
-\subsection{SLiM}
-
-To model the stickleback populations, we used a flexible evolutionary framework, SLiM. (cite SLiM 2) 
+For this study, we used a flexible evolutionary framework, SLiM \citep{Haller2017}.
 The model-type we used was based off of an extended wright-fisher. 
 The simulation life cycle of each generation is as follows.
-\plr{This can be merged with the next section, perhaps as a section called ``Life cycle''.  
-    Here and below, avoid using SLiM jargon (don't say ``callback'').}
-First, 
-generate offspring by:
+First, generate offspring by:
 (1) choosing parents based on cached fitness values and migration rates, 
@@
 (4) modify each child by a defined callback. 
-Next, 
-offspring become individuals before fitness is evaluated for the next generation.
-Finally,
-the generation counter is incremented. 
+Next, offspring become individuals before fitness is evaluated for the next generation.
+Finally, the generation counter is incremented. 
 The population structure in SLiM can be arranged in any number of subpopulations, 
@@
 
-\subsection{Model}
-
-\subsubsection*{Geography}
+\subsection*{Geography}
 
@@
 and a set introduced later in the simulation as a subset of the marine. 
-In SLiM, these freshwater sets are defined as just two continuous subpopulations. 
-Using SLiM?s $modifyChild()$ callback, we?ve modeled both to act as 10 separate discrete subpopulations. 
-\plr{no need to say this: we will provide the SLiM recipes as supplementary material
-    so that someone interested can see exactly how it works.}
 Each of the $10$ lakes, $i$, is located along the marine at $i - 0.5$, 
@@
 
-\subsubsection*{Selective Pressure}
+\subsection*{Selective Pressure}
+
+%(TALK ABOUT ADAPTEDNESS and TRADE-OFF), 
+% abstracting real fish phenotype with `` trait value "
+
+%Trait value is equal to the sum of the contribution of the effect loci and the loci may be 
+%add, dom, recc. Uniformly chosen
 
@@
 Freshwater and Marine environments are distinguished by a single numeric value. 
-This value is the optimum phenotype for each environment
+This number is the optimum trait value for each environment
 and acts as a representation of lateral plate number and opercle shape in the stickleback populations.
-\plr{More concisely, these could be called the ``optimal trait values''.}
-The fitness of an individual is then determined by a probability density for a normal distribution 
-at the quantile (?) of the difference between the optimum and the individual's phenotype
-\plr{instead, we can say that we use a ``Gaussian fitness function'', 
-    where the fitness of an individual with trait $x$ in a population with optimal trait value $x_0$ is:}
+%The fitness of an individual is then determined by a probability density for a normal distribution 
+%at the quantile (?) of the difference between the optimum and the individual's phenotype
+We use a Gaussian fitness function, 
+where the fitness of an individual with trait $x$ in a population with optimal trait value $x_0$ is:
 \begin{align}
@@
 
+Where a trait value, $x$, is also defined by a single numeric value; 
+determined as a summation of the selection coefficients of all mutations an individual possesses. 
+% no empty lines before equations, usually - they should read like part of the text itself.
 %	\[
-% \textnormal{Individual.Fitness = } X \sim \mathcal{N}(\textnormal{Optimum - Individual.Phenotype},\,0.0\,15.0).
+%{\textnormal{Individual.Phenotype = } \sum_{m \in mutations} \textnormal{m.selectionCoeff}}
 %	\]
-
-Where an individual?s phenotype is also defined by a single numeric value; 
-determined as a summation of the selection coefficients of all mutations an individual possesses. 
-\plr{Hm. Should we say ``phenotype'' or ``trait''? I've been saying ``trait'', which is less jargon-ey?}
-% no empty lines before equations, usually - they should read like part of the text itself.
-	\[
-{\textnormal{Individual.Phenotype = } \sum_{m \in mutations} \textnormal{m.selectionCoeff}}
-	\]
-
-\todo{This isn't quite right because of how we calculate additive, recessive, and dominant}
 
@@
 Among the effect mutations; 
-\plr{``effect mutation'' is not a standard term; should say ``mutations that affect the phenotype'' here,
-    and be careful about what we call them elsewhere.}
-there are additive, dominant, and recessive mutations which effect phenotype
+there are additive, dominant, and recessive mutations which a ffect phenotype
 in either the positive or negative direction. 
-The selection coefficients are pulled from a beta distribution 
+The selection coefficients are pulled from a gamma distribution 
 with a shape parameter of $1$ and a mean of either $0.5$, or ${-0.5}$. 
@@
     Also, a Gamma with shape 1 and mean 0.5 is an Exponential with mean 0.5.}
-	\[
-\textnormal{Mutation.selectionCoeff = } X \sim \Gamma (1,\pm0.5)
-	\]
+
+%	\[
+%\textnormal{Mutation.selectionCoeff = } X \sim \Gamma (1,\pm0.5)
+%	\]
 
@@
 
-\subsubsection*{Genomic Architecture}
+\subsection*{Genomic Architecture}
 
@@
 To mimic this and compare dynamics of neutral vs. effect under selection, 
-we create regions of effect. 
+we create regions of effect. (???)
 In these genomic elements of effect, we uniformly allow additive, recessive, and dominant, mutation that can push phenotype in either the positive or negative direction to arise. 
@@
 The genome has a recombination rate of $10^{-5}$, which is approximately 1 breakpoint event, per individual, per generation.
-
-\todo{finish this}
 
@@
 Throughout the simulation, we measure;
-mutation frequencies, 
+%mutation frequencies, 
 $F_{ST}$ values, \plr{how is $F_{ST}$ computed?}
@@
 \plr{Define ``mean percentage''; no need to say that it is the sum of the frequencies, I think.}
+%word this out better. 
 \plr{Here FAA are defined in terms of the original populations, explaining why the introduced populations can have MPFAI of 0.
@@
 \subsection{Local Adaptation: differentiation with gene flow}
+
+%Talk about the genetic basis of trait. 
+%Talk about the progression of adaptation. How you can see introduction new alleles
 
@@
 As can be seen in Figure \ref{fig:TimeToAdaptation} at the lowest migration rate of $5 * 10^{-5}$,
-It took the introduced population over 30 thousand generations for the average phenotype of all the lakes to 
+It took the introduced population over 20 thousand generations for the average phenotype of all the lakes to 
 get to within 0.5 of the original lakes original phenotypes. 
@@
   		\caption{ 
+		%SP - suppose you have an Fst Peak, how likely is it that that region is causal. 
 		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
@@
 \end{figure}
+
+%Suppose you have a Fst Peak, ___ how likely is it that that region is causal. and conversely, what proportion of 
+%the causal loci are under Fst Peaks.
+
+%
 
@@
 %\subsubsection*{Low Recombination Causes Clustering}
-%\begin{comment}
+\begin{comment}
 \begin{figure}[h!tb]
@@
 \end{figure}
-%\end{comment}
+\end{comment}
 

### 2018-07-09 Stickleback_Paper.tex
latest

@@
 \usepackage{xcolor}
+%\documentclass{article}
+%\usepackage{subcaption}
+%\documentclass[smallextended]{svjour3}
+%\usepackage{lipsum}% Just for this example
 
@@
 
-effect loci | QTL | causal: = ??
+effect loci (locus) | QTL | causal: = ?
 
-Original Lakes | $P_{1}$, $P_{3}$, $P_{3}$ = ?
+Original Lakes | initial | $P_{1}$, $P_{3}$, $P_{3}$ = ?
 
 Migration Rate | Introgression = ?
+
+pre-existing freshwater adapted alleles | standing genetic variants = ?
 
@@
 If selection on SGV is key in rapid adaptation, 
-many questions are brought to light concerning the surrounding population genetics parameters and underlying genomic architecture.
+many questions are exposed concerning the surrounding population genetics parameters and underlying genomic architecture.
 How rapidly can selection act on standing genetic variants at a given value of migration, $M$?
@@
 In this paper, we explore time to local adaptation as a function of migration rate through an environment (marine) with opposing selective pressure. 
-Next, we compare to real data and explore how $F_{st}$ data can be impacted by introgression, or lack of.
+Next, we compare to real data and explore how $F_{st}$ data can be impacted by introgression.
 %Explain what we do and what we find that is important
@@
 of coastal marine and freshwater lake populations. 
-Here, we describe the specifics of out model.
+Here, we describe the specifics and parameter values of our model.
 
@@
 
-For this study, we used a flexible evolutionary framework, SLiM \citep{Haller2017}.
-The model-type we used was based off of an extended wright-fisher. 
+For this study, we used an evolutionary framework, SLiM \citep{Haller2017}.
+The model-type we used was derived from an extended wright-fisher. 
 The simulation life cycle of each generation is as follows.
@@
 (4) modify each child by a defined callback. 
-Next, offspring become individuals before fitness is evaluated for the next generation.
-Finally, the generation counter is incremented. 
+Next, offspring become individuals.
+And finally, fitness is evaluated for the next generation
+and the generation counter is incremented.
+
 The population structure in SLiM can be arranged in any number of subpopulations, 
 continuous or discrete, connected by any rate of migration. 
+Individual's (diploidy) genomes are represented by a linear array of loci. 
+Each locus is positionally defined with recombination and mutation rates,
+as well as mutation types allowed to arise.
 
@@
 
+\jgg{do we need this section or should we just give a really good description in the geography diagram?}
 What we've modeled geographically is a coastal marine population 
@@
 The population structure for the marine is modeled as a one-dimensional, continuous population
-ranging from 0.0 to 10.0.
+ranging from [0.0 to 10.0].
 There are two sets of lakes which represent freshwater populations; 
@@
   		\caption{A representation of the geographic and evolutionary history of all populations throughout the simulation. 
-		The marine is a one-dimensional, continuous population with spatial positions ranging from [0.0 - 10.0]. Each 
-		lake$_{i}$ is a discrete population connected by migration only through the marine at position $i - 0.5$. The introduced lakes 
-		have the same location and selective pressure as the original lakes, but arise at 40K generations.
+		The marine is a one-dimensional, continuous population with spatial positions ranging from [0.0 - 10.0]. 
+		Each freshwater lake$_{i}$ for both the introduced and freshwater populations
+		is a discrete population connected by migration only through the marine, at position $i - 0.5$. 
+		The introduced lakes have the same spatial location (but separate?) and selective pressure as the original lakes, but arise at 40K generations.
+		The marine has $2000$ diploid individuals while each of the lakes fluctuate around $200$. 
 		the introduced population is initialized as a copy of all marine individuals to model marine 
@@
 
-\subsection*{Selective Pressure}
-
-%(TALK ABOUT ADAPTEDNESS and TRADE-OFF), 
-% abstracting real fish phenotype with `` trait value "
-
-%Trait value is equal to the sum of the contribution of the effect loci and the loci may be 
-%add, dom, recc. Uniformly chosen
-
-To emulate the freshwater and marine selective pressure, 
-we set up a quantitative genetics model in which fitness is phenotypically based. 
-Freshwater and Marine environments are distinguished by a single numeric value. 
-This number is the optimum trait value for each environment
-and acts as a representation of lateral plate number and opercle shape in the stickleback populations.
-%The fitness of an individual is then determined by a probability density for a normal distribution 
-%at the quantile (?) of the difference between the optimum and the individual's phenotype
-We use a Gaussian fitness function, 
-where the fitness of an individual with trait $x$ in a population with optimal trait value $x_0$ is:
-\begin{align}
-    f(x; x_0) &= \exp\left(\frac{(x - x_0)^2 }{ 2 \sigma^2}\right),
-\end{align}
-with $\sigma = 15$.
-
-Where a trait value, $x$, is also defined by a single numeric value; 
-determined as a summation of the selection coefficients of all mutations an individual possesses. 
-% no empty lines before equations, usually - they should read like part of the text itself.
-%	\[
-%{\textnormal{Individual.Phenotype = } \sum_{m \in mutations} \textnormal{m.selectionCoeff}}
-%	\]
-
-We allow six mutation types which affect phenotype along with one neutral mutation type.
-Among the effect mutations; 
-there are additive, dominant, and recessive mutations which a ffect phenotype
-in either the positive or negative direction. 
-The selection coefficients are pulled from a gamma distribution 
-with a shape parameter of $1$ and a mean of either $0.5$, or ${-0.5}$. 
-\plr{Can just say this in words, no need to have an equation.
-    Also, a Gamma with shape 1 and mean 0.5 is an Exponential with mean 0.5.}
-
-%	\[
-%\textnormal{Mutation.selectionCoeff = } X \sim \Gamma (1,\pm0.5)
-%	\]
-
-\todo{maybe the order of these should be reversed?}
-\plr{I agree: first explain how phenotypes are calculated, then say what fitness is.
-    Notation suggestion: phenotype for an individual that is heterozygous at loci $H$
-    and homozygous derived at loci $D$ is $x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
-    where $h_\ell$ is the dominance coefficient for locus $\ell$ (0 for recessive, 1 for dominant, 0.5 for additive),
-    and $s_\ell$ is the effect of the derived mutation on the phenotype}
-
 \subsection*{Genomic Architecture}
 
-\plr{This should maybe come before the phenotype.}
 In SLiM, the genome is conceptually a linear array of loci defining locations for 
@@
 
-\subsection{Sampling}
-\plr{I think the purpose of this section is to define terms and statistics used later;
-    so maybe a better title is ``Descriptive statistics''?}
+\subsection*{Selective Pressure}
+
+%(TALK ABOUT ADAPTEDNESS and TRADE-OFF), 
+% abstracting real fish phenotype with `` trait value "
+
+%Trait value is equal to the sum of the contribution of the effect loci and the loci may be 
+%add, dom, recc. Uniformly chosen
+
+To emulate the freshwater and marine selective pressure, 
+we set up a quantitative genetics model in which fitness is phenotypically based. 
+phenotype for an individual that is heterozygous at locus $H$
+and homozygous derived at loci $D$ is $x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
+where $h_\ell$ is the dominance coefficient for locus $\ell$ (0 for recessive, 1 for dominant, 0.5 for additive),
+and $s_\ell$ is the effect of the derived mutation on the phenotype
+
+Freshwater and Marine environments are distinguished by a single numeric value. 
+This number is the optimum trait value for each environment
+and acts as a representation of lateral plate number and opercle shape in the stickleback populations.
+%The fitness of an individual is then determined by a probability density for a normal distribution 
+%at the quantile (?) of the difference between the optimum and the individual's phenotype
+We use a Gaussian fitness function, 
+where the fitness of an individual with trait $x$ in a population with optimal trait value $x_0$ is:
+\begin{align}
+    f(x; x_0) &= \exp\left(\frac{(x - x_0)^2 }{ 2 \sigma^2}\right),
+\end{align}
+with $\sigma = 15$.
+
+%Where a trait value, $x$, is also defined by a single numeric value; 
+%determined as a summation of the selection coefficients of all mutations an individual possesses.
+
+%	\[
+%{\textnormal{Individual.Phenotype = } \sum_{m \in mutations} \textnormal{m.selectionCoeff}}
+%	\]
+
+%We allow six mutation types which affect phenotype along with one neutral mutation type.
+%Among the effect mutations; 
+%there are additive, dominant, and recessive mutations which a ffect phenotype
+%in either the positive or negative direction. 
+%The selection coefficients are pulled from a gamma distribution 
+%with a shape parameter of $1$ and a mean of either $0.5$, or ${-0.5}$.
+
+%	\[
+%\textnormal{Mutation.selectionCoeff = } X \sim \Gamma (1,\pm0.5)
+%	\]
+
+%\plr{I agree: first explain how phenotypes are calculated, then say what fitness is.
+%    Notation suggestion: phenotype for an individual that is heterozygous at loci $H$
+%    and homozygous derived at loci $D$ is $x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
+%    where $h_\ell$ is the dominance coefficient for locus $\ell$ (0 for recessive, 1 for dominant, 0.5 for additive),
+%    and $s_\ell$ is the effect of the derived mutation on the phenotype}
+
+\subsection*{Descriptive statistics}
+%\plr{I think the purpose of this section is to define terms and statistics used later;
+%    so maybe a better title is ``Descriptive statistics''?}
 
@@
 impact the sharing of alleles between populations. 
-\plr{Below, I think we should make sure to use ``population'' to mean either ``lake'' or ``ocean'';
-    so when we compute freshwater-marine $F_{ST}$ this measures differentiation between habitats,
-    not between populations.}
-we define freshwater adapted alleles (FAA),
-\plr{I vote to remove the abbreviations. Also, should make the dependence on generation more clear: maybe ``freshwater adapted alleles at time $t$''?} 
-at any given generation,
+%\plr{Below, I think we should make sure to use ``population'' to mean either ``lake'' or ``ocean'';
+%    so when we compute freshwater-marine $F_{ST}$ this measures differentiation between habitats,
+%    not between populations.}
+We define pre-existing freshwater adapted alleles at time $t$,
+%\plr{I vote to remove the abbreviations. Also, should make the dependence on generation more clear: maybe ``freshwater adapted alleles at time $t$''?} 
 to be a mutations with a frequency higher than $0.5$ in \textit{any} of the original lakes, 
 while remaining lower than $0.5$ in the marine. 
-This is because the transportation hypothesis does not specify where or when an advantageous mutation arises; 
-but simply suggests that when one comes to high enough frequency, it too, could participate in the transportation process. \cite{schluter2009genetics}
+They are defined this way because the transportation hypothesis does not specify where or when an advantageous mutation arises,
+but simply suggests that when a freshwater adapted alleles comes to high enough frequency, it too, could participate in the transportation process \cite{schluter2009genetics}.
 To observe transportation of freshwater alleles in our simulations, we use a variety of metrics when sampling.
-Throughout the simulation, we measure;
-%mutation frequencies, 
-$F_{ST}$ values, \plr{how is $F_{ST}$ computed?}
-average phenotype, 
-mean number of original lakes each FAA appears in,
-and total number of unique FAA. 
-to understand where, as well as how much, of these freshwater haplotypes are being transported,
-we track the mean percentage of FAA per individual (MPFAI) across all populations throughout the simulation:
-computed as the sum of the FAA mutation frequencies among each population.
-\plr{Define ``mean percentage''; no need to say that it is the sum of the frequencies, I think.}
-%word this out better. 
-\plr{Here FAA are defined in terms of the original populations, explaining why the introduced populations can have MPFAI of 0.
-    This is counter-intuitive from the name (FAA); maybe we should call them ``pre-existing freshwater alleles''?
-    ``standing genetic variants''?}
-we're particularly interested in using MPFAI to measure the marine population' capacity to carry FAA,
-as well as observing the introduced populations' ability to select upon them. 
-\plr{omit this last sentence?}
+let $P_{1}$ and $P_{2}$ be the vector of mutation frequencies in the first and second populations, respectively.
+with $\hat{P} = (P_{1} + P_{2}) / 2$ and
+$PQ = (P_{1} \times (1 - P_{1}) + P_{2} \times (1 - P_{2})) / 2$,
+we compute $F_{ST} = 1 - (PQ / (\hat{P} \times (1 - \hat{P})))$.
+%mean number of original lakes each FAA appears in,
+%and total number of unique FAA. 
+Time to adaptation ($T_{adapt}$) of the introduced population is the generation at which
+the difference between the average phenotype of the original and the introduced freshwater populations is less than 0.5. 
+To understand where, as well as how much, of these pre-existing freshwater adapted alleles are being transported,
+we track the the average percentage of freshwater adapted alleles at time $t$ across all populations throughout the simulation.
+%\plr{Here FAA are defined in terms of the original populations, explaining why the introduced populations can have MPFAI of 0.
+%    This is counter-intuitive from the name (FAA); maybe we should call them ``pre-existing freshwater alleles''?
+%    ``standing genetic variants''?}
+%we're particularly interested in using MPFAI to measure the marine population' capacity to carry FAA,
+%as well as observing the introduced populations' ability to select upon them. 
+%\plr{omit this last sentence?}
+%Because we are interested in the local adaptation of the introduced freshwater populations,
 
-Because we are interested in the local adaptation of the introduced freshwater populations,
-we define time to adaptation ($T_{adapt}$) of the introduced population to be the generation at which
-the difference between the average phenotype of the original and the introduced freshwater populations is less than 0.5. 
-At this generation, we look at:
-the correlation (B. Pearson) between the mutation frequencies of all effect loci and FAA,
-and the number of shared high frequency alleles. 
-\plr{Here we need to define terms; we don't need to say exactly what was computed,
-    as that will be redundant with the Results.}
-These metrics tells us how much of the FAA  were used in the lakes once they have locally adapted.
+%At this generation, we look at:
+%the correlation (B. Pearson) between the mutation frequencies of all effect loci and FAA,
+%and the number of shared high frequency alleles. 
+%\plr{Here we need to define terms; we don't need to say exactly what was computed,
+%    as that will be redundant with the Results.}
+%These metrics tells us how much of the FAA  were used in the lakes once they have locally adapted.
 
 \section{Results}
+
+Across a range of introgression from $\approx 0.01 - 100.0$ migrants per generation in populations of 
+size 2000 diploid individuals, we encountered large influence on 
+time to adaptation,
+sharing of pre-existing freshwater adapted alleles,
+and population genetic signals left behind in the simulated populations.
+A larger amount of introgression from all lakes to the marine population helped both the initial and introduced freshwater populations 
+share pre-existing freshwater adapted alleles.
+this resulted in rapid adaptation ($\approx 100$ generations) of the introduced population to adapt to the freshwater environment.
+We also found that larger amounts of migration allow for more statistical power and less false positives in the resulting population genetics data ($F_{st}$ per SNP across the genome).
 
@@
 
-\subsection{Local Adaptation: differentiation with gene flow}
+\subsection*{Local Adaptation: differentiation with gene flow}
 
@@
 Local adaptation occurred across all simulated parameter values.
-Starting from the same baseline, freshwater and marine populations diverged phenotypically, 
-until they reached an equilibrium,
+Starting from the same baseline, freshwater and marine populations diverged phenotypically
+until they reached an equilibrium
 where the population means were close to or at the optimal phenotype. 
@@
 Polymorphic loci that affected the phenotype,
-showed greater differentiation compared to neutral loci between marine and freshwater populations, with the exception 
-of migration rate at a value of $5e-5$. 
-$F_{st}$ plots along the genome, show peaks at loci underlying local adaptation.
-\plr{Rather than referring to e.g., ``migration rate of $5 \times 10^{-5}$'',
-    it'd be easier to follow if we had names for the different migration rate situations.
-    Maybe ``low, medium-low, medium-high, and high''?
-    Or talk about the number of migrants per generation per lake, e.g.,
-    ``with the exception of the lowest migration rate of 0.01 migrants per generation per lake''?}
+showed greater differentiation compared to neutral loci between marine and freshwater populations, 
+with the exception of the lowest migration rate of 0.01 migrants per generation per lake. 
+%$F_{st}$ plots along the genome, show peaks at loci underlying local adaptation.
+
+%\plr{Rather than referring to e.g., ``migration rate of $5 \times 10^{-5}$'',
+%    it'd be easier to follow if we had names for the different migration rate situations.
+%    Maybe ``low, medium-low, medium-high, and high''?
+%    Or talk about the number of migrants per generation per lake, e.g.,
+%    ``with the exception of the lowest migration rate of 0.01 migrants per generation per lake''?}
 
@@
     so that when someone wonders ``What was the population size again?'' they could easily find it.}
+\jgg{Re: agreed}
 
-\plr{I vote to put figures \ref{fig:phenotype_ts2} and \ref{fig:phenotype_ts3} 
-    together as one figure and put them here.
-    There's a lot in those we could talk about: you can see the stepwise introduction
-    of individual alleles in figure \ref{fig:phenotype_ts2}, for instance.}
+%\plr{I vote to put figures \ref{fig:phenotype_ts2} and \ref{fig:phenotype_ts3} 
+%    together as one figure and put them here.
+%    There's a lot in those we could talk about: you can see the stepwise introduction
+%    of individual alleles in figure \ref{fig:phenotype_ts2}, for instance.}
 
-\subsection{The Effect of Migration Rate}
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
+  		\caption{ Average phenotype in each population as a function of time $t$, in generations. 
+		This is a single simulation run at a migration rate of $5 \times 10^{-4} \approx 1$ mig/gen. 
+		The dark black line represents the mean phenotype of the single marine population with an optimum trait value of + 10. 
+		The pink represents the entire initial freshwater population mean, 
+		as a sum of all the lakes (light green) with and optimum trait value of -10.
+		The orange represents the entire introduced freshwater population mean
+		as a sum of all the lakes (dark green) with an optimum trait value of -10 as well.
+		}
+  		\label{fig:phenotype_ts2}
+	\end{center}
+\end{figure}
 
-\subsubsection*{Migration is Mixing}
-\plr{Title: ``Migration homogenizes genetic variation''}
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
+  		\caption{Average phenotype in each population as a function of time $t$, in generations. 
+		This is a single simulation run at a migration rate of $5 \times 10^{-3} \approx 1$ mig/gen. 
+		The dark black line represents the mean phenotype of the single marine population with an optimum trait value of + 10. 
+		The pink represents the entire initial freshwater population mean, 
+		as a sum of all the lakes (light green) with and optimum trait value of -10.
+		The orange represents the entire introduced freshwater population mean
+		as a sum of all the lakes (dark green) with an optimum trait value of -10 as well.
+		}
+  		\label{fig:phenotype_ts3}
+	\end{center}
+\end{figure}
+
+\jgg{we should make one caption and the two above should be top-bottom sub-figures}
+\jgg{also, interesting point to make on second plot, you can see the initial lakes taking the same size step. sharing}
+%I could zoom in on the relevent portions of the figure for this. (begging and introduction)
+% a 2 by 2 with the important sections (0 - 20K) $ (40-50K) for both above figures. share axes
+
+\jgg{finished with most edits up to here}
+
+\subsection*{The Effect of Migration Rate}
+
+\subsubsection*{Migration homogenizes genetic variation}
 
@@
 upon by selection, in all populations.
-Additionally, standing genetic variance \todo{computed as?}
+Additionally, standing genetic variance
 (seen in Figure \ref{fig:SGV})
 values in the marine population steady increased with migration which suggests that more alleles from the 
-lakes surrounding were exporting alleles into the marine individuals.
-\todo{Something about SGV rates for the the other population}
+lakes surrounding were exporting alleles into the marine individual.
 
@@
 get to within 0.5 of the original lakes original phenotypes. 
-\plr{From the plots it looks more like adaptation happens (at least in most lakes) by like 10K or 15K.}
+%\plr{From the plots it looks more like adaptation happens (at least in most lakes) by like 10K or 15K.}
 This suggests that selection was acting primarily on new mutations, meaning the 
@@
 it was selected upon. 
-\plr{We should do the calculation to see if this makes sense: 
-    there's one new effect mutation in the right direction every 100 generations in the lakes,
-    one of effect size $s$ has probability of about $s/2$ of not being lost to drift...}
+%\plr{We should do the calculation to see if this makes sense: 
+%    there's one new effect mutation in the right direction every 100 generations in the lakes,
+%    one of effect size $s$ has probability of about $s/2$ of not being lost to drift...}
 There was a positive correlation between the amount of standing genetic variation in the 
@@
 		\label{fig:MPFAI}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=\linewidth]{matplotlibPlots/MFAI.png}
-  		\caption{Distributions of mean number of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
-		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
-		freshwater adapted alleles.
-		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
-		}
-  		\label{fig:MNFAI}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=\linewidth]{plotlyPlots/NumFAA.png}
-  		\caption{(PLACEHOLDER)}
-		\label{fig:NumFAA}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=\linewidth]{plotlyPlots/StandingGeneticVariation.png}
-  		\caption{(PLACEHOLDER)}
-		\label{fig:SGV}
 	\end{center}
@@
 
-\subsection{Genomic Architecture}
+\subsection*{Population genetics data signals}
+
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3.png}
+  		\caption{(Decription and cleaning needed)
+		}
+  		\label{fig:Fst3}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-4.png}
+  		\caption{(Description and cleaning needed)
+		}
+  		\label{fig:Fst2}
+	\end{center}
+\end{figure}
 
@@
 %\subsubsection*{Low Recombination Causes Clustering}
-\begin{comment}
+
+\section{Conclusion \& Discussion}
+
+% at low M we should be able to predict something ---
+% at high M we should be able to predict something ----
+
+%Here, we suggest a range of migration rate (introgression) parameter values which would allow for the rapid (and in turn, parallel) adaptation 
+%of marine stickleback introduced intro a freshwater environment.
+%While this range is heavily effected by all other selections of parameter values.
+
+%(1) There's a threshold of introgression for both rapid adaptation and migration load. 
+\subsection*{thresholds}
+
+%(2) Do the math, discuss number ( and possibly type ) of mutations underlying the trait value.
+\subsection*{do the math}
+
+%(3) discuss what signals researchers could possibly look out for or connect to real data
+\subsection*{signals and real data}
+
+\clearpage
+%\plr{some problems because you had bibliographystyle before bibliography}
+\bibliographystyle{plainnat}
+\bibliography{Citations}{}
+
+\section{Supp.}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=\linewidth]{plotlyPlots/NumFAA.png}
+  		\caption{(PLACEHOLDER)}
+		\label{fig:NumFAA}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=\linewidth]{plotlyPlots/StandingGeneticVariation.png}
+  		\caption{(PLACEHOLDER)}
+		\label{fig:SGV}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=\linewidth]{matplotlibPlots/MFAI.png}
+  		\caption{Distributions of mean number of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
+		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
+		freshwater adapted alleles.
+		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
+		}
+  		\label{fig:MNFAI}
+	\end{center}
+\end{figure}
+
 \begin{figure}[h!tb]
@@
   		\label{fig:Fst1}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-4.png}
-  		\caption{(PLACEHOLDER)
-		}
-  		\label{fig:Fst2}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3.png}
-  		\caption{(PLACEHOLDER)
-		}
-  		\label{fig:Fst3}
 	\end{center}
@@
 	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-5.png}
-  		\caption{ (PLACEHOLDER)
-		}
-  		\label{fig:Haplo1}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-4.png}
-  		\caption{(PLACEHOLDER)
-		}
-  		\label{fig:Haplo2}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-3.png}
-  		\caption{(PLACEHOLDER)
-		}
-  		\label{fig:Haplo3}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-2.png}
-  		\caption{(PLACEHOLDER)
-		}
-  		\label{fig:Haplo4}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
   		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-5.png}
@@
   		\label{fig:phenotype_ts1}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
-  		\caption{(PLACEHOLDER)
-		}
-  		\label{fig:phenotype_ts2}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
-  		\caption{(PLACEHOLDER)
-		}
-  		\label{fig:phenotype_ts3}
 	\end{center}
@@
 \end{figure}
-\end{comment}
-
-\section{Conclusion \& Discussion}
-
-% at low M we should be able to predict something ---
-% at high M we should be able to predict something ----
-
-Here, we suggest a range of migration rate (introgression) parameter values which would allow for the rapid (and in turn, parallel) adaptation 
-of marine stickleback introduced intro a freshwater environment.
-While this range is heavily effected by all other selections of parameter values.
-
-\newpage
-\plr{some problems because you had bibliographystyle before bibliography}
-\bibliographystyle{plainnat}
-\bibliography{Citations}{}
 

### 2018-07-11 Stickleback_Paper.tex
Finished cleaning up results

@@
 
-Migration Rate | Introgression = ?
+Migration Rate | Historical Introgression = ?
 
@@
 
-If selection on SGV is key in rapid adaptation, 
+If selection on standing genetic variants is key in rapid adaptation, 
 many questions are exposed concerning the surrounding population genetics parameters and underlying genomic architecture.
-How rapidly can selection act on standing genetic variants at a given value of migration, $M$?
-How many alleles underly a certain trait? (expand here)
+%How rapidly can selection act on standing genetic variants at a given value of migration, $M$?
+What scale of historical introgression between populations with selective pressures $X$ and $Y$, allows for rapid and parallel local adaptation
+of a population derived from $X$ to an introduced environment with selective pressure $Y$?
+What is the origin of pre-existing adaptive variants?
+How are the variants structured, historically and across the genome underlying the trait?
+%How does population genetics data influenced by historical introgression and selection on standing genetic variants?
+%How many alleles underly a certain trait and how many are swept away by drift? 
 Furthermore, how can we infer causal loci for regions of the genome that must be driving the rapid adaptation, from real data.
-Many biologist today make use of genome wide association studies (GWAS) and 
+Many biologists today make use of genome wide association studies (GWAS) and 
 $F_{st}$ across the genome to estimate regions responsible for certain traits. 
@@
 explored the dichotomy of selection on new alleles and those brought in by migrants from a similar selective pressure.
-This lead to an understanding of the process by which alleles move from an existing 'patch' to an introduced 'patch' with the 
-same selective pressure. 
-%read though Peter's paper more and connect it to what we're doing here.
+Answering the fundamental questions surrounding spatial resolution of convergent evolution. 
+%This lead to an understanding of the process by which alleles move from an existing 'patch' to an introduced 'patch' with the 
+%same selective pressure. 
+\jgg{It seems like this could be a good parallel to what we're doing here, or maybe for discussion?}
 
@@
 They are defined this way because the transportation hypothesis does not specify where or when an advantageous mutation arises,
-but simply suggests that when a freshwater adapted alleles comes to high enough frequency, it too, could participate in the transportation process \cite{schluter2009genetics}.
+but simply suggests that when a freshwater adapted alleles comes to high enough frequency, it too, could participate in the transportation process \citep{schluter2009genetics}.
 To observe transportation of freshwater alleles in our simulations, we use a variety of metrics when sampling.
@@
 
-Across a range of introgression from $\approx 0.01 - 100.0$ migrants per generation in populations of 
+Across a range of historical introgression from $\approx 0.01 - 100.0$ migrants per generation in populations of 
 size 2000 diploid individuals, we encountered large influence on 
@@
 and population genetic signals left behind in the simulated populations.
-A larger amount of introgression from all lakes to the marine population helped both the initial and introduced freshwater populations 
+More gene flow is experienced with greater historical introgression between the two selective pressures.
+Interestingly, the ability for separate populations to locally adapt to their own selective pressure was relatively unaffected 
+until the highest rate of migration between the marine and freshwater environments.
+All rates of historical introgression aside from the lowest, helped both the initial and introduced freshwater populations 
 share pre-existing freshwater adapted alleles.
-this resulted in rapid adaptation ($\approx 100$ generations) of the introduced population to adapt to the freshwater environment.
+The sharing of alleles resulted in rapid adaptation ($\approx 100$ generations) of the introduced population (split from the marine population) to adapt to the freshwater environment.
 We also found that larger amounts of migration allow for more statistical power and less false positives in the resulting population genetics data ($F_{st}$ per SNP across the genome).
 
-%We found this
-%--justify it
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
+  		\caption{ Mean individual trait value in each population as a function of time $t$, in generations. 
+		This is a single simulation run at a migration rate of $5 \times 10^{-4} \approx 1$ mig/gen in both 
+		directions for every freshwater population to the single marine population. 
+		The dark black line represents the mean phenotype of the single marine population with an optimum trait value of + 10. 
+		The pink represents the entire initial freshwater population mean
+		of all the lakes (light green) with and optimum trait value of -10.
+		The orange represents the entire introduced freshwater population mean
+		of all the lakes (dark green) with an optimum trait value of -10 as well.
+		}
+  		\label{fig:phenotype_ts2}
+	\end{center}
+\end{figure}
 
-% A LIST OF PLOTS
-% GENERAL OBSERVATIONS
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
+  		\caption{Mean individual trait value in each population as a function of time $t$, in generations. 
+		This is a single simulation run at a migration rate of $5 \times 10^{-3} \approx 1$ mig/gen in both 
+		directions for every freshwater population to the single marine population. 
+		The dark black line represents the mean phenotype of the single marine population with an optimum trait value of + 10. 
+		The pink represents the entire initial freshwater population mean
+		of all the lakes (light green) with and optimum trait value of -10.
+		The orange represents the entire introduced freshwater population mean
+		of all the lakes (dark green) with an optimum trait value of -10 as well.
+		}
+  		\label{fig:phenotype_ts3}
+	\end{center}
+\end{figure}
 
-% [1]. Fst Neutral 
-% [2]. Fst Effect 
-% [3]. Fst Across The 
-% [4]. Median Number of FAA Per individual / Total Number of FAA
-% [5]. Phenotype Distribution
-% [6]. Percentage of shared High Frequency Alleles Shared Between Original and Introduced
-% [7]. Correlation of Allele Frequency 
-% [8]. Time until Adaptation
-% [9]. Total Num FWAA - Average number of lakes each allele appears in.
-% [10]. Distribution of # of FWAA Per marine individual.
+\jgg{we should make one caption and the two above should be top-bottom sub-figures}
+\jgg{also, interesting point to make on second plot, you can see the initial lakes taking the same size step. sharing}
+%I could zoom in on the relevent portions of the figure for this. (begging and introduction)
+% a 2 by 2 with the important sections (0 - 20K) $ (40-50K) for both above figures. share axes
+%\jgg{finished with most edits up to here}
 
@@
 where the population means were close to or at the optimal phenotype. 
-phenotypic variation within each population was small compared to the difference between populations.
-Polymorphic loci that affected the phenotype,
+The distribution of mean number of freshwater adapted alleles suggested that there was 
+$\approx 8$ SNPs underlying the freshwater phenotype. 
+This was true for all rates of introgression except for the highest at $100$ mig/gen
+where the mean trait value in the populations equilibrated well below the optimum.
+We can see the introduction of new alleles in Figure \ref{fig:phenotype_ts2}. 
+In generation (0 - 20K) The sweeps that cause jumps in the mean phenotype of the freshwater population 
+have a relatively large effect size equal to or greater than $1$.
+%In Figure \ref{fig:counts} we can see that the number of effect regions with zero 
+%pre-existing freshwater adapted alleles increased with introgression while
+%regions of effect containing more pre-existing freshwater adapted alleles increase.
+Phenotypic variation within each population was small compared to the difference between populations.
+Polymorphic loci that impact the individual trait value,
 showed greater differentiation compared to neutral loci between marine and freshwater populations, 
 with the exception of the lowest migration rate of 0.01 migrants per generation per lake.
-%$F_{st}$ plots along the genome, show peaks at loci underlying local adaptation.
-
-%\plr{Rather than referring to e.g., ``migration rate of $5 \times 10^{-5}$'',
-%    it'd be easier to follow if we had names for the different migration rate situations.
-%    Maybe ``low, medium-low, medium-high, and high''?
-%    Or talk about the number of migrants per generation per lake, e.g.,
-%    ``with the exception of the lowest migration rate of 0.01 migrants per generation per lake''?}
 
 \todo{ add more general observations and caveats about not all parameter values. }
-
 \plr{It would be nice to have a small table of parameter values,
@@
 
-%\plr{I vote to put figures \ref{fig:phenotype_ts2} and \ref{fig:phenotype_ts3} 
-%    together as one figure and put them here.
-%    There's a lot in those we could talk about: you can see the stepwise introduction
-%    of individual alleles in figure \ref{fig:phenotype_ts2}, for instance.}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
-  		\caption{ Average phenotype in each population as a function of time $t$, in generations. 
-		This is a single simulation run at a migration rate of $5 \times 10^{-4} \approx 1$ mig/gen. 
-		The dark black line represents the mean phenotype of the single marine population with an optimum trait value of + 10. 
-		The pink represents the entire initial freshwater population mean, 
-		as a sum of all the lakes (light green) with and optimum trait value of -10.
-		The orange represents the entire introduced freshwater population mean
-		as a sum of all the lakes (dark green) with an optimum trait value of -10 as well.
-		}
-  		\label{fig:phenotype_ts2}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
-  		\caption{Average phenotype in each population as a function of time $t$, in generations. 
-		This is a single simulation run at a migration rate of $5 \times 10^{-3} \approx 1$ mig/gen. 
-		The dark black line represents the mean phenotype of the single marine population with an optimum trait value of + 10. 
-		The pink represents the entire initial freshwater population mean, 
-		as a sum of all the lakes (light green) with and optimum trait value of -10.
-		The orange represents the entire introduced freshwater population mean
-		as a sum of all the lakes (dark green) with an optimum trait value of -10 as well.
-		}
-  		\label{fig:phenotype_ts3}
-	\end{center}
-\end{figure}
-
-\jgg{we should make one caption and the two above should be top-bottom sub-figures}
-\jgg{also, interesting point to make on second plot, you can see the initial lakes taking the same size step. sharing}
-%I could zoom in on the relevent portions of the figure for this. (begging and introduction)
-% a 2 by 2 with the important sections (0 - 20K) $ (40-50K) for both above figures. share axes
-
-\jgg{finished with most edits up to here}
-
-\subsection*{The Effect of Migration Rate}
-
-\subsubsection*{Migration homogenizes genetic variation}
+\subsection*{Migration homogenizes genetic variation}
 
@@
 $F_{st}$ values for neutral alleles steadily decline between all sets of populations as 
-migration increases, this illustrates less difference between sections of the genome which are not acted up 
+migration increases, this illustrates less difference between sections of the genome which are not acted
 upon by selection, in all populations.
@@
 
-\subsubsection*{migration affects the speed of adaptation}
+\subsection*{migration affects the speed of adaptation}
 
@@
 as migration rates increased. 
-As can be seen in Figure \ref{fig:TimeToAdaptation} at the lowest migration rate of $5 * 10^{-5}$,
-It took the introduced population over 20 thousand generations for the average phenotype of all the lakes to 
-get to within 0.5 of the original lakes original phenotypes. 
+As can be seen in Figure \ref{fig:TimeToAdaptation} at the lowest migration rate of $5 \times 10^{-5}$,
+It took the introduced population over 20 thousand generations for the mean trait value, $X_{I}$ of all the lakes to 
+get to within 0.5 of the original lakes mean trait value, $X_{O}$
 %\plr{From the plots it looks more like adaptation happens (at least in most lakes) by like 10K or 15K.}
-This suggests that selection was acting primarily on new mutations, meaning the 
+This suggests that selection was acting primarily on new mutations, and that the 
 introduced lakes needed to wait for a beneficial mutation to arise before 
@@
 
+%(2) Do the math, discuss number ( and possibly type ) of mutations underlying the trait value.
+\subsection*{do the math}
+\jgg{peter? :)}
+
 %TIES TO NATURE?
 
-\subsubsection*{sharing of freshwater adapted alleles}
+\subsection*{sharing of freshwater adapted alleles}
 
@@
 
-To further investigate this correlation, we take a look at the FAA driving 
+To further investigate this correlation, we take a look at the freshwater adapted alleles driving 
 local adaptation of the introduced population. 
-One common metric we investigate is the percentage of FAA per individual in each of the populations. 
-This gives us a general look at the where these alleles are being distributed. 
-Looking at the lowest migration rate for the original lakes population In Figure (MPAA / IND) 
-We see that the average individual has almost exactly $1/10^{th}$ of the total defined FWAA throughout the simulation. 
+One common metric we investigate is the percentage of freshwater adapted alleles per individual in each of the populations any time $t$. 
+This gives us a idea of where the where these alleles are being distributed throughout the simulation run. 
+Looking at the lowest migration rate for the original lakes population In Figure \ref{fig:MPFAI}
+We see that the average individual has almost exactly $1/10^{th}$ of the total defined freshwater adapted alleles throughout the simulation. 
 Because FAA are defined among \textit{any} population, this suggests that each one of the $10$ 
@@
 
-As we can see in Figure (Total/Avg shared)
-as migration rate parameter increases for the simulations, 
-we see the distribution of the total number of FAA's decrease, and 
+As we can see in Figure \ref{fig:NumFAA}
+as historical introgression increases for the simulations, 
+we see the distribution of the total number of freshwater adapted alleles decrease, and 
 the average number of lakes each allele appears in at high frequency ($p > 0.5$), increase.
@@
 
-\subsubsection*{We see Migration Load after a certain threshold of migration}
+\subsection*{We see Migration Load after a certain threshold of migration}
 
@@
 We have experienced the most dramatic shifts of the population dynamics at $M = 5x10^{-3}$.
-After a threshold between this and $M = 5x10^{-2}$, we start to experience migration load. 
+After a threshold between this and $M = 5x10^{-2}$, all populations begin to experience migration load. 
 Significant gene flow constricts local adaptation
@@
 In Figure \ref{fig:MeanPhenotype} at $M = 5x10^{-2}$, we can see the distribution of average phenotype throughout the simulation
-pull towards the opposing selective pressure value and away from the local optimum in all subpopulations.
+pull towards the opposing selective pressure value and away from the local optimum trait value in all populations.
 
@@
 
-\subsubsection*{There is a Qualitative Threshold of migration rates}
-
-%FIGURES THAT SHOW THIS:
-%all
-%TIES TO NATURE?
-%SPECIFICALLY (interesting stuff)
-
-%GENERAL
-We have found that too little migration leads to selection upon new mutations in all subpopulations and lakes alike. 
-In contrast, at high migration rates we have seen that migration load limits 
-the ability for species to locally adapt to the selective pressure of their environment.
-This leads us to consider a window (Goldilocks Zone) of introgression which allows for the transportation
-of FAA's without migration load. In a species that commonly is subjected to two general types of selective pressure 
-such as marine and 
-\plr{Maybe this goes in the Discussion? Remaining observations could merge with the previous section.}
-
 \subsection*{Population genetics data signals}
 
-\begin{figure}[h!tb]
+\begin{figure}
 	\begin{center}
   		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3.png}
-  		\caption{(Decription and cleaning needed)
+  		\caption{(Description and cleaning needed)
 		}
@@
 
-\begin{figure}[h!tb]
+\begin{figure}
 	\begin{center}
@@
 \end{figure}
+
+Now we take a closer look at the genomic architecture of local adaptation between freshwater and marine individuals. 
+Often when looking at real datas, biologists observe patterns of $F_{st}$ and clustering of alleles that suggest causative loci for certain traits.
+However, researchers would like to know how parameters such as migration and recombination impact the results from natural populations.
+To give an idea about the impact of these parameters in our model, we examine the distribution, location, and false positive rate on predicting effect loci
+given the $F_{st}$ per SNP across the genome.
 
@@
 
-%
+\subsection*{False-Positives \& Statistical Power}
 
-Here, we take a closer look at the genomic architecture of local adaptation between freshwater and marine individuals. 
-Often when looking at real datas, biologists observe patterns of $F_{st}$ and clustering of alleles that suggest causative loci for certain traits.
-However, researchers would like to know how parameters such as migration and recombination impact the results from natural populations.
-To give an idea about the impact of these parameters in our model, we examine the distribution, location, and false positive rate on predicting effect loci
-given the $F_{st}$ per SNP across the genome.
+%\plr{Start off by talking about Fst along the genome, then say ``if we identify QTL as the outliers, then...''}
+%\plr{I vote for combining and including figures \ref{fig:Fst2} and \ref{fig:Fst3}.}
 
-\subsubsection*{False-Positives \& Statistical Power}
+At higher rates of introgression in our simulations, we saw $F_{st}$ peaks at regions of the genome underlying individual trait value (phenotype). 
+In contrast, the lower rates of introgression showed higher $F_{st}$ across the entire genome such that the regions under selection were indistinguishable. 
+For every $F_{st}$ peak, we saw the same regions of the genome express high $F_{st}$ when comparing either the initial \textit{or} introduced freshwater populations to the marine population.
+If we identify quantitative trait loci as the outliers, then this is again suggestive of both freshwater populations sharing regions of divergence from the marine population.
 
-\plr{Start off by talking about Fst along the genome, then say ``if we identify QTL as the outliers, then...''}
-\plr{I vote for combining and including figures \ref{fig:Fst2} and \ref{fig:Fst3}.}
+With 10 effect regions across the genome, we can see in \ref{fig:Fst3} that only about 1 or 2 of the region contain peaks. 
+%SOFT STATEMENT - MORE (Peak) EVALUATION?
+When looking closely at the peaks we see that each is usually a composition of 7-8 SNPs close in proximity.
+%SOFT STATEMENT -
+Because some of those mutations are selectively neutral (we can see this because the counts of FAA are not as high as the number in the clusters),
+this is suggestive of hitchhiking alleles along with the SNPs being brought to high frequency by a selective sweeps.
+
+In regions of the genome underlying individual trait value, we observed 
 Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$. 
@@
 %The lowest migration rate of $10^{-5}$ also has a very low false positive rate across all $F_{st}$ threshold values.
-
 %\subsubsection*{Low Recombination Causes Clustering}
@@
 \section{Conclusion \& Discussion}
+
+We have shown that historical introgression, at our given parameter sets,
+is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations.
+Selection is able to rebuild the freshwater haplotype from marine populations.
+Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load 
+limited the ability of the populations to reach the local optimum.
+
+We also have also explored the genomic architecture as a consequence of the nature of selection in our scenario. 
+The alleles underlying individual trait value were of large effect and a low number. 
+With a total of $10^{5}$ loci, to be realistic, each loci should represent 1000 $Kb$ in real data.
+
+We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
+%wouldn't it be cool if we could use this fact and simulate introgression, starting with real data that is noisy to extrapolate the areas we care about?
+%this could be nonsense
 
@@
 %(1) There's a threshold of introgression for both rapid adaptation and migration load. 
+%(1.5) Maybe refer to the literature for migration load.  
 \subsection*{thresholds}
 
-%(2) Do the math, discuss number ( and possibly type ) of mutations underlying the trait value.
-\subsection*{do the math}
+We have found that too little migration leads to selection upon new mutations in all subpopulations and lakes alike. 
+In contrast, at high migration rates we have seen that migration load limits 
+the ability for species to locally adapt to the selective pressure of their environment.
+This leads us to consider a window (Goldilocks Zone) of introgression which allows for the transportation
+of FAA's without migration load. In a species that commonly is subjected to two general types of selective pressure 
+such as marine and freshwater stickleback this is suggestive of the benefit of hybridization in all subpopulations before
+reaching a point of pre or post zygotic
 
@@
   		\label{fig:MNFAI}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=\linewidth]{matplotlibPlots/effectRegionCounts.png}
+  		\caption{(PLACEHOLDER)}
+		\label{fig:counts}
 	\end{center}

### 2018-07-11 Stickleback_Paper.tex
latest

@@
 %(2) Do the math, discuss number ( and possibly type ) of mutations underlying the trait value.
-\subsection*{do the math}
-\jgg{peter? :)}
 
@@
 %SPECIFICALLY (interesting stuff)
+
+\subsection*{do the math}
+\jgg{peter? :)}
+
+(1) - What is the mean fitness of a population
+
+(2) - assuming everybody is at the mean fitness, what is the expected num of offspring after t gens in marine
+
+(3) - expected num of establishing alleles in new lake. = (num outmigrants) $\times$ (num per outmigrant) $\times$ (prob migrant) $\times$ $2\hat{x}s $ mean fit in new lake
 
@@
 With a total of $10^{5}$ loci, to be realistic, each loci should represent 1000 $Kb$ in real data.
+\jgg{This should be expanded upon, not sure what this means in terms of the genomic architecture}
 
 We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
+This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
+differences between populations that have little gene flow between them.
+It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
+mistaken for being causative as they often display the same amount of divergence.
 %wouldn't it be cool if we could use this fact and simulate introgression, starting with real data that is noisy to extrapolate the areas we care about?
@@
 the ability for species to locally adapt to the selective pressure of their environment.
-This leads us to consider a window (Goldilocks Zone) of introgression which allows for the transportation
-of FAA's without migration load. In a species that commonly is subjected to two general types of selective pressure 
-such as marine and freshwater stickleback this is suggestive of the benefit of hybridization in all subpopulations before
-reaching a point of pre or post zygotic
+This leads us to consider a window of introgression which allows for the transportation
+of FAA's without migration load.
+
+\subsection*{connect results back to real data?}
+
+%In a species that commonly is subjected to two general types of selective pressure 
+%such as marine and freshwater stickleback this is suggestive of the benefit of hybridization in all subpopulations before
+%reaching a point of pre or post zygotic
 
 %(3) discuss what signals researchers could possibly look out for or connect to real data
-\subsection*{signals and real data}
 

### 2018-07-11 Stickleback_Paper.tex
latest

@@
 
+should we be using the word ``patch"?
+
 \section{Abstract}
@@
 While this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently.
-Novel evidence from natural populations has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern. \cite{Nelson2017} 
+Novel evidence from natural populations has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{Nelson2017}.
 \citet{schluter2009genetics} 
@@
 allows the marine to maintain the freshwater haplotype dispersed in low frequency among its' individuals. 
-Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection reconstructs the freshwater haplotype. \cite{schluter2009genetics} 
+Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection reconstructs the freshwater haplotype \citep{schluter2009genetics}.
 An alternative to this hypothesis is that individuals from other freshwater environments migrate directly to the new environments, 
@@
 
-%Where a trait value, $x$, is also defined by a single numeric value; 
-%determined as a summation of the selection coefficients of all mutations an individual possesses.
-
-%	\[
-%{\textnormal{Individual.Phenotype = } \sum_{m \in mutations} \textnormal{m.selectionCoeff}}
-%	\]
-
-%We allow six mutation types which affect phenotype along with one neutral mutation type.
-%Among the effect mutations; 
-%there are additive, dominant, and recessive mutations which a ffect phenotype
-%in either the positive or negative direction. 
-%The selection coefficients are pulled from a gamma distribution 
-%with a shape parameter of $1$ and a mean of either $0.5$, or ${-0.5}$.
-
-%	\[
-%\textnormal{Mutation.selectionCoeff = } X \sim \Gamma (1,\pm0.5)
-%	\]
-
-%\plr{I agree: first explain how phenotypes are calculated, then say what fitness is.
-%    Notation suggestion: phenotype for an individual that is heterozygous at loci $H$
-%    and homozygous derived at loci $D$ is $x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
-%    where $h_\ell$ is the dominance coefficient for locus $\ell$ (0 for recessive, 1 for dominant, 0.5 for additive),
-%    and $s_\ell$ is the effect of the derived mutation on the phenotype}
-
 \subsection*{Descriptive statistics}
-%\plr{I think the purpose of this section is to define terms and statistics used later;
-%    so maybe a better title is ``Descriptive statistics''?}
-
-%how did we use slim 
-%More question based, Why did we measure what we measued 
-%are there ways to have two or three subheaders 
-%decriptive statist. 
-%genetic basis -- 
-%process of plotting. 
-%if it's something new
 
@@
 impact the sharing of alleles between populations. 
-%\plr{Below, I think we should make sure to use ``population'' to mean either ``lake'' or ``ocean'';
-%    so when we compute freshwater-marine $F_{ST}$ this measures differentiation between habitats,
-%    not between populations.}
 We define pre-existing freshwater adapted alleles at time $t$,
-%\plr{I vote to remove the abbreviations. Also, should make the dependence on generation more clear: maybe ``freshwater adapted alleles at time $t$''?} 
 to be a mutations with a frequency higher than $0.5$ in \textit{any} of the original lakes, 
@@
 we track the the average percentage of freshwater adapted alleles at time $t$ across all populations throughout the simulation.
-%\plr{Here FAA are defined in terms of the original populations, explaining why the introduced populations can have MPFAI of 0.
-%    This is counter-intuitive from the name (FAA); maybe we should call them ``pre-existing freshwater alleles''?
-%    ``standing genetic variants''?}
-%we're particularly interested in using MPFAI to measure the marine population' capacity to carry FAA,
-%as well as observing the introduced populations' ability to select upon them. 
-%\plr{omit this last sentence?}
-%Because we are interested in the local adaptation of the introduced freshwater populations,
-
-%At this generation, we look at:
-%the correlation (B. Pearson) between the mutation frequencies of all effect loci and FAA,
-%and the number of shared high frequency alleles. 
-%\plr{Here we need to define terms; we don't need to say exactly what was computed,
-%    as that will be redundant with the Results.}
-%These metrics tells us how much of the FAA  were used in the lakes once they have locally adapted.
 
@@
 %    one of effect size $s$ has probability of about $s/2$ of not being lost to drift...}
-There was a positive correlation between the amount of standing genetic variation in the 
+In Figure \ref{fig:MPFAI} we can see there was a correlation between the amount of pre-existing freshwater adapted alleles in the 
 marine and the rapid adaptation of the introduced populations. 
+In the same figure we can see the distributions
+for both the original and Introduced freshwater population become more balanced (similar?) as introgression increases.
+This is further evidence that the introduced population is taking advantage of the alleles from the initial freshwater populations
 
@@
 local adaptation of the introduced population. 
-One common metric we investigate is the percentage of freshwater adapted alleles per individual in each of the populations any time $t$. 
+One common metric we investigate is the percentage of freshwater adapted alleles per individual in each of the populations, at any time $t$. 
 This gives us a idea of where the where these alleles are being distributed throughout the simulation run. 
@@
 We have shown that historical introgression, at our given parameter sets,
-is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations.
-Selection is able to rebuild the freshwater haplotype from marine populations.
+is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island.
+Selection is able to rebuild the freshwater haplotype from marine populations as a medium between all freshwater populations.
 Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load 

### 2018-07-14 Stickleback_Paper.tex
latest
reworked methods

@@
 \usepackage{xcolor}
-%\documentclass{article}
-%\usepackage{subcaption}
-%\documentclass[smallextended]{svjour3}
-%\usepackage{lipsum}% Just for this example
 
@@
 %\doublespacing
-%\setstretch{1.1}
 
@@
 
-\linespread{1.5}
+\title{The sweet spot for local polygenic adaptation: a case study motivated by stickleback}
+\author{Jared Galloway, Bill Cresko, and Peter Ralph}
+\maketitle
 
-JARGON WE SHOULD THINK ON ??
+\textbf{Other title ideas?} (Bad ideas allowed.)
+
+How do stickleback share adaptations between lakes? A simulation study.
+
+A simulation study of gene flow and local polygenic adaptation
+
+\textbf{JARGON WE SHOULD THINK ON ??}
 
@@
 % Jared's general notes for formatting
-
-%\plr{``citep'' for ``cite in parentheses''. (and, citations come before the punctuation)}
-%\plr{in latex, use ` `words' ' to get the open- and closed-double quotes (not the double-quote character)}
-%\plr{"citet" for "cite, as part of the text"}
+%
+% - ``citep'' for ``cite in parentheses''. (and, citations come before the punctuation)
+% - in latex, use ` `words' ' to get the open- and closed-double quotes (not the double-quote character)
+% - "citet" for "cite, as part of the text"
 
@@
 Quickly inhabited by the surrounding marine population of Stickleback,
-\citet{LescakE7204} observed significant phenotypic changes in less than 50 years that appear to be 
+\citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be 
 parallel to freshwater stickleback that have been separated for over $13,000$ years.
@@
 While this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently.
-Novel evidence from natural populations has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{Nelson2017}.
+Novel evidence from natural populations has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
 \citet{schluter2009genetics} 
@@
 There are many different parameters that carry a large impact on the questions above.
-\citet{Ralph2015} 
+\citet{ralph2015convergent} 
 explored the dichotomy of selection on new alleles and those brought in by migrants from a similar selective pressure.
@@
 
-\section{Simulation Model}
+%%%%%%%%%%%%%%%%
+\section{Methods}
 
-Our simulations are forward-moving evolutionary models which
-emulate the geography, selective pressure, and genomic architecture 
-of coastal marine and freshwater lake populations. 
-Here, we describe the specifics and parameter values of our model.
+We explored these questions
+using forwards-time simulations with explicit genomic representation of a quantitative trait
+in SLiM \citep{haller2017slim}.
+The details (described below) are motivated by the nine-spine stickleback,
+but remain simplistic in some aspects due to computational constraints.
+Possibly the most important caveat is that simulated population sizes are much smaller
+than the census size of the nine-spine stickleback population;
+we will consider the likely effects of this in the Discussion.
+\plr{TODO}
 
-\subsection*{life cycle}
+\paragraph{Habitat and geography}
+Our model has two habitat types, which are separate populations
+with a fixed total size of 2,000 diploid individuals each.
+The scheme, depicted in Figure \ref{fig:Geo},
+roughly models a set of freshwater habitats
+along a stretch of coastline.
+The ``marine'' habitat  is a continuous, one-dimensional range of 10 units of length,
+and the ``freshwater'' habitat is divided into ten subpopulations
+(which we call ``lakes''), each connected to the marine habitat
+at positions $i - 1/2$ for $1 \le i \le 10$.
 
-For this study, we used an evolutionary framework, SLiM \citep{Haller2017}.
-The model-type we used was derived from an extended wright-fisher. 
-The simulation life cycle of each generation is as follows.
-First, generate offspring by:
-(1) choosing parents based on cached fitness values and migration rates, 
-(2) performing recombination of parental genomes, 
-(3) allow for mutation, 
-(4) modify each child by a defined callback. 
-Next, offspring become individuals.
-And finally, fitness is evaluated for the next generation
-and the generation counter is incremented.
-
-The population structure in SLiM can be arranged in any number of subpopulations, 
-continuous or discrete, connected by any rate of migration. 
-Individual's (diploidy) genomes are represented by a linear array of loci. 
-Each locus is positionally defined with recombination and mutation rates,
-as well as mutation types allowed to arise.
-
-\subsection*{Geography}
-
-\jgg{do we need this section or should we just give a really good description in the geography diagram?}
-What we've modeled geographically is a coastal marine population 
-connected by migration to multiple, smaller freshwater populations that are located along the coast.
-The population structure for the marine is modeled as a one-dimensional, continuous population
-ranging from [0.0 to 10.0].
-There are two sets of lakes which represent freshwater populations; 
-the first which evolves in parallel to the marine, 
-and a set introduced later in the simulation as a subset of the marine. 
-Each of the $10$ lakes, $i$, is located along the marine at $i - 0.5$, 
-and connected by migration only through the marine environment. 
-The marine has $2000$ individuals while 
-each of the lakes fluctuate around $200$.
+Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff,
+to roughly model the cumulative effect of the various phenotypes
+such as lateral plating and opercle shape
+on which divergent selection acts in the two environments.
+The optimal trait value in the marine habitat is $+10$,
+and in the freshwater habitat is $-10$,
+and fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$ is
+Gaussian with standard deviation, i.e.,
+\begin{align*}
+    f(x; x_0)
+    &=
+    \exp\left\{
+        \frac{1}{2}
+            \left(
+            \frac{x-x_\text{opt}}{15}
+            \right)^2
+        \right\} .
+\end{align*}
 
@@
 
-\subsection*{Genomic Architecture}
+\paragraph{Potential genomic architecture}
+Each individual carries two copies of a linear chromosome of $10^5$ loci
+(which could be thought of roughly as a kilobase each).
+New mutations occur with probability $10^{-7}$ per locus per generation.
+Ten ``effect regions'' of 100 loci each
+are spread evenly along the chromosome (separated by 4950 loci)
+in which mutations affect the trait;
+mutations in other parts of the genome are neutral.
+Each mutation in these regions is either additive, completely recessive,
+or completely dominant (with equal probability),
+and with an effect sizes chosen randomly from an Exponential distribution with mean $1/2$,
+either positive or negative with equal probability.
+Individual trait value are determined additively from the diploid genotypes.
+Concretely, an individual that is heterozygous and homozygous for mutations
+at loci $H$ and $D$ respectively
+has trait value
+$x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
+where $s_\ell$ is the effect size of the mutation at locus $\ell$.
+Subsequent mutations at the same locus replace the previous allele.
 
-In SLiM, the genome is conceptually a linear array of loci defining locations for 
-mutations that arise and breakpoints for recombination. 
-Genomic elements define a range depicting the types of mutations that can arise, and the ratio at which different mutation types can arise. 
-Although still uncertain about how much of the genome is directly associated with the distinguished phenotypes, 
-GWAS has indicated clusters of loci (linkage groups? Operon?) along the genome to be causative. 
-To mimic this and compare dynamics of neutral vs. effect under selection, 
-we create regions of effect. (???)
-In these genomic elements of effect, we uniformly allow additive, recessive, and dominant, mutation that can push phenotype in either the positive or negative direction to arise. 
-Every genomic position outside of these effect regions only allows for selectively neutral mutations to arise.
+\paragraph{Life cycle}
+Each generation, the two parents of each new offspring are chosen 
+proportional to their fitness (all individuals are hermaphroditic),
+and the contributing genomes are produced by Poisson recombination
+with an average of one crossover per chromosome per generation
+($10^{-7}$ per locus per generation).
+The model is Wright-Fisher, so that each generation, in each habitat,
+2000 new individuals are produced,
+and so to keep population sizes roughly constant within each lake,
+at the start of each generation we rescale fitness values in the freshwater habitat
+so that the sum of the fitness values within each lake are equal.
 
-In our simulations, there are 10 effect regions of size 100 loci, uniformly placed between buffer regions of size 4950 loci. 
-This means the effect regions make up exactly $1\%$ of the entire genome for every individual throughout the simulation. 
-The entire genome has a mutation rate of $10^{-7}$ meaning approximately 1 in every 100 individuals, per generation, is subject to a new mutation. 
-The genome has a recombination rate of $10^{-5}$, which is approximately 1 breakpoint event, per individual, per generation.
+Migration occurs both locally along the coastline in the marine habitat
+and between the marine habitat and the lakes,
+and can be thought of as occurring at the juvenile stage.
+The between-habitat migration rate is denoted $m$,
+and will be called simply the ``migration rate''.
+Each new individual in the marine habitat has freshwater parents with probability $m$;
+to obtain the pair, a first parent is chosen proportional to fitness,
+and a mate is chosen from the same lake as the first, also proportional to fitness.
+The resulting offspring is given a spatial location in the marine habitat
+at the location of the parent's lake.
+Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$):
+first, a single parent is chosen proportionally to fitness in the marine habitat,
+and then a mate is chosen, 
+also proportionally to fitness but reweighted by a Gaussian function 
+of the distance separating the two, with stanadard deviation $1/2$: 
+concretely, if the first parent is marine individual $i$,
+then marine individual $j$ is chosen as the mate
+with probability proportional to $f(x_i) \exp(-2d_{ij}^2)$,
+where $d_{ij}$ is the distance between the two locations.
+Finally, each new marine offspring is given a position 
+displaced from the first parent's position by a random Gaussian distance
+with mean 0 and standard deviation 0.02, and reflected to stay within the population.
+New offspring in the freshwater habitat are chosen in the same way,
+except the probability that the parents are marine individuals is $m$;
+any new freshwater offpsring produced by marine individuals
+are assigned to the lake nearest to the position of the first marine parent.
 
-\subsection*{Selective Pressure}
+\paragraph{New lakes}
+To study adaptation in newly appearing habitats,
+we introduce a new set of lakes midway through the simulation.
+The initial set of individuals have parents chosen as above from the marine habitat,
+and act like an independent copy of the original set of lakes -- in particular,
+the two sets of lakes each have 2,000 individuals at all times.
+Since this doubles the number of lake-to-marine immigrants,
+after this happens the probability that a new marine individual has freshwater parents
+is $2m$ instead of $m$.
 
-%(TALK ABOUT ADAPTEDNESS and TRADE-OFF), 
-% abstracting real fish phenotype with `` trait value "
+\paragraph{Descriptive statistics}
+To assess whether new lakes adapt using existing genetic diversity,
+we define a \emph{pre-existing freshwater adapted allele}
+to be an effect mutation that has frequency higher than $0.5$ in at least one of the original lakes,
+while remaining lower than $0.5$ in the marine. 
+This categorization is made for each generation using the allele frequencies from that generation,
+and so changes with time.
+Alleles common in the newly introduced lakes do not count
+if they are not also common in the original lakes.
+They are defined this way because the transportation hypothesis 
+does not specify where or when an advantageous mutation arises,
+but simply suggests that any sufficiently common freshwater adapted allele 
+could participate in adaptaiton in new habitat \citep{schluter2009genetics}.
 
-%Trait value is equal to the sum of the contribution of the effect loci and the loci may be 
-%add, dom, recc. Uniformly chosen
+\emph{Time to adaptation} of the introduced population, denoted $T_\text{adapt}$, 
+is defined to be the generation at which 
+the difference between the average trait value in the original and the introduced freshwater populations 
+is less than 0.5.
 
-To emulate the freshwater and marine selective pressure, 
-we set up a quantitative genetics model in which fitness is phenotypically based. 
-phenotype for an individual that is heterozygous at locus $H$
-and homozygous derived at loci $D$ is $x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
-where $h_\ell$ is the dominance coefficient for locus $\ell$ (0 for recessive, 1 for dominant, 0.5 for additive),
-and $s_\ell$ is the effect of the derived mutation on the phenotype
+We describe overall genetic differentiation between the habitats using $F_{ST}$,
+calculated on a per-locus basis.
+Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele
+in the freshwater and marine habitats, respectively,
+and $\bar p = (p_f + p_m)/2$,
+then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
 
-Freshwater and Marine environments are distinguished by a single numeric value. 
-This number is the optimum trait value for each environment
-and acts as a representation of lateral plate number and opercle shape in the stickleback populations.
-%The fitness of an individual is then determined by a probability density for a normal distribution 
-%at the quantile (?) of the difference between the optimum and the individual's phenotype
-We use a Gaussian fitness function, 
-where the fitness of an individual with trait $x$ in a population with optimal trait value $x_0$ is:
-\begin{align}
-    f(x; x_0) &= \exp\left(\frac{(x - x_0)^2 }{ 2 \sigma^2}\right),
-\end{align}
-with $\sigma = 15$.
-
-\subsection*{Descriptive statistics}
-
-Here, we're interested in observing how different parameter values
-(mainly migrations rates and recombination rates)
-impact the sharing of alleles between populations. 
-We define pre-existing freshwater adapted alleles at time $t$,
-to be a mutations with a frequency higher than $0.5$ in \textit{any} of the original lakes, 
-while remaining lower than $0.5$ in the marine. 
-They are defined this way because the transportation hypothesis does not specify where or when an advantageous mutation arises,
-but simply suggests that when a freshwater adapted alleles comes to high enough frequency, it too, could participate in the transportation process \citep{schluter2009genetics}.
-To observe transportation of freshwater alleles in our simulations, we use a variety of metrics when sampling.
-let $P_{1}$ and $P_{2}$ be the vector of mutation frequencies in the first and second populations, respectively.
-with $\hat{P} = (P_{1} + P_{2}) / 2$ and
-$PQ = (P_{1} \times (1 - P_{1}) + P_{2} \times (1 - P_{2})) / 2$,
-we compute $F_{ST} = 1 - (PQ / (\hat{P} \times (1 - \hat{P})))$.
-%mean number of original lakes each FAA appears in,
-%and total number of unique FAA. 
-Time to adaptation ($T_{adapt}$) of the introduced population is the generation at which
-the difference between the average phenotype of the original and the introduced freshwater populations is less than 0.5. 
-To understand where, as well as how much, of these pre-existing freshwater adapted alleles are being transported,
-we track the the average percentage of freshwater adapted alleles at time $t$ across all populations throughout the simulation.
-
+%%%%%%%%%%%%%%%%%
 \section{Results}
@@
 
-\subsection*{Population genetics data signals}
+\subsection*{Realized genomic architecture and signals}
 

### 2018-07-17 Stickleback_Paper.tex
edits through most of results

@@
 
-Across a range of historical introgression from $\approx 0.01 - 100.0$ migrants per generation in populations of 
-size 2000 diploid individuals, we encountered large influence on 
-time to adaptation,
+% Overview
+We varied the migration rate, $m$,
+from $5 \times 10^{-5}$ to $5 \times 10^{-2}$ per individual,
+i.e., between 0.01 and 10 migrants per year to and from each of the ten lakes.
+This had a strong effect on many aspects of adaptation: 
+including time to adaptation,
 sharing of pre-existing freshwater adapted alleles,
-and population genetic signals left behind in the simulated populations.
-More gene flow is experienced with greater historical introgression between the two selective pressures.
+and the population genetic signals left behind.
+At the lowest migration rate, lakes adpated nearly independently,
+while at the highest migration rate, the habitats are beginning to act like a single metapopulation
+with substantial migration load.
+We will now dissect what happens between these extremes.
+
 Interestingly, the ability for separate populations to locally adapt to their own selective pressure was relatively unaffected 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Local Adaptation: differentiation with gene flow}
+
+% The progression of adaptation. How you can see introduction new alleles
+
+Local adaptation occurred in all simulations:
+as shown in Figure \ref{fig:phenotype_ts2},
+freshwater and marine populations diverged
+until the trait means were close to the optimal values in each habitat. 
+(Remaining simulations are shown in Figures \ref{phenotype_ts1} and \ref{phenotype_ts4}.)
+Greater migration decreased adaptedness only weakly
+except for at the highest migration rate (10 migrants per generation per lake),
+where the mean trait value equilibrated at roughly half its optimal value 
+(Figures \ref{fig:MeanPhenotype} and \ref{phenotype_ts4}).
+With less than one migrant per generation per lake,
+adaptation happened substantially faster
+in the marine habitat than in the lakes.
+The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
+as jumps in the mean phenotype curve;
+in the two simulations these occur on a time scale of 100 (lower migration) to 1000 (higher migration)
+generations,
+and move the phenotype by an amount of order 1.
+Trait variation within each population was small compared to the difference between populations,
+with interquartile ranges of around XXX.\plr{TODO}
+Across all parameter values, differences at around 16 commonly polymorphic sites 
+(eight the shift the trait in each direction)
+were responsible for most of the adaptive differences between freshwater and marine habitats.
+\plr{ALSO TODO: CHECK THIS}
+
 \begin{figure}
 	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
-  		\caption{ Mean individual trait value in each population as a function of time $t$, in generations. 
-		This is a single simulation run at a migration rate of $5 \times 10^{-4} \approx 1$ mig/gen in both 
-		directions for every freshwater population to the single marine population. 
-		The dark black line represents the mean phenotype of the single marine population with an optimum trait value of + 10. 
-		The pink represents the entire initial freshwater population mean
-		of all the lakes (light green) with and optimum trait value of -10.
-		The orange represents the entire introduced freshwater population mean
-		of all the lakes (dark green) with an optimum trait value of -10 as well.
+        \includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4}
+        \includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3}
+  		\caption{ 
+        Mean individual trait values in the marine habitat (black line),
+        the original lakes (light green lines; average in pink),
+        and the introduced lakes (dark green lines; average in orange)
+        across the course of two simulations, with migration rates of
+        \textbf{(top)} $m=5 \times 10^{-4}$ and
+        \textbf{(bottom)} $m=5 \times 10^{-3}$ per generation individual
+        (i.e., 0.1 and 1.0 migrants per generation per lake, respectively).
+        Optimal trait values in the two habitats are at $\pm 10$.
 		}
@@
 
-\begin{figure}
+\jgg{also, interesting point to make on second plot, you can see the initial lakes taking the same size step, showing sharing of alleles.
+could zoom in on the relevent portions of the figure for this. (begging and introduction)
+ a 2 by 2 with the important sections (0 - 20K) or (40-50K) for both above figures?}
+
+\begin{figure}[h!tb]
 	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
-  		\caption{Mean individual trait value in each population as a function of time $t$, in generations. 
-		This is a single simulation run at a migration rate of $5 \times 10^{-3} \approx 1$ mig/gen in both 
-		directions for every freshwater population to the single marine population. 
-		The dark black line represents the mean phenotype of the single marine population with an optimum trait value of + 10. 
-		The pink represents the entire initial freshwater population mean
-		of all the lakes (light green) with and optimum trait value of -10.
-		The orange represents the entire introduced freshwater population mean
-		of all the lakes (dark green) with an optimum trait value of -10 as well.
+  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MeanPhenotype1.png}
+  		\caption{
+            Distribution of mean trait values throughout each simulation.
+            The dashed pink and purple lines show the optima ($\pm 10$).
 		}
-  		\label{fig:phenotype_ts3}
+  		\label{fig:MeanPhenotype}
 	\end{center}
 \end{figure}
+\plr{Is this the whole simulation? what about the initial adaptation phase?
+     Is the horizontal line at the median? what are the whiskers at?}
 
-\jgg{we should make one caption and the two above should be top-bottom sub-figures}
-\jgg{also, interesting point to make on second plot, you can see the initial lakes taking the same size step. sharing}
-%I could zoom in on the relevent portions of the figure for this. (begging and introduction)
-% a 2 by 2 with the important sections (0 - 20K) $ (40-50K) for both above figures. share axes
-%\jgg{finished with most edits up to here}
-
-\subsection*{Local Adaptation: differentiation with gene flow}
-
-%Talk about the genetic basis of trait. 
-%Talk about the progression of adaptation. How you can see introduction new alleles
-
-Local adaptation occurred across all simulated parameter values.
-Starting from the same baseline, freshwater and marine populations diverged phenotypically
-until they reached an equilibrium
-where the population means were close to or at the optimal phenotype. 
-The distribution of mean number of freshwater adapted alleles suggested that there was 
-$\approx 8$ SNPs underlying the freshwater phenotype. 
-This was true for all rates of introgression except for the highest at $100$ mig/gen
-where the mean trait value in the populations equilibrated well below the optimum.
-We can see the introduction of new alleles in Figure \ref{fig:phenotype_ts2}. 
-In generation (0 - 20K) The sweeps that cause jumps in the mean phenotype of the freshwater population 
-have a relatively large effect size equal to or greater than $1$.
-%In Figure \ref{fig:counts} we can see that the number of effect regions with zero 
-%pre-existing freshwater adapted alleles increased with introgression while
-%regions of effect containing more pre-existing freshwater adapted alleles increase.
-Phenotypic variation within each population was small compared to the difference between populations.
-Polymorphic loci that impact the individual trait value,
-showed greater differentiation compared to neutral loci between marine and freshwater populations, 
-with the exception of the lowest migration rate of 0.01 migrants per generation per lake.
-
-\todo{ add more general observations and caveats about not all parameter values. }
-\plr{It would be nice to have a small table of parameter values,
-    so that when someone wonders ``What was the population size again?'' they could easily find it.}
-\jgg{Re: agreed}
-
-\subsection*{Migration homogenizes genetic variation}
-
-%FIGURES THAT SHOW THIS:
-%- Fst Plots.
-%- SGV Plots.
-%- Haplotype plots
+As expected, increasing migration rate decreased differentiation between habitats.
+As seen in Figure \ref{fig:Fst},
+$F_{ST}$ between marine and freshwater habitats at neutral sites 
+steadily declines as migration increases.
+Local adaptation was still able to occur despite overall homogenization:
+$F_{ST}$ between marine and freshwater habitats 
+at sites with alleles affecting the trait (``effect mutations'')
+was relatively constant across migration values.
 
@@
   		\includegraphics[width=0.8\linewidth]{matplotlibPlots/FST_HRR.png}
-  		\caption{Distributions of $F_{st}$  values between the three subpopulations throughout the simulations.
-		On the left, $F_{st}$ was calculated only on neutral mutation frequencies. On the right, $F_{st}$ was calculated
-		only on effect (Impact on phenotype) mutation frequencies. 
-		Effect mutations are acted upon by selection as they have an affect on fitness where neutral mutations  }
+  		\caption{
+            Distributions of $F_{st}$ values between the three habitats,
+            accumulated across the course of the simulation, calculated 
+            \textbf{(left)} using only neutral mutations, and
+            \textbf{(right)} using effect mutations.
+        }
   		\label{fig:Fst}
@@
 
-Across increasing parameters of migration, we observed more gene flow across the populations. 
-As can be seen in Figure \ref{fig:Fst},
-$F_{st}$ values for neutral alleles steadily decline between all sets of populations as 
-migration increases, this illustrates less difference between sections of the genome which are not acted
-upon by selection, in all populations.
-Additionally, standing genetic variance
-(seen in Figure \ref{fig:SGV})
-values in the marine population steady increased with migration which suggests that more alleles from the 
-lakes surrounding were exporting alleles into the marine individual.
+\paragraph*{Speed of adaptation}
+Increasing migration rate strongly decreased the time until freshwater habitats could adapt,
+both in the initial and introduced sets of lakes.
+As shown in Figure \ref{fig:TimeToAdaptation}, at the lowest migration rate 
+it took over 20 thousand generations for the mean trait value across introduced lakes to 
+get to within 0.5 of the original lakes value.
+Although many lakes had adapted before this time,
+the different rates of introduction of effect alleles is clearly seen in the traces
+of trait values against time (e.g., Figure \ref{fig:phenotype_ts2}).
+However, higher migration rates allowed the freshwater habitat to adapt
+nearly as quickly as the marine habitat.
 
-%TIES TO NATURE?
-%SPECIFICALLY (interesting stuff)
-
-\subsection*{migration affects the speed of adaptation}
-
-%FIGURES THAT SHOW THIS:
-%-Time to adaptation
 \begin{figure}
 	\begin{center}
-  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/TimeToAdaptation.png}
-  		\caption{Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
-		it takes for the introduced population's mean phenotype to come within 0.5 of the original lakes average phenotype. 
-		Each point represents a simulation run at some value of $M$. 
-		The yellow dashed line is the average of all points at each respective parameter value}
+        \includegraphics[width=0.8\linewidth]{matplotlibPlots/TimeToAdaptation}
+        \caption{
+            Time to adaptation as a function of migration rate ($m$),
+            measured as the number of generations
+            until the introduced population's mean phenotype is within 0.5 
+            of the original lakes average phenotype. 
+            Each point represents a single simulation, and
+            the line shows the average of all points at each value of $m$.
+        }
   		\label{fig:TimeToAdaptation}
@@
 
-We observed a dramatic shift in time until adaptation for the introduced populations
-as migration rates increased. 
-As can be seen in Figure \ref{fig:TimeToAdaptation} at the lowest migration rate of $5 \times 10^{-5}$,
-It took the introduced population over 20 thousand generations for the mean trait value, $X_{I}$ of all the lakes to 
-get to within 0.5 of the original lakes mean trait value, $X_{O}$
-%\plr{From the plots it looks more like adaptation happens (at least in most lakes) by like 10K or 15K.}
-This suggests that selection was acting primarily on new mutations, and that the 
-introduced lakes needed to wait for a beneficial mutation to arise before 
-it was selected upon. 
-%\plr{We should do the calculation to see if this makes sense: 
-%    there's one new effect mutation in the right direction every 100 generations in the lakes,
-%    one of effect size $s$ has probability of about $s/2$ of not being lost to drift...}
-In Figure \ref{fig:MPFAI} we can see there was a correlation between the amount of pre-existing freshwater adapted alleles in the 
-marine and the rapid adaptation of the introduced populations. 
-In the same figure we can see the distributions
-for both the original and Introduced freshwater population become more balanced (similar?) as introgression increases.
-This is further evidence that the introduced population is taking advantage of the alleles from the initial freshwater populations
-
-%(2) Do the math, discuss number ( and possibly type ) of mutations underlying the trait value.
-
-%TIES TO NATURE?
-
-\subsection*{sharing of freshwater adapted alleles}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Sharing of freshwater adapted alleles}
 
@@
 %-Total/Avg shared
+
+As expected,
+lakes were able to adapt faster at higher migration rates
+because greater mixing allowed them to share alleles
+instead of developing their own genetic basis of adaptation.
+As a first indication of this, Figure \ref{fig:Fst} shows that
+$F_{ST}$ between the two sets of lakes at effect mutations decreased with migration rate.
+
+To further investigate sharing of locally adaptive alleles, 
+we investigate the ``pre-existing freshwater adapted alleles'',
+defined for a particular generation
+to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat
+in that generation.
+Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles,
+across generations,
+as well as the average number of lakes each is found in.
+As migration rates increase, the number of these alleles decreases steadily,
+and each is concurrently found in a greater number of lakes.
+
+As the number of common, locally adaptive alleles decreases,
+the genetic basis of adaptation is more commonly shared.
+Figure \ref{fig:MPFAI}B shows
+the distribution of the mean percentage of currently-defined freshwater adapted alleles 
+that each genome in each of the populations carries,
+averaged across time and individuals.
+If all individuals across lakes carried the same set of alleles determining their trait value,
+this would be 100\%.
+This value is nearly reached at the highest migration rate;
+but it is lower due to migration load.
+At the lowest migration rate, 
+each genome in the original lakes have almost exactly $1/10^{th}$ 
+of the total freshwater adapted alleles --
+since there are 10 lakes, this implies that each lake has adapted with a unique set of alleles.
+Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
+
 \begin{figure}
 	\begin{center}
-  		\includegraphics[width=\linewidth]{matplotlibPlots/MPFAI.png}
-  		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
+        \includegraphics[width=\linewidth]{plotlyPlots/NumFAA}
+        \includegraphics[width=\linewidth]{matplotlibPlots/MPFAI}
+  		\caption{
+            \textbf{(A)}
+            Distribution of the number of freshwater adapted alleles (green)
+            and the average number of lakes each allele is present in (red)
+            across the course of the simulation.
+            \textbf{(B)}
+            Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
 		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
@@
 
-To further investigate this correlation, we take a look at the freshwater adapted alleles driving 
-local adaptation of the introduced population. 
-One common metric we investigate is the percentage of freshwater adapted alleles per individual in each of the populations, at any time $t$. 
-This gives us a idea of where the where these alleles are being distributed throughout the simulation run. 
-Looking at the lowest migration rate for the original lakes population In Figure \ref{fig:MPFAI}
-We see that the average individual has almost exactly $1/10^{th}$ of the total defined freshwater adapted alleles throughout the simulation. 
-Because FAA are defined among \textit{any} population, this suggests that each one of the $10$ 
-lakes has created their own solution to adaptation of the the freshwater selective pressure.
-
-As we can see in Figure \ref{fig:NumFAA}
-as historical introgression increases for the simulations, 
-we see the distribution of the total number of freshwater adapted alleles decrease, and 
-the average number of lakes each allele appears in at high frequency ($p > 0.5$), increase.
-With the distribution of effect size remaining the same across all simulations, 
-these are both suggestive of the original lakes sharing solutions to the same selective pressure.
-
-%TIES TO NATURE?
-
-%SPECIFICALLY (interesting stuff)
-
-\subsection*{do the math}
-\jgg{peter? :)}
-
-(1) - What is the mean fitness of a population
-
-(2) - assuming everybody is at the mean fitness, what is the expected num of offspring after t gens in marine
-
-(3) - expected num of establishing alleles in new lake. = (num outmigrants) $\times$ (num per outmigrant) $\times$ (prob migrant) $\times$ $2\hat{x}s $ mean fit in new lake
-
-\subsection*{We see Migration Load after a certain threshold of migration}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Migration load and genetic variation}
 
@@
 %- Phenotype Distribution
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MeanPhenotype1.png}
-  		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
-		The dashed pink line (Pheno = +10) is the optimum phenotype any individual in the marine environment.
-		In contrast the purple line at (Pheno = -10) represents the optimum for any individual in the freshwater environment. 
-		All individuals at generation 0 (beginning of the simulation) 
-		}
-  		\label{fig:MeanPhenotype}
-	\end{center}
-\end{figure}
-
 %GENERAL
 
-In the distributions we have shown across migration rate parameter values, 
-We have experienced the most dramatic shifts of the population dynamics at $M = 5x10^{-3}$.
-After a threshold between this and $M = 5x10^{-2}$, all populations begin to experience migration load. 
+At first, increased migration allows sharing of adaptive alleles between lakes,
+but at the highest migration rate,
+the constant influx of alleles between the habitats creates substantial migration load.
+This rate, $m=.05$, only replaces 5\% of each population each generation
+with migrants from the other habitat, but this is sufficient to shift the mean trait values
+to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
 Significant gene flow constricts local adaptation
 as a consequence of a large number of offspring through hybridization events between subpopulations.
-In Figure \ref{fig:MeanPhenotype} at $M = 5x10^{-2}$, we can see the distribution of average phenotype throughout the simulation
-pull towards the opposing selective pressure value and away from the local optimum trait value in all populations.
 
-%TIES TO NATURE?
+\paragraph{Standing genetic variation}
+Despite a dramatic difference in the amount of allelic sharing between lakes,
+standing genetic variance (Figure \ref{fig:SGV}) was around 0.05 in the freshwater habitat
+across all but the highest migration rates.
+Concurrently,
+genetic variation in the marine population steady increased with migration
+to a similar value.
+On the one hand, it is not surprising that lakes, as a group,
+carry more genetic variation than the marine habitat,
+since population subdivision allows different alleles to become common in each.
+However, it is striking that at $m=.005$, the marine habitat carries as much genetic variation,
+despite a lack of any substantial migration load.
 
-%SPECIFICALLY (interesting stuff)
-
-\subsection*{Realized genomic architecture and signals}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Realized genomic architecture}
 
@@
 	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3.png}
+        \includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3}
+        \includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-4}
   		\caption{(Description and cleaning needed)
@@
 
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-4.png}
-  		\caption{(Description and cleaning needed)
-		}
-  		\label{fig:Fst2}
-	\end{center}
-\end{figure}
+Now we take a closer look at the genomic architecture of local adaptation 
+between the two habitats.
+Do the alleles underlying trait differences cluster along the genome?
+Do measures of local differentiation identify the causal loci?
+Figure \ref{fig:Fst3} shows plots along the genome of per-locus $F_{ST}$ values
+between the marine and original freshwater habitats
+at the two intermediate migration rates.
+(Note that we are pooling freshwater habitats; 
+a single lake would provide substantially less power.)
 
-Now we take a closer look at the genomic architecture of local adaptation between freshwater and marine individuals. 
-Often when looking at real datas, biologists observe patterns of $F_{st}$ and clustering of alleles that suggest causative loci for certain traits.
-However, researchers would like to know how parameters such as migration and recombination impact the results from natural populations.
-To give an idea about the impact of these parameters in our model, we examine the distribution, location, and false positive rate on predicting effect loci
-given the $F_{st}$ per SNP across the genome.
+\plr{need to know where the FAA are in these figures}
+
+Higher migration rates showed more distinct $F_{ST}$ peaks
+over polymorphic loci underlying trait differences between the habitats.
+As migration rate decreased, the ``background'' levels of $F_{ST}$ increased,
+swamping out this signal until the regions under selection were indistinguishable. 
+This is likely due to two reasons: 
+first, stronger genetic drift with less migration leading to higher background $F_{ST}$,
+and second, greater sharing of adaptive alleles providing a shared signal across populations.
+
+This suggests that genome scans for local adaptation
+based purely on measures of differentiation
+will only be successful given enough migration between habitats.
+To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates
+that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value
+to be a causal locus.
+\plr{Labels are wrong on that figure: false pos rate should go down with cutoff.
+     Is it true positive rate?}
 
@@
 		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
+        TODO: x-axis label ($F_{ST}$).
 		}
@@
 \end{figure}
-
-%Suppose you have a Fst Peak, ___ how likely is it that that region is causal. and conversely, what proportion of 
-%the causal loci are under Fst Peaks.
-
-\subsection*{False-Positives \& Statistical Power}
-
-%\plr{Start off by talking about Fst along the genome, then say ``if we identify QTL as the outliers, then...''}
-%\plr{I vote for combining and including figures \ref{fig:Fst2} and \ref{fig:Fst3}.}
-
-At higher rates of introgression in our simulations, we saw $F_{st}$ peaks at regions of the genome underlying individual trait value (phenotype). 
-In contrast, the lower rates of introgression showed higher $F_{st}$ across the entire genome such that the regions under selection were indistinguishable. 
-For every $F_{st}$ peak, we saw the same regions of the genome express high $F_{st}$ when comparing either the initial \textit{or} introduced freshwater populations to the marine population.
-If we identify quantitative trait loci as the outliers, then this is again suggestive of both freshwater populations sharing regions of divergence from the marine population.
 
@@
 
-\section{Conclusion \& Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Theoretical expectations}
+\plr{does this go first or second?}
+
+Here's some rough calculations to get a sense for what should be going on.
+Everything is done in more detail in OTHER PLACES WE SHOULD CITE.
+Fisher, Chevin, etc.
+
+Suppose a new allele enters a lake, either by migration or mutation.
+If, when it is rare but present in $n$ copies,
+it has fitness advantage $s$
+-- i.e., the expected number of copies in the next generation is $(1+s)n$ --
+then the probability that it escapes demographic stochasticity to become common in the population
+is approximately $2s$ \citep{fisher,prob_fixation}.
+If the current population all differed from the optimum trait by $z$,
+and the allele has effect size $-u$ in heterozygotes,
+then the fitness advantage of the allele would be
+$s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$,
+where in our parameterization, $\beta = 1 / 450$.
+This tells us two things:
+(1) the rate of adaptation decreases as the population approaches the optimum,
+and (2) larger mutations (in the right direction) are more likely to fix.
+
+\paragraph{New mutations}
+The total rate of appearance of new mutations per lake is $\mu_L = 0.04$,
+which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive
+in either direction.
+This implies that a new additive or dominant effect mutation appears once every 87.5 generations,
+on average.
+Effect sizes are Exponential with mean $1/2$, and as above the probability
+and so the probability that a dominant mutation manages to establish in a population differing from
+the optimum by $z$ is roughly
+$\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$,
+and so the rate of establishment of dominant mutations is $\beta z / 87.5$,
+i.e., about one such mutation every $2461/z$ generations.
+The distribution of these successfully established mutations 
+has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2.
+Since additive alleles have half the effect in heterozygotes,
+they have half the probability of establishment.
+Combining these, we expect during the initial phase of adaptation,
+when the populations begin at around distance $z=10$ from the optimum,
+adaptive alleles to appear through mutation on a time scale of 250 generations,
+and each to move the trait by a distance of order 1.
+
+\paragraph{Migration}
+A allele that moves the trait $z=1$ units in the freshwater direction
+has fitness roughly $\exp(-\beta) \approx 1 - \beta$ in the marine environment
+(which is close to optimal).
+The product of population size and fitness differential in the marine environment
+is therefore $2Ns = 8.9$, implying that these alleles are strongly selected against
+but might occasionally drift to moderate frequency.
+
+TODO: estimate rate at which migrant alleles move between lakes
+
+%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
 
@@
 
-\clearpage
-%\plr{some problems because you had bibliographystyle before bibliography}
+\paragraph{The adaptive filter?}
+RAMBLING THOUGHTS HERE
+Since larger effect alleles are more likely to establish,
+be it by mutation or migration,
+repeated colonization of new freshwater habitats will select for larger alleles,
+be it single alleles or haplotypes bound together by an inversion.
+However, these are more strongly selected against in the interstitial time.
+Being recessive would help with this,
+but would also make it more difficult to establish.
+
 \bibliographystyle{plainnat}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
+\appendix
 \section{Supp.}
-
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=\linewidth]{plotlyPlots/NumFAA.png}
-  		\caption{(PLACEHOLDER)}
-		\label{fig:NumFAA}
-	\end{center}
-\end{figure}
 

### 2018-07-17 Stickleback_Paper.tex
abstract

@@
 
-Trait | Phenotype = Probably Trait
-
 effect loci (locus) | QTL | causal: = ?
@@
 
-should we be using the word ``patch"?
+\textbf{Pending things to do?} that maybe we do, maybe not
+
+longer genomes ($10^8$ not $10^5$ but rescaling everything accordingly)
+
+verify that adaptation of introduced pops is due to initial diversity not subsequent migration
 
@@
 
-\todo{Do last.}
+Ninespine stickleback provide one of the most striking examples of local adaptation,
+as the hemisphere-wide metapopulation includes both marine populations
+and a large number of smaller freshwater populations
+that have apparently repeatedly adapted to freshwater conditions
+often using the same genetic basis.
+In this paper, we use simulations motivated by stickleback populations
+to examine what amounts of gene flow favor stable metapopulation polymorphism
+with allele sharing,
+and to further dissect the underlying dynamics.
+We find that rapid, repeated adaptation 
+using alleles maintained at low frequency by migration-selection balance
+(the ``transporter hypothesis'')
+occurs over a realistic range of intermediate rates of gene flow,
+between slow, independent adaptation with low gene flow
+and large migration load at high gene flow.
+We do not see evidence for strong effects encouraging genomic clustering of causal alleles
+or towards particular dominance coefficients.
+$F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow,
+The results support existing theory of local adaptation,
+and provide a more concrete look at a particular, empirically motivated example.
 
 \section{Introduction}
-
-\todo{Take out any "we's" and be more specific}
 

### 2018-07-17 Stickleback_Paper.tex

@@
 and large migration load at high gene flow.
+This is mediated mainly by the total downstream influx of alleles \emph{(maybe??)}.
 We do not see evidence for strong effects encouraging genomic clustering of causal alleles
@@
 
-\jgg{Starting a paper is awkward, help}
-The rate and tempo of evolution remains largely unknown.
-Today, we have found multiple examples as evidence of rapid evolution 
-in tens of generations or fewer.
-Biologists are now starting to address the complexity behind ecological speciation. 
-\jgg{Peter and Bill: Could use help with examples of rapid evolution and the story leading up to stickleback right here}
+\emph{intro outline:}
+Start off with stickleback: repeated, rapid, parallel adaptation with obvious phenotypes.
+But, most species don't seem to have such obviously structured polymorphism: 
+what's special about stickleback?
+The transporter hypothesis is a mechanism; what's required to make it work?
+Other species adapt rapidly in similar ways:
+e.g., Darwin's finches and island mice;
+do these differ through not having such a big, connected reservoir of genetic variation?
+or by not having such a large total population size of the ``other'' habitat?
+Or, is this situation more common than we know, with less obvious phenotypes?
+Also, can we actually tell what's happening from genetic data?
 
 Recently, multiple instances of similar (parallel) underlying genetic basis of rapid local adaptation has brought about 
-questions surrounding the origins of adaptive alleles.
-A empirical model is the alaskan populations of freshwater and marine Ninespine Stickleback fish.
+questions surrounding the origins and maintenance of genetic polymorphism
+A empirical model is the Alaskan populations of freshwater and marine Ninespine Stickleback fish.
 In 1964, The Great Alaskan Earthquake caused an uplift of Middleton island and in turn,
@@
 i.e., between 0.01 and 10 migrants per year to and from each of the ten lakes.
-This had a strong effect on many aspects of adaptation: 
-including time to adaptation,
-sharing of pre-existing freshwater adapted alleles,
+This had a strong effect on many aspects of adaptation,
+including how fast adaptation occurred in each lake,
+how much alleles were shared between lakes,
 and the population genetic signals left behind.
@@
 
+\emph{merge with the above}
 Interestingly, the ability for separate populations to locally adapt to their own selective pressure was relatively unaffected 
@@
 until the trait means were close to the optimal values in each habitat. 
-(Remaining simulations are shown in Figures \ref{phenotype_ts1} and \ref{phenotype_ts4}.)
+(Remaining simulations are shown in Figures \ref{fig:phenotype_ts1} and \ref{fig:phenotype_ts4}.)
 Greater migration decreased adaptedness only weakly
@@
 where the mean trait value equilibrated at roughly half its optimal value 
-(Figures \ref{fig:MeanPhenotype} and \ref{phenotype_ts4}).
-With less than one migrant per generation per lake,
-adaptation happened substantially faster
-in the marine habitat than in the lakes.
+(Figures \ref{fig:MeanPhenotype} and \ref{fig:phenotype_ts4}).
 The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
-as jumps in the mean phenotype curve;
+as jumps in the mean trait value;
 in the two simulations these occur on a time scale of 100 (lower migration) to 1000 (higher migration)
 generations,
-and move the phenotype by an amount of order 1.
+and move the trait by an amount of order 1.
 Trait variation within each population was small compared to the difference between populations,
@@
 Local adaptation was still able to occur despite overall homogenization:
-$F_{ST}$ between marine and freshwater habitats 
-at sites with alleles affecting the trait (``effect mutations'')
-was relatively constant across migration values.
+if computed using only sites with alleles affecting the trait (``effect mutations''),
+$F_{ST}$ between habitats was relatively constant across migration values.
 
@@
 
-As expected,
-lakes were able to adapt faster at higher migration rates
-because greater mixing allowed them to share alleles
+The tenfold difference in speed of adaptation occurs because
+at low migration rates, adaptation occurs independently in each lake,
+and the marine habitat has ten times the influx of new alleles than any one lake.
+In other words, 
+greater mixing at higher migration rates allows lakes to share alleles
 instead of developing their own genetic basis of adaptation.
 As a first indication of this, Figure \ref{fig:Fst} shows that
-$F_{ST}$ between the two sets of lakes at effect mutations decreased with migration rate.
+$F_{ST}$ between the ``original'' and ``introduced'' sets of lakes 
+at effect mutations decreased with migration rate.
 
-To further investigate sharing of locally adaptive alleles, 
+To further investigate sharing of locally adaptive alleles
+and the transporter hypothesis,
 we investigate the ``pre-existing freshwater adapted alleles'',

### 2018-07-17 Stickleback_Paper.tex
Update on Overleaf.

@@
 \newcommand{\jgg}[1]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=black]{#1}}
+\newcommand{\bill}[1]{\todo[linecolor=green,backgroundcolor=yellow!25,bordercolor=black]{#1}}
 
@@
 Recently, multiple instances of similar (parallel) underlying genetic basis of rapid local adaptation has brought about 
-questions surrounding the origins and maintenance of genetic polymorphism
+questions surrounding the origins and maintenance of genetic polymorphism.
 A empirical model is the Alaskan populations of freshwater and marine Ninespine Stickleback fish.
@@
 \jgg{Bill - more specifics about Bill, Suzie, Kristin  \& or Thom's papers .. ? }
+\bill{no way}
 These results leave us with questions surrounding the nature of rapid adaptation.

### 2018-07-18 Stickleback_Paper.tex
Update on Overleaf.

@@
 \title{The sweet spot for local polygenic adaptation: a case study motivated by stickleback}
-\author{Jared Galloway, Bill Cresko, and Peter Ralph}
+\author{Jared Galloway, William A. Cresko, and Peter Ralph}
 \maketitle
@@
 A simulation study of gene flow and local polygenic adaptation
+
+\bill{A simulation study of gene flow and local polygenic adaptation in stickleback fish}
 
@@
 
-Ninespine stickleback provide one of the most striking examples of local adaptation,
-as the hemisphere-wide metapopulation includes both marine populations
-and a large number of smaller freshwater populations
+Threespine stickleback fish provide one of the most striking examples of local adaptation.
+This hemisphere-wide metapopulation includes both marine populations and a large number of smaller freshwater populations
 that have apparently repeatedly adapted to freshwater conditions
@@
 and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region.
-\jgg{Bill - more specifics about Bill, Suzie, Kristin  \& or Thom's papers .. ? }
-\bill{no way}
+% \jgg{Bill - more specifics about Bill, Suzie, Kristin  \& or Thom's papers .. ? }
 These results leave us with questions surrounding the nature of rapid adaptation.
@@
 in SLiM \citep{haller2017slim}.
-The details (described below) are motivated by the nine-spine stickleback,
+The details (described below) are motivated by previous population genomic studies of threespine stickleback,
 but remain simplistic in some aspects due to computational constraints.
 Possibly the most important caveat is that simulated population sizes are much smaller
-than the census size of the nine-spine stickleback population;
+than the census size of the threespine stickleback population;
 we will consider the likely effects of this in the Discussion.
@@
 \paragraph{Habitat and geography}
-Our model has two habitat types, which are separate populations
-with a fixed total size of 2,000 diploid individuals each.
+Our model has two habitat types, which are considered separate populations
+with a fixed total size of 2,000 diploid individuals each. The marine population is treated as a single large deme, but the freshwater population is subdivided into local subpopulations or demes.
 The scheme, depicted in Figure \ref{fig:Geo},
@@
 
-Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff,
-to roughly model the cumulative effect of the various phenotypes
-such as lateral plating and opercle shape
+Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff such that alternative values of the trait have high or low fitness in tht two different habitats.
+This situation roughly models the cumulative effect of the various phenotypes
+such as armor morphology, body size, craniofacial variation and opercle shape (among others)
 on which divergent selection acts in the two environments.
 The optimal trait value in the marine habitat is $+10$,
-and in the freshwater habitat is $-10$,
-and fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$ is
+and in the freshwater habitat is $-10$. Fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$ is
 Gaussian with standard deviation, i.e.,
@@
 
-\paragraph{Potential genomic architecture}
+\paragraph{Potential genomic architecture}\bill{Initial genomic architecture? The header is a little confusing to me}
 Each individual carries two copies of a linear chromosome of $10^5$ loci
@@
 also proportionally to fitness but reweighted by a Gaussian function 
-of the distance separating the two, with stanadard deviation $1/2$: 
-concretely, if the first parent is marine individual $i$,
+of the distance separating the two, with stanadard deviation $1/2$. 
+Concretely, if the first parent is marine individual $i$,
 then marine individual $j$ is chosen as the mate
@@
 \paragraph{New lakes}
-To study adaptation in newly appearing habitats,
+To study adaptation in newly appearing freshwater habitats,
 we introduce a new set of lakes midway through the simulation.
@@
 
-\paragraph{Descriptive statistics}
+\paragraph{Descriptive statistics}\bill{Population genetic analyses? Instead of the more general 'Descriptive statistics'?}
 To assess whether new lakes adapt using existing genetic diversity,

### 2018-07-26 Stickleback_Paper.tex
more theory

--- 2018-07-17 Stickleback_Paper.tex
@@
 on average.
-Effect sizes are Exponential with mean $1/2$, and as above the probability
+Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, 
 and so the probability that a dominant mutation manages to establish in a population differing from
@@
 they have half the probability of establishment.
-Combining these, we expect during the initial phase of adaptation,
-when the populations begin at around distance $z=10$ from the optimum,
-adaptive alleles to appear through mutation on a time scale of 250 generations,
+During the initial phase of adaptation,
+the populations begin at around distance $z=10$ from the optimum.
+Combining these facts, we expect 
+adaptive alleles to appear through mutation at first on a time scale of 250 generations,
+with the time between local fixation of new alleles increasing as adaptation progresses,
 and each to move the trait by a distance of order 1.
 
-\paragraph{Migration}
-A allele that moves the trait $z=1$ units in the freshwater direction
-has fitness roughly $\exp(-\beta) \approx 1 - \beta$ in the marine environment
+\paragraph{Standing variation}
+An allele that moves the trait $z$ units in the freshwater direction in heterozygotes
+has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment
 (which is close to optimal).
 The product of population size and fitness differential in the marine environment
+for a mutation with $z=1$
 is therefore $2Ns = 8.9$, implying that these alleles are strongly selected against
 but might occasionally drift to moderate frequency.
+The average frequency of such an allele in the marine environment at migration-selection equilibrium
+is equal to the total influx of alleles per generation divided by the selective disadvantage,
+which if $M = 2000 m$ is the number of immigrants per generation,
+is $2 M / \beta z^2$. 
+With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$:
+since lakes have 400 genomes, 
+as long as $M \ge 1$, the chances are good that any particular lake-adapted allele
+that is present in all pre-existing lakes
+will appear at least once in the fish that colonize a new lake.
+However, an allele with $z=1$ only has probability of around 1/20 of establishing locally,
+suggesting that we'd need $M \ge 10$ to ensure enough pre-existing genetic variation
+that adaptation would happen entirely using the initial set of colonizers.
+This corresponds to our highest two migration rates,
+as in e.g., Figure \ref{fig:phenotype_ts2}B.
 
-TODO: estimate rate at which migrant alleles move between lakes
+\paragraph{Migration}
+If sufficient genetic variation is not present in a new lake initially,
+it must appear by new mutation or migration.
+Since a proportion $m$ of each lake is composed of migrants each generation,
+it takes $1/m$ generations until the genetic variation introduced by migrants
+equals the amount initially present at colonization.
+This implies a dichotomy: either 
+(a) adaptation is possible using variants present at colonization
+or arriving shortly thereafter, or 
+(b) adaptation takes many multiples of $1/m$ generations.
+
+These calculations depend on there being no bottleneck in colonization of the lake;
+if there is a bottleneck, then an additional factor must be added.
+
+At what point do we expect new mutation to be more important than migration for adaptation?
+By the calculations above, if $M \ge 10$, we expect initial diversity in a lake to be sufficient 
+for adaptation, corresponding to our third-highest migration rate.
+If this does not happen, then we expect adaptation to take a multiple of $1/m$ generations;
+with our values, $1/m$ ranges from 20,000 to 20 generations.
+Above we estimated that adaptive alleles due to new mutation fix locally about every 2,000 generations,
+suggesting that at our second-lowest migration rate (where $1/m = 2,000$),
+the two contributions of migration and new mutation are roughly equal.
+This is in fact what we see: in Figure \ref{fig:MPFAI},
+we see that at the second migration rate, alleles start to be shared between lakes,
+while by the third migration rate, they are almost entirely shared.
 

### 2018-08-17 Stickleback_Paper.tex
rm DS file

--- 2018-07-26 Stickleback_Paper.tex
@@
 
+Trait | Phenotype = Probably Trait
+
 effect loci (locus) | QTL | causal: = ?
@@
 
-\textbf{Pending things to do?} that maybe we do, maybe not
-
-longer genomes ($10^8$ not $10^5$ but rescaling everything accordingly)
-
-verify that adaptation of introduced pops is due to initial diversity not subsequent migration
+should we be using the word ``patch"?
 
@@
 
-Ninespine stickleback provide one of the most striking examples of local adaptation,
-as the hemisphere-wide metapopulation includes both marine populations
-and a large number of smaller freshwater populations
-that have apparently repeatedly adapted to freshwater conditions
-often using the same genetic basis.
-In this paper, we use simulations motivated by stickleback populations
-to examine what amounts of gene flow favor stable metapopulation polymorphism
-with allele sharing,
-and to further dissect the underlying dynamics.
-We find that rapid, repeated adaptation 
-using alleles maintained at low frequency by migration-selection balance
-(the ``transporter hypothesis'')
-occurs over a realistic range of intermediate rates of gene flow,
-between slow, independent adaptation with low gene flow
-and large migration load at high gene flow.
-This is mediated mainly by the total downstream influx of alleles \emph{(maybe??)}.
-We do not see evidence for strong effects encouraging genomic clustering of causal alleles
-or towards particular dominance coefficients.
-$F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow,
-The results support existing theory of local adaptation,
-and provide a more concrete look at a particular, empirically motivated example.
+\todo{Do last.}
 
 \section{Introduction}
+
+\todo{Take out any "we's" and be more specific}
 
@@
 
-\emph{intro outline:}
-Start off with stickleback: repeated, rapid, parallel adaptation with obvious phenotypes.
-But, most species don't seem to have such obviously structured polymorphism: 
-what's special about stickleback?
-The transporter hypothesis is a mechanism; what's required to make it work?
-Other species adapt rapidly in similar ways:
-e.g., Darwin's finches and island mice;
-do these differ through not having such a big, connected reservoir of genetic variation?
-or by not having such a large total population size of the ``other'' habitat?
-Or, is this situation more common than we know, with less obvious phenotypes?
-Also, can we actually tell what's happening from genetic data?
+\jgg{Starting a paper is awkward, help}
+The rate and tempo of evolution remains largely unknown.
+Today, we have found multiple examples as evidence of rapid evolution 
+in tens of generations or fewer.
+Biologists are now starting to address the complexity behind ecological speciation. 
+\jgg{Peter and Bill: Could use help with examples of rapid evolution and the story leading up to stickleback right here}
 
 Recently, multiple instances of similar (parallel) underlying genetic basis of rapid local adaptation has brought about 
-questions surrounding the origins and maintenance of genetic polymorphism
-A empirical model is the Alaskan populations of freshwater and marine Ninespine Stickleback fish.
+questions surrounding the origins of adaptive alleles.
+A empirical model is the alaskan populations of freshwater and marine Ninespine Stickleback fish.
 In 1964, The Great Alaskan Earthquake caused an uplift of Middleton island and in turn,
@@
 
-% Overview
-We varied the migration rate, $m$,
-from $5 \times 10^{-5}$ to $5 \times 10^{-2}$ per individual,
-i.e., between 0.01 and 10 migrants per year to and from each of the ten lakes.
-This had a strong effect on many aspects of adaptation,
-including how fast adaptation occurred in each lake,
-how much alleles were shared between lakes,
-and the population genetic signals left behind.
-At the lowest migration rate, lakes adpated nearly independently,
-while at the highest migration rate, the habitats are beginning to act like a single metapopulation
-with substantial migration load.
-We will now dissect what happens between these extremes.
-
-\emph{merge with the above}
+Across a range of historical introgression from $\approx 0.01 - 100.0$ migrants per generation in populations of 
+size 2000 diploid individuals, we encountered large influence on 
+time to adaptation,
+sharing of pre-existing freshwater adapted alleles,
+and population genetic signals left behind in the simulated populations.
+More gene flow is experienced with greater historical introgression between the two selective pressures.
 Interestingly, the ability for separate populations to locally adapt to their own selective pressure was relatively unaffected 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Local Adaptation: differentiation with gene flow}
-
-% The progression of adaptation. How you can see introduction new alleles
-
-Local adaptation occurred in all simulations:
-as shown in Figure \ref{fig:phenotype_ts2},
-freshwater and marine populations diverged
-until the trait means were close to the optimal values in each habitat. 
-(Remaining simulations are shown in Figures \ref{fig:phenotype_ts1} and \ref{fig:phenotype_ts4}.)
-Greater migration decreased adaptedness only weakly
-except for at the highest migration rate (10 migrants per generation per lake),
-where the mean trait value equilibrated at roughly half its optimal value 
-(Figures \ref{fig:MeanPhenotype} and \ref{fig:phenotype_ts4}).
-The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
-as jumps in the mean trait value;
-in the two simulations these occur on a time scale of 100 (lower migration) to 1000 (higher migration)
-generations,
-and move the trait by an amount of order 1.
-Trait variation within each population was small compared to the difference between populations,
-with interquartile ranges of around XXX.\plr{TODO}
-Across all parameter values, differences at around 16 commonly polymorphic sites 
-(eight the shift the trait in each direction)
-were responsible for most of the adaptive differences between freshwater and marine habitats.
-\plr{ALSO TODO: CHECK THIS}
-
 \begin{figure}
 	\begin{center}
-        \includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4}
-        \includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3}
-  		\caption{ 
-        Mean individual trait values in the marine habitat (black line),
-        the original lakes (light green lines; average in pink),
-        and the introduced lakes (dark green lines; average in orange)
-        across the course of two simulations, with migration rates of
-        \textbf{(top)} $m=5 \times 10^{-4}$ and
-        \textbf{(bottom)} $m=5 \times 10^{-3}$ per generation individual
-        (i.e., 0.1 and 1.0 migrants per generation per lake, respectively).
-        Optimal trait values in the two habitats are at $\pm 10$.
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenoThroughout5e-5.png}
+  		\caption{ Mean individual trait value in each population as a function of time $t$, in generations. 
+		This is a single simulation run at a migration rate of $5 \times 10^{-4} \approx 1$ mig/gen in both 
+		directions for every freshwater population to the single marine population. 
+		The dark black line represents the mean phenotype of the single marine population with an optimum trait value of + 10. 
+		The pink represents the entire initial freshwater population mean
+		of all the lakes (light green) with and optimum trait value of -10.
+		The orange represents the entire introduced freshwater population mean
+		of all the lakes (dark green) with an optimum trait value of -10 as well.
 		}
@@
 
-\jgg{also, interesting point to make on second plot, you can see the initial lakes taking the same size step, showing sharing of alleles.
-could zoom in on the relevent portions of the figure for this. (begging and introduction)
- a 2 by 2 with the important sections (0 - 20K) or (40-50K) for both above figures?}
-
-\begin{figure}[h!tb]
+\begin{figure}
 	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MeanPhenotype1.png}
-  		\caption{
-            Distribution of mean trait values throughout each simulation.
-            The dashed pink and purple lines show the optima ($\pm 10$).
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenoThroughout5e-4.png}
+  		\caption{Mean individual trait value in each population as a function of time $t$, in generations. 
+		This is a single simulation run at a migration rate of $5 \times 10^{-3} \approx 1$ mig/gen in both 
+		directions for every freshwater population to the single marine population. 
+		The dark black line represents the mean phenotype of the single marine population with an optimum trait value of + 10. 
+		The pink represents the entire initial freshwater population mean
+		of all the lakes (light green) with and optimum trait value of -10.
+		The orange represents the entire introduced freshwater population mean
+		of all the lakes (dark green) with an optimum trait value of -10 as well.
 		}
-  		\label{fig:MeanPhenotype}
+  		\label{fig:phenotype_ts3}
 	\end{center}
 \end{figure}
-\plr{Is this the whole simulation? what about the initial adaptation phase?
-     Is the horizontal line at the median? what are the whiskers at?}
 
-As expected, increasing migration rate decreased differentiation between habitats.
-As seen in Figure \ref{fig:Fst},
-$F_{ST}$ between marine and freshwater habitats at neutral sites 
-steadily declines as migration increases.
-Local adaptation was still able to occur despite overall homogenization:
-if computed using only sites with alleles affecting the trait (``effect mutations''),
-$F_{ST}$ between habitats was relatively constant across migration values.
+\jgg{we should make one caption and the two above should be top-bottom sub-figures}
+\jgg{also, interesting point to make on second plot, you can see the initial lakes taking the same size step. sharing}
+%I could zoom in on the relevent portions of the figure for this. (begging and introduction)
+% a 2 by 2 with the important sections (0 - 20K) $ (40-50K) for both above figures. share axes
+%\jgg{finished with most edits up to here}
+
+\subsection*{Local Adaptation: differentiation with gene flow}
+
+%Talk about the genetic basis of trait. 
+%Talk about the progression of adaptation. How you can see introduction new alleles
+
+Local adaptation occurred across all simulated parameter values.
+Starting from the same baseline, freshwater and marine populations diverged phenotypically
+until they reached an equilibrium
+where the population means were close to or at the optimal phenotype. 
+The distribution of mean number of freshwater adapted alleles suggested that there was 
+$\approx 8$ SNPs underlying the freshwater phenotype. 
+This was true for all rates of introgression except for the highest at $100$ mig/gen
+where the mean trait value in the populations equilibrated well below the optimum.
+We can see the introduction of new alleles in Figure \ref{fig:phenotype_ts2}. 
+In generation (0 - 20K) The sweeps that cause jumps in the mean phenotype of the freshwater population 
+have a relatively large effect size equal to or greater than $1$.
+%In Figure \ref{fig:counts} we can see that the number of effect regions with zero 
+%pre-existing freshwater adapted alleles increased with introgression while
+%regions of effect containing more pre-existing freshwater adapted alleles increase.
+Phenotypic variation within each population was small compared to the difference between populations.
+Polymorphic loci that impact the individual trait value,
+showed greater differentiation compared to neutral loci between marine and freshwater populations, 
+with the exception of the lowest migration rate of 0.01 migrants per generation per lake.
+
+\todo{ add more general observations and caveats about not all parameter values. }
+\plr{It would be nice to have a small table of parameter values,
+    so that when someone wonders ``What was the population size again?'' they could easily find it.}
+\jgg{Re: agreed}
+
+\subsection*{Migration homogenizes genetic variation}
+
+%FIGURES THAT SHOW THIS:
+%- Fst Plots.
+%- SGV Plots.
+%- Haplotype plots
 
@@
 	\begin{center}
-  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/FST_HRR.png}
-  		\caption{
-            Distributions of $F_{st}$ values between the three habitats,
-            accumulated across the course of the simulation, calculated 
-            \textbf{(left)} using only neutral mutations, and
-            \textbf{(right)} using effect mutations.
-        }
+  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/Fst_Adaptation.png}
+  		\caption{Distributions of $F_{st}$  values between the three subpopulations throughout the simulations.
+		On the left, $F_{st}$ was calculated only on neutral mutation frequencies. On the right, $F_{st}$ was calculated
+		only on effect (Impact on phenotype) mutation frequencies. 
+		Effect mutations are acted upon by selection as they have an affect on fitness where neutral mutations  }
   		\label{fig:Fst}
@@
 
-\paragraph*{Speed of adaptation}
-Increasing migration rate strongly decreased the time until freshwater habitats could adapt,
-both in the initial and introduced sets of lakes.
-As shown in Figure \ref{fig:TimeToAdaptation}, at the lowest migration rate 
-it took over 20 thousand generations for the mean trait value across introduced lakes to 
-get to within 0.5 of the original lakes value.
-Although many lakes had adapted before this time,
-the different rates of introduction of effect alleles is clearly seen in the traces
-of trait values against time (e.g., Figure \ref{fig:phenotype_ts2}).
-However, higher migration rates allowed the freshwater habitat to adapt
-nearly as quickly as the marine habitat.
+Across increasing parameters of migration, we observed more gene flow across the populations. 
+As can be seen in Figure \ref{fig:Fst},
+$F_{st}$ values for neutral alleles steadily decline between all sets of populations as 
+migration increases, this illustrates less difference between sections of the genome which are not acted
+upon by selection, in all populations.
+Additionally, standing genetic variance
+(seen in Figure \ref{fig:SGV})
+values in the marine population steady increased with migration which suggests that more alleles from the 
+lakes surrounding were exporting alleles into the marine individual.
 
+%TIES TO NATURE?
+%SPECIFICALLY (interesting stuff)
+
+\subsection*{migration affects the speed of adaptation}
+
+%FIGURES THAT SHOW THIS:
+%-Time to adaptation
 \begin{figure}
 	\begin{center}
-        \includegraphics[width=0.8\linewidth]{matplotlibPlots/TimeToAdaptation}
-        \caption{
-            Time to adaptation as a function of migration rate ($m$),
-            measured as the number of generations
-            until the introduced population's mean phenotype is within 0.5 
-            of the original lakes average phenotype. 
-            Each point represents a single simulation, and
-            the line shows the average of all points at each value of $m$.
-        }
+  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/Adaptation_2_matplot.png}
+  		\caption{Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
+		it takes for the introduced population's mean phenotype to come within 0.5 of the original lakes average phenotype. 
+		Each point represents a simulation run at some value of $M$. 
+		The yellow dashed line is the average of all points at each respective parameter value}
   		\label{fig:TimeToAdaptation}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Sharing of freshwater adapted alleles}
+We observed a dramatic shift in time until adaptation for the introduced populations
+as migration rates increased. 
+As can be seen in Figure \ref{fig:TimeToAdaptation} at the lowest migration rate of $5 \times 10^{-5}$,
+It took the introduced population over 20 thousand generations for the mean trait value, $X_{I}$ of all the lakes to 
+get to within 0.5 of the original lakes mean trait value, $X_{O}$
+%\plr{From the plots it looks more like adaptation happens (at least in most lakes) by like 10K or 15K.}
+This suggests that selection was acting primarily on new mutations, and that the 
+introduced lakes needed to wait for a beneficial mutation to arise before 
+it was selected upon. 
+%\plr{We should do the calculation to see if this makes sense: 
+%    there's one new effect mutation in the right direction every 100 generations in the lakes,
+%    one of effect size $s$ has probability of about $s/2$ of not being lost to drift...}
+In Figure \ref{fig:MPFAI} we can see there was a correlation between the amount of pre-existing freshwater adapted alleles in the 
+marine and the rapid adaptation of the introduced populations. 
+In the same figure we can see the distributions
+for both the original and Introduced freshwater population become more balanced (similar?) as introgression increases.
+This is further evidence that the introduced population is taking advantage of the alleles from the initial freshwater populations
+
+%(2) Do the math, discuss number ( and possibly type ) of mutations underlying the trait value.
+
+%TIES TO NATURE?
+
+\subsection*{sharing of freshwater adapted alleles}
 
@@
 %-Total/Avg shared
-
-The tenfold difference in speed of adaptation occurs because
-at low migration rates, adaptation occurs independently in each lake,
-and the marine habitat has ten times the influx of new alleles than any one lake.
-In other words, 
-greater mixing at higher migration rates allows lakes to share alleles
-instead of developing their own genetic basis of adaptation.
-As a first indication of this, Figure \ref{fig:Fst} shows that
-$F_{ST}$ between the ``original'' and ``introduced'' sets of lakes 
-at effect mutations decreased with migration rate.
-
-To further investigate sharing of locally adaptive alleles
-and the transporter hypothesis,
-we investigate the ``pre-existing freshwater adapted alleles'',
-defined for a particular generation
-to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat
-in that generation.
-Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles,
-across generations,
-as well as the average number of lakes each is found in.
-As migration rates increase, the number of these alleles decreases steadily,
-and each is concurrently found in a greater number of lakes.
-
-As the number of common, locally adaptive alleles decreases,
-the genetic basis of adaptation is more commonly shared.
-Figure \ref{fig:MPFAI}B shows
-the distribution of the mean percentage of currently-defined freshwater adapted alleles 
-that each genome in each of the populations carries,
-averaged across time and individuals.
-If all individuals across lakes carried the same set of alleles determining their trait value,
-this would be 100\%.
-This value is nearly reached at the highest migration rate;
-but it is lower due to migration load.
-At the lowest migration rate, 
-each genome in the original lakes have almost exactly $1/10^{th}$ 
-of the total freshwater adapted alleles --
-since there are 10 lakes, this implies that each lake has adapted with a unique set of alleles.
-Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
-
 \begin{figure}
 	\begin{center}
-        \includegraphics[width=\linewidth]{plotlyPlots/NumFAA}
-        \includegraphics[width=\linewidth]{matplotlibPlots/MPFAI}
-  		\caption{
-            \textbf{(A)}
-            Distribution of the number of freshwater adapted alleles (green)
-            and the average number of lakes each allele is present in (red)
-            across the course of the simulation.
-            \textbf{(B)}
-            Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
+  		\includegraphics[width=\linewidth]{matplotlibPlots/MPFAI_2_matplot.png}
+  		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
 		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Migration load and genetic variation}
+To further investigate this correlation, we take a look at the freshwater adapted alleles driving 
+local adaptation of the introduced population. 
+One common metric we investigate is the percentage of freshwater adapted alleles per individual in each of the populations, at any time $t$. 
+This gives us a idea of where the where these alleles are being distributed throughout the simulation run. 
+Looking at the lowest migration rate for the original lakes population In Figure \ref{fig:MPFAI}
+We see that the average individual has almost exactly $1/10^{th}$ of the total defined freshwater adapted alleles throughout the simulation. 
+Because FAA are defined among \textit{any} population, this suggests that each one of the $10$ 
+lakes has created their own solution to adaptation of the the freshwater selective pressure.
+
+As we can see in Figure \ref{fig:NumFAA}
+as historical introgression increases for the simulations, 
+we see the distribution of the total number of freshwater adapted alleles decrease, and 
+the average number of lakes each allele appears in at high frequency ($p > 0.5$), increase.
+With the distribution of effect size remaining the same across all simulations, 
+these are both suggestive of the original lakes sharing solutions to the same selective pressure.
+
+%TIES TO NATURE?
+
+%SPECIFICALLY (interesting stuff)
+
+\subsection*{do the math}
+\jgg{peter? :)}
+
+(1) - What is the mean fitness of a population
+
+(2) - assuming everybody is at the mean fitness, what is the expected num of offspring after t gens in marine
+
+(3) - expected num of establishing alleles in new lake. = (num outmigrants) $\times$ (num per outmigrant) $\times$ (prob migrant) $\times$ $2\hat{x}s $ mean fit in new lake
+
+\subsection*{We see Migration Load after a certain threshold of migration}
 
@@
 %- Phenotype Distribution
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/Pheno_2_matplot.png}
+  		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
+		The dashed pink line (Pheno = +10) is the optimum phenotype any individual in the marine environment.
+		In contrast the purple line at (Pheno = -10) represents the optimum for any individual in the freshwater environment. 
+		All individuals at generation 0 (beginning of the simulation) 
+		}
+  		\label{fig:MeanPhenotype}
+	\end{center}
+\end{figure}
+
 %GENERAL
 
-At first, increased migration allows sharing of adaptive alleles between lakes,
-but at the highest migration rate,
-the constant influx of alleles between the habitats creates substantial migration load.
-This rate, $m=.05$, only replaces 5\% of each population each generation
-with migrants from the other habitat, but this is sufficient to shift the mean trait values
-to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
+In the distributions we have shown across migration rate parameter values, 
+We have experienced the most dramatic shifts of the population dynamics at $M = 5x10^{-3}$.
+After a threshold between this and $M = 5x10^{-2}$, all populations begin to experience migration load. 
 Significant gene flow constricts local adaptation
 as a consequence of a large number of offspring through hybridization events between subpopulations.
+In Figure \ref{fig:MeanPhenotype} at $M = 5x10^{-2}$, we can see the distribution of average phenotype throughout the simulation
+pull towards the opposing selective pressure value and away from the local optimum trait value in all populations.
 
-\paragraph{Standing genetic variation}
-Despite a dramatic difference in the amount of allelic sharing between lakes,
-standing genetic variance (Figure \ref{fig:SGV}) was around 0.05 in the freshwater habitat
-across all but the highest migration rates.
-Concurrently,
-genetic variation in the marine population steady increased with migration
-to a similar value.
-On the one hand, it is not surprising that lakes, as a group,
-carry more genetic variation than the marine habitat,
-since population subdivision allows different alleles to become common in each.
-However, it is striking that at $m=.005$, the marine habitat carries as much genetic variation,
-despite a lack of any substantial migration load.
+%TIES TO NATURE?
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Realized genomic architecture}
+%SPECIFICALLY (interesting stuff)
+
+\subsection*{Realized genomic architecture and signals}
+
+%\begin{figure}
+%	\begin{center}
+%  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3.png}
+%  		\caption{(Description and cleaning needed)
+%		}
+%  		\label{fig:Fst3}
+%	\end{center}
+%\end{figure}
+
+%\begin{figure}
+%	\begin{center}
+%  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-4.png}
+%  		\caption{(Description and cleaning needed)
+%		}
+%  		\label{fig:Fst2}
+%	\end{center}
+%\end{figure}
+
+Now we take a closer look at the genomic architecture of local adaptation between freshwater and marine individuals. 
+Often when looking at real datas, biologists observe patterns of $F_{st}$ and clustering of alleles that suggest causative loci for certain traits.
+However, researchers would like to know how parameters such as migration and recombination impact the results from natural populations.
+To give an idea about the impact of these parameters in our model, we examine the distribution, location, and false positive rate on predicting effect loci
+given the $F_{st}$ per SNP across the genome.
 
@@
 	\begin{center}
-        \includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3}
-        \includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-4}
-  		\caption{(Description and cleaning needed)
-		}
-  		\label{fig:Fst3}
-	\end{center}
-\end{figure}
-
-Now we take a closer look at the genomic architecture of local adaptation 
-between the two habitats.
-Do the alleles underlying trait differences cluster along the genome?
-Do measures of local differentiation identify the causal loci?
-Figure \ref{fig:Fst3} shows plots along the genome of per-locus $F_{ST}$ values
-between the marine and original freshwater habitats
-at the two intermediate migration rates.
-(Note that we are pooling freshwater habitats; 
-a single lake would provide substantially less power.)
-
-\plr{need to know where the FAA are in these figures}
-
-Higher migration rates showed more distinct $F_{ST}$ peaks
-over polymorphic loci underlying trait differences between the habitats.
-As migration rate decreased, the ``background'' levels of $F_{ST}$ increased,
-swamping out this signal until the regions under selection were indistinguishable. 
-This is likely due to two reasons: 
-first, stronger genetic drift with less migration leading to higher background $F_{ST}$,
-and second, greater sharing of adaptive alleles providing a shared signal across populations.
-
-This suggests that genome scans for local adaptation
-based purely on measures of differentiation
-will only be successful given enough migration between habitats.
-To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates
-that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value
-to be a causal locus.
-\plr{Labels are wrong on that figure: false pos rate should go down with cutoff.
-     Is it true positive rate?}
-
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{matplotlibPlots/Power_FP.png}
+  		\includegraphics[width=0.7\linewidth]{matplotlibPlots/0_25_power.png}
   		\caption{ 
@@
 		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
-        TODO: x-axis label ($F_{ST}$).
 		}
@@
 \end{figure}
+
+%Suppose you have a Fst Peak, ___ how likely is it that that region is causal. and conversely, what proportion of 
+%the causal loci are under Fst Peaks.
+
+\subsection*{False-Positives \& Statistical Power}
+
+%\plr{Start off by talking about Fst along the genome, then say ``if we identify QTL as the outliers, then...''}
+%\plr{I vote for combining and including figures \ref{fig:Fst2} and \ref{fig:Fst3}.}
+
+At higher rates of introgression in our simulations, we saw $F_{st}$ peaks at regions of the genome underlying individual trait value (phenotype). 
+In contrast, the lower rates of introgression showed higher $F_{st}$ across the entire genome such that the regions under selection were indistinguishable. 
+For every $F_{st}$ peak, we saw the same regions of the genome express high $F_{st}$ when comparing either the initial \textit{or} introduced freshwater populations to the marine population.
+If we identify quantitative trait loci as the outliers, then this is again suggestive of both freshwater populations sharing regions of divergence from the marine population.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Theoretical expectations}
-\plr{does this go first or second?}
-
-Here's some rough calculations to get a sense for what should be going on.
-Everything is done in more detail in OTHER PLACES WE SHOULD CITE.
-Fisher, Chevin, etc.
-
-Suppose a new allele enters a lake, either by migration or mutation.
-If, when it is rare but present in $n$ copies,
-it has fitness advantage $s$
--- i.e., the expected number of copies in the next generation is $(1+s)n$ --
-then the probability that it escapes demographic stochasticity to become common in the population
-is approximately $2s$ \citep{fisher,prob_fixation}.
-If the current population all differed from the optimum trait by $z$,
-and the allele has effect size $-u$ in heterozygotes,
-then the fitness advantage of the allele would be
-$s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$,
-where in our parameterization, $\beta = 1 / 450$.
-This tells us two things:
-(1) the rate of adaptation decreases as the population approaches the optimum,
-and (2) larger mutations (in the right direction) are more likely to fix.
-
-\paragraph{New mutations}
-The total rate of appearance of new mutations per lake is $\mu_L = 0.04$,
-which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive
-in either direction.
-This implies that a new additive or dominant effect mutation appears once every 87.5 generations,
-on average.
-Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, 
-and so the probability that a dominant mutation manages to establish in a population differing from
-the optimum by $z$ is roughly
-$\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$,
-and so the rate of establishment of dominant mutations is $\beta z / 87.5$,
-i.e., about one such mutation every $2461/z$ generations.
-The distribution of these successfully established mutations 
-has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2.
-Since additive alleles have half the effect in heterozygotes,
-they have half the probability of establishment.
-During the initial phase of adaptation,
-the populations begin at around distance $z=10$ from the optimum.
-Combining these facts, we expect 
-adaptive alleles to appear through mutation at first on a time scale of 250 generations,
-with the time between local fixation of new alleles increasing as adaptation progresses,
-and each to move the trait by a distance of order 1.
-
-\paragraph{Standing variation}
-An allele that moves the trait $z$ units in the freshwater direction in heterozygotes
-has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment
-(which is close to optimal).
-The product of population size and fitness differential in the marine environment
-for a mutation with $z=1$
-is therefore $2Ns = 8.9$, implying that these alleles are strongly selected against
-but might occasionally drift to moderate frequency.
-The average frequency of such an allele in the marine environment at migration-selection equilibrium
-is equal to the total influx of alleles per generation divided by the selective disadvantage,
-which if $M = 2000 m$ is the number of immigrants per generation,
-is $2 M / \beta z^2$. 
-With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$:
-since lakes have 400 genomes, 
-as long as $M \ge 1$, the chances are good that any particular lake-adapted allele
-that is present in all pre-existing lakes
-will appear at least once in the fish that colonize a new lake.
-However, an allele with $z=1$ only has probability of around 1/20 of establishing locally,
-suggesting that we'd need $M \ge 10$ to ensure enough pre-existing genetic variation
-that adaptation would happen entirely using the initial set of colonizers.
-This corresponds to our highest two migration rates,
-as in e.g., Figure \ref{fig:phenotype_ts2}B.
-
-\paragraph{Migration}
-If sufficient genetic variation is not present in a new lake initially,
-it must appear by new mutation or migration.
-Since a proportion $m$ of each lake is composed of migrants each generation,
-it takes $1/m$ generations until the genetic variation introduced by migrants
-equals the amount initially present at colonization.
-This implies a dichotomy: either 
-(a) adaptation is possible using variants present at colonization
-or arriving shortly thereafter, or 
-(b) adaptation takes many multiples of $1/m$ generations.
-
-These calculations depend on there being no bottleneck in colonization of the lake;
-if there is a bottleneck, then an additional factor must be added.
-
-At what point do we expect new mutation to be more important than migration for adaptation?
-By the calculations above, if $M \ge 10$, we expect initial diversity in a lake to be sufficient 
-for adaptation, corresponding to our third-highest migration rate.
-If this does not happen, then we expect adaptation to take a multiple of $1/m$ generations;
-with our values, $1/m$ ranges from 20,000 to 20 generations.
-Above we estimated that adaptive alleles due to new mutation fix locally about every 2,000 generations,
-suggesting that at our second-lowest migration rate (where $1/m = 2,000$),
-the two contributions of migration and new mutation are roughly equal.
-This is in fact what we see: in Figure \ref{fig:MPFAI},
-we see that at the second migration rate, alleles start to be shared between lakes,
-while by the third migration rate, they are almost entirely shared.
-
-%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
+\section{Conclusion \& Discussion}
 
@@
 
-\paragraph{The adaptive filter?}
-RAMBLING THOUGHTS HERE
-Since larger effect alleles are more likely to establish,
-be it by mutation or migration,
-repeated colonization of new freshwater habitats will select for larger alleles,
-be it single alleles or haplotypes bound together by an inversion.
-However, these are more strongly selected against in the interstitial time.
-Being recessive would help with this,
-but would also make it more difficult to establish.
-
+\clearpage
+%\plr{some problems because you had bibliographystyle before bibliography}
 \bibliographystyle{plainnat}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-\appendix
 \section{Supp.}
@@
 	\begin{center}
-  		\includegraphics[width=\linewidth]{plotlyPlots/StandingGeneticVariation.png}
-  		\caption{(PLACEHOLDER)}
-		\label{fig:SGV}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=\linewidth]{matplotlibPlots/MFAI.png}
+  		\includegraphics[width=\linewidth]{matplotlibPlots/MFAI_2_matplot.png}
   		\caption{Distributions of mean number of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
@@
 
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=\linewidth]{matplotlibPlots/effectRegionCounts.png}
-  		\caption{(PLACEHOLDER)}
-		\label{fig:counts}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-5.png}
-  		\caption{ (PLACEHOLDER)
-		}
-  		\label{fig:Fst1}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-2.png}
-  		\caption{(PLACEHOLDER)
-		}
-  		\label{fig:Fst4}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-5.png}
-  		\caption{ (PLACEHOLDER)
-		}
-  		\label{fig:phenotype_ts1}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-2.png}
-  		\caption{(PLACEHOLDER)
-		}
-  		\label{fig:phenotype_ts4}
-	\end{center}
-\end{figure}
-
 \end{document}

### 2018-08-17 Stickleback_Paper.tex

--- 2018-07-26 Stickleback_Paper.tex
@@
 
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MeanPhenotype1.png}
-  		\caption{
-            Distribution of mean trait values throughout each simulation.
-            The dashed pink and purple lines show the optima ($\pm 10$).
-		}
-  		\label{fig:MeanPhenotype}
-	\end{center}
-\end{figure}
-\plr{Is this the whole simulation? what about the initial adaptation phase?
-     Is the horizontal line at the median? what are the whiskers at?}
-
 As expected, increasing migration rate decreased differentiation between habitats.
@@
 	\begin{center}
-        \includegraphics[width=0.8\linewidth]{matplotlibPlots/TimeToAdaptation}
-        \caption{
-            Time to adaptation as a function of migration rate ($m$),
-            measured as the number of generations
-            until the introduced population's mean phenotype is within 0.5 
-            of the original lakes average phenotype. 
-            Each point represents a single simulation, and
-            the line shows the average of all points at each value of $m$.
-        }
+  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/Adaptation_2_matplot.png}
+  		\caption{Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
+		it takes for the introduced population's mean phenotype to come within 0.5 of the original lakes average phenotype. 
+		Each point represents a simulation run at some value of $M$. 
+		The yellow dashed line is the average of all points at each respective parameter value}
   		\label{fig:TimeToAdaptation}
@@
 	\begin{center}
-        \includegraphics[width=\linewidth]{plotlyPlots/NumFAA}
-        \includegraphics[width=\linewidth]{matplotlibPlots/MPFAI}
-  		\caption{
-            \textbf{(A)}
-            Distribution of the number of freshwater adapted alleles (green)
-            and the average number of lakes each allele is present in (red)
-            across the course of the simulation.
-            \textbf{(B)}
-            Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
+  		\includegraphics[width=\linewidth]{matplotlibPlots/MPFAI_2_matplot.png}
+  		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
 		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
@@
 %- Phenotype Distribution
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/Pheno_2_matplot.png}
+  		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
+		The dashed pink line (Pheno = +10) is the optimum phenotype any individual in the marine environment.
+		In contrast the purple line at (Pheno = -10) represents the optimum for any individual in the freshwater environment. 
+		All individuals at generation 0 (beginning of the simulation) 
+		}
+  		\label{fig:MeanPhenotype}
+	\end{center}
+\end{figure}
+
 %GENERAL
@@
 \subsection*{Realized genomic architecture}
-
-\begin{figure}
-	\begin{center}
-        \includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3}
-        \includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-4}
-  		\caption{(Description and cleaning needed)
-		}
-  		\label{fig:Fst3}
-	\end{center}
-\end{figure}
 
@@
 	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{matplotlibPlots/Power_FP.png}
+  		\includegraphics[width=0.7\linewidth]{matplotlibPlots/0_25_power.png}
   		\caption{ 

### 2018-08-17 matplotScripts/Fst_Dist_matplot.py
new files

@@ -1,0 +1,53 @@
+import plotly
+from plotly.graph_objs import Scatter, Layout
+import plotly.graph_objs as go
+import plotly.plotly as py
+import sys
+
+num = sys.argv[1]
+recipes = ['8_'+num+'_0','8_'+num+'_1','8_'+num+'_2','8_'+num+'_3']
+
+def readHeader(firstLine):
+
+	numberOfSamplingPoints,samplingInterval,introduction,endofSim = (int(i) for i in firstLine.split())
+	samplesUntilIntroduction = introduction/samplingInterval - 1
+	return numberOfSamplingPoints,samplingInterval,introduction
+
+Original_Intro = []
+Marine_Original = []
+Marine_Intro = []
+
+data = [[][][]]
+
+	
+for i in range(4):
+
+	file1 = "../../Output1/MyRecipe"+recipes[i]+"/FreshwaterFreshwater2Fst.txt"
+	#,"Freshwater","IntroducedFreshwater",'rgb(60,179,113)'
+	file2 = "../../Output1/MyRecipe"+recipes[i]+"/OceanFreshwaterFst.txt"
+	#,"Ocean","Freshwater",'rgb(127,255,212)'
+	file3 = "../../Output1/MyRecipe"+recipes[i]+"/OceanFreshwater2Fst.txt"
+	#,"Ocean","IntroducedFreshwater",'rgb(138,43,226)'
+
+	
+
+	for k,f in enumerate([file1,file2,file3]):
+
+		File = open(f,"r")
+		first = File.readline()
+
+		#read in respective Fst/Pos pairs - they are in order.
+		Fst = list(map(float,File.readline().split()))
+		Pos = list(map(int,File.readline().split()))
+		
+		#sort both lists by Fst value
+		Pos_Sorted = [fst for _,fst in sorted(zip(Fst,Pos))]
+		Fst_Sorted = sorted(Fst)
+
+		data[k].append()
+		
+	
+
+
+
+

### 2018-08-17 matplotScripts/boxplot_MFAI_matplot.py

@@ -1,0 +1,74 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import sys
+
+num = sys.argv[1]
+
+Traces = []
+
+recipes = ['8_'+num+'_0','8_'+num+'_1','8_'+num+'_2','8_'+num+'_3']
+mig_rates = ['mig = 5e-5 (1/10 Ind/Gen)', 'mig = .0005 (1 Ind/Gen)','mig = .005 (10 Ind/Gen)','mig = .05 (100 Ind/Gen)']
+
+xAxis = []
+AvgFreshAllelesPerMarineInd = []
+AvgFreshAllelesPerOrigInd = []
+AvgFreshAllelesPerIntroInd = []	
+
+for i in range(4):
+
+	M = []
+	O = []
+	I = []
+
+	File = open("../../Output1/MyRecipe"+recipes[i]+"/AvgFWAA.txt","r")
+	first = File.readline()
+	count = 0
+
+	for line in File:
+		n1,n2,n3 = (float(s) for s in line.split())
+		M.append(n1)
+		O.append(n2)
+		I.append(n3)
+
+	AvgFreshAllelesPerMarineInd.append(M)
+	AvgFreshAllelesPerOrigInd.append(O)
+	AvgFreshAllelesPerIntroInd.append(I)
+
+data_a = AvgFreshAllelesPerMarineInd
+data_b = AvgFreshAllelesPerOrigInd
+data_c = AvgFreshAllelesPerIntroInd
+
+ticks = ['5e-5','5e-4','5e-3','5e-2']
+
+def set_box_color(bp, color):
+    plt.setp(bp['boxes'], color=color)
+    plt.setp(bp['whiskers'], color=color)
+    plt.setp(bp['caps'], color=color)
+    plt.setp(bp['medians'], color=color)
+
+#plt.figure()
+fig1,ax1 = plt.subplots()
+bpl = ax1.boxplot(data_a, positions=np.array(range(len(data_a)))*3.0-0.6, sym='', widths=0.4)
+bpm = ax1.boxplot(data_b, positions=np.array(range(len(data_b)))*3.0, sym='', widths=0.4)
+bpr = ax1.boxplot(data_c, positions=np.array(range(len(data_c)))*3.0+0.6, sym='', widths=0.4)
+set_box_color(bpl, '#D7191C') # colors are from http://colorbrewer2.org/
+set_box_color(bpm, '#2C7BB6')
+set_box_color(bpr, '#5D7FF6')
+
[…]

### 2018-08-17 matplotScripts/boxplot_pheno_matplot.py

@@ -1,0 +1,108 @@
+import sys
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+num = sys.argv[1]
+title = sys.argv[2]
+
+Traces = []
+
+recipes = ['8_'+num+'_0','8_'+num+'_1','8_'+num+'_2','8_'+num+'_3']
+mig_rates = ['mig = 5e-5 (1/10 Ind/Gen)', 'mig = .0005 (1 Ind/Gen)','mig = .005 (10 Ind/Gen)','mig = .05 (100 Ind/Gen)']
+
+xAxis = []
+xAxis2 = []
+col1 = []
+col2 = []
+col3 = []
+
+numSamplingPoints = 0
+interval = 0 
+intro = 0 
+end = 0
+newLakesIntroduced = 0 
+
+for i in range(4):
+
+	c1 = []
+	c2 = []
+	c3 = []
+
+	File = open("../../Output1/MyRecipe"+recipes[i]+"/AveragePhenotypeThroughout.txt","r")
+	first = File.readline()
+	numSamplingPoints,interval,intro,end = (int(s) for s in first.split()) 
+
+	newLakesIntroduced = ( intro / interval ) -1
+
+	count = 0
+	for line in File:
+		if(count < newLakesIntroduced):
+			#n1, n2= (float(s) for s in line.split())
+			li = line.split()
+			c1.append(float(li[0]))
+			c2.append(float(li[1]))
+		else:
+			li = line.split()
+			c1.append(float(li[0]))
+			c2.append(float(li[1]))
+			c3.append(float(li[12]))
+
+		count += 1;		
+
+	col1.append(c1)
+	col2.append(c2)
+	col3.append(c3)
+
+
+#xAxis = xAxis[200:1600] + xAxis[1800:3200] + xAxis[3400:4800] + xAxis[5000:]
+#xAxis2 = xAxis2[100:800] + xAxis2[900:1600] + xAxis2[1700:2400] + xAxis2[2500:]
+#col1 = col1[200:1600] + col1[1800:3200] + col1[3400:4800] + col1[5000:]
+#col2 = col2[200:1600] + col2[1800:3200] + col2[3400:4800] + col2[5000:]
+#col3 = col3[100:800] + col3[900:1600] + col3[1700:2400] + col3[2500:]
+
+data_a = col1
+data_b = col2
+data_c = col3
+
+ticks = ['5e-5','5e-4','5e-3','5e-2']
+
+def set_box_color(bp, color):
[…]

### 2018-08-21 Stickleback_Paper.tex
semi final plots and edits

@@
 but remain simplistic in some aspects due to computational constraints.
+\jgg{should we brag about treeSeq here a little?}
 Possibly the most important caveat is that simulated population sizes are much smaller
@@
 	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{GeographyDiagram}
+  		\includegraphics[width=0.6\linewidth]{GeographyFigure}
   		\caption{A representation of the geographic and evolutionary history of all populations throughout the simulation. 
@@
 	\begin{center}
-        \includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4}
-        \includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3}
+        \includegraphics[width=1.0\linewidth]{semi_final_plots/Pheno_semi_final.png}
   		\caption{ 
@@
 	\begin{center}
-  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/Fst_Adaptation.png}
+  		\includegraphics[width=1.0\linewidth]{semi_final_plots/Fst_semi_final.png}
   		\caption{Distributions of $F_{st}$  values between the three subpopulations throughout the simulations.
@@
 	\begin{center}
-  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/Adaptation_2_matplot.png}
+  		\includegraphics[width=1.0\linewidth]{semi_final_plots/T_adapt_semi_final.png}
   		\caption{Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
@@
 	\begin{center}
-  		\includegraphics[width=\linewidth]{matplotlibPlots/MPFAI_2_matplot.png}
+  		\includegraphics[width=\linewidth]{semi_final_plots/FAA_semi_final.png}
   		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
@@
 	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/Pheno_2_matplot.png}
+  		\includegraphics[width=1.0\linewidth]{semi_final_plots/PhenoDist_semi_final.png}
   		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
@@
 
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=1.0\linewidth]{semi_final_plots/Haplo_semi_final.png}
+  		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
+		The dashed pink line (Pheno = +10) is the optimum phenotype any individual in the marine environment.
+		In contrast the purple line at (Pheno = -10) represents the optimum for any individual in the freshwater environment. 
+		All individuals at generation 0 (beginning of the simulation) 
+		}
+  		\label{fig:MeanPhenotype}
+	\end{center}
+\end{figure}
+
 Now we take a closer look at the genomic architecture of local adaptation 
@@
 	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{matplotlibPlots/0_25_power.png}
+  		\includegraphics[width=0.7\linewidth]{semi_final_plots/TP_SP_semi_final.png}
   		\caption{ 
@@
 
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=\linewidth]{matplotlibPlots/MFAI_2_matplot.png}
-  		\caption{Distributions of mean number of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
-		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
-		freshwater adapted alleles.
-		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
-		}
-  		\label{fig:MNFAI}
-	\end{center}
-\end{figure}
-
 \end{document}

### 2018-08-21 Stickleback_Paper.tex
latest changes

@@
 Our model has two habitat types, which are considered separate populations
-with a fixed total size of 2,000 diploid individuals each. The marine population is treated as a single large deme, but the freshwater population is subdivided into local subpopulations or demes.
+with a fixed total size of 5,000 diploid individuals each. The marine population is treated as a single large deme, but the freshwater population is subdivided into local subpopulations or demes.
 The scheme, depicted in Figure \ref{fig:Geo},
@@
 along a stretch of coastline.
-The ``marine'' habitat  is a continuous, one-dimensional range of 10 units of length,
+The ``marine'' habitat  is a continuous, one-dimensional range of 25 units of length,
+%%%%
 and the ``freshwater'' habitat is divided into ten subpopulations
@@
 	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{GeographyFigure}
+  		\includegraphics[width=0.6\linewidth]{GeographyFigure.pdf}
   		\caption{A representation of the geographic and evolutionary history of all populations throughout the simulation. 

### 2018-08-21 matplotScripts/boxplot_pheno_matplot.py

@@ -1,108 +1,0 @@
-import sys
-import matplotlib.pyplot as plt
-import numpy as np
-
-
-num = sys.argv[1]
-title = sys.argv[2]
-
-Traces = []
-
-recipes = ['8_'+num+'_0','8_'+num+'_1','8_'+num+'_2','8_'+num+'_3']
-mig_rates = ['mig = 5e-5 (1/10 Ind/Gen)', 'mig = .0005 (1 Ind/Gen)','mig = .005 (10 Ind/Gen)','mig = .05 (100 Ind/Gen)']
-
-xAxis = []
-xAxis2 = []
-col1 = []
-col2 = []
-col3 = []
-
-numSamplingPoints = 0
-interval = 0 
-intro = 0 
-end = 0
-newLakesIntroduced = 0 
-
-for i in range(4):
-
-	c1 = []
-	c2 = []
-	c3 = []
-
-	File = open("../../Output1/MyRecipe"+recipes[i]+"/AveragePhenotypeThroughout.txt","r")
-	first = File.readline()
-	numSamplingPoints,interval,intro,end = (int(s) for s in first.split()) 
-
-	newLakesIntroduced = ( intro / interval ) -1
-
-	count = 0
-	for line in File:
-		if(count < newLakesIntroduced):
-			#n1, n2= (float(s) for s in line.split())
-			li = line.split()
-			c1.append(float(li[0]))
-			c2.append(float(li[1]))
-		else:
-			li = line.split()
-			c1.append(float(li[0]))
-			c2.append(float(li[1]))
-			c3.append(float(li[12]))
-
-		count += 1;		
-
-	col1.append(c1)
-	col2.append(c2)
-	col3.append(c3)
-
-
-#xAxis = xAxis[200:1600] + xAxis[1800:3200] + xAxis[3400:4800] + xAxis[5000:]
-#xAxis2 = xAxis2[100:800] + xAxis2[900:1600] + xAxis2[1700:2400] + xAxis2[2500:]
-#col1 = col1[200:1600] + col1[1800:3200] + col1[3400:4800] + col1[5000:]
-#col2 = col2[200:1600] + col2[1800:3200] + col2[3400:4800] + col2[5000:]
-#col3 = col3[100:800] + col3[900:1600] + col3[1700:2400] + col3[2500:]
-
-data_a = col1
-data_b = col2
-data_c = col3
-
-ticks = ['5e-5','5e-4','5e-3','5e-2']
-
-def set_box_color(bp, color):
[…]

### 2018-08-22 Stickleback_Paper.tex
Jared's edits

@@
 verify that adaptation of introduced pops is due to initial diversity not subsequent migration
+\jgg{one way to do this is check the mean number of samples under the introduced populations?}
 
@@
 The ``marine'' habitat  is a continuous, one-dimensional range of 25 units of length,
-%%%%
 and the ``freshwater'' habitat is divided into ten subpopulations
 (which we call ``lakes''), each connected to the marine habitat
-at positions $i - 1/2$ for $1 \le i \le 10$.
+at positions $i - 1/2$ for $1 \le i \le 25$.
 
-Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff such that alternative values of the trait have high or low fitness in tht two different habitats.
+Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff such that alternative values of the trait have high or low fitness in the two different habitats.
 This situation roughly models the cumulative effect of the various phenotypes
@@
   		\includegraphics[width=0.6\linewidth]{GeographyFigure.pdf}
-  		\caption{A representation of the geographic and evolutionary history of all populations throughout the simulation. 
-		The marine is a one-dimensional, continuous population with spatial positions ranging from [0.0 - 10.0]. 
+  		\caption{
+			A representation of the geographic and evolutionary history of all populations throughout the simulation. 
+		The marine is a one-dimensional, continuous population with spatial positions ranging from [0.0 - 25.0]. 
 		Each freshwater lake$_{i}$ for both the introduced and freshwater populations
 		is a discrete population connected by migration only through the marine, at position $i - 0.5$. 
-		The introduced lakes have the same spatial location (but separate?) and selective pressure as the original lakes, but arise at 40K generations.
-		The marine has $2000$ diploid individuals while each of the lakes fluctuate around $200$. 
+		Each of the introduced lakes have the same spatial dynamic and selective pressure as the pre-existing lakes, but arise at 100K generations into the simulation.
+		The marine has $5000$ diploid individuals while each of the lakes fluctuate  $200$ diploid individuals. 
 		the introduced population is initialized as a copy of all marine individuals to model marine 
@@
 		with the new selective pressure across a range of parameter values. 
-		%Migration between the marine and original lakes is 
-		%set to $0$ the generation before introduction to avoid direct migrants from original $->$ introduced lakes.
 			}
@@
 \paragraph{Potential genomic architecture}\bill{Initial genomic architecture? The header is a little confusing to me}
-Each individual carries two copies of a linear chromosome of $10^5$ loci
-(which could be thought of roughly as a kilobase each).
-New mutations occur with probability $10^{-7}$ per locus per generation.
-Ten ``effect regions'' of 100 loci each
-are spread evenly along the chromosome (separated by 4950 loci)
+Each individual carries two copies of a linear chromosome of $10^8$ loci representing a realistic size genome.
+New mutations occur with probability $10^{-10}$ per locus per generation in regions affecting phenotype and $10^{-8}$ in neutral regions.
+Ten ``effect regions'' of $100,000$ loci each
+are spread evenly along the chromosome (separated by $495,0000$ loci)
 in which mutations affect the trait;
@@
 or completely dominant (with equal probability),
-and with an effect sizes chosen randomly from an Exponential distribution with mean $1/2$,
+and with effect sizes chosen randomly from an Exponential distribution with mean $1/2$,
 either positive or negative with equal probability.
-Individual trait value are determined additively from the diploid genotypes.
+Individual trait values are determined additively from the diploid genotypes.
 Concretely, an individual that is heterozygous and homozygous for mutations
@@
 with an average of one crossover per chromosome per generation
-($10^{-7}$ per locus per generation).
+($10^{-8}$ per locus per generation).
 The model is Wright-Fisher, so that each generation, in each habitat,
-2000 new individuals are produced,
+5000 new individuals are produced,
 and so to keep population sizes roughly constant within each lake,
@@
 of the distance separating the two, with stanadard deviation $1/2$. 
-Concretely, if the first parent is marine individual $i$,
+More specifically, if the first parent is marine individual $i$,
 then marine individual $j$ is chosen as the mate
@@
 except the probability that the parents are marine individuals is $m$;
-any new freshwater offpsring produced by marine individuals
+any new freshwater offspring produced by marine individuals
 are assigned to the lake nearest to the position of the first marine parent.
@@
 and act like an independent copy of the original set of lakes -- in particular,
-the two sets of lakes each have 2,000 individuals at all times.
+the two sets of lakes each have 5,000 individuals at all times.
 Since this doubles the number of lake-to-marine immigrants,
@@
 % Overview
-We varied the migration rate, $m$,
-from $5 \times 10^{-5}$ to $5 \times 10^{-2}$ per individual,
-i.e., between 0.01 and 10 migrants per year to and from each of the ten lakes.
+We varied the migration rate, $m$, across separate simulations
+from $5 \times 10^{-5}$ to $5 \times 10^{-1}$ per individual,
+i.e., between 0.01 and 100 migrants per year to and from each of the freshwater lakes.
 This had a strong effect on many aspects of adaptation,
@@
 and the population genetic signals left behind.
-At the lowest migration rate, lakes adpated nearly independently,
+At the lowest migration rate, lakes adapted nearly independently,
 while at the highest migration rate, the habitats are beginning to act like a single metapopulation
@@
 until the trait means were close to the optimal values in each habitat. 
-(Remaining simulations are shown in Figures \ref{fig:phenotype_ts1} and \ref{fig:phenotype_ts4}.)
+%(Remaining simulations are shown in Figures \ref{fig:phenotype_ts1} and \ref{fig:phenotype_ts4}.)
+%Create these once all plots are finalized
 Greater migration decreased adaptedness only weakly
-except for at the highest migration rate (10 migrants per generation per lake),
-where the mean trait value equilibrated at roughly half its optimal value 
+except for at the two highest migration rates (10 and 100 migrants per generation per lake),
+where the mean trait value equilibrated at roughly half its optimal value and experienced panmixia, respectively.
 (Figures \ref{fig:MeanPhenotype} and \ref{fig:phenotype_ts4}).
@@
 generations,
+\jgg{I'm not sure what this means? how do we know how long the jumps take?}
 and move the trait by an amount of order 1.
@@
   		\caption{ 
-        Mean individual trait values in the marine habitat (black line),
-        the original lakes (light green lines; average in pink),
-        and the introduced lakes (dark green lines; average in orange)
+        Mean individual trait values in the marine habitat (blue line),
+        the original lakes (light green lines; average in dark green),
+        and the introduced lakes (yellow lines; average in orange)
         across the course of two simulations, with migration rates of
-        \textbf{(top)} $m=5 \times 10^{-4}$ and
-        \textbf{(bottom)} $m=5 \times 10^{-3}$ per generation individual
-        (i.e., 0.1 and 1.0 migrants per generation per lake, respectively).
+        \textbf{(top)} $m=5 \times 10^{-5}$ and
+        \textbf{(bottom)} $m=5 \times 10^{-4}$ per generation individual
+        (i.e., 0.01 and 0.1 migrants per lake per generation, respectively).
         Optimal trait values in the two habitats are at $\pm 10$.
@@
 \end{figure}
-
-\jgg{also, interesting point to make on second plot, you can see the initial lakes taking the same size step, showing sharing of alleles.
-could zoom in on the relevent portions of the figure for this. (begging and introduction)
- a 2 by 2 with the important sections (0 - 20K) or (40-50K) for both above figures?}
 
@@
   		\includegraphics[width=1.0\linewidth]{semi_final_plots/Fst_semi_final.png}
-  		\caption{Distributions of $F_{st}$  values between the three subpopulations throughout the simulations.
-		On the left, $F_{st}$ was calculated only on neutral mutation frequencies. On the right, $F_{st}$ was calculated
-		only on effect (Impact on phenotype) mutation frequencies. 
-		Effect mutations are acted upon by selection as they have an affect on fitness where neutral mutations  }
+  		\caption{
+		Finish this when we finalize plots!
+		 }
   		\label{fig:Fst}
@@
 As shown in Figure \ref{fig:TimeToAdaptation}, at the lowest migration rate 
-it took over 20 thousand generations for the mean trait value across introduced lakes to 
+it took over $20,000$ generations for the mean trait value across introduced lakes to 
 get to within 0.5 of the original lakes value.
@@
 of trait values against time (e.g., Figure \ref{fig:phenotype_ts2}).
-However, higher migration rates allowed the freshwater habitat to adapt
+However, higher migration rates allowed the freshwater habitats to adapt
 nearly as quickly as the marine habitat.
@@
 
-The tenfold difference in speed of adaptation occurs because
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=\linewidth]{semi_final_plots/FAA_semi_final.png}
+  		\caption{
+			finish this after plots are finalized
+		}
+		\label{fig:MPFAI}
+	\end{center}
+\end{figure}
+
+This dramatic increase in speed of local adaptation occurs because
 at low migration rates, adaptation occurs independently in each lake,
@@
 instead of developing their own genetic basis of adaptation.
-As a first indication of this, Figure \ref{fig:Fst} shows that
-$F_{ST}$ between the ``original'' and ``introduced'' sets of lakes 
-at effect mutations decreased with migration rate.
-
-To further investigate sharing of locally adaptive alleles
-and the transporter hypothesis,
-we investigate the ``pre-existing freshwater adapted alleles'',
-defined for a particular generation
-to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat
-in that generation.
-Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles,
-across generations,
-as well as the average number of lakes each is found in.
-As migration rates increase, the number of these alleles decreases steadily,
-and each is concurrently found in a greater number of lakes.
+%As a first indication of this, Figure \ref{fig:Fst} shows that
+%$F_{ST}$ between the ``original'' and ``pre-existing'' sets of lakes 
+%at effect mutations decreased with migration rate.
+To further investigate sharing of locally adaptive alleles and the transporter hypothesis,
+we investigate the ``pre-existing freshwater adapted alleles'', defined for a particular generation
+to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat in that generation.
 
@@
 At the lowest migration rate, 
-each genome in the original lakes have almost exactly $1/10^{th}$ 
+each genome in the original lakes have almost exactly $1/25^{th} = 0.04\%$   
 of the total freshwater adapted alleles --
-since there are 10 lakes, this implies that each lake has adapted with a unique set of alleles.
+since there are 25 lakes, this implies that each lake has adapted with a unique set of alleles.
 Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
 
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=\linewidth]{semi_final_plots/FAA_semi_final.png}
-  		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
-		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
-		freshwater adapted alleles.
-		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
-		while the percentage tells us the variance of the haplotype}
-		\label{fig:MPFAI}
-	\end{center}
-\end{figure}
+In Figure \ref{fig:MPFAI}C we can see that; as migration rates increase, the number of these alleles decreases steadily.
+On the left in Figure \ref{fig:MPFAI}B, we see the mean percentage of alleles per individual rapidly increases. 
+Together, these two results strongly suggest the sharing genetic basis.
+\jgg{TODO, reference the number of lakes that each allele appears in, make plot for supp}
+%As a first indication of this, Figure \ref{fig:MPFAI}B shows that the introduced 
+%population utilizes a higher percentage of "pre-existing freshwater adapted alleles"
+%with higher rates of introgression between the populations.
+Figure \ref{fig:MPFAI}A shows the distribution of the mean number of pre-existing freshwater adapted alleles,
+Throughout the simulations, in each deme.
+Interestingly, the genetic basis of the freshwater phenotype 
+seems to simplify with greater $m$. 
+This would seem to suggest higher rates of migration allowed 
+adaptive alleles of higher effect to travel more efficiently through the populations
+\jgg{is this right?}
 
@@
 At first, increased migration allows sharing of adaptive alleles between lakes,
-but at the highest migration rate,
+but at the second from highest migration rate,
 the constant influx of alleles between the habitats creates substantial migration load.
@@
 as a consequence of a large number of offspring through hybridization events between subpopulations.
+at $m=0.5$, all populations experience panmixia
 
 \paragraph{Standing genetic variation}
+\jgg{todo, make this figure for the supp}
+\jgg{this will need updated in I make the plot for it}
 Despite a dramatic difference in the amount of allelic sharing between lakes,
-standing genetic variance (Figure \ref{fig:SGV}) was around 0.05 in the freshwater habitat
+standing genetic variance (Figure \ref{fig:SGV}) 
+was around 0.05 in the freshwater habitat
 across all but the highest migration rates.
@@
   		\includegraphics[width=1.0\linewidth]{semi_final_plots/Haplo_semi_final.png}
-  		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
-		The dashed pink line (Pheno = +10) is the optimum phenotype any individual in the marine environment.
-		In contrast the purple line at (Pheno = -10) represents the optimum for any individual in the freshwater environment. 
-		All individuals at generation 0 (beginning of the simulation) 
+  		\caption{
+			Finish when plot is finalized. 
 		}
@@
 to be a causal locus.
-\plr{Labels are wrong on that figure: false pos rate should go down with cutoff.
-     Is it true positive rate?}
 
@@
 
-We also have also explored the genomic architecture as a consequence of the nature of selection in our scenario. 
-The alleles underlying individual trait value were of large effect and a low number. 
-With a total of $10^{5}$ loci, to be realistic, each loci should represent 1000 $Kb$ in real data.
-\jgg{This should be expanded upon, not sure what this means in terms of the genomic architecture}
-
 We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 

### 2018-08-22 Stickleback_Paper.tex
little edits

@@
 Figure \ref{fig:MPFAI}B shows
-the distribution of the mean percentage of currently-defined freshwater adapted alleles 
+the percentage of currently-defined, freshwater-adapted alleles 
 that each genome in each of the populations carries,
@@
 In Figure \ref{fig:MPFAI}C we can see that; as migration rates increase, the number of these alleles decreases steadily.
-On the left in Figure \ref{fig:MPFAI}B, we see the mean percentage of alleles per individual rapidly increases. 
+On the left in Figure \ref{fig:MPFAI}B, the mean percentage of alleles per individual rapidly increases. 
 Together, these two results strongly suggest the sharing genetic basis.

### 2018-09-07 Stickleback_Paper.tex
origin of adaptive variants

@@
 pre-existing freshwater adapted alleles | standing genetic variants = ?
-
-\textbf{Pending things to do?} that maybe we do, maybe not
-
-longer genomes ($10^8$ not $10^5$ but rescaling everything accordingly)
-
-verify that adaptation of introduced pops is due to initial diversity not subsequent migration
-\jgg{one way to do this is check the mean number of samples under the introduced populations?}
 
@@
 
-%%%%%%%%%%%%%%%%%
 \section{Results}
 
-% Overview
 We varied the migration rate, $m$, across separate simulations
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Local Adaptation: differentiation with gene flow}
@@
 %\subsubsection*{Low Recombination Causes Clustering}
+\subsubsection*{Origin of introduced freshwater adaptive alleles}
+Up to this point, we have quantified rapid and parallel adaptation and found strong evidence that sharing of alleles is (more than likely?) responsible.
+However at $m$ equal to $0.1$ \emph{mig/gen/lake}, in \ref{fig:MPFAI} we see sharing of alleles with between populations,
+yet \emph{Time to adaptation} is still fairly slow ($\approx 2,000$ gens).
+This is surprising when considering we see a similar quantity of allele sharing at  $m$ equal to $1.0$ \emph{mig/gen/lake}, 
+but much more rapid adaptation ($\approx 60$ gens). 
+This $30X$ speedup in \emph{time to adaptation} seems to be the result of either subsequent migration
+from the marine after the point of freshwater introduction, or, the higher saturation of pre-existing freshwater adapted alleles 
+in the marine at \emph{time to introduction} allowed for selection to effectively propagate the pre-existing freshwater adapted alleles.
+To explore this further, we have looked at the genealogical origins using SLiM's \emph{treeSeq} (CITE thesis, treeSeq paper)
+of the adaptive variants in the introduced population (p3) at \emph{time of introduction}, when the introduced population's mean phenotype
+came within 0.5 of the pre-existing freshwater population's mean phenotype.
+The adaptive variants in the introduced population were defined as effect mutations at higher than 0.5 frequency in the 
+introduced population, and below 0.5 frequency in the marine population.
+We classified the origin of each adaptive variant, (freshwater adapted allele in the introduced population) to be:
+(1) pre-existing, propagating from an ancestor that was in the initial generation introduced to the new freshwater selective pressure,
+(2) subsequent migration, propagating from and ancestor who was a migrant from the ocean \emph{after} introduction,
+We then averaged these 3 numbers, for each variant existing in the population.
+In Figure (blah), (FINISH AFTER PLOT)
 

### 2018-09-12 Stickleback_Paper.tex
latest changes

--- 2018-08-22 Stickleback_Paper.tex
@@
 
-\textbf{Pending things to do?} that maybe we do, maybe not
-
-longer genomes ($10^8$ not $10^5$ but rescaling everything accordingly)
-
-verify that adaptation of introduced pops is due to initial diversity not subsequent migration
-\jgg{one way to do this is check the mean number of samples under the introduced populations?}
-
 \section{Abstract}
@@
 and large migration load at high gene flow.
-This is mediated mainly by the total downstream influx of alleles \emph{(maybe??)}.
+This is mediated mainly by the total downstream influx of alleles \emph{(maybe??)(No)}.
 We do not see evidence for strong effects encouraging genomic clustering of causal alleles
@@
 and then a mate is chosen, 
-also proportionally to fitness but reweighted by a Gaussian function 
-of the distance separating the two, with stanadard deviation $1/2$. 
+also proportionally to fitness but reweighed by a Gaussian function 
+of the distance separating the two, with standard deviation $1/2$. 
 More specifically, if the first parent is marine individual $i$,
@@
 is $2m$ instead of $m$.
+
+\paragraph{Using SLiM's TreeSeq}
+This case study involves analyzing large simulations which model populations under the influence of selection. 
+To fully understand the origins of adaptive variants, calculate statistics, and avoid the computationally expensive task of 
+simulating neutral mutations, we have used SLiM's \emph{treeSeq} feature to track the genealogical history of all samples throughout the simulation. 
+This feature outputs a compact representation the genomic outcome of every meiosis relevant to tracked samples (Often thought of as the ARG).
+\jgg{maybe the below should be in descriptive statistics}
+To understand recomposition of the freshwater haplotype in the new lakes through selection, 
+we were able to classify each adaptive variant, in each genome in the population at the generation at time of adaptation,
+as deriving from one of four origins:
+(1) Migrants, brought in from a migrant after the introduction of the new lakes 
+(2) Pre-existing, passed down from an individual in the original generation of the new lakes, 
+however, not defined to be an adaptive variant in the original lakes
+(3) Captured, Pre-existing, passed down from an individual in the original generation of the new lakes, 
+and defined to be adaptive in the original lakes
+(4) De novo, mutated and propagated from an individual existing in the introduced lakes population, after introduction.
+
+We also used the tree sequence to count the "fitness" of individuals in the initial generation of the new lakes. 
+The "fitness" of an individual in this case, is defined to be the total number off eventual offspring, in the generation at the time of adaptation, 
+which have inherited a adaptive variant from that individual. 
+To calculate this, we only considered gene trees which overlapped an adaptive variant site (defined at time of adaptation).
+We then counted the total number of offspring which had inherited from any adaptive variant site, 
+for each individual in the initial generation.
 
@@
 We will now dissect what happens between these extremes.
-
-\emph{merge with the above}
-Interestingly, the ability for separate populations to locally adapt to their own selective pressure was relatively unaffected 
-until the highest rate of migration between the marine and freshwater environments.
+Interestingly, all rates of $m$ aside from the two highest had little affect on populations ability to locally adapt.
 All rates of historical introgression aside from the lowest, helped both the initial and introduced freshwater populations 
 share pre-existing freshwater adapted alleles.
-The sharing of alleles resulted in rapid adaptation ($\approx 100$ generations) of the introduced population (split from the marine population) to adapt to the freshwater environment.
+The sharing of alleles resulted in rapid adaptation ($\approx 60$ generations) of the introduced population (split from the marine population) to adapt to the freshwater environment.
 We also found that larger amounts of migration allow for more statistical power and less false positives in the resulting population genetics data ($F_{st}$ per SNP across the genome).
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Local Adaptation: differentiation with gene flow}
 
-% The progression of adaptation. How you can see introduction new alleles
-
+\jgg{This needs some work}
 Local adaptation occurred in all simulations:
@@
 until the trait means were close to the optimal values in each habitat. 
-%(Remaining simulations are shown in Figures \ref{fig:phenotype_ts1} and \ref{fig:phenotype_ts4}.)
-%Create these once all plots are finalized
-Greater migration decreased adaptedness only weakly
-except for at the two highest migration rates (10 and 100 migrants per generation per lake),
-where the mean trait value equilibrated at roughly half its optimal value and experienced panmixia, respectively.
-(Figures \ref{fig:MeanPhenotype} and \ref{fig:phenotype_ts4}).
 The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
 as jumps in the mean trait value;
-in the two simulations these occur on a time scale of 100 (lower migration) to 1000 (higher migration)
-generations,
+in the two simulations these occur on a time scale of 100 (lower migration) to 1000 (higher migration) generations,
 \jgg{I'm not sure what this means? how do we know how long the jumps take?}
@@
   		\caption{
-		Finish this when we finalize plots!
+		$F_{st}$ as a function of genomic location between all combinations of the three subpopulations; 
+		the old lakes, the new lakes, and the marine.
+		$F_{st}$ was calculated based upon allele frequencies in genomic windows of 500 mb. 
+		The four columns of subplot represent separate simulations with increasing $m$. 
+		The three rows represent the combinations of the three separate populations.
+		Ten vertical dotted pink lines in each subplot represent regions which have the potential to affect phenotype
 		 }
@@
 Increasing migration rate strongly decreased the time until freshwater habitats could adapt,
-both in the initial and introduced sets of lakes.
+both in the old and new sets of lakes.
+The different rates of introduction of effect alleles is clearly seen in the traces
+of trait values against time (e.g., Figure \ref{fig:phenotype_ts2}).
 As shown in Figure \ref{fig:TimeToAdaptation}, at the lowest migration rate 
@@
 get to within 0.5 of the original lakes value.
-Although many lakes had adapted before this time,
-the different rates of introduction of effect alleles is clearly seen in the traces
-of trait values against time (e.g., Figure \ref{fig:phenotype_ts2}).
-However, higher migration rates allowed the freshwater habitats to adapt
-nearly as quickly as the marine habitat.
+At $m = 1.0$, we saw the introduced population adapt in just 32 generations. 
+This rate of adaptation follows what we've observed in natural populations of Threespine stickleback.
+At the highest migration rate, $m = 100$, all individuals were unable to locally adapt.
 
@@
   		\includegraphics[width=1.0\linewidth]{semi_final_plots/T_adapt_semi_final.png}
-  		\caption{Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
+  		\caption{
+		Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
 		it takes for the introduced population's mean phenotype to come within 0.5 of the original lakes average phenotype. 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Sharing of freshwater adapted alleles}
-
-%FIGURES THAT SHOW THIS:
-%-MPAA / IND
-%-Total/Avg shared
 
@@
   		\caption{
-			finish this after plots are finalized
+		Distribution of freshwater adapted alleles 
 		}
@@
 
-\paragraph{Standing genetic variation}
-\jgg{todo, make this figure for the supp}
-\jgg{this will need updated in I make the plot for it}
-Despite a dramatic difference in the amount of allelic sharing between lakes,
-standing genetic variance (Figure \ref{fig:SGV}) 
-was around 0.05 in the freshwater habitat
-across all but the highest migration rates.
-Concurrently,
-genetic variation in the marine population steady increased with migration
-to a similar value.
-On the one hand, it is not surprising that lakes, as a group,
-carry more genetic variation than the marine habitat,
-since population subdivision allows different alleles to become common in each.
-However, it is striking that at $m=.005$, the marine habitat carries as much genetic variation,
-despite a lack of any substantial migration load.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 With 10 effect regions across the genome, we can see in \ref{fig:Fst3} that only about 1 or 2 of the region contain peaks. 
-%SOFT STATEMENT - MORE (Peak) EVALUATION?
 When looking closely at the peaks we see that each is usually a composition of 7-8 SNPs close in proximity.
-%SOFT STATEMENT -
 Because some of those mutations are selectively neutral (we can see this because the counts of FAA are not as high as the number in the clusters),
@@
 
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{semi_final_plots/OriginRatio.png}
+  		\caption{ 
+		Ratio of Origins of adaptive alleles in all genomes at time of local adaptation for different values of $m$
+		}
+  		\label{fig:Origin}
+	\end{center}
+\end{figure}
+
 %The lowest migration rate of $10^{-5}$ also has a very low false positive rate across all $F_{st}$ threshold values.
 %\subsubsection*{Low Recombination Causes Clustering}
+\subsubsection*{Origin of introduced freshwater adaptive alleles: a conundrum}
+Up to this point, we have quantified rapid and parallel adaptation and found strong evidence that sharing of alleles is (more than likely?) responsible.
+However in Figure \ref{fig:MPFAI} at $m$ equal to $0.1$ \emph{mig/gen/lake}, we see sharing of alleles between populations,
+yet \emph{Time to adaptation} is still fairly slow ($\approx 2,000$ gens).
+This is surprising when considering we see a similar quantity of allele sharing at  $m$ equal to $1.0$ \emph{mig/gen/lake}, 
+but much more rapid adaptation ($\approx 60$ gens). 
+Initially, this 30X speedup of adaptation might seem to be the result of subsequent migration from the marine after the point of freshwater introduction. 
+This theory would assert that the continued influx of alleles is necessary for rapid local adaptation at the rate which we've observed in Nature.
+However, in Figure \ref{fig:Origin}, we find that at both $m = 0.1$ and $m = 1.0$, the vast majority of adaptive alleles origins were found to propagate 
+from individuals in the original generation ("Captured" OR "Pre-Existing").
+
+The alternative explanation is that at a higher migration rate, a higher saturation of marine adapted alleles are linked to freshwater adapted alleles.
+This linkage makes rebuilding the freshwater haplotype, in the new lakes, much more difficult as it much wait for recombination to break the links.
+With more historical introgression between the two selective pressures, the freshwater haplotype would remain more "intact" and much more efficient
+to rebuild in the lake.
+
+%Now lets make this more concrete
 

### 2018-09-12 Stickleback_Paper.tex
subplot captions

@@
 		$F_{st}$ was calculated based upon allele frequencies in genomic windows of 500 mb. 
-		The four columns of subplot represent separate simulations with increasing $m$. 
-		The three rows represent the combinations of the three separate populations.
+		The four columns of subplots are separate simulations with increasing $m$. 
+		The three rows are the combinations of the three separate populations.
 		Ten vertical dotted pink lines in each subplot represent regions which have the potential to affect phenotype
@@
   		\caption{
-		Distribution of freshwater adapted alleles 
+		Distribution of freshwater adapted alleles among the three populations. 
+		A freshwater adapted variant is defined as, at any generation throughout the simulation,
+		a variant affecting phenotype that is greater than 0.5 frequency in any of the old lakes, 
+		and less than 0.5 frequency in the marine population. 
+		Subplot $A$ is the distribution throughout the simulation, of the mean number of
+		freshwater adapted variants per individual, in each population.
+		Subplot $B$ is the distribution throughout the simulation, of the mean \emph{percentage} of
+		freshwater adapted variants per individual, in each population. 
+		Subplot $C$ is the distribution throughout the simulation, of the total number of
+		freshwater adapted alleles defined at any one generation.
+		In other words, $B  =  A  /  C$.
 		}

### 2018-09-13 Stickleback_Paper.tex
Captions and Cleanup

@@
 
-\paragraph{Using SLiM's TreeSeq}
+\paragraph{TreeSeq}
 This case study involves analyzing large simulations which model populations under the influence of selection. 
@@
 get to within 0.5 of the original lakes value.
-At $m = 1.0$, we saw the introduced population adapt in just 32 generations. 
+At $m = 1.0$, we saw the introduced population adapt in just $32$ generations. 
 This rate of adaptation follows what we've observed in natural populations of Threespine stickleback.
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% I'M HERE
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Migration load and genetic variation}
 
-%FIGURES THAT SHOW THIS:
-%- Phenotype Distribution
 \begin{figure}[h!tb]
 	\begin{center}
-  		\includegraphics[width=1.0\linewidth]{semi_final_plots/PhenoDist_semi_final.png}
+  		\includegraphics[width=0.6\linewidth]{semi_final_plots/PhenoDist_semi_final.png}
   		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
@@
 \end{figure}
-
-%GENERAL
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Realized genomic architecture}
@@
   		\caption{
-			Finish when plot is finalized. 
+			Phenotypes (\textbf{Top}) (effect size and mutation type) and haplotypes \textbf{(Botton)} as a function of spatial position of the individual 
+			possessing the genome. The phenotype of each haplotype is, by definition, the sum of the 
+			effect sizes of each mutation in the genome. 
 		}
@@
   		\caption{ 
-		%SP - suppose you have an Fst Peak, how likely is it that that region is causal. 
 		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
@@
   		\caption{ 
-		Ratio of Origins of adaptive alleles in all genomes at time of local adaptation for different values of $m$
+		Origin ratio of adaptive alleles in all genomes at time of local adaptation for different values of $m$
+		For all alleles at a frequency $> 0.5$ in any one of the new lakes individuals at time of adaptation, 
+		we traced the origin of the segment of DNA back in time to classify them as 
+		\textbf{(Blue)} Migrants, brought in from a migrant after the introduction of the new lakes 
+		\textbf{(Orange)} Pre-existing, passed down from an individual in the original generation of the new lakes, 
+		however, not defined to be an adaptive variant in the original lakes
+		\textbf{(Green)} Captured, Pre-existing, passed down from an individual in the original generation of the new lakes, 
+		and defined to be adaptive in the original lakes
+		\textbf{(Red)} De novo, mutated and propagated from an individual existing in the introduced lakes population, after introduction. 
 		}

### 2018-09-17 Stickleback_Paper.tex
production plots + edits

@@
 
-\bill{A simulation study of gene flow and local polygenic adaptation in stickleback fish}
+A simulation study of gene flow and local polygenic adaptation in stickleback fish
 
-\textbf{JARGON WE SHOULD THINK ON ??}
+effect loci (locus) | QTL | causal: = ? REGIONS UNDERLYING TRAIT VALUE , GENES, ---
 
-effect loci (locus) | QTL | causal: = ?
-
-Original Lakes | initial | $P_{1}$, $P_{3}$, $P_{3}$ = ?
-
-Migration Rate | Historical Introgression = ?
-
-pre-existing freshwater adapted alleles | standing genetic variants = ?
-
-\section{Abstract}
+\section*{Abstract}
 
@@
 and large migration load at high gene flow.
-This is mediated mainly by the total downstream influx of alleles \emph{(maybe??)(No)}.
-We do not see evidence for strong effects encouraging genomic clustering of causal alleles
-or towards particular dominance coefficients.
+Rapid adaptation occurs when the freshwater haplotype has been maintained,
+through historical introgression, in marine environments without tight linkage to 
+marine adapted alleles.
+%This is mediated mainly by the total downstream influx of alleles \emph{(maybe??)(No)}.
+%We do not see evidence for strong effects encouraging genomic clustering of causal alleles
+%or towards particular dominance coefficients.
 $F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow,
@@
 
-\section{Introduction}
+\section*{Introduction}
 
@@
 %%%%%%%%%%%%%%%%
-\section{Methods}
+\section*{Methods}
 
@@
 Possibly the most important caveat is that simulated population sizes are much smaller
-than the census size of the threespine stickleback population;
+than the census size of the Threespine Stickleback population;
 we will consider the likely effects of this in the Discussion.
@@
 New mutations occur with probability $10^{-10}$ per locus per generation in regions affecting phenotype and $10^{-8}$ in neutral regions.
-Ten ``effect regions'' of $100,000$ loci each
-are spread evenly along the chromosome (separated by $495,0000$ loci)
-in which mutations affect the trait;
-mutations in other parts of the genome are neutral.
+Ten ``effect regions'' of $100 Kb$  loci each are spread evenly along the chromosome; 
+mutations in these regions have the potential to affect trait value of the individual which posses them
 Each mutation in these regions is either additive, completely recessive,
-or completely dominant (with equal probability),
-and with effect sizes chosen randomly from an Exponential distribution with mean $1/2$,
+or completely dominant (with equal probability).
+Effect sizes are chosen randomly from an exponential distribution with mean $1/2$,
 either positive or negative with equal probability.
+Between the effect regions is $9.9 Mb$ of neutral loci.
 Individual trait values are determined additively from the diploid genotypes.
 Concretely, an individual that is heterozygous and homozygous for mutations
-at loci $H$ and $D$ respectively
-has trait value
+at loci $H$ and $D$ respectively has trait value
 $x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
@@
 The between-habitat migration rate is denoted $m$,
-and will be called simply the ``migration rate''.
-Each new individual in the marine habitat has freshwater parents with probability $m$;
-to obtain the pair, a first parent is chosen proportional to fitness,
+and will be called simply the ``gene flow''.
+$\approx m$ individuals \emph{per} lake \emph{per} generation, have parents from the opposing selective pressure.
+A first parent is chosen proportional to fitness,
 and a mate is chosen from the same lake as the first, also proportional to fitness.
@@
 we introduce a new set of lakes midway through the simulation.
-The initial set of individuals have parents chosen as above from the marine habitat,
-and act like an independent copy of the original set of lakes -- in particular,
-the two sets of lakes each have 5,000 individuals at all times.
+As with the old lakes, these are a set of discrete demes connected by migration \emph{only} to the marine.
+The initial generation of individuals in these new lakes were created as a copy of the marine individuals 
+at the generation of introduction.
+The new lake's geography acts like an independent copy of the original set of lakes -- in particular,
+the two sets of lakes each have 5,000 individuals at all times 
+(Gene flow constricted to within the lakes and through migration to the marine environment).
 Since this doubles the number of lake-to-marine immigrants,
@@
 This case study involves analyzing large simulations which model populations under the influence of selection. 
-To fully understand the origins of adaptive variants, calculate statistics, and avoid the computationally expensive task of 
+To fully understand the origins of freshwater alleles, calculate statistics, and avoid the computationally expensive task of 
 simulating neutral mutations, we have used SLiM's \emph{treeSeq} feature to track the genealogical history of all samples throughout the simulation. 
 This feature outputs a compact representation the genomic outcome of every meiosis relevant to tracked samples (Often thought of as the ARG).
-\jgg{maybe the below should be in descriptive statistics}
-To understand recomposition of the freshwater haplotype in the new lakes through selection, 
-we were able to classify each adaptive variant, in each genome in the population at the generation at time of adaptation,
+
+Using the output tree sequence, 
+we were able to classify each freshwater allele in each genome of the new lakes population, at time of adaptation,
 as deriving from one of four origins:
@@
 (4) De novo, mutated and propagated from an individual existing in the introduced lakes population, after introduction. 
+\jgg{We should discuss further what these categories aught to be named}
 
+\begin{comment}
 We also used the tree sequence to count the "fitness" of individuals in the initial generation of the new lakes. 
@@
 for each individual in the initial generation. 
+\end{comment}
+
+To determine the ability for the new lakes to select upon the standing variation in the marine at different rates of historical introgression,
+We were able compute the trait value of all effect regions in the initial population.
+Given the probability that effect region fixes at a given trait value, $2 * x / 45$, where $x$ is the effect size of a haplotype,
+we determined the expected total effect sizes of the haplotypes that fix.
+Concretely, it is the sum across the 10 effect regions of
+$$\sum_{i=1}^N \prod_{j < i} (1 - p_j) p_i x_i$$
+where $N$ is the number of genomes per lake, $p_i$ is the probability of fixation of the 
+$i^{th}$ haplotype, $x_{i}$ is the effect size of the $i^{th}$ haplotype, and the haplotypes are sorted in decreasing order by $x_{i}$.
+These numbers were computed assuming additivity of mutations, meaning the numbers produced are a slight overestimate.
 
@@
 To assess whether new lakes adapt using existing genetic diversity,
-we define a \emph{pre-existing freshwater adapted allele}
+we define a \emph{freshwater allele}
 to be an effect mutation that has frequency higher than $0.5$ in at least one of the original lakes,
@@
 but simply suggests that any sufficiently common freshwater adapted allele 
-could participate in adaptaiton in new habitat \citep{schluter2009genetics}.
+could participate in adaptation in new habitat \citep{schluter2009genetics}.
 
@@
 
-\section{Results}
+\section*{Results}
 
-We varied the migration rate, $m$, across separate simulations
-from $5 \times 10^{-5}$ to $5 \times 10^{-1}$ per individual,
-i.e., between 0.01 and 100 migrants per year to and from each of the freshwater lakes.
+We varied the migration rate, $m$ across separate simulations
+from $5 \times 10^{-5}$ to $5 \times 10^{-1}$ by steps one order of magnitude.
+From here on out we will refer to these rates of gene flow as running simulations ranging from {0.01 - 100.0}
+migrants \emph{per} lake \emph{per} generation.
+%i.e., between 0.01 and 100 migrants per generation to and from each of the freshwater lakes.
 This had a strong effect on many aspects of adaptation,
 including how fast adaptation occurred in each lake,
-how much alleles were shared between lakes,
+how many alleles were shared between lakes,
 and the population genetic signals left behind.
+We find a threshold of gene flow which allows 
+the freshwater haplotype to be maintained in the marine without creating linkage to marine alleles. 
+below this threshold, selection on pre-existing freshwater alleles to be much less efficient 
+due to background selection. \jgg{did I get this right?}
+
+Interestingly, all rates of $m$ aside from the two highest had little affect on populations ability to locally adapt.
+All rates of gene flow aside from the lowest, helped both the old lake's and the new lake's freshwater populations share pre-existing freshwater adapted alleles.
+The sharing of alleles at $m = 1.0$ resulted in rapid adaptation ($\approx 32$ generations) of the introduced population to adapt to the freshwater environment.
 At the lowest migration rate, lakes adapted nearly independently,
-while at the highest migration rate, the habitats are beginning to act like a single metapopulation
+while at the highest migration rate, the habitats act like a single metapopulation
 with substantial migration load.
 We will now dissect what happens between these extremes.
-Interestingly, all rates of $m$ aside from the two highest had little affect on populations ability to locally adapt.
-All rates of historical introgression aside from the lowest, helped both the initial and introduced freshwater populations 
-share pre-existing freshwater adapted alleles.
-The sharing of alleles resulted in rapid adaptation ($\approx 60$ generations) of the introduced population (split from the marine population) to adapt to the freshwater environment.
-We also found that larger amounts of migration allow for more statistical power and less false positives in the resulting population genetics data ($F_{st}$ per SNP across the genome).
 
@@
 
-\jgg{This needs some work}
+\begin{figure}
+	\begin{center}
+        \includegraphics{Final_Plots/Pheno_Time.pdf}
+  		\caption{ 
+        		Mean individual trait values in the marine habitat (blue line),
+        		the original lakes (light green lines; average in dark green),
+        		and the introduced lakes (yellow lines; average in orange)
+        		across the course of two simulations, with migration rates of
+        		\textbf{(top)} $m=5 \times 10^{-5}$ and
+        		\textbf{(bottom)} $m=5 \times 10^{-4}$ per generation individual
+        		(i.e., 0.01 and 0.1 migrants per lake per generation, respectively).
+        		Optimal trait values in the two habitats are at $\pm 10$.
+		}
+  		\label{fig:phenotype_ts2}
+	\end{center}
+\end{figure}
+
 Local adaptation occurred in all simulations:
@@
 were responsible for most of the adaptive differences between freshwater and marine habitats.
-\plr{ALSO TODO: CHECK THIS}
-
-\begin{figure}
-	\begin{center}
-        \includegraphics[width=1.0\linewidth]{semi_final_plots/Pheno_semi_final.png}
-  		\caption{ 
-        		Mean individual trait values in the marine habitat (blue line),
-        		the original lakes (light green lines; average in dark green),
-        		and the introduced lakes (yellow lines; average in orange)
-        		across the course of two simulations, with migration rates of
-        		\textbf{(top)} $m=5 \times 10^{-5}$ and
-        		\textbf{(bottom)} $m=5 \times 10^{-4}$ per generation individual
-        		(i.e., 0.01 and 0.1 migrants per lake per generation, respectively).
-        		Optimal trait values in the two habitats are at $\pm 10$.
-		}
-  		\label{fig:phenotype_ts2}
-	\end{center}
-\end{figure}
 
@@
 	\begin{center}
-  		\includegraphics[width=1.0\linewidth]{semi_final_plots/Fst_semi_final.png}
+  		\includegraphics{Final_Plots/Fst_Genome.pdf}
   		\caption{
@@
 As shown in Figure \ref{fig:TimeToAdaptation}, at the lowest migration rate 
-it took over $20,000$ generations for the mean trait value across introduced lakes to 
+it took over $18,000$ generations for the mean trait value across introduced lakes to 
 get to within 0.5 of the original lakes value.
 At $m = 1.0$, we saw the introduced population adapt in just $32$ generations. 
-This rate of adaptation follows what we've observed in natural populations of Threespine stickleback.
+This rate of adaptation follows what we've observed in natural populations of threespine stickleback.
 At the highest migration rate, $m = 100$, all individuals were unable to locally adapt.
@@
 	\begin{center}
-  		\includegraphics[width=1.0\linewidth]{semi_final_plots/T_adapt_semi_final.png}
+  		\includegraphics{Final_Plots/Time_Adapt.pdf}
   		\caption{
@@
 
-\subsection*{Sharing of freshwater adapted alleles}
+\subsection*{Sharing of freshwater alleles}
+
+We find that this dramatic increase in speed of local adaptation occurs because
+higher gene between populations allow sharing of freshwater alleles more effectively when compared to de novo mutation.
+At low migration rates, adaptation occurs independently in each lake.
+In Figure \ref{fig:Origin}, at $0.01$, we can see that the large majority of adaptive alleles derived from de novo mutation in the 
+population rather than selection on standing variants in the ocean. 
+As The migration increases, we see a larger ratio of the origins of adaptive alleles 
+derive from pre-existing variation in the marine population at the time of introduction.
+In other words, 
+greater mixing at higher migration rates allows lakes to share alleles
+instead of developing their own genetic basis of adaptation.
+To further investigate sharing of locally adaptive alleles and the transporter hypothesis,
+we investigate the ``freshwater alleles'', defined for a particular generation
+to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat in that generation.
+
+As the number of common, locally adaptive alleles decreases, as seen in Figure \ref{fig:MPFAI}C,
+the genetic basis of adaptation is more commonly shared.
+Figure \ref{fig:MPFAI}B shows
+the percentage of currently-defined, freshwater-adapted alleles 
+that each genome in each of the populations carries,
+averaged across time and individuals.
+If all individuals across lakes carried the same set of alleles determining their trait value,
+this would be 100\%.
+This value is nearly reached at the fourth migration rate;
+but it is lower due to migration load.
+At the lowest migration rate, 
+each genome in the original lakes have almost exactly $1/25^{th} = 0.04\%$   
+of the total freshwater adapted alleles --
+since there are 25 lakes, this implies that each lake has adapted with a unique set of alleles.
+Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
+Figure \ref{fig:MPFAI}A shows us that by $0.1$ migrants per lake per generation,
+the average individual across the new lakes has nearly the same amount of 
+``pre-existing freshwater adapted alleles" as individuals across the old lakes.
+Interestingly, the genetic basis of the freshwater phenotype 
+seems to simplify with greater $m$. 
+This would seem to suggest higher rates of migration allowed 
+adaptive alleles of higher effect to travel more efficiently through the populations
+\jgg{is this right?}
 
@@
 	\begin{center}
-  		\includegraphics[width=\linewidth]{semi_final_plots/FAA_semi_final.png}
+  		\includegraphics{Final_Plots/Freshwater_Alleles.pdf}
   		\caption{
-		Distribution of freshwater adapted alleles among the three populations. 
+		Distribution of freshwater alleles among the three populations. 
 		A freshwater adapted variant is defined as, at any generation throughout the simulation,
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% I'M HERE
-
-This dramatic increase in speed of local adaptation occurs because
-at low migration rates, adaptation occurs independently in each lake,
-and the marine habitat has ten times the influx of new alleles than any one lake.
-In other words, 
-greater mixing at higher migration rates allows lakes to share alleles
-instead of developing their own genetic basis of adaptation.
-%As a first indication of this, Figure \ref{fig:Fst} shows that
-%$F_{ST}$ between the ``original'' and ``pre-existing'' sets of lakes 
-%at effect mutations decreased with migration rate.
-To further investigate sharing of locally adaptive alleles and the transporter hypothesis,
-we investigate the ``pre-existing freshwater adapted alleles'', defined for a particular generation
-to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat in that generation.
-
-As the number of common, locally adaptive alleles decreases,
-the genetic basis of adaptation is more commonly shared.
-Figure \ref{fig:MPFAI}B shows
-the percentage of currently-defined, freshwater-adapted alleles 
-that each genome in each of the populations carries,
-averaged across time and individuals.
-If all individuals across lakes carried the same set of alleles determining their trait value,
-this would be 100\%.
-This value is nearly reached at the highest migration rate;
-but it is lower due to migration load.
-At the lowest migration rate, 
-each genome in the original lakes have almost exactly $1/25^{th} = 0.04\%$   
-of the total freshwater adapted alleles --
-since there are 25 lakes, this implies that each lake has adapted with a unique set of alleles.
-Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
-
-In Figure \ref{fig:MPFAI}C we can see that; as migration rates increase, the number of these alleles decreases steadily.
-On the left in Figure \ref{fig:MPFAI}B, the mean percentage of alleles per individual rapidly increases. 
-Together, these two results strongly suggest the sharing genetic basis.
-\jgg{TODO, reference the number of lakes that each allele appears in, make plot for supp}
-%As a first indication of this, Figure \ref{fig:MPFAI}B shows that the introduced 
-%population utilizes a higher percentage of "pre-existing freshwater adapted alleles"
-%with higher rates of introgression between the populations.
-Figure \ref{fig:MPFAI}A shows the distribution of the mean number of pre-existing freshwater adapted alleles,
-Throughout the simulations, in each deme.
-Interestingly, the genetic basis of the freshwater phenotype 
-seems to simplify with greater $m$. 
-This would seem to suggest higher rates of migration allowed 
-adaptive alleles of higher effect to travel more efficiently through the populations
-\jgg{is this right?}
-
 \subsection*{Migration load and genetic variation}
 
-\begin{figure}[h!tb]
-	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{semi_final_plots/PhenoDist_semi_final.png}
-  		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
-		The dashed pink line (Pheno = +10) is the optimum phenotype any individual in the marine environment.
-		In contrast the purple line at (Pheno = -10) represents the optimum for any individual in the freshwater environment. 
-		All individuals at generation 0 (beginning of the simulation) 
-		}
-  		\label{fig:MeanPhenotype}
-	\end{center}
-\end{figure}
-
 At first, increased migration allows sharing of adaptive alleles between lakes,
-but at the second from highest migration rate,
-the constant influx of alleles between the habitats creates substantial migration load.
+but at $m = 5 \times 10^{-2}$
+the constant influx of alleles between the habitats creates noticeable migration load.
 This rate, $m=.05$, only replaces 5\% of each population each generation
@@
 
-\subsection*{Realized genomic architecture}
-
-\begin{figure}[h!tb]
+\begin{figure}
 	\begin{center}
-  		\includegraphics[width=1.0\linewidth]{semi_final_plots/Haplo_semi_final.png}
-  		\caption{
-			Phenotypes (\textbf{Top}) (effect size and mutation type) and haplotypes \textbf{(Botton)} as a function of spatial position of the individual 
-			possessing the genome. The phenotype of each haplotype is, by definition, the sum of the 
-			effect sizes of each mutation in the genome. 
+  		\includegraphics{Final_Plots/Pheno_Dist.pdf}
+  		\caption{Distribution of mean trait value throughout simulation runs at separate migration rate $(m)$ parameter values, for each population.
+		The dashed pink line (Trait value = +10) is the optimum phenotype any individual in the marine environment.
+		In contrast the purple line at (Trait value = -10) represents the optimum for any individual in the freshwater environment. 
+		All individuals at generation 0 (beginning of the simulation) 
 		}
@@
 
-Now we take a closer look at the genomic architecture of local adaptation 
-between the two habitats.
-Do the alleles underlying trait differences cluster along the genome?
-Do measures of local differentiation identify the causal loci?
-Figure \ref{fig:Fst3} shows plots along the genome of per-locus $F_{ST}$ values
-between the marine and original freshwater habitats
-at the two intermediate migration rates.
-(Note that we are pooling freshwater habitats; 
-a single lake would provide substantially less power.)
-
-\plr{need to know where the FAA are in these figures}
+\subsection*{Realized genomic architecture}
 
@@
 and second, greater sharing of adaptive alleles providing a shared signal across populations.
-
 This suggests that genome scans for local adaptation
@@
 will only be successful given enough migration between habitats.
-To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates
-that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value
-to be a causal locus.
-
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{semi_final_plots/TP_SP_semi_final.png}
-  		\caption{ 
-		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
-		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
-		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
-        		TODO: x-axis label ($F_{ST}$).
-		}
-  		\label{fig:Power_FP}
-	\end{center}
-\end{figure}
-
-With 10 effect regions across the genome, we can see in \ref{fig:Fst3} that only about 1 or 2 of the region contain peaks. 
+With 10 effect regions across the genome, we can see in \ref{fig:Fst} that only about 1 or 2 of the region contain peaks. 
 When looking closely at the peaks we see that each is usually a composition of 7-8 SNPs close in proximity.
@@
 
-In regions of the genome underlying individual trait value, we observed 
-Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$. 
+Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$?
 Knowing exactly which mutations effect phenotype in our simulations, 
 we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome. 
-In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power. 
+In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, 
+we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power. 
 This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.
+
+\subsubsection*{Origin of introduced freshwater adaptive alleles: a conundrum}
+Up to this point, we have quantified rapid and parallel adaptation and found strong evidence that sharing of alleles is (more than likely?) responsible.
+However in Figure \ref{fig:MPFAI}A at $0.1$ \emph{mig/gen/lake}, we see sharing of alleles between populations,
+yet \emph{Time to adaptation} is still fairly slow ($\approx 2,000$ gens).
+This is surprising when considering we see a similar quantity of allele sharing at  $m$ equal to $1.0$ \emph{mig/gen/lake}, 
+but much more rapid adaptation ($\approx 30$ gens). 
+Initially, this 60X speedup of adaptation might seem to be the result of subsequent migration from the marine after the point of freshwater introduction. 
+To investigate this further we have track the origins of all adaptive variants in all new lake extant individuals at the time of adaptation. 
+However, in Figure \ref{fig:Origin}, we find that at both $m = 0.1$ and $m = 1.0$ \emph{migs/gen/lake}, the vast majority of adaptive alleles origins were found to propagate 
+from individuals in the original generation ("Captured" OR "Pre-Existing").
 
@@
 	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{semi_final_plots/OriginRatio.png}
+  		\includegraphics[width=0.7\linewidth]{Final_Plots/Allele_Origin.pdf}
   		\caption{ 
@@
 
-%The lowest migration rate of $10^{-5}$ also has a very low false positive rate across all $F_{st}$ threshold values.
-%\subsubsection*{Low Recombination Causes Clustering}
-\subsubsection*{Origin of introduced freshwater adaptive alleles: a conundrum}
-Up to this point, we have quantified rapid and parallel adaptation and found strong evidence that sharing of alleles is (more than likely?) responsible.
-However in Figure \ref{fig:MPFAI} at $m$ equal to $0.1$ \emph{mig/gen/lake}, we see sharing of alleles between populations,
-yet \emph{Time to adaptation} is still fairly slow ($\approx 2,000$ gens).
-This is surprising when considering we see a similar quantity of allele sharing at  $m$ equal to $1.0$ \emph{mig/gen/lake}, 
-but much more rapid adaptation ($\approx 60$ gens). 
-Initially, this 30X speedup of adaptation might seem to be the result of subsequent migration from the marine after the point of freshwater introduction. 
-This theory would assert that the continued influx of alleles is necessary for rapid local adaptation at the rate which we've observed in Nature.
-However, in Figure \ref{fig:Origin}, we find that at both $m = 0.1$ and $m = 1.0$, the vast majority of adaptive alleles origins were found to propagate 
-from individuals in the original generation ("Captured" OR "Pre-Existing").
+The alternative explanation is that at a higher migration rate, a higher saturation of marine adapted alleles are linked to freshwater adapted alleles.
+In turn, this linkage acts as a barrier rebuilding the freshwater haplotype, in the new lakes, much more difficult as it much wait for recombination to break the links.
+With more historical introgression between the two selective pressures, the freshwater haplotype would remain more "intact" and much more efficient to rebuild in the lake.
+To confirm this, we decided to look at the trait values of the ten affect regions in each genome of the initial population.
+Given the probability that effect region fixes at a given trait value, $2 * x / 45$ where $x$ is the effect size of a haplotype
+we were able to determine expected total effect sizes of the haplotypes that fix (detail about computation in methods).
+Below, we see the mean of these expected total effect sizes across all lakes.
 
-The alternative explanation is that at a higher migration rate, a higher saturation of marine adapted alleles are linked to freshwater adapted alleles.
-This linkage makes rebuilding the freshwater haplotype, in the new lakes, much more difficult as it much wait for recombination to break the links.
-With more historical introgression between the two selective pressures, the freshwater haplotype would remain more "intact" and much more efficient
-to rebuild in the lake.
+\begin{center}
+\begin{tabular}{ | c | c  | c  | c | }
+ \hline
+ $m = 0.01$ & $m = 0.1$ & $m = 1.0$ & $m = 10.0$ \\ 
+ $0.1227$ &  $4.6976$ & $12.9298$  & $12.2710$\\
+ \hline
+\end{tabular}
+\end{center}
 
-%Now lets make this more concrete
+Here, we see a dramatic difference in the total effect sizes of the haplotypes that are likely to fix.
+This provides evidence that linkage to marine alleles has a major impact in the introduced population's ability 
+to select and rebuild a pre-existing haplotype given the standing variation at the time of introduction. 
+\jgg{Below may belong in the discussion / future work}
+This also brings to light the role of recombination in a system of migration selection balance. 
+On one hand, a lower recombination rate would allow for the freshwater haplotype to remain "intact" 
+within the marine environment, however, without significant introgression the haplotype would quickly get selected against in marine populations. 
+On the other hand a higher recombination rate would, in theory, allow freshwater adapted alleles more longevity 
+as they would hitchhike with marine adapted haplotypes.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Theoretical expectations}
@@
 %%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
+\section*{Discussion}
+
+\jgg{TODO: effect of smaller than realistic population sizes?}
+
+\jgg{TODO: Talk about the role recombination plays}
 
@@
 \appendix
-\section{Supp.}
+\section*{Supp.}
+\begin{figure}
+	\begin{center}
+  		\includegraphics{Final_Plots/True_Power.pdf}
+  		\caption{ 
+		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
+		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
+		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
+        		TODO: x-axis label ($F_{ST}$).
+		}
+  		\label{fig:Power_FP}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=1.0\linewidth]{semi_final_plots/Haplo_semi_final.png}
+  		\caption{
+			Phenotypes (\textbf{Top}) (effect size and mutation type) and haplotypes \textbf{(Botton)} as a function of spatial position of the individual 
+			possessing the genome. The phenotype of each haplotype is, by definition, the sum of the 
+			effect sizes of each mutation in the genome. 
+		}
+  		\label{fig:Haplo_Pheno}
+	\end{center}
+\end{figure}
 

### 2018-09-17 Stickleback_Paper.tex
last of jared edits, rough draft 1

@@
 marine adapted alleles.
-%This is mediated mainly by the total downstream influx of alleles \emph{(maybe??)(No)}.
-%We do not see evidence for strong effects encouraging genomic clustering of causal alleles
-%or towards particular dominance coefficients.
-$F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow,
+We find $F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow.
 The results support existing theory of local adaptation,
@@
 and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region.
-% \jgg{Bill - more specifics about Bill, Suzie, Kristin  \& or Thom's papers .. ? }
 These results leave us with questions surrounding the nature of rapid adaptation.
@@
 The new lake's geography acts like an independent copy of the original set of lakes -- in particular,
-the two sets of lakes each have 5,000 individuals at all times 
-(Gene flow constricted to within the lakes and through migration to the marine environment).
+the two sets of lakes each have 5,000 individuals at all times.
+%(Gene flow constricted to within the lakes and through migration to the marine environment).
 Since this doubles the number of lake-to-marine immigrants,
-after this happens the probability that a new marine individual has freshwater parents
+after this happens the probability that a new marine individual has parents from old lakes or the new lakes
 is $2m$ instead of $m$.
@@
 
-To determine the ability for the new lakes to select upon the standing variation in the marine at different rates of historical introgression,
-We were able compute the trait value of all effect regions in the initial population.
+We were also able to use the tree sequence to get information about the individual ``effect regions" 
+in all initial genomes of the new lakes at time of introduction.
+From this we determined the ability for the new lakes to select upon the standing variation in the marine.
 Given the probability that effect region fixes at a given trait value, $2 * x / 45$, where $x$ is the effect size of a haplotype,
-we determined the expected total effect sizes of the haplotypes that fix.
+we calculated the expected total effect sizes of the haplotypes that fix.
 Concretely, it is the sum across the 10 effect regions of
@@
 We varied the migration rate, $m$ across separate simulations
-from $5 \times 10^{-5}$ to $5 \times 10^{-1}$ by steps one order of magnitude.
-From here on out we will refer to these rates of gene flow as running simulations ranging from {0.01 - 100.0}
+from $5 \times 10^{-5}$ to $5 \times 10^{-1}$ by one order of magnitude steps.
+For clarity given our populations sizes of 5,000 diploid individuals, 200 in each lakes,
+we will refer to these rates of gene flow as running simulations ranging from {0.01 - 100.0}
 migrants \emph{per} lake \emph{per} generation.
-%i.e., between 0.01 and 100 migrants per generation to and from each of the freshwater lakes.
-This had a strong effect on many aspects of adaptation,
+The differing parameter values had a strong effect on many aspects of adaptation;
 including how fast adaptation occurred in each lake,
@@
 
-Interestingly, all rates of $m$ aside from the two highest had little affect on populations ability to locally adapt.
+Interestingly, all rates of $m$ aside from the two highest had little affect on population's ability to locally adapt.
 All rates of gene flow aside from the lowest, helped both the old lake's and the new lake's freshwater populations share pre-existing freshwater adapted alleles.
@@
 
-\subsubsection*{Origin of introduced freshwater adaptive alleles: a conundrum}
+\subsubsection*{Origin of introduced freshwater adaptive alleles}
 Up to this point, we have quantified rapid and parallel adaptation and found strong evidence that sharing of alleles is (more than likely?) responsible.
-However in Figure \ref{fig:MPFAI}A at $0.1$ \emph{mig/gen/lake}, we see sharing of alleles between populations,
+Interestingly, in Figure \ref{fig:MPFAI}A at $0.1$ \emph{mig/gen/lake}, we see sharing of alleles between populations,
 yet \emph{Time to adaptation} is still fairly slow ($\approx 2,000$ gens).
@@
 Initially, this 60X speedup of adaptation might seem to be the result of subsequent migration from the marine after the point of freshwater introduction. 
+perhaps a larger influx of adapted alleles are needed in addition to standing variation in the initial generation.
 To investigate this further we have track the origins of all adaptive variants in all new lake extant individuals at the time of adaptation. 
-However, in Figure \ref{fig:Origin}, we find that at both $m = 0.1$ and $m = 1.0$ \emph{migs/gen/lake}, the vast majority of adaptive alleles origins were found to propagate 
-from individuals in the original generation ("Captured" OR "Pre-Existing").
+Surprisingly in Figure \ref{fig:Origin}, we find that at both $m = 0.1$ and $m = 1.0$ \emph{migs/gen/lake},
+the vast majority of adaptive alleles origins were found to propagate from individuals in the original generation ("Captured" OR "Pre-Existing").
 
@@
 
-The alternative explanation is that at a higher migration rate, a higher saturation of marine adapted alleles are linked to freshwater adapted alleles.
-In turn, this linkage acts as a barrier rebuilding the freshwater haplotype, in the new lakes, much more difficult as it much wait for recombination to break the links.
-With more historical introgression between the two selective pressures, the freshwater haplotype would remain more "intact" and much more efficient to rebuild in the lake.
+The alternative explanation, is that at a higher migration rate, a higher saturation of marine adapted alleles are linked to freshwater adapted alleles.
+In turn, the equilibrium of freshwater alleles in marine genomes created by migration-selection balance at lower rates of gene flow acts as a barrier when rebuilding the freshwater haplotype. 
+With more gene flow between the two selective pressures, the freshwater haplotype would remain more "intact" and much more efficient to rebuild in the introduced environment.
 To confirm this, we decided to look at the trait values of the ten affect regions in each genome of the initial population.
-Given the probability that effect region fixes at a given trait value, $2 * x / 45$ where $x$ is the effect size of a haplotype
+Given the probability that effect region fixes at a given trait value, $2 \times x / 45$ where $x$ is the effect size of a haplotype
 we were able to determine expected total effect sizes of the haplotypes that fix (detail about computation in methods).
@@
  $m = 0.01$ & $m = 0.1$ & $m = 1.0$ & $m = 10.0$ \\ 
- $0.1227$ &  $4.6976$ & $12.9298$  & $12.2710$\\
+ $-0.1227$ &  $-4.6976$ & $-12.9298$  & $-12.2710$\\
  \hline
@@
 
-Here, we see a dramatic difference in the total effect sizes of the haplotypes that are likely to fix.
+Between $m = 0.1$ and $m = 1.0$ we see a large difference in the total effect sizes of the haplotypes that are expected to fix.
+Given these numbers were computed assuming additivity (slight overestimate) and individuals are diploid,
+at $m = 0.1$ the mean total genotype expected to fix in the populations puts individuals at a trait value of $\approx -8$ 
+after selection on alleles haplotypes in the initial population. 
+In turn, the population must wait for novel mutation or subsequent migration to provide the alleles for the remaining 2 units before 
+attaining a mean average trait value at the optimum (-10).
+It follows in Figure \ref{fig:phenotype_ts2}B that the majority time to adaptation is spent within 2 units of mean trait value.
+The initial adaptation on selection of standing variation up to that point happens quite rapidly.
 This provides evidence that linkage to marine alleles has a major impact in the introduced population's ability 
-to select and rebuild a pre-existing haplotype given the standing variation at the time of introduction. 
-\jgg{Below may belong in the discussion / future work}
-This also brings to light the role of recombination in a system of migration selection balance. 
-On one hand, a lower recombination rate would allow for the freshwater haplotype to remain "intact" 
-within the marine environment, however, without significant introgression the haplotype would quickly get selected against in marine populations. 
-On the other hand a higher recombination rate would, in theory, allow freshwater adapted alleles more longevity 
-as they would hitchhike with marine adapted haplotypes.
+to select and rebuild a pre-existing haplotype given the standing variation at time of introduction.
 
@@
 
-\jgg{TODO: effect of smaller than realistic population sizes?}
-
-\jgg{TODO: Talk about the role recombination plays}
-
 We have shown that historical introgression, at our given parameter sets,
@@
 
+\jgg{TODO: effect of smaller than realistic population sizes?}
+
+\jgg{TODO: Talk about the role recombination plays}
+\jgg{Below may belong in the discussion / future work}
+This also brings to light the role of recombination in a system of migration selection balance. 
+On one hand, a lower recombination rate would allow for the freshwater haplotype to remain "intact" 
+within the marine environment, however, without significant introgression the haplotype would quickly get selected against in marine populations. 
+On the other hand a higher recombination rate would, in theory, allow freshwater adapted alleles more longevity 
+as they would hitchhike with marine adapted haplotypes.
+
 \bibliographystyle{plainnat}
@@
 	\begin{center}
-  		\includegraphics[width=1.0\linewidth]{semi_final_plots/Haplo_semi_final.png}
+  		\includegraphics[width=1.0\linewidth]{Final_Plots/Haplo_small.pdf}
   		\caption{

### 2018-09-24 Stickleback_Paper.tex
latest

@@
 
-effect loci (locus) | QTL | causal: = ? REGIONS UNDERLYING TRAIT VALUE , GENES, ---
+\textbf{Jargon}
+
+effect loci? regions underlying trait value?
+
+names of four allele origin categories
 
@@
 Ten ``effect regions'' of $100 Kb$  loci each are spread evenly along the chromosome; 
-mutations in these regions have the potential to affect trait value of the individual which posses them
+mutations in these regions have the potential to affect trait value of the individual which posses them.
 Each mutation in these regions is either additive, completely recessive,
@@
 \subsubsection*{Origin of introduced freshwater adaptive alleles}
-Up to this point, we have quantified rapid and parallel adaptation and found strong evidence that sharing of alleles is (more than likely?) responsible.
-Interestingly, in Figure \ref{fig:MPFAI}A at $0.1$ \emph{mig/gen/lake}, we see sharing of alleles between populations,
+Up to this point, we have quantified rapid adaptation and found strong evidence that sharing of alleles is responsible in this system.
+Interestingly, in Figure \ref{fig:MPFAI}A at $m = 0.1$, we see sharing of alleles between populations,
 yet \emph{Time to adaptation} is still fairly slow ($\approx 2,000$ gens).
-This is surprising when considering we see a similar quantity of allele sharing at  $m$ equal to $1.0$ \emph{mig/gen/lake}, 
+This is surprising when considering we see a similar quantity of allele sharing at  $m = 1.0$, 
 but much more rapid adaptation ($\approx 30$ gens). 
-Initially, this 60X speedup of adaptation might seem to be the result of subsequent migration from the marine after the point of freshwater introduction. 
-perhaps a larger influx of adapted alleles are needed in addition to standing variation in the initial generation.
-To investigate this further we have track the origins of all adaptive variants in all new lake extant individuals at the time of adaptation. 
-Surprisingly in Figure \ref{fig:Origin}, we find that at both $m = 0.1$ and $m = 1.0$ \emph{migs/gen/lake},
-the vast majority of adaptive alleles origins were found to propagate from individuals in the original generation ("Captured" OR "Pre-Existing").
+Initially, this 60X speedup of adaptation seems to be the result of subsequent gene flow of freshwater alleles. 
+Perhaps a larger influx of freshwater alleles are needed in addition to standing variation in the initial generation.
+To investigate this further we have tracked the origins of all freshwater alleles, in all new lake individuals at the time of adaptation. 
+Surprisingly in Figure \ref{fig:Origin}, we find that at both $m = 0.1$ and $m = 1.0$,
+the vast majority of adaptive alleles were found to propagate from individuals in the original generation ("Captured" OR "Pre-Existing").
 
@@
 
-The alternative explanation, is that at a higher migration rate, a higher saturation of marine adapted alleles are linked to freshwater adapted alleles.
-In turn, the equilibrium of freshwater alleles in marine genomes created by migration-selection balance at lower rates of gene flow acts as a barrier when rebuilding the freshwater haplotype. 
-With more gene flow between the two selective pressures, the freshwater haplotype would remain more "intact" and much more efficient to rebuild in the introduced environment.
+The alternative explanation, is that at a lower migration rate, a higher saturation of marine adapted alleles are linked to freshwater adapted alleles.
+In turn, the equilibrium of freshwater alleles in marine genomes created by migration-selection balance acts as a barrier when rebuilding the freshwater haplotype. 
+%With more gene flow between the two selective pressures, the freshwater haplotype remains more "intact" and much more efficient to rebuild in the introduced environment.
 To confirm this, we decided to look at the trait values of the ten affect regions in each genome of the initial population.
-Given the probability that effect region fixes at a given trait value, $2 \times x / 45$ where $x$ is the effect size of a haplotype
+Given the probability that an ``effect region" fixes at a given trait value, $2 \times x / 45$ where $x$ is the effect size of a haplotype,
 we were able to determine expected total effect sizes of the haplotypes that fix (detail about computation in methods).

### 2018-10-22 Stickleback_Paper.tex
Update on Overleaf.

--- 2018-09-17 Stickleback_Paper.tex
@@
 
-\title{The sweet spot for local polygenic adaptation: a case study motivated by stickleback}
+\title{A simulation study of gene flow and local polygenic adaptation motivated by threespine stickleback evolution}
 \author{Jared Galloway, William A. Cresko, and Peter Ralph}
@@
 
-How do stickleback share adaptations between lakes? A simulation study.
+\textbf{JARGON WE SHOULD THINK ON ??}
 
-A simulation study of gene flow and local polygenic adaptation
+effect loci (locus) | QTL | causal: = ?
 
-A simulation study of gene flow and local polygenic adaptation in stickleback fish
+Original Lakes | initial | $P_{1}$, $P_{3}$, $P_{3}$ = ?
 
-effect loci (locus) | QTL | causal: = ? REGIONS UNDERLYING TRAIT VALUE , GENES, ---
+Migration Rate | Historical Introgression = ?
 
-\section*{Abstract}
+pre-existing freshwater adapted alleles | standing genetic variants = ?
 
-Threespine stickleback fish provide one of the most striking examples of local adaptation.
-This hemisphere-wide metapopulation includes both marine populations and a large number of smaller freshwater populations
-that have apparently repeatedly adapted to freshwater conditions
-often using the same genetic basis.
-In this paper, we use simulations motivated by stickleback populations
-to examine what amounts of gene flow favor stable metapopulation polymorphism
-with allele sharing,
-and to further dissect the underlying dynamics.
-We find that rapid, repeated adaptation 
-using alleles maintained at low frequency by migration-selection balance
-(the ``transporter hypothesis'')
-occurs over a realistic range of intermediate rates of gene flow,
-between slow, independent adaptation with low gene flow
-and large migration load at high gene flow.
-Rapid adaptation occurs when the freshwater haplotype has been maintained,
-through historical introgression, in marine environments without tight linkage to 
-marine adapted alleles.
-%This is mediated mainly by the total downstream influx of alleles \emph{(maybe??)(No)}.
-%We do not see evidence for strong effects encouraging genomic clustering of causal alleles
-%or towards particular dominance coefficients.
-$F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow,
-The results support existing theory of local adaptation,
-and provide a more concrete look at a particular, empirically motivated example.
+\textbf{Pending things to do?} that maybe we do, maybe not
 
-\section*{Introduction}
+longer genomes ($10^8$ not $10^5$ but rescaling everything accordingly)
+
+verify that adaptation of introduced pops is due to initial diversity not subsequent migration
+
+\section{Abstract}
+
+Threespine stickleback fish provide a striking example of local adaptation in the context of recurrent gene flow.
+This Northern hemisphere-wide metapopulation includes both marine populations and a large number of smaller freshwater populations that have repeatedly adapted to freshwater conditions, often by using standing genetic variation.
+Here we use genealogical simulations to determine the levels of gene flow can best match the observed patterns of allele sharing among habitats in stickleback, and to provide a framework for better understanding of the dynamics of gene flow and local adaptation for the maintenance and reuse of standing genetic variation.
+We find that rapid, repeated adaptation using alleles maintained at low frequency by migration-selection balance occurs over a realistic range of intermediate rates of gene flow.
+Slow and independent adaptation occurs with low gene flow, whereas large migration load occurs at high gene flow, mediated mainly by the total downstream influx of alleles \emph{(maybe??)}.
+We do not see evidence for strong effects encouraging genomic clustering of causal alleles or towards particular dominance coefficients.
+$F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow, The results support existing theory of local adaptation, and provide a more concrete look at a particular, empirically motivated example.
+
+\textbf{**We need to be more crystal clear in the abstract about the findings from the paper**}
+
+\section{Introduction}
 
@@
 
-\emph{intro outline:}
-Start off with stickleback: repeated, rapid, parallel adaptation with obvious phenotypes.
-But, most species don't seem to have such obviously structured polymorphism: 
-what's special about stickleback?
-The transporter hypothesis is a mechanism; what's required to make it work?
-Other species adapt rapidly in similar ways:
-e.g., Darwin's finches and island mice;
-do these differ through not having such a big, connected reservoir of genetic variation?
-or by not having such a large total population size of the ``other'' habitat?
-Or, is this situation more common than we know, with less obvious phenotypes?
-Also, can we actually tell what's happening from genetic data?
+The canonical model for the genetics of adaptation, first formulated by Fisher in the early part of the 20th century, involves the sequential fixation of new mutations.
+While it has proven valid in numerous studies in the field and the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures.
+A growing number of studies have identified instances of convergent or parallel adaptive evolution that utilize standing genetic variation.
+What evolutionary processes promote the maintenance of re-use of such genetic polymorphisms?
+How do these processes interact with diverse geographic contexts, as well as variation in genetic processes such as recombination, to change the likelihood of so-called standing genetic variation (SGV).
 
-Recently, multiple instances of similar (parallel) underlying genetic basis of rapid local adaptation has brought about 
-questions surrounding the origins and maintenance of genetic polymorphism.
-A empirical model is the Alaskan populations of freshwater and marine Ninespine Stickleback fish.
-In 1964, The Great Alaskan Earthquake caused an uplift of Middleton island and in turn,
-introduced a group of freshwater ponds around the perimeter of the island. 
-Quickly inhabited by the surrounding marine population of Stickleback,
-\citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be 
-parallel to freshwater stickleback that have been separated for over $13,000$ years.
-In freshwater stickleback, the number of lateral plates are reduced
-and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region.
-% \jgg{Bill - more specifics about Bill, Suzie, Kristin  \& or Thom's papers .. ? }
-These results leave us with questions surrounding the nature of rapid adaptation.
-Does convergent evolution breed it's own solution (haplotype) for every new selective pressure, 
-or can these solutions can be efficiently shared across multiple sub - populations facing similar selective pressures.
+Clues come from population genomic studies in the wild that have recently been powered by advances in sequencing technologies.
+One such organism is the threespine stickleback.
+This small fish has long been used for studies of behavior, ecology and evolution in the wild, and has more recently become a model for understanding the genetic basis of adaptive evolution.
+The ancestral marine form has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere, which have each evolved a similar suite - or 'freshwater syndrome' - of phenotypes despite local differences.
+This adaptive phenotypic evolution to freshwater can occur in mere millenia, in many cases since the end of the last ice age about 13,00 years ago.
+Studies of the fossil record also clearly show that invasion of oceanic stickleback and local adaptation in freshwater has occurred repeatedly over millions of years, often involving very similar phenotypic shifts as seen in contemporary populations.
 
-The leading hypothesis for stickleback is that rather than acting on new mutations, 
-adaptation to freshwater environments is sped up through selection on standing genetic variation (SGV) found in marine populations. 
-\jgg{Bill: maybe you could make this section a little more specific}
-One clear example of this is the gene \textit{eda} 
-which has been shown to regulate the number of lateral plates. 
-While this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently.
-Novel evidence from natural populations has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
-\citet{schluter2009genetics} 
-suggested the ``transporter''-hypothesis.  
-This outlined the flow of freshwater alleles into marine populations through offspring of hybridization events.
-It suggests freshwater haplotypes are distributed though marine individuals and are continually selected upon in introduced freshwater populations.
-The continued selection on freshwater favored alleles and introgression between the two sub-species,
-allows the marine to maintain the freshwater haplotype dispersed in low frequency among its' individuals. 
-Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection reconstructs the freshwater haplotype \citep{schluter2009genetics}.
-An alternative to this hypothesis is that individuals from other freshwater environments migrate directly to the new environments, 
-and their haplotype is passed down directly. 
-This hypothesis has been shown to be unlikely due to finding a high frequency of freshwater alleles in the ocean,
-and almost no freshwater individuals.
+A finding from recent stickleback population genomic studies is that the same genomic regions, and in some cases the same haplotypes, are implicated in patterns of repeated evolution.
+These studies show that divergent stickleback populations are clearly still connected via bi-directional gene flow.
+This reuse of standing genetic variation motivated empirical studies to ask whether this phenotypic and genomic evolution could occur in decades.
+The answer is a clear yes.
+For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island.
+Quickly inhabited by the surrounding marine population of stickleback, \citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be parallel to freshwater stickleback that have been separated for over thousands of years.
+In freshwater stickleback, the number of lateral plates are reduced and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region.
 
-\plr{Bring up the possibility of more than one freshwater haplotype here.}
-\jgg{Re: should I? Doesn't Thom's work and parallel above suggest they have the same alleles?}
+ But how can evolution occur so quickly?
+The leading hypothesis for stickleback is that rather than acting on new mutations, adaptation to freshwater environments is accelerated by selection on SGV found in marine populations.
+The first clear example of the global reuse of SGV was the gene \textit{eda} which has been shown to be an important regulator for the number of lateral plates.
+While the low lateral plate version of this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently.
+More recently, population genomic studies employing genome-wide haplotype analyses has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
+Genomic regions that exhibit local adaptation, as evidenced by increased relative population genomic divergence, also show much older haplotypes that are enriched in the alternative habitats.
 
-If selection on standing genetic variants is key in rapid adaptation, 
-many questions are exposed concerning the surrounding population genetics parameters and underlying genomic architecture.
-%How rapidly can selection act on standing genetic variants at a given value of migration, $M$?
-What scale of historical introgression between populations with selective pressures $X$ and $Y$, allows for rapid and parallel local adaptation
-of a population derived from $X$ to an introduced environment with selective pressure $Y$?
-What is the origin of pre-existing adaptive variants?
+These empirical data generally support the ``transporter''-hypothesis proposed by Conte and Schluter in 2009 \citet{schluter2009genetics}, which is a conceptual model for the flow of freshwater alleles from multiple smaller freshwater populations into much larger and less structured marine populations through hybridization events.
+Alleles in this asymetrically structured metapopulation can then be recycled for subsequent freshwater adaptation.
+Several questions remain, however, about the manner and degree to which these alleles are scattered among the marine fish and must be re-assembled.
+Is it more akin to the atomization and rebuilding that occurred in the Star Trek series that motivated this hypothesis?
+Alternatively does it occur in more of a patchwork by individuals, or their early generation hybrids, in geographically adjacent freshwater habitats moving quickly through the marine environment to seed new freshwater populations.
+More generally, how do different levels of gene flow and local adaptation interact with the asymmetric nature of stickleback population structure affect the dynamics of the origin and rate of adaptive alleles, as well as their eventual organization into marine and freshwater typical genomes?
+
+Here we develop a forward simulation approach in SLiM to model the stickleback population structure in marine and freshwater habitats to document the effect of variation in gene flow and recombination on the genetic and genomic architecture of local adaptation.
+We address how rapidly can selection act on standing genetic variants at a given value of migration, $M$, and what the precise origin is of those variants.
+We ask what scale of historical introgression between populations with selective pressures $X$ and $Y$, allows for rapid and parallel local adaptation of a population derived from $X$ to an introduced environment with selective pressure $Y$?
 How are the variants structured, historically and across the genome underlying the trait?
-%How does population genetics data influenced by historical introgression and selection on standing genetic variants?
-%How many alleles underly a certain trait and how many are swept away by drift? 
 Furthermore, how can we infer causal loci for regions of the genome that must be driving the rapid adaptation, from real data.
-Many biologists today make use of genome wide association studies (GWAS) and 
-$F_{st}$ across the genome to estimate regions responsible for certain traits. 
+Many biologists today make use of genome wide association studies (GWAS) and $F_{st}$ across the genome to estimate regions responsible for certain traits.
 Unfortunately, the data can be heavily influenced by the level of introgression and population structure of the samples.
@@
 
-There are many different parameters that carry a large impact on the questions above.
-\citet{ralph2015convergent} 
-explored the dichotomy of selection on new alleles and those brought in by migrants from a similar selective pressure.
-Answering the fundamental questions surrounding spatial resolution of convergent evolution. 
-%This lead to an understanding of the process by which alleles move from an existing 'patch' to an introduced 'patch' with the 
-%same selective pressure. 
-\jgg{It seems like this could be a good parallel to what we're doing here, or maybe for discussion?}
-
-In this paper, we explore time to local adaptation as a function of migration rate through an environment (marine) with opposing selective pressure. 
-Next, we compare to real data and explore how $F_{st}$ data can be impacted by introgression.
-%Explain what we do and what we find that is important
-
-%We use forward moving simulations to model marine and freshwater populations of stickleback.
-%We then observe the effect (focusing on migration and recombination, for now) 
-%of different parameters on local adaptation of introduced freshwater environments.
+We find that ....
+\textbf{1.
+rewrite the previous paragraph to be more concise and tight, and 2.
+briefly restate our major findings which should line up well with those presented in the abstract.
+The following I grabbed from the beginning of the Results section}
+*****
+We varied the migration rate, $m$, from $5 \times 10^{-5}$ to $5 \times 10^{-2}$ per individual, i.e., between 0.01 and 10 migrants per year to and from each of the ten lakes.
+This had a strong effect on many aspects of adaptation, including how fast adaptation occurred in each lake, how much alleles were shared between lakes,
+and the population genetic signals left behind.
+At the lowest migration rate, lakes adapted nearly independently, while at the highest migration rate, the habitats are beginning to act like a single metapopulation with substantial migration load.
+We will now dissect what happens between these extremes.
+Interestingly, the ability for separate populations to locally adapt to their own selective pressure was relatively unaffected until the highest rate of migration between the marine and freshwater environments.
+All rates of historical introgression aside from the lowest, helped both the initial and introduced freshwater populations share pre-existing freshwater adapted alleles.
+The sharing of alleles resulted in rapid adaptation ($\approx 100$ generations) of the introduced population (split from the marine population) to adapt to the freshwater environment.
+We also found that larger amounts of migration allow for more statistical power and less false positives in the resulting population genetics data ($F_{st}$ per SNP across the genome).
+****
 
 %%%%%%%%%%%%%%%%
-\section*{Methods}
+\section{Methods}
 
-We explored these questions
-using forwards-time simulations with explicit genomic representation of a quantitative trait
-in SLiM \citep{haller2017slim}.
-The details (described below) are motivated by previous population genomic studies of threespine stickleback,
-but remain simplistic in some aspects due to computational constraints.
-\jgg{should we brag about treeSeq here a little?}
-Possibly the most important caveat is that simulated population sizes are much smaller
-than the census size of the Threespine Stickleback population;
-we will consider the likely effects of this in the Discussion.
-\plr{TODO}
+We explored these questions using forwards-time simulations with explicit genomic representation of a quantitative trait in SLiM \citep{haller2017slim}.
+The details (described below) are motivated by previous population genomic studies of threespine stickleback, but remain simplistic in some aspects due to computational constraints.
+Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback population;
+we will consider the likely effects of this in the Discussion. \plr{TODO}
 
 \paragraph{Habitat and geography}
-Our model has two habitat types, which are considered separate populations
-with a fixed total size of 5,000 diploid individuals each. The marine population is treated as a single large deme, but the freshwater population is subdivided into local subpopulations or demes.
-The scheme, depicted in Figure \ref{fig:Geo},
-roughly models a set of freshwater habitats
-along a stretch of coastline.
-The ``marine'' habitat  is a continuous, one-dimensional range of 25 units of length,
-and the ``freshwater'' habitat is divided into ten subpopulations
-(which we call ``lakes''), each connected to the marine habitat
-at positions $i - 1/2$ for $1 \le i \le 25$.
+Our model has two habitat types, which are considered separate populations with a fixed total size of 2,000 diploid individuals each.
+The marine population is treated as a single large deme, but the freshwater population is subdivided into local subpopulations or demes.
+The scheme, depicted in Figure \ref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline.
+The ``marine'' habitat  is a continuous, one-dimensional range of 10 units of length, and the ``freshwater'' habitat is divided into ten subpopulations (which we call ``lakes''), each connected to the marine habitat at positions $i - 1/2$ for $1 \le i \le 10$.
 
-Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff such that alternative values of the trait have high or low fitness in the two different habitats.
-This situation roughly models the cumulative effect of the various phenotypes
-such as armor morphology, body size, craniofacial variation and opercle shape (among others)
-on which divergent selection acts in the two environments.
-The optimal trait value in the marine habitat is $+10$,
-and in the freshwater habitat is $-10$. Fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$ is
-Gaussian with standard deviation, i.e.,
+Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff such that alternative values of the trait have high or low fitness in tht two different habitats.
+This situation roughly models the cumulative effect of the various phenotypes such as armor morphology, body size, craniofacial variation and opercle shape (among others) on which divergent selection acts in the two environments.
+The optimal trait value in the marine habitat is $+10$, and in the freshwater habitat is $-10$.
+Fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$ is Gaussian with standard deviation, i.e.,
+
 \begin{align*}
@@
 	\begin{center}
-  		\includegraphics[width=0.6\linewidth]{GeographyFigure.pdf}
-  		\caption{
-			A representation of the geographic and evolutionary history of all populations throughout the simulation. 
-		The marine is a one-dimensional, continuous population with spatial positions ranging from [0.0 - 25.0]. 
-		Each freshwater lake$_{i}$ for both the introduced and freshwater populations
-		is a discrete population connected by migration only through the marine, at position $i - 0.5$. 
-		Each of the introduced lakes have the same spatial dynamic and selective pressure as the pre-existing lakes, but arise at 100K generations into the simulation.
-		The marine has $5000$ diploid individuals while each of the lakes fluctuate  $200$ diploid individuals. 
-		the introduced population is initialized as a copy of all marine individuals to model marine 
-		individuals inhabiting a newly created freshwater environment such as the ponds around Middleton Island.
-		We then observe the selection process of marine individuals and following generations 
-		with the new selective pressure across a range of parameter values. 
+  		\includegraphics[width=0.6\linewidth]{GeographyDiagram}
+  		\caption{A representation of the geographic and evolutionary history of all populations throughout the simulation.
+The marine is a continuous population with spatial positions ranging from [0.0 - 10.0].
+Each freshwater lake$_{i}$ for both the introduced and freshwater populations is a discrete population connected by migration only through the marine, at position $i - 0.5$.
+The introduced lakes have the same spatial location (but separate?) and selective pressure as the original lakes, but arise at 40K generations.
+The marine has $2000$ diploid individuals while each of the lakes fluctuate around $200$.
+The introduced population is initialized as a copy of all marine individuals to model marine individuals inhabiting a newly created freshwater environment such as the ponds around Middleton Island.
+We then observe the selection process of marine individuals and following generations with the new selective pressure across a range of parameter values.
+		%Migration between the marine and original lakes is 
+		%set to $0$ the generation before introduction to avoid direct migrants from original $->$ introduced lakes.
 		}
@@
 
-\paragraph{Potential genomic architecture}\bill{Initial genomic architecture? The header is a little confusing to me}
-Each individual carries two copies of a linear chromosome of $10^8$ loci representing a realistic size genome.
-New mutations occur with probability $10^{-10}$ per locus per generation in regions affecting phenotype and $10^{-8}$ in neutral regions.
-Ten ``effect regions'' of $100 Kb$  loci each are spread evenly along the chromosome; 
-mutations in these regions have the potential to affect trait value of the individual which posses them
-Each mutation in these regions is either additive, completely recessive,
-or completely dominant (with equal probability).
-Effect sizes are chosen randomly from an exponential distribution with mean $1/2$,
-either positive or negative with equal probability.
-Between the effect regions is $9.9 Mb$ of neutral loci.
-Individual trait values are determined additively from the diploid genotypes.
-Concretely, an individual that is heterozygous and homozygous for mutations
-at loci $H$ and $D$ respectively has trait value
-$x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
-where $s_\ell$ is the effect size of the mutation at locus $\ell$.
+\paragraph{Initial genomic architecture} Each individual carries two copies of a linear chromosome of $10^5$ loci (which could be thought of roughly as a kilobase each).
+New mutations occur with probability $10^{-7}$ per locus per generation.
+Ten ``effect regions'' of 100 loci each are spread evenly along the chromosome (separated by 4950 loci) in which mutations affect the trait; mutations in other parts of the genome are neutral.
+Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability), and with an effect sizes chosen randomly from an Exponential distribution with mean $1/2$, either positive or negative with equal probability.
+Individual trait value are determined additively from the diploid genotypes.
+Concretely, an individual that is heterozygous and homozygous for mutations at loci $H$ and $D$ respectively has trait value $x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$, where $s_\ell$ is the effect size of the mutation at locus $\ell$.
 Subsequent mutations at the same locus replace the previous allele.
 
-\paragraph{Life cycle}
-Each generation, the two parents of each new offspring are chosen 
-proportional to their fitness (all individuals are hermaphroditic),
-and the contributing genomes are produced by Poisson recombination
-with an average of one crossover per chromosome per generation
-($10^{-8}$ per locus per generation).
-The model is Wright-Fisher, so that each generation, in each habitat,
-5000 new individuals are produced,
-and so to keep population sizes roughly constant within each lake,
-at the start of each generation we rescale fitness values in the freshwater habitat
-so that the sum of the fitness values within each lake are equal.
+\paragraph{Equilibrium in initial marine and freshwater populations} Each generation, the two parents of each new offspring are chosen proportional to their fitness (all individuals are hermaphroditic),and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-7}$ per locus per generation).
+The model is Wright-Fisher, so that each generation, in each habitat, 2000 new individuals are produced, and so to keep population sizes roughly constant within each lake, at the start of each generation we rescale fitness values in the freshwater habitat so that the sum of the fitness values within each lake are equal.
 
-Migration occurs both locally along the coastline in the marine habitat
-and between the marine habitat and the lakes,
-and can be thought of as occurring at the juvenile stage.
-The between-habitat migration rate is denoted $m$,
-and will be called simply the ``gene flow''.
-$\approx m$ individuals \emph{per} lake \emph{per} generation, have parents from the opposing selective pressure.
-A first parent is chosen proportional to fitness,
-and a mate is chosen from the same lake as the first, also proportional to fitness.
-The resulting offspring is given a spatial location in the marine habitat
-at the location of the parent's lake.
-Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$):
-first, a single parent is chosen proportionally to fitness in the marine habitat,
-and then a mate is chosen, 
-also proportionally to fitness but reweighed by a Gaussian function 
-of the distance separating the two, with standard deviation $1/2$. 
-More specifically, if the first parent is marine individual $i$,
-then marine individual $j$ is chosen as the mate
-with probability proportional to $f(x_i) \exp(-2d_{ij}^2)$,
+Migration occurs both locally along the coastline in the marine habitat and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage.
+The between-habitat migration rate is denoted $m$, and will be called simply the ``migration rate''.
+Each new individual in the marine habitat has freshwater parents with probability $m$; to obtain the pair, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness.
+The resulting offspring is given a spatial location in the marine habitat at the location of the parent's lake.
+Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$): first, a single parent is chosen proportionally to fitness in the marine habitat,
+and then a mate is chosen,  also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with stanadard deviation $1/2$.
+
+Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_i) \exp(-2d_{ij}^2)$,
 where $d_{ij}$ is the distance between the two locations.
-Finally, each new marine offspring is given a position 
-displaced from the first parent's position by a random Gaussian distance
-with mean 0 and standard deviation 0.02, and reflected to stay within the population.
-New offspring in the freshwater habitat are chosen in the same way,
-except the probability that the parents are marine individuals is $m$;
-any new freshwater offspring produced by marine individuals
-are assigned to the lake nearest to the position of the first marine parent.
+Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.02, and reflected to stay within the population.
+New offspring in the freshwater habitat are chosen in the same way, except the probability that the parents are marine individuals is $m$; any new freshwater offpsring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.
 
-\paragraph{New lakes}
-To study adaptation in newly appearing freshwater habitats,
-we introduce a new set of lakes midway through the simulation.
-As with the old lakes, these are a set of discrete demes connected by migration \emph{only} to the marine.
-The initial generation of individuals in these new lakes were created as a copy of the marine individuals 
-at the generation of introduction.
-The new lake's geography acts like an independent copy of the original set of lakes -- in particular,
-the two sets of lakes each have 5,000 individuals at all times 
-(Gene flow constricted to within the lakes and through migration to the marine environment).
-Since this doubles the number of lake-to-marine immigrants,
-after this happens the probability that a new marine individual has freshwater parents
-is $2m$ instead of $m$.
+\paragraph{Population genomic analyses of new freshwater populations} To study adaptation in newly appearing freshwater habitats, we introduce a new set of lakes midway through the simulation.
+The initial set of individuals have parents chosen as above from the marine habitat, and act like an independent copy of the original set of lakes -- in particular, the two sets of lakes each have 2,000 individuals at all times.
+Since this doubles the number of lake-to-marine immigrants, after this happens the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.
 
-\paragraph{TreeSeq}
-This case study involves analyzing large simulations which model populations under the influence of selection. 
-To fully understand the origins of freshwater alleles, calculate statistics, and avoid the computationally expensive task of 
-simulating neutral mutations, we have used SLiM's \emph{treeSeq} feature to track the genealogical history of all samples throughout the simulation. 
-This feature outputs a compact representation the genomic outcome of every meiosis relevant to tracked samples (Often thought of as the ARG).
-
-Using the output tree sequence, 
-we were able to classify each freshwater allele in each genome of the new lakes population, at time of adaptation,
-as deriving from one of four origins:
-(1) Migrants, brought in from a migrant after the introduction of the new lakes 
-(2) Pre-existing, passed down from an individual in the original generation of the new lakes, 
-however, not defined to be an adaptive variant in the original lakes
-(3) Captured, Pre-existing, passed down from an individual in the original generation of the new lakes, 
-and defined to be adaptive in the original lakes
-(4) De novo, mutated and propagated from an individual existing in the introduced lakes population, after introduction. 
-\jgg{We should discuss further what these categories aught to be named}
-
-\begin{comment}
-We also used the tree sequence to count the "fitness" of individuals in the initial generation of the new lakes. 
-The "fitness" of an individual in this case, is defined to be the total number off eventual offspring, in the generation at the time of adaptation, 
-which have inherited a adaptive variant from that individual. 
-To calculate this, we only considered gene trees which overlapped an adaptive variant site (defined at time of adaptation).
-We then counted the total number of offspring which had inherited from any adaptive variant site, 
-for each individual in the initial generation. 
-\end{comment}
-
-To determine the ability for the new lakes to select upon the standing variation in the marine at different rates of historical introgression,
-We were able compute the trait value of all effect regions in the initial population.
-Given the probability that effect region fixes at a given trait value, $2 * x / 45$, where $x$ is the effect size of a haplotype,
-we determined the expected total effect sizes of the haplotypes that fix.
-Concretely, it is the sum across the 10 effect regions of
-$$\sum_{i=1}^N \prod_{j < i} (1 - p_j) p_i x_i$$
-where $N$ is the number of genomes per lake, $p_i$ is the probability of fixation of the 
-$i^{th}$ haplotype, $x_{i}$ is the effect size of the $i^{th}$ haplotype, and the haplotypes are sorted in decreasing order by $x_{i}$.
-These numbers were computed assuming additivity of mutations, meaning the numbers produced are a slight overestimate.
-
-\paragraph{Descriptive statistics}\bill{Population genetic analyses? Instead of the more general 'Descriptive statistics'?}
-To assess whether new lakes adapt using existing genetic diversity,
-we define a \emph{freshwater allele}
-to be an effect mutation that has frequency higher than $0.5$ in at least one of the original lakes,
-while remaining lower than $0.5$ in the marine. 
-This categorization is made for each generation using the allele frequencies from that generation,
-and so changes with time.
+To assess whether new lakes adapt using existing genetic diversity, we define a \emph{pre-existing freshwater adapted allele}
+to be an effect mutation that has frequency higher than $0.5$ in at least one of the original lakes, while remaining lower than $0.5$ in the marine.
+This categorization is made for each generation using the allele frequencies from that generation, and so changes with time.
 Alleles common in the newly introduced lakes do not count
 if they are not also common in the original lakes.
-They are defined this way because the transportation hypothesis 
-does not specify where or when an advantageous mutation arises,
-but simply suggests that any sufficiently common freshwater adapted allele 
-could participate in adaptation in new habitat \citep{schluter2009genetics}.
+They are defined this way because the transportation hypothesis does not specify where or when an advantageous mutation arises, but simply suggests that any sufficiently common freshwater adapted allele could participate in adaptaiton in new habitat \citep{schluter2009genetics}.
 
-\emph{Time to adaptation} of the introduced population, denoted $T_\text{adapt}$, 
-is defined to be the generation at which 
-the difference between the average trait value in the original and the introduced freshwater populations 
-is less than 0.5.
+\emph{Time to adaptation} of the introduced population, denoted $T_\text{adapt}$, is defined to be the generation at which the difference between the average trait value in the original and the introduced freshwater populations is less than 0.5.
+We describe overall genetic differentiation between the habitats using $F_{ST}$,
+calculated on a per-locus basis. Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively,
+and $\bar p = (p_f + p_m)/2$, then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
 
-We describe overall genetic differentiation between the habitats using $F_{ST}$,
-calculated on a per-locus basis.
-Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele
-in the freshwater and marine habitats, respectively,
-and $\bar p = (p_f + p_m)/2$,
-then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
-
-\section*{Results}
-
-We varied the migration rate, $m$ across separate simulations
-from $5 \times 10^{-5}$ to $5 \times 10^{-1}$ by steps one order of magnitude.
-From here on out we will refer to these rates of gene flow as running simulations ranging from {0.01 - 100.0}
-migrants \emph{per} lake \emph{per} generation.
-%i.e., between 0.01 and 100 migrants per generation to and from each of the freshwater lakes.
-This had a strong effect on many aspects of adaptation,
-including how fast adaptation occurred in each lake,
-how many alleles were shared between lakes,
-and the population genetic signals left behind.
-We find a threshold of gene flow which allows 
-the freshwater haplotype to be maintained in the marine without creating linkage to marine alleles. 
-below this threshold, selection on pre-existing freshwater alleles to be much less efficient 
-due to background selection. \jgg{did I get this right?}
-
-Interestingly, all rates of $m$ aside from the two highest had little affect on populations ability to locally adapt.
-All rates of gene flow aside from the lowest, helped both the old lake's and the new lake's freshwater populations share pre-existing freshwater adapted alleles.
-The sharing of alleles at $m = 1.0$ resulted in rapid adaptation ($\approx 32$ generations) of the introduced population to adapt to the freshwater environment.
-At the lowest migration rate, lakes adapted nearly independently,
-while at the highest migration rate, the habitats act like a single metapopulation
-with substantial migration load.
-We will now dissect what happens between these extremes.
+%%%%%%%%%%%%%%%%%
+\section{Results}
 
@@
 
+% The progression of adaptation. How you can see introduction new alleles
+
+Local adaptation occurred in all simulations: as shown in Figure \ref{fig:phenotype_ts2}, freshwater and marine populations diverged until the trait means were close to the optimal values in each habitat.
+(Remaining simulations are shown in Figures \ref{fig:phenotype_ts1} and \ref{fig:phenotype_ts4}.) Greater migration decreased adaptedness only weakly except for at the highest migration rate (10 migrants per generation per lake), where the mean trait value equilibrated at roughly half its optimal value (Figures \ref{fig:MeanPhenotype} and \ref{fig:phenotype_ts4}).
+The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
+as jumps in the mean trait value; in the two simulations these occur on a time scale of 100 (lower migration) to 1000 (higher migration) generations, and move the trait by an amount of order 1.
+Trait variation within each population was small compared to the difference between populations, with interquartile ranges of around XXX.\plr{TODO} Across all parameter values, differences at around 16 commonly polymorphic sites (eight the shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats.
+
+\plr{ALSO TODO: CHECK THIS}
+
 \begin{figure}
 	\begin{center}
-        \includegraphics{Final_Plots/Pheno_Time.pdf}
+        \includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4}
+        \includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3}
   		\caption{ 
-        		Mean individual trait values in the marine habitat (blue line),
-        		the original lakes (light green lines; average in dark green),
-        		and the introduced lakes (yellow lines; average in orange)
+        Mean individual trait values in the marine habitat (black line),
+        the original lakes (light green lines; average in pink),
+        and the introduced lakes (dark green lines; average in orange)
         across the course of two simulations, with migration rates of
-        		\textbf{(top)} $m=5 \times 10^{-5}$ and
-        		\textbf{(bottom)} $m=5 \times 10^{-4}$ per generation individual
-        		(i.e., 0.01 and 0.1 migrants per lake per generation, respectively).
+        \textbf{(top)} $m=5 \times 10^{-4}$ and
+        \textbf{(bottom)} $m=5 \times 10^{-3}$ per generation individual
+        (i.e., 0.1 and 1.0 migrants per generation per lake, respectively).
         Optimal trait values in the two habitats are at $\pm 10$.
@@
 
-Local adaptation occurred in all simulations:
-as shown in Figure \ref{fig:phenotype_ts2},
-freshwater and marine populations diverged
-until the trait means were close to the optimal values in each habitat. 
-The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
-as jumps in the mean trait value;
-in the two simulations these occur on a time scale of 100 (lower migration) to 1000 (higher migration) generations,
-\jgg{I'm not sure what this means? how do we know how long the jumps take?}
-and move the trait by an amount of order 1.
-Trait variation within each population was small compared to the difference between populations,
-with interquartile ranges of around XXX.\plr{TODO}
-Across all parameter values, differences at around 16 commonly polymorphic sites 
-(eight the shift the trait in each direction)
-were responsible for most of the adaptive differences between freshwater and marine habitats.
+\jgg{also, interesting point to make on second plot, you can see the initial lakes taking the same size step, showing sharing of alleles.
+could zoom in on the relevent portions of the figure for this.
+(begging and introduction) a 2 by 2 with the important sections (0 - 20K) or (40-50K) for both above figures?}
+
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MeanPhenotype1.png}
+  		\caption{
+            Distribution of mean trait values throughout each simulation.
+            The dashed pink and purple lines show the optima ($\pm 10$).
+		}
+  		\label{fig:MeanPhenotype}
+	\end{center}
+\end{figure}
+\plr{Is this the whole simulation? what about the initial adaptation phase?
+     Is the horizontal line at the median? what are the whiskers at?}
 
 As expected, increasing migration rate decreased differentiation between habitats.
-As seen in Figure \ref{fig:Fst},
-$F_{ST}$ between marine and freshwater habitats at neutral sites 
-steadily declines as migration increases.
-Local adaptation was still able to occur despite overall homogenization:
-if computed using only sites with alleles affecting the trait (``effect mutations''),
-$F_{ST}$ between habitats was relatively constant across migration values.
+As seen in Figure \ref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.
+Local adaptation was still able to occur despite overall homogenization: if computed using only sites with alleles affecting the trait (``effect mutations''), $F_{ST}$ between habitats was relatively constant across migration values.
 
@@
 	\begin{center}
-  		\includegraphics{Final_Plots/Fst_Genome.pdf}
+  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/FST_HRR.png}
   		\caption{
-		$F_{st}$ as a function of genomic location between all combinations of the three subpopulations; 
-		the old lakes, the new lakes, and the marine.
-		$F_{st}$ was calculated based upon allele frequencies in genomic windows of 500 mb. 
-		The four columns of subplots are separate simulations with increasing $m$. 
-		The three rows are the combinations of the three separate populations.
-		Ten vertical dotted pink lines in each subplot represent regions which have the potential to affect phenotype
+            Distributions of $F_{st}$ values between the three habitats,
+            accumulated across the course of the simulation, calculated 
+            \textbf{(left)} using only neutral mutations, and
+            \textbf{(right)} using effect mutations.
         }
@@
 
-\paragraph*{Speed of adaptation}
-Increasing migration rate strongly decreased the time until freshwater habitats could adapt,
-both in the old and new sets of lakes.
-The different rates of introduction of effect alleles is clearly seen in the traces
-of trait values against time (e.g., Figure \ref{fig:phenotype_ts2}).
-As shown in Figure \ref{fig:TimeToAdaptation}, at the lowest migration rate 
-it took over $18,000$ generations for the mean trait value across introduced lakes to 
-get to within 0.5 of the original lakes value.
-At $m = 1.0$, we saw the introduced population adapt in just $32$ generations. 
-This rate of adaptation follows what we've observed in natural populations of threespine stickleback.
-At the highest migration rate, $m = 100$, all individuals were unable to locally adapt.
+\paragraph*{Rate of adaptation}
+Increasing migration rate strongly decreased the time until freshwater habitats could adapt, both in the initial and introduced sets of lakes.
+As shown in Figure \ref{fig:TimeToAdaptation}, at the lowest migration rate it took over 20 thousand generations for the mean trait value across introduced lakes to get to within 0.5 of the original lakes value.
+Although many lakes had adapted before this time, the different rates of introduction of effect alleles is clearly seen in the traces of trait values against time (e.g., Figure \ref{fig:phenotype_ts2}).
+However, higher migration rates allowed the freshwater habitat to adapt nearly as quickly as the marine habitat.
 
@@
 	\begin{center}
-  		\includegraphics{Final_Plots/Time_Adapt.pdf}
+        \includegraphics[width=0.8\linewidth]{matplotlibPlots/TimeToAdaptation}
         \caption{
-		Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
-		it takes for the introduced population's mean phenotype to come within 0.5 of the original lakes average phenotype. 
-		Each point represents a simulation run at some value of $M$. 
-		The yellow dashed line is the average of all points at each respective parameter value}
+            Time to adaptation as a function of migration rate ($m$),
+            measured as the number of generations
+            until the introduced population's mean phenotype is within 0.5 
+            of the original lakes average phenotype. 
+            Each point represents a single simulation, and
+            the line shows the average of all points at each value of $m$.
+        }
   		\label{fig:TimeToAdaptation}
@@
 
-\subsection*{Sharing of freshwater alleles}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Sharing of freshwater adapted alleles}
 
-We find that this dramatic increase in speed of local adaptation occurs because
-higher gene between populations allow sharing of freshwater alleles more effectively when compared to de novo mutation.
-At low migration rates, adaptation occurs independently in each lake.
-In Figure \ref{fig:Origin}, at $0.01$, we can see that the large majority of adaptive alleles derived from de novo mutation in the 
-population rather than selection on standing variants in the ocean. 
-As The migration increases, we see a larger ratio of the origins of adaptive alleles 
-derive from pre-existing variation in the marine population at the time of introduction.
-In other words, 
-greater mixing at higher migration rates allows lakes to share alleles
-instead of developing their own genetic basis of adaptation.
-To further investigate sharing of locally adaptive alleles and the transporter hypothesis,
-we investigate the ``freshwater alleles'', defined for a particular generation
-to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat in that generation.
+%FIGURES THAT SHOW THIS:
+%-MPAA / IND
+%-Total/Avg shared
 
-As the number of common, locally adaptive alleles decreases, as seen in Figure \ref{fig:MPFAI}C,
-the genetic basis of adaptation is more commonly shared.
-Figure \ref{fig:MPFAI}B shows
-the percentage of currently-defined, freshwater-adapted alleles 
-that each genome in each of the populations carries,
-averaged across time and individuals.
-If all individuals across lakes carried the same set of alleles determining their trait value,
-this would be 100\%.
-This value is nearly reached at the fourth migration rate;
-but it is lower due to migration load.
-At the lowest migration rate, 
-each genome in the original lakes have almost exactly $1/25^{th} = 0.04\%$   
-of the total freshwater adapted alleles --
-since there are 25 lakes, this implies that each lake has adapted with a unique set of alleles.
+The tenfold difference in speed of adaptation occurs because at low migration rates, adaptation occurs independently in each lake,and the marine habitat has ten times the influx of new alleles than any one lake.
+In other words, greater mixing at higher migration rates allows lakes to share alleles instead of developing their own genetic basis of adaptation.
+As a first indication of this, Figure \ref{fig:Fst} shows that $F_{ST}$ between the ``original'' and ``introduced'' sets of lakes at effect mutations decreased with migration rate.
+
+To further investigate sharing of locally adaptive alleles and the transporter hypothesis, we investigate the ``pre-existing freshwater adapted alleles'',
+defined for a particular generation to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat in that generation.
+Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles, across generations, as well as the average number of lakes each is found in.
+As migration rates increase, the number of these alleles decreases steadily, and each is concurrently found in a greater number of lakes.
+
+As the number of common, locally adaptive alleles decreases, the genetic basis of adaptation is more commonly shared.
+Figure \ref{fig:MPFAI}B shows the distribution of the mean percentage of currently-defined freshwater adapted alleles that each genome in each of the populations carries, averaged across time and individuals.
+If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%.
+This value is nearly reached at the highest migration rate; but it is lower due to migration load.
+At the lowest migration rate, each genome in the original lakes have almost exactly $1/10^{th}$ of the total freshwater adapted alleles -- since there are 10 lakes, this implies that each lake has adapted with a unique set of alleles.
 Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
-Figure \ref{fig:MPFAI}A shows us that by $0.1$ migrants per lake per generation,
-the average individual across the new lakes has nearly the same amount of 
-``pre-existing freshwater adapted alleles" as individuals across the old lakes.
-Interestingly, the genetic basis of the freshwater phenotype 
-seems to simplify with greater $m$. 
-This would seem to suggest higher rates of migration allowed 
-adaptive alleles of higher effect to travel more efficiently through the populations
-\jgg{is this right?}
 
@@
 	\begin{center}
-  		\includegraphics{Final_Plots/Freshwater_Alleles.pdf}
+        \includegraphics[width=\linewidth]{plotlyPlots/NumFAA}
+        \includegraphics[width=\linewidth]{matplotlibPlots/MPFAI}
   		\caption{
-		Distribution of freshwater alleles among the three populations. 
-		A freshwater adapted variant is defined as, at any generation throughout the simulation,
-		a variant affecting phenotype that is greater than 0.5 frequency in any of the old lakes, 
-		and less than 0.5 frequency in the marine population. 
-		Subplot $A$ is the distribution throughout the simulation, of the mean number of
-		freshwater adapted variants per individual, in each population.
-		Subplot $B$ is the distribution throughout the simulation, of the mean \emph{percentage} of
-		freshwater adapted variants per individual, in each population. 
-		Subplot $C$ is the distribution throughout the simulation, of the total number of
-		freshwater adapted alleles defined at any one generation.
-		In other words, $B  =  A  /  C$.
-		}
+            \textbf{(A)}
+            Distribution of the number of freshwater adapted alleles (green)
+            and the average number of lakes each allele is present in (red)
+            across the course of the simulation.
+            \textbf{(B)}
+            Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
+		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
+		freshwater adapted alleles.
+		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
+		while the percentage tells us the variance of the haplotype}
 		\label{fig:MPFAI}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Migration load and genetic variation}
 
-At first, increased migration allows sharing of adaptive alleles between lakes,
-but at $m = 5 \times 10^{-2}$
-the constant influx of alleles between the habitats creates noticeable migration load.
-This rate, $m=.05$, only replaces 5\% of each population each generation
-with migrants from the other habitat, but this is sufficient to shift the mean trait values
-to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
-Significant gene flow constricts local adaptation
-as a consequence of a large number of offspring through hybridization events between subpopulations.
-at $m=0.5$, all populations experience panmixia
+%FIGURES THAT SHOW THIS:
+%- Phenotype Distribution
+%GENERAL
+
+At first, increased migration allows sharing of adaptive alleles between lakes, but at the highest migration rate, the constant influx of alleles between the habitats creates substantial migration load.
+This rate, $m=.05$, only replaces 5\% of each population each generation with migrants from the other habitat, but this is sufficient to shift the mean trait values to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
+Significant gene flow constricts local adaptation as a consequence of a large number of offspring through hybridization events between subpopulations.
+
+\paragraph{Standing genetic variation} Despite a dramatic difference in the amount of allelic sharing between lakes, standing genetic variance (Figure \ref{fig:SGV}) was around 0.05 in the freshwater habitat across all but the highest migration rates.
+Concurrently, genetic variation in the marine population steady increased with migration
+to a similar value.
+On the one hand, it is not surprising that lakes, as a group, carry more genetic variation than the marine habitat, since population subdivision allows different alleles to become common in each.
+However, it is striking that at $m=.005$, the marine habitat carries as much genetic variation, despite a lack of any substantial migration load.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Realized genomic architecture}
 
@@
 	\begin{center}
-  		\includegraphics{Final_Plots/Pheno_Dist.pdf}
-  		\caption{Distribution of mean trait value throughout simulation runs at separate migration rate $(m)$ parameter values, for each population.
-		The dashed pink line (Trait value = +10) is the optimum phenotype any individual in the marine environment.
-		In contrast the purple line at (Trait value = -10) represents the optimum for any individual in the freshwater environment. 
-		All individuals at generation 0 (beginning of the simulation) 
+        \includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3}
+        \includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-4}
+  		\caption{(Description and cleaning needed)
 		}
-  		\label{fig:MeanPhenotype}
+  		\label{fig:Fst3}
 	\end{center}
@@
 
-\subsection*{Realized genomic architecture}
+Now we take a closer look at the genomic architecture of local adaptation between the two habitats. Do the alleles underlying trait differences cluster along the genome?
+Do measures of local differentiation identify the causal loci?
+Figure \ref{fig:Fst3} shows plots along the genome of per-locus $F_{ST}$ values between the marine and original freshwater habitats at the two intermediate migration rates.
+(Note that we are pooling freshwater habitats; a single lake would provide substantially less power.)
 
-Higher migration rates showed more distinct $F_{ST}$ peaks
-over polymorphic loci underlying trait differences between the habitats.
-As migration rate decreased, the ``background'' levels of $F_{ST}$ increased,
-swamping out this signal until the regions under selection were indistinguishable. 
-This is likely due to two reasons: 
-first, stronger genetic drift with less migration leading to higher background $F_{ST}$,
-and second, greater sharing of adaptive alleles providing a shared signal across populations.
-This suggests that genome scans for local adaptation
-based purely on measures of differentiation
-will only be successful given enough migration between habitats.
-With 10 effect regions across the genome, we can see in \ref{fig:Fst} that only about 1 or 2 of the region contain peaks. 
-When looking closely at the peaks we see that each is usually a composition of 7-8 SNPs close in proximity.
-Because some of those mutations are selectively neutral (we can see this because the counts of FAA are not as high as the number in the clusters),
-this is suggestive of hitchhiking alleles along with the SNPs being brought to high frequency by a selective sweeps.
+\plr{need to know where the FAA are in these figures}
 
-Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$?
-Knowing exactly which mutations effect phenotype in our simulations, 
-we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome. 
-In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, 
-we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power. 
-This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.
+Higher migration rates showed more distinct $F_{ST}$ peaks over polymorphic loci underlying trait differences between the habitats.
+As migration rate decreased, the ``background'' levels of $F_{ST}$ increased, swamping out this signal until the regions under selection were indistinguishable.
+This is likely due to two reasons: first, stronger genetic drift with less migration leading to higher background $F_{ST}$, and second, greater sharing of adaptive alleles providing a shared signal across populations.
 
-\subsubsection*{Origin of introduced freshwater adaptive alleles: a conundrum}
-Up to this point, we have quantified rapid and parallel adaptation and found strong evidence that sharing of alleles is (more than likely?) responsible.
-However in Figure \ref{fig:MPFAI}A at $0.1$ \emph{mig/gen/lake}, we see sharing of alleles between populations,
-yet \emph{Time to adaptation} is still fairly slow ($\approx 2,000$ gens).
-This is surprising when considering we see a similar quantity of allele sharing at  $m$ equal to $1.0$ \emph{mig/gen/lake}, 
-but much more rapid adaptation ($\approx 30$ gens). 
-Initially, this 60X speedup of adaptation might seem to be the result of subsequent migration from the marine after the point of freshwater introduction. 
-To investigate this further we have track the origins of all adaptive variants in all new lake extant individuals at the time of adaptation. 
-However, in Figure \ref{fig:Origin}, we find that at both $m = 0.1$ and $m = 1.0$ \emph{migs/gen/lake}, the vast majority of adaptive alleles origins were found to propagate 
-from individuals in the original generation ("Captured" OR "Pre-Existing").
+This suggests that genome scans for local adaptation based purely on measures of differentiation will only be successful given enough migration between habitats.
+To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.
+
+\plr{Labels are wrong on that figure: false pos rate should go down with cutoff.
+     Is it true positive rate?}
 
@@
 	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{Final_Plots/Allele_Origin.pdf}
+  		\includegraphics[width=0.7\linewidth]{matplotlibPlots/Power_FP.png}
   		\caption{ 
-		Origin ratio of adaptive alleles in all genomes at time of local adaptation for different values of $m$
-		For all alleles at a frequency $> 0.5$ in any one of the new lakes individuals at time of adaptation, 
-		we traced the origin of the segment of DNA back in time to classify them as 
-		\textbf{(Blue)} Migrants, brought in from a migrant after the introduction of the new lakes 
-		\textbf{(Orange)} Pre-existing, passed down from an individual in the original generation of the new lakes, 
-		however, not defined to be an adaptive variant in the original lakes
-		\textbf{(Green)} Captured, Pre-existing, passed down from an individual in the original generation of the new lakes, 
-		and defined to be adaptive in the original lakes
-		\textbf{(Red)} De novo, mutated and propagated from an individual existing in the introduced lakes population, after introduction. 
+		%SP - suppose you have an Fst Peak, how likely is it that that region is causal. 
+		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
+		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
+		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
+        TODO: x-axis label ($F_{ST}$).
 		}
-  		\label{fig:Origin}
+  		\label{fig:Power_FP}
 	\end{center}
@@
 
-The alternative explanation is that at a higher migration rate, a higher saturation of marine adapted alleles are linked to freshwater adapted alleles.
-In turn, this linkage acts as a barrier rebuilding the freshwater haplotype, in the new lakes, much more difficult as it much wait for recombination to break the links.
-With more historical introgression between the two selective pressures, the freshwater haplotype would remain more "intact" and much more efficient to rebuild in the lake.
-To confirm this, we decided to look at the trait values of the ten affect regions in each genome of the initial population.
-Given the probability that effect region fixes at a given trait value, $2 * x / 45$ where $x$ is the effect size of a haplotype
-we were able to determine expected total effect sizes of the haplotypes that fix (detail about computation in methods).
-Below, we see the mean of these expected total effect sizes across all lakes.
+With 10 effect regions across the genome, we can see in \ref{fig:Fst3} that only about 1 or 2 of the region contain peaks. 
+%SOFT STATEMENT - MORE (Peak) EVALUATION?
+When looking closely at the peaks we see that each is usually a composition of 7-8 SNPs close in proximity.
+%SOFT STATEMENT -
+Because some of those mutations are selectively neutral (we can see this because the counts of FAA are not as high as the number in the clusters), this is suggestive of hitchhiking alleles along with the SNPs being brought to high frequency by a selective sweeps.
 
-\begin{center}
-\begin{tabular}{ | c | c  | c  | c | }
- \hline
- $m = 0.01$ & $m = 0.1$ & $m = 1.0$ & $m = 10.0$ \\ 
- $0.1227$ &  $4.6976$ & $12.9298$  & $12.2710$\\
- \hline
-\end{tabular}
-\end{center}
+In regions of the genome underlying individual trait value, we observed Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$.
+Knowing exactly which mutations effect phenotype in our simulations, we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome.
+In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power.
+This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.
 
-Here, we see a dramatic difference in the total effect sizes of the haplotypes that are likely to fix.
-This provides evidence that linkage to marine alleles has a major impact in the introduced population's ability 
-to select and rebuild a pre-existing haplotype given the standing variation at the time of introduction. 
-\jgg{Below may belong in the discussion / future work}
-This also brings to light the role of recombination in a system of migration selection balance. 
-On one hand, a lower recombination rate would allow for the freshwater haplotype to remain "intact" 
-within the marine environment, however, without significant introgression the haplotype would quickly get selected against in marine populations. 
-On the other hand a higher recombination rate would, in theory, allow freshwater adapted alleles more longevity 
-as they would hitchhike with marine adapted haplotypes.
+%The lowest migration rate of $10^{-5}$ also has a very low false positive rate across all $F_{st}$ threshold values.
+%\subsubsection*{Low Recombination Causes Clustering}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Theoretical expectations}
@@
 
-Here's some rough calculations to get a sense for what should be going on.
-Everything is done in more detail in OTHER PLACES WE SHOULD CITE.
+Here's some rough calculations to get a sense for what should be going on. Everything is done in more detail in OTHER PLACES WE SHOULD CITE.
 Fisher, Chevin, etc.
@@
 Suppose a new allele enters a lake, either by migration or mutation.
-If, when it is rare but present in $n$ copies,
-it has fitness advantage $s$
--- i.e., the expected number of copies in the next generation is $(1+s)n$ --
-then the probability that it escapes demographic stochasticity to become common in the population
-is approximately $2s$ \citep{fisher,prob_fixation}.
-If the current population all differed from the optimum trait by $z$,
-and the allele has effect size $-u$ in heterozygotes,
-then the fitness advantage of the allele would be
-$s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$,
-where in our parameterization, $\beta = 1 / 450$.
-This tells us two things:
-(1) the rate of adaptation decreases as the population approaches the optimum,
-and (2) larger mutations (in the right direction) are more likely to fix.
+If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ --then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{fisher,prob_fixation}.
+If the current population all differed from the optimum trait by $z$, and the allele has effect size $-u$ in heterozygotes, then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
+This tells us two things: (1) the rate of adaptation decreases as the population approaches the optimum, and (2) larger mutations (in the right direction) are more likely to fix.
 
 \paragraph{New mutations}
-The total rate of appearance of new mutations per lake is $\mu_L = 0.04$,
-which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive
-in either direction.
-This implies that a new additive or dominant effect mutation appears once every 87.5 generations,
-on average.
-Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, 
-and so the probability that a dominant mutation manages to establish in a population differing from
-the optimum by $z$ is roughly
-$\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$,
-and so the rate of establishment of dominant mutations is $\beta z / 87.5$,
-i.e., about one such mutation every $2461/z$ generations.
-The distribution of these successfully established mutations 
-has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2.
-Since additive alleles have half the effect in heterozygotes,
-they have half the probability of establishment.
-During the initial phase of adaptation,
-the populations begin at around distance $z=10$ from the optimum.
-Combining these facts, we expect 
-adaptive alleles to appear through mutation at first on a time scale of 250 generations,
-with the time between local fixation of new alleles increasing as adaptation progresses,
-and each to move the trait by a distance of order 1.
+The total rate of appearance of new mutations per lake is $\mu_L = 0.04$, which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive in either direction.
+This implies that a new additive or dominant effect mutation appears once every 87.5 generations, on average.
+Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, and so the probability that a dominant mutation manages to establish in a population differing from the optimum by $z$ is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$, and so the rate of establishment of dominant mutations is $\beta z / 87.5$, i.e., about one such mutation every $2461/z$ generations.
+The distribution of these successfully established mutations  has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2.
+Since additive alleles have half the effect in heterozygotes, they have half the probability of establishment.
+During the initial phase of adaptation, the populations begin at around distance $z=10$ from the optimum.
+Combining these facts, we expect adaptive alleles to appear through mutation at first on a time scale of 250 generations, with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
 
 \paragraph{Standing variation}
-An allele that moves the trait $z$ units in the freshwater direction in heterozygotes
-has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment
-(which is close to optimal).
-The product of population size and fitness differential in the marine environment
-for a mutation with $z=1$
-is therefore $2Ns = 8.9$, implying that these alleles are strongly selected against
-but might occasionally drift to moderate frequency.
-The average frequency of such an allele in the marine environment at migration-selection equilibrium
-is equal to the total influx of alleles per generation divided by the selective disadvantage,
-which if $M = 2000 m$ is the number of immigrants per generation,
-is $2 M / \beta z^2$. 
-With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$:
-since lakes have 400 genomes, 
-as long as $M \ge 1$, the chances are good that any particular lake-adapted allele
-that is present in all pre-existing lakes
-will appear at least once in the fish that colonize a new lake.
-However, an allele with $z=1$ only has probability of around 1/20 of establishing locally,
-suggesting that we'd need $M \ge 10$ to ensure enough pre-existing genetic variation
-that adaptation would happen entirely using the initial set of colonizers.
-This corresponds to our highest two migration rates,
-as in e.g., Figure \ref{fig:phenotype_ts2}B.
+An allele that moves the trait $z$ units in the freshwater direction in heterozygotes has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment (which is close to optimal).
+The product of population size and fitness differential in the marine environment for a mutation with $z=1$ is therefore $2Ns = 8.9$, implying that these alleles are strongly selected against but might occasionally drift to moderate frequency.
+The average frequency of such an allele in the marine environment at migration-selection equilibrium is equal to the total influx of alleles per generation divided by the selective disadvantage, which if $M = 2000 m$ is the number of immigrants per generation, is $2 M / \beta z^2$.
+With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$: since lakes have 400 genomes, as long as $M \ge 1$, the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes will appear at least once in the fish that colonize a new lake.
+However, an allele with $z=1$ only has probability of around 1/20 of establishing locally, suggesting that we'd need $M \ge 10$ to ensure enough pre-existing genetic variation that adaptation would happen entirely using the initial set of colonizers.
+This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.
 
 \paragraph{Migration}
-If sufficient genetic variation is not present in a new lake initially,
-it must appear by new mutation or migration.
-Since a proportion $m$ of each lake is composed of migrants each generation,
-it takes $1/m$ generations until the genetic variation introduced by migrants
-equals the amount initially present at colonization.
-This implies a dichotomy: either 
-(a) adaptation is possible using variants present at colonization
-or arriving shortly thereafter, or 
-(b) adaptation takes many multiples of $1/m$ generations.
+If sufficient genetic variation is not present in a new lake initially, it must appear by new mutation or migration.
+Since a proportion $m$ of each lake is composed of migrants each generation, it takes $1/m$ generations until the genetic variation introduced by migrants equals the amount initially present at colonization.
+This implies a dichotomy: either (a) adaptation is possible using variants present at colonization or arriving shortly thereafter, or (b) adaptation takes many multiples of $1/m$ generations.
 
-These calculations depend on there being no bottleneck in colonization of the lake;
-if there is a bottleneck, then an additional factor must be added.
-
+These calculations depend on there being no bottleneck in colonization of the lake; if there is a bottleneck, then an additional factor must be added.
 At what point do we expect new mutation to be more important than migration for adaptation?
-By the calculations above, if $M \ge 10$, we expect initial diversity in a lake to be sufficient 
-for adaptation, corresponding to our third-highest migration rate.
+By the calculations above, if $M \ge 10$, we expect initial diversity in a lake to be sufficient  for adaptation, corresponding to our third-highest migration rate.
 If this does not happen, then we expect adaptation to take a multiple of $1/m$ generations;
 with our values, $1/m$ ranges from 20,000 to 20 generations.
-Above we estimated that adaptive alleles due to new mutation fix locally about every 2,000 generations,
-suggesting that at our second-lowest migration rate (where $1/m = 2,000$),
-the two contributions of migration and new mutation are roughly equal.
-This is in fact what we see: in Figure \ref{fig:MPFAI},
-we see that at the second migration rate, alleles start to be shared between lakes,
-while by the third migration rate, they are almost entirely shared.
+Above we estimated that adaptive alleles due to new mutation fix locally about every 2,000 generations, suggesting that at our second-lowest migration rate (where $1/m = 2,000$), the two contributions of migration and new mutation are roughly equal.
+This is in fact what we see: in Figure \ref{fig:MPFAI}, we see that at the second migration rate, alleles start to be shared between lakes, while by the third migration rate, they are almost entirely shared.
 
 %%%%%%%%%%%%%%%%%%%%
-\section*{Discussion}
+\section{Discussion}
 
-\jgg{TODO: effect of smaller than realistic population sizes?}
+We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island.
+Selection is able to rebuild the freshwater haplotype from marine populations as a medium between all freshwater populations.
+Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.
 
-\jgg{TODO: Talk about the role recombination plays}
+We also have also explored the genomic architecture as a consequence of the nature of selection in our scenario.
+The alleles underlying individual trait value were of large effect and a low number.
+With a total of $10^{5}$ loci, to be realistic, each loci should represent 1000 $Kb$ in real data.
 
-We have shown that historical introgression, at our given parameter sets,
-is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island.
-Selection is able to rebuild the freshwater haplotype from marine populations as a medium between all freshwater populations.
-Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load 
-limited the ability of the populations to reach the local optimum.
+\jgg{This should be expanded upon, not sure what this means in terms of the genomic architecture}
 
 We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome.
-This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
-differences between populations that have little gene flow between them.
-It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
-mistaken for being causative as they often display the same amount of divergence.
+This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more differences between populations that have little gene flow between them.
+It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be mistaken for being causative as they often display the same amount of divergence.
 %wouldn't it be cool if we could use this fact and simulate introgression, starting with real data that is noisy to extrapolate the areas we care about?
@@
 We have found that too little migration leads to selection upon new mutations in all subpopulations and lakes alike.
-In contrast, at high migration rates we have seen that migration load limits 
-the ability for species to locally adapt to the selective pressure of their environment.
-This leads us to consider a window of introgression which allows for the transportation
-of FAA's without migration load.
+In contrast, at high migration rates we have seen that migration load limits the ability for species to locally adapt to the selective pressure of their environment.
+This leads us to consider a window of introgression which allows for the transportation of FAA's without migration load.
 
@@
 \appendix
-\section*{Supp.}
+\section{Supp.}
+
 \begin{figure}
 	\begin{center}
-  		\includegraphics{Final_Plots/True_Power.pdf}
-  		\caption{ 
-		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
-		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
-		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
-        		TODO: x-axis label ($F_{ST}$).
-		}
-  		\label{fig:Power_FP}
+  		\includegraphics[width=\linewidth]{plotlyPlots/StandingGeneticVariation.png}
+  		\caption{(PLACEHOLDER)}
+		\label{fig:SGV}
 	\end{center}
@@
 	\begin{center}
-  		\includegraphics[width=1.0\linewidth]{semi_final_plots/Haplo_semi_final.png}
-  		\caption{
-			Phenotypes (\textbf{Top}) (effect size and mutation type) and haplotypes \textbf{(Botton)} as a function of spatial position of the individual 
-			possessing the genome. The phenotype of each haplotype is, by definition, the sum of the 
-			effect sizes of each mutation in the genome. 
+  		\includegraphics[width=\linewidth]{matplotlibPlots/MFAI.png}
+  		\caption{Distributions of mean number of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
+		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
+		freshwater adapted alleles.
+		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
 		}
-  		\label{fig:Haplo_Pheno}
+  		\label{fig:MNFAI}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=\linewidth]{matplotlibPlots/effectRegionCounts.png}
+  		\caption{(PLACEHOLDER)}
+		\label{fig:counts}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-5.png}
+  		\caption{ (PLACEHOLDER)
+		}
+  		\label{fig:Fst1}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-2.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:Fst4}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-5.png}
+  		\caption{ (PLACEHOLDER)
+		}
+  		\label{fig:phenotype_ts1}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h!tb]
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-2.png}
+  		\caption{(PLACEHOLDER)
+		}
+  		\label{fig:phenotype_ts4}
 	\end{center}

### 2018-10-22 Stickleback_Paper.tex

--- 2018-09-24 Stickleback_Paper.tex
@@
 
-\textbf{Jargon}
-
-effect loci? regions underlying trait value?
-
-names of four allele origin categories
-
 \section*{Abstract}
 
-Threespine stickleback fish provide one of the most striking examples of local adaptation.
-This hemisphere-wide metapopulation includes both marine populations and a large number of smaller freshwater populations
-that have apparently repeatedly adapted to freshwater conditions
-often using the same genetic basis.
-In this paper, we use simulations motivated by stickleback populations
-to examine what amounts of gene flow favor stable metapopulation polymorphism
-with allele sharing,
-and to further dissect the underlying dynamics.
-We find that rapid, repeated adaptation 
-using alleles maintained at low frequency by migration-selection balance
-(the ``transporter hypothesis'')
-occurs over a realistic range of intermediate rates of gene flow,
-between slow, independent adaptation with low gene flow
-and large migration load at high gene flow.
-Rapid adaptation occurs when the freshwater haplotype has been maintained,
-through historical introgression, in marine environments without tight linkage to 
-marine adapted alleles.
-We find $F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow.
-The results support existing theory of local adaptation,
-and provide a more concrete look at a particular, empirically motivated example.
+Threespine stickleback fish provide a striking example of local adaptation in the context of recurrent gene flow.
+This Northern hemisphere-wide metapopulation includes both marine populations and a large number of smaller freshwater populations that have repeatedly adapted to freshwater conditions, often by using standing genetic variation.
+Here we use genealogical simulations to determine the levels of gene flow can best match the observed patterns of allele sharing among habitats in stickleback, and to provide a framework for better understanding of the dynamics of gene flow and local adaptation for the maintenance and reuse of standing genetic variation.
+We find that rapid, repeated adaptation using alleles maintained at low frequency by migration-selection balance occurs over a realistic range of intermediate rates of gene flow.
+Slow and independent adaptation occurs with low gene flow, whereas large migration load occurs at high gene flow, mediated mainly by the total downstream influx of alleles \emph{(maybe??)}.
+We do not see evidence for strong effects encouraging genomic clustering of causal alleles or towards particular dominance coefficients.
+$F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow, The results support existing theory of local adaptation, and provide a more concrete look at a particular, empirically motivated example.
+
+\textbf{**We need to be more crystal clear in the abstract about the findings from the paper**}
 
@@
 
-\emph{intro outline:}
-Start off with stickleback: repeated, rapid, parallel adaptation with obvious phenotypes.
-But, most species don't seem to have such obviously structured polymorphism: 
-what's special about stickleback?
-The transporter hypothesis is a mechanism; what's required to make it work?
-Other species adapt rapidly in similar ways:
-e.g., Darwin's finches and island mice;
-do these differ through not having such a big, connected reservoir of genetic variation?
-or by not having such a large total population size of the ``other'' habitat?
-Or, is this situation more common than we know, with less obvious phenotypes?
-Also, can we actually tell what's happening from genetic data?
+The canonical model for the genetics of adaptation, first formulated by Fisher in the early part of the 20th century, involves the sequential fixation of new mutations.
+While it has proven valid in numerous studies in the field and the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures.
+A growing number of studies have identified instances of convergent or parallel adaptive evolution that utilize standing genetic variation.
+What evolutionary processes promote the maintenance of re-use of such genetic polymorphisms?
+How do these processes interact with diverse geographic contexts, as well as variation in genetic processes such as recombination, to change the likelihood of so-called standing genetic variation (SGV).
 
-Recently, multiple instances of similar (parallel) underlying genetic basis of rapid local adaptation has brought about 
-questions surrounding the origins and maintenance of genetic polymorphism.
-A empirical model is the Alaskan populations of freshwater and marine Ninespine Stickleback fish.
-In 1964, The Great Alaskan Earthquake caused an uplift of Middleton island and in turn,
-introduced a group of freshwater ponds around the perimeter of the island. 
-Quickly inhabited by the surrounding marine population of Stickleback,
-\citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be 
-parallel to freshwater stickleback that have been separated for over $13,000$ years.
-In freshwater stickleback, the number of lateral plates are reduced
-and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region.
-These results leave us with questions surrounding the nature of rapid adaptation.
-Does convergent evolution breed it's own solution (haplotype) for every new selective pressure, 
-or can these solutions can be efficiently shared across multiple sub - populations facing similar selective pressures.
+Clues come from population genomic studies in the wild that have recently been powered by advances in sequencing technologies.
+One such organism is the threespine stickleback.
+This small fish has long been used for studies of behavior, ecology and evolution in the wild, and has more recently become a model for understanding the genetic basis of adaptive evolution.
+The ancestral marine form has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere, which have each evolved a similar suite - or 'freshwater syndrome' - of phenotypes despite local differences.
+This adaptive phenotypic evolution to freshwater can occur in mere millenia, in many cases since the end of the last ice age about 13,00 years ago.
+Studies of the fossil record also clearly show that invasion of oceanic stickleback and local adaptation in freshwater has occurred repeatedly over millions of years, often involving very similar phenotypic shifts as seen in contemporary populations.
 
-The leading hypothesis for stickleback is that rather than acting on new mutations, 
-adaptation to freshwater environments is sped up through selection on standing genetic variation (SGV) found in marine populations. 
-\jgg{Bill: maybe you could make this section a little more specific}
-One clear example of this is the gene \textit{eda} 
-which has been shown to regulate the number of lateral plates. 
-While this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently.
-Novel evidence from natural populations has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
-\citet{schluter2009genetics} 
-suggested the ``transporter''-hypothesis.  
-This outlined the flow of freshwater alleles into marine populations through offspring of hybridization events.
-It suggests freshwater haplotypes are distributed though marine individuals and are continually selected upon in introduced freshwater populations.
-The continued selection on freshwater favored alleles and introgression between the two sub-species,
-allows the marine to maintain the freshwater haplotype dispersed in low frequency among its' individuals. 
-Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection reconstructs the freshwater haplotype \citep{schluter2009genetics}.
-An alternative to this hypothesis is that individuals from other freshwater environments migrate directly to the new environments, 
-and their haplotype is passed down directly. 
-This hypothesis has been shown to be unlikely due to finding a high frequency of freshwater alleles in the ocean,
-and almost no freshwater individuals.
+A finding from recent stickleback population genomic studies is that the same genomic regions, and in some cases the same haplotypes, are implicated in patterns of repeated evolution.
+These studies show that divergent stickleback populations are clearly still connected via bi-directional gene flow.
+This reuse of standing genetic variation motivated empirical studies to ask whether this phenotypic and genomic evolution could occur in decades.
+The answer is a clear yes.
+For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island.
+Quickly inhabited by the surrounding marine population of stickleback, \citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be parallel to freshwater stickleback that have been separated for over thousands of years.
+In freshwater stickleback, the number of lateral plates are reduced and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region.
 
-\plr{Bring up the possibility of more than one freshwater haplotype here.}
-\jgg{Re: should I? Doesn't Thom's work and parallel above suggest they have the same alleles?}
+ But how can evolution occur so quickly?
+The leading hypothesis for stickleback is that rather than acting on new mutations, adaptation to freshwater environments is accelerated by selection on SGV found in marine populations.
+The first clear example of the global reuse of SGV was the gene \textit{eda} which has been shown to be an important regulator for the number of lateral plates.
+While the low lateral plate version of this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently.
+More recently, population genomic studies employing genome-wide haplotype analyses has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
+Genomic regions that exhibit local adaptation, as evidenced by increased relative population genomic divergence, also show much older haplotypes that are enriched in the alternative habitats.
 
-If selection on standing genetic variants is key in rapid adaptation, 
-many questions are exposed concerning the surrounding population genetics parameters and underlying genomic architecture.
-%How rapidly can selection act on standing genetic variants at a given value of migration, $M$?
-What scale of historical introgression between populations with selective pressures $X$ and $Y$, allows for rapid and parallel local adaptation
-of a population derived from $X$ to an introduced environment with selective pressure $Y$?
-What is the origin of pre-existing adaptive variants?
+These empirical data generally support the ``transporter''-hypothesis proposed by Conte and Schluter in 2009 \citet{schluter2009genetics}, which is a conceptual model for the flow of freshwater alleles from multiple smaller freshwater populations into much larger and less structured marine populations through hybridization events.
+Alleles in this asymetrically structured metapopulation can then be recycled for subsequent freshwater adaptation.
+Several questions remain, however, about the manner and degree to which these alleles are scattered among the marine fish and must be re-assembled.
+Is it more akin to the atomization and rebuilding that occurred in the Star Trek series that motivated this hypothesis?
+Alternatively does it occur in more of a patchwork by individuals, or their early generation hybrids, in geographically adjacent freshwater habitats moving quickly through the marine environment to seed new freshwater populations.
+More generally, how do different levels of gene flow and local adaptation interact with the asymmetric nature of stickleback population structure affect the dynamics of the origin and rate of adaptive alleles, as well as their eventual organization into marine and freshwater typical genomes?
+
+Here we develop a forward simulation approach in SLiM to model the stickleback population structure in marine and freshwater habitats to document the effect of variation in gene flow and recombination on the genetic and genomic architecture of local adaptation.
+We address how rapidly can selection act on standing genetic variants at a given value of migration, $M$, and what the precise origin is of those variants.
+We ask what scale of historical introgression between populations with selective pressures $X$ and $Y$, allows for rapid and parallel local adaptation of a population derived from $X$ to an introduced environment with selective pressure $Y$?
 How are the variants structured, historically and across the genome underlying the trait?
-%How does population genetics data influenced by historical introgression and selection on standing genetic variants?
-%How many alleles underly a certain trait and how many are swept away by drift? 
 Furthermore, how can we infer causal loci for regions of the genome that must be driving the rapid adaptation, from real data.
-Many biologists today make use of genome wide association studies (GWAS) and 
-$F_{st}$ across the genome to estimate regions responsible for certain traits. 
+Many biologists today make use of genome wide association studies (GWAS) and $F_{st}$ across the genome to estimate regions responsible for certain traits.
 Unfortunately, the data can be heavily influenced by the level of introgression and population structure of the samples.
@@
 
-There are many different parameters that carry a large impact on the questions above.
-\citet{ralph2015convergent} 
-explored the dichotomy of selection on new alleles and those brought in by migrants from a similar selective pressure.
-Answering the fundamental questions surrounding spatial resolution of convergent evolution. 
-%This lead to an understanding of the process by which alleles move from an existing 'patch' to an introduced 'patch' with the 
-%same selective pressure. 
-\jgg{It seems like this could be a good parallel to what we're doing here, or maybe for discussion?}
-
-In this paper, we explore time to local adaptation as a function of migration rate through an environment (marine) with opposing selective pressure. 
-Next, we compare to real data and explore how $F_{st}$ data can be impacted by introgression.
-%Explain what we do and what we find that is important
-
-%We use forward moving simulations to model marine and freshwater populations of stickleback.
-%We then observe the effect (focusing on migration and recombination, for now) 
-%of different parameters on local adaptation of introduced freshwater environments.
+We find that ....
+\textbf{1.
+rewrite the previous paragraph to be more concise and tight, and 2.
+briefly restate our major findings which should line up well with those presented in the abstract.
+The following I grabbed from the beginning of the Results section}
+*****
+We varied the migration rate, $m$, from $5 \times 10^{-5}$ to $5 \times 10^{-2}$ per individual, i.e., between 0.01 and 10 migrants per year to and from each of the ten lakes.
+This had a strong effect on many aspects of adaptation, including how fast adaptation occurred in each lake, how much alleles were shared between lakes,
+and the population genetic signals left behind.
+At the lowest migration rate, lakes adapted nearly independently, while at the highest migration rate, the habitats are beginning to act like a single metapopulation with substantial migration load.
+We will now dissect what happens between these extremes.
+Interestingly, the ability for separate populations to locally adapt to their own selective pressure was relatively unaffected until the highest rate of migration between the marine and freshwater environments.
+All rates of historical introgression aside from the lowest, helped both the initial and introduced freshwater populations share pre-existing freshwater adapted alleles.
+The sharing of alleles resulted in rapid adaptation ($\approx 100$ generations) of the introduced population (split from the marine population) to adapt to the freshwater environment.
+We also found that larger amounts of migration allow for more statistical power and less false positives in the resulting population genetics data ($F_{st}$ per SNP across the genome).
+****
 
@@
 
-We explored these questions
-using forwards-time simulations with explicit genomic representation of a quantitative trait
-in SLiM \citep{haller2017slim}.
-The details (described below) are motivated by previous population genomic studies of threespine stickleback,
-but remain simplistic in some aspects due to computational constraints.
-\jgg{should we brag about treeSeq here a little?}
-Possibly the most important caveat is that simulated population sizes are much smaller
-than the census size of the Threespine Stickleback population;
-we will consider the likely effects of this in the Discussion.
-\plr{TODO}
+We explored these questions using forwards-time simulations with explicit genomic representation of a quantitative trait in SLiM \citep{haller2017slim}. 
+The details (described below) are motivated by previous population genomic studies of threespine stickleback, but remain simplistic in some aspects due to computational constraints.
+Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback population;
+we will consider the likely effects of this in the Discussion. \plr{TODO}
 
 \paragraph{Habitat and geography}
-Our model has two habitat types, which are considered separate populations
-with a fixed total size of 5,000 diploid individuals each. The marine population is treated as a single large deme, but the freshwater population is subdivided into local subpopulations or demes.
-The scheme, depicted in Figure \ref{fig:Geo},
-roughly models a set of freshwater habitats
-along a stretch of coastline.
-The ``marine'' habitat  is a continuous, one-dimensional range of 25 units of length,
-and the ``freshwater'' habitat is divided into ten subpopulations
-(which we call ``lakes''), each connected to the marine habitat
-at positions $i - 1/2$ for $1 \le i \le 25$.
+Our model has two habitat types, which are considered separate populations with a fixed total size of 5,000 diploid individuals each. 
+The marine population is treated as a single large deme, but the freshwater population is subdivided into local subpopulations or demes. 
+The scheme, depicted in Figure \ref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. 
+The ``marine'' habitat  is a continuous, one-dimensional range of 25 units of length, and the 
+``freshwater'' habitat is divided into ten subpopulations (which we call ``lakes''), each connected to the marine habitat at positions $i - 1/2$ for $1 \le i \le 25$.
 
-Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff such that alternative values of the trait have high or low fitness in the two different habitats.
-This situation roughly models the cumulative effect of the various phenotypes
-such as armor morphology, body size, craniofacial variation and opercle shape (among others)
-on which divergent selection acts in the two environments.
-The optimal trait value in the marine habitat is $+10$,
-and in the freshwater habitat is $-10$. Fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$ is
-Gaussian with standard deviation, i.e.,
+Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff such that alternative values of the trait have high or low fitness in tht two different habitats. 
+This situation roughly models the cumulative effect of the various phenotypes such as armor morphology, 
+body size, craniofacial variation and opercle shape (among others) on which divergent selection acts in the two environments. 
+The optimal trait value in the marine habitat is $+10$, and in the freshwater habitat is $-10$. 
+Fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$ is Gaussian with standard deviation, i.e.,
+
 \begin{align*}
@@
 \end{figure}
+
+\paragraph{Initial genomic architecture} Each individual carries two copies of a linear chromosome of $10^8$ loci(which could be thought of roughly as a kilobase each). 
+New mutations occur with probability $10^{-7}$ per locus per generation. 
+Ten ``effect regions'' of $100Kb$ loci each are spread evenly along the chromosome (separated by 4950 loci) in which mutations affect the trait; 
+mutations in other parts of the genome are neutral. 
+Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability), 
+and with an effect sizes chosen randomly from an Exponential distribution with mean $1/2$, either positive or negative with equal probability. 
+Individual trait value are determined additively from the diploid genotypes. 
+Concretely, an individual that is heterozygous and homozygous for mutations at loci $H$ and $D$ respectively has trait value 
+$x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$, 
+where $s_\ell$ is the effect size of the mutation at locus $\ell$. Subsequent mutations at the same locus replace the previous allele.
+
+\paragraph{Equilibrium in initial marine and freshwater populations} 
+Each generation, the two parents of each new offspring are chosen proportional to their fitness 
+(all individuals are hermaphroditic),
+and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation 
+($10^{-7}$ per locus per generation). 
+The model is Wright-Fisher, so that each generation, in each habitat, 
+2000 new individuals are produced, and so to keep population sizes roughly constant within each lake, 
+at the start of each generation we rescale fitness values in the freshwater habitat so that the sum of the fitness values within each lake are equal.
+
+Migration occurs both locally along the coastline in the marine habitat and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
+The between-habitat migration rate is denoted $m$, and will be called simply the ``migration rate''. 
+Each new individual in the marine habitat has freshwater parents with probability $m$; 
+to obtain the pair, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. 
+The resulting offspring is given a spatial location in the marine habitat at the location of the parent's lake. 
+Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$): 
+first, a single parent is chosen proportionally to fitness in the marine habitat,
+and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with stanadard deviation $1/2$.
+
+Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_i) \exp(-2d_{ij}^2)$,
+where $d_{ij}$ is the distance between the two locations. 
+Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.02, 
+and reflected to stay within the population. 
+New offspring in the freshwater habitat are chosen in the same way, except the probability that the parents are marine individuals is $m$; 
+any new freshwater offpsring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.
+
+\paragraph{Population genomic analyses of new freshwater populations} 
+To study adaptation in newly appearing freshwater habitats, we introduce a new set of lakes midway through the simulation. 
+The initial set of individuals have parents chosen as above from the marine habitat, 
+and act like an independent copy of the original set of lakes -- in particular, 
+the two sets of lakes each have 2,000 individuals at all times. 
+Since this doubles the number of lake-to-marine immigrants, after this happens the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.
+
+To assess whether new lakes adapt using existing genetic diversity, we define a \emph{pre-existing freshwater adapted allele}
+to be an effect mutation that has frequency higher than $0.5$ in at least one of the original lakes, while remaining lower than $0.5$ in the marine. 
+This categorization is made for each generation using the allele frequencies from that generation, and so changes with time. Alleles common in the newly introduced lakes do not count
+if they are not also common in the original lakes. 
+They are defined this way because the transportation hypothesis does not specify where or when an advantageous mutation arises, 
+but simply suggests that any sufficiently common freshwater adapted allele could participate in adaptaiton in new habitat \citep{schluter2009genetics}.
+
+\emph{Time to adaptation} of the introduced population, denoted $T_\text{adapt}$, is defined to be the generation at which the difference between the average trait value in the original and the introduced freshwater populations is less than 0.5.
+We describe overall genetic differentiation between the habitats using $F_{ST}$,
+calculated on a per-locus basis. Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively,
+and $\bar p = (p_f + p_m)/2$, then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
 
@@
 As expected, increasing migration rate decreased differentiation between habitats. 
-As seen in Figure \ref{fig:Fst},
-$F_{ST}$ between marine and freshwater habitats at neutral sites 
-steadily declines as migration increases.
+As seen in Figure \ref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases. 
 Local adaptation was still able to occur despite overall homogenization: 
-if computed using only sites with alleles affecting the trait (``effect mutations''),
-$F_{ST}$ between habitats was relatively constant across migration values.
+if computed using only sites with alleles affecting the trait (``effect mutations''), $F_{ST}$ between habitats was relatively constant across migration values.
 
@@
 \end{figure}
+
+\begin{comment}
+\paragraph*{Rate of adaptation}
+Increasing migration rate strongly decreased the time until freshwater habitats could adapt, both in the initial and introduced sets of lakes.
+As shown in Figure \ref{fig:TimeToAdaptation}, at the lowest migration rate it took over 20 thousand generations for the mean trait value across introduced lakes to get to within 0.5 of the original lakes value.
+Although many lakes had adapted before this time, the different rates of introduction of effect alleles is clearly seen in the traces of trait values against time (e.g., Figure \ref{fig:phenotype_ts2}).
+However, higher migration rates allowed the freshwater habitat to adapt nearly as quickly as the marine habitat.
+\end{comment}
 
@@
 \subsection*{Sharing of freshwater alleles}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Sharing of freshwater adapted alleles}
+
+%FIGURES THAT SHOW THIS:
+%-MPAA / IND
+%-Total/Avg shared
+\begin{comment}
+The tenfold difference in speed of adaptation occurs because at low migration rates, adaptation occurs independently in each lake,and the marine habitat has ten times the influx of new alleles than any one lake.
+In other words, greater mixing at higher migration rates allows lakes to share alleles instead of developing their own genetic basis of adaptation.
+As a first indication of this, Figure \ref{fig:Fst} shows that $F_{ST}$ between the ``original'' and ``introduced'' sets of lakes at effect mutations decreased with migration rate.
+
+To further investigate sharing of locally adaptive alleles and the transporter hypothesis, we investigate the ``pre-existing freshwater adapted alleles'',
+defined for a particular generation to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat in that generation.
+Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles, across generations, as well as the average number of lakes each is found in.
+As migration rates increase, the number of these alleles decreases steadily, and each is concurrently found in a greater number of lakes.
+
+As the number of common, locally adaptive alleles decreases, the genetic basis of adaptation is more commonly shared.
+Figure \ref{fig:MPFAI}B shows the distribution of the mean percentage of currently-defined freshwater adapted alleles that each genome in each of the populations carries, averaged across time and individuals.
+If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%.
+This value is nearly reached at the highest migration rate; but it is lower due to migration load.
+At the lowest migration rate, each genome in the original lakes have almost exactly $1/10^{th}$ of the total freshwater adapted alleles -- since there are 10 lakes, this implies that each lake has adapted with a unique set of alleles.
+Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
+\end{comment}
 
@@
 
+\begin{comment}
+At first, increased migration allows sharing of adaptive alleles between lakes, but at the highest migration rate, the constant influx of alleles between the habitats creates substantial migration load.
+This rate, $m=.05$, only replaces 5\% of each population each generation with migrants from the other habitat, but this is sufficient to shift the mean trait values to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
+Significant gene flow constricts local adaptation as a consequence of a large number of offspring through hybridization events between subpopulations.
+
+\paragraph{Standing genetic variation} Despite a dramatic difference in the amount of allelic sharing between lakes, standing genetic variance (Figure \ref{fig:SGV}) was around 0.05 in the freshwater habitat across all but the highest migration rates.
+Concurrently, genetic variation in the marine population steady increased with migration
+to a similar value.
+On the one hand, it is not surprising that lakes, as a group, carry more genetic variation than the marine habitat, since population subdivision allows different alleles to become common in each.
+However, it is striking that at $m=.005$, the marine habitat carries as much genetic variation, despite a lack of any substantial migration load.
+
+\end{comment}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection*{Migration load and genetic variation}
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{comment}
+\subsection*{Realized genomic architecture}
+Now we take a closer look at the genomic architecture of local adaptation between the two habitats. Do the alleles underlying trait differences cluster along the genome?
+Do measures of local differentiation identify the causal loci?
+Figure \ref{fig:Fst3} shows plots along the genome of per-locus $F_{ST}$ values between the marine and original freshwater habitats at the two intermediate migration rates.
+(Note that we are pooling freshwater habitats; a single lake would provide substantially less power.)
+
+\plr{need to know where the FAA are in these figures}
+
+Higher migration rates showed more distinct $F_{ST}$ peaks over polymorphic loci underlying trait differences between the habitats.
+As migration rate decreased, the ``background'' levels of $F_{ST}$ increased, swamping out this signal until the regions under selection were indistinguishable.
+This is likely due to two reasons: first, stronger genetic drift with less migration leading to higher background $F_{ST}$, and second, greater sharing of adaptive alleles providing a shared signal across populations.
+
+This suggests that genome scans for local adaptation based purely on measures of differentiation will only be successful given enough migration between habitats.
+To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.
+
+\plr{Labels are wrong on that figure: false pos rate should go down with cutoff.
+     Is it true positive rate?}
+
+In regions of the genome underlying individual trait value, we observed Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$.
+Knowing exactly which mutations effect phenotype in our simulations, we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome.
+In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power.
+This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.
+\end{comment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Suppose a new allele enters a lake, either by migration or mutation.
-If, when it is rare but present in $n$ copies,
-it has fitness advantage $s$
--- i.e., the expected number of copies in the next generation is $(1+s)n$ --
-then the probability that it escapes demographic stochasticity to become common in the population
-is approximately $2s$ \citep{fisher,prob_fixation}.
-If the current population all differed from the optimum trait by $z$,
-and the allele has effect size $-u$ in heterozygotes,
-then the fitness advantage of the allele would be
-$s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$,
-where in our parameterization, $\beta = 1 / 450$.
-This tells us two things:
-(1) the rate of adaptation decreases as the population approaches the optimum,
-and (2) larger mutations (in the right direction) are more likely to fix.
+If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ --then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{fisher,prob_fixation}.
+If the current population all differed from the optimum trait by $z$, and the allele has effect size $-u$ in heterozygotes, then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
+This tells us two things: (1) the rate of adaptation decreases as the population approaches the optimum, and (2) larger mutations (in the right direction) are more likely to fix.
 
 \paragraph{New mutations}
-The total rate of appearance of new mutations per lake is $\mu_L = 0.04$,
-which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive
-in either direction.
-This implies that a new additive or dominant effect mutation appears once every 87.5 generations,
-on average.
-Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, 
-and so the probability that a dominant mutation manages to establish in a population differing from
-the optimum by $z$ is roughly
-$\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$,
-and so the rate of establishment of dominant mutations is $\beta z / 87.5$,
-i.e., about one such mutation every $2461/z$ generations.
-The distribution of these successfully established mutations 
-has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2.
-Since additive alleles have half the effect in heterozygotes,
-they have half the probability of establishment.
-During the initial phase of adaptation,
-the populations begin at around distance $z=10$ from the optimum.
-Combining these facts, we expect 
-adaptive alleles to appear through mutation at first on a time scale of 250 generations,
-with the time between local fixation of new alleles increasing as adaptation progresses,
-and each to move the trait by a distance of order 1.
+The total rate of appearance of new mutations per lake is $\mu_L = 0.04$, which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive in either direction.
+This implies that a new additive or dominant effect mutation appears once every 87.5 generations, on average.
+Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, and so the probability that a dominant mutation manages to establish in a population differing from the optimum by $z$ is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$, and so the rate of establishment of dominant mutations is $\beta z / 87.5$, i.e., about one such mutation every $2461/z$ generations.
+The distribution of these successfully established mutations  has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2.
+Since additive alleles have half the effect in heterozygotes, they have half the probability of establishment.
+During the initial phase of adaptation, the populations begin at around distance $z=10$ from the optimum.
+Combining these facts, we expect adaptive alleles to appear through mutation at first on a time scale of 250 generations, with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
 
 \paragraph{Standing variation}
-An allele that moves the trait $z$ units in the freshwater direction in heterozygotes
-has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment
-(which is close to optimal).
-The product of population size and fitness differential in the marine environment
-for a mutation with $z=1$
-is therefore $2Ns = 8.9$, implying that these alleles are strongly selected against
-but might occasionally drift to moderate frequency.
-The average frequency of such an allele in the marine environment at migration-selection equilibrium
-is equal to the total influx of alleles per generation divided by the selective disadvantage,
-which if $M = 2000 m$ is the number of immigrants per generation,
-is $2 M / \beta z^2$. 
-With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$:
-since lakes have 400 genomes, 
-as long as $M \ge 1$, the chances are good that any particular lake-adapted allele
-that is present in all pre-existing lakes
-will appear at least once in the fish that colonize a new lake.
-However, an allele with $z=1$ only has probability of around 1/20 of establishing locally,
-suggesting that we'd need $M \ge 10$ to ensure enough pre-existing genetic variation
-that adaptation would happen entirely using the initial set of colonizers.
-This corresponds to our highest two migration rates,
-as in e.g., Figure \ref{fig:phenotype_ts2}B.
+An allele that moves the trait $z$ units in the freshwater direction in heterozygotes has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment (which is close to optimal).
+The product of population size and fitness differential in the marine environment for a mutation with $z=1$ is therefore $2Ns = 8.9$, implying that these alleles are strongly selected against but might occasionally drift to moderate frequency.
+The average frequency of such an allele in the marine environment at migration-selection equilibrium is equal to the total influx of alleles per generation divided by the selective disadvantage, which if $M = 2000 m$ is the number of immigrants per generation, is $2 M / \beta z^2$.
+With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$: since lakes have 400 genomes, as long as $M \ge 1$, the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes will appear at least once in the fish that colonize a new lake.
+However, an allele with $z=1$ only has probability of around 1/20 of establishing locally, suggesting that we'd need $M \ge 10$ to ensure enough pre-existing genetic variation that adaptation would happen entirely using the initial set of colonizers.
+This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.
 
 \paragraph{Migration}
-If sufficient genetic variation is not present in a new lake initially,
-it must appear by new mutation or migration.
-Since a proportion $m$ of each lake is composed of migrants each generation,
-it takes $1/m$ generations until the genetic variation introduced by migrants
-equals the amount initially present at colonization.
-This implies a dichotomy: either 
-(a) adaptation is possible using variants present at colonization
-or arriving shortly thereafter, or 
-(b) adaptation takes many multiples of $1/m$ generations.
+If sufficient genetic variation is not present in a new lake initially, it must appear by new mutation or migration.
+Since a proportion $m$ of each lake is composed of migrants each generation, it takes $1/m$ generations until the genetic variation introduced by migrants equals the amount initially present at colonization.
+This implies a dichotomy: either (a) adaptation is possible using variants present at colonization or arriving shortly thereafter, or (b) adaptation takes many multiples of $1/m$ generations.
 
-These calculations depend on there being no bottleneck in colonization of the lake;
-if there is a bottleneck, then an additional factor must be added.
-
+These calculations depend on there being no bottleneck in colonization of the lake; if there is a bottleneck, then an additional factor must be added.
 At what point do we expect new mutation to be more important than migration for adaptation?
-By the calculations above, if $M \ge 10$, we expect initial diversity in a lake to be sufficient 
-for adaptation, corresponding to our third-highest migration rate.
+By the calculations above, if $M \ge 10$, we expect initial diversity in a lake to be sufficient  for adaptation, corresponding to our third-highest migration rate.
 If this does not happen, then we expect adaptation to take a multiple of $1/m$ generations;
 with our values, $1/m$ ranges from 20,000 to 20 generations.
-Above we estimated that adaptive alleles due to new mutation fix locally about every 2,000 generations,
-suggesting that at our second-lowest migration rate (where $1/m = 2,000$),
-the two contributions of migration and new mutation are roughly equal.
-This is in fact what we see: in Figure \ref{fig:MPFAI},
-we see that at the second migration rate, alleles start to be shared between lakes,
-while by the third migration rate, they are almost entirely shared.
+Above we estimated that adaptive alleles due to new mutation fix locally about every 2,000 generations, suggesting that at our second-lowest migration rate (where $1/m = 2,000$), the two contributions of migration and new mutation are roughly equal.
+This is in fact what we see: in Figure \ref{fig:MPFAI}, we see that at the second migration rate, alleles start to be shared between lakes, while by the third migration rate, they are almost entirely shared.
 
@@
 
-We have shown that historical introgression, at our given parameter sets,
-is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island.
+We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island.
 Selection is able to rebuild the freshwater haplotype from marine populations as a medium between all freshwater populations.
-Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load 
-limited the ability of the populations to reach the local optimum.
+Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.
+
+We also have also explored the genomic architecture as a consequence of the nature of selection in our scenario.
+The alleles underlying individual trait value were of large effect and a low number.
+With a total of $10^{5}$ loci, to be realistic, each loci should represent 1000 $Kb$ in real data.
 
@@
 mistaken for being causative as they often display the same amount of divergence.
+>>>>>>> master
 %wouldn't it be cool if we could use this fact and simulate introgression, starting with real data that is noisy to extrapolate the areas we care about?

### 2018-10-22 Stickleback_Paper.tex
fixed header

@@
 
+\begin{comment}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %-MPAA / IND
-%-Total/Avg shared
-\begin{comment}
 The tenfold difference in speed of adaptation occurs because at low migration rates, adaptation occurs independently in each lake,and the marine habitat has ten times the influx of new alleles than any one lake.

### 2018-12-07 Stickleback_Paper.tex
Update on Overleaf.

@@
 
-\title{A simulation study of gene flow and local polygenic adaptation motivated by threespine stickleback evolution}
+\title{A few stickleback suffice to transport adaptive alleles to new lakes}
 \author{Jared Galloway, William A. Cresko, and Peter Ralph}
@@
 
-\textbf{Other title ideas?} (Bad ideas allowed.)
-
-How do stickleback share adaptations between lakes? A simulation study.
-
-A simulation study of gene flow and local polygenic adaptation
-
-A simulation study of gene flow and local polygenic adaptation in stickleback fish
-
 \section*{Abstract}
 
-Threespine stickleback fish provide a striking example of local adaptation in the context of recurrent gene flow.
-This Northern hemisphere-wide metapopulation includes both marine populations and a large number of smaller freshwater populations that have repeatedly adapted to freshwater conditions, often by using standing genetic variation.
-Here we use genealogical simulations to determine the levels of gene flow can best match the observed patterns of allele sharing among habitats in stickleback, and to provide a framework for better understanding of the dynamics of gene flow and local adaptation for the maintenance and reuse of standing genetic variation.
-We find that rapid, repeated adaptation using alleles maintained at low frequency by migration-selection balance occurs over a realistic range of intermediate rates of gene flow.
-Slow and independent adaptation occurs with low gene flow, whereas large migration load occurs at high gene flow, mediated mainly by the total downstream influx of alleles \emph{(maybe??)}.
-We do not see evidence for strong effects encouraging genomic clustering of causal alleles or towards particular dominance coefficients.
-$F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow, The results support existing theory of local adaptation, and provide a more concrete look at a particular, empirically motivated example.
+Threespine stickleback fish provide a striking example of local adaptation despite recurrent gene flow.
+The species is distributed around the Northern hemisphere in both marine and freshwater habitats.
+It is thought that these numerous, smaller freshwater populations
+have been established "de novo" from marine fish,
+and that a shared freshwater phenotype
+is often established using standing genetic variation.
+Here we use genealogical simulations to determine the levels of gene flow
+that best matches observed patterns of allele sharing among habitats in stickleback, 
+and more generally to better understand how gene flow and local adaptation in large metapopulations
+determine speed of adaptation and reuse of standing genetic variation.
+We find that rapid, repeated adaptation using a shared set of alleles
+maintained at low frequency by migration-selection balance
+occurs over a realistic range of intermediate rates of gene flow.
+Low gene flow leads to slow, independent adaptation of distinct habitats,
+whereas high gene flow leads to large migration load.
+We do not see evidence for strong effects encouraging genomic clustering of causal alleles
+or towards particular dominance coefficients.
+$F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow.
+The results support existing theory of local adaptation,
+and provide a more concrete look at a particular, empirically motivated example.
 
@@
 
-We explored these questions using forwards-time simulations with explicit genomic representation of a quantitative trait in SLiM \citep{haller2017slim}. 
-The details (described below) are motivated by previous population genomic studies of threespine stickleback, but remain simplistic in some aspects due to computational constraints.
-Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback population;
-we will consider the likely effects of this in the Discussion. \plr{TODO}
+We explored these questions using forwards-time simulations with explicit genomic representation of a quantitative trait in SLiM \citep{haller2017slim,haller2018slim3}.
+The details (described below) are motivated by current understanding of threespine stickleback history and demography, but remain simplistic in some aspects due to computational constraints.
+Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback population
+(a caveat we revisit in the Discussion).
 
@@
 		is a discrete population connected by migration only through the marine, at position $i - 0.5$. 
-		Each of the introduced lakes have the same spatial dynamic and selective pressure as the pre-existing lakes, but arise at 100K generations into the simulation.
+		Each of the introduced lakes have the same spatial dynamic and selective pressure 
+        as the pre-existing lakes, but arise at 100K generations into the simulation.
 		The marine has $5000$ diploid individuals while each of the lakes fluctuate  $200$ diploid individuals. 

### 2018-12-07 Stickleback_Paper.tex
Update on Overleaf.

@@
 \paragraph{Habitat and geography}
-Our model has two habitat types, which are considered separate populations with a fixed total size of 5,000 diploid individuals each. 
-The marine population is treated as a single large deme, but the freshwater population is subdivided into local subpopulations or demes. 
-The scheme, depicted in Figure \ref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. 
-The ``marine'' habitat  is a continuous, one-dimensional range of 25 units of length, and the 
-``freshwater'' habitat is divided into ten subpopulations (which we call ``lakes''), each connected to the marine habitat at positions $i - 1/2$ for $1 \le i \le 25$.
+Each habitat type -- marine and freshwater -- have a fixed total of 5,000 diploid individuals each. 
+The arrangement of these habitats, depicted in Figure \ref{fig:Geo},
+roughly models a set of freshwater habitats along a stretch of coastline. 
+The marine habitat  is a continuous, one-dimensional range of 25 units of length,
+while the freshwater habitat is divided into ten subpopulations (which we call ``lakes''),
+each connected to the marine habitat at regularly spaced intervals
+(positions $i - 1/2$ for $1 \le i \le 25$).
 
-Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff such that alternative values of the trait have high or low fitness in tht two different habitats. 
+Divergent selection is mediated by a single quantitative trait
+with different optima in marine and freshwater habitats.
 This situation roughly models the cumulative effect of the various phenotypes such as armor morphology, 
-body size, craniofacial variation and opercle shape (among others) on which divergent selection acts in the two environments. 
-The optimal trait value in the marine habitat is $+10$, and in the freshwater habitat is $-10$. 
-Fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$ is Gaussian with standard deviation, i.e.,
-
+body size, craniofacial variation and opercle shape on which divergent selection is thought to act
+in the two environments. 
+Concretely, the optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively,
+and fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$
+is determined by a Gaussian kernel with standard deviation 15, i.e.,
 \begin{align*}
-    f(x; x_0)
+    f(x; x_\text{opt})
     &=
@@
 \end{align*}
+Note that only the the scale of trait values relative to mutation effect size is relevant.
+We chose the difference between optima and strength of stabilizing selection in each habitat
+so that 
+(a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,
+and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.
 
@@
 mistaken for being causative as they often display the same amount of divergence.
->>>>>>> master
+
 %wouldn't it be cool if we could use this fact and simulate introgression, starting with real data that is noisy to extrapolate the areas we care about?
@@
 
+\paragraph{Modeling assumptions}
+Our simulations had much smaller population sizes than real populations.
+How might this affect things?
+
 \bibliographystyle{plainnat}

### 2018-12-08 Stickleback_Paper.tex
peter edits

@@
 We explored these questions using forwards-time simulations with explicit genomic representation of a quantitative trait in SLiM \citep{haller2017slim,haller2018slim3}.
-The details (described below) are motivated by current understanding of threespine stickleback history and demography, but remain simplistic in some aspects due to computational constraints.
-Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback population
-(a caveat we revisit in the Discussion).
+The details of the model were motivated by current understanding of threespine stickleback history and demography,
+but remain simplistic in some aspects due to computational constraints.
+Possibly the most important caveat
+is that simulated population sizes are much smaller than the census size of the threespine stickleback population
+(see the Discussion for more on this).
 
@@
   		\caption{
-			A representation of the geographic and evolutionary history of all populations throughout the simulation. 
-		The marine is a one-dimensional, continuous population with spatial positions ranging from [0.0 - 25.0]. 
-		Each freshwater lake$_{i}$ for both the introduced and freshwater populations
-		is a discrete population connected by migration only through the marine, at position $i - 0.5$. 
-		Each of the introduced lakes have the same spatial dynamic and selective pressure 
-        as the pre-existing lakes, but arise at 100K generations into the simulation.
-		The marine has $5000$ diploid individuals while each of the lakes fluctuate  $200$ diploid individuals. 
-		the introduced population is initialized as a copy of all marine individuals to model marine 
-		individuals inhabiting a newly created freshwater environment such as the ponds around Middleton Island.
-		We then observe the selection process of marine individuals and following generations 
-		with the new selective pressure across a range of parameter values. 
+            \textbf{Diagram of simulated populations:}
+            a single, continuous, one-dimensional marine habitat (blue)
+            is coupled to randomly mating ``lakes'' at discrete locations.
+            After an inital period of 100K generations with 25 lakes,
+            an additional 25 lakes are added (at the same set of locations)
+            to simulate the appearance of newly accessible freshwater habitats.
+            The marine habitat, and each set of 25 lakes, each contain 5,000 individuals at all times.
 			}
@@
 
-\paragraph{Initial genomic architecture} Each individual carries two copies of a linear chromosome of $10^8$ loci(which could be thought of roughly as a kilobase each). 
-New mutations occur with probability $10^{-7}$ per locus per generation. 
-Ten ``effect regions'' of $100Kb$ loci each are spread evenly along the chromosome (separated by 4950 loci) in which mutations affect the trait; 
-mutations in other parts of the genome are neutral. 
+\paragraph{Genetic architecture of the trait}
+Each individual carries two copies of a linear chromosome of $10^8$ loci.
+Mutations that can affect the trait under selection can occur at rate $10^{-7}$ per locus per generation
+in ten ``effect regions'' of $100$ loci each,
+spread evenly along the chromosome (separated by 4950 loci).
 Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability), 
 and with an effect sizes chosen randomly from an Exponential distribution with mean $1/2$, either positive or negative with equal probability. 
-Individual trait value are determined additively from the diploid genotypes. 
-Concretely, an individual that is heterozygous and homozygous for mutations at loci $H$ and $D$ respectively has trait value 
-$x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$, 
-where $s_\ell$ is the effect size of the mutation at locus $\ell$. Subsequent mutations at the same locus replace the previous allele.
+Individual trait values are determined additively from the diploid genotypes. 
+Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value 
+$x = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, 
+where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$.
+Subsequent mutations at the same locus replace the previous allele.
 
-\paragraph{Equilibrium in initial marine and freshwater populations} 
+\paragraph{Population dynamics}
+We use SLiM to simulate a Wright--Fisher population with nonoverlapping generations 
+and a fixed population size of 5000 diploid individuals in each habitat.
 Each generation, the two parents of each new offspring are chosen proportional to their fitness 
@@
 ($10^{-7}$ per locus per generation). 
-The model is Wright-Fisher, so that each generation, in each habitat, 
-2000 new individuals are produced, and so to keep population sizes roughly constant within each lake, 
-at the start of each generation we rescale fitness values in the freshwater habitat so that the sum of the fitness values within each lake are equal.
+Since the total population across \emph{all} lakes is regulated,
+to keep population sizes roughly constant within each lake, 
+before selection we divide fitness values of each freshwater individual by the mean fitness in their lake,
+so that the mean fitnesses of all lakes are equal.
 
-Migration occurs both locally along the coastline in the marine habitat and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
-The between-habitat migration rate is denoted $m$, and will be called simply the ``migration rate''. 
+Dispersal occurs both locally along the coastline in the marine habitat 
+and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
+There is no dispersal directly between lakes.
+The lake--ocean migration rate is denoted $m$, and will be called simply the ``migration rate''. 
 Each new individual in the marine habitat has freshwater parents with probability $m$; 
@@
 and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with stanadard deviation $1/2$. 
-
-Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_i) \exp(-2d_{ij}^2)$,
+Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
 where $d_{ij}$ is the distance between the two locations. 
@@
 
-\paragraph{Population genomic analyses of new freshwater populations} 
-To study adaptation in newly appearing freshwater habitats, we introduce a new set of lakes midway through the simulation. 
-The initial set of individuals have parents chosen as above from the marine habitat, 
+\paragraph{New freshwater populations} 
+To study how newly appearing freshwater habitats adapt, we introduce a new set of 25 lakes midway through the simulation.
+The initial set of individuals in these new lakes have parents chosen in the same way from the marine habitat as ocean-to-lake migrants,
 and act like an independent copy of the original set of lakes -- in particular, 
-the two sets of lakes each have 2,000 individuals at all times. 
-Since this doubles the number of lake-to-marine immigrants, after this happens the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.
+the two sets of lakes each have 5,000 individuals at all times. 
+(Since this doubles the number of lake-to-marine immigrants, 
+after this happens the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.)
 
-To assess whether new lakes adapt using existing genetic diversity, we define a \emph{pre-existing freshwater adapted allele}
-to be an effect mutation that has frequency higher than $0.5$ in at least one of the original lakes, while remaining lower than $0.5$ in the marine. 
-This categorization is made for each generation using the allele frequencies from that generation, and so changes with time. Alleles common in the newly introduced lakes do not count
+We quantify how quickly these new populations adapt to their new environment
+by the \textbf{time to adaptation} (denoted $T_\text{adapt}$), 
+which we define to be the number of generations until 
+the difference between the average trait value in the original and the introduced freshwater populations is less than 0.5. 
+We also quantify genetic differentiation between the habitats with per-locus $F_{ST}$:
+if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively,
+and $\bar p = (p_f + p_m)/2$, then $F_{ST} = 1 - (p_f (1-p_f) + p_m (1-p_m))/ (2 \bar p (1-\bar p))$.
+
+We also quantify the degree to which adaptation in new lakes uses alleles 
+already present and specific to the pre-existing lakes,
+defining a \emph{pre-existing freshwater adapted allele}
+to be a mutation (with an effect on the trait)
+that at the time of introduction of the new populations
+has frequency higher than $0.5$ in at least one of the original lakes, 
+while remaining lower than $0.5$ in the marine habitat. 
+These alleles represent standing variation (rather than new mutations),
+but furthermore are presumably responsible for local adaptation
+in at least some lakes,
+and present only at migration-selection balance in the ocean.
+Alleles common in the newly introduced lakes do not count
 if they are not also common in the original lakes. 
-They are defined this way because the transportation hypothesis does not specify where or when an advantageous mutation arises, 
-but simply suggests that any sufficiently common freshwater adapted allele could participate in adaptaiton in new habitat \citep{schluter2009genetics}.
-
-\emph{Time to adaptation} of the introduced population, denoted $T_\text{adapt}$, is defined to be the generation at which the difference between the average trait value in the original and the introduced freshwater populations is less than 0.5.
-We describe overall genetic differentiation between the habitats using $F_{ST}$,
-calculated on a per-locus basis. Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively,
-and $\bar p = (p_f + p_m)/2$, then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
-
-\paragraph{Potential genomic architecture}\bill{Initial genomic architecture? The header is a little confusing to me}
-Each individual carries two copies of a linear chromosome of $10^8$ loci representing a realistic size genome.
-New mutations occur with probability $10^{-10}$ per locus per generation in regions affecting phenotype and $10^{-8}$ in neutral regions.
-Ten ``effect regions'' of $100 Kb$  loci each are spread evenly along the chromosome; 
-mutations in these regions have the potential to affect trait value of the individual which posses them.
-Each mutation in these regions is either additive, completely recessive,
-or completely dominant (with equal probability).
-Effect sizes are chosen randomly from an exponential distribution with mean $1/2$,
-either positive or negative with equal probability.
-Between the effect regions is $9.9 Mb$ of neutral loci.
-Individual trait values are determined additively from the diploid genotypes.
-Concretely, an individual that is heterozygous and homozygous for mutations
-at loci $H$ and $D$ respectively has trait value
-$x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
-where $s_\ell$ is the effect size of the mutation at locus $\ell$.
-Subsequent mutations at the same locus replace the previous allele.
+They are defined this way because 
+the transportation hypothesis does not specify where or when an advantageous mutation arises, 
+but simply suggests that any sufficiently common freshwater adapted allele 
+could participate in adaptation in new habitat \citep{schluter2009genetics}.
+\plr{I think we use this definition with respect to different times? 
+     But I don't see where we do this.}
 

### 2018-12-08 Stickleback_Paper.tex
gradually...

@@
 There is no dispersal directly between lakes.
-The lake--ocean migration rate is denoted $m$, and will be called simply the ``migration rate''. 
+The lake--ocean migration rate is denoted $m$, 
+which we refer to as the rate of \emph{gene flow} between habitats.
 Each new individual in the marine habitat has freshwater parents with probability $m$; 
@@
 
-We quantify how quickly these new populations adapt to their new environment
-by the \textbf{time to adaptation} (denoted $T_\text{adapt}$), 
-which we define to be the number of generations until 
-the difference between the average trait value in the original and the introduced freshwater populations is less than 0.5. 
-We also quantify genetic differentiation between the habitats with per-locus $F_{ST}$:
+We quantify genetic differentiation between the habitats with per-locus $F_{ST}$:
 if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively,
@@
 
-\paragraph{Life cycle}
-Each generation, the two parents of each new offspring are chosen 
-proportional to their fitness (all individuals are hermaphroditic),
-and the contributing genomes are produced by Poisson recombination
-with an average of one crossover per chromosome per generation
-($10^{-8}$ per locus per generation).
-The model is Wright-Fisher, so that each generation, in each habitat,
-5000 new individuals are produced,
-and so to keep population sizes roughly constant within each lake,
-at the start of each generation we rescale fitness values in the freshwater habitat
-so that the sum of the fitness values within each lake are equal.
+\paragraph{Recording genealogical history}
+We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treeseq}
+to record the genealogical history of all individuals alive 
+at the time of introduction of new lakes, the time of adaptation, and the end of the simulation.
+This allowed us to directly query the true origins of adaptive alleles,
+as well as allowing much larger simulations
+by avoiding the computationally expensive task of simulating neutral mutations
+(which were added after the fact at a rate of $10^{-7}$ per locus per generation, 
+as described in \citet{kelleher2018efficient}).
 
-Migration occurs both locally along the coastline in the marine habitat
-and between the marine habitat and the lakes,
-and can be thought of as occurring at the juvenile stage.
-The between-habitat migration rate is denoted $m$,
-and will be called simply the ``gene flow''.
-$\approx m$ individuals \emph{per} lake \emph{per} generation, have parents from the opposing selective pressure.
-A first parent is chosen proportional to fitness,
-and a mate is chosen from the same lake as the first, also proportional to fitness.
-The resulting offspring is given a spatial location in the marine habitat
-at the location of the parent's lake.
-Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$):
-first, a single parent is chosen proportionally to fitness in the marine habitat,
-and then a mate is chosen, 
-also proportionally to fitness but reweighed by a Gaussian function 
-of the distance separating the two, with standard deviation $1/2$. 
-More specifically, if the first parent is marine individual $i$,
-then marine individual $j$ is chosen as the mate
-with probability proportional to $f(x_i) \exp(-2d_{ij}^2)$,
-where $d_{ij}$ is the distance between the two locations.
-Finally, each new marine offspring is given a position 
-displaced from the first parent's position by a random Gaussian distance
-with mean 0 and standard deviation 0.02, and reflected to stay within the population.
-New offspring in the freshwater habitat are chosen in the same way,
-except the probability that the parents are marine individuals is $m$;
-any new freshwater offspring produced by marine individuals
-are assigned to the lake nearest to the position of the first marine parent.
+The output tree sequence from each simulation allows us to,
+at each locus along the genome,
+construct the genealogical tree relating all extant chromosomes
+as well as all chromosomes present at time of introduction of the new lakes.
+Using these trees,
+we could therefore classify each freshwater allele found in the new lakes at time of adaptation
+into four categories:
+\begin{itemize}
+    \item \emph{Migrants}, not present in the set of individuals that initially colonized
+        the new lake, but brought in by migration at a later time;
+    \item \emph{Pre-existing}, passed down from an individual in the original generation of the new lakes, 
+however, not defined to be an adaptive variant in the original lakes
+    \item \emph{Captured}, Pre-existing, passed down from an individual in the original generation of the new lakes, 
+and defined to be adaptive in the original lakes
+    \item \emph{De novo}, mutated and propagated from an individual existing in the introduced lakes population, after introduction. 
+\end{itemize}
 
-\paragraph{New lakes}
-To study adaptation in newly appearing freshwater habitats,
-we introduce a new set of lakes midway through the simulation.
-As with the old lakes, these are a set of discrete demes connected by migration \emph{only} to the marine.
-The initial generation of individuals in these new lakes were created as a copy of the marine individuals 
-at the generation of introduction.
-The new lake's geography acts like an independent copy of the original set of lakes -- in particular,
-the two sets of lakes each have 5,000 individuals at all times.
-%(Gene flow constricted to within the lakes and through migration to the marine environment).
-Since this doubles the number of lake-to-marine immigrants,
-after this happens the probability that a new marine individual has parents from old lakes or the new lakes
-is $2m$ instead of $m$.
-
-\paragraph{TreeSeq}
-This case study involves analyzing large simulations which model populations under the influence of selection. 
-To fully understand the origins of freshwater alleles, calculate statistics, and avoid the computationally expensive task of 
-simulating neutral mutations, we have used SLiM's \emph{treeSeq} feature to track the genealogical history of all samples throughout the simulation. 
-This feature outputs a compact representation the genomic outcome of every meiosis relevant to tracked samples (Often thought of as the ARG).
-
-Using the output tree sequence, 
-we were able to classify each freshwater allele in each genome of the new lakes population, at time of adaptation,
-as deriving from one of four origins:
-(1) Migrants, brought in from a migrant after the introduction of the new lakes 
-(2) Pre-existing, passed down from an individual in the original generation of the new lakes, 
-however, not defined to be an adaptive variant in the original lakes
-(3) Captured, Pre-existing, passed down from an individual in the original generation of the new lakes, 
-and defined to be adaptive in the original lakes
-(4) De novo, mutated and propagated from an individual existing in the introduced lakes population, after introduction. 
-\jgg{We should discuss further what these categories aught to be named}
-
-\begin{comment}
-We also used the tree sequence to count the "fitness" of individuals in the initial generation of the new lakes. 
+We also used the tree sequence to count the ``fitness'' of individuals in the initial generation of the new lakes. 
 The "fitness" of an individual in this case, is defined to be the total number off eventual offspring, in the generation at the time of adaptation, 
@@
 for each individual in the initial generation.
-\end{comment}
 
@@
 
+%%%%%%%%%%%%%%%%%%
 \section*{Results}
 
-We varied the migration rate, $m$ across separate simulations
-from $5 \times 10^{-5}$ to $5 \times 10^{-1}$ by one order of magnitude steps.
-For clarity given our populations sizes of 5,000 diploid individuals, 200 in each lakes,
-we will refer to these rates of gene flow as running simulations ranging from {0.01 - 100.0}
-migrants \emph{per} lake \emph{per} generation.
-The differing parameter values had a strong effect on many aspects of adaptation;
-including how fast adaptation occurred in each lake,
-how many alleles were shared between lakes,
+We varied the ocean--lake migration rate, $m$, 
+across separate simulations
+from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
+Since each lake contains 200 individuals,
+this corresponds to between 0.01 and 100 migrants per lake per generation.
+Many aspects of adaptation changed substantially across this range,
+including the speed of adaptation,
+degree of sharing of adaptive alleles between lakes,
 and the population genetic signals left behind.
-We find a threshold of gene flow which allows 
-the freshwater haplotype to be maintained in the marine without creating linkage to marine alleles. 
-below this threshold, selection on pre-existing freshwater alleles to be much less efficient 
-due to background selection. \jgg{did I get this right?}
+At very low rates of gene flow,
+each new lake's population adapted \emph{de novo},
+which took a very long time.
+At very high rates of gene flow, local adaptation was almost impossible.
+Between these two extremes,
+genetic variation that allowed adaptation to freshwater habitats
+could move relatively easily between lakes --
+despite being deleterious in the intervening marine habitat --
+which allowed populations arriving in new lakes to adapt quickly,
+reassembling collections of alleles responsible for freshwater adaptation
+in other lakes.
 
-Interestingly, all rates of $m$ aside from the two highest had little affect on population's ability to locally adapt.
-All rates of gene flow aside from the lowest, helped both the old lake's and the new lake's freshwater populations share pre-existing freshwater adapted alleles.
-The sharing of alleles at $m = 1.0$ resulted in rapid adaptation ($\approx 32$ generations) of the introduced population to adapt to the freshwater environment.
-At the lowest migration rate, lakes adapted nearly independently,
-while at the highest migration rate, the habitats act like a single metapopulation
-with substantial migration load.
-We will now dissect what happens between these extremes.
-
+%%%%%%%%%%%%
 \subsection*{Local Adaptation: differentiation with gene flow}
@@
         		Mean individual trait values in the marine habitat (blue line),
-        		the original lakes (light green lines; average in dark green),
-        		and the introduced lakes (yellow lines; average in orange)
+        		the original lakes (yellow lines; average in orange),
+        		and the new lakes (light green lines; average in dark green),
         		across the course of two simulations, with migration rates of
         		\textbf{(top)} $m=5 \times 10^{-5}$ and
-        		\textbf{(bottom)} $m=5 \times 10^{-4}$ per generation individual
+        		\textbf{(bottom)} $m=5 \times 10^{-4}$ per generation per individual
         		(i.e., 0.01 and 0.1 migrants per lake per generation, respectively).
         		Optimal trait values in the two habitats are at $\pm 10$.
+                Analogous plots for other migration rates
+                are shown in Figures \ref{fig:pheno_by_time1} and \ref{fig:pheno_by_time2}.
 		}
@@
 
-Local adaptation occurred in all simulations:
-as shown in Figure \ref{fig:phenotype_ts2},
+At the highest migration rate (at which half of each population was composed of migrants),
+populations had only slightly different average trait values.
+At lower migration rates (10 migrants per generation and below),
+populations adapted to local conditions:
+as shown in Figures \ref{fig:phenotype_ts2},
+\ref{fig:pheno_by_time1} and \ref{fig:pheno_by_time2},
 freshwater and marine populations diverged
@@
 The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
-as jumps in the mean trait value;
-in the two simulations these occur on a time scale of 100 (lower migration) to 1000 (higher migration) generations,
-\jgg{I'm not sure what this means? how do we know how long the jumps take?}
-and move the trait by an amount of order 1.
-Trait variation within each population was small compared to the difference between populations,
-with interquartile ranges of around XXX.\plr{TODO}
+as jumps in the mean trait value,
+which move the trait by an amount of order 1 every few hundred generations.
+Trait variation within each population was small compared to the difference between populations.
 Across all parameter values, differences at around 16 commonly polymorphic sites 
-(eight the shift the trait in each direction)
+(eight that shift the trait in each direction)
 were responsible for most of the adaptive differences between freshwater and marine habitats.
@@
 As expected, increasing migration rate decreased differentiation between habitats. 
-As seen in Figure \ref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases. 
-Local adaptation was still able to occur despite overall homogenization: 
-if computed using only sites with alleles affecting the trait (``effect mutations''), $F_{ST}$ between habitats was relatively constant across migration values.
+As seen in Figure \ref{fig:Fst}, 
+$F_{ST}$ between marine and freshwater habitats 
+at neutral sites steadily declines as migration increases. 
+However, local adaptation was still able to occur despite overall homogenization 
+at up to 10 migrants per lake per generation.
+% if computed using only sites with alleles affecting the trait (``effect mutations''),
+% $F_{ST}$ between habitats was relatively constant across migration values.
 
@@
   		\caption{
-		$F_{st}$ as a function of genomic location between all combinations of the three subpopulations; 
-		the old lakes, the new lakes, and the marine.
-		$F_{st}$ was calculated based upon allele frequencies in genomic windows of 500 mb. 
-		The four columns of subplots are separate simulations with increasing $m$. 
-		The three rows are the combinations of the three separate populations.
-		Ten vertical dotted pink lines in each subplot represent regions which have the potential to affect phenotype
-		 }
-  		\label{fig:Fst}
+		Average $F_{ST}$ in windows of 500Mb between:
+        \textbf{(top)} marine habitat and old lakes;
+        \textbf{(middle)} marine habitat and new lakes; and
+        \textbf{(bottom)} old and new lakes.
+        Each plot shows $F_{ST}$ values for a separate simulation,
+        with columns corresponding to increasing gene flow from left to right.
+		Ten vertical dotted pink lines in each subplot 
+        show regions which have the potential to affect phenotype.
+     } \label{fig:Fst}
 	\end{center}
@@
 
-\begin{comment}
-\paragraph*{Rate of adaptation}
-Increasing migration rate strongly decreased the time until freshwater habitats could adapt, both in the initial and introduced sets of lakes.
-As shown in Figure \ref{fig:TimeToAdaptation}, at the lowest migration rate it took over 20 thousand generations for the mean trait value across introduced lakes to get to within 0.5 of the original lakes value.
-Although many lakes had adapted before this time, the different rates of introduction of effect alleles is clearly seen in the traces of trait values against time (e.g., Figure \ref{fig:phenotype_ts2}).
-However, higher migration rates allowed the freshwater habitat to adapt nearly as quickly as the marine habitat.
-\end{comment}
-
 \paragraph*{Speed of adaptation}
+To quantify how quickly new lakes adapt,
+we measured the ``time to adaptation'',
+defined as the number of generations until 
+average trait values in old and new lakes were within 0.5 of each other.
 Increasing migration rate strongly decreased the time until freshwater habitats could adapt,
 both in the old and new sets of lakes.
-The different rates of introduction of effect alleles is clearly seen in the traces
-of trait values against time (e.g., Figure \ref{fig:phenotype_ts2}).
-As shown in Figure \ref{fig:TimeToAdaptation}, at the lowest migration rate 
-it took over $18,000$ generations for the mean trait value across introduced lakes to 
-get to within 0.5 of the original lakes value.
-At $m = 1.0$, we saw the introduced population adapt in just $32$ generations. 
-This rate of adaptation follows what we've observed in natural populations of threespine stickleback.
-At the highest migration rate, $m = 100$, all individuals were unable to locally adapt.
+Mean times to adaptation are shown across different rates of gene flow 
+in Figure \ref{fig:TimeToAdaptation}.
+Adaptation of new lakes took over 18,000 generations at the lowest migration rate,
+while at one migrant per lake per generation
+new lakes managed to adapt in only 32 generations.
 
@@
   		\caption{
-		Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
-		it takes for the introduced population's mean phenotype to come within 0.5 of the original lakes average phenotype. 
-		Each point represents a simulation run at some value of $M$. 
-		The yellow dashed line is the average of all points at each respective parameter value}
-  		\label{fig:TimeToAdaptation}
+		Time to adaptation as a function of migration rate.
+        The time to adaptation is measured as the number of generations until
+		the introduced population's mean phenotype 
+        comes within 0.5 of the original lakes average phenotype. 
+		Each point represents a single simulation run,
+		and the yellow dashed line is the average of all points at each respective parameter value.
+        (Adaptation did not occur at the highest rate of gene flow.)
+        \textbf{where are the multiple points?}
+        } \label{fig:TimeToAdaptation}
 	\end{center}
@@
 
-\subsection*{Sharing of freshwater alleles}
-
-\begin{comment}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %-MPAA / IND
-The tenfold difference in speed of adaptation occurs because at low migration rates, adaptation occurs independently in each lake,and the marine habitat has ten times the influx of new alleles than any one lake.
-In other words, greater mixing at higher migration rates allows lakes to share alleles instead of developing their own genetic basis of adaptation.
-As a first indication of this, Figure \ref{fig:Fst} shows that $F_{ST}$ between the ``original'' and ``introduced'' sets of lakes at effect mutations decreased with migration rate.
+The tenfold difference in speed of adaptation occurs because at low migration rates, 
+adaptation occurs independently in each lake,
+and the marine habitat has ten times the influx of new alleles than any one lake. 
+In other words, greater mixing at higher migration rates 
+allows lakes to share alleles instead of developing their own genetic basis of adaptation. 
+As a first indication of this, Figure \ref{fig:Fst} 
+shows that $F_{ST}$ between the ``original'' and ``introduced'' sets of lakes 
+at effect mutations decreased with migration rate.
 
@@
 Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
-\end{comment}
 
@@
 
-\begin{comment}
 At first, increased migration allows sharing of adaptive alleles between lakes, but at the highest migration rate, the constant influx of alleles between the habitats creates substantial migration load.
@@
 However, it is striking that at $m=.005$, the marine habitat carries as much genetic variation, despite a lack of any substantial migration load.
-
-\end{comment}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{comment}
 \subsection*{Realized genomic architecture}
@@
 This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.
-\end{comment}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \subsection*{Realized genomic architecture}
@@
 
+\begin{figure}
+	\begin{center}
+        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-2.png}
+        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
+  		\caption{
+            \textbf{Mean phenotypes through time},
+            at migration rates of
+            \textbf{(top)} 10 and
+            \textbf{(bottom)} 1 migrants per lake per generation.
+		}
+  		\label{fig:pheno_by_time1}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
+        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-5.png}
+  		\caption{
+            \textbf{Mean phenotypes through time},
+            at migration rates of
+            \textbf{(top)} 0.1 and
+            \textbf{(bottom)} 0.01 migrants per lake per generation.
+		}
+  		\label{fig:pheno_by_time2}
+	\end{center}
+\end{figure}
+
 \end{document}

### 2018-12-08 Stickleback_Paper.tex
.

@@
 \paragraph{Recording genealogical history}
-We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treeseq}
+We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence}
 to record the genealogical history of all individuals alive 

### 2018-12-08 Stickleback_Paper.tex
some more editing...

@@
 (all individuals are hermaphroditic),
-and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation 
+and the contributing genomes are produced by Poisson recombination 
+with an average of one crossover per chromosome per generation 
 ($10^{-7}$ per locus per generation). 
@@
 to keep population sizes roughly constant within each lake, 
-before selection we divide fitness values of each freshwater individual by the mean fitness in their lake,
+before selection we divide fitness values of each freshwater individual 
+by the mean fitness in their lake,
 so that the mean fitnesses of all lakes are equal.
@@
 We quantify genetic differentiation between the habitats with per-locus $F_{ST}$:
-if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively,
-and $\bar p = (p_f + p_m)/2$, then $F_{ST} = 1 - (p_f (1-p_f) + p_m (1-p_m))/ (2 \bar p (1-\bar p))$.
-
-We also quantify the degree to which adaptation in new lakes uses alleles 
-already present and specific to the pre-existing lakes,
-defining a \emph{pre-existing freshwater adapted allele}
-to be a mutation (with an effect on the trait)
-that at the time of introduction of the new populations
-has frequency higher than $0.5$ in at least one of the original lakes, 
-while remaining lower than $0.5$ in the marine habitat. 
-These alleles represent standing variation (rather than new mutations),
-but furthermore are presumably responsible for local adaptation
-in at least some lakes,
-and present only at migration-selection balance in the ocean.
-Alleles common in the newly introduced lakes do not count
-if they are not also common in the original lakes. 
-They are defined this way because 
-the transportation hypothesis does not specify where or when an advantageous mutation arises, 
-but simply suggests that any sufficiently common freshwater adapted allele 
-could participate in adaptation in new habitat \citep{schluter2009genetics}.
-\plr{I think we use this definition with respect to different times? 
-     But I don't see where we do this.}
+if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats,
+respectively,
+and $\bar p = (p_f + p_m)/2$,
+then $F_{ST} = 1 - (p_f (1-p_f) + p_m (1-p_m))/ (2 \bar p (1-\bar p))$.
 
@@
 Using these trees,
-we could therefore classify each freshwater allele found in the new lakes at time of adaptation
+we classified each freshwater allele found in the new lakes at time of adaptation
 into four categories:
-\begin{itemize}
-    \item \emph{Migrants}, not present in the set of individuals that initially colonized
-        the new lake, but brought in by migration at a later time;
-    \item \emph{Pre-existing}, passed down from an individual in the original generation of the new lakes, 
-however, not defined to be an adaptive variant in the original lakes
-    \item \emph{Captured}, Pre-existing, passed down from an individual in the original generation of the new lakes, 
-and defined to be adaptive in the original lakes
-    \item \emph{De novo}, mutated and propagated from an individual existing in the introduced lakes population, after introduction. 
-\end{itemize}
+\begin{enumerate}
+    \item \emph{De novo alleles:}
+        deriving from a new mutation that occurred in a new lake.
+    \item \emph{Migrant alleles:} 
+        not present in the individuals that initially colonized the new lake, 
+        but brought in by migration at a later time.
+    \item \emph{Captured alleles:}
+        present in the individuals that initially colonized the new lake, 
+        and both common (above 50\%) in the original lakes,
+        and uncommon (below 50\%) in the ocean.
+    \item \emph{Marine alleles:} 
+        present in the individuals that initially colonized the new lake, 
+        and not a captured allele.
+\end{enumerate}
+Roughly speaking, these four categories describe the contribution of
+(1) new mutation, 
+(2) post-colonization migration, 
+(3) standing variation at migration--selection balance, and
+(4) standing variation at mutation--selection balance.
+We could partition migrant alleles into finer categories
+(e.g., we do not distinguish whether they are common in the original lakes or not),
+but this category turns out to be small.
 
@@
 
+\begin{figure}
+	\begin{center}
+  		\includegraphics{Final_Plots/Pheno_Dist.pdf}
+  		\caption{
+            Distribution of mean trait values across generations of the simulation,
+            for different migration rates.
+            The dashed pink and purple lines at $\pm 10$ give the optimum phenotypes
+            in the marine and freshwater environments, respectively.
+		}
+  		\label{fig:MeanPhenotype}
+	\end{center}
+\end{figure}
+
+Local adaptation occurred at most migration rates,
+as shown in Figure~\ref{fig:MeanPhenotype}.
 At the highest migration rate (at which half of each population was composed of migrants),
@@
 As expected, increasing migration rate decreased differentiation between habitats. 
-As seen in Figure \ref{fig:Fst}, 
+As seen in Figure~\ref{fig:Fst}, 
 $F_{ST}$ between marine and freshwater habitats 
@@
 \paragraph*{Speed of adaptation}
-To quantify how quickly new lakes adapt,
-we measured the ``time to adaptation'',
-defined as the number of generations until 
-average trait values in old and new lakes were within 0.5 of each other.
-Increasing migration rate strongly decreased the time until freshwater habitats could adapt,
+Adaptation occurred much more quickly
+at higher migration rates,
 both in the old and new sets of lakes.
-Mean times to adaptation are shown across different rates of gene flow 
-in Figure \ref{fig:TimeToAdaptation}.
-Adaptation of new lakes took over 18,000 generations at the lowest migration rate,
+We measured this ``time to adaptation''
+as the number of generations until 
+average trait values in old and new lakes were within 0.5 of each other,
+shown in Figure \ref{fig:TimeToAdaptation} for different rates of gene flow.
+Adaptation of new lakes took over 18,000 generations at the lowest rate,
 while at one migrant per lake per generation
@@
 %-MPAA / IND
-The tenfold difference in speed of adaptation occurs because at low migration rates, 
+At low migration rates, the \emph{initial} period of adaptation
+takes roughly ten times longer for lakes as it does for the ocean.
+This difference occurs because at low migration rates, 
 adaptation occurs independently in each lake,
-and the marine habitat has ten times the influx of new alleles than any one lake. 
-In other words, greater mixing at higher migration rates 
+and the marine habitat has ten times as many individuals,
+and therefore ten times the influx of new alleles than any one lake. 
+At higher migration rates, greater mixing
 allows lakes to share alleles instead of developing their own genetic basis of adaptation. 
@@
 
-To further investigate sharing of locally adaptive alleles and the transporter hypothesis, we investigate the ``pre-existing freshwater adapted alleles'',
-defined for a particular generation to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat in that generation.
-Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles, across generations, as well as the average number of lakes each is found in.
-As migration rates increase, the number of these alleles decreases steadily, and each is concurrently found in a greater number of lakes.
+To investigate in more depth how locally adaptive alleles found in the original lakes
+are shared between lakes and spread to the new lakes,
+we counted ``pre-existing freshwater adapted alleles''.
+These we define for a particular generation 
+to be any mutation with an effect on the trait
+whose frequency is above 50\% in at least one original lake 
+and below 50\% in the marine habitat.
+Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles, 
+across generations.
+At the lowest migration rate (one migrant per 100 generations per lake), 
+each lake has a private set of about 10 mutations
+nearly fixed in that lake but not elsewhere:
+new lakes acquire new adaptive alleles and so have none of these.
+At one migrant per 10 generations per lake,
+the original set of lakes each developed private alleles,
+but these are present at low frequency in the ocean,
+and so new lakes adapt 
+mostly using a subset of this large repertoire of standing variation.
+As migration rate increases past this, 
+the number of pre-existing freshwater adapted alleles declines
+because alleles move between populations by migration
+before they can appear by new mutation,
+while the frequency of these alleles in the ocean
+stays relatively constant.
 
-As the number of common, locally adaptive alleles decreases, the genetic basis of adaptation is more commonly shared.
-Figure \ref{fig:MPFAI}B shows the distribution of the mean percentage of currently-defined freshwater adapted alleles that each genome in each of the populations carries, averaged across time and individuals.
-If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%.
-This value is nearly reached at the highest migration rate; but it is lower due to migration load.
-At the lowest migration rate, each genome in the original lakes have almost exactly $1/10^{th}$ of the total freshwater adapted alleles -- since there are 10 lakes, this implies that each lake has adapted with a unique set of alleles.
+Figure~\ref{fig:MPFAI}B shows the distribution of 
+the mean percentage of currently-defined freshwater adapted alleles 
+that each genome in each of the populations carries, 
+averaged across time and individuals. 
+If all individuals across lakes carried the same set of alleles determining their trait value, 
+this would be 100\%. 
+At the lowest migration rates,
+each genome in the original lakes have almost exactly~$1/25^{th}$ 
+of the total number of pre-existing freshwater adapted alleles -- 
+this is because each of the 25 lakes has adapted with a unique set of alleles. 
 Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
 
-We find that this dramatic increase in speed of local adaptation occurs because
-higher gene between populations allow sharing of freshwater alleles more effectively when compared to de novo mutation.
+The numbers of Figure~\ref{fig:MPFAI} suggest that
+dramatic increase in speed of local adaptation we observed above occurs because
+higher gene between populations allow sharing of freshwater alleles more effectively 
+when compared to \textit{de novo} mutation.
+We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting
+allele common in the new lakes at time of adaptation,
+as defined in the Methods.
 At low migration rates, adaptation occurs independently in each lake.
-In Figure \ref{fig:Origin}, at $0.01$, we can see that the large majority of adaptive alleles derived from de novo mutation in the 
+In Figure~\ref{fig:Origin}, at $0.01$, we can see that the large majority of adaptive alleles derived from de novo mutation in the 
 population rather than selection on standing variants in the ocean. 
@@
   		\caption{
-		Distribution of freshwater alleles among the three populations. 
-		A freshwater adapted variant is defined as, at any generation throughout the simulation,
-		a variant affecting phenotype that is greater than 0.5 frequency in any of the old lakes, 
-		and less than 0.5 frequency in the marine population. 
-		Subplot $A$ is the distribution throughout the simulation, of the mean number of
-		freshwater adapted variants per individual, in each population.
-		Subplot $B$ is the distribution throughout the simulation, of the mean \emph{percentage} of
-		freshwater adapted variants per individual, in each population. 
-		Subplot $C$ is the distribution throughout the simulation, of the total number of
-		freshwater adapted alleles defined at any one generation.
-		In other words, $B  =  A  /  C$.
+            Amount of standing freshwater variation by habitat, across migration rates.
+            Each plot counts ``pre-existing freshwater adapted alleles'',
+            that are common in the original lakes but rare in the ocean
+            (see text for definition).
+            \textbf{(A)} Mean number of these alleles per individual.
+            \textbf{(B)} Mean percentage of these alleles per individual.
+            \textbf{(C)} Total number of these alleles (so, $B = A/C$).
+            The number of alleles meeting these conditions changes over the course of the simulation,
+            and each plot shows distributions of these values across generations.
+            The horizontal axis shows $M$, the mean number of migrants per lake per generation.
 		}
@@
 at $m=0.5$, all populations experience panmixia
-
-\begin{figure}
-	\begin{center}
-  		\includegraphics{Final_Plots/Pheno_Dist.pdf}
-  		\caption{Distribution of mean trait value throughout simulation runs at separate migration rate $(m)$ parameter values, for each population.
-		The dashed pink line (Trait value = +10) is the optimum phenotype any individual in the marine environment.
-		In contrast the purple line at (Trait value = -10) represents the optimum for any individual in the freshwater environment. 
-		All individuals at generation 0 (beginning of the simulation) 
-		}
-  		\label{fig:MeanPhenotype}
-	\end{center}
-\end{figure}
 
@@
   		\caption{ 
-		Origin ratio of adaptive alleles in all genomes at time of local adaptation for different values of $m$
-		For all alleles at a frequency $> 0.5$ in any one of the new lakes individuals at time of adaptation, 
-		we traced the origin of the segment of DNA back in time to classify them as 
-		\textbf{(Blue)} Migrants, brought in from a migrant after the introduction of the new lakes 
-		\textbf{(Orange)} Pre-existing, passed down from an individual in the original generation of the new lakes, 
-		however, not defined to be an adaptive variant in the original lakes
-		\textbf{(Green)} Captured, Pre-existing, passed down from an individual in the original generation of the new lakes, 
-		and defined to be adaptive in the original lakes
-		\textbf{(Red)} De novo, mutated and propagated from an individual existing in the introduced lakes population, after introduction. 
+        \textbf{(Origin of adaptive alleles:)}
+        Each bar plot shows the origins
+		of all trait-affecting alleles above frequency 50\% in at least one new lake,
+        classified as
+		\textbf{(red)} new mutations,
+		\textbf{(blue)} post-colonization migrants,
+		\textbf{(green)} ``captured'' from pre-existing lakes, or
+		\textbf{(orange)} standing marine variation.
+        See Methods for precise definitions of these categories.
 		}

### 2018-12-09 Stickleback_Paper.tex
.

@@
 \end{align*}
-Note that only the the scale of trait values relative to mutation effect size is relevant.
+Note that only the scale of trait values relative to mutation effect size is relevant.
 We chose the difference between optima and strength of stabilizing selection in each habitat
@@
             is coupled to randomly mating ``lakes'' at discrete locations.
-            After an inital period of 100K generations with 25 lakes,
+            After an initial period of 100K generations with 25 lakes,
             an additional 25 lakes are added (at the same set of locations)
@@
 first, a single parent is chosen proportionally to fitness in the marine habitat,
-and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with stanadard deviation $1/2$. 
+and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. 
 Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
@@
 
-We also used the tree sequence to count the ``fitness'' of individuals in the initial generation of the new lakes. 
-The "fitness" of an individual in this case, is defined to be the total number off eventual offspring, in the generation at the time of adaptation, 
-which have inherited a adaptive variant from that individual. 
-To calculate this, we only considered gene trees which overlapped an adaptive variant site (defined at time of adaptation).
-We then counted the total number of offspring which had inherited from any adaptive variant site, 
-for each individual in the initial generation.
+%% we did this, but it is not used in the results
+% We also used the tree sequence to measure the realized fitness of individuals 
+% in the initial generation of the new lakes,
+% measured as
+% the total number of eventual offspring, in the generation at the time of adaptation, 
+% which have inherited a adaptive variant from that individual. 
+% To calculate this, we only considered gene trees which overlapped an adaptive variant site (defined at time of adaptation).
+% We then counted the total number of offspring which had inherited from any adaptive variant site, 
+% for each individual in the initial generation.
 
-We were also able to use the tree sequence to get information about the individual ``effect regions" 
+We were also able to use the tree sequence to get information about the individual ``effect regions''
 in all initial genomes of the new lakes at time of introduction.
@@
 To investigate in more depth how locally adaptive alleles found in the original lakes
-are shared between lakes and spread to the new lakes,
+are shared between lakes, and how they spread to the new lakes,
 we counted ``pre-existing freshwater adapted alleles''.
@@
 Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
+Figure \ref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above,
+the average individual across the new lakes has nearly the same amount of 
+pre-existing freshwater adapted alleles as individuals across the old lakes.
+As expected, the genetic basis of the freshwater phenotype 
+seems to simplify as migration increases --
+higher rates of migration allow
+adaptive alleles of higher effect to travel more efficiently through the population,
+even though they are deleterious in the ocean.
 
 The numbers of Figure~\ref{fig:MPFAI} suggest that
-dramatic increase in speed of local adaptation we observed above occurs because
-higher gene between populations allow sharing of freshwater alleles more effectively 
-when compared to \textit{de novo} mutation.
+the dramatic increase in speed of local adaptation we observed above occurs because
+higher gene flow between populations allows sharing of freshwater alleles between populations.
 We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting
@@
 as defined in the Methods.
-At low migration rates, adaptation occurs independently in each lake.
-In Figure~\ref{fig:Origin}, at $0.01$, we can see that the large majority of adaptive alleles derived from de novo mutation in the 
-population rather than selection on standing variants in the ocean. 
-As The migration increases, we see a larger ratio of the origins of adaptive alleles 
+Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow,
+the majority of adaptive alleles are derived from de novo mutation.
+As gene flow increases, a larger fraction of adaptive alleles 
 derive from pre-existing variation in the marine population at the time of introduction.
@@
 instead of developing their own genetic basis of adaptation.
-To further investigate sharing of locally adaptive alleles and the transporter hypothesis,
-we investigate the ``freshwater alleles'', defined for a particular generation
-to be effect mutations above 50\% in at least one original lake and below 50\% in the marine habitat in that generation.
-
-As the number of common, locally adaptive alleles decreases, as seen in Figure \ref{fig:MPFAI}C,
-the genetic basis of adaptation is more commonly shared.
-Figure \ref{fig:MPFAI}B shows
-the percentage of currently-defined, freshwater-adapted alleles 
-that each genome in each of the populations carries,
-averaged across time and individuals.
-If all individuals across lakes carried the same set of alleles determining their trait value,
-this would be 100\%.
-This value is nearly reached at the fourth migration rate;
-but it is lower due to migration load.
-At the lowest migration rate, 
-each genome in the original lakes have almost exactly $1/25^{th} = 0.04\%$   
-of the total freshwater adapted alleles --
-since there are 25 lakes, this implies that each lake has adapted with a unique set of alleles.
-Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
-Figure \ref{fig:MPFAI}A shows us that by $0.1$ migrants per lake per generation,
-the average individual across the new lakes has nearly the same amount of 
-``pre-existing freshwater adapted alleles" as individuals across the old lakes.
-Interestingly, the genetic basis of the freshwater phenotype 
-seems to simplify with greater $m$. 
-This would seem to suggest higher rates of migration allowed 
-adaptive alleles of higher effect to travel more efficiently through the populations
-\jgg{is this right?}
 
@@
 
-At first, increased migration allows sharing of adaptive alleles between lakes, but at the highest migration rate, the constant influx of alleles between the habitats creates substantial migration load.
-This rate, $m=.05$, only replaces 5\% of each population each generation with migrants from the other habitat, but this is sufficient to shift the mean trait values to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
-Significant gene flow constricts local adaptation as a consequence of a large number of offspring through hybridization events between subpopulations.
+At first, increased migration allows sharing of adaptive alleles between lakes, 
+but at the highest migration rate, 
+the constant influx of alleles between the habitats creates substantial migration load. 
+The rate at which migration load becomes substantial,
+10 migrants per lake per generation,
+only replaces 5\% of each population each generation with migrants from the other habitat, 
+but this is sufficient to shift the mean trait values to nearly half their optimal values, 
+as seen in Figure \ref{fig:MeanPhenotype}.
 
-\paragraph{Standing genetic variation} Despite a dramatic difference in the amount of allelic sharing between lakes, standing genetic variance (Figure \ref{fig:SGV}) was around 0.05 in the freshwater habitat across all but the highest migration rates.
-Concurrently, genetic variation in the marine population steady increased with migration
-to a similar value.
-On the one hand, it is not surprising that lakes, as a group, carry more genetic variation than the marine habitat, since population subdivision allows different alleles to become common in each.
-However, it is striking that at $m=.005$, the marine habitat carries as much genetic variation, despite a lack of any substantial migration load.
+\paragraph{Standing genetic variation} 
+Despite a dramatic difference in the amount of allelic sharing between lakes, 
+standing genetic variance (Figure \ref{fig:SGV}) 
+was around 0.05 in the freshwater habitat across all but the highest migration rates. 
+Concurrently, genetic variation in the marine population 
+steadily increased with migration to a similar value. 
+On the one hand, it is not surprising that lakes, as a group, 
+carry more genetic variation than the marine habitat, 
+since population subdivision allows different alleles to become common in each. 
+However, it is striking that at $m=.005$, 
+the marine habitat carries as much genetic variation, 
+despite a lack of any substantial migration load.
+\plr{perhaps we want to remove this bit?}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Realized genomic architecture}
+\subsection*{Realized genetic architecture}
 Now we take a closer look at the genomic architecture of local adaptation between the two habitats. Do the alleles underlying trait differences cluster along the genome?
@@
 
+\begin{figure}
+	\begin{center}
+        \includegraphics[width=1.0\linewidth]{plotlyPlots/StandingGeneticVariation.png}
+  		\caption{
+            \textbf{Standing genetic variance},
+            computed as XXX write if we leave this in XXX.
+		}
+  		\label{fig:pheno_by_time2}
+	\end{center}
+\end{figure}
+
 \end{document}

### 2018-12-09 Stickleback_Paper.tex
added r code

@@
 Mutations that can affect the trait under selection can occur at rate $10^{-7}$ per locus per generation
-in ten ``effect regions'' of $100$ loci each,
-spread evenly along the chromosome (separated by 4950 loci).
+in ten ``effect regions'' of $10^5$ loci each,
+spread evenly along the chromosome.
 Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability), 
@@
 	\begin{center}
-  		\includegraphics{Final_Plots/Fst_Genome.pdf}
+        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome.pdf}
   		\caption{
@@
 
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=0.7\linewidth]{Final_Plots/Allele_Origin.pdf}
+  		\caption{ 
+        \textbf{(Origin of adaptive alleles:)}
+        Each bar plot shows the origins
+		of all trait-affecting alleles above frequency 50\% in at least one new lake,
+        classified as
+		\textbf{(red)} new mutations,
+		\textbf{(blue)} post-colonization migrants,
+		\textbf{(green)} ``captured'' from pre-existing lakes, or
+		\textbf{(orange)} standing marine variation.
+        See Methods for precise definitions of these categories.
+		}
+  		\label{fig:Origin}
+	\end{center}
+\end{figure}
+
 At first, increased migration allows sharing of adaptive alleles between lakes, 
@@
 despite a lack of any substantial migration load.
-\plr{perhaps we want to remove this bit?}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection*{Migration load and genetic variation}
-
-At first, increased migration allows sharing of adaptive alleles between lakes,
-but at $m = 5 \times 10^{-2}$
-the constant influx of alleles between the habitats creates noticeable migration load.
-This rate, $m=.05$, only replaces 5\% of each population each generation
-with migrants from the other habitat, but this is sufficient to shift the mean trait values
-to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
-Significant gene flow constricts local adaptation
-as a consequence of a large number of offspring through hybridization events between subpopulations.
-at $m=0.5$, all populations experience panmixia
+\plr{perhaps we want to remove this bit? revisit if it is telling us something important.}
 
@@
 \subsection*{Realized genetic architecture}
-Now we take a closer look at the genomic architecture of local adaptation between the two habitats. Do the alleles underlying trait differences cluster along the genome?
+
+Now we take a closer look at the genomic architecture of local adaptation between the two habitats. 
+Do the alleles underlying trait differences cluster along the genome?
 Do measures of local differentiation identify the causal loci? 
-Figure \ref{fig:Fst3} shows plots along the genome of per-locus $F_{ST}$ values between the marine and original freshwater habitats at the two intermediate migration rates.
+Figure \ref{fig:Fst} shows plots along the genome of per-locus $F_{ST}$ values between habitats.
 (Note that we are pooling freshwater habitats; a single lake would provide substantially less power.)
+\plr{need to know where the FAA are in these figures}
+Higher migration rates showed more distinct $F_{ST}$ peaks over polymorphic loci 
+underlying trait differences between the habitats. 
+``Background'' levels of $F_{ST}$ increase as gene flow decreases,
+swamping out this signal until the regions under selection are indistinguishable. 
+This is likely due to two reasons: 
+first, stronger genetic drift with less migration leads to higher background $F_{ST}$, and 
+second, greater sharing of adaptive alleles providing a shared signal across populations.
 
-\plr{need to know where the FAA are in these figures}
+This suggests that genome scans for local adaptation based purely on measures of differentiation 
+will only be successful given enough migration between habitats. 
+To quantify this, 
+Figure \ref{fig:Power_FP} shows the power and false positive rates 
+that would be obtained by an $F_{ST}$ cutoff 
+that declared everything above a certain value to be a causal locus.
 
-Higher migration rates showed more distinct $F_{ST}$ peaks over polymorphic loci underlying trait differences between the habitats.
-As migration rate decreased, the ``background'' levels of $F_{ST}$ increased, swamping out this signal until the regions under selection were indistinguishable.
-This is likely due to two reasons: first, stronger genetic drift with less migration leading to higher background $F_{ST}$, and second, greater sharing of adaptive alleles providing a shared signal across populations.
-
-This suggests that genome scans for local adaptation based purely on measures of differentiation will only be successful given enough migration between habitats.
-To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.
-
-\plr{Labels are wrong on that figure: false pos rate should go down with cutoff.
-     Is it true positive rate?}
+\plr{Need to fix caption on figure \ref{fig:Power_FP} before editing this bit.}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Realized genomic architecture}
+%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Origin of introduced freshwater adaptive alleles}
 
-Higher migration rates showed more distinct $F_{ST}$ peaks
-over polymorphic loci underlying trait differences between the habitats.
-As migration rate decreased, the ``background'' levels of $F_{ST}$ increased,
-swamping out this signal until the regions under selection were indistinguishable. 
-This is likely due to two reasons: 
-first, stronger genetic drift with less migration leading to higher background $F_{ST}$,
-and second, greater sharing of adaptive alleles providing a shared signal across populations.
-This suggests that genome scans for local adaptation
-based purely on measures of differentiation
-will only be successful given enough migration between habitats.
-With 10 effect regions across the genome, we can see in \ref{fig:Fst} that only about 1 or 2 of the region contain peaks. 
-When looking closely at the peaks we see that each is usually a composition of 7-8 SNPs close in proximity.
-Because some of those mutations are selectively neutral (we can see this because the counts of FAA are not as high as the number in the clusters),
-this is suggestive of hitchhiking alleles along with the SNPs being brought to high frequency by a selective sweeps.
+We have thus far found that the speed of adaptation depends strongly
+on the degree to which alleles can be shared between populations.
+However, there remains a mystery.
+Figure~\ref{fig:MPFAI}A shows that with $m = 0.1$ migrants per lake per generation, 
+there is substantial sharing of alleles between populations,
+and yet Figure~\ref{fig:TimeToAdaptation} shows that new lakes 
+take around 2,000 generations to adapt.
+This is surprising because we see a similar quantity of allele sharing at  $m = 1.0$, 
+but much more rapid adaptation ($\approx 30$ generations). 
+Perhaps
+this 60$\times$ discrepancy occurs because 
+new lakes must wait for an additional influx of freshwater alleles through migration
+beyond what is present in the initial generation?
+However, Figure~\ref{fig:Origin} shows that at both $m = 0.1$ and $m = 1.0$,
+the vast majority of alleles underlying the local adaptation
+derived from individuals in the original generation (``captured'' or ``pre-existing'').
 
-Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$?
-Knowing exactly which mutations effect phenotype in our simulations, 
-we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome. 
-In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, 
-we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power. 
-This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.
-
-\subsubsection*{Origin of introduced freshwater adaptive alleles}
-Up to this point, we have quantified rapid adaptation and found strong evidence that sharing of alleles is responsible in this system.
-Interestingly, in Figure \ref{fig:MPFAI}A at $m = 0.1$, we see sharing of alleles between populations,
-yet \emph{Time to adaptation} is still fairly slow ($\approx 2,000$ gens).
-This is surprising when considering we see a similar quantity of allele sharing at  $m = 1.0$, 
-but much more rapid adaptation ($\approx 30$ gens). 
-Initially, this 60X speedup of adaptation seems to be the result of subsequent gene flow of freshwater alleles. 
-Perhaps a larger influx of freshwater alleles are needed in addition to standing variation in the initial generation.
-To investigate this further we have tracked the origins of all freshwater alleles, in all new lake individuals at the time of adaptation. 
-Surprisingly in Figure \ref{fig:Origin}, we find that at both $m = 0.1$ and $m = 1.0$,
-the vast majority of adaptive alleles were found to propagate from individuals in the original generation ("Captured" OR "Pre-Existing").
-
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{Final_Plots/Allele_Origin.pdf}
-  		\caption{ 
-        \textbf{(Origin of adaptive alleles:)}
-        Each bar plot shows the origins
-		of all trait-affecting alleles above frequency 50\% in at least one new lake,
-        classified as
-		\textbf{(red)} new mutations,
-		\textbf{(blue)} post-colonization migrants,
-		\textbf{(green)} ``captured'' from pre-existing lakes, or
-		\textbf{(orange)} standing marine variation.
-        See Methods for precise definitions of these categories.
-		}
-  		\label{fig:Origin}
-	\end{center}
-\end{figure}
-
-The alternative explanation, is that at a lower migration rate, a higher saturation of marine adapted alleles are linked to freshwater adapted alleles.
-In turn, the equilibrium of freshwater alleles in marine genomes created by migration-selection balance acts as a barrier when rebuilding the freshwater haplotype. 
-%With more gene flow between the two selective pressures, the freshwater haplotype remains more "intact" and much more efficient to rebuild in the introduced environment.
-To confirm this, we decided to look at the trait values of the ten affect regions in each genome of the initial population.
-Given the probability that an ``effect region" fixes at a given trait value, $2 \times x / 45$ where $x$ is the effect size of a haplotype,
+An alternative explanation is that at lower rates of gene flow,
+freshwater alleles present as standing variation in the marine habitat
+are more tightly linked to marine alleles,
+and so adaptation in the new lakes must wait for recombination.
+This might be expected since freshwater alleles present in the ocean,
+to remain present at migration-selection balance at low migration rates,
+must be ``masked'' by nearby compensatory alleles.
+Reversing this ``masking'' then slows the process of adaptation that uses this genetic variation.
+Said another way,
+higher gene flow from freshwater to the ocean
+maintains relatively intact freshwater haplotypes
+that can be more easily rebuilt.
+To confirm this, we calculated for each individual in the initial generation of the new lakes
+the total contribution to trait values of each of the ten genomic regions 
+that containing trait-affecting mutations.
+The recombination rate within each of these ten regions is $10^{-2}$,
+so these regions act initially more-or-less like single loci.
+If a region moves the trait an average of $x$ units towards the freshwater optimum
+(averaging over dominance relationships),
+Given the probability that an ``effect region'' fixes at a given trait value, 
+$2 \times x / 45$ where $x$ is the effect size of a haplotype,
 we were able to determine expected total effect sizes of the haplotypes that fix (detail about computation in methods).
@@
 \appendix
-\section*{Supp.}
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+\section*{Supplementary material}
+
 \begin{figure}
@@
 		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
-        		TODO: x-axis label ($F_{ST}$).
+            \textbf{
+                TODO: fix this caption; x-axis label ($F_{ST}$).}
 		}
@@
 		}
-  		\label{fig:pheno_by_time2}
+        \label{fig:SGV}
 	\end{center}

### 2018-12-09 r_calcs/plot_linkage.R

@@ -1,0 +1,152 @@
+library(matrixStats)
+library(xtable)
+
+everything <- lapply(0:3, function (run_num) {
+
+    nodes <- read.table(sprintf("nodes_positions_%d.txt", run_num))
+    colnames(nodes) <- c("id", "location")
+    nodes$pop <- cut(nodes$location, 25)
+    if (FALSE) plot(nodes$location, col=nodes$pop) # sanity check
+
+    # has one row per genome and one column per variant
+    genotypes <- matrix(scan(sprintf("genotype_matrix_%d.txt", run_num)), nrow=nrow(nodes))
+
+    # initializeMutationType("m1", 0.5, "g", 0.5, 1.0);  //Positive Direction, Dominant 
+    # initializeMutationType("m2", 0.5, "g", -0.5, 1.0); //Negative Direction, Dominant
+    # //initializeMutationType("m3", 0.5, "f", 0.0);       //Neither Direction, Neutral
+    # initializeMutationType("m4", 0.5, "g", 0.5, 1.0);  //Positive Direction, Recessive
+    # initializeMutationType("m5", 0.5, "g", -0.5, 1.0); //Negative Direction, Recessive
+    # initializeMutationType("m6", 0.5, "g", 0.6, 1.0);  //Positive Direction, Additive
+    # initializeMutationType("m7", 0.5, "g", -0.6, 1.0); //Negative Direction, Additive
+
+    variants <- read.table(sprintf("variants_%d.txt", run_num))
+    colnames(variants) <- c("id", "size", "type", "pos")
+    stopifnot(all(variants$type != 3))
+    stopifnot(ncol(genotypes) == nrow(variants))
+    variants$freq <- colMeans(genotypes)
+    variants$region <- cut(variants$pos, 10)
+    variants$direction <- ifelse(variants$type %in% c(1, 4, 6), +1, -1)
+    stopifnot(all(sign(variants$size) == variants$direction))
+    variants$dominance <- c(1, 1, NA, 0, 0, 0.5, 0.5)[variants$type]
+
+    # now we compute net effects by effect region and genome
+    effects <- genotypes * variants$size[col(genotypes)]
+    region_effects <- t(apply(effects, 1, tapply, variants$region, sum))
+    nodes$phenotype <- rowSums(region_effects)
+
+    # distribution of per-genome phenotypes
+    if (FALSE) hist(nodes$phenotype)
+
+    pos_effects <- t(apply(ifelse(effects > 0, effects, 0), 1, tapply, variants$region, sum))
[…]

### 2018-12-09 Stickleback_Paper.tex
.

@@
 so these regions act initially more-or-less like single loci.
+Even though the trait is additive across loci,
+the presence of sufficiently many distinct alleles to attain, in principle,
+the optimal freshwater trait value
+does not ensure adaptation will actually occur with these alleles.
+On the contrary, many of these will likely be lost to drift.
+To get a rough idea of whether the genetic variation present in the initial population
+is sufficient for the population to adapt,
+we translated the total effect size $x$ of each region 
+to a probability of fixation
+(by $p(x) \approx 2x/45$),
+and then calculated $\sum_i x_i p(x_i)$.
 If a region moves the trait an average of $x$ units towards the freshwater optimum

### 2018-12-09 Stickleback_Paper.tex
lots of words about prob of fixation that we do not need now, grr

@@
 %\doublespacing
+
+\newcommand{\E}{\mathbb{E}}
+\renewcommand{\P}{\mathbb{P}}
 
@@
 that can be more easily rebuilt.
-To confirm this, we calculated for each individual in the initial generation of the new lakes
+Recall that trait-affecting mutations only occur in relatively small regions of $10^5$ loci
+in which recombination occurs at rate $10^{-2}$.
+Even if there is a sufficient amount of variation in the initial population of a lake
+to shift the trait from the marine optimum ($+10$) to the freshwater optimum ($-10$),
+these may be tightly linked to alleles that counteract their effect.
+For example, suppose there are 10 variants segregating at low frequency with effect size $-1$ each,
+but each is paired with a compensatory allele with effect size $+1$.
+Each local haplotype is therefore neutral, 
+and so likely to be at higher frequency in the marine population.
+
+Suppose that the alleles present in the initial population of a new lake
+have effect sizes $(x_1, \ldots, x_n)$,
+and that the probability of fixation of a single allele of effect size $x$ is $p(x)$.
+Let $X_i$ be the average per capita effect of the $i^\text{th}$ allele after fixation,
+so that $X_i = 0$ if the allele is lost,
+and $X_i = 2x_i$ if the allele fixes (ignoring dominance).
+Therefore, the expected contribution to the final trait value of allele $i$
+is $\E[X_i] = 2 x_i p(x_i)$.
+The trait value after the initial round of adaptation is $\sum_i X_i$,
+and so the expected trait value is $2 \sum_i x_i p(x_i)$.
+Since there are many alleles that contribute more-or-less additively to the trait,
+the probability of fixation of a particular allele is not easy to find,
+but we can roughly approximate it as follows:
+if the current population trait mean $z$,
+then an allele that moves the trait a distance $x$ closer to the optimum
+has a fitness advantage in heterozygotes of 
+$\exp(-\beta((z - x)^2) / \exp( - \beta z^2)) \approx 1 + 2 \beta z x$, 
+where in our parameterization, $\beta = 1 / 450$.
+Since the probability of fixation of a single allele with relative advantage $s$
+is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical},
+then with $z=10$ we can use $p(x) \approx 4 x / 45$.
+Putting these considerations together,
+a rough estimate for the total shift in trait value resulting from selection
+on standing variation in the initial population is
+$R = \sum_i 8 x_i / 45$,
+where the sum is over all trait-affecting alleles.
+This approximation should be good if each trait-affecting allele is rare (present in only a few copies),
+effects are additive, and most are unlinked.
+
+We computed $R$ to confirm our suspicion that linkage slowed down adaptation
+by showing that there is sufficient allelic variation in the initial populations
+to complete adaptation across both migration rates (0.1 and 1 migrants per lake per generations),
+but not if the haplotypes of each genomic region are assumed to not recombine.
+
+We calculated for each individual in the initial generation of the new lakes
 the total contribution to trait values of each of the ten genomic regions 
@@
 On the contrary, many of these will likely be lost to drift.
+
 To get a rough idea of whether the genetic variation present in the initial population
@@
 to a probability of fixation
-(by $p(x) \approx 2x/45$),
+(as described below $p(x) \approx 2x/45$),
 and then calculated $\sum_i x_i p(x_i)$.
@@
 Suppose a new allele enters a lake, either by migration or mutation. 
-If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ --then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{fisher,prob_fixation}.
-If the current population all differed from the optimum trait by $z$, and the allele has effect size $-u$ in heterozygotes, then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
+If, when it is rare but present in $n$ copies, it has fitness advantage $s$ 
+-- i.e., the expected number of copies in the next generation is $(1+s)n$ 
+-- then the probability that it escapes demographic stochasticity to become common in the population 
+is approximately $2s$ \citep{fisher,prob_fixation}. 
+If the current population all differed from the optimum trait by $z$, 
+and the allele has effect size $-u$ in heterozygotes, 
+then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
 This tells us two things: (1) the rate of adaptation decreases as the population approaches the optimum, and (2) larger mutations (in the right direction) are more likely to fix.

### 2018-12-10 r_calcs/plot_linkage.R
.

@@ -34,6 +34,8 @@ everything <- lapply(0:3, function (run_num) {
     region_effects <- t(apply(effects, 1, tapply, variants$region, sum))
     nodes$phenotype <- rowSums(region_effects)
 
+    frequencies <- do.call(rbind, 
+                           tapply(1:nrow(genotypes), nodes$pop, function (k) colMeans(genotypes[k,])))
     # distribution of per-genome phenotypes
     if (FALSE) hist(nodes$phenotype)
 
@@ -46,7 +48,8 @@ everything <- lapply(0:3, function (run_num) {
                 effects=effects,
                 region_effects=region_effects,
                 pos_effects=pos_effects,
-                neg_effects=neg_effects))
+                neg_effects=neg_effects,
+                frequencies=frequencies))
   } )
 
 pdf(file='frequencies.pdf', width=8, height=5, pointsize=10)
@@ -57,7 +60,7 @@ lapply(everything, function (x) with(list2env(x), {
 
     layout(t(1:2))
 
-    # this shows how frequences alleles vary across lakes (one line per allele)
+    # this shows how allele frequences vary across lakes (one line per allele)
     matplot(lake_freqs, type='l',
             xlab='lake position',
             ylab='allele frequency')
@@ -123,30 +126,51 @@ expected_fix <- sapply(everything, function (x) with(list2env(x), {
               } )
           } ), sum)
      }))
-colnames(expected_fix) <- paste("m=", c(0.01, 0.1, 1, 10))
-
-## This is the table of the text
-xtable(data.frame(t(colMeans(expected_fix)), check.names=FALSE))
-
-# \begin{tabular}{rrrr}
-#   \hline
-#   m= 0.01 & m= 0.1 & m= 1 & m= 10 \\ 
-#   \hline
-#   0.12 & 4.70 & 12.93 & 12.27 \\ 
-#   \hline
-# \end{tabular}
+colnames(expected_fix) <- paste0("m=", c(0.01, 0.1, 1, 10))
 
 # sum of haplotype effect size, weighted by fitness
-# with fitness beta * z * u, where u is the effect size; z is the deviation from optimum,
+# with fitness 2 * beta * z * u, where u is the net effect size; z is the deviation from optimum,
 # and beta is 1/450
 haplotype_fitness <- sapply(everything, function (x) with(list2env(x), {
[…]

### 2018-12-10 Stickleback_Paper.tex

@@
 
-\begin{center}
-\begin{tabular}{ | c | c  | c  | c | }
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{rrrrr}
       \hline
- $m = 0.01$ & $m = 0.1$ & $m = 1.0$ & $m = 10.0$ \\ 
- $-0.1227$ &  $-4.6976$ & $-12.9298$  & $-12.2710$\\
+     & m=0.01 & m=0.1 & m=1 & m=10 \\ 
+      \hline
+          best & 2.77 & -12.02 & -21.02 & -7.02 \\ 
+      expected & 4.22 & -4.78 & -14.64 & -5.47 \\ 
        \hline
     \end{tabular}
-\end{center}
+    \caption{
+        Haplotypic variation present in the new lakes at time of colonization,
+        across rates of gene flow.
+        \textbf{``Best''} shows the most negative trait value achievable
+            with intact haplotypes, averaged across populations.
+        \textbf{``Expected''} shows the trait value
+            achieved by averaging haplotypic values weighted by fitness
+            (see text for details).
+        This suggests that populations at $m=1$ have sufficient genetic variation
+        to completely adapt (reach $-10$) using intact haplotypes,
+        while at $m=0.1$ variation must be released from haplotypes by recombination.
+    } \label{tab:linkage}
+\end{table}
 

### 2018-12-10 Stickleback_Paper.tex
Jared's Edits

@@
 determine speed of adaptation and reuse of standing genetic variation.
-We find that rapid, repeated adaptation using a shared set of alleles
-maintained at low frequency by migration-selection balance
-occurs over a realistic range of intermediate rates of gene flow.
-Low gene flow leads to slow, independent adaptation of distinct habitats,
-whereas high gene flow leads to large migration load.
-We do not see evidence for strong effects encouraging genomic clustering of causal alleles
-or towards particular dominance coefficients.
-$F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow.
-The results support existing theory of local adaptation,
-and provide a more concrete look at a particular, empirically motivated example.
-
-\textbf{**We need to be more crystal clear in the abstract about the findings from the paper**}
+(1) We find that rapid, repeated adaptation using a shared set of alleles maintained at low frequency by migration-selection balance occurs over a realistic range of intermediate rates of gene flow.
+(2) Low gene flow leads to slow, independent adaptation of distinct habitats, whereas high gene flow leads to large migration load.
+%We do not see evidence for strong effects encouraging genomic clustering of causal alleles
+%or towards particular dominance coefficients.
+(3) We quantify how $F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow.
+(4) In addition, we find the origin of many freshwater adapted alleles in the introduced lakes to be propagated from the original generation of marine individuals that had inhabited the lake.
+The results support existing theory of local adaptation, and provide a more concrete look at a particular, empirically motivated example.
 
@@
 
-Here we develop a forward simulation approach in SLiM to model the stickleback population structure in marine and freshwater habitats to document the effect of variation in gene flow and recombination on the genetic and genomic architecture of local adaptation.
-We address how rapidly can selection act on standing genetic variants at a given value of migration, $M$, and what the precise origin is of those variants.
-We ask what scale of historical introgression between populations with selective pressures $X$ and $Y$, allows for rapid and parallel local adaptation of a population derived from $X$ to an introduced environment with selective pressure $Y$?
-How are the variants structured, historically and across the genome underlying the trait?
-Furthermore, how can we infer causal loci for regions of the genome that must be driving the rapid adaptation, from real data.
-Many biologists today make use of genome wide association studies (GWAS) and $F_{st}$ across the genome to estimate regions responsible for certain traits.
-Unfortunately, the data can be heavily influenced by the level of introgression and population structure of the samples.
-This being the case, what are the biases we can expect to see in real data for certain levels of introgression.
+Here we develop a forward simulation approach in SLiM to model the stickleback evolutionary history in marine and freshwater habitats.
+We then record the effect of variation in gene flow on the genetic and genomic architecture of local adaptation.
+We ask, how rapidly can selection act on standing genetic variants at a given value of migration, $M$?
+%Furthermore, how can we infer causal (adaptive?) loci for regions of the genome that must be driving the rapid adaptation, from real data. 
+Furthermore, how might the inference of causal loci through the use of genome wide association studies (GWAS) and $F_{st}$ scans across the genome be skewed by (lack of) introgression between sub-populations.
+%The use of genome wide association studies (GWAS) and $F_{st}$ scans across the genome to estimate regions responsible for certain traits are popular methods that can be skewed by gene flow and population structure.
+%This being the case, what are the biases we can expect to see in real data for certain levels of introgression. 
+Finally, we trace back through the genealogical history of adaptive variants in new lakes to find the precise origin is of those variants.
+\textbf{Jared - My attempt at concisely stating are questions}
 
-We find that ....
+We find that only a few stickleback per generation, per lake, suffice to maintain freshwater alleles at an acceptable frequency in marine environments for efficient ``transportation" of the freshwater haplotype.
+The selection on SGV at this level results in rapid local adaptation in $\approx 60$ generations, this follows closely to what we observe in nature (such as middleton island, cite this).
+With lower $M < 1$ individuals per lake per generation we find either the lakes adapt indipendently by ``\textit{de novo}" mutation, or the selection on SGV is slow. (greater than 2000 gens.)
+This low level of introgression also results in false positive $F_{st}$ peaks when scanning across the genome suggesting introgression plays a key role in inferring causal loci . 
+With higher $M > 10$ individuals per lake per generation we see the impact of migration load, preventing the demes from fully adapting to their selective pressure.
+When tracing back to the origin of adaptive alleles being *used* in new lakes that had adapted, we find the large majority of the alleles had been ``captured" meaning they were present in the individuals that initially colonized the new lake,
+common in the original lakes, and uncommon in the ocean. 
+This finding is in opposition to the claim that freshwater alleles are transported through subsequent migration, after the lakes have been colonized. 
+\textbf{Jared - My attempt at concisely stating the results}
+
+%We ask what scale of historical introgression between populations with selective pressures $X$ and $Y$, allows for rapid and parallel local adaptation of a population derived from $X$ to an introduced environment with selective pressure $Y$?
+%We ask how the variants are structured, historically and across the genome underlying the trait? 
+%Furthermore, how can we infer causal (adaptive?) loci for regions of the genome that must be driving the rapid adaptation, from real data. 
+%Many biologists today make use of genome wide association studies (GWAS) and $F_{st}$ across the genome to estimate regions responsible for certain traits. 
+%Unfortunately, the data can be heavily influenced by the level of introgression and population structure of the samples. 
+%This being the case, what are the biases we can expect to see in real data for certain levels of introgression.
+
+%We find that ....
 \textbf{1.
@@
 The following I grabbed from the beginning of the Results section}
-*****
-We varied the migration rate, $m$, from $5 \times 10^{-5}$ to $5 \times 10^{-2}$ per individual, i.e., between 0.01 and 10 migrants per year to and from each of the ten lakes.
+%*****
+%We varied the migration rate, $m$, from $5 \times 10^{-5}$ to $5 \times 10^{-2}$ per individual, i.e., between 0.01 and 10 migrants per year to and from each of the ten lakes.
 This had a strong effect on many aspects of adaptation, including how fast adaptation occurred in each lake, how much alleles were shared between lakes,
-and the population genetic signals left behind.
+%and the population genetic signals left behind.
 At the lowest migration rate, lakes adapted nearly independently, while at the highest migration rate, the habitats are beginning to act like a single metapopulation with substantial migration load.
@@
 We also found that larger amounts of migration allow for more statistical power and less false positives in the resulting population genetics data ($F_{st}$ per SNP across the genome).
-****
+%****
 
@@
             a single, continuous, one-dimensional marine habitat (blue)
-            is coupled to randomly mating ``lakes'' at discrete locations.
+            is coupled to randomly mating ``lakes'' at discrete locations with dark green arrows representing migration patterns.
+            %It is important to note there is no migration from any one lake to another, therefor, freshwater alleles must travel through the marine habitat as is often the case in nature
             After an initial period of 100K generations with 25 lakes,
@@
 New offspring in the freshwater habitat are chosen in the same way, except the probability that the parents are marine individuals is $m$; 
-any new freshwater offpsring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.
+any new freshwater offspring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.
 
@@
         but brought in by migration at a later time.
+        %\textbf{The allele CAN exist in individuals at time of introduction, this category mean the linage of a specific allele in an individual happened to come from a migrant}
     \item \emph{Captured alleles:}
@@
 \end{enumerate}
+\textbf{I may be conf}
+
 Roughly speaking, these four categories describe the contribution of
@@
         (Adaptation did not occur at the highest rate of gene flow.)
-        \textbf{where are the multiple points?}
+        \textbf{where are the multiple points?, They do not exist, we only ran the one set. They each take $\approx$ 36 hours to run; but I can run them again if you want.} 
         } \label{fig:TimeToAdaptation}
@@
 \section*{Supplementary material}
-
+\begin{comment}
 \begin{figure}
@@
 \end{figure}
+\end{comment}
 

### 2018-12-11 Stickleback_Paper.tex
Jared's edits to the introduction

@@
 It is thought that these numerous, smaller freshwater populations
-have been established "de novo" from marine fish,
+have been established ``de novo" from marine fish,
 and that a shared freshwater phenotype
@@
 determine speed of adaptation and reuse of standing genetic variation.
-(1) We find that rapid, repeated adaptation using a shared set of alleles maintained at low frequency by migration-selection balance occurs over a realistic range of intermediate rates of gene flow.
-(2) Low gene flow leads to slow, independent adaptation of distinct habitats, whereas high gene flow leads to large migration load.
-%We do not see evidence for strong effects encouraging genomic clustering of causal alleles
-%or towards particular dominance coefficients.
-(3) We quantify how $F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow.
-(4) In addition, we find the origin of many freshwater adapted alleles in the introduced lakes to be propagated from the original generation of marine individuals that had inhabited the lake.
+We find that rapid, repeated adaptation using a shared set of alleles maintained at low frequency by migration-selection balance occurs over a realistic range of intermediate rates of gene flow.
+Low gene flow leads to slow, independent adaptation of distinct habitats, whereas high gene flow leads to large migration load.
+We quantify how $F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow.
+In addition, we find the origin of many freshwater adapted alleles in the introduced lakes to be propagated from the original generation of marine individuals that had inhabited the lake.
 The results support existing theory of local adaptation, and provide a more concrete look at a particular, empirically motivated example.
@@
 The canonical model for the genetics of adaptation, first formulated by Fisher in the early part of the 20th century, involves the sequential fixation of new mutations. 
-While it has proven valid in numerous studies in the field and the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures.
-A growing number of studies have identified instances of convergent or parallel adaptive evolution that utilize standing genetic variation.
-What evolutionary processes promote the maintenance of re-use of such genetic polymorphisms?
-How do these processes interact with diverse geographic contexts, as well as variation in genetic processes such as recombination, to change the likelihood of so-called standing genetic variation (SGV).
+While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures.
+A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split.
+%Which evolutionary processes promote the maintenance of re-use of such genetic polymorphisms? 
+%How do these processes interact with diverse geographic contexts, as well as variation in populations genetic processes such as gene flow, recombination, selection to fully take advantage of SGV to locally adapt.
+However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
+How do complex geographic systems with a variety of selective pressures interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations?
 
-Clues come from population genomic studies in the wild that have recently been powered by advances in sequencing technologies.
-One such organism is the threespine stickleback.
-This small fish has long been used for studies of behavior, ecology and evolution in the wild, and has more recently become a model for understanding the genetic basis of adaptive evolution.
-The ancestral marine form has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere, which have each evolved a similar suite - or 'freshwater syndrome' - of phenotypes despite local differences.
-This adaptive phenotypic evolution to freshwater can occur in mere millenia, in many cases since the end of the last ice age about 13,00 years ago.
-Studies of the fossil record also clearly show that invasion of oceanic stickleback and local adaptation in freshwater has occurred repeatedly over millions of years, often involving very similar phenotypic shifts as seen in contemporary populations.
+Recently, a large flood of data resulting from systems such as this come from population genomic studies in nature that have been powered by advances in sequencing technologies. 
+%One such organism is the Threespine Stickleback.
+One such organism is the Threespine Stickleback which has exhibited a long standing evolutionary history of migration-selection balance between marine and freshwater populations across the globe.
+%This small fish has long been used for studies of behavior, ecology and evolution in the wild, and has more recently become a model for understanding the genetic basis of adaptive evolution. 
+%The ancestral marine form has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
+The rigidity of the species' ability to frequently prosper in newly created freshwater environments has made this fish a good model for understanding the genetic basis of adaptive evolution. 
+The ancestral marine form of Stickleback has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
+%which have each evolved a similar suite - or 'freshwater syndrome' - of phenotypes despite local differences. 
+%Phnotypically, these freshwater patches which have 
+Geographic isolation prevents direct migration between a large majority of these patches of freshwater populations,
+yet we observe very similar phenotypes and many shared adaptive alleles often found to be identical by decent (IBD). 
+%Impressively, this adaptive phenotypic evolution to freshwater environments can occur in mere millenia, in many cases since the end of the last ice age about 13,000 years ago. 
+%Studies of the fossil record also clearly show that invasion of oceanic stickleback and local adaptation in freshwater has occurred repeatedly over millions of years, 
+%often involving very similar phenotypic shifts as seen in contemporary populations.
+Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in tens of generations, 
+and the adaptive alleles are found in freshwater populations that have been geographically isolated since the end of the last ice age $\approx 13,000$ years ago (Cite Kristin).
+%A finding from recent stickleback population genomic studies is that the same genomic regions, and in some cases the same haplotypes, are implicated in patterns of repeated evolution. (Cite?)
+%These studies show that divergent stickleback populations are clearly still connected via bi-directional gene flow. 
+%This reuse of standing genetic variation motivated empirical studies to ask whether this phenotypic and genomic evolution could occur in decades. 
+%The answer is a clear yes. 
+For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island. 
+Quickly inhabited by the surrounding marine population of stickleback, 
+\citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be parallel to freshwater stickleback that have been separated for over thousands of years. 
+In these freshwater stickleback, the number of lateral plates are reduced and the opercle shapes shows
+the same expansion of the dorsal region and reduction of the ventral region as observed in the large majority of freshwater demes.
 
-A finding from recent stickleback population genomic studies is that the same genomic regions, and in some cases the same haplotypes, are implicated in patterns of repeated evolution.
-These studies show that divergent stickleback populations are clearly still connected via bi-directional gene flow.
-This reuse of standing genetic variation motivated empirical studies to ask whether this phenotypic and genomic evolution could occur in decades.
-The answer is a clear yes.
-For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island.
-Quickly inhabited by the surrounding marine population of stickleback, \citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be parallel to freshwater stickleback that have been separated for over thousands of years.
-In freshwater stickleback, the number of lateral plates are reduced and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region.
-
- But how can evolution occur so quickly?
-The leading hypothesis for stickleback is that rather than acting on new mutations, adaptation to freshwater environments is accelerated by selection on SGV found in marine populations.
+But how can evolution occur at such a rapid pace? 
+Waiting for new mutations to arise in each lake would take much longer, and genotypes being identical by state is even more improbable. 
+This would seem to suggest that freshwater alleles are maintained in the marine individuals by repeated adaptation to freshwater environments and downstream gene flow from those locally adapted sub-populations.
+%This suggests that adaptation to freshwater environments is accelerated by selection on SGV found in marine populations. 
 The first clear example of the global reuse of SGV was the gene \textit{eda} which has been shown to be an important regulator for the number of lateral plates. 
 While the low lateral plate version of this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently. 
-More recently, population genomic studies employing genome-wide haplotype analyses has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
-Genomic regions that exhibit local adaptation, as evidenced by increased relative population genomic divergence, also show much older haplotypes that are enriched in the alternative habitats.
+More recently, population genomic studies employing genome-wide haplotype analyses has provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
+%Genomic regions that exhibit local adaptation, as evidenced by increased relative population genomic divergence, also show much older haplotypes that are enriched in the alternative habitats.
 
-These empirical data generally support the ``transporter''-hypothesis proposed by Conte and Schluter in 2009 \citet{schluter2009genetics}, which is a conceptual model for the flow of freshwater alleles from multiple smaller freshwater populations into much larger and less structured marine populations through hybridization events.
-Alleles in this asymetrically structured metapopulation can then be recycled for subsequent freshwater adaptation.
-Several questions remain, however, about the manner and degree to which these alleles are scattered among the marine fish and must be re-assembled.
+These empirical data generally support the ``transporter''-hypothesis proposed by Conte and Schluter in 2009 \citet{schluter2009genetics}, 
+which is a conceptual model for the flow of freshwater alleles from multiple smaller freshwater populations into much larger and less structured marine populations through hybridization events. 
+Alleles in this asymmetrically structured meta-population can then be recycled for subsequent freshwater adaptation. 
+Several questions remain, however, about the manner and degree to which these alleles are scattered among the marine fish allowing the haplotype to be re-assembled. 
 Is it more akin to the atomization and rebuilding that occurred in the Star Trek series that motivated this hypothesis? 
-Alternatively does it occur in more of a patchwork by individuals, or their early generation hybrids, in geographically adjacent freshwater habitats moving quickly through the marine environment to seed new freshwater populations.
-More generally, how do different levels of gene flow and local adaptation interact with the asymmetric nature of stickleback population structure affect the dynamics of the origin and rate of adaptive alleles, as well as their eventual organization into marine and freshwater typical genomes?
+Alternatively does it occur in more of a patchwork by individuals, 
+or their early generation hybrids in geographically adjacent freshwater habitats moving quickly through the marine environment to seed new freshwater populations. 
+More generally, how do different levels of gene flow and local adaptation interact with the asymmetric nature of stickleback population structure affect the dynamics of the origin and rate of adaptive alleles,
+as well as their eventual organization into marine and freshwater typical genomes?
 
@@
 Finally, we trace back through the genealogical history of adaptive variants in new lakes to find the precise origin is of those variants.
-\textbf{Jared - My attempt at concisely stating are questions}
+%\textbf{Jared - My attempt at concisely stating are questions}
 
@@
 The selection on SGV at this level results in rapid local adaptation in $\approx 60$ generations, this follows closely to what we observe in nature (such as middleton island, cite this).
-With lower $M < 1$ individuals per lake per generation we find either the lakes adapt indipendently by ``\textit{de novo}" mutation, or the selection on SGV is slow. (greater than 2000 gens.)
-This low level of introgression also results in false positive $F_{st}$ peaks when scanning across the genome suggesting introgression plays a key role in inferring causal loci . 
+Surprisingly, we find that the the continued gene flow (subsequent migration) of marine individuals is less important than the initial amount of SGV in the initial generation marine individuals at time of colonization.
+This suggests downstream gene flow from freshwater populations is key to this system of ``transportation".
+%With lower $M < 1$ individuals per lake per generation we find either the lakes adapt indipendently by ``\textit{de novo}" mutation, or the selection on SGV is slow. (greater than 2000 gens.)
+We show that low levels of introgression also results in noise that often presents itself as $F_{st}$ peaks when scanning across the genome suggesting introgression plays a key role in inferring causal loci .
 With higher $M > 10$ individuals per lake per generation we see the impact of migration load, preventing the demes from fully adapting to their selective pressure.
-When tracing back to the origin of adaptive alleles being *used* in new lakes that had adapted, we find the large majority of the alleles had been ``captured" meaning they were present in the individuals that initially colonized the new lake,
-common in the original lakes, and uncommon in the ocean. 
-This finding is in opposition to the claim that freshwater alleles are transported through subsequent migration, after the lakes have been colonized. 
-\textbf{Jared - My attempt at concisely stating the results}
+%When tracing back to the origin of adaptive alleles being *used* in new lakes that had adapted, we find the large majority of the alleles had been ``captured" meaning they were present in the individuals that initially colonized the new lake,
+%common in the original lakes, and uncommon in the ocean. 
+%This finding is in opposition to the claim that freshwater alleles are transported through subsequent migration, after the lakes have been colonized. 
+%\textbf{Jared - My attempt at concisely stating the results}
 
@@
 Using these trees,
-we classified each freshwater allele found in the new lakes at time of adaptation
+%we classified each freshwater allele found in the new lakes at time of adaptation
+we classified each adaptive allele, in each genome,
 into four categories:
@@
     \item \emph{Migrant alleles:} 
-        not present in the individuals that initially colonized the new lake, 
-        but brought in by migration at a later time.
+        %not present in the individuals that initially colonized the new lake, 
+        %but brought in by migration at a later time.
         %\textbf{The allele CAN exist in individuals at time of introduction, this category mean the linage of a specific allele in an individual happened to come from a migrant}
+        deriving from a migrant that was not in the initial generation that colonized the lake
     \item \emph{Captured alleles:}
@@
 \end{enumerate}
-\textbf{I may be conf}
+%\textbf{I may be conf}
 
@@
 
+\begin{comment}
 \paragraph{Standing genetic variation} 
@@
 \plr{perhaps we want to remove this bit? revisit if it is telling us something important.}
+\end{comment}
 
@@
 
+%sum up the system
+
+%dynamic. and static.
+
+% not subsequent
+
+% reasonable amounts of gene flow result, 
+% SGV is powerful, and onlt needs a small amount of gene flow.
+
+% downstream migration not as important. 
+% doesn't take a strange environment.
+
+% look up and cite migration load.
+
 We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
@@
 
-We also have also explored the genomic architecture as a consequence of the nature of selection in our scenario.
-The alleles underlying individual trait value were of large effect and a low number.
-With a total of $10^{5}$ loci, to be realistic, each loci should represent 1000 $Kb$ in real data.
+%Systems in which there is one, static selective pressure with a wide range connected to many, much smaller systems
 

### 2018-12-11 Stickleback_Paper.tex
Jared's edits to the introduction 2

@@
 A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split.
-%Which evolutionary processes promote the maintenance of re-use of such genetic polymorphisms? 
-%How do these processes interact with diverse geographic contexts, as well as variation in populations genetic processes such as gene flow, recombination, selection to fully take advantage of SGV to locally adapt.
 However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
@@
 Recently, a large flood of data resulting from systems such as this come from population genomic studies in nature that have been powered by advances in sequencing technologies. 
-%One such organism is the Threespine Stickleback.
 One such organism is the Threespine Stickleback which has exhibited a long standing evolutionary history of migration-selection balance between marine and freshwater populations across the globe.
-%This small fish has long been used for studies of behavior, ecology and evolution in the wild, and has more recently become a model for understanding the genetic basis of adaptive evolution. 
-%The ancestral marine form has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
 The rigidity of the species' ability to frequently prosper in newly created freshwater environments has made this fish a good model for understanding the genetic basis of adaptive evolution. 
 The ancestral marine form of Stickleback has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
-%which have each evolved a similar suite - or 'freshwater syndrome' - of phenotypes despite local differences. 
-%Phnotypically, these freshwater patches which have 
 Geographic isolation prevents direct migration between a large majority of these patches of freshwater populations,
 yet we observe very similar phenotypes and many shared adaptive alleles often found to be identical by decent (IBD). 
-%Impressively, this adaptive phenotypic evolution to freshwater environments can occur in mere millenia, in many cases since the end of the last ice age about 13,000 years ago. 
-%Studies of the fossil record also clearly show that invasion of oceanic stickleback and local adaptation in freshwater has occurred repeatedly over millions of years, 
-%often involving very similar phenotypic shifts as seen in contemporary populations.
 Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in tens of generations, 
 and the adaptive alleles are found in freshwater populations that have been geographically isolated since the end of the last ice age $\approx 13,000$ years ago (Cite Kristin).
-%A finding from recent stickleback population genomic studies is that the same genomic regions, and in some cases the same haplotypes, are implicated in patterns of repeated evolution. (Cite?)
-%These studies show that divergent stickleback populations are clearly still connected via bi-directional gene flow. 
-%This reuse of standing genetic variation motivated empirical studies to ask whether this phenotypic and genomic evolution could occur in decades. 
-%The answer is a clear yes. 
 For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island. 
@@
 Waiting for new mutations to arise in each lake would take much longer, and genotypes being identical by state is even more improbable. 
-This would seem to suggest that freshwater alleles are maintained in the marine individuals by repeated adaptation to freshwater environments and downstream gene flow from those locally adapted sub-populations.
+This would seem to suggest that freshwater alleles are maintained in the marine individuals allowing the accelerated selection on SGV found in marine individuals which colonize the lake. 
+%by repeated adaptation to freshwater environments and downstream gene flow from those locally adapted sub-populations.
 %This suggests that adaptation to freshwater environments is accelerated by selection on SGV found in marine populations. 

### 2018-12-11 Stickleback_Paper.tex
peter edits through theory

@@
 the vast majority of alleles underlying the local adaptation
-derived from individuals in the original generation (``captured'' or ``pre-existing'').
+derived from individuals in the original generation (``captured'' or ``marine'').
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{rrrrr}
+      \hline
+     & m=0.01 & m=0.1 & m=1 & m=10 \\ 
+      \hline
+          best & 2.77 & -12.02 & -21.02 & -7.02 \\ 
+      % expected & 4.22 & -4.78 & -14.64 & -5.47 \\ 
+       \hline
+    \end{tabular}
+    \caption{
+        Haplotypic variation present in the new lakes at time of colonization,
+        across rates of gene flow,
+        % \textbf{``Best''} 
+        quantified as the most negative trait value achievable
+            with intact haplotypes, averaged across populations
+            (see text for details).
+        % \textbf{``Expected''} shows the trait value
+        %     achieved by averaging haplotypic values weighted by fitness
+        %     (see text for details).
+    } \label{tab:linkage}
+\end{table}
 
@@
 but each is paired with a compensatory allele with effect size $+1$.
-Each local haplotype is therefore neutral, 
-and so likely to be at higher frequency in the marine population.
-
+Each local haplotype is therefore neutral
+(and so likely to be at higher frequency in the marine population).
+
+Indeed, this appears to be the case.
+To quantify the genetic variation available \emph{without} recombining within the ten genomic regions,
+we first found, within each population, the haplotype with the largest net negative effect
+at each of the ten genomic regions.
+Summing these ten numbers, 
+we get the maximum amount that selection could move the population in the freshwater direction
+without recombining within regions.
+The mean of this value across the 25 populations is shown 
+in the first row (``best'') of Table~\ref{tab:linkage} --
+typical populations at $m=0.1$ migrants per lake per generation could shift to a phenotype of -12.02
+(and so have sufficient variation to adapt without recombination),
+but the ``best'' haplotypes populations at $m=1.0$ have effect sizes nearly twice as big 
+(a mean total of -21.02).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Theoretical expectations}
+
+We now revisit our observations above
+in the light of population genetics theory.
+The discussion will be rough,
+with a focus on intuition rather than precise calculations.
+Since within each population we model stabilizing selection on an additive trait
+controlled by a moderately large number of loci,
+more precise expectations might be obtained through quantitative genetics \citep{svardal2014general}
+or even Fisher's geometric model \citep{barton2001hybridization,chevin2014niche}.
+
+Suppose a new allele enters a lake, either by migration or mutation. 
+If, when it is rare but present in $n$ copies, it has fitness advantage $s$ 
+-- i.e., the expected number of copies in the next generation is $(1+s)n$ 
+-- then the probability that it escapes demographic stochasticity to become common in the population 
+is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical}
+(assuming Poisson reproduction, as we roughly have here).
+If the current population all differed from the optimum trait by $z$, 
+and the allele has effect size $-u$ in heterozygotes, 
+then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$. 
+This tells us two things: (1) the rate of adaptation decreases as the population approaches the optimum, and (2) larger mutations (in the right direction) are more likely to fix.
+
+\paragraph{New mutations}
+The total rate of appearance of new mutations per lake is $\mu_L = 0.04$, 
+which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive in either direction. 
+This implies that a new additive or dominant effect mutation appears once every 87.5 generations, on average. 
+Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, 
+and so the probability that a dominant mutation manages to establish in a population differing from the optimum by $z$ 
+is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$, 
+and so the rate of establishment of dominant mutations is $\beta z / 87.5$, 
+i.e., about one such mutation every $2461/z$ generations. 
+The distribution of these successfully established mutations 
+has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2. 
+Since additive alleles have half the effect in heterozygotes, 
+they have half the probability of establishment. 
+During the initial phase of adaptation, the populations begin at around distance $z=10$ from the optimum. 
+Combining these facts, we expect adaptive alleles to appear through mutation within lakes at first on a time scale of 250 generations, 
+with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
+This agrees roughly with what we see in Figures~\ref{fig:phenotype_ts2} and XXX.
+\plr{refer to additional trace plots in the supp}
+
+\paragraph{Standing variation}
+An allele that moves the trait $z$ units in the freshwater direction in heterozygotes 
+has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment. 
+The product of population size and fitness differential in the marine environment for a mutation with $z=1$ is therefore $2Ns = 8.9$, 
+implying that these alleles are strongly selected against but might occasionally drift to moderate frequency. 
+The average frequency of such an allele in the marine environment at migration-selection equilibrium 
+is equal to the total influx of alleles into the ocean per generation divided by the selective disadvantage, 
+which if there are $M$ immigrants per generation, is $2 M / \beta z^2$. 
+\plr{This $M$ is the total number in the ocean! Make sure this differs from notation above for the number per lake.}
+Each new lake is likely to contain a few copies of alleles at frequency above 1/400
+(since each new lake is begun with 400 genomes).
+With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$,
+so if $M \ge 1$, 
+the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes 
+will appear at least once in the fish that colonize a new lake. 
+However, an allele with $z=1$ only has probability of around 1/20 of establishing locally, 
+suggesting that the migration rate should be somewhat higher to ensure enough pre-existing genetic variation 
+that adaptation would happen entirely using the initial set of colonizers. 
+% This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.
+However, this calculation treats each allele independently;
+in practice we found that standing freshwater variation in the ocean
+were masked by linkage to compensatory marine variants.
+
+\paragraph{Migration}
+The key quantity regulating the amount of standing variation in the ocean was the \emph{downstream} migration rate,
+from lakes to the ocean.
+How important is the upstream migration rate?
+If sufficient genetic variation is not present in a new lake initially, 
+it must appear either by new mutation or by migration. 
+Since a proportion $m$ of each lake is composed of migrants each generation, 
+it takes $1/m$ generations until the genetic variation introduced by migrants equals the amount initially present at colonization. 
+This implies a dichotomy: either 
+(a) migration is high, and adaptation is possible using variants present at colonization or arriving shortly thereafter, or 
+(b) migration is low, so adaptation takes many multiples of $1/m$ generations.
+Since in our model lower migration also reduces the amount of variation available in the ocean,
+we expect very little contribution of subsequent migration across any value of $m$,
+as seen in Figure~\ref{fig:Origin}.
+
+% These calculations depend on there being no bottleneck in colonization of the lake; if there is a bottleneck, then an additional factor must be added.
+At what point do we expect new mutation to be more important than migration for adaptation?
+By the calculations above, if $M \ge 10$, we expect initial diversity in a lake to be sufficient  for adaptation, corresponding to our third-highest migration rate.
+If this does not happen, then we expect adaptation to take a multiple of $1/m$ generations;
+% with our values, $1/m$ ranges from 20,000 to 20 generations.
+Above we estimated that adaptive alleles due to new mutation fix locally about every 2,000 generations, suggesting that at our second-lowest migration rate (where $1/m = 2,000$), the two contributions of migration and new mutation are roughly equal.
+% This is in fact what we see: in Figure \ref{fig:MPFAI}, we see that at the second migration rate, alleles start to be shared between lakes, while by the third migration rate, they are almost entirely shared.
+
+%%%%%%%%%%%%%%%%%%%%
+\section*{Discussion}
+
+%sum up the system
+
+%dynamic. and static.
+
+% not subsequent
+
+% reasonable amounts of gene flow result, 
+% SGV is powerful, and onlt needs a small amount of gene flow.
+
+% downstream migration not as important. 
+% doesn't take a strange environment.
+
+% look up and cite migration load.
+
+We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
+Selection is able to rebuild the freshwater haplotype from marine populations as a medium between all freshwater populations. 
+Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.
+
+%Systems in which there is one, static selective pressure with a wide range connected to many, much smaller systems
+
+We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
+This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
+differences between populations that have little gene flow between them.
+It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
+mistaken for being causative as they often display the same amount of divergence.
+
+%wouldn't it be cool if we could use this fact and simulate introgression, starting with real data that is noisy to extrapolate the areas we care about?
+%this could be nonsense
+
+% at low M we should be able to predict something ---
+% at high M we should be able to predict something ----
+
+%Here, we suggest a range of migration rate (introgression) parameter values which would allow for the rapid (and in turn, parallel) adaptation 
+%of marine stickleback introduced intro a freshwater environment.
+%While this range is heavily effected by all other selections of parameter values.
+
+%(1) There's a threshold of introgression for both rapid adaptation and migration load. 
+%(1.5) Maybe refer to the literature for migration load.  
+\subsection*{thresholds}
+
+We have found that too little migration leads to selection upon new mutations in all subpopulations and lakes alike.
+In contrast, at high migration rates we have seen that migration load limits the ability for species to locally adapt to the selective pressure of their environment.
+This leads us to consider a window of introgression which allows for the transportation of FAA's without migration load.
+
+\subsection*{connect results back to real data?}
+
+%In a species that commonly is subjected to two general types of selective pressure 
+%such as marine and freshwater stickleback this is suggestive of the benefit of hybridization in all subpopulations before
+%reaching a point of pre or post zygotic
+
+%(3) discuss what signals researchers could possibly look out for or connect to real data
+
+\paragraph{The adaptive filter?}
+RAMBLING THOUGHTS HERE
+Since larger effect alleles are more likely to establish,
+be it by mutation or migration,
+repeated colonization of new freshwater habitats will select for larger alleles,
+be it single alleles or haplotypes bound together by an inversion.
+However, these are more strongly selected against in the interstitial time.
+Being recessive would help with this,
+but would also make it more difficult to establish.
+
+\jgg{TODO: effect of smaller than realistic population sizes?}
+
+\jgg{TODO: Talk about the role recombination plays}
+This also brings to light the role of recombination in a system of migration selection balance. 
+On one hand, a lower recombination rate would allow for the freshwater haplotype to remain "intact" 
+within the marine environment, however, without significant introgression the haplotype would quickly get selected against in marine populations. 
+On the other hand a higher recombination rate would, in theory, allow freshwater adapted alleles more longevity 
+as they would hitchhike with marine adapted haplotypes.
+
+\paragraph{Modeling assumptions}
+Our simulations had much smaller population sizes than real populations.
+How might this affect things?
+
+\bibliographystyle{plainnat}
+\bibliography{Citations}{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
+\appendix
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+\section*{Supplementary material}
+\begin{comment}
+\begin{figure}
+	\begin{center}
+  		\includegraphics{Final_Plots/True_Power.pdf}
+  		\caption{ 
+		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
+		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
+		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
+            \textbf{
+                TODO: fix this caption; x-axis label ($F_{ST}$).}
+		}
+  		\label{fig:Power_FP}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=1.0\linewidth]{Final_Plots/Haplo_small.pdf}
+  		\caption{
+			Phenotypes (\textbf{Top}) (effect size and mutation type) and haplotypes \textbf{(Botton)} as a function of spatial position of the individual 
+			possessing the genome. The phenotype of each haplotype is, by definition, the sum of the 
+			effect sizes of each mutation in the genome. 
+		}
+  		\label{fig:Haplo_Pheno}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-2.png}
+        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
+  		\caption{
+            \textbf{Mean phenotypes through time},
+            at migration rates of
+            \textbf{(top)} 10 and
+            \textbf{(bottom)} 1 migrants per lake per generation.
+		}
+  		\label{fig:pheno_by_time1}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
+        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-5.png}
+  		\caption{
+            \textbf{Mean phenotypes through time},
+            at migration rates of
+            \textbf{(top)} 0.1 and
+            \textbf{(bottom)} 0.01 migrants per lake per generation.
+		}
+  		\label{fig:pheno_by_time2}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+        \includegraphics[width=1.0\linewidth]{plotlyPlots/StandingGeneticVariation.png}
+  		\caption{
+            \textbf{Standing genetic variance},
+            computed as XXX write if we leave this in XXX.
+		}
+        \label{fig:SGV}
+	\end{center}
+\end{figure}
+\end{comment}
+
+\end{document}
+
+The calculation above ignores allele frequencies -- 
+even if a population \emph{could} adapt entirely using the best possible haplotypes,
+many of these may be lost to drift.
+To account for this, we calculated a similar quantity but accounting for the probability of fixation.
 Suppose that the alleles present in the initial population of a new lake
@@
 
-To get a rough idea of whether the genetic variation present in the initial population
-is sufficient for the population to adapt,
-we translated the total effect size $x$ of each region 
-to a probability of fixation
-(as described below $p(x) \approx 2x/45$),
-and then calculated $\sum_i x_i p(x_i)$.
-If a region moves the trait an average of $x$ units towards the freshwater optimum
-(averaging over dominance relationships),
-Given the probability that an ``effect region'' fixes at a given trait value, 
-$2 \times x / 45$ where $x$ is the effect size of a haplotype,
-we were able to determine expected total effect sizes of the haplotypes that fix (detail about computation in methods).
-Below, we see the mean of these expected total effect sizes across all lakes.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{rrrrr}
-      \hline
-     & m=0.01 & m=0.1 & m=1 & m=10 \\ 
-      \hline
-          best & 2.77 & -12.02 & -21.02 & -7.02 \\ 
-      expected & 4.22 & -4.78 & -14.64 & -5.47 \\ 
-       \hline
-    \end{tabular}
-    \caption{
-        Haplotypic variation present in the new lakes at time of colonization,
-        across rates of gene flow.
-        \textbf{``Best''} shows the most negative trait value achievable
-            with intact haplotypes, averaged across populations.
-        \textbf{``Expected''} shows the trait value
-            achieved by averaging haplotypic values weighted by fitness
-            (see text for details).
-        This suggests that populations at $m=1$ have sufficient genetic variation
-        to completely adapt (reach $-10$) using intact haplotypes,
-        while at $m=0.1$ variation must be released from haplotypes by recombination.
-    } \label{tab:linkage}
-\end{table}
-
 Between $m = 0.1$ and $m = 1.0$ we see a large difference in the total effect sizes of the haplotypes that are expected to fix.
@@
 to select and rebuild a pre-existing haplotype given the standing variation at time of introduction.
-
-\subsection*{Theoretical expectations}
-\plr{does this go first or second?}
-
-Here's some rough calculations to get a sense for what should be going on. Everything is done in more detail in OTHER PLACES WE SHOULD CITE.
-Fisher, Chevin, etc.
-
-Suppose a new allele enters a lake, either by migration or mutation. 
-If, when it is rare but present in $n$ copies, it has fitness advantage $s$ 
--- i.e., the expected number of copies in the next generation is $(1+s)n$ 
--- then the probability that it escapes demographic stochasticity to become common in the population 
-is approximately $2s$ \citep{fisher,prob_fixation}. 
-If the current population all differed from the optimum trait by $z$, 
-and the allele has effect size $-u$ in heterozygotes, 
-then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
-This tells us two things: (1) the rate of adaptation decreases as the population approaches the optimum, and (2) larger mutations (in the right direction) are more likely to fix.
-
-\paragraph{New mutations}
-The total rate of appearance of new mutations per lake is $\mu_L = 0.04$, which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive in either direction.
-This implies that a new additive or dominant effect mutation appears once every 87.5 generations, on average.
-Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, and so the probability that a dominant mutation manages to establish in a population differing from the optimum by $z$ is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$, and so the rate of establishment of dominant mutations is $\beta z / 87.5$, i.e., about one such mutation every $2461/z$ generations.
-The distribution of these successfully established mutations  has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2.
-Since additive alleles have half the effect in heterozygotes, they have half the probability of establishment.
-During the initial phase of adaptation, the populations begin at around distance $z=10$ from the optimum.
-Combining these facts, we expect adaptive alleles to appear through mutation at first on a time scale of 250 generations, with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
-
-\paragraph{Standing variation}
-An allele that moves the trait $z$ units in the freshwater direction in heterozygotes has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment (which is close to optimal).
-The product of population size and fitness differential in the marine environment for a mutation with $z=1$ is therefore $2Ns = 8.9$, implying that these alleles are strongly selected against but might occasionally drift to moderate frequency.
-The average frequency of such an allele in the marine environment at migration-selection equilibrium is equal to the total influx of alleles per generation divided by the selective disadvantage, which if $M = 2000 m$ is the number of immigrants per generation, is $2 M / \beta z^2$.
-With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$: since lakes have 400 genomes, as long as $M \ge 1$, the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes will appear at least once in the fish that colonize a new lake.
-However, an allele with $z=1$ only has probability of around 1/20 of establishing locally, suggesting that we'd need $M \ge 10$ to ensure enough pre-existing genetic variation that adaptation would happen entirely using the initial set of colonizers.
-This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.
-
-\paragraph{Migration}
-If sufficient genetic variation is not present in a new lake initially, it must appear by new mutation or migration.
-Since a proportion $m$ of each lake is composed of migrants each generation, it takes $1/m$ generations until the genetic variation introduced by migrants equals the amount initially present at colonization.
-This implies a dichotomy: either (a) adaptation is possible using variants present at colonization or arriving shortly thereafter, or (b) adaptation takes many multiples of $1/m$ generations.
-
-These calculations depend on there being no bottleneck in colonization of the lake; if there is a bottleneck, then an additional factor must be added.
-At what point do we expect new mutation to be more important than migration for adaptation?
-By the calculations above, if $M \ge 10$, we expect initial diversity in a lake to be sufficient  for adaptation, corresponding to our third-highest migration rate.
-If this does not happen, then we expect adaptation to take a multiple of $1/m$ generations;
-with our values, $1/m$ ranges from 20,000 to 20 generations.
-Above we estimated that adaptive alleles due to new mutation fix locally about every 2,000 generations, suggesting that at our second-lowest migration rate (where $1/m = 2,000$), the two contributions of migration and new mutation are roughly equal.
-This is in fact what we see: in Figure \ref{fig:MPFAI}, we see that at the second migration rate, alleles start to be shared between lakes, while by the third migration rate, they are almost entirely shared.
-
-%%%%%%%%%%%%%%%%%%%%
-\section*{Discussion}
-
-%sum up the system
-
-%dynamic. and static.
-
-% not subsequent
-
-% reasonable amounts of gene flow result, 
-% SGV is powerful, and onlt needs a small amount of gene flow.
-
-% downstream migration not as important. 
-% doesn't take a strange environment.
-
-% look up and cite migration load.
-
-We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
-Selection is able to rebuild the freshwater haplotype from marine populations as a medium between all freshwater populations. 
-Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.
-
-%Systems in which there is one, static selective pressure with a wide range connected to many, much smaller systems
-
-We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
-This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
-differences between populations that have little gene flow between them.
-It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
-mistaken for being causative as they often display the same amount of divergence.
-
-%wouldn't it be cool if we could use this fact and simulate introgression, starting with real data that is noisy to extrapolate the areas we care about?
-%this could be nonsense
-
-% at low M we should be able to predict something ---
-% at high M we should be able to predict something ----
-
-%Here, we suggest a range of migration rate (introgression) parameter values which would allow for the rapid (and in turn, parallel) adaptation 
-%of marine stickleback introduced intro a freshwater environment.
-%While this range is heavily effected by all other selections of parameter values.
-
-%(1) There's a threshold of introgression for both rapid adaptation and migration load. 
-%(1.5) Maybe refer to the literature for migration load.  
-\subsection*{thresholds}
-
-We have found that too little migration leads to selection upon new mutations in all subpopulations and lakes alike.
-In contrast, at high migration rates we have seen that migration load limits the ability for species to locally adapt to the selective pressure of their environment.
-This leads us to consider a window of introgression which allows for the transportation of FAA's without migration load.
-
-\subsection*{connect results back to real data?}
-
-%In a species that commonly is subjected to two general types of selective pressure 
-%such as marine and freshwater stickleback this is suggestive of the benefit of hybridization in all subpopulations before
-%reaching a point of pre or post zygotic
-
-%(3) discuss what signals researchers could possibly look out for or connect to real data
-
-\paragraph{The adaptive filter?}
-RAMBLING THOUGHTS HERE
-Since larger effect alleles are more likely to establish,
-be it by mutation or migration,
-repeated colonization of new freshwater habitats will select for larger alleles,
-be it single alleles or haplotypes bound together by an inversion.
-However, these are more strongly selected against in the interstitial time.
-Being recessive would help with this,
-but would also make it more difficult to establish.
-
-\jgg{TODO: effect of smaller than realistic population sizes?}
-
-\jgg{TODO: Talk about the role recombination plays}
-\jgg{Below may belong in the discussion / future work}
-This also brings to light the role of recombination in a system of migration selection balance. 
-On one hand, a lower recombination rate would allow for the freshwater haplotype to remain "intact" 
-within the marine environment, however, without significant introgression the haplotype would quickly get selected against in marine populations. 
-On the other hand a higher recombination rate would, in theory, allow freshwater adapted alleles more longevity 
-as they would hitchhike with marine adapted haplotypes.
-
-\paragraph{Modeling assumptions}
-Our simulations had much smaller population sizes than real populations.
-How might this affect things?
-
-\bibliographystyle{plainnat}
-\bibliography{Citations}{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-\appendix
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}
-
-\section*{Supplementary material}
-\begin{comment}
-\begin{figure}
-	\begin{center}
-  		\includegraphics{Final_Plots/True_Power.pdf}
-  		\caption{ 
-		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
-		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
-		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
-            \textbf{
-                TODO: fix this caption; x-axis label ($F_{ST}$).}
-		}
-  		\label{fig:Power_FP}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-  		\includegraphics[width=1.0\linewidth]{Final_Plots/Haplo_small.pdf}
-  		\caption{
-			Phenotypes (\textbf{Top}) (effect size and mutation type) and haplotypes \textbf{(Botton)} as a function of spatial position of the individual 
-			possessing the genome. The phenotype of each haplotype is, by definition, the sum of the 
-			effect sizes of each mutation in the genome. 
-		}
-  		\label{fig:Haplo_Pheno}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-2.png}
-        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
-  		\caption{
-            \textbf{Mean phenotypes through time},
-            at migration rates of
-            \textbf{(top)} 10 and
-            \textbf{(bottom)} 1 migrants per lake per generation.
-		}
-  		\label{fig:pheno_by_time1}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
-        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-5.png}
-  		\caption{
-            \textbf{Mean phenotypes through time},
-            at migration rates of
-            \textbf{(top)} 0.1 and
-            \textbf{(bottom)} 0.01 migrants per lake per generation.
-		}
-  		\label{fig:pheno_by_time2}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-        \includegraphics[width=1.0\linewidth]{plotlyPlots/StandingGeneticVariation.png}
-  		\caption{
-            \textbf{Standing genetic variance},
-            computed as XXX write if we leave this in XXX.
-		}
-        \label{fig:SGV}
-	\end{center}
-\end{figure}
-\end{comment}
-
-\end{document}

### 2018-12-11 Stickleback_Paper.tex
Jared's edits on Methods

--- 2018-12-11 Stickleback_Paper.tex
@@
 Recently, a large flood of data resulting from systems such as this come from population genomic studies in nature that have been powered by advances in sequencing technologies. 
-One such organism is the Threespine Stickleback which has exhibited a long standing evolutionary history of migration-selection balance between marine and freshwater populations across the globe.
+One such organism is the threespine stickleback which has exhibited a long standing evolutionary history of migration-selection balance between marine and freshwater populations across the globe.
 The rigidity of the species' ability to frequently prosper in newly created freshwater environments has made this fish a good model for understanding the genetic basis of adaptive evolution. 
@@
 This would seem to suggest that freshwater alleles are maintained in the marine individuals allowing the accelerated selection on SGV found in marine individuals which colonize the lake. 
-%by repeated adaptation to freshwater environments and downstream gene flow from those locally adapted sub-populations.
-%This suggests that adaptation to freshwater environments is accelerated by selection on SGV found in marine populations. 
 The first clear example of the global reuse of SGV was the gene \textit{eda} which has been shown to be an important regulator for the number of lateral plates. 
@@
 More recently, population genomic studies employing genome-wide haplotype analyses has provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
-%Genomic regions that exhibit local adaptation, as evidenced by increased relative population genomic divergence, also show much older haplotypes that are enriched in the alternative habitats.
 
@@
 We ask, how rapidly can selection act on standing genetic variants at a given value of migration, $M$?
-%Furthermore, how can we infer causal (adaptive?) loci for regions of the genome that must be driving the rapid adaptation, from real data. 
 Furthermore, how might the inference of causal loci through the use of genome wide association studies (GWAS) and $F_{st}$ scans across the genome be skewed by (lack of) introgression between sub-populations.
-%The use of genome wide association studies (GWAS) and $F_{st}$ scans across the genome to estimate regions responsible for certain traits are popular methods that can be skewed by gene flow and population structure.
-%This being the case, what are the biases we can expect to see in real data for certain levels of introgression. 
 Finally, we trace back through the genealogical history of adaptive variants in new lakes to find the precise origin is of those variants.
-%\textbf{Jared - My attempt at concisely stating are questions}
 
@@
 This suggests downstream gene flow from freshwater populations is key to this system of ``transportation".
-%With lower $M < 1$ individuals per lake per generation we find either the lakes adapt indipendently by ``\textit{de novo}" mutation, or the selection on SGV is slow. (greater than 2000 gens.)
 We show that low levels of introgression also results in noise that often presents itself as $F_{st}$ peaks when scanning across the genome suggesting introgression plays a key role in inferring causal loci .
 With higher $M > 10$ individuals per lake per generation we see the impact of migration load, preventing the demes from fully adapting to their selective pressure.
-%When tracing back to the origin of adaptive alleles being *used* in new lakes that had adapted, we find the large majority of the alleles had been ``captured" meaning they were present in the individuals that initially colonized the new lake,
-%common in the original lakes, and uncommon in the ocean. 
-%This finding is in opposition to the claim that freshwater alleles are transported through subsequent migration, after the lakes have been colonized. 
-%\textbf{Jared - My attempt at concisely stating the results}
-
-%We ask what scale of historical introgression between populations with selective pressures $X$ and $Y$, allows for rapid and parallel local adaptation of a population derived from $X$ to an introduced environment with selective pressure $Y$?
-%We ask how the variants are structured, historically and across the genome underlying the trait? 
-%Furthermore, how can we infer causal (adaptive?) loci for regions of the genome that must be driving the rapid adaptation, from real data. 
-%Many biologists today make use of genome wide association studies (GWAS) and $F_{st}$ across the genome to estimate regions responsible for certain traits. 
-%Unfortunately, the data can be heavily influenced by the level of introgression and population structure of the samples. 
-%This being the case, what are the biases we can expect to see in real data for certain levels of introgression.
-
-%We find that ....
-\textbf{1.
-rewrite the previous paragraph to be more concise and tight, and 2.
-briefly restate our major findings which should line up well with those presented in the abstract.
-The following I grabbed from the beginning of the Results section}
-%*****
-%We varied the migration rate, $m$, from $5 \times 10^{-5}$ to $5 \times 10^{-2}$ per individual, i.e., between 0.01 and 10 migrants per year to and from each of the ten lakes.
-This had a strong effect on many aspects of adaptation, including how fast adaptation occurred in each lake, how much alleles were shared between lakes,
-%and the population genetic signals left behind.
-At the lowest migration rate, lakes adapted nearly independently, while at the highest migration rate, the habitats are beginning to act like a single metapopulation with substantial migration load.
-We will now dissect what happens between these extremes.
-Interestingly, the ability for separate populations to locally adapt to their own selective pressure was relatively unaffected until the highest rate of migration between the marine and freshwater environments.
-All rates of historical introgression aside from the lowest, helped both the initial and introduced freshwater populations share pre-existing freshwater adapted alleles.
-The sharing of alleles resulted in rapid adaptation ($\approx 100$ generations) of the introduced population (split from the marine population) to adapt to the freshwater environment.
-We also found that larger amounts of migration allow for more statistical power and less false positives in the resulting population genetics data ($F_{st}$ per SNP across the genome).
-%****
 
@@
 
-We explored these questions using forwards-time simulations with explicit genomic representation of a quantitative trait in SLiM \citep{haller2017slim,haller2018slim3}.
-The details of the model were motivated by current understanding of threespine stickleback history and demography,
-but remain simplistic in some aspects due to computational constraints.
-Possibly the most important caveat
-is that simulated population sizes are much smaller than the census size of the threespine stickleback population
+We explored these questions using forwards-time simulations with explicit genomic representation of individuals selecting upon a single continuous quantitative trait using SLiM \citep{haller2017slim,haller2018slim3}.
+The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography.
+We set up a realistic model of the relationship between marine and freshwater selective pressures with continuous space and migration rates between the populations to emulate gene flow. 
+Certain aspects of the the model remain simplistic due to computational constraints.
+Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback populations observed in nature
 (see the Discussion for more on this).
@@
 \paragraph{Habitat and geography}
-Each habitat type -- marine and freshwater -- have a fixed total of 5,000 diploid individuals each. 
+%The simulation begins with two subpopulations, one of both the
+Throughout the simulation there are two habitat types defined by their selective pressures.
+We start the simulations with two populations of size 5,000 diploid individuals in each of the habitats, for a total of 10,000 diploid individuals.
+%The two habitat types -- made to model the marine and freshwater selective pressures --  a fixed total populations size of 5,000 diploid individuals each. 
+%The habitat types differ in the optimal trait value.
+%The simulations begin with two subpopulations, each of size 5,000 diploid individuals, one set living in marine and freshwater selective pressures.
 The arrangement of these habitats, depicted in Figure \ref{fig:Geo},
 roughly models a set of freshwater habitats along a stretch of coastline. 
-The marine habitat  is a continuous, one-dimensional range of 25 units of length,
-while the freshwater habitat is divided into ten subpopulations (which we call ``lakes''),
+The marine habitat is a continuous, one-dimensional range of length 25 units,
+while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''),
 each connected to the marine habitat at regularly spaced intervals
@@
 This situation roughly models the cumulative effect of the various phenotypes such as armor morphology, 
-body size, craniofacial variation and opercle shape on which divergent selection is thought to act
-in the two environments. 
+body size, craniofacial variation and opercle shape on which divergent selection is thought to act on in the two environments. 
 Concretely, the optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively,
-and fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$
+and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$
 is determined by a Gaussian kernel with standard deviation 15, i.e.,
 \begin{align*}
-    f(x; x_\text{opt})
+    f(x_\text{ind}; x_\text{opt})
     &=
@@
             \left(
-            \frac{x-x_\text{opt}}{15}
+            \frac{x_\text{ind}-x_\text{opt}}{15}
             \right)^2
@@
 Note that only the scale of trait values relative to mutation effect size is relevant.
+In short, the difference between an individual's trait value and the optimum is the main determinant of an individuals fitness.
 We chose the difference between optima and strength of stabilizing selection in each habitat
@@
             After an initial period of 100K generations with 25 lakes,
-            an additional 25 lakes are added (at the same set of locations)
-            to simulate the appearance of newly accessible freshwater habitats.
+            an additional 25 lakes are added (at the same set of locations) and populated with a copy of the \textit{marine} individuals
+            to simulate the appearance of newly accessible freshwater habitats colonized by marine stickleback.
             The marine habitat, and each set of 25 lakes, each contain 5,000 individuals at all times.
@@
 \paragraph{Genetic architecture of the trait}
-Each individual carries two copies of a linear chromosome of $10^8$ loci.
+Each individual carries two linear chromosomes, each of size $10^8$ loci.
 Mutations that can affect the trait under selection can occur at rate $10^{-7}$ per locus per generation
@@
 spread evenly along the chromosome.
-Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability), 
-and with an effect sizes chosen randomly from an Exponential distribution with mean $1/2$, either positive or negative with equal probability. 
-Individual trait values are determined additively from the diploid genotypes. 
+Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
+Effect sizes for these mutations are chosen randomly from an exponential distribution with mean $1/2$, either positive or negative with equal probability. 
+Individual trait values ($x_\text{ind}$) are determined additively from the diploid genotypes. 
 Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value 
-$x = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, 
+$x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, 
 where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$.
@@
 \paragraph{Population dynamics}
-We use SLiM to simulate a Wright--Fisher population with nonoverlapping generations 
-and a fixed population size of 5000 diploid individuals in each habitat.
+We use SLiM to simulate a Wright--Fisher population with non-overlapping generations 
+and a fixed population size of 5,000 diploid individuals in each habitat.
 Each generation, the two parents of each new offspring are chosen proportional to their fitness 
@@
 ($10^{-7}$ per locus per generation). 
-Since the total population across \emph{all} lakes is regulated,
-to keep population sizes roughly constant within each lake, 
-before selection we divide fitness values of each freshwater individual 
-by the mean fitness in their lake,
-so that the mean fitnesses of all lakes are equal.
+Since the total population across \emph{all} 25 lakes is fixed at 5000,
+we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation.
+to do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, 
+so that the mean fitnesses of all lakes are equal before selection happens.
 
-Dispersal occurs both locally along the coastline in the marine habitat 
-and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
-There is no dispersal directly between lakes.
-The lake--ocean migration rate is denoted $m$, 
+\textbf{This next paragraph needs work, skipping for now}
+Dispersal of individuals occurs both locally along the coastline in the marine habitat 
+and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage (before the selection of parents). 
+There is no dispersal directly between, or within, lakes.
+The bi-directional lake--ocean migration rate is denoted $m$, 
 which we refer to as the rate of \emph{gene flow} between habitats.
-Each new individual in the marine habitat has freshwater parents with probability $m$; 
-to obtain the pair, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. 
+Parents for a new lake individual are chosen from the marine habitat with probability $m$:
+to obtain the pair of freshwater parents, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. 
 The resulting offspring is given a spatial location in the marine habitat at the location of the parent's lake. 
@@
 \paragraph{New freshwater populations} 
-To study how newly appearing freshwater habitats adapt, we introduce a new set of 25 lakes midway through the simulation.
-The initial set of individuals in these new lakes have parents chosen in the same way from the marine habitat as ocean-to-lake migrants,
-and act like an independent copy of the original set of lakes -- in particular, 
-the two sets of lakes each have 5,000 individuals at all times. 
-(Since this doubles the number of lake-to-marine immigrants, 
-after this happens the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.)
-
-We quantify genetic differentiation between the habitats with per-locus $F_{ST}$:
-if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats,
-respectively,
-and $\bar p = (p_f + p_m)/2$,
-then $F_{ST} = 1 - (p_f (1-p_f) + p_m (1-p_m))/ (2 \bar p (1-\bar p))$.
-
-\paragraph{Recording genealogical history}
-We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence}
-to record the genealogical history of all individuals alive 
-at the time of introduction of new lakes, the time of adaptation, and the end of the simulation.
-This allowed us to directly query the true origins of adaptive alleles,
-as well as allowing much larger simulations
-by avoiding the computationally expensive task of simulating neutral mutations
-(which were added after the fact at a rate of $10^{-7}$ per locus per generation, 
-as described in \citet{kelleher2018efficient}).
-
-The output tree sequence from each simulation allows us to,
-at each locus along the genome,
-construct the genealogical tree relating all extant chromosomes
-as well as all chromosomes present at time of introduction of the new lakes.
-Using these trees,
-%we classified each freshwater allele found in the new lakes at time of adaptation
-we classified each adaptive allele, in each genome,
-into four categories:
-\begin{enumerate}
-    \item \emph{De novo alleles:}
-        deriving from a new mutation that occurred in a new lake.
-    \item \emph{Migrant alleles:} 
-        %not present in the individuals that initially colonized the new lake, 
-        %but brought in by migration at a later time.
-        %\textbf{The allele CAN exist in individuals at time of introduction, this category mean the linage of a specific allele in an individual happened to come from a migrant}
-        deriving from a migrant that was not in the initial generation that colonized the lake
-    \item \emph{Captured alleles:}
-        present in the individuals that initially colonized the new lake, 
-        and both common (above 50\%) in the original lakes,
-        and uncommon (below 50\%) in the ocean.
-    \item \emph{Marine alleles:} 
-        present in the individuals that initially colonized the new lake, 
-        and not a captured allele.
-\end{enumerate}
-%\textbf{I may be conf}
-
-Roughly speaking, these four categories describe the contribution of
-(1) new mutation, 
-(2) post-colonization migration, 
-(3) standing variation at migration--selection balance, and
-(4) standing variation at mutation--selection balance.
-We could partition migrant alleles into finer categories
-(e.g., we do not distinguish whether they are common in the original lakes or not),
-but this category turns out to be small.
-
-%% we did this, but it is not used in the results
-% We also used the tree sequence to measure the realized fitness of individuals 
-% in the initial generation of the new lakes,
-% measured as
-% the total number of eventual offspring, in the generation at the time of adaptation, 
-% which have inherited a adaptive variant from that individual. 
-% To calculate this, we only considered gene trees which overlapped an adaptive variant site (defined at time of adaptation).
-% We then counted the total number of offspring which had inherited from any adaptive variant site, 
-% for each individual in the initial generation.
-
-We were also able to use the tree sequence to get information about the individual ``effect regions''
-in all initial genomes of the new lakes at time of introduction.
-From this we determined the ability for the new lakes to select upon the standing variation in the marine.
-Given the probability that effect region fixes at a given trait value, $2 * x / 45$, where $x$ is the effect size of a haplotype,
-we calculated the expected total effect sizes of the haplotypes that fix.
-Concretely, it is the sum across the 10 effect regions of
-$$\sum_{i=1}^N \prod_{j < i} (1 - p_j) p_i x_i$$
-where $N$ is the number of genomes per lake, $p_i$ is the probability of fixation of the 
-$i^{th}$ haplotype, $x_{i}$ is the effect size of the $i^{th}$ haplotype, and the haplotypes are sorted in decreasing order by $x_{i}$.
-These numbers were computed assuming additivity of mutations, meaning the numbers produced are a slight overestimate.
+To study how newly appearing freshwater habitats adapt and select on the standing genetic variation in the marine, we introduce a new set of 25 lakes midway through the simulation at 100K generations.
+These new lakes are populated with a copy of the marine individuals to emulate a freshwater lake being colonized by its neighboring marine population. 
+In particular, after this point of introduction in the simulation there are:
+two \emph{sets} of lakes, and one marine population, each with 5,000 individuals for a total of 15,000 individuals being tracked in the simulation. 
+Since this introduction of lakes doubles the number of lake-to-marine immigrants, 
+the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.
 
@@
 then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
+
+\paragraph{Recording genealogical history}
+We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence}
+to output the genealogical history of all individuals
+at the time of introduction of new lakes, at the time of adaptation, and at the end of the simulation.
+This allowed us to directly query the true origins of adaptive alleles.
+In addition it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations
+which were retroactively added to the gene trees at a rate of $10^{-7}$ per locus per generation, as described in \citet{kelleher2018efficient}).
+
+The output tree sequence from each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
+To do this, we constructed the genealogical tree relating all extant chromosomes at each locus along the genome.
+Using these trees we classified each adaptive allele, in each genome in the new lakes at the time of adaptation, into four categories:
+\begin{enumerate}
+    \item \emph{a ``De novo" allele:}
+        deriving from a new mutation that occurred in a new lake.
+    \item \emph{a ``Migrant" allele:} 
+        deriving from a migrant that was not in the initial generation that colonized the lake
+    \item \emph{a ``Captured" allele:}
+        present in the individual that initially colonized the new lake, 
+        and both common (above 50\%) in the original lakes,
+        and uncommon (below 50\%) in the ocean.
+    \item \emph{a ``Marine" allele:} 
+        present in the individual that initially colonized the new lake, 
+        and not a captured allele.
+\end{enumerate}
+%\textbf{I may be conf}
+
+We then looked at the ratio of alleles from each origin to the total number of alleles being used as the basis for local adaptation in the new lakes.
+This essentially gives a measure on whether selection in the new environments made use of
+(1) new mutation, 
+(2) post-colonization migration, 
+(3) standing variation at migration--selection balance, and
+(4) standing variation at mutation--selection balance.
+In other words, we were able to quantify the contributions of these possible origins 
+of adaptive alleles to understand where the genomic basis of adaptation had derived from.
+
+We were also able to use the tree sequence to get information about the individual ``effect regions''
+in all initial genomes of the new lakes at time of introduction.
+From this we determined the ability for the new lakes to select upon the standing variation in the marine.
+Given the probability that effect region fixes at a given trait value, $2 * x / 45$, where $x$ is the effect size of a haplotype,
+we calculated the expected total effect sizes of the haplotypes that fix.
+Concretely, it is the sum across the 10 effect regions of
+$$\sum_{i=1}^N \prod_{j < i} (1 - p_j) p_i x_i$$
+where $N$ is the number of genomes per lake, $p_i$ is the probability of fixation of the 
+$i^{th}$ haplotype, $x_{i}$ is the effect size of the $i^{th}$ haplotype, and the haplotypes are sorted in decreasing order by $x_{i}$.
+These numbers were computed assuming additivity of mutations, meaning the numbers produced are a slight overestimate.
 
@@
 	\begin{center}
-  		\includegraphics[width=0.7\linewidth]{Final_Plots/Allele_Origin.pdf}
+  		\includegraphics[width=0.7\linewidth]{Final_Plots/Allele_Origin_2.pdf}
   		\caption{ 

### 2018-12-11 Stickleback_Paper.tex

@@
 \paragraph{Habitat and geography}
-%The simulation begins with two subpopulations, one of both the
-Throughout the simulation there are two habitat types defined by their selective pressures.
+Throughout the simulation there are two habitat types -- Marine and Freshwater-- defined by their selective pressures.
 We start the simulations with two populations of size 5,000 diploid individuals in each of the habitats, for a total of 10,000 diploid individuals.
-%The two habitat types -- made to model the marine and freshwater selective pressures --  a fixed total populations size of 5,000 diploid individuals each. 
-%The habitat types differ in the optimal trait value.
-%The simulations begin with two subpopulations, each of size 5,000 diploid individuals, one set living in marine and freshwater selective pressures.
 The arrangement of these habitats, depicted in Figure \ref{fig:Geo},
@@
 \end{align*}
-Note that only the scale of trait values relative to mutation effect size is relevant.
-In short, the difference between an individual's trait value and the optimum is the main determinant of an individuals fitness.
+
+In short, the difference between an individual's trait value and the optimum largely determine's that individuals fitness. 
+Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
 We chose the difference between optima and strength of stabilizing selection in each habitat
@@
 
-\textbf{This next paragraph needs work, skipping for now}
-Dispersal of individuals occurs both locally along the coastline in the marine habitat 
-and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage (before the selection of parents). 
-There is no dispersal directly between, or within, lakes.
-The bi-directional lake--ocean migration rate is denoted $m$, 
+Dispersal occurs both locally along the coastline in the marine habitat 
+and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
+There is no dispersal directly between lakes.
+The lake--ocean migration rate is denoted $m$, 
 which we refer to as the rate of \emph{gene flow} between habitats.
-Parents for a new lake individual are chosen from the marine habitat with probability $m$:
-to obtain the pair of freshwater parents, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. 
+Each new individual in the marine habitat has freshwater parents with probability $m$; 
+to obtain the pair, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. 
 The resulting offspring is given a spatial location in the marine habitat at the location of the parent's lake. 
@@
 any new freshwater offspring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.
+\textbf{The paragraph above has a confusing structure, skipping for now}
 
@@
 
-%%%%%%%%%%%%%%%%%%
 \section*{Results}
 
-We varied the ocean--lake migration rate, $m$, 
-across separate simulations
-from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
-Since each lake contains 200 individuals,
-this corresponds to between 0.01 and 100 migrants per lake per generation.
-Many aspects of adaptation changed substantially across this range,
-including the speed of adaptation,
-degree of sharing of adaptive alleles between lakes,
-and the population genetic signals left behind.
-At very low rates of gene flow,
-each new lake's population adapted \emph{de novo},
-which took a very long time.
+To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
+Since each lake contains 200 individuals, we refer to these migration rates as between 0.01 and 100 migrants per lake, per generation.
+Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
+At very low rates of gene flow, each new lake's population adapted completely separately from \emph{de novo} mutation, which took a very long time ( $\approx 20,000$ generations).
 At very high rates of gene flow, local adaptation was almost impossible.
-Between these two extremes,
-genetic variation that allowed adaptation to freshwater habitats
-could move relatively easily between lakes --
+Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes.
+Surprisingly, only a few migrant per generation from the lakes to the marine were needed to see the effects of rapid local adaptation accelerated by selection on standing genetic variants in the new lakes.
 despite being deleterious in the intervening marine habitat --
-which allowed populations arriving in new lakes to adapt quickly,
-reassembling collections of alleles responsible for freshwater adaptation
-in other lakes.
+which allowed populations arriving in new lakes to adapt quickly, reassembling collections of alleles responsible for freshwater adaptation in other lakes.
 
-%%%%%%%%%%%%
 \subsection*{Local Adaptation: differentiation with gene flow}
@@
 at up to 10 migrants per lake per generation.
-% if computed using only sites with alleles affecting the trait (``effect mutations''),
-% $F_{ST}$ between habitats was relatively constant across migration values.
 
@@
         		comes within 0.5 of the original lakes average phenotype. 
-		Each point represents a single simulation run,
-		and the yellow dashed line is the average of all points at each respective parameter value.
+		Each point represents a single simulation run.
         		(Adaptation did not occur at the highest rate of gene flow.)
-        \textbf{where are the multiple points?, They do not exist, we only ran the one set. They each take $\approx$ 36 hours to run; but I can run them again if you want.} 
         } \label{fig:TimeToAdaptation}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Sharing of freshwater adapted alleles}
 
-%FIGURES THAT SHOW THIS:
-%-MPAA / IND
 At low migration rates, the \emph{initial} period of adaptation
@@
 but each is paired with a compensatory allele with effect size $+1$.
-Each local haplotype is therefore neutral, 
-and so likely to be at higher frequency in the marine population.
+Each local haplotype is therefore neutral
+(and so likely to be at higher frequency in the marine population).
 
-Suppose that the alleles present in the initial population of a new lake
-have effect sizes $(x_1, \ldots, x_n)$,
-and that the probability of fixation of a single allele of effect size $x$ is $p(x)$.
-Let $X_i$ be the average per capita effect of the $i^\text{th}$ allele after fixation,
-so that $X_i = 0$ if the allele is lost,
-and $X_i = 2x_i$ if the allele fixes (ignoring dominance).
-Therefore, the expected contribution to the final trait value of allele $i$
-is $\E[X_i] = 2 x_i p(x_i)$.
-The trait value after the initial round of adaptation is $\sum_i X_i$,
-and so the expected trait value is $2 \sum_i x_i p(x_i)$.
-Since there are many alleles that contribute more-or-less additively to the trait,
-the probability of fixation of a particular allele is not easy to find,
-but we can roughly approximate it as follows:
-if the current population trait mean $z$,
-then an allele that moves the trait a distance $x$ closer to the optimum
-has a fitness advantage in heterozygotes of 
-$\exp(-\beta((z - x)^2) / \exp( - \beta z^2)) \approx 1 + 2 \beta z x$, 
-where in our parameterization, $\beta = 1 / 450$.
-Since the probability of fixation of a single allele with relative advantage $s$
-is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical},
-then with $z=10$ we can use $p(x) \approx 4 x / 45$.
-Putting these considerations together,
-a rough estimate for the total shift in trait value resulting from selection
-on standing variation in the initial population is
-$R = \sum_i 8 x_i / 45$,
-where the sum is over all trait-affecting alleles.
-This approximation should be good if each trait-affecting allele is rare (present in only a few copies),
-effects are additive, and most are unlinked.
+Indeed, this appears to be the case.
+To quantify the genetic variation available \emph{without} recombining within the ten genomic regions,
+we first found, within each population, the haplotype with the largest net negative effect
+at each of the ten genomic regions.
+Summing these ten numbers, 
+we get the maximum amount that selection could move the population in the freshwater direction
+without recombining within regions.
+The mean of this value across the 25 populations is shown 
+in the first row (``best'') of Table~\ref{tab:linkage} --
+typical populations at $m=0.1$ migrants per lake per generation could shift to a phenotype of -12.02
+(and so have sufficient variation to adapt without recombination),
+but the ``best'' haplotypes populations at $m=1.0$ have effect sizes nearly twice as big 
+(a mean total of -21.02).
 
-We computed $R$ to confirm our suspicion that linkage slowed down adaptation
-by showing that there is sufficient allelic variation in the initial populations
-to complete adaptation across both migration rates (0.1 and 1 migrants per lake per generations),
-but not if the haplotypes of each genomic region are assumed to not recombine.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Theoretical expectations}
 
-We calculated for each individual in the initial generation of the new lakes
-the total contribution to trait values of each of the ten genomic regions 
-that containing trait-affecting mutations.
-The recombination rate within each of these ten regions is $10^{-2}$,
-so these regions act initially more-or-less like single loci.
-Even though the trait is additive across loci,
-the presence of sufficiently many distinct alleles to attain, in principle,
-the optimal freshwater trait value
-does not ensure adaptation will actually occur with these alleles.
-On the contrary, many of these will likely be lost to drift.
-
-To get a rough idea of whether the genetic variation present in the initial population
-is sufficient for the population to adapt,
-we translated the total effect size $x$ of each region 
-to a probability of fixation
-(as described below $p(x) \approx 2x/45$),
-and then calculated $\sum_i x_i p(x_i)$.
-If a region moves the trait an average of $x$ units towards the freshwater optimum
-(averaging over dominance relationships),
-Given the probability that an ``effect region'' fixes at a given trait value, 
-$2 \times x / 45$ where $x$ is the effect size of a haplotype,
-we were able to determine expected total effect sizes of the haplotypes that fix (detail about computation in methods).
-Below, we see the mean of these expected total effect sizes across all lakes.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{rrrrr}
-      \hline
-     & m=0.01 & m=0.1 & m=1 & m=10 \\ 
-      \hline
-          best & 2.77 & -12.02 & -21.02 & -7.02 \\ 
-      expected & 4.22 & -4.78 & -14.64 & -5.47 \\ 
-       \hline
-    \end{tabular}
-    \caption{
-        Haplotypic variation present in the new lakes at time of colonization,
-        across rates of gene flow.
-        \textbf{``Best''} shows the most negative trait value achievable
-            with intact haplotypes, averaged across populations.
-        \textbf{``Expected''} shows the trait value
-            achieved by averaging haplotypic values weighted by fitness
-            (see text for details).
-        This suggests that populations at $m=1$ have sufficient genetic variation
-        to completely adapt (reach $-10$) using intact haplotypes,
-        while at $m=0.1$ variation must be released from haplotypes by recombination.
-    } \label{tab:linkage}
-\end{table}
-
-Between $m = 0.1$ and $m = 1.0$ we see a large difference in the total effect sizes of the haplotypes that are expected to fix.
-Given these numbers were computed assuming additivity (slight overestimate) and individuals are diploid,
-at $m = 0.1$ the mean total genotype expected to fix in the populations puts individuals at a trait value of $\approx -8$ 
-after selection on alleles haplotypes in the initial population. 
-In turn, the population must wait for novel mutation or subsequent migration to provide the alleles for the remaining 2 units before 
-attaining a mean average trait value at the optimum (-10).
-It follows in Figure \ref{fig:phenotype_ts2}B that the majority time to adaptation is spent within 2 units of mean trait value.
-The initial adaptation on selection of standing variation up to that point happens quite rapidly.
-This provides evidence that linkage to marine alleles has a major impact in the introduced population's ability 
-to select and rebuild a pre-existing haplotype given the standing variation at time of introduction.
-
-\subsection*{Theoretical expectations}
-\plr{does this go first or second?}
-
-Here's some rough calculations to get a sense for what should be going on. Everything is done in more detail in OTHER PLACES WE SHOULD CITE.
-Fisher, Chevin, etc.
+We now revisit our observations above
+in the light of population genetics theory.
+The discussion will be rough,
+with a focus on intuition rather than precise calculations.
+Since within each population we model stabilizing selection on an additive trait
+controlled by a moderately large number of loci,
+more precise expectations might be obtained through quantitative genetics \citep{svardal2014general}
+or even Fisher's geometric model \citep{barton2001hybridization,chevin2014niche}.
 
@@
 -- then the probability that it escapes demographic stochasticity to become common in the population 
-is approximately $2s$ \citep{fisher,prob_fixation}. 
+is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical}
+(assuming Poisson reproduction, as we roughly have here).
 If the current population all differed from the optimum trait by $z$, 
@@
 \paragraph{New mutations}
-The total rate of appearance of new mutations per lake is $\mu_L = 0.04$, which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive in either direction.
+The total rate of appearance of new mutations per lake is $\mu_L = 0.04$, 
+which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive in either direction. 
 This implies that a new additive or dominant effect mutation appears once every 87.5 generations, on average. 
-Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, and so the probability that a dominant mutation manages to establish in a population differing from the optimum by $z$ is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$, and so the rate of establishment of dominant mutations is $\beta z / 87.5$, i.e., about one such mutation every $2461/z$ generations.
-The distribution of these successfully established mutations  has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2.
-Since additive alleles have half the effect in heterozygotes, they have half the probability of establishment.
+Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, 
+and so the probability that a dominant mutation manages to establish in a population differing from the optimum by $z$ 
+is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$, 
+and so the rate of establishment of dominant mutations is $\beta z / 87.5$, 
+i.e., about one such mutation every $2461/z$ generations. 
+The distribution of these successfully established mutations 
+has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2. 
+Since additive alleles have half the effect in heterozygotes, 
+they have half the probability of establishment. 
 During the initial phase of adaptation, the populations begin at around distance $z=10$ from the optimum. 
-Combining these facts, we expect adaptive alleles to appear through mutation at first on a time scale of 250 generations, with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
+Combining these facts, we expect adaptive alleles to appear through mutation within lakes at first on a time scale of 250 generations, 
+with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
+This agrees roughly with what we see in Figures~\ref{fig:phenotype_ts2} and XXX.
+\plr{refer to additional trace plots in the supp}
 
 \paragraph{Standing variation}
-An allele that moves the trait $z$ units in the freshwater direction in heterozygotes has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment (which is close to optimal).
-The product of population size and fitness differential in the marine environment for a mutation with $z=1$ is therefore $2Ns = 8.9$, implying that these alleles are strongly selected against but might occasionally drift to moderate frequency.
-The average frequency of such an allele in the marine environment at migration-selection equilibrium is equal to the total influx of alleles per generation divided by the selective disadvantage, which if $M = 2000 m$ is the number of immigrants per generation, is $2 M / \beta z^2$.
-With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$: since lakes have 400 genomes, as long as $M \ge 1$, the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes will appear at least once in the fish that colonize a new lake.
-However, an allele with $z=1$ only has probability of around 1/20 of establishing locally, suggesting that we'd need $M \ge 10$ to ensure enough pre-existing genetic variation that adaptation would happen entirely using the initial set of colonizers.
-This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.
+An allele that moves the trait $z$ units in the freshwater direction in heterozygotes 
+has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment. 
+The product of population size and fitness differential in the marine environment for a mutation with $z=1$ is therefore $2Ns = 8.9$, 
+implying that these alleles are strongly selected against but might occasionally drift to moderate frequency. 
+The average frequency of such an allele in the marine environment at migration-selection equilibrium 
+is equal to the total influx of alleles into the ocean per generation divided by the selective disadvantage, 
+which if there are $M$ immigrants per generation, is $2 M / \beta z^2$. 
+\plr{This $M$ is the total number in the ocean! Make sure this differs from notation above for the number per lake.}
+Each new lake is likely to contain a few copies of alleles at frequency above 1/400
+(since each new lake is begun with 400 genomes).
+With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$,
+so if $M \ge 1$, 
+the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes 
+will appear at least once in the fish that colonize a new lake. 
+However, an allele with $z=1$ only has probability of around 1/20 of establishing locally, 
+suggesting that the migration rate should be somewhat higher to ensure enough pre-existing genetic variation 
+that adaptation would happen entirely using the initial set of colonizers. 
+% This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.
+However, this calculation treats each allele independently;
+in practice we found that standing freshwater variation in the ocean
+were masked by linkage to compensatory marine variants.
 
 \paragraph{Migration}
-If sufficient genetic variation is not present in a new lake initially, it must appear by new mutation or migration.
-Since a proportion $m$ of each lake is composed of migrants each generation, it takes $1/m$ generations until the genetic variation introduced by migrants equals the amount initially present at colonization.
-This implies a dichotomy: either (a) adaptation is possible using variants present at colonization or arriving shortly thereafter, or (b) adaptation takes many multiples of $1/m$ generations.
+The key quantity regulating the amount of standing variation in the ocean was the \emph{downstream} migration rate,
+from lakes to the ocean.
+How important is the upstream migration rate?
+If sufficient genetic variation is not present in a new lake initially, 
+it must appear either by new mutation or by migration. 
+Since a proportion $m$ of each lake is composed of migrants each generation, 
+it takes $1/m$ generations until the genetic variation introduced by migrants equals the amount initially present at colonization. 
+This implies a dichotomy: either 
+(a) migration is high, and adaptation is possible using variants present at colonization or arriving shortly thereafter, or 
+(b) migration is low, so adaptation takes many multiples of $1/m$ generations.
+Since in our model lower migration also reduces the amount of variation available in the ocean,
+we expect very little contribution of subsequent migration across any value of $m$,
+as seen in Figure~\ref{fig:Origin}.
 
-These calculations depend on there being no bottleneck in colonization of the lake; if there is a bottleneck, then an additional factor must be added.
+% These calculations depend on there being no bottleneck in colonization of the lake; if there is a bottleneck, then an additional factor must be added.
 At what point do we expect new mutation to be more important than migration for adaptation?
@@
 If this does not happen, then we expect adaptation to take a multiple of $1/m$ generations;
-with our values, $1/m$ ranges from 20,000 to 20 generations.
+% with our values, $1/m$ ranges from 20,000 to 20 generations.
 Above we estimated that adaptive alleles due to new mutation fix locally about every 2,000 generations, suggesting that at our second-lowest migration rate (where $1/m = 2,000$), the two contributions of migration and new mutation are roughly equal.
-This is in fact what we see: in Figure \ref{fig:MPFAI}, we see that at the second migration rate, alleles start to be shared between lakes, while by the third migration rate, they are almost entirely shared.
+% This is in fact what we see: in Figure \ref{fig:MPFAI}, we see that at the second migration rate, alleles start to be shared between lakes, while by the third migration rate, they are almost entirely shared.
 
@@
 We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
-Selection is able to rebuild the freshwater haplotype from marine populations as a medium between all freshwater populations. 
+Selection is able to rebuild the freshwater haplotype at a rapid pace (in tens of generations) from marine populations as a medium between all freshwater populations. 
 Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.
-
-%Systems in which there is one, static selective pressure with a wide range connected to many, much smaller systems
-
+Surprisingly, we have discovered the genetic basis for local adaptation which accelerated by standing genetic variation to be derived largely from the initial generation of inhabitants. 
+This is in opposition to the assumption that subsequent migration is also necessary for this process to take place.
+In turn, this means that the downstream movement of alleles \emph{from} the lakes \emph{to} the oceans is much more important in this system than upstream migration events. 
 We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
@@
 \jgg{TODO: Talk about the role recombination plays}
-\jgg{Below may belong in the discussion / future work}
 This also brings to light the role of recombination in a system of migration selection balance. 
@@
 \end{document}
+
+The calculation above ignores allele frequencies -- 
+even if a population \emph{could} adapt entirely using the best possible haplotypes,
+many of these may be lost to drift.
+To account for this, we calculated a similar quantity but accounting for the probability of fixation.
+Suppose that the alleles present in the initial population of a new lake
+have effect sizes $(x_1, \ldots, x_n)$,
+and that the probability of fixation of a single allele of effect size $x$ is $p(x)$.
+Let $X_i$ be the average per capita effect of the $i^\text{th}$ allele after fixation,
+so that $X_i = 0$ if the allele is lost,
+and $X_i = 2x_i$ if the allele fixes (ignoring dominance).
+Therefore, the expected contribution to the final trait value of allele $i$
+is $\E[X_i] = 2 x_i p(x_i)$.
+The trait value after the initial round of adaptation is $\sum_i X_i$,
+and so the expected trait value is $2 \sum_i x_i p(x_i)$.
+Since there are many alleles that contribute more-or-less additively to the trait,
+the probability of fixation of a particular allele is not easy to find,
+but we can roughly approximate it as follows:
+if the current population trait mean $z$,
+then an allele that moves the trait a distance $x$ closer to the optimum
+has a fitness advantage in heterozygotes of 
+$\exp(-\beta((z - x)^2) / \exp( - \beta z^2)) \approx 1 + 2 \beta z x$, 
+where in our parameterization, $\beta = 1 / 450$.
+Since the probability of fixation of a single allele with relative advantage $s$
+is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical},
+then with $z=10$ we can use $p(x) \approx 4 x / 45$.
+Putting these considerations together,
+a rough estimate for the total shift in trait value resulting from selection
+on standing variation in the initial population is
+$R = \sum_i 8 x_i / 45$,
+where the sum is over all trait-affecting alleles.
+This approximation should be good if each trait-affecting allele is rare (present in only a few copies),
+effects are additive, and most are unlinked.
+
+We computed $R$ to confirm our suspicion that linkage slowed down adaptation
+by showing that there is sufficient allelic variation in the initial populations
+to complete adaptation across both migration rates (0.1 and 1 migrants per lake per generations),
+but not if the haplotypes of each genomic region are assumed to not recombine.
+
+We calculated for each individual in the initial generation of the new lakes
+the total contribution to trait values of each of the ten genomic regions 
+that containing trait-affecting mutations.
+The recombination rate within each of these ten regions is $10^{-2}$,
+so these regions act initially more-or-less like single loci.
+Even though the trait is additive across loci,
+the presence of sufficiently many distinct alleles to attain, in principle,
+the optimal freshwater trait value
+does not ensure adaptation will actually occur with these alleles.
+On the contrary, many of these will likely be lost to drift.
+
+Between $m = 0.1$ and $m = 1.0$ we see a large difference in the total effect sizes of the haplotypes that are expected to fix.
+Given these numbers were computed assuming additivity (slight overestimate) and individuals are diploid,
+at $m = 0.1$ the mean total genotype expected to fix in the populations puts individuals at a trait value of $\approx -8$ 
+after selection on alleles haplotypes in the initial population. 
+In turn, the population must wait for novel mutation or subsequent migration to provide the alleles for the remaining 2 units before 
+attaining a mean average trait value at the optimum (-10).
+It follows in Figure \ref{fig:phenotype_ts2}B that the majority time to adaptation is spent within 2 units of mean trait value.
+The initial adaptation on selection of standing variation up to that point happens quite rapidly.
+This provides evidence that linkage to marine alleles has a major impact in the introduced population's ability 
+to select and rebuild a pre-existing haplotype given the standing variation at time of introduction.

### 2018-12-11 Stickleback_Paper.tex
Supp and revisions to results

@@
 
-In short, the difference between an individual's trait value and the optimum largely determine's that individuals fitness. 
+In short, the difference between an individual's trait value and the optimum largely determines that individual's fitness. 
 Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
@@
 
+\textbf{The paragraph below has a confusing structure, skipping for now ... }
 Dispersal occurs both locally along the coastline in the marine habitat 
@@
 any new freshwater offspring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.
-\textbf{The paragraph above has a confusing structure, skipping for now}
 
@@
 \end{enumerate}
-%\textbf{I may be conf}
 
@@
 Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
-At very low rates of gene flow, each new lake's population adapted completely separately from \emph{de novo} mutation, which took a very long time ( $\approx 20,000$ generations).
-At very high rates of gene flow, local adaptation was almost impossible.
+At very low rates of gene flow, each new lake's population adapted completely separately from \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations).
+At very high rates of gene flow, local adaptation was constrained by the large influx of maladaptive alleles.
 Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes.
-Surprisingly, only a few migrant per generation from the lakes to the marine were needed to see the effects of rapid local adaptation accelerated by selection on standing genetic variants in the new lakes.
-despite being deleterious in the intervening marine habitat --
-which allowed populations arriving in new lakes to adapt quickly, reassembling collections of alleles responsible for freshwater adaptation in other lakes.
+Surprisingly, only a few migrants per generation from the lakes to the marine were needed
+to see the effects of rapid local adaptation accelerated by selection on standing genetic variants in the new lakes -- despite being deleterious in the intervening marine habitat.
 
 \subsection*{Local Adaptation: differentiation with gene flow}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics{Final_Plots/Pheno_Dist.pdf}
+  		\caption{
+            Distribution of mean trait values across generations of the simulation,
+            for different migration rates.
+            The dashed pink and purple lines at $\pm 10$ give the optimum phenotypes
+            in the marine and freshwater environments, respectively.
+		}
+  		\label{fig:MeanPhenotype}
+	\end{center}
+\end{figure}
 
@@
                 	Analogous plots for other migration rates
-                are shown in Figures \ref{fig:pheno_by_time1} and \ref{fig:pheno_by_time2}.
+                	are shown in Figure \ref{fig:phenotype_time_supp}.
 		}
-  		\label{fig:phenotype_ts2}
+  		\label{fig:phenotype_time}
 	\end{center}
@@
 
-\begin{figure}
-	\begin{center}
-  		\includegraphics{Final_Plots/Pheno_Dist.pdf}
-  		\caption{
-            Distribution of mean trait values across generations of the simulation,
-            for different migration rates.
-            The dashed pink and purple lines at $\pm 10$ give the optimum phenotypes
-            in the marine and freshwater environments, respectively.
-		}
-  		\label{fig:MeanPhenotype}
-	\end{center}
-\end{figure}
-
-Local adaptation occurred at most migration rates,
-as shown in Figure~\ref{fig:MeanPhenotype}.
-At the highest migration rate (at which half of each population was composed of migrants),
-populations had only slightly different average trait values.
-At lower migration rates (10 migrants per generation and below),
-populations adapted to local conditions:
-as shown in Figures \ref{fig:phenotype_ts2},
-\ref{fig:pheno_by_time1} and \ref{fig:pheno_by_time2},
-freshwater and marine populations diverged
-until the trait means were close to the optimal values in each habitat. 
-The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
-as jumps in the mean trait value,
+As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred at most migration rates with the exception of the highest migration at which half of each population was composed of migrants.
+At lower migration rates (10 migrants per generation and below), populations adapted to local conditions.
+As shown in Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp},
+freshwater and marine populations diverged until the trait means were close to the optimal values in each habitat. 
+The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_time} as jumps in the mean trait value,
 which move the trait by an amount of order 1 every few hundred generations.
@@
 
-\begin{comment}
-\paragraph{Standing genetic variation} 
-Despite a dramatic difference in the amount of allelic sharing between lakes, 
-standing genetic variance (Figure \ref{fig:SGV}) 
-was around 0.05 in the freshwater habitat across all but the highest migration rates. 
-Concurrently, genetic variation in the marine population 
-steadily increased with migration to a similar value. 
-On the one hand, it is not surprising that lakes, as a group, 
-carry more genetic variation than the marine habitat, 
-since population subdivision allows different alleles to become common in each. 
-However, it is striking that at $m=.005$, 
-the marine habitat carries as much genetic variation, 
-despite a lack of any substantial migration load.
-\plr{perhaps we want to remove this bit? revisit if it is telling us something important.}
-\end{comment}
+%\begin{comment}
+%\paragraph{Standing genetic variation} 
+%Despite a dramatic difference in the amount of allelic sharing between lakes, 
+%standing genetic variance (Figure \ref{fig:SGV}) 
+%was around 0.05 in the freshwater habitat across all but the highest migration rates. 
+%Concurrently, genetic variation in the marine population 
+%steadily increased with migration to a similar value. 
+%On the one hand, it is not surprising that lakes, as a group, 
+%carry more genetic variation than the marine habitat, 
+%since population subdivision allows different alleles to become common in each. 
+%However, it is striking that at $m=.005$, 
+%the marine habitat carries as much genetic variation, 
+%despite a lack of any substantial migration load.
+%\plr{perhaps we want to remove this bit? revisit if it is telling us something important.}
+%\end{comment}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Realized genetic architecture}
@@
 
-%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Origin of introduced freshwater adaptive alleles}
@@
 
-%wouldn't it be cool if we could use this fact and simulate introgression, starting with real data that is noisy to extrapolate the areas we care about?
-%this could be nonsense
-
-% at low M we should be able to predict something ---
-% at high M we should be able to predict something ----
-
-%Here, we suggest a range of migration rate (introgression) parameter values which would allow for the rapid (and in turn, parallel) adaptation 
-%of marine stickleback introduced intro a freshwater environment.
-%While this range is heavily effected by all other selections of parameter values.
-
-%(1) There's a threshold of introgression for both rapid adaptation and migration load. 
-%(1.5) Maybe refer to the literature for migration load.  
 \subsection*{thresholds}
@@
 \subsection*{connect results back to real data?}
-
-%In a species that commonly is subjected to two general types of selective pressure 
-%such as marine and freshwater stickleback this is suggestive of the benefit of hybridization in all subpopulations before
-%reaching a point of pre or post zygotic
-
-%(3) discuss what signals researchers could possibly look out for or connect to real data
 
@@
 \section*{Supplementary material}
-\begin{comment}
+
 \begin{figure}
@@
 	\begin{center}
-        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-2.png}
-        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
+  		\includegraphics[width=1.0\linewidth]{Final_Plots/Pheno_Time_supp.pdf}
   		\caption{
-            \textbf{Mean phenotypes through time},
-            at migration rates of
-            \textbf{(top)} 10 and
-            \textbf{(bottom)} 1 migrants per lake per generation.
+		Mean individual trait values in the marine habitat (blue line),
+        		the original lakes (yellow lines; average in orange),
+        		and the new lakes (light green lines; average in dark green),
+        		across the course of two simulations, with migration rates of
+        		\textbf{(A)} $m=5 \times 10^{-3}$,
+        		\textbf{(B)} $m=5 \times 10^{-2}$ and
+		\textbf{(B)} $m=5 \times 10^{-1}$.
+        		(i.e., 1, 10, and 100 migrants per lake per generation, respectively).
+        		Optimal trait values in the two habitats are at $\pm 10$.
 		}
-  		\label{fig:pheno_by_time1}
+  		\label{fig:phenotype_time_supp}
 	\end{center}
 \end{figure}
-
-\begin{figure}
-	\begin{center}
-        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
-        \includegraphics[width=1.0\linewidth]{plotlyPlots/PhenotypeThroughout5e-5.png}
-  		\caption{
-            \textbf{Mean phenotypes through time},
-            at migration rates of
-            \textbf{(top)} 0.1 and
-            \textbf{(bottom)} 0.01 migrants per lake per generation.
-		}
-  		\label{fig:pheno_by_time2}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-        \includegraphics[width=1.0\linewidth]{plotlyPlots/StandingGeneticVariation.png}
-  		\caption{
-            \textbf{Standing genetic variance},
-            computed as XXX write if we leave this in XXX.
-		}
-        \label{fig:SGV}
-	\end{center}
-\end{figure}
-\end{comment}
 

### 2018-12-12 Stickleback_Paper.tex
latest

@@
 However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
-How do complex geographic systems with a variety of selective pressures interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations?
+In addition, we would like to know how complex geographic systems, with a variety of selective pressures, interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations.
 
-Recently, a large flood of data resulting from systems such as this come from population genomic studies in nature that have been powered by advances in sequencing technologies. 
-One such organism is the threespine stickleback which has exhibited a long standing evolutionary history of migration-selection balance between marine and freshwater populations across the globe.
-The rigidity of the species' ability to frequently prosper in newly created freshwater environments has made this fish a good model for understanding the genetic basis of adaptive evolution. 
-The ancestral marine form of Stickleback has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
-Geographic isolation prevents direct migration between a large majority of these patches of freshwater populations,
-yet we observe very similar phenotypes and many shared adaptive alleles often found to be identical by decent (IBD). 
+Recently, a large flood of data resulting from evolutionary systems of organisms have come from population genomic studies powered by advances in sequencing technologies. 
+One such organism is the threespine stickleback. 
+Many studies of species have exhibited a long standing evolutionary history of migration-selection balance between marine and freshwater populations across the globe.
+The rigidity of the stickleback's ability to frequently prosper in newly created freshwater environments has made this fish a good model for understanding the genetic basis of adaptive evolution. 
+The ancestral marine form of stickleback has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
+While geographic isolation prevents direct migration between a large majority of these patches of freshwater populations,
+They appear to exhibit very similar phenotypes, as well as many shared, adaptive alleles often found to be identical by decent (IBD). 
 Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in tens of generations, 
@@
 But how can evolution occur at such a rapid pace? 
-Waiting for new mutations to arise in each lake would take much longer, and genotypes being identical by state is even more improbable. 
+Waiting for new mutations to arise in each lake would take much longer, and the genotypes being identical by state is even more improbable. 
 This would seem to suggest that freshwater alleles are maintained in the marine individuals allowing the accelerated selection on SGV found in marine individuals which colonize the lake. 
@@
 While the low lateral plate version of this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently. 
-More recently, population genomic studies employing genome-wide haplotype analyses has provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
+Contemporary population genomic studies employing genome-wide haplotype analyses has provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
 
@@
 
-We find that only a few stickleback per generation, per lake, suffice to maintain freshwater alleles at an acceptable frequency in marine environments for efficient ``transportation" of the freshwater haplotype.
+We find that only a few stickleback per generation, per lake, 
+suffice to maintain freshwater alleles from a (global?) haplotype at an acceptable frequency in marine environments for efficient ``transportation" of the freshwater haplotype.
 The selection on SGV at this level results in rapid local adaptation in $\approx 60$ generations, this follows closely to what we observe in nature (such as middleton island, cite this).
@@
             is coupled to randomly mating ``lakes'' at discrete locations with dark green arrows representing migration patterns.
-            %It is important to note there is no migration from any one lake to another, therefor, freshwater alleles must travel through the marine habitat as is often the case in nature
             After an initial period of 100K generations with 25 lakes,
@@
 
-As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred at most migration rates with the exception of the highest migration at which half of each population was composed of migrants.
-At lower migration rates (10 migrants per generation and below), populations adapted to local conditions.
-As shown in Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp},
-freshwater and marine populations diverged until the trait means were close to the optimal values in each habitat. 
-The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_time} as jumps in the mean trait value,
-which move the trait by an amount of order 1 every few hundred generations.
-Trait variation within each population was small compared to the difference between populations.
-Across all parameter values, differences at around 16 commonly polymorphic sites 
-(eight that shift the trait in each direction)
+As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred at migration rates of $m \le 10$.
+% with the exception of the highest gene flow at which half of each population was composed of migrants.
+%At lower migration rates (10 migrants per generation and below), populations adapted to local conditions.
+Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp}
+show freshwater and marine populations diverging until the trait means were close to the optimal values in each habitat. 
+The establishment of new alleles in the lakes is visible in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value.
+%which move the trait by an amount of order 1 every few hundred generations.
+%Trait variation within each population was small compared to the difference between populations.
+At the lowest rate of gene flow, $m = 0.01$, differences at $\approx 20$ commonly polymorphic sites 
+($\approx 10$ that shift the trait in each direction)
 were responsible for most of the adaptive differences between freshwater and marine habitats.
-
+In Figure \ref{fig:MPFAI}A, we can see that more gene flow simplified the genetic basis. %does this belong here?
 As expected, increasing migration rate decreased differentiation between habitats. 
@@
 at neutral sites steadily declines as migration increases.
-However, local adaptation was still able to occur despite overall homogenization 
-at up to 10 migrants per lake per generation.
 
@@
 shown in Figure \ref{fig:TimeToAdaptation} for different rates of gene flow.
-Adaptation of new lakes took over 18,000 generations at the lowest rate,
-while at one migrant per lake per generation
-new lakes managed to adapt in only 32 generations.
+Adaptation of new lakes took over 18,000 generations at the lowest rate of $m = 0.01$,
+while at $m = 1$ migrant per, lake per generation,
+new lakes managed to adapt in just under 60 generations. 
+%This dramatic decrease in time to local adaptation is suggestive of just how quickly
+%selection can act on standing variation. 
+%It also suggests a rough threshold of gene flow required 
+%to efficiently select upon the alleles present in the marine individuals.
 
@@
 At low migration rates, the \emph{initial} period of adaptation
-takes roughly ten times longer for lakes as it does for the ocean.
+takes roughly 25 times longer for lakes than it does for the ocean.
 This difference occurs because at low migration rates,
-adaptation occurs independently in each lake,
-and the marine habitat has ten times as many individuals,
-and therefore ten times the influx of new alleles than any one lake. 
+all subpopulation must wait for novel mutations to arise in order to adapt.
+Because the marine is one continuous population with 25 times more individuals than any one lake, 
+there is a much larger influx of new mutations to be selected upon. 
 At higher migration rates, greater mixing
-allows lakes to share alleles instead of developing their own genetic basis of adaptation. 
+allows the initial lakes to share alleles instead of developing their own genetic basis of adaptation. 
 As a first indication of this, Figure \ref{fig:Fst} 
@@
 at effect mutations decreased with migration rate.
+%We are interested in this shared genetic basis of adaptation because it is commonly thought that this is
+%a requirement for the efficient ``transportation" of the freshwater haplotype in to new lakes \citet{schluter2009genetics}
 
 To investigate in more depth how locally adaptive alleles found in the original lakes
-are shared between lakes, and how they spread to the new lakes,
-we counted ``pre-existing freshwater adapted alleles''.
-These we define for a particular generation 
-to be any mutation with an effect on the trait
-whose frequency is above 50\% in at least one original lake 
-and below 50\% in the marine habitat.
-Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles, 
-across generations.
-At the lowest migration rate (one migrant per 100 generations per lake), 
-each lake has a private set of about 10 mutations
-nearly fixed in that lake but not elsewhere:
-new lakes acquire new adaptive alleles and so have none of these.
-At one migrant per 10 generations per lake,
-the original set of lakes each developed private alleles,
-but these are present at low frequency in the ocean,
-and so new lakes adapt 
-mostly using a subset of this large repertoire of standing variation.
-As migration rate increases past this, 
-the number of pre-existing freshwater adapted alleles declines
-because alleles move between populations by migration
-before they can appear by new mutation,
-while the frequency of these alleles in the ocean
-stays relatively constant.
+are shared between lakes, as well as how they spread to the new lakes,
+we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation.
+To be considered in this category, a mutation must participate in the genetic basis of any freshwater adaptation.
+Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lakes and below 50\% in the marine habitat.
+Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles across generations.
+At $m = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere:
+the new lakes acquire new adaptive alleles and therefor have little to none of these adaptive alleles.
+At $m = 0.1$, we again observe the original lakes adapting nearly independently from each other.
+However, in Figure \ref{fig:MPFAI}A we can see the average marine individual is shown to hold $\approx 2$ pre-existing freshwater adapted alleles. 
+This being the case, we can see new lakes adapt using a subset of this large repertoire of standing variation deriving from a multitude of independently derived haplotypes.
+As migration rate increases past this, %we can see migration selection balance keeps the average capacity for marine individuals to carry freshwater alleles at about 2.
+we can see in Figure \ref{fig:MPFAI}C that the total number of pre-existing freshwater adapted alleles declines.
+%while the number of alleles which make up the genetic basis of adaptation in freshwater individuals remains relatively similar.
+%Although the genetic basis seems to simplify slightly due to more , this 
+This suggests that the alleles move between populations by migration 
+before selection acts on new mutation for neighboring lakes.
+Interestingly, the frequency of these alleles in the ocean
+stays relatively constant across the reasonable rates of gene flow.
 
-Figure~\ref{fig:MPFAI}B shows the distribution of 
+Figure~\ref{fig:MPFAI}B shows the distribution, through time, of 
 the mean percentage of currently-defined freshwater adapted alleles 
-that each genome in each of the populations carries, 
-averaged across time and individuals. 
+that each genome in each of the populations carries.
 If all individuals across lakes carried the same set of alleles determining their trait value, 
 this would be 100\%. 
-At the lowest migration rates,
+At the lowest migration rate $m = 0.01$,
 each genome in the original lakes have almost exactly~$1/25^{th}$ 
 of the total number of pre-existing freshwater adapted alleles -- 
-this is because each of the 25 lakes has adapted with a unique set of alleles. 
+this is because each one of the 25 lakes has adapted with a unique set of alleles. 
 Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
@@
 higher gene flow between populations allows sharing of freshwater alleles between populations.
+%however, it is not enough for the marine population to hold bits and peices from independently formed haplotypes,
+%For selection to act on SGV on the scale we observe in nature, the gene flow 
+%initial generation to be able to rebuild 
 We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting
-allele common in the new lakes at time of adaptation,
+allele, in each individual, common in any of the new lakes at time of adaptation
 as defined in the Methods.
-Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow,
-the majority of adaptive alleles are derived from de novo mutation.
+Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo"} mutation.
 As gene flow increases, a larger fraction of adaptive alleles 
@@
 
-At first, increased migration allows sharing of adaptive alleles between lakes, 
-but at the highest migration rate, 
-the constant influx of alleles between the habitats creates substantial migration load. 
+While increased migration allows sharing of adaptive alleles between lakes, 
+at $m = 10.0$ the constant influx of alleles between the habitats creates substantial migration load. 
 The rate at which migration load becomes substantial,
@@
 Now we take a closer look at the genomic architecture of local adaptation between the two habitats. 
-Do the alleles underlying trait differences cluster along the genome?
+%Do the alleles underlying trait differences cluster along the genome?
 Do measures of local differentiation identify the causal loci? 
@@
 (Note that we are pooling freshwater habitats; a single lake would provide substantially less power.)
-\plr{need to know where the FAA are in these figures}
+%\plr{need to know where the FAA are in these figures}
 Higher migration rates showed more distinct $F_{ST}$ peaks over polymorphic loci 
@@
 will only be successful given enough migration between habitats. 
-To quantify this, 
-Figure \ref{fig:Power_FP} shows the power and false positive rates 
+To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates 
 that would be obtained by an $F_{ST}$ cutoff 
@@
 
-In regions of the genome underlying individual trait value, we observed Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$.
+In regions of the genome underlying individual trait value, we observed 
+Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$. 
 Knowing exactly which mutations effect phenotype in our simulations, we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome.

### 2018-12-12 Stickleback_Paper.tex
plot mod and first attempt at discussion

@@
 While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures.
+What role does genomic variation play in a metapopulation with connection to an array of selective pressures?
+How could migration-selection balance actually benefit a super species as a whole in tandem with it experiencing divergent selection?
+\jgg{these q's make sense to me, but could be utter non-sense to an experienced pop geneticist. lmk}
 A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split.
@@
 	\begin{center}
-        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome.pdf}
+        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome_faa.pdf}
   		\caption{
@@
         		show regions which have the potential to affect phenotype.
+		All locations of pre-existing freshwater adapted alleles have been highlighted by 
+		blue vertical lines with an alpha value of 0.3, so darker shades of blue imply more 
+		freshwater adapted alleles at that location.
      } \label{fig:Fst}
@@
 
-Now we take a closer look at the genomic architecture of local adaptation between the two habitats. 
-%Do the alleles underlying trait differences cluster along the genome?
+Here, we take a closer look at the genomic architecture of local adaptation between the two habitats. 
 Do measures of local differentiation identify the causal loci? 
@@
 (Note that we are pooling freshwater habitats; a single lake would provide substantially less power.)
-%\plr{need to know where the FAA are in these figures}
+\plr{need to know where the FAA are in these figures}
+\jgg{check the new plot}
 Higher migration rates showed more distinct $F_{ST}$ peaks over polymorphic loci 
@@
 \plr{Need to fix caption on figure \ref{fig:Power_FP} before editing this bit.}
+\jgg{There is still something wrong with this figure besides the labels ... }
 
@@
 but much more rapid adaptation ($\approx 30$ generations). 
-Perhaps
-this 60$\times$ discrepancy occurs because 
+Perhaps this 60$\times$ discrepancy occurs because 
 new lakes must wait for an additional influx of freshwater alleles through migration
 beyond what is present in the initial generation?
-However, Figure~\ref{fig:Origin} shows that at both $m = 0.1$ and $m = 1.0$,
+This hypothesis is disproven by
+Figure~\ref{fig:Origin}, which shows that at both $m = 0.1$ and $m = 1.0$,
 the vast majority of alleles underlying the local adaptation
-derived from individuals in the original generation (``captured'' or ``marine'').
+derived from individuals in the original generation (``captured'').
 
@@
 maintains relatively intact freshwater haplotypes
-that can be more easily rebuilt.
+that can be more easily rebuilt in the marine environments.
 Recall that trait-affecting mutations only occur in relatively small regions of $10^5$ loci
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Theoretical expectations}
@@
 
-%sum up the system
-
-%dynamic. and static.
-
-% not subsequent
-
-% reasonable amounts of gene flow result, 
-% SGV is powerful, and onlt needs a small amount of gene flow.
-
-% downstream migration not as important. 
-% doesn't take a strange environment.
-
-% look up and cite migration load.
-
-We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
+In this paper, we have described a realistic simulation model to observe the effects of selection of genetic variation (SGV) across a wide range of gene flow rates.
+we have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
 Selection is able to rebuild the freshwater haplotype at a rapid pace (in tens of generations) from marine populations as a medium between all freshwater populations. 
 Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.
-Surprisingly, we have discovered the genetic basis for local adaptation which accelerated by standing genetic variation to be derived largely from the initial generation of inhabitants. 
-This is in opposition to the assumption that subsequent migration is also necessary for this process to take place.
-In turn, this means that the downstream movement of alleles \emph{from} the lakes \emph{to} the oceans is much more important in this system than upstream migration events. 
-We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
+
+While our findings prove useful to the understanding of stickleback, they are not autonomous to this species. 
+Now, we will discuss the broader implications of our findings on the more general framework of similar systems, we also explore what other questions remain. 
+Small patches of similar selective pressures can be found in many places such as high altitudes, underwater volcanic vents, or post forest fire environments. 
+%We are still finding new subspecies, that seem to have adapted to extreme environments presented in small patches connected to larger spanning biospheres where the ancestral species naturally inhabits.
+Recently, (Cite Nelson Ferretti et al), discovered seven new species of Hapalotremus (Theraphosinae), living at the extremely high altitudes distributed along the Andes and Yungas in western South America.
+With the ancestral species at much lower altitudes, beneficial high altitude alleles could potentially be carried to help in the colonization of neighboring mountain environments. 
+Systems of metapopulations like this could potentially arise in many places
+where the ancestral species carries around with it beneficial alleles for
+ selective pressures it has encountered in the past and remains connected to through hybridization events.
+
+%(1) the subspeciesthat inhabits a different selective pressure, and ancestral species are not constrained by pre, or post-zygotic isolation to participate in hybridization events, and
+%(2) the ancestral species spans across a wide range. 
+%For brevity, I will call evolutionary systems which satisfy these conditions, 
+%The vast number of species which encou
+
+ \paragraph{Atmosphere's ability to carry whoever scotty is beaming up ... }
+Perhaps the most notable finding in this study is that large, ancestral populations have the capacity to maintain alleles which are potentially deleterious for an extended amount of time.  
+%This system seems analagous the marvel superhero \emph{Rogue}, who can steal superpowers or memories from another  for a limited amount of time as long as she has contact with the indivi
+This principle, deriving solely from the properties of sex and natural selection, could be applicable to almost any species experiencing the influence of divergent selection with introgression.
+And to explore further how this process is carried out exactly, many of the questions lie in understanding which properties of genomic architecture and demographics allow this transportation to be possible.
+In terms of Star Trek's ``transporter" referred to by \citet{schluter2009genetics}, 
+This is largely analogous to exploring the specific properties of the atmosphere / space which allow for the heroes to be disassembled, then put back together.
+%While the perspective of our 
+In our simulations, the continued gene flow from freshwater populations meant that the alleles were maintained at a constant rate with fixed demographics. 
+Two possible questions that might arise from this are:
+(1) How do population sizes, demographics, and evolutionary histories impact an ancestral population's capacity for carrying maladaptive alleles in its opposing environment.
+(2) What role does genomic architecture play in the ability for a haplotype to be rebuilt
+
+\paragraph{Origins of genetic basis of local adaptation in new lakes}
+In Figure~\ref{fig:Origin} we found that the genetic basis for local adaptation, when selecting upon SGV, is propogated largely from the initial generation of inhabitants. 
+This contrasts the belief that subsequent migration is a necessity in newly established populations ability to rebuild a haplotype.
+This finding suggests that with high probability, a randomly chosen subset from ancestral species in the metapopulations like marine stickleback, 
+carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
+
+\paragraph{Realized genomic architecture}
+Additionally have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
 This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
@@
 
+%\paragraph{Implications for the ``Transporter"--hypothesis}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%DO WE NEED THIS?
+\begin{comment}
 \subsection*{thresholds}
@@
 How might this affect things?
+\end{comment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-12-12 Stickleback_Paper.tex
plot twist

@@
 	\begin{center}
-        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome_faa.pdf}
+        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome_faa_pos.pdf}
   		\caption{
@@
         		with columns corresponding to increasing gene flow from left to right.
-		Ten vertical dotted pink lines in each subplot 
-        		show regions which have the potential to affect phenotype.
+		%Ten vertical dotted pink lines in each subplot 
+        		%show regions which have the potential to affect phenotype.
 		All locations of pre-existing freshwater adapted alleles have been highlighted by 
-		blue vertical lines with an alpha value of 0.3, so darker shades of blue imply more 
+		pink dashed vertical lines with an alpha value of 0.3, so darker shades of blue imply more 
 		freshwater adapted alleles at that location.

### 2018-12-12 Stickleback_Paper.tex
edit to intro

@@
 While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures.
-What role does genomic variation play in a metapopulation with connection to an array of selective pressures?
-How could migration-selection balance actually benefit a super species as a whole in tandem with it experiencing divergent selection?
-\jgg{these q's make sense to me, but could be utter non-sense to an experienced pop geneticist. lmk}
+What role does genomic variation play in metapopulations with connection to an array of selective pressures?
+How could migration-selection balance and ecological speciation actually benefit species as a whole?
+\jgg{these q's make sense to me, but could be utter non-sense to an experienced pop geneticist.}
 A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split.

### 2018-12-12 Stickleback_Paper.tex
gramm

@@
 
-In this paper, we have described a realistic simulation model to observe the effects of selection of genetic variation (SGV) across a wide range of gene flow rates.
-we have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
+In this paper, we have described a realistic simulation model of a coastal metapopulation to observe the effects of selection of genetic variation (SGV) across a wide range of gene flow rates.
+We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
 Selection is able to rebuild the freshwater haplotype at a rapid pace (in tens of generations) from marine populations as a medium between all freshwater populations. 
@@
 Systems of metapopulations like this could potentially arise in many places
-where the ancestral species carries around with it beneficial alleles for
- selective pressures it has encountered in the past and remains connected to through hybridization events.
+where the ancestral species carries around with it beneficial alleles forselective pressures it has encountered in the past and remains connected to through hybridization events.
 
@@
 And to explore further how this process is carried out exactly, many of the questions lie in understanding which properties of genomic architecture and demographics allow this transportation to be possible.
-In terms of Star Trek's ``transporter" referred to by \citet{schluter2009genetics}, 
-This is largely analogous to exploring the specific properties of the atmosphere / space which allow for the heroes to be disassembled, then put back together.
+%In terms of Star Trek's ``transporter" referred to by \citet{schluter2009genetics}, 
+his is largely analogous to exploring the specific properties of the atmosphere / space which allow for the heroes to be disassembled, then put back together.
 %While the perspective of our 
@@
 Two possible questions that might arise from this are:
-(1) How do population sizes, demographics, and evolutionary histories impact an ancestral population's capacity for carrying maladaptive alleles in its opposing environment.
-(2) What role does genomic architecture play in the ability for a haplotype to be rebuilt
+(1) How do population sizes, demographics, and evolutionary histories impact an ancestral population's capacity for carrying maladaptive alleles in its opposing environment, and 
+(2) What role does genomic architecture play in the ability for a haplotype to be rebuilt.
 

### 2018-12-12 Stickleback_Paper.tex
rm note

@@
 How could migration-selection balance and ecological speciation actually benefit species as a whole?
-\jgg{these q's make sense to me, but could be utter non-sense to an experienced pop geneticist.}
 A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split.

### 2018-12-12 Stickleback_Paper.tex
rewrite pop dyn para

@@
 
-\textbf{The paragraph below has a confusing structure, skipping for now ... }
 Dispersal occurs both locally along the coastline in the marine habitat 
@@
 which we refer to as the rate of \emph{gene flow} between habitats.
-Each new individual in the marine habitat has freshwater parents with probability $m$; 
-to obtain the pair, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. 
-The resulting offspring is given a spatial location in the marine habitat at the location of the parent's lake. 
-Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$): 
+Each new individual in each habitat 
+has parents from the other habitat with probability $m$ 
+(in which case we call it a ``migrant''),
+and parents from the same habitat with probability $1-m$.
+% how do we get nonmigrants?
+The first parent of each nonmigrant individual in the freshwater habitat
+is chosen from the freshwater habitat proportional to fitness, 
+and a mate is chosen from the same lake as the first, also proportional to fitness. 
+The offspring lives in the same lake as the parents.
+Parents for each nonmigrant marine individual are chosen similarly:
 first, a single parent is chosen proportionally to fitness in the marine habitat,
 and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. 
-Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
-where $d_{ij}$ is the distance between the two locations. 
+Concretely, if the first parent is marine individual $i$, 
+then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
+where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
 Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.02, 
 and reflected to stay within the population.
-New offspring in the freshwater habitat are chosen in the same way, except the probability that the parents are marine individuals is $m$; 
-any new freshwater offspring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.
+% what about migrants?
+Parents for each freshwater migrant are chosen in the same way
+as for nonmigrant marine individuals,
+and are assigned to the lake nearest to the position of the first marine parent.
+Similarly, parents for each marine migrant are both chosen from the same lake as before,
+and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.
 

### 2018-12-12 Stickleback_Paper.tex
merge

--- 2018-12-12 Stickleback_Paper.tex
@@
 
-\textbf{The paragraph below has a confusing structure, skipping for now ... }
+%\textbf{The paragraph below has a confusing structure, skipping for now ... }
 Dispersal occurs both locally along the coastline in the marine habitat
+%When an offspring is generated, it's spatial location is calculated by 
 and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
-There is no dispersal directly between lakes.
+%There is no dispersal directly between lakes.
 The lake--ocean migration rate is denoted $m$, 
-which we refer to as the rate of \emph{gene flow} between habitats.
+%which we refer to as the rate of \emph{gene flow} between habitats.
 Each new individual in the marine habitat has freshwater parents with probability $m$; 
@@
 where $d_{ij}$ is the distance between the two locations. 
-Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.02, 
+Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, 
 and reflected to stay within the population.

### 2018-12-12 Stickleback_Paper.tex

--- 2018-12-12 Stickleback_Paper.tex
@@
 and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
-There is no dispersal directly between lakes.
 The lake--ocean migration rate is denoted $m$, 
@@
 where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
-Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.02, 
+Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, 
 and reflected to stay within the population.

### 2018-12-12 Stickleback_Paper.tex
Edits to realized genomic architecture

@@
 
-Dispersal occurs both locally along the coastline in the marine habitat 
-and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
+As shown in Figure~\ref{fig:Geo} as lines and arrows, dispersal occurs both locally along the coastline in the marine habitat
+as well as between the marine habitat and the lakes.
+All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
 The lake--ocean migration rate is denoted $m$, 
@@
 and a mate is chosen from the same lake as the first, also proportional to fitness. 
-The offspring lives in the same lake as the parents.
+The offspring then lives in the same lake as the parents.
 Parents for each nonmigrant marine individual are chosen similarly:
@@
 To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
-Since each lake contains 200 individuals, we refer to these migration rates as between 0.01 and 100 migrants per lake, per generation.
+Since each lake contains 200 individuals, we refer to these migration rates in terms of \emph{dispersal} ($M$) \jgg{($D?$)} ranging from $0.01$ to $100$ migrants per lake, per generation.
 Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
@@
         		with columns corresponding to increasing gene flow from left to right.
+		$F_{st}$ is calculated between all marine individuals, and all lakes pooled together.
 		%Ten vertical dotted pink lines in each subplot 
@@
 Here, we take a closer look at the genomic architecture of local adaptation between the two habitats. 
-Do measures of local differentiation identify the causal loci? 
-Figure \ref{fig:Fst} shows plots along the genome of per-locus $F_{ST}$ values between habitats.
-(Note that we are pooling freshwater habitats; a single lake would provide substantially less power.)
-\plr{need to know where the FAA are in these figures}
-\jgg{check the new plot}
-Higher migration rates showed more distinct $F_{ST}$ peaks over polymorphic loci 
-underlying trait differences between the habitats. 
+Do measures of local differentiation such as $F_{st}$ scans or Genome Wide Association Studies (GWAS) identify the causal loci? 
+Figure \ref{fig:Fst} shows plots along the genome of per-locus $F_{ST}$ values between the marine habitat and all freshwater habitats pooled together.
+Higher rates of dispersal showed more distinct $F_{ST}$ peaks over polymorphic loci 
 ``Background'' levels of $F_{ST}$ increase as gene flow decreases,
 swamping out this signal until the regions under selection are indistinguishable. 
-This is likely due to two reasons: 
+This is likely the contribution of two separate forces of natural selection: 
 first, stronger genetic drift with less migration leads to higher background $F_{ST}$, and 
@@
 
-This suggests that genome scans for local adaptation based purely on measures of differentiation 
+At first glance, this suggests that genome scans for local adaptation based purely on measures of differentiation 
 will only be successful given enough migration between habitats. 
+But how many of these peaks are actually underlying trait differences that form the basis for local adaptation?
 To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates 
@@
 that declared everything above a certain value to be a causal locus.
+What we observe most plainly in this graph is that $M = 1.0 \& M = 10.0$ migrants per lake, per generation
+are the only rates of dispersal which provide reliable true negatives when observing peaks
+This means that a large percentage of peaks past an $F_{ST}$ cutoff of $0.05$ 
+actually lie on top of regions which could impact trait differences. 
+Unfortunately, the low statistical power for all rates of $M$ would seem to suggest
+that many regions which impact trait differences do not appear as $F_{st}$ peaks. 
+This may imply that genome scans for local adaptation based purely on measures of differentiation may not be as reliable as previously thought.
+To summarize, our study find that with high rates of dispersal between population, $F_{st}$ peaks may be a good indication of causal loci.
+However, this is a one way implication, meaning even with high dispersal, not all causal loci will appear as $F_{st}$ peaks.
 
-\plr{Need to fix caption on figure \ref{fig:Power_FP} before editing this bit.}
-\jgg{There is still something wrong with this figure besides the labels ... }
+It's important to note that our study in Figures~\ref{fig:Fst} \& \ref{fig:Power_FP} pooled together all lakes and all marine individuals for the calculations. 
+For comparison, Figures~\ref{fig:fst_5lakes} \& \ref{fig:Power_FP_5lakes} shows the same scans but calculated 
+only on between individuals in the marine with spacial location $\le 5.0$ and the individuals in the respective 5 lakes.
+These graphs show even less power and True Positives only at $ M = 10.0$ migrants per lake, per generation
 
-In regions of the genome underlying individual trait value, we observed 
-Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$. 
-Knowing exactly which mutations effect phenotype in our simulations, we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome.
-In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power. 
-This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.
+%In regions of the genome underlying individual trait value, we observed 
+%Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$. 
+%Knowing exactly which mutations effect phenotype in our simulations, we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome.
+%In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power. 
+%This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.
 
@@
 	\begin{center}
-  		\includegraphics{Final_Plots/True_Power.pdf}
+  		\includegraphics{Final_Plots/True_Power_0_25_500.pdf}
   		\caption{ 
@@
   		\label{fig:Power_FP}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics{Final_Plots/True_Power_0_5_500.pdf}
+  		\caption{ 
+		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
+		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
+		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
+            \textbf{
+                TODO: fix this caption; x-axis label ($F_{ST}$).}
+		}
+  		\label{fig:Power_FP_5lakes}
 	\end{center}
@@
   		\label{fig:phenotype_time_supp}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+  		\includegraphics[width=1.0\linewidth]{Final_Plots/Fst_Genome_faa_0_5_500.pdf}
+  		\caption{
+		Average $F_{ST}$ in windows of 500Mb between:
+        		\textbf{(top)} marine habitat and old lakes;
+        		\textbf{(middle)} marine habitat and new lakes; and
+        		\textbf{(bottom)} old and new lakes.
+        		Each plot shows $F_{ST}$ values for a separate simulation,
+        		with columns corresponding to increasing gene flow from left to right.
+		$F_{st}$ is calculated between all marine individuals with spatial location $\le 5.0$, and the five corresponding lakes.
+		All locations of pre-existing freshwater adapted alleles have been highlighted by 
+		pink dashed vertical lines with an alpha value of 0.3, so darker shades of blue imply more 
+		freshwater adapted alleles at that location.
+		}
+  		\label{fig:fst_5lakes}
 	\end{center}

### 2018-12-12 Stickleback_Paper.tex
revisions to origin of introduced f alleles

@@
 on the degree to which alleles can be shared between populations.
-However, there remains a mystery.
-Figure~\ref{fig:MPFAI}A shows that with $m = 0.1$ migrants per lake per generation, 
+However, the large increase in speed of adaptation between dispersal rates of $M = 0.1 \& M = 1.0$, seen in Figure \ref{fig:TimeToAdaptation}, remains a mystery. 
+Figure~\ref{fig:MPFAI}A shows that with $M = 0.1$ migrants per lake per generation, 
 there is substantial sharing of alleles between populations,
-and yet Figure~\ref{fig:TimeToAdaptation} shows that new lakes 
-take around 2,000 generations to adapt.
+and yet at this rate new lakes take around 2,000 generations to adapt.
 This is surprising because we see a similar quantity of allele sharing at  $m = 1.0$, 
-but much more rapid adaptation ($\approx 30$ generations). 
-Perhaps this 60$\times$ discrepancy occurs because 
+but much more rapid adaptation ($\approx 60$ generations). 
+A reasonable explanation for this 33$\times$ discrepancy is that 
 new lakes must wait for an additional influx of freshwater alleles through migration
-beyond what is present in the initial generation?
+beyond what is present in the initial generation.
 This hypothesis is disproven by
@@
 freshwater alleles present as standing variation in the marine habitat
-are more tightly linked to marine alleles,
-and so adaptation in the new lakes must wait for recombination.
-This might be expected since freshwater alleles present in the ocean,
-to remain present at migration-selection balance at low migration rates,
+are more tightly linked to marine alleles.
+This implies the adaptation in the new lakes must wait for recombination to break freshwater adapted alleles from nearby marine adapted alleles.
+This might be expected at low dispersal rates since freshwater alleles present in the ocean
+%to remain present at migration-selection balance at low migration rates,
 must be ``masked'' by nearby compensatory alleles.
+Without masking their affect on phenotype, they would likely be pushed out by drift.
 Reversing this ``masking'' then slows the process of adaptation that uses this genetic variation.
-Said another way,
-higher gene flow from freshwater to the ocean
-maintains relatively intact freshwater haplotypes
+In other words, higher dispersal from freshwater to the ocean maintains relatively intact freshwater haplotypes
 that can be more easily rebuilt in the marine environments.
-Recall that trait-affecting mutations only occur in relatively small regions of $10^5$ loci
-in which recombination occurs at rate $10^{-2}$.
-Even if there is a sufficient amount of variation in the initial population of a lake
+Recall from our description in methods, that trait-affecting mutations only occur in relatively small regions of $10^5$ loci
+in which recombination occurs in these regions only once in every one thousand meioses.
+This implies that even if there is a sufficient amount of variation in the initial population of a lake
 to shift the trait from the marine optimum ($+10$) to the freshwater optimum ($-10$),
-these may be tightly linked to alleles that counteract their effect.
+rebuilding the most beneficial haplotype may prove to be a non-trivial task for selection.
+%these may be tightly linked to alleles that counteract their effect.
 For example, suppose there are 10 variants segregating at low frequency with effect size $-1$ each,
 but each is paired with a compensatory allele with effect size $+1$.
-Each local haplotype is therefore neutral
-(and so likely to be at higher frequency in the marine population).
+Each local haplotype is therefore neutral.
+This would also explain why at low rates of dispersal 
+the marine still has the capacity to hold a large amount maladaptive alleles within the individuals of its population, 
+as seen in \ref{fig:MPFAI}A.
+%(and so likely to be at higher frequency in the marine population).
 
-Indeed, this appears to be the case.
+%Indeed, this appears to be the case.
 To quantify the genetic variation available \emph{without} recombining within the ten genomic regions,
@@
 we get the maximum amount that selection could move the population in the freshwater direction
-without recombining within regions.
-The mean of this value across the 25 populations is shown 
-in the first row (``best'') of Table~\ref{tab:linkage} --
-typical populations at $m=0.1$ migrants per lake per generation could shift to a phenotype of -12.02
+without recombining \emph{within} selectively non-neutral regions.
+The mean of this value across the 25 lake populations is shown 
+in the (``best'') row of Table~\ref{tab:linkage} --
+typical populations at $M=0.1$ migrants per lake per generation could shift to a phenotype of -12.02
 (and so have sufficient variation to adapt without recombination),
-but the ``best'' haplotypes populations at $m=1.0$ have effect sizes nearly twice as big 
-(a mean total of -21.02).
+but the ``best'' haplotypes in the populations at $M=1.0$ have effect sizes nearly twice as big at a mean total of -21.02.
+Confusingly, at $M=10.0$ the best haplotype then raises to an unimpressive -7.02.
+However, this can simply be explained by the migration load the populations are experiencing, explained earlier and seen in Figure~\ref{fig:MeanPhenotype}
 

### 2018-12-13 Stickleback_Paper.tex
edits through intro

@@
 The results support existing theory of local adaptation, and provide a more concrete look at a particular, empirically motivated example.  
+\plr{say transporter hypothesis?}
+\plr{talk about linkage and haplotypes}
 
@@
 In addition, we would like to know how complex geographic systems, with a variety of selective pressures, interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations.
+\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
 
@@
 
-These empirical data generally support the ``transporter''-hypothesis proposed by Conte and Schluter in 2009 \citet{schluter2009genetics}, 
-which is a conceptual model for the flow of freshwater alleles from multiple smaller freshwater populations into much larger and less structured marine populations through hybridization events. 
-Alleles in this asymmetrically structured meta-population can then be recycled for subsequent freshwater adaptation. 
-Several questions remain, however, about the manner and degree to which these alleles are scattered among the marine fish allowing the haplotype to be re-assembled. 
-Is it more akin to the atomization and rebuilding that occurred in the Star Trek series that motivated this hypothesis? 
-Alternatively does it occur in more of a patchwork by individuals, 
-or their early generation hybrids in geographically adjacent freshwater habitats moving quickly through the marine environment to seed new freshwater populations. 
-More generally, how do different levels of gene flow and local adaptation interact with the asymmetric nature of stickleback population structure affect the dynamics of the origin and rate of adaptive alleles,
-as well as their eventual organization into marine and freshwater typical genomes?
-
-Here we develop a forward simulation approach in SLiM to model the stickleback evolutionary history in marine and freshwater habitats.
-We then record the effect of variation in gene flow on the genetic and genomic architecture of local adaptation.
-We ask, how rapidly can selection act on standing genetic variants at a given value of migration, $M$?
-Furthermore, how might the inference of causal loci through the use of genome wide association studies (GWAS) and $F_{st}$ scans across the genome be skewed by (lack of) introgression between sub-populations.
-Finally, we trace back through the genealogical history of adaptive variants in new lakes to find the precise origin is of those variants.
+\citet{schluter2009genetics} formulated the ``transporter'' hypothesis to explain these observations,
+a conceptual model in which alleles benefical in freshwater populations
+are maintained at migration-selection balance in the larger oceanic population
+and thus available for subsequent adaptation into new freshwater habitats.
+The hypothesis seems fits available data,
+but a number of quantitative questions remain.
+Are the actual population sizes, migration rates, and fitness differentials
+consistent with this hypothesis?
+How many of these differentially selected alleles are there,
+and how are they arranged within the genome?
+Does adaptation to a new lake rely on presence of early-generation hybrids between freshwater and marine fish,
+i.e., on long haplotypes recently inherited from freshwater-dwelling stickleback?
+Or, is the process
+more akin to the atomization and reconstruction effected by the Star Trek technology that motivated the hypothesis' name?
+More generally, 
+how do the dynamics and eventual genomic organization
+depend on the level of gene flow and strength of local adaptation,
+both across the entire stickleback metapopulation
+and separately within each lake?
+
+In this paper, we used forward simulations 
+modeled on the threespine stickleback metapopulation
+to ask how gene flow affects the genetic and genomic architecture of local adaptation.
+We know in practice that stickleback populations in new habitats can adapt in only tens of generations \citep{middleton};
+what level of between-habitat gene flow is required for such rapid adaptation?
+Genome scans for between-habitat differentiation have detected a number of causal QTL;
+what proportion of the loci actually underlying adaptation do we expect to be detected by such scans?
+How does our power to detect causal loci depend on the amount of gene flow?
 
 We find that only a few stickleback per generation, per lake, 
-suffice to maintain freshwater alleles from a (global?) haplotype at an acceptable frequency in marine environments for efficient ``transportation" of the freshwater haplotype.
-The selection on SGV at this level results in rapid local adaptation in $\approx 60$ generations, this follows closely to what we observe in nature (such as middleton island, cite this).
-Surprisingly, we find that the the continued gene flow (subsequent migration) of marine individuals is less important than the initial amount of SGV in the initial generation marine individuals at time of colonization.
-This suggests downstream gene flow from freshwater populations is key to this system of ``transportation".
-We show that low levels of introgression also results in noise that often presents itself as $F_{st}$ peaks when scanning across the genome suggesting introgression plays a key role in inferring causal loci .
-With higher $M > 10$ individuals per lake per generation we see the impact of migration load, preventing the demes from fully adapting to their selective pressure.
+suffices to maintain sufficient genetic variation in the ocean
+so that new freshwater populations can adapt from standing variation in only tens of generations.
+We also find that continued gene flow of marine individuals into recently established lakes
+is not important for adaptation,
+and so it is only ``downstream'' gene flow that allows freshwater adaptations to be ``transported'' between lakes.
+% We show that low levels of introgression also results in noise that often presents itself as $F_{st}$ peaks when scanning across the genome suggesting introgression plays a key role in inferring causal loci .
+This rapid, local adaptation can only occur over an intermediate range of levels of gene flow --
+low gene flow prevents variation from being transported,
+while high gene flow prevents local adaptation.
 
@@
 
-We explored these questions using forwards-time simulations with explicit genomic representation of individuals selecting upon a single continuous quantitative trait using SLiM \citep{haller2017slim,haller2018slim3}.
-The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography.
-We set up a realistic model of the relationship between marine and freshwater selective pressures with continuous space and migration rates between the populations to emulate gene flow. 
-Certain aspects of the the model remain simplistic due to computational constraints.
-Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback populations observed in nature
+To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} 
+to implement forwards-time simulation of populations of individuals with explicitly represented genomes
+in which selection acted upon a single continuous quantitative trait.
+The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography,
+and included divergent selection in the two habitats which had substantial spatial structure.
+Certain aspects of the model remain simplistic due to computational constraints.
+Possibly the most important caveat is that simulated population sizes 
+are much smaller than the census size of the threespine stickleback populations observed in nature
 (see the Discussion for more on this).

### 2018-12-13 Stickleback_Paper.tex
theoretical expectation

--- 2018-12-12 Stickleback_Paper.tex
@@
 in the light of population genetics theory.
-The discussion will be rough,
+The discussion will be roughly based on estimates,
 with a focus on intuition rather than precise calculations.
-Since within each population we model stabilizing selection on an additive trait
-controlled by a moderately large number of loci,
+Because we model stabilizing selection on an additive trait
+controlled by a moderately large number of loci within each population,
 more precise expectations might be obtained through quantitative genetics \citep{svardal2014general}
@@
 and so the rate of establishment of dominant mutations is $\beta z / 87.5$, 
-i.e., about one such mutation every $2461/z$ generations. 
+i.e., about one such mutation every $39375/z$ generations. 
+A the begging of the simulation where z = 10, we would then expect the fixation of alleles to take $\approx 5000$ generations.
+It follows that when looking at Figure~\ref{fig:phenotype_time}A, we can see the establishment of alleles in jumps roughly this size.
 The distribution of these successfully established mutations 
@@
 with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
-This agrees roughly with what we see in Figures~\ref{fig:phenotype_ts2} and XXX.
-\plr{refer to additional trace plots in the supp}
+This agrees roughly with what we see in Figures~\ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp}.
 
 \paragraph{Standing variation}
-An allele that moves the trait $z$ units in the freshwater direction in heterozygotes 
-has fitness roughly $\exp(-\beta z^2) \approx 1 - \beta z^2$ in the marine environment. 
-The product of population size and fitness differential in the marine environment for a mutation with $z=1$ is therefore $2Ns = 8.9$, 
+An allele that moves the trait $u$ units in the freshwater direction, in heterozygotes, 
+has fitness roughly $\exp(-\beta u^2) \approx 1 - \beta u^2$ in the marine environment, i.e a fitness differential of $\beta u^2$. 
+The product of population size and fitness differential in the marine environment for a mutation with $u=1$ is therefore $2Ns = 8.9$, 
 implying that these alleles are strongly selected against but might occasionally drift to moderate frequency. 
 The average frequency of such an allele in the marine environment at migration-selection equilibrium 
-is equal to the total influx of alleles into the ocean per generation divided by the selective disadvantage, 
-which if there are $M$ immigrants per generation, is $2 M / \beta z^2$. 
-\plr{This $M$ is the total number in the ocean! Make sure this differs from notation above for the number per lake.}
+is equal to the total influx of alleles into the ocean per generation divided by the selective disadvantage, i.e.
+if there are $I$ immigrants per generation, the frequency is $2 I / \beta u^2$. 
+%\plr{This $M$ is the total number in the ocean! Make sure this differs from notation above for the number per lake.}
 Each new lake is likely to contain a few copies of alleles at frequency above 1/400
-(since each new lake is begun with 400 genomes).
-With $z=1$, the factor multiplying $M$ is $2/\beta z^2 \approx 1/200$,
-so if $M \ge 1$, 
+(since each new lake is initialized with 400 randomly selected genomes).
+With $u=1$, the factor multiplying $I$ is $2/\beta u^2 \approx 1/200$,
+so if $I \ge 1$, 
 the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes 
 will appear at least once in the fish that colonize a new lake. 
-However, an allele with $z=1$ only has probability of around 1/20 of establishing locally, 
+However, an allele with $u=1$ only has probability of around 1/20 of establishing locally, 
 suggesting that the migration rate should be somewhat higher to ensure enough pre-existing genetic variation 
@@
 Systems of metapopulations like this could potentially arise in many places
-where the ancestral species carries around with it beneficial alleles forselective pressures it has encountered in the past and remains connected to through hybridization events.
+where the ancestral species carries around with it beneficial alleles for selective pressures it has encountered in the past and remains connected to through hybridization events.
 

### 2018-12-13 Stickleback_Paper.tex
peter edits until disco

@@
 \paragraph{Habitat and geography}
-Throughout the simulation there are two habitat types -- Marine and Freshwater-- defined by their selective pressures.
-We start the simulations with two populations of size 5,000 diploid individuals in each of the habitats, for a total of 10,000 diploid individuals.
-The arrangement of these habitats, depicted in Figure \ref{fig:Geo},
+Simulations have two habitat types -- Marine and Freshwater -- defined by their selective pressures,
+each with 5,000 diploid individuals.
+The arrangement of these habitats, depicted in Figure~\ref{fig:Geo},
 roughly models a set of freshwater habitats along a stretch of coastline. 
@@
 with different optima in marine and freshwater habitats.
-This situation roughly models the cumulative effect of the various phenotypes such as armor morphology, 
-body size, craniofacial variation and opercle shape on which divergent selection is thought to act on in the two environments. 
-Concretely, the optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively,
+This situation roughly models the cumulative effect of the various phenotypes 
+thought to be under divergent selection between the habitats,
+such as armor morphology, body size, craniofacial variation and opercle shape.
+The optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively,
 and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$
@@
 \end{align*}
-
-In short, the difference between an individual's trait value and the optimum largely determines that individual's fitness. 
-Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
+In short, the difference between an individual's trait value and the optimum determines that individual's fitness. 
+% Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
 We chose the difference between optima and strength of stabilizing selection in each habitat
@@
             a single, continuous, one-dimensional marine habitat (blue)
-            is coupled to randomly mating ``lakes'' at discrete locations with dark green arrows representing migration patterns.
+            is coupled to randomly mating ``lakes'' at discrete locations with arrows representing migration patterns.
             After an initial period of 100K generations with 25 lakes,
-            an additional 25 lakes are added (at the same set of locations) and populated with a copy of the \textit{marine} individuals
+            an additional 25 lakes are added (at the same set of locations) and populated with marine individuals
             to simulate the appearance of newly accessible freshwater habitats colonized by marine stickleback.
@@
 Each individual carries two linear chromosomes, each of size $10^8$ loci.
-Mutations that can affect the trait under selection can occur at rate $10^{-7}$ per locus per generation
-in ten ``effect regions'' of $10^5$ loci each,
+Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation
+in ten regions of $10^5$ loci each,
 spread evenly along the chromosome.
@@
 Subsequent mutations at the same locus replace the previous allele.
+We also used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence}
+to add neutral mutations across the entire genome at a rate of $10^{-7}$ per locus per generation, 
+as described in \citet{kelleher2018efficient}.
 
@@
 Since the total population across \emph{all} 25 lakes is fixed at 5000,
+and the Wright--Fisher model assumes unrealistic global population regulation,
 we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation.
@@
 
-As shown in Figure~\ref{fig:Geo} as lines and arrows, dispersal occurs both locally along the coastline in the marine habitat
-as well as between the marine habitat and the lakes.
+As depicted in Figure~\ref{fig:Geo}, dispersal occurs both locally along the coastline in the marine habitat
+as well as between the marine habitat and the lakes,
+with a lake--ocean migration rate denoted $m$.
 All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
-The lake--ocean migration rate is denoted $m$, 
-which we refer to as the rate of \emph{gene flow} between habitats.
 Each new individual in each habitat 
@@
 Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, 
-and reflected to stay within the population.
+and reflected to stay within the habitat range.
 % what about migrants?
@@
 
-\paragraph{New freshwater populations} 
-To study how newly appearing freshwater habitats adapt and select on the standing genetic variation in the marine, we introduce a new set of 25 lakes midway through the simulation at 100K generations.
-These new lakes are populated with a copy of the marine individuals to emulate a freshwater lake being colonized by its neighboring marine population. 
-In particular, after this point of introduction in the simulation there are:
-two \emph{sets} of lakes, and one marine population, each with 5,000 individuals for a total of 15,000 individuals being tracked in the simulation. 
+\paragraph{New lakes} 
+To study how marine-derived populations adapt after colonizing newly appearing freshwater habitats,
+we introduce a new set of 25 lakes after 100,000 generations.
+These new lakes are populated with marine individuals to emulate a freshwater lake being colonized by
+oceanic stickleback.
+This creates two \emph{sets} of lakes, which along with the marine population have a total of 15,000 individuals.
 Since this introduction of lakes doubles the number of lake-to-marine immigrants, 
 the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.
-
-\paragraph{Descriptive statistics}\bill{Population genetic analyses? Instead of the more general 'Descriptive statistics'?}
-To assess whether new lakes adapt using existing genetic diversity,
-we define a \emph{freshwater allele}
-to be an effect mutation that has frequency higher than $0.5$ in at least one of the original lakes,
-while remaining lower than $0.5$ in the marine. 
-This categorization is made for each generation using the allele frequencies from that generation,
-and so changes with time.
-Alleles common in the newly introduced lakes do not count
-if they are not also common in the original lakes.
-They are defined this way because the transportation hypothesis 
-does not specify where or when an advantageous mutation arises,
-but simply suggests that any sufficiently common freshwater adapted allele 
-could participate in adaptation in new habitat \citep{schluter2009genetics}.
-
-\emph{Time to adaptation} of the introduced population, denoted $T_\text{adapt}$, 
-is defined to be the generation at which 
-the difference between the average trait value in the original and the introduced freshwater populations 
-is less than 0.5.
-
-We describe overall genetic differentiation between the habitats using $F_{ST}$,
-calculated on a per-locus basis.
-Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele
-in the freshwater and marine habitats, respectively,
-and $\bar p = (p_f + p_m)/2$,
-then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
 
@@
 This allowed us to directly query the true origins of adaptive alleles.
-In addition it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations
-which were retroactively added to the gene trees at a rate of $10^{-7}$ per locus per generation, as described in \citet{kelleher2018efficient}).
+In addition, it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations
+which were retroactively added to the gene trees at a rate of $10^{-7}$ per locus per generation, as described in \citet{kelleher2018efficient}.
 
@@
     \item \emph{a ``Migrant" allele:} 
-        deriving from a migrant that was not in the initial generation that colonized the lake
+        deriving from a migrant not in the initial generation that colonized the lake
     \item \emph{a ``Captured" allele:}
-        present in the individual that initially colonized the new lake, 
+        present in initial colonists of the new lake, 
         and both common (above 50\%) in the original lakes,
@@
     \item \emph{a ``Marine" allele:} 
-        present in the individual that initially colonized the new lake, 
-        and not a captured allele.
+        present in initial colonists of the new lake, 
+        but not a ``captured'' allele.
 \end{enumerate}
-
-We then looked at the ratio of alleles from each origin to the total number of alleles being used as the basis for local adaptation in the new lakes.
-This essentially gives a measure on whether selection in the new environments made use of
+The proportion of trait-affecting alleles in new lakes that fall in these categories
+measures the degree to which selection in the new environments made use of
 (1) new mutation, 
@@
 (4) standing variation at mutation--selection balance.
-In other words, we were able to quantify the contributions of these possible origins 
-of adaptive alleles to understand where the genomic basis of adaptation had derived from.
-
-We were also able to use the tree sequence to get information about the individual ``effect regions''
-in all initial genomes of the new lakes at time of introduction.
-From this we determined the ability for the new lakes to select upon the standing variation in the marine.
-Given the probability that effect region fixes at a given trait value, $2 * x / 45$, where $x$ is the effect size of a haplotype,
-we calculated the expected total effect sizes of the haplotypes that fix.
-Concretely, it is the sum across the 10 effect regions of
-$$\sum_{i=1}^N \prod_{j < i} (1 - p_j) p_i x_i$$
-where $N$ is the number of genomes per lake, $p_i$ is the probability of fixation of the 
-$i^{th}$ haplotype, $x_{i}$ is the effect size of the $i^{th}$ haplotype, and the haplotypes are sorted in decreasing order by $x_{i}$.
-These numbers were computed assuming additivity of mutations, meaning the numbers produced are a slight overestimate.
-
+
+We also used neutral mutations to calculate $F_{ST}$,
+calculated on a per-locus basis and then averaged in windows 
+as a measure of between-population relative differentiation.
+Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele
+in the freshwater and marine habitats, respectively,
+and $\bar p = (p_f + p_m)/2$,
+then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
+
+%%%%%%%%%%%%%%%%%%
 \section*{Results}
 
-To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
-Since each lake contains 200 individuals, we refer to these migration rates in terms of \emph{dispersal} ($M$) \jgg{($D?$)} ranging from $0.01$ to $100$ migrants per lake, per generation.
-Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
-At very low rates of gene flow, each new lake's population adapted completely separately from \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations).
-At very high rates of gene flow, local adaptation was constrained by the large influx of maladaptive alleles.
+To observe the impact of gene flow on selection in the new lakes, 
+we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
+Below, we often refer to these migration rates in terms of the number of migrants per lake, per generation, which we denote $M$.
+Since each lake contains 200 individuals, $M=200m$.
+Many aspects of adaptation changed substantially across this range, including the speed of adaptation, 
+degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
+At very low rates of gene flow, each new lake's population adapted almost completely independently throught \emph{de novo} mutation, 
+which took a very long time ($\approx 20,000$ generations).
+At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles.
 Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes.
-Surprisingly, only a few migrants per generation from the lakes to the marine were needed
-to see the effects of rapid local adaptation accelerated by selection on standing genetic variants in the new lakes -- despite being deleterious in the intervening marine habitat.
-
+Perhaps surprisingly, only a few migrants per generation from the lakes to the ocean were needed
+to maintain sufficient genetic variation in the ocean to accelerate adaptation in new lakes.
+
+%%%%%%%%%%%%
 \subsection*{Local Adaptation: differentiation with gene flow}
@@
                 across the course of two simulations, with migration rates of
-        		\textbf{(top)} $m=5 \times 10^{-5}$ and
-        		\textbf{(bottom)} $m=5 \times 10^{-4}$ per generation per individual
-        		(i.e., 0.01 and 0.1 migrants per lake per generation, respectively).
+                \textbf{(top)} $M=0.01$ and
+                \textbf{(bottom)} $M=0.1$ 
+                migrants per lake per generation, respectively.
                 Optimal trait values in the two habitats are at $\pm 10$.
@@
 
-As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred at migration rates of $m \le 10$.
-% with the exception of the highest gene flow at which half of each population was composed of migrants.
-%At lower migration rates (10 migrants per generation and below), populations adapted to local conditions.
+As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred in all simulations
+with the exception of the highest gene flow at which half of each population was composed of migrants.
 Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp}
-show freshwater and marine populations diverging until the trait means were close to the optimal values in each habitat. 
-The establishment of new alleles in the lakes is visible in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value.
-%which move the trait by an amount of order 1 every few hundred generations.
+show how mean trait values in freshwater and marine populations diverged over time,
+until the trait means were close to the optimal values in each habitat. 
+Establishment of new alleles in the lakes is visible 
+in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value
+which move the trait by an amount of order 1 every few hundred generations.
 %Trait variation within each population was small compared to the difference between populations.
-At the lowest rate of gene flow, $m = 0.01$, differences at $\approx 20$ commonly polymorphic sites 
-($\approx 10$ that shift the trait in each direction)
+At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites 
+(about 10 that shift the trait in each direction)
 were responsible for most of the adaptive differences between freshwater and marine habitats.
-In Figure \ref{fig:MPFAI}A, we can see that more gene flow simplified the genetic basis. %does this belong here?
-As expected, increasing migration rate decreased differentiation between habitats. 
-As seen in Figure~\ref{fig:Fst}, 
+As expected, increasing migration rate decreased differentiation between habitats:
+as seen in Figure~\ref{fig:Fst}, 
 $F_{ST}$ between marine and freshwater habitats 
@@
 
-\paragraph*{Speed of adaptation}
 Adaptation occurred much more quickly
@@
 average trait values in old and new lakes were within 0.5 of each other,
-shown in Figure \ref{fig:TimeToAdaptation} for different rates of gene flow.
+shown in Figure~\ref{fig:TimeToAdaptation} for different rates of gene flow.
 Adaptation of new lakes took over 18,000 generations at the lowest rate of $m = 0.01$,
@@
 new lakes managed to adapt in just under 60 generations.
-%This dramatic decrease in time to local adaptation is suggestive of just how quickly
-%selection can act on standing variation. 
-%It also suggests a rough threshold of gene flow required 
-%to efficiently select upon the alleles present in the marine individuals.
 
@@
 
+%%%%%%%%%%%%
 \subsection*{Sharing of freshwater adapted alleles}
@@
 This difference occurs because at low migration rates,
-all subpopulation must wait for novel mutations to arise in order to adapt.
-Because the marine is one continuous population with 25 times more individuals than any one lake, 
+each lake must wait for its own novel mutations to arise in order to adapt.
+Because the marine habitat is continuous, with 25 times more individuals than any one lake, 
 there is a much larger influx of new mutations to be selected upon. 
 At higher migration rates, greater mixing
-allows the initial lakes to share alleles instead of developing their own genetic basis of adaptation. 
-As a first indication of this, Figure \ref{fig:Fst} 
-shows that $F_{ST}$ between the ``original'' and ``introduced'' sets of lakes 
-at effect mutations decreased with migration rate.
-%We are interested in this shared genetic basis of adaptation because it is commonly thought that this is
-%a requirement for the efficient ``transportation" of the freshwater haplotype in to new lakes \citet{schluter2009genetics}
+allows the initial lakes to share alleles instead of developing their own genetic basis for adaptation.
 
@@
 we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation.
-To be considered in this category, a mutation must participate in the genetic basis of any freshwater adaptation.
-Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lakes and below 50\% in the marine habitat.
+To be considered in this category, 
+an allele must participate in the genetic basis of local adaptation for at least one of the original lakes.
+Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lake and below 50\% in the marine habitat.
 Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles across generations.
-At $m = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere:
-the new lakes acquire new adaptive alleles and therefor have little to none of these adaptive alleles.
-At $m = 0.1$, we again observe the original lakes adapting nearly independently from each other.
-However, in Figure \ref{fig:MPFAI}A we can see the average marine individual is shown to hold $\approx 2$ pre-existing freshwater adapted alleles. 
-This being the case, we can see new lakes adapt using a subset of this large repertoire of standing variation deriving from a multitude of independently derived haplotypes.
-As migration rate increases past this, %we can see migration selection balance keeps the average capacity for marine individuals to carry freshwater alleles at about 2.
-we can see in Figure \ref{fig:MPFAI}C that the total number of pre-existing freshwater adapted alleles declines.
-%while the number of alleles which make up the genetic basis of adaptation in freshwater individuals remains relatively similar.
-%Although the genetic basis seems to simplify slightly due to more , this 
-This suggests that the alleles move between populations by migration 
-before selection acts on new mutation for neighboring lakes.
+At $M = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere:
+new lakes independently acquire new adaptive alleles rather than pre-existing ones.
+At $M = 0.1$, we again observe the original lakes adapting nearly independently from each other,
+but now the new lakes adapt using pre-existing alleles present in the original set of lakes.
+Concurrently, the average marine individual carries $\approx 2$ pre-existing freshwater adapted alleles,
+standing variation which was nearly absent at $M=0.01$.
+As migration rate increases past this, 
+the total number of pre-existing freshwater adapted alleles declines (Figure \ref{fig:MPFAI}C).
 Interestingly, the frequency of these alleles in the ocean
@@
 Figure~\ref{fig:MPFAI}B shows the distribution, through time, of 
-the mean percentage of currently-defined freshwater adapted alleles 
+the mean \emph{percentage} of currently-defined freshwater adapted alleles 
 that each genome in each of the populations carries.
@@
 higher rates of migration allow
-adaptive alleles of higher effect to travel more efficiently through the population,
+adaptive alleles of larger effect to travel more efficiently through the population,
 even though they are deleterious in the ocean.
 
-The numbers of Figure~\ref{fig:MPFAI} suggest that
+The numbers in Figure~\ref{fig:MPFAI} suggest that
 the dramatic increase in speed of local adaptation we observed above occurs because
@@
 %initial generation to be able to rebuild 
-We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting
-allele, in each individual, common in any of the new lakes at time of adaptation
-as defined in the Methods.
-Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo"} mutation.
+We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele
+in each individual in the new lakes, as defined in the Methods.
+Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo''} mutation.
 As gene flow increases, a larger fraction of adaptive alleles 
@@
         See Methods for precise definitions of these categories.
-		}
-  		\label{fig:Origin}
+    } \label{fig:Origin}
     \end{center}
@@
 
-\subsection*{Realized genetic architecture}
+%%%%%%%%%%%%
+\subsection*{Scans for selection}
 
 Here, we take a closer look at the genomic architecture of local adaptation between the two habitats. 
-Do measures of local differentiation such as $F_{st}$ scans or Genome Wide Association Studies (GWAS) identify the causal loci? 
-Figure \ref{fig:Fst} shows plots along the genome of per-locus $F_{ST}$ values between the marine habitat and all freshwater habitats pooled together.
-Higher rates of dispersal showed more distinct $F_{ST}$ peaks over polymorphic loci 
-``Background'' levels of $F_{ST}$ increase as gene flow decreases,
+Can measures of local differentiation such as $F_{ST}$ be used to identify the causal loci? 
+Figure \ref{fig:Fst} shows plots along the genome of average per-locus $F_{ST}$ values 
+in 500bp windows between the marine habitat and all freshwater habitats pooled together.
+\plr{500bp windows, right?}
+Higher rates of dispersal showed more distinct $F_{ST}$ peaks over polymorphic loci,
+while ``background'' levels of $F_{ST}$ increase as gene flow decreases,
 swamping out this signal until the regions under selection are indistinguishable. 
@@
 that declared everything above a certain value to be a causal locus.
-What we observe most plainly in this graph is that $M = 1.0 \& M = 10.0$ migrants per lake, per generation
-are the only rates of dispersal which provide reliable true negatives when observing peaks
+\plr{need to say what true and false positive windows are}
+What we observe most plainly in this graph is that $M = 1.0$ and $M = 10.0$ migrants per lake, per generation
+are the only rates of dispersal which provide reliable true negatives when observing peaks.
 This means that a large percentage of peaks past an $F_{ST}$ cutoff of $0.05$ 
@@
 
-It's important to note that our study in Figures~\ref{fig:Fst} \& \ref{fig:Power_FP} pooled together all lakes and all marine individuals for the calculations. 
+It's important to note that our study in Figures~\ref{fig:Fst} and \ref{fig:Power_FP} pooled together all lakes and all marine individuals for the calculations. 
 For comparison, Figures~\ref{fig:fst_5lakes} \& \ref{fig:Power_FP_5lakes} shows the same scans but calculated 
@@
 on the degree to which alleles can be shared between populations.
-However, the large increase in speed of adaptation between dispersal rates of $M = 0.1 \& M = 1.0$, seen in Figure \ref{fig:TimeToAdaptation}, remains a mystery. 
+However, the large increase in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$, 
+seen in Figure \ref{fig:TimeToAdaptation}, remains mysterious.
 Figure~\ref{fig:MPFAI}A shows that with $M = 0.1$ migrants per lake per generation, 
@@
 and yet at this rate new lakes take around 2,000 generations to adapt.
-This is surprising because we see a similar quantity of allele sharing at  $m = 1.0$, 
+This is surprising because we see a similar quantity of allele sharing at  $M = 1.0$, 
 but much more rapid adaptation ($\approx 60$ generations).
-A reasonable explanation for this 33$\times$ discrepancy is that 
+A reasonable explanation for this 33-fold discrepancy is that 
 new lakes must wait for an additional influx of freshwater alleles through migration
 beyond what is present in the initial generation.
-This hypothesis is disproven by
-Figure~\ref{fig:Origin}, which shows that at both $m = 0.1$ and $m = 1.0$,
-the vast majority of alleles underlying the local adaptation
-derived from individuals in the original generation (``captured'').
+However, Figure~\ref{fig:Origin} shows this is not the case: at both $m = 0.1$ and $m = 1.0$,
+the vast majority of alleles responsible for local adaptation
+came from individuals in the original generation (``captured'').
 
@@
 are more tightly linked to marine alleles.
-This implies the adaptation in the new lakes must wait for recombination to break freshwater adapted alleles from nearby marine adapted alleles.
-This might be expected at low dispersal rates since freshwater alleles present in the ocean
-%to remain present at migration-selection balance at low migration rates,
-must be ``masked'' by nearby compensatory alleles.
-Without masking their affect on phenotype, they would likely be pushed out by drift.
+This implies the adaptation in the new lakes must wait for recombination 
+to separate freshwater adapted alleles from nearby marine adapted alleles.
+This might be expected at low dispersal rates since freshwater alleles 
+may need to be ``masked'' by nearby compensatory alleles
+to remain in the ocean for long periods of time.
 Reversing this ``masking'' then slows the process of adaptation that uses this genetic variation.
+\plr{cite \citet{yeaman2011genetic}?}
 In other words, higher dispersal from freshwater to the ocean maintains relatively intact freshwater haplotypes
 that can be more easily rebuilt in the marine environments.
-Recall from our description in methods, that trait-affecting mutations only occur in relatively small regions of $10^5$ loci
-in which recombination occurs in these regions only once in every one thousand meioses.
+Recall that trait-affecting mutations only occur in relatively small regions of $10^5$ loci
+in which recombination occurs only once in every one thousand meioses.
 This implies that even if there is a sufficient amount of variation in the initial population of a lake
@@
 Each local haplotype is therefore neutral.
-This would also explain why at low rates of dispersal 
-the marine still has the capacity to hold a large amount maladaptive alleles within the individuals of its population, 
-as seen in \ref{fig:MPFAI}A.
+This might also explain why at $M=0.1$,
+marine individuals still hold on average several alleles that shift the trait in the freshwater direction.
+(Figure \ref{fig:MPFAI}A).
 %(and so likely to be at higher frequency in the marine population).
@@
 %Indeed, this appears to be the case.
-To quantify the genetic variation available \emph{without} recombining within the ten genomic regions,
+To quantify the genetic variation available \emph{without} recombination within the ten genomic regions,
 we first found, within each population, the haplotype with the largest net negative effect
@@
 we get the maximum amount that selection could move the population in the freshwater direction
-without recombining \emph{within} selectively non-neutral regions.
+without recombining {within} these regions.
 The mean of this value across the 25 lake populations is shown 
-in the (``best'') row of Table~\ref{tab:linkage} --
+in Table~\ref{tab:linkage} --
 typical populations at $M=0.1$ migrants per lake per generation could shift to a phenotype of -12.02
@@
 but the ``best'' haplotypes in the populations at $M=1.0$ have effect sizes nearly twice as big at a mean total of -21.02.
-Confusingly, at $M=10.0$ the best haplotype then raises to an unimpressive -7.02.
-However, this can simply be explained by the migration load the populations are experiencing, explained earlier and seen in Figure~\ref{fig:MeanPhenotype}
-
+The amount of variation available at $M=10.0$ is lower (only -7.02),
+presumably because of migration load, while at $M=0.01$ almost no alleles with negative effect are present.
+Note that some of these haplotypes will likely be lost to drift
+-- indeed, if they did all fix, then populations at $M=0.1$ would adapt much more quickly.
+However, the comparison still provides a good comparison.
+
+%%%%%%%%%%%%
 \subsection*{Theoretical expectations}
@@
 (assuming Poisson reproduction, as we roughly have here).
-If the current population all differed from the optimum trait by $z$, 
+If the current population all differ from the optimum trait by $z$, 
 and the allele has effect size $-u$ in heterozygotes, 
@@
 The total rate of appearance of new mutations per lake is $\mu_L = 0.04$, 
-which are divided evenly in seven categories: neutral, and then additive, dominant, and recessive in either direction. 
-This implies that a new additive or dominant effect mutation appears once every 87.5 generations, on average. 
+which are divided evenly in six categories: additive, dominant, and recessive, in either direction. 
+This implies that a new additive or dominant effect mutation appears once every 75 generations, on average. 
 Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$, 
@@
 is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$, 
-and so the rate of establishment of dominant mutations is $\beta z / 87.5$, 
-i.e., about one such mutation every $39375/z$ generations. 
-A the begging of the simulation where z = 10, we would then expect the fixation of alleles to take $\approx 5000$ generations.
-It follows that when looking at Figure~\ref{fig:phenotype_time}A, we can see the establishment of alleles in jumps roughly this size.
-The distribution of these successfully established mutations 
+and so the rate of establishment of dominant mutations is $\beta z / 75$, 
+i.e., about one such mutation every $33750/z$ generations. 
+At the beginning of the simulation where $z = 10$, we would then expect the fixation of alleles to take around 3,000 generations.
+The distribution of the effect sizes of these successfully established mutations 
 has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2. 
@@
 During the initial phase of adaptation, the populations begin at around distance $z=10$ from the optimum. 
-Combining these facts, we expect adaptive alleles to appear through mutation within lakes at first on a time scale of 250 generations, 
+Combining these facts, we expect adaptive alleles to appear through mutation within lakes at first on a time scale of 3,000 generations, 
 with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
@@
 An allele that moves the trait $u$ units in the freshwater direction, in heterozygotes, 
-has fitness roughly $\exp(-\beta u^2) \approx 1 - \beta u^2$ in the marine environment, i.e a fitness differential of $\beta u^2$. 
-The product of population size and fitness differential in the marine environment for a mutation with $u=1$ is therefore $2Ns = 8.9$, 
-implying that these alleles are strongly selected against but might occasionally drift to moderate frequency. 
+has fitness roughly $\exp(-\beta u^2) \approx 1 - \beta u^2$ in the marine environment, i.e, a fitness differential of $s = \beta u^2$. 
+The product of population size and fitness differential in the marine environment for a mutation with $u=1$ is therefore $2Ns \approx 22$,
+implying that these alleles are strongly selected against but might drift to moderate frequency if recessive.
 The average frequency of such an allele in the marine environment at migration-selection equilibrium 
-is equal to the total influx of alleles into the ocean per generation divided by the selective disadvantage, i.e.
-if there are $I$ immigrants per generation, the frequency is $2 I / \beta u^2$. 
-%\plr{This $M$ is the total number in the ocean! Make sure this differs from notation above for the number per lake.}
+is equal to the proportion of individuals in the ocean replaced by migrants per generation divided by the selective disadvantage, i.e.,
+around $m / \beta u^2$. 
 Each new lake is likely to contain a few copies of alleles at frequency above 1/400
 (since each new lake is initialized with 400 randomly selected genomes).
-With $u=1$, the factor multiplying $I$ is $2/\beta u^2 \approx 1/200$,
-so if $I \ge 1$, 
+With $u=1$, this occurs if $m/\beta > 1/400$,
+and so since $1/\beta = 450$, if $m \ge 5 \times 10^{-6}$
 the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes 
@@
 However, an allele with $u=1$ only has probability of around 1/20 of establishing locally, 
-suggesting that the migration rate should be somewhat higher to ensure enough pre-existing genetic variation 
-that adaptation would happen entirely using the initial set of colonizers. 
+so an allele of this size must be present in about 20 copies to ensure adaptation.
+Putting these together, we expect migration-selection balance to maintain sufficient genetic variation
+for new lakes to adapt if $m \ge 10^{-4}$, which corresponds to $M \ge 0.4$,
+agreeing with our results.
 % This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.

### 2018-12-14 Stickleback_Paper.tex
more citations, added to disc

@@
 While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures.
-What role does genomic variation play in metapopulations with connection to an array of selective pressures?
+What role does genomic variation play in meta-populations with connection to an array of selective pressures?
 How could migration-selection balance and ecological speciation actually benefit species as a whole?
-A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split.
+A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split \citet{nelson2017ancient,Barrett2008,Schrider2017}.
 However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
@@
 Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in tens of generations, 
-and the adaptive alleles are found in freshwater populations that have been geographically isolated since the end of the last ice age $\approx 13,000$ years ago (Cite Kristin).
+and the adaptive alleles are found in freshwater populations that have been geographically isolated since the end of the last ice age $\approx 13,000$ years ago.
 For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island. 
@@
 depend on the level of gene flow and strength of local adaptation,
-both across the entire stickleback metapopulation
+both across the entire stickleback meta-population
 and separately within each lake?
@@
 In this paper, we used forward simulations 
-modeled on the threespine stickleback metapopulation
+modeled on the threespine stickleback meta-population
 to ask how gene flow affects the genetic and genomic architecture of local adaptation.
-We know in practice that stickleback populations in new habitats can adapt in only tens of generations \citep{middleton};
+We know in practice that stickleback populations in new habitats can adapt in only tens of generations \citep{lescak2015evolution};
 what level of between-habitat gene flow is required for such rapid adaptation?
@@
 \paragraph{Recording genealogical history}
-We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence}
+We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence,Kelleher2016}
 to output the genealogical history of all individuals
@@
 which took a very long time ($\approx 20,000$ generations).
-At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles.
+At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles \citet{bolnick2007natural}.
 Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes.
@@
 in 500bp windows between the marine habitat and all freshwater habitats pooled together.
-\plr{500bp windows, right?}
 Higher rates of dispersal showed more distinct $F_{ST}$ peaks over polymorphic loci,
@@
 
-In this paper, we have described a realistic simulation model of a coastal metapopulation to observe the effects of selection of genetic variation (SGV) across a wide range of gene flow rates.
+In this paper, we have described a realistic simulation model of a coastal meta-population to observe the effects of selection of genetic variation (SGV) across a wide range of gene flow rates.
 We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
@@
 Small patches of similar selective pressures can be found in many places such as high altitudes, underwater volcanic vents, or post forest fire environments. 
-%We are still finding new subspecies, that seem to have adapted to extreme environments presented in small patches connected to larger spanning biospheres where the ancestral species naturally inhabits.
-Recently, (Cite Nelson Ferretti et al), discovered seven new species of Hapalotremus (Theraphosinae), living at the extremely high altitudes distributed along the Andes and Yungas in western South America.
-With the ancestral species at much lower altitudes, beneficial high altitude alleles could potentially be carried to help in the colonization of neighboring mountain environments. 
-Systems of metapopulations like this could potentially arise in many places
+Recently,  \citep{Ferretti2018}, discovered seven new species of Hapalotremus (Theraphosinae), living at the extremely high altitudes distributed along the Andes and Yungas in western South America.
+With the ancestral species at much lower altitudes, one could image high altitude alleles that could potentially be carried to help in the colonization of neighboring mountain environments. 
+Systems of meta-populations like this could potentially arise in many places
 where the ancestral species carries around with it beneficial alleles for selective pressures it has encountered in the past and remains connected to through hybridization events.
-
-%(1) the subspeciesthat inhabits a different selective pressure, and ancestral species are not constrained by pre, or post-zygotic isolation to participate in hybridization events, and
-%(2) the ancestral species spans across a wide range. 
-%For brevity, I will call evolutionary systems which satisfy these conditions, 
-%The vast number of species which encou
 
@@
 Perhaps the most notable finding in this study is that large, ancestral populations have the capacity to maintain alleles which are potentially deleterious for an extended amount of time.  
-%This system seems analagous the marvel superhero \emph{Rogue}, who can steal superpowers or memories from another  for a limited amount of time as long as she has contact with the indivi
 This principle, deriving solely from the properties of sex and natural selection, could be applicable to almost any species experiencing the influence of divergent selection with introgression.
 And to explore further how this process is carried out exactly, many of the questions lie in understanding which properties of genomic architecture and demographics allow this transportation to be possible.
-%In terms of Star Trek's ``transporter" referred to by \citet{schluter2009genetics}, 
 his is largely analogous to exploring the specific properties of the atmosphere / space which allow for the heroes to be disassembled, then put back together.
-%While the perspective of our 
 In our simulations, the continued gene flow from freshwater populations meant that the alleles were maintained at a constant rate with fixed demographics. 
@@
 (2) What role does genomic architecture play in the ability for a haplotype to be rebuilt.
-
-\paragraph{Origins of genetic basis of local adaptation in new lakes}
-In Figure~\ref{fig:Origin} we found that the genetic basis for local adaptation, when selecting upon SGV, is propogated largely from the initial generation of inhabitants. 
-This contrasts the belief that subsequent migration is a necessity in newly established populations ability to rebuild a haplotype.
-This finding suggests that with high probability, a randomly chosen subset from ancestral species in the metapopulations like marine stickleback, 
-carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
 
@@
 mistaken for being causative as they often display the same amount of divergence.
+In general, our study may suggest that diversity scans may not prove as useful in causal loci inference as previously thought.
+It is important to note that these summary statics could prove to be very useful if used beyond just the peaks seen when looked at as a function of genomic position. 
+Many methods, such as advanced machine learning, have proved to be very good at taking in noisy data and parsing out important information \citet{Schrider2018}.
+
+\paragraph{Implications for the ``Transporter"-hypothesis}
+This study was in large, supportive of the ``Transporter"-hypothesis suggested by \citep{schluter2009genetics}.
+Provided there is enough dispersal, we found that repeated colonization and adaptation of freshwater stickleback was able to produce beneficial alleles which could then be 
+supported by the marine stickleback population through the offspring of hybridization events. 
+When new freshwater environments were colonized by the marine stickleback, selection was able to rebuild the freshwater haplotype. 
+Shown in Table~\ref{tab:linkage}, we found that a populations capacity to cary maladaptive alleles remains pretty constant over realistic ranges of dispersal, but low rates of gene flow may 
+"mask" the alleles through linkage which counteracts the fitness disadvantage. 
+This linkage prevents the ability for a sub-population to rapidly select on the beneficial SGV as recombination acts as a bottleneck to evolution in this situation. 
+Additionally, In Figure~\ref{fig:Origin} we found that the genetic basis for local adaptation, when selecting upon SGV, is propagated largely from the initial generation of inhabitants. 
+This contrasts the belief that subsequent migration is a necessity in newly established populations ability to rebuild a haplotype.
+In other words, this finding suggests that with high probability, a randomly chosen subset from ancestral species in the meta-populations like marine stickleback, 
+carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
 

### 2018-12-21 Stickleback_Paper.tex
latest edits

@@
 How could migration-selection balance and ecological speciation actually benefit species as a whole?
-A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split \citet{nelson2017ancient,Barrett2008,Schrider2017}.
+A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split \citet{nelson2017ancient,Schrider2017,Barrett2008}.
 However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
@@
 And to explore further how this process is carried out exactly, many of the questions lie in understanding which properties of genomic architecture and demographics allow this transportation to be possible.
-his is largely analogous to exploring the specific properties of the atmosphere / space which allow for the heroes to be disassembled, then put back together.
 In our simulations, the continued gene flow from freshwater populations meant that the alleles were maintained at a constant rate with fixed demographics. 

### 2019-01-26 Stickleback_Paper.tex
Fig update
Friday edits, finish on Saturday

@@
 
-Threespine stickleback fish provide a striking example of local adaptation despite recurrent gene flow.
-The species is distributed around the Northern hemisphere in both marine and freshwater habitats.
-It is thought that these numerous, smaller freshwater populations
-have been established ``de novo" from marine fish,
-and that a shared freshwater phenotype
-is often established using standing genetic variation.
-Here we use genealogical simulations to determine the levels of gene flow
-that best matches observed patterns of allele sharing among habitats in stickleback, 
-and more generally to better understand how gene flow and local adaptation in large metapopulations
-determine speed of adaptation and reuse of standing genetic variation.
-We find that rapid, repeated adaptation using a shared set of alleles maintained at low frequency by migration-selection balance occurs over a realistic range of intermediate rates of gene flow.
-Low gene flow leads to slow, independent adaptation of distinct habitats, whereas high gene flow leads to large migration load.
-We quantify how $F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow.
-In addition, we find the origin of many freshwater adapted alleles in the introduced lakes to be propagated from the original generation of marine individuals that had inhabited the lake.
-The results support existing theory of local adaptation, and provide a more concrete look at a particular, empirically motivated example.  
-\plr{say transporter hypothesis?}
-\plr{talk about linkage and haplotypes}
+Threespine stickleback provide a striking example of local adaptation to divergent habitats while still being contected by recurrent gene flow.
+These small fish occurin marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases smaller freshwater populations have been established ``de novo'' from marine colonists.
+Previous results show that a similar freshwater phenotype often evolves in independent populations soon after colonization using largely the same standing genetic variation, even in very distant geographic regions.
+Here we use genealogical simulations to determine the levels of gene flow that best matches observed patterns of allele sharing among habitats in stickleback, and to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and reuse of standing genetic variation.
+We find that rapid, repeated adaptation using a shared set of alleles that are maintained at low frequency by migration-selection balance in ocean populations occurs over a realistic range of intermediate rates of gene flow that match previous empirical population genomic studies in stickleback.
+Using these realistic simulations we discovered that low gene flow leads to slow, independent adaption to district habitats, whereas high gene flow leads to significant mutation load.
+A consequence is that population genomic scans for locally adapted alleles are more likely to succeed with higher rates of gene flow.
+Lastly, we find that the origin of many freshwater adapted alleles in the introduced lakes is the original generation of marine individuals that colonized the lake.
+Thuses simulations provide deeper context for existing studies of stickleback evolutionary genomics, and provide guidance for future empirical studies in this model.
+More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.
+
+%\plr{say transporter hypothesis?}
+%\plr{talk about linkage and haplotypes}
 
@@
 % Jared's general notes for formatting
-%
 % - ``citep'' for ``cite in parentheses''. (and, citations come before the punctuation)
@@
 
-The canonical model for the genetics of adaptation, first formulated by Fisher in the early part of the 20th century, involves the sequential fixation of new mutations. 
+The sequential fixation of new mutations has long been the canonical model for the genetics of adaptation (\textbf{cite}).
 While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures.
@@
 However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
-In addition, we would like to know how complex geographic systems, with a variety of selective pressures, interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations.
-\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
+In addition, we would like to know how complex geographic systems, with a variety of selective pressures, interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations
+
+%\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
 
@@
 The ancestral marine form of stickleback has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
-While geographic isolation prevents direct migration between a large majority of these patches of freshwater populations,
-They appear to exhibit very similar phenotypes, as well as many shared, adaptive alleles often found to be identical by decent (IBD). 
-Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in tens of generations, 
-and the adaptive alleles are found in freshwater populations that have been geographically isolated since the end of the last ice age $\approx 13,000$ years ago.
+While geographic isolation prevents direct migration between a large majority of these patches of freshwater populations, They appear to exhibit very similar phenotypes, as well as many shared, adaptive alleles often found to be identical by decent (IBD).
+Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in tens of generations, and the adaptive alleles are found in freshwater populations that have been geographically isolated since the end of the last ice age $\approx 13,000$ years ago.
 For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island.
-Quickly inhabited by the surrounding marine population of stickleback, 
-\citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be parallel to freshwater stickleback that have been separated for over thousands of years. 
-In these freshwater stickleback, the number of lateral plates are reduced and the opercle shapes shows
-the same expansion of the dorsal region and reduction of the ventral region as observed in the large majority of freshwater demes.
+Quickly inhabited by the surrounding marine population of stickleback, \citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be parallel to freshwater stickleback that have been separated for over thousands of years.
+In these freshwater stickleback, the number of lateral plates are reduced and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region as observed in the large majority of freshwater demes.
 

### 2019-04-24 Stickleback_Paper.tex
latest paper updates

@@
 Threespine stickleback provide a striking example of local adaptation to divergent habitats while still being contected by recurrent gene flow.
-These small fish occurin marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases smaller freshwater populations have been established ``de novo'' from marine colonists.
+These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases smaller freshwater populations have been established ``de novo'' from marine colonists.
 Previous results show that a similar freshwater phenotype often evolves in independent populations soon after colonization using largely the same standing genetic variation, even in very distant geographic regions.
-Here we use genealogical simulations to determine the levels of gene flow that best matches observed patterns of allele sharing among habitats in stickleback, and to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and reuse of standing genetic variation.
-We find that rapid, repeated adaptation using a shared set of alleles that are maintained at low frequency by migration-selection balance in ocean populations occurs over a realistic range of intermediate rates of gene flow that match previous empirical population genomic studies in stickleback.
-Using these realistic simulations we discovered that low gene flow leads to slow, independent adaption to district habitats, whereas high gene flow leads to significant mutation load.
-A consequence is that population genomic scans for locally adapted alleles are more likely to succeed with higher rates of gene flow.
-Lastly, we find that the origin of many freshwater adapted alleles in the introduced lakes is the original generation of marine individuals that colonized the lake.
-Thuses simulations provide deeper context for existing studies of stickleback evolutionary genomics, and provide guidance for future empirical studies in this model.
+Here we use genealogical simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback, and to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and re-use of standing genetic variation.
+In our simulations we find that repeated adaptation uses a shared set of alleles that are maintained at low frequency by migration-selection balance in ocean populations.
+This process occurs over a realistic range of intermediate rates of gene flow that match previous empirical population genomic studies in stickleback.
+Examining these simulations more deeply reveals that low gene flow leads to slow, independent adaption to district habitats, whereas high gene flow leads to significant mutation load but an increase probability of success for population genomic scans for locally adapted alleles.
+We also find that the original generation of marine individuals that colonized the lakes is most often the sole source of most freshwater adapted alleles.
+These simulations provide deeper context for existing studies of stickleback evolutionary genomics, and provide guidance for future empirical studies in this model.
 More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.
@@
 
-The sequential fixation of new mutations has long been the canonical model for the genetics of adaptation (\textbf{cite}).
-While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures.
-What role does genomic variation play in meta-populations with connection to an array of selective pressures?
-How could migration-selection balance and ecological speciation actually benefit species as a whole?
-A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split \citet{nelson2017ancient,Schrider2017,Barrett2008}.
-However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
-In addition, we would like to know how complex geographic systems, with a variety of selective pressures, interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations
+The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations (\textbf{cite}).
+While this model has been very useful heuristically for guiding studies in the field and lab (\textbf{cite}), it is now rightfully understood as incomplete for many species in nature that have more complicated population structures (\textbf{cite}).
+In particular, empirical studies have increasingly pointed to the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats and experiencing diverse selective pressures while still exhibiting significant gene flow (\textbf{cite}).
+A growing number of empirical studies have identified instances of convergent evolution using SGV \citep{nelson2017ancient,Schrider2017,Barrett2008}.
+However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during invasion and adaptation to new environments (\textbf{cite}).
+It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via SGV.
+While emprical studies in the field and lab are essential, the use of simulation models of specific systems can greatly advanced our understanding of the adaptation genetic dynamics using SGV.
 
 %\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
+%\wac{I did a little editing but Peter can still totally rewrite. Regardless it needs to be streamlined a fair bit}
 
-Recently, a large flood of data resulting from evolutionary systems of organisms have come from population genomic studies powered by advances in sequencing technologies.
-One such organism is the threespine stickleback.
-Many studies of species have exhibited a long standing evolutionary history of migration-selection balance between marine and freshwater populations across the globe.
-The rigidity of the stickleback's ability to frequently prosper in newly created freshwater environments has made this fish a good model for understanding the genetic basis of adaptive evolution.
-The ancestral marine form of stickleback has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
-While geographic isolation prevents direct migration between a large majority of these patches of freshwater populations, They appear to exhibit very similar phenotypes, as well as many shared, adaptive alleles often found to be identical by decent (IBD).
-Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in tens of generations, and the adaptive alleles are found in freshwater populations that have been geographically isolated since the end of the last ice age $\approx 13,000$ years ago.
-For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island.
-Quickly inhabited by the surrounding marine population of stickleback, \citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be parallel to freshwater stickleback that have been separated for over thousands of years.
-In these freshwater stickleback, the number of lateral plates are reduced and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region as observed in the large majority of freshwater demes.
+The threespine stickleback fish has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes (\textbf{cite}) and the much more extensive use of SGV than once thought (\textbf{cite}).
+The ancestral marine form of stickleback has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere (\textbf{cite}).
+While geographic isolation prevents direct migration among freshwater populations, they often evolve similar (but not exactly the same) phenotypes (\textbf{cite}).
+The most recent evolutionary genomic studies on stickleback document that while adaptation to freshwater can occur over decades (\textbf{cite}), the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years (\textbf{cite}).
+Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations (\textbf{cite}).
+For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska (\textbf{cite}).
+Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years (\textbf{cite}).
+Amazingly, nearly a quarter of the genome diverged significantly in this same time but - with very little change in the rest of the genome - in a pattern that mirrored previous studies of freshwater stickleback populations thousands of years old (\textbf{cite}).
 
-But how can evolution occur at such a rapid pace?
-Waiting for new mutations to arise in each lake would take much longer, and the genotypes being identical by state is even more improbable.
-This would seem to suggest that freshwater alleles are maintained in the marine individuals allowing the accelerated selection on SGV found in marine individuals which colonize the lake.
-The first clear example of the global reuse of SGV was the gene \textit{eda} which has been shown to be an important regulator for the number of lateral plates.
-While the low lateral plate version of this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently.
-Contemporary population genomic studies employing genome-wide haplotype analyses has provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
+But how can evolution occur at such a rapid pace in this and other species?
+Waiting for new mutations to arise in each lake or pond would take much longer than the decades since the Alaskan earthquake (\textbf{cite}).
+Even more improbable is having divergence cluster in such similar genomic regions across independent populations.
+An alternative hypothesis, supported by the most recent population genomic analyses, is that the marjority of alleles important for freshwater adaptation are maintained in the marine individuals due to recurrent gene flow from freshwater back in to marine populations.
+In 2009 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter hypothesis'' to describe the process of alleles benefical in freshwater populations are maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation into new freshwater habitats.
+The first clear example of the global reuse of SGV in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates (\textbf{cite}).
+While the low lateral plate version of this gene arose millions of years ago, it is found in much younger freshwater ponds around the Northern Hemisphere (\textbf{cite}).
+More recently, genome-wide haplotype analyses have provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
+These data support the transporter hypothesis, but a number of questions remain.
+What are the actual population sizes, migration rates, and fitness differentials consistent with this hypothesis?
+How many differentially selected alleles exist, how many are used at any one time, and how are they arranged within the genome?
 
-\citet{schluter2009genetics} formulated the ``transporter'' hypothesis to explain these observations,
-a conceptual model in which alleles benefical in freshwater populations
-are maintained at migration-selection balance in the larger oceanic population
-and thus available for subsequent adaptation into new freshwater habitats.
-The hypothesis seems fits available data,
-but a number of quantitative questions remain.
-Are the actual population sizes, migration rates, and fitness differentials
-consistent with this hypothesis?
-How many of these differentially selected alleles are there,
-and how are they arranged within the genome?
-Does adaptation to a new lake rely on presence of early-generation hybrids between freshwater and marine fish,
-i.e., on long haplotypes recently inherited from freshwater-dwelling stickleback?
-Or, is the process
-more akin to the atomization and reconstruction effected by the Star Trek technology that motivated the hypothesis' name?
-More generally, 
-how do the dynamics and eventual genomic organization
-depend on the level of gene flow and strength of local adaptation,
-both across the entire stickleback meta-population
-and separately within each lake?
-
-In this paper, we used forward simulations 
-modeled on the threespine stickleback meta-population
-to ask how gene flow affects the genetic and genomic architecture of local adaptation.
-We know in practice that stickleback populations in new habitats can adapt in only tens of generations \citep{lescak2015evolution};
-what level of between-habitat gene flow is required for such rapid adaptation?
-Genome scans for between-habitat differentiation have detected a number of causal QTL;
-what proportion of the loci actually underlying adaptation do we expect to be detected by such scans?
-How does our power to detect causal loci depend on the amount of gene flow?
-
-We find that only a few stickleback per generation, per lake, 
-suffices to maintain sufficient genetic variation in the ocean
-so that new freshwater populations can adapt from standing variation in only tens of generations.
-We also find that continued gene flow of marine individuals into recently established lakes
-is not important for adaptation,
-and so it is only ``downstream'' gene flow that allows freshwater adaptations to be ``transported'' between lakes.
-% We show that low levels of introgression also results in noise that often presents itself as $F_{st}$ peaks when scanning across the genome suggesting introgression plays a key role in inferring causal loci .
-This rapid, local adaptation can only occur over an intermediate range of levels of gene flow --
-low gene flow prevents variation from being transported,
-while high gene flow prevents local adaptation.
+Here we use forward simulations incorporating linkage to model the stickleback metapopulation and address these questions.
+We ask how variation in amount of gene flow affects the genetic architecture of local adaptation to newly created freshwater ponds.
+Because we are documenting the entire genealogy of all alleles, we can determine the distribution and abundance of haplotypes across all in silico marine and freshwater populations to know the timing and proportion of potentially adaptive alleles that are actually utilized in each population.
+In addition, we can document how these adaptive alleles are distributed across the genomes, and as a consequence determine how this may affect the efficacy of genome scans for between-habitat differentiation.
+We find that only a few stickleback migrants per lake per generation are sufficient to maintain SGV in the ocean, allowing new freshwater populations can adapt in only tens of generations.
+We also find that continued gene flow of marine individuals into recently established lakes is not necessary for adaptation as the majority of alleles are present in the initial marine colonizers.
+Finally, we find that this rapid, local adaptation can only occur over an intermediate range of levels of gene flow.
+Low gene flow prevents variation from being transported, while high gene flow prevents local adaptation.
+Together these simulation findings enhance and extend the existing rich empirical work on stickleback evolutionary genomics.
 
 %%%%%%%%%%%%%%%%
+
 \section*{Methods}

### 2019-04-24 Stickleback_Paper.tex
updated file and added Thom's MS

@@
 
-\title{A few stickleback suffice to transport adaptive alleles to new lakes}
+\title{A few stickleback can transport adaptive alleles to new lakes}
 \author{Jared Galloway, William A. Cresko, and Peter Ralph}
@@
 Previous results show that a similar freshwater phenotype often evolves in independent populations soon after colonization using largely the same standing genetic variation, even in very distant geographic regions.
+But how many stickleback migrants are needed to capture a sufficient amount of standing variation to fuel adaptation?
 Here we use genealogical simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback, and to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and re-use of standing genetic variation.
 In our simulations we find that repeated adaptation uses a shared set of alleles that are maintained at low frequency by migration-selection balance in ocean populations.
-This process occurs over a realistic range of intermediate rates of gene flow that match previous empirical population genomic studies in stickleback.
-Examining these simulations more deeply reveals that low gene flow leads to slow, independent adaption to district habitats, whereas high gene flow leads to significant mutation load but an increase probability of success for population genomic scans for locally adapted alleles.
+This process occurs over a realistic range of intermediate levels of gene flow that match previous empirical population genomic studies in stickleback.
+Examining these simulations more deeply reveals that low gene flow leads to slow, independent adaption to district habitats, whereas high gene flow leads to significant mutation load but an increased probability of successful population genomic scans for locally adapted alleles.
 We also find that the original generation of marine individuals that colonized the lakes is most often the sole source of most freshwater adapted alleles.
@@
 More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.
-
-%\plr{say transporter hypothesis?}
-%\plr{talk about linkage and haplotypes}
 
@@
 The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations (\textbf{cite}).
-While this model has been very useful heuristically for guiding studies in the field and lab (\textbf{cite}), it is now rightfully understood as incomplete for many species in nature that have more complicated population structures (\textbf{cite}).
+While it has been very useful heuristically for guiding studies in the field and lab (\textbf{cite}), this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures (\textbf{cite}).
 In particular, empirical studies have increasingly pointed to the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats and experiencing diverse selective pressures while still exhibiting significant gene flow (\textbf{cite}).
@@
 The ancestral marine form of stickleback has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere (\textbf{cite}).
-While geographic isolation prevents direct migration among freshwater populations, they often evolve similar (but not exactly the same) phenotypes (\textbf{cite}).
+While geographic isolation often prevents direct migration among freshwater populations, they frequently evolve similar (but not exactly the same) phenotypes (\textbf{cite}).
 The most recent evolutionary genomic studies on stickleback document that while adaptation to freshwater can occur over decades (\textbf{cite}), the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years (\textbf{cite}).
@@
 More recently, genome-wide haplotype analyses have provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
-These data support the transporter hypothesis, but a number of questions remain.
+
+While this growing body of population genomic data on stickleback evolution support the transporter hypothesis, a number of questions remain.
 What are the actual population sizes, migration rates, and fitness differentials consistent with this hypothesis?
 How many differentially selected alleles exist, how many are used at any one time, and how are they arranged within the genome?
-
-Here we use forward simulations incorporating linkage to model the stickleback metapopulation and address these questions.
+A curious natural history observation undergirding many of these questions is the fact that some newly formed freshwater habitats, such as the ponds on Middleton Island, are quite small and presumably the number of initial marine migrants is few.
+Here we use forward simulations implementing in SLiM and incorporating linkage to model the stickleback metapopulation and address these questions.
 We ask how variation in amount of gene flow affects the genetic architecture of local adaptation to newly created freshwater ponds.
@@
 
-To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} 
-to implement forwards-time simulation of populations of individuals with explicitly represented genomes
-in which selection acted upon a single continuous quantitative trait.
-The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography,
-and included divergent selection in the two habitats which had substantial spatial structure.
+To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait.
+The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography, and included divergent selection in the two habitats which had substantial spatial structure.
 Certain aspects of the model remain simplistic due to computational constraints.
-Possibly the most important caveat is that simulated population sizes 
-are much smaller than the census size of the threespine stickleback populations observed in nature
-(see the Discussion for more on this).
+Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback populations observed in nature (see the Discussion for more on this).
 
 \paragraph{Habitat and geography}
-Simulations have two habitat types -- Marine and Freshwater -- defined by their selective pressures,
-each with 5,000 diploid individuals.
-The arrangement of these habitats, depicted in Figure~\ref{fig:Geo},
-roughly models a set of freshwater habitats along a stretch of coastline. 
-The marine habitat is a continuous, one-dimensional range of length 25 units,
-while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''),
-each connected to the marine habitat at regularly spaced intervals
-(positions $i - 1/2$ for $1 \le i \le 25$).
 
-Divergent selection is mediated by a single quantitative trait
-with different optima in marine and freshwater habitats.
-This situation roughly models the cumulative effect of the various phenotypes 
-thought to be under divergent selection between the habitats,
-such as armor morphology, body size, craniofacial variation and opercle shape.
-The optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively,
-and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$
-is determined by a Gaussian kernel with standard deviation 15, i.e.,
+Simulations have two habitat types -- Marine and Freshwater -- defined by their selective pressures, each with 5,000 diploid individuals.
+The arrangement of these habitats, depicted in Figure~\ref{fig:Geo} roughly models a set of freshwater habitats along a stretch of coastline.
+The marine habitat is a continuous, one-dimensional range of length 25 units, while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''), each connected to the marine habitat at regularly spaced intervals (positions $i - 1/2$ for $1 \le i \le 25$).
+
+Divergent selection is mediated by a single quantitative trait with different optima in marine and freshwater habitats.
+This situation roughly models the cumulative effect of the various phenotypes thought to be under divergent selection between the habitats, such as armor morphology, body size, craniofacial variation and opercle shape.
+The optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively, and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$ is determined by a Gaussian kernel with standard deviation 15, i.e.,
+
 \begin{align*}
@@
 \end{align*}
+
 In short, the difference between an individual's trait value and the optimum determines that individual's fitness.
+We chose the difference between optima and strength of stabilizing selection in each habitat so that (a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.
 % Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
-We chose the difference between optima and strength of stabilizing selection in each habitat
-so that 
-(a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,
-and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.
 
@@
 \paragraph{Genetic architecture of the trait}
+
 Each individual carries two linear chromosomes, each of size $10^8$ loci.
-Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation
-in ten regions of $10^5$ loci each,
-spread evenly along the chromosome.
+Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation in ten regions of $10^5$ loci each, spread evenly along the chromosome.
 Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability).
@@
 Individual trait values ($x_\text{ind}$) are determined additively from the diploid genotypes.
-Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value 
-$x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, 
-where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$.
+Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value $x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$.
 Subsequent mutations at the same locus replace the previous allele.
-We also used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence}
-to add neutral mutations across the entire genome at a rate of $10^{-7}$ per locus per generation, 
-as described in \citet{kelleher2018efficient}.
+We also used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence} to add neutral mutations across the entire genome at a rate of $10^{-7}$ per locus per generation,  as described in \citet{kelleher2018efficient}.
 
 \paragraph{Population dynamics}
-We use SLiM to simulate a Wright--Fisher population with non-overlapping generations 
-and a fixed population size of 5,000 diploid individuals in each habitat.
-Each generation, the two parents of each new offspring are chosen proportional to their fitness 
-(all individuals are hermaphroditic),
-and the contributing genomes are produced by Poisson recombination 
-with an average of one crossover per chromosome per generation 
-($10^{-7}$ per locus per generation). 
-Since the total population across \emph{all} 25 lakes is fixed at 5000,
-and the Wright--Fisher model assumes unrealistic global population regulation,
-we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation.
-to do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, 
-so that the mean fitnesses of all lakes are equal before selection happens.
 
-As depicted in Figure~\ref{fig:Geo}, dispersal occurs both locally along the coastline in the marine habitat
-as well as between the marine habitat and the lakes,
-with a lake--ocean migration rate denoted $m$.
-All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
-Each new individual in each habitat 
-has parents from the other habitat with probability $m$ 
-(in which case we call it a ``migrant''),
-and parents from the same habitat with probability $1-m$.
+We use SLiM to simulate a Wright--Fisher population with non-overlapping generations and a fixed population size of 5,000 diploid individuals in each habitat.
+Each generation, the two parents of each new offspring are chosen proportional to their fitness (all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombinationwith an average of one crossover per chromosome per generation ($10^{-7}$ per locus per generation).
+Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation.
+To do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, so that the mean fitnesses of all lakes are equal before selection happens.
+
+As depicted in Figure~\ref{fig:Geo}, dispersal occurs both locally along the coastline in the marine habitat as well as between the marine habitat and the lakes,
+with a lake--ocean migration rate denoted $m$.All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
+Each new individual in each habitat has parents from the other habitat with probability $m$ (in which case we call it a ``migrant''),and parents from the same habitat with probability $1-m$.
 % how do we get nonmigrants?
-The first parent of each nonmigrant individual in the freshwater habitat
-is chosen from the freshwater habitat proportional to fitness, 
-and a mate is chosen from the same lake as the first, also proportional to fitness. 
+The first parent of each nonmigrant individual in the freshwater habitat is chosen from the freshwater habitat proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness.
 The offspring then lives in the same lake as the parents.
 Parents for each nonmigrant marine individual are chosen similarly:
-first, a single parent is chosen proportionally to fitness in the marine habitat,
-and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. 
-Concretely, if the first parent is marine individual $i$, 
-then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
-where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
-Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, 
-and reflected to stay within the habitat range.
+first, a single parent is chosen proportionally to fitness in the marine habitat, and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$.
+Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$, where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations.
+Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, and reflected to stay within the habitat range.
 % what about migrants?
-Parents for each freshwater migrant are chosen in the same way
-as for nonmigrant marine individuals,
-and are assigned to the lake nearest to the position of the first marine parent.
-Similarly, parents for each marine migrant are both chosen from the same lake as before,
-and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.
+Parents for each freshwater migrant are chosen in the same way as for nonmigrant marine individuals, and are assigned to the lake nearest to the position of the first marine parent.
+Similarly, parents for each marine migrant are both chosen from the same lake as before, and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.
 
 \paragraph{New lakes}
-To study how marine-derived populations adapt after colonizing newly appearing freshwater habitats,
-we introduce a new set of 25 lakes after 100,000 generations.
-These new lakes are populated with marine individuals to emulate a freshwater lake being colonized by
-oceanic stickleback.
+
+To study how marine-derived populations adapt after colonizing newly appearing freshwater habitats, we introduce a new set of 25 lakes after 100,000 generations.
+These new lakes are populated with marine individuals to emulate a freshwater lake being colonized by oceanic stickleback.
 This creates two \emph{sets} of lakes, which along with the marine population have a total of 15,000 individuals.
-Since this introduction of lakes doubles the number of lake-to-marine immigrants, 
-the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.
+Since this introduction of lakes doubles the number of lake-to-marine immigrants,the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.
 
 \paragraph{Recording genealogical history}
-We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence,Kelleher2016}
-to output the genealogical history of all individuals
-at the time of introduction of new lakes, at the time of adaptation, and at the end of the simulation.
+We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence,Kelleher2016} to output the genealogical history of all individuals at the time of introduction of new lakes, at the time of adaptation, and at the end of the simulation.
 This allowed us to directly query the true origins of adaptive alleles.
-In addition, it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations
-which were retroactively added to the gene trees at a rate of $10^{-7}$ per locus per generation, as described in \citet{kelleher2018efficient}.
+In addition, it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations which were retroactively added to the gene trees at a rate of $10^{-7}$ per locus per generation, as described in \citet{kelleher2018efficient}.
 
@@
 Using these trees we classified each adaptive allele, in each genome in the new lakes at the time of adaptation, into four categories:
+
 \begin{enumerate}
@@
 \end{enumerate}
-The proportion of trait-affecting alleles in new lakes that fall in these categories
-measures the degree to which selection in the new environments made use of
+
+The proportion of trait-affecting alleles in new lakes that fall in these categories measures the degree to which selection in the new environments made use of
 (1) new mutation, 
@@
 
-We also used neutral mutations to calculate $F_{ST}$,
-calculated on a per-locus basis and then averaged in windows 
-as a measure of between-population relative differentiation.
-Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele
-in the freshwater and marine habitats, respectively,
-and $\bar p = (p_f + p_m)/2$,
-then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
+We also used neutral mutations to calculate $F_{ST}$, calculated on a per-locus basis and then averaged in windows as a measure of between-population relative differentiation.
+Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively, and $\bar p = (p_f + p_m)/2$, then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
 
@@
 
-To observe the impact of gene flow on selection in the new lakes, 
-we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
+To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
 Below, we often refer to these migration rates in terms of the number of migrants per lake, per generation, which we denote $M$.
 Since each lake contains 200 individuals, $M=200m$.
-Many aspects of adaptation changed substantially across this range, including the speed of adaptation, 
-degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
-At very low rates of gene flow, each new lake's population adapted almost completely independently throught \emph{de novo} mutation, 
-which took a very long time ($\approx 20,000$ generations).
+Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
+At very low rates of gene flow, each new lake's population adapted almost completely independently throught \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations).
 At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles \citet{bolnick2007natural}.
 Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes.
-Perhaps surprisingly, only a few migrants per generation from the lakes to the ocean were needed
-to maintain sufficient genetic variation in the ocean to accelerate adaptation in new lakes.
+Perhaps surprisingly, only a few migrants per generation from the lakes to the ocean were needed to maintain sufficient genetic variation in the ocean to accelerate adaptation in new lakes.
 
@@
 
-As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred in all simulations
-with the exception of the highest gene flow at which half of each population was composed of migrants.
-Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp}
-show how mean trait values in freshwater and marine populations diverged over time,
-until the trait means were close to the optimal values in each habitat. 
-Establishment of new alleles in the lakes is visible 
-in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value
-which move the trait by an amount of order 1 every few hundred generations.
+As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred in all simulations with the exception of the highest gene flow at which half of each population was composed of migrants.
+Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat.
+Establishment of new alleles in the lakes is visible in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations.
 %Trait variation within each population was small compared to the difference between populations.
-At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites 
-(about 10 that shift the trait in each direction)
-were responsible for most of the adaptive differences between freshwater and marine habitats.
-As expected, increasing migration rate decreased differentiation between habitats:
-as seen in Figure~\ref{fig:Fst}, 
-$F_{ST}$ between marine and freshwater habitats 
-at neutral sites steadily declines as migration increases.
+At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites(about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats.
+As expected, increasing migration rate decreased differentiation between habitats: as seen in Figure~\ref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.
 
@@
 
-Adaptation occurred much more quickly
-at higher migration rates,
-both in the old and new sets of lakes.
-We measured this ``time to adaptation''
-as the number of generations until 
-average trait values in old and new lakes were within 0.5 of each other,
-shown in Figure~\ref{fig:TimeToAdaptation} for different rates of gene flow.
-Adaptation of new lakes took over 18,000 generations at the lowest rate of $m = 0.01$,
-while at $m = 1$ migrant per, lake per generation,
-new lakes managed to adapt in just under 60 generations.
+Adaptation occurred much more quickly at higher migration rates, both in the old and new sets of lakes.
+We measured this ``time to adaptation'' as the number of generations until average trait values in old and new lakes were within 0.5 of each other, shown in Figure~\ref{fig:TimeToAdaptation} for different rates of gene flow.
+Adaptation of new lakes took over 18,000 generations at the lowest rate of $m = 0.01$, while at $m = 1$ migrant per, lake per generation, new lakes managed to adapt in just under 60 generations.
 
@@
 
-At low migration rates, the \emph{initial} period of adaptation
-takes roughly 25 times longer for lakes than it does for the ocean.
-This difference occurs because at low migration rates,
-each lake must wait for its own novel mutations to arise in order to adapt.
-Because the marine habitat is continuous, with 25 times more individuals than any one lake, 
-there is a much larger influx of new mutations to be selected upon. 
-At higher migration rates, greater mixing
-allows the initial lakes to share alleles instead of developing their own genetic basis for adaptation.
+At low migration rates, the \emph{initial} period of adaptation takes roughly 25 times longer for lakes than it does for the ocean.
+This difference occurs because at low migration rates, each lake must wait for its own novel mutations to arise in order to adapt.
+Because the marine habitat is continuous, with 25 times more individuals than any one lake, there is a much larger influx of new mutations to be selected upon.
+At higher migration rates, greater mixing allows the initial lakes to share alleles instead of developing their own genetic basis for adaptation.
 
-To investigate in more depth how locally adaptive alleles found in the original lakes
-are shared between lakes, as well as how they spread to the new lakes,
-we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation.
-To be considered in this category, 
-an allele must participate in the genetic basis of local adaptation for at least one of the original lakes.
+To investigate in more depth how locally adaptive alleles found in the original lakes are shared between lakes, as well as how they spread to the new lakes, we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation.
+To be considered in this category, an allele must participate in the genetic basis of local adaptation for at least one of the original lakes.
 Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lake and below 50\% in the marine habitat.
 Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles across generations.
-At $M = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere:
-new lakes independently acquire new adaptive alleles rather than pre-existing ones.
-At $M = 0.1$, we again observe the original lakes adapting nearly independently from each other,
-but now the new lakes adapt using pre-existing alleles present in the original set of lakes.
-Concurrently, the average marine individual carries $\approx 2$ pre-existing freshwater adapted alleles,
-standing variation which was nearly absent at $M=0.01$.
-As migration rate increases past this, 
-the total number of pre-existing freshwater adapted alleles declines (Figure \ref{fig:MPFAI}C).
+At $M = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere: new lakes independently acquire new adaptive alleles rather than pre-existing ones.
+At $M = 0.1$, we again observe the original lakes adapting nearly independently from each other, but now the new lakes adapt using pre-existing alleles present in the original set of lakes.
+Concurrently, the average marine individual carries $\approx 2$ pre-existing freshwater adapted alleles, standing variation which was nearly absent at $M=0.01$.
+As migration rate increases past this, the total number of pre-existing freshwater adapted alleles declines (Figure \ref{fig:MPFAI}C).
 Interestingly, the frequency of these alleles in the ocean
@@
 
-Figure~\ref{fig:MPFAI}B shows the distribution, through time, of 
-the mean \emph{percentage} of currently-defined freshwater adapted alleles 
-that each genome in each of the populations carries.
-If all individuals across lakes carried the same set of alleles determining their trait value, 
-this would be 100\%. 
-At the lowest migration rate $m = 0.01$,
-each genome in the original lakes have almost exactly~$1/25^{th}$ 
-of the total number of pre-existing freshwater adapted alleles -- 
-this is because each one of the 25 lakes has adapted with a unique set of alleles. 
+Figure~\ref{fig:MPFAI}B shows the distribution, through time, of the mean \emph{percentage} of currently-defined freshwater adapted alleles that each genome in each of the populations carries.
+If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%.
+At the lowest migration rate $m = 0.01$, each genome in the original lakes have almost exactly~$1/25^{th}$ of the total number of pre-existing freshwater adapted alleles -- this is because each one of the 25 lakes has adapted with a unique set of alleles.
 Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
-Figure \ref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above,
-the average individual across the new lakes has nearly the same amount of 
-pre-existing freshwater adapted alleles as individuals across the old lakes.
-As expected, the genetic basis of the freshwater phenotype 
-seems to simplify as migration increases --
-higher rates of migration allow
-adaptive alleles of larger effect to travel more efficiently through the population,
-even though they are deleterious in the ocean.
+Figure \ref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above, the average individual across the new lakes has nearly the same amount of pre-existing freshwater adapted alleles as individuals across the old lakes.
+As expected, the genetic basis of the freshwater phenotype seems to simplify as migration increases --
+higher rates of migration allow adaptive alleles of larger effect to travel more efficiently through the population, even though they are deleterious in the ocean.
 
-The numbers in Figure~\ref{fig:MPFAI} suggest that
-the dramatic increase in speed of local adaptation we observed above occurs because
-higher gene flow between populations allows sharing of freshwater alleles between populations.
+The numbers in Figure~\ref{fig:MPFAI} suggest that the dramatic increase in speed of local adaptation we observed above occurs because higher gene flow between populations allows sharing of freshwater alleles between populations.
+
 %however, it is not enough for the marine population to hold bits and peices from independently formed haplotypes,
@@
 %initial generation to be able to rebuild
-We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele
-in each individual in the new lakes, as defined in the Methods.
+
+We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele in each individual in the new lakes, as defined in the Methods.
 Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo''} mutation.
-As gene flow increases, a larger fraction of adaptive alleles 
-derive from pre-existing variation in the marine population at the time of introduction.
-In other words, 
-greater mixing at higher migration rates allows lakes to share alleles
-instead of developing their own genetic basis of adaptation.
+As gene flow increases, a larger fraction of adaptive alleles derive from pre-existing variation in the marine population at the time of introduction.
+In other words, greater mixing at higher migration rates allows lakes to share alleles instead of developing their own genetic basis of adaptation.
 
@@
 
-While increased migration allows sharing of adaptive alleles between lakes, 
-at $m = 10.0$ the constant influx of alleles between the habitats creates substantial migration load. 
-The rate at which migration load becomes substantial,
-10 migrants per lake per generation,
-only replaces 5\% of each population each generation with migrants from the other habitat, 
-but this is sufficient to shift the mean trait values to nearly half their optimal values, 
-as seen in Figure \ref{fig:MeanPhenotype}.
+While increased migration allows sharing of adaptive alleles between lakes, at $m = 10.0$ the constant influx of alleles between the habitats creates substantial migration load.
+The rate at which migration load becomes substantial, 10 migrants per lake per generation, only replaces 5\% of each population each generation with migrants from the other habitat, but this is sufficient to shift the mean trait values to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
 
@@
 Can measures of local differentiation such as $F_{ST}$ be used to identify the causal loci?
-Figure \ref{fig:Fst} shows plots along the genome of average per-locus $F_{ST}$ values 
-in 500bp windows between the marine habitat and all freshwater habitats pooled together.
-Higher rates of dispersal showed more distinct $F_{ST}$ peaks over polymorphic loci,
-while ``background'' levels of $F_{ST}$ increase as gene flow decreases,
-swamping out this signal until the regions under selection are indistinguishable. 
-This is likely the contribution of two separate forces of natural selection: 
-first, stronger genetic drift with less migration leads to higher background $F_{ST}$, and 
-second, greater sharing of adaptive alleles providing a shared signal across populations.
+Figure \ref{fig:Fst} shows plots along the genome of average per-locus $F_{ST}$ values in 500bp windows between the marine habitat and all freshwater habitats pooled together.
+Higher rates of dispersal showed more distinct $F_{ST}$ peaks over polymorphic loci, while ``background'' levels of $F_{ST}$ increase as gene flow decreases, swamping out this signal until the regions under selection are indistinguishable.
+This is likely the contribution of two separate forces of natural selection: first, stronger genetic drift with less migration leads to higher background $F_{ST}$, and second, greater sharing of adaptive alleles providing a shared signal across populations.
 
-At first glance, this suggests that genome scans for local adaptation based purely on measures of differentiation 
-will only be successful given enough migration between habitats. 
+At first glance, this suggests that genome scans for local adaptation based purely on measures of differentiation will only be successful given enough migration between habitats.
 But how many of these peaks are actually underlying trait differences that form the basis for local adaptation?
 To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates
-that would be obtained by an $F_{ST}$ cutoff 
-that declared everything above a certain value to be a causal locus.
+that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.
+
 \plr{need to say what true and false positive windows are}
-What we observe most plainly in this graph is that $M = 1.0$ and $M = 10.0$ migrants per lake, per generation
-are the only rates of dispersal which provide reliable true negatives when observing peaks.
-This means that a large percentage of peaks past an $F_{ST}$ cutoff of $0.05$ 
-actually lie on top of regions which could impact trait differences. 
-Unfortunately, the low statistical power for all rates of $M$ would seem to suggest
-that many regions which impact trait differences do not appear as $F_{st}$ peaks. 
+
+What we observe most plainly in this graph is that $M = 1.0$ and $M = 10.0$ migrants per lake, per generation are the only rates of dispersal which provide reliable true negatives when observing peaks.
+This means that a large percentage of peaks past an $F_{ST}$ cutoff of $0.05$ actually lie on top of regions which could impact trait differences.
+Unfortunately, the low statistical power for all rates of $M$ would seem to suggest that many regions which impact trait differences do not appear as $F_{st}$ peaks.
 This may imply that genome scans for local adaptation based purely on measures of differentiation may not be as reliable as previously thought.
@@
 It's important to note that our study in Figures~\ref{fig:Fst} and \ref{fig:Power_FP} pooled together all lakes and all marine individuals for the calculations.
-For comparison, Figures~\ref{fig:fst_5lakes} \& \ref{fig:Power_FP_5lakes} shows the same scans but calculated 
-only on between individuals in the marine with spacial location $\le 5.0$ and the individuals in the respective 5 lakes.
+For comparison, Figures~\ref{fig:fst_5lakes} \& \ref{fig:Power_FP_5lakes} shows the same scans but calculated only on between individuals in the marine with spacial location $\le 5.0$ and the individuals in the respective 5 lakes.
 These graphs show even less power and True Positives only at $ M = 10.0$ migrants per lake, per generation
@@
 
-We have thus far found that the speed of adaptation depends strongly
-on the degree to which alleles can be shared between populations.
-However, the large increase in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$, 
-seen in Figure \ref{fig:TimeToAdaptation}, remains mysterious.
-Figure~\ref{fig:MPFAI}A shows that with $M = 0.1$ migrants per lake per generation, 
-there is substantial sharing of alleles between populations,
-and yet at this rate new lakes take around 2,000 generations to adapt.
-This is surprising because we see a similar quantity of allele sharing at  $M = 1.0$, 
-but much more rapid adaptation ($\approx 60$ generations).
-A reasonable explanation for this 33-fold discrepancy is that 
-new lakes must wait for an additional influx of freshwater alleles through migration
-beyond what is present in the initial generation.
-However, Figure~\ref{fig:Origin} shows this is not the case: at both $m = 0.1$ and $m = 1.0$,
-the vast majority of alleles responsible for local adaptation
-came from individuals in the original generation (``captured'').
+We have thus far found that the speed of adaptation depends strongly on the degree to which alleles can be shared between populations.
+However, the large increase in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$, seen in Figure \ref{fig:TimeToAdaptation}, remains mysterious.
+Figure~\ref{fig:MPFAI}A shows that with $M = 0.1$ migrants per lake per generation, there is substantial sharing of alleles between populations, and yet at this rate new lakes take around 2,000 generations to adapt.
+This is surprising because we see a similar quantity of allele sharing at  $M = 1.0$, but much more rapid adaptation ($\approx 60$ generations).
+A reasonable explanation for this 33-fold discrepancy is that new lakes must wait for an additional influx of freshwater alleles through migration beyond what is present in the initial generation.
+However, Figure~\ref{fig:Origin} shows this is not the case: at both $m = 0.1$ and $m = 1.0$, the vast majority of alleles responsible for local adaptation came from individuals in the original generation (``captured'').
 
@@
 
-An alternative explanation is that at lower rates of gene flow,
-freshwater alleles present as standing variation in the marine habitat
-are more tightly linked to marine alleles.
-This implies the adaptation in the new lakes must wait for recombination 
-to separate freshwater adapted alleles from nearby marine adapted alleles.
-This might be expected at low dispersal rates since freshwater alleles 
-may need to be ``masked'' by nearby compensatory alleles
-to remain in the ocean for long periods of time.
+An alternative explanation is that at lower rates of gene flow, freshwater alleles present as standing variation in the marine habitat are more tightly linked to marine alleles.
+This implies the adaptation in the new lakes must wait for recombination to separate freshwater adapted alleles from nearby marine adapted alleles.
+This might be expected at low dispersal rates since freshwater alleles may need to be ``masked'' by nearby compensatory alleles to remain in the ocean for long periods of time.
 Reversing this ``masking'' then slows the process of adaptation that uses this genetic variation.
+
 \plr{cite \citet{yeaman2011genetic}?}
-In other words, higher dispersal from freshwater to the ocean maintains relatively intact freshwater haplotypes
-that can be more easily rebuilt in the marine environments.
-Recall that trait-affecting mutations only occur in relatively small regions of $10^5$ loci
-in which recombination occurs only once in every one thousand meioses.
-This implies that even if there is a sufficient amount of variation in the initial population of a lake
-to shift the trait from the marine optimum ($+10$) to the freshwater optimum ($-10$),
-rebuilding the most beneficial haplotype may prove to be a non-trivial task for selection.
+
+In other words, higher dispersal from freshwater to the ocean maintains relatively intact freshwater haplotypes that can be more easily rebuilt in the marine environments.
+Recall that trait-affecting mutations only occur in relatively small regions of $10^5$ loci in which recombination occurs only once in every one thousand meioses.
+This implies that even if there is a sufficient amount of variation in the initial population of a lake to shift the trait from the marine optimum ($+10$) to the freshwater optimum ($-10$), rebuilding the most beneficial haplotype may prove to be a non-trivial task for selection.
 %these may be tightly linked to alleles that counteract their effect.
-For example, suppose there are 10 variants segregating at low frequency with effect size $-1$ each,
-but each is paired with a compensatory allele with effect size $+1$.
+For example, suppose there are 10 variants segregating at low frequency with effect size $-1$ each, but each is paired with a compensatory allele with effect size $+1$.
 Each local haplotype is therefore neutral.
-This might also explain why at $M=0.1$,
-marine individuals still hold on average several alleles that shift the trait in the freshwater direction.
+This might also explain why at $M=0.1$, marine individuals still hold on average several alleles that shift the trait in the freshwater direction.
 (Figure \ref{fig:MPFAI}A).
@@
 %Indeed, this appears to be the case.
-To quantify the genetic variation available \emph{without} recombination within the ten genomic regions,
-we first found, within each population, the haplotype with the largest net negative effect
-at each of the ten genomic regions.
-Summing these ten numbers, 
-we get the maximum amount that selection could move the population in the freshwater direction
-without recombining {within} these regions.
-The mean of this value across the 25 lake populations is shown 
-in Table~\ref{tab:linkage} --
-typical populations at $M=0.1$ migrants per lake per generation could shift to a phenotype of -12.02
-(and so have sufficient variation to adapt without recombination),
-but the ``best'' haplotypes in the populations at $M=1.0$ have effect sizes nearly twice as big at a mean total of -21.02.
-The amount of variation available at $M=10.0$ is lower (only -7.02),
-presumably because of migration load, while at $M=0.01$ almost no alleles with negative effect are present.
-Note that some of these haplotypes will likely be lost to drift
--- indeed, if they did all fix, then populations at $M=0.1$ would adapt much more quickly.
+To quantify the genetic variation available \emph{without} recombination within the ten genomic regions, we first found, within each population, the haplotype with the largest net negative effect at each of the ten genomic regions.
+Summing these ten numbers, we get the maximum amount that selection could move the population in the freshwater direction without recombining {within} these regions.
+The mean of this value across the 25 lake populations is shown in Table~\ref{tab:linkage} -- typical populations at $M=0.1$ migrants per lake per generation could shift to a phenotype of -12.02
+(and so have sufficient variation to adapt without recombination), but the ``best'' haplotypes in the populations at $M=1.0$ have effect sizes nearly twice as big at a mean total of -21.02.
+The amount of variation available at $M=10.0$ is lower (only -7.02), presumably because of migration load, while at $M=0.01$ almost no alleles with negative effect are present.
+Note that some of these haplotypes will likely be lost to drift -- indeed, if they did all fix, then populations at $M=0.1$ would adapt much more quickly.
 However, the comparison still provides a good comparison.
@@
 
-We now revisit our observations above
-in the light of population genetics theory.
-The discussion will be roughly based on estimates,
-with a focus on intuition rather than precise calculations.
-Because we model stabilizing selection on an additive trait
-controlled by a moderately large number of loci within each population,
-more precise expectations might be obtained through quantitative genetics \citep{svardal2014general}
-or even Fisher's geometric model \citep{barton2001hybridization,chevin2014niche}.
+We now revisit our observations above in the light of population genetics theory.
+The discussion will be roughly based on estimates, with a focus on intuition rather than precise calculations.
+Because we model stabilizing selection on an additive trait controlled by a moderately large number of loci within each population, more precise expectations might be obtained through quantitative genetics \citep{svardal2014general} or even Fisher's geometric model \citep{barton2001hybridization,chevin2014niche}.
 
 Suppose a new allele enters a lake, either by migration or mutation.
-If, when it is rare but present in $n$ copies, it has fitness advantage $s$ 
--- i.e., the expected number of copies in the next generation is $(1+s)n$ 
--- then the probability that it escapes demographic stochasticity to become common in the population 
-is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical}
-(assuming Poisson reproduction, as we roughly have here).
-If the current population all differ from the optimum trait by $z$, 
-and the allele has effect size $-u$ in heterozygotes, 
-then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$. 
+If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ -- then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical(assuming Poisson reproduction, as we roughly have here).
+If the current population all differ from the optimum trait by $z$, and the allele has effect size $-u$ in heterozygotes,  then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
 This tells us two things: (1) the rate of adaptation decreases as the population approaches the optimum, and (2) larger mutations (in the right direction) are more likely to fix.
@@
 \paragraph{New mutations}
-The total rate of appearance of new mutations per lake is $\mu_L = 0.04$, 
-which are divided evenly in six categories: additive, dominant, and recessive, in either direction. 
+
+The total rate of appearance of new mutations per lake is $\mu_L = 0.04$, which are divided evenly in six categories: additive, dominant, and recessive, in either direction.
 This implies that a new additive or dominant effect mutation appears once every 75 generations, on average.
 Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$,
-and so the probability that a dominant mutation manages to establish in a population differing from the optimum by $z$ 
-is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$, 
-and so the rate of establishment of dominant mutations is $\beta z / 75$, 
-i.e., about one such mutation every $33750/z$ generations. 
+and so the probability that a dominant mutation manages to establish in a population differing from the optimum by $z$ is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$, 
+and so the rate of establishment of dominant mutations is $\beta z / 75$, i.e., about one such mutation every $33750/z$ generations.
 At the beginning of the simulation where $z = 10$, we would then expect the fixation of alleles to take around 3,000 generations.
@@
 has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2.
-Since additive alleles have half the effect in heterozygotes, 
-they have half the probability of establishment. 
+Since additive alleles have half the effect in heterozygotes, they have half the probability of establishment.
 During the initial phase of adaptation, the populations begin at around distance $z=10$ from the optimum.
-Combining these facts, we expect adaptive alleles to appear through mutation within lakes at first on a time scale of 3,000 generations, 
-with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
+Combining these facts, we expect adaptive alleles to appear through mutation within lakes at first on a time scale of 3,000 generations, with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
 This agrees roughly with what we see in Figures~\ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp}.
@@
 \paragraph{Standing variation}
-An allele that moves the trait $u$ units in the freshwater direction, in heterozygotes, 
-has fitness roughly $\exp(-\beta u^2) \approx 1 - \beta u^2$ in the marine environment, i.e, a fitness differential of $s = \beta u^2$. 
-The product of population size and fitness differential in the marine environment for a mutation with $u=1$ is therefore $2Ns \approx 22$,
-implying that these alleles are strongly selected against but might drift to moderate frequency if recessive.
-The average frequency of such an allele in the marine environment at migration-selection equilibrium 
-is equal to the proportion of individuals in the ocean replaced by migrants per generation divided by the selective disadvantage, i.e.,
-around $m / \beta u^2$. 
-Each new lake is likely to contain a few copies of alleles at frequency above 1/400
-(since each new lake is initialized with 400 randomly selected genomes).
+An allele that moves the trait $u$ units in the freshwater direction, in heterozygotes, has fitness roughly $\exp(-\beta u^2) \approx 1 - \beta u^2$ in the marine environment, i.e, a fitness differential of $s = \beta u^2$.
+The product of population size and fitness differential in the marine environment for a mutation with $u=1$ is therefore $2Ns \approx 22$, implying that these alleles are strongly selected against but might drift to moderate frequency if recessive.
+The average frequency of such an allele in the marine environment at migration-selection equilibrium is equal to the proportion of individuals in the ocean replaced by migrants per generation divided by the selective disadvantage, i.e., around $m / \beta u^2$.
+Each new lake is likely to contain a few copies of alleles at frequency above 1/400 (since each new lake is initialized with 400 randomly selected genomes).
 With $u=1$, this occurs if $m/\beta > 1/400$,
-and so since $1/\beta = 450$, if $m \ge 5 \times 10^{-6}$
-the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes 
-will appear at least once in the fish that colonize a new lake. 
-However, an allele with $u=1$ only has probability of around 1/20 of establishing locally, 
-so an allele of this size must be present in about 20 copies to ensure adaptation.
-Putting these together, we expect migration-selection balance to maintain sufficient genetic variation
-for new lakes to adapt if $m \ge 10^{-4}$, which corresponds to $M \ge 0.4$,
+and so since $1/\beta = 450$, if $m \ge 5 \times 10^{-6}$ the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes will appear at least once in the fish that colonize a new lake.
+However, an allele with $u=1$ only has probability of around 1/20 of establishing locally, so an allele of this size must be present in about 20 copies to ensure adaptation.
+Putting these together, we expect migration-selection balance to maintain sufficient genetic variation for new lakes to adapt if $m \ge 10^{-4}$, which corresponds to $M \ge 0.4$,
 agreeing with our results.
 % This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.
-However, this calculation treats each allele independently;
-in practice we found that standing freshwater variation in the ocean
-were masked by linkage to compensatory marine variants.
+However, this calculation treats each allele independently; in practice we found that standing freshwater variation in the ocean were masked by linkage to compensatory marine variants.
 
 \paragraph{Migration}
-The key quantity regulating the amount of standing variation in the ocean was the \emph{downstream} migration rate,
-from lakes to the ocean.
-How important is the upstream migration rate?
-If sufficient genetic variation is not present in a new lake initially, 
-it must appear either by new mutation or by migration. 
+The key quantity regulating the amount of standing variation in the ocean was the \emph{downstream} migration rate, from lakes to the ocean. How important is the upstream migration rate?
+If sufficient genetic variation is not present in a new lake initially, it must appear either by new mutation or by migration.
 Since a proportion $m$ of each lake is composed of migrants each generation,
 it takes $1/m$ generations until the genetic variation introduced by migrants equals the amount initially present at colonization.
-This implies a dichotomy: either 
-(a) migration is high, and adaptation is possible using variants present at colonization or arriving shortly thereafter, or 
-(b) migration is low, so adaptation takes many multiples of $1/m$ generations.
-Since in our model lower migration also reduces the amount of variation available in the ocean,
-we expect very little contribution of subsequent migration across any value of $m$,
-as seen in Figure~\ref{fig:Origin}.
+This implies a dichotomy: either (a) migration is high, and adaptation is possible using variants present at colonization or arriving shortly thereafter, or (b) migration is low, so adaptation takes many multiples of $1/m$ generations.
+Since in our model lower migration also reduces the amount of variation available in the ocean, we expect very little contribution of subsequent migration across any value of $m$, as seen in Figure~\ref{fig:Origin}.
 

### 2019-05-08 Stickleback_Paper.tex
Removed unneccessary files & updated figured top have larger font size

@@
 
-Threespine stickleback provide a striking example of local adaptation to divergent habitats while still being contected by recurrent gene flow.
+Threespine stickleback provide a striking example of local adaptation to divergent habitats while still being connected by recurrent gene flow.
 These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases smaller freshwater populations have been established ``de novo'' from marine colonists.
@@
 In short, the difference between an individual's trait value and the optimum determines that individual's fitness.
-We chose the difference between optima and strength of stabilizing selection in each habitat so that (a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.
-% Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
+We chose the difference between optima and strength of stabilizing selection in each habitat so that (a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other, and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.
 
@@
     \begin{center}
-          \includegraphics[width=0.6\linewidth]{GeographyFigure.pdf}
+          \includegraphics[width=0.8\linewidth]{Final_Plots/GeographyFigure.pdf}
           \caption{
@@
 We use SLiM to simulate a Wright--Fisher population with non-overlapping generations and a fixed population size of 5,000 diploid individuals in each habitat.
-Each generation, the two parents of each new offspring are chosen proportional to their fitness (all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombinationwith an average of one crossover per chromosome per generation ($10^{-7}$ per locus per generation).
+Each generation, the two parents of each new offspring are chosen proportional to their fitness (all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-7}$ per locus per generation).
 Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation.
@@
 Each new individual in each habitat has parents from the other habitat with probability $m$ (in which case we call it a ``migrant''),and parents from the same habitat with probability $1-m$.
-% how do we get nonmigrants?
 The first parent of each nonmigrant individual in the freshwater habitat is chosen from the freshwater habitat proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness.
@@
 Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, and reflected to stay within the habitat range.
-% what about migrants?
 Parents for each freshwater migrant are chosen in the same way as for nonmigrant marine individuals, and are assigned to the lake nearest to the position of the first marine parent.
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome_faa_pos.pdf}
+        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome_faa.pdf}
           \caption{
@@
         All locations of pre-existing freshwater adapted alleles have been highlighted by 
-        pink dashed vertical lines with an alpha value of 0.3, so darker shades of blue imply more 
+        pink dashed vertical lines with an alpha value of 0.3, so darker shades of pink imply more 
         freshwater adapted alleles at that location.
@@
             The number of alleles meeting these conditions changes over the course of the simulation,
-            and each plot shows distributions of these values across generations.
-            The horizontal axis shows $M$, the mean number of migrants per lake per generation.
-        }
+            and each plot shows distributions of these values throughout each simulations.
+            The horizontal axis shows $M$, the mean number of migrants per lake per generation.}
         \label{fig:MPFAI}
@@
     \begin{center}
-    \includegraphics[width=0.7\linewidth]{Final_Plots/Allele_Origin_2.pdf}
+    \includegraphics{Final_Plots/Allele_Origin_2.pdf}
     \caption{ 
@@
 Suppose a new allele enters a lake, either by migration or mutation.
-If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ -- then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical(assuming Poisson reproduction, as we roughly have here).
+If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ -- then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical} (assuming Poisson reproduction, as we roughly have here).
 If the current population all differ from the optimum trait by $z$, and the allele has effect size $-u$ in heterozygotes,  then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
@@
 
-%\paragraph{Implications for the ``Transporter"--hypothesis}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\begin{comment}
 \begin{figure}
@@
     \begin{center}
-          \includegraphics[width=1.0\linewidth]{Final_Plots/Fst_Genome_faa_0_5_500.pdf}
+          \includegraphics[width=1.0\linewidth]{Final_Plots/Fst_Genome_faa0_5_500.pdf}
           \caption{

### 2019-05-08 matplotScripts/Adaptation_matplot.py

@@ -1,64 +1,0 @@
-
-
-#read in the arguments from each run across paramters values and plot all the time to adaptations.
-
-import sys
-#import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-
-num = sys.argv[1]
-
-x = [0.00005,0.0005,0.005,0.05]
-x2 = ['5e-5','5e-4','5e-3','5e-2']
-color_shape = ['go','b^','rO']
-
-timeToAdaptation = []
-numShared = []
-totalNumFWAA = []
-numShared_DivTotal = []
-corFWAA = []
-corEffect = []
-
-for j in sys.argv[1:]:
-
-	TTA = []
-	num = j
-	recipes = ['8_'+num+'_0','8_'+num+'_1','8_'+num+'_2','8_'+num+'_3']
-	
-	for i in recipes:
-
-		File = open("../../Output1/MyRecipe"+i+"/Adaptation.txt","r")
-		header = File.readline()
-		FWAA_P2 = File.readline()
-		FWAA_P3 = File.readline()
-
-		TTA.append(int(File.readline().split()[0]))
-		numShared_ = File.readline().split()[0]
-		totalNumFWAA_ = File.readline().split()[0]
-
-		numShared_DivTotal.append(float(numShared_ )/float(totalNumFWAA_))
-
-		corFWAA.append(File.readline().split()[0])
-		corEffect.append(File.readline().split()[0])
-
-	timeToAdaptation.append(TTA)
-
-timeToAdaptation = np.array(timeToAdaptation)	
-means = np.mean(timeToAdaptation,axis=0)
-
-f1, ax1 = plt.subplots()
-for k in range(len(timeToAdaptation)):
-	ax1.plot(x2,timeToAdaptation[k],color_shape[k])
-ax1.plot(x2,means,"y--")
-
-ax1.grid(axis = 'y')
-ax1.set_ylabel('Time (Generation)')
-ax1.set_xlabel('Migration Rate')
-ax1.set_title('Time to Adaptation')
-
-ax1.set_yscale('log')
-plt.show()
-
-
-

### 2019-05-08 matplotScripts/Fst_Dist_matplot.py

@@ -1,53 +1,0 @@
-import plotly
-from plotly.graph_objs import Scatter, Layout
-import plotly.graph_objs as go
-import plotly.plotly as py
-import sys
-
-num = sys.argv[1]
-recipes = ['8_'+num+'_0','8_'+num+'_1','8_'+num+'_2','8_'+num+'_3']
-
-def readHeader(firstLine):
-
-	numberOfSamplingPoints,samplingInterval,introduction,endofSim = (int(i) for i in firstLine.split())
-	samplesUntilIntroduction = introduction/samplingInterval - 1
-	return numberOfSamplingPoints,samplingInterval,introduction
-
-Original_Intro = []
-Marine_Original = []
-Marine_Intro = []
-
-data = [[][][]]
-
-	
-for i in range(4):
-
-	file1 = "../../Output1/MyRecipe"+recipes[i]+"/FreshwaterFreshwater2Fst.txt"
-	#,"Freshwater","IntroducedFreshwater",'rgb(60,179,113)'
-	file2 = "../../Output1/MyRecipe"+recipes[i]+"/OceanFreshwaterFst.txt"
-	#,"Ocean","Freshwater",'rgb(127,255,212)'
-	file3 = "../../Output1/MyRecipe"+recipes[i]+"/OceanFreshwater2Fst.txt"
-	#,"Ocean","IntroducedFreshwater",'rgb(138,43,226)'
-
-	
-
-	for k,f in enumerate([file1,file2,file3]):
-
-		File = open(f,"r")
-		first = File.readline()
-
-		#read in respective Fst/Pos pairs - they are in order.
-		Fst = list(map(float,File.readline().split()))
-		Pos = list(map(int,File.readline().split()))
-		
-		#sort both lists by Fst value
-		Pos_Sorted = [fst for _,fst in sorted(zip(Fst,Pos))]
-		Fst_Sorted = sorted(Fst)
-
-		data[k].append()
-		
-	
-
-
-
-

### 2019-05-08 matplotScripts/boxplot_Fst_matplot.py

@@ -1,87 +1,0 @@
-#read in one set of runs across parameter values and plot the Fst boxplots for neutral and effect mutations
-
-
-import sys
-import matplotlib.pyplot as plt
-import numpy as np
-
-num = sys.argv[1]
-
-recipes = ['8_'+num+'_0','8_'+num+'_1','8_'+num+'_2','8_'+num+'_3']
-ticks = ['5e-5','5e-4','5e-3','5e-2']
-titles = ['Neutral','Effect']
-
-data_a = []
-data_b = []
-data_c = []
-
-for l in range(2):
-
-	FstMarineOrig = []
-	FstMarineIntro = []
-	FstOrigIntro = []
-
-	for i in range(len(recipes)):
-		
-		MO1 = []
-		MI1 = []
-		OI1 = []
-
-		File = open("../../Output1/MyRecipe"+recipes[i]+"/MeanFstThroughout"+titles[l]+".txt","r")
-		first = File.readline()
-		count = 0
-		for line in File:
-			#cut out the first 50 data points to cut tails from boxplot
-			if(count < 50):
-				count+=1
-				continue
-			else:
-				n1,n2,n3 = (float(s) for s in line.split())
-				MO1.append(n1)
-				MI1.append(n2)
-				OI1.append(n3)
-		
-		FstMarineOrig.append(MO1)
-		FstMarineIntro.append(MI1)
-		FstOrigIntro.append(OI1)
-
-	data_a.append(FstMarineOrig)
-	data_b.append(FstMarineIntro)
-	data_c.append(FstOrigIntro)
-
-f1, ax1 = plt.subplots(1,2,sharey=True,sharex=True,linewidth=.01)
-
-def set_box_color(bp, color):
-    plt.setp(bp['boxes'], color=color)
-    plt.setp(bp['whiskers'], color=color)
-    plt.setp(bp['caps'], color=color)
-    plt.setp(bp['medians'], color=color)
-
-for k in range(2):
-
-	bpl = ax1[k].boxplot(data_a[k], positions=np.array(range(len(data_a[k])))*3.0-0.6, sym='', widths=0.4)
-	bpm = ax1[k].boxplot(data_b[k], positions=np.array(range(len(data_b[k])))*3.0, sym='', widths=0.4)
-	bpr = ax1[k].boxplot(data_c[k], positions=np.array(range(len(data_c[k])))*3.0+0.6, sym='', widths=0.4)
[…]

### 2019-05-08 matplotScripts/boxplot_MFAI_matplot.py

@@ -1,74 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import sys
-
-num = sys.argv[1]
-
-Traces = []
-
-recipes = ['8_'+num+'_0','8_'+num+'_1','8_'+num+'_2','8_'+num+'_3']
-mig_rates = ['mig = 5e-5 (1/10 Ind/Gen)', 'mig = .0005 (1 Ind/Gen)','mig = .005 (10 Ind/Gen)','mig = .05 (100 Ind/Gen)']
-
-xAxis = []
-AvgFreshAllelesPerMarineInd = []
-AvgFreshAllelesPerOrigInd = []
-AvgFreshAllelesPerIntroInd = []	
-
-for i in range(4):
-
-	M = []
-	O = []
-	I = []
-
-	File = open("../../Output1/MyRecipe"+recipes[i]+"/AvgFWAA.txt","r")
-	first = File.readline()
-	count = 0
-
-	for line in File:
-		n1,n2,n3 = (float(s) for s in line.split())
-		M.append(n1)
-		O.append(n2)
-		I.append(n3)
-
-	AvgFreshAllelesPerMarineInd.append(M)
-	AvgFreshAllelesPerOrigInd.append(O)
-	AvgFreshAllelesPerIntroInd.append(I)
-
-data_a = AvgFreshAllelesPerMarineInd
-data_b = AvgFreshAllelesPerOrigInd
-data_c = AvgFreshAllelesPerIntroInd
-
-ticks = ['5e-5','5e-4','5e-3','5e-2']
-
-def set_box_color(bp, color):
-    plt.setp(bp['boxes'], color=color)
-    plt.setp(bp['whiskers'], color=color)
-    plt.setp(bp['caps'], color=color)
-    plt.setp(bp['medians'], color=color)
-
-#plt.figure()
-fig1,ax1 = plt.subplots()
-bpl = ax1.boxplot(data_a, positions=np.array(range(len(data_a)))*3.0-0.6, sym='', widths=0.4)
-bpm = ax1.boxplot(data_b, positions=np.array(range(len(data_b)))*3.0, sym='', widths=0.4)
-bpr = ax1.boxplot(data_c, positions=np.array(range(len(data_c)))*3.0+0.6, sym='', widths=0.4)
-set_box_color(bpl, '#D7191C') # colors are from http://colorbrewer2.org/
-set_box_color(bpm, '#2C7BB6')
-set_box_color(bpr, '#5D7FF6')
-
[…]

### 2019-05-08 matplotScripts/boxplot_MPFAI_matplot.py

@@ -1,74 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import sys
-
-num = sys.argv[1]
-
-Traces = []
-
-recipes = ['8_'+num+'_0','8_'+num+'_1','8_'+num+'_2','8_'+num+'_3']
-mig_rates = ['mig = 5e-5 (1/10 Ind/Gen)', 'mig = .0005 (1 Ind/Gen)','mig = .005 (10 Ind/Gen)','mig = .05 (100 Ind/Gen)']
-
-xAxis = []
-AvgFreshAllelesPerMarineInd = []
-AvgFreshAllelesPerOrigInd = []
-AvgFreshAllelesPerIntroInd = []	
-
-for i in range(4):
-
-	M = []
-	O = []
-	I = []
-
-	File = open("../../Output1/MyRecipe"+recipes[i]+"/AvgFWAA_divTotal.txt","r")
-	first = File.readline()
-	count = 0
-
-	for line in File:
-		n1,n2,n3 = (float(s) for s in line.split())
-		M.append(n1)
-		O.append(n2)
-		I.append(n3)
-
-	AvgFreshAllelesPerMarineInd.append(M)
-	AvgFreshAllelesPerOrigInd.append(O)
-	AvgFreshAllelesPerIntroInd.append(I)
-
-data_a = AvgFreshAllelesPerMarineInd
-data_b = AvgFreshAllelesPerOrigInd
-data_c = AvgFreshAllelesPerIntroInd
-
-ticks = ['5e-5','5e-4','5e-3','5e-2']
-
-def set_box_color(bp, color):
-    plt.setp(bp['boxes'], color=color)
-    plt.setp(bp['whiskers'], color=color)
-    plt.setp(bp['caps'], color=color)
-    plt.setp(bp['medians'], color=color)
-
-#plt.figure()
-fig1,ax1 = plt.subplots()
-bpl = ax1.boxplot(data_a, positions=np.array(range(len(data_a)))*3.0-0.6, sym='', widths=0.4)
-bpm = ax1.boxplot(data_b, positions=np.array(range(len(data_b)))*3.0, sym='', widths=0.4)
-bpr = ax1.boxplot(data_c, positions=np.array(range(len(data_c)))*3.0+0.6, sym='', widths=0.4)
-set_box_color(bpl, '#D7191C') # colors are from http://colorbrewer2.org/
-set_box_color(bpm, '#2C7BB6')
-set_box_color(bpr, '#5D7FF6')
-
[…]

### 2019-05-08 matplotScripts/haplo_stitch_matplot.py

@@ -1,37 +1,0 @@
-'''
-import matplotlib.pyplot as plt
-import matplotlib.image as mpimg
-import numpy as np
-
-
-
-fig, axs = plt.subplots(3, 1)
-
-img1 = mpimg.imread("/Users/jaredgalloway/Desktop/HAPSREAL/5e_5p1.png")
-img2 = mpimg.imread("/Users/jaredgalloway/Desktop/HAPSREAL/5e_5p2.png")
-img3 = mpimg.imread("/Users/jaredgalloway/Desktop/HAPSREAL/5e_5p3.png")
-
-axs[0] = plt.imshow(img1)
-axs[1] = plt.imshow(img2)
-axs[2] = plt.imshow(img3)
-
-plt.show()
-'''
-
-
-
-import numpy as np
-import matplotlib.pyplot as plt
-
-w=10
-h=10
-fig=plt.figure(figsize=(8, 8))
-columns = 4
-rows = 5
-for i in range(1, columns*rows +1):
-    img = np.random.randint(10, size=(h,w))
-    fig.add_subplot(rows, columns, i)
-    plt.imshow(img)
-plt.show()
-
-

### 2019-05-08 r_calcs/plot_linkage.R

@@ -1,176 +1,0 @@
-library(matrixStats)
-library(xtable)
-
-everything <- lapply(0:3, function (run_num) {
-
-    nodes <- read.table(sprintf("nodes_positions_%d.txt", run_num))
-    colnames(nodes) <- c("id", "location")
-    nodes$pop <- cut(nodes$location, 25)
-    if (FALSE) plot(nodes$location, col=nodes$pop) # sanity check
-
-    # has one row per genome and one column per variant
-    genotypes <- matrix(scan(sprintf("genotype_matrix_%d.txt", run_num)), nrow=nrow(nodes))
-
-    # initializeMutationType("m1", 0.5, "g", 0.5, 1.0);  //Positive Direction, Dominant 
-    # initializeMutationType("m2", 0.5, "g", -0.5, 1.0); //Negative Direction, Dominant
-    # //initializeMutationType("m3", 0.5, "f", 0.0);       //Neither Direction, Neutral
-    # initializeMutationType("m4", 0.5, "g", 0.5, 1.0);  //Positive Direction, Recessive
-    # initializeMutationType("m5", 0.5, "g", -0.5, 1.0); //Negative Direction, Recessive
-    # initializeMutationType("m6", 0.5, "g", 0.6, 1.0);  //Positive Direction, Additive
-    # initializeMutationType("m7", 0.5, "g", -0.6, 1.0); //Negative Direction, Additive
-
-    variants <- read.table(sprintf("variants_%d.txt", run_num))
-    colnames(variants) <- c("id", "size", "type", "pos")
-    stopifnot(all(variants$type != 3))
-    stopifnot(ncol(genotypes) == nrow(variants))
-    variants$freq <- colMeans(genotypes)
-    variants$region <- cut(variants$pos, 10)
-    variants$direction <- ifelse(variants$type %in% c(1, 4, 6), +1, -1)
-    stopifnot(all(sign(variants$size) == variants$direction))
-    variants$dominance <- c(1, 1, NA, 0, 0, 0.5, 0.5)[variants$type]
-
-    # now we compute net effects by effect region and genome
-    effects <- genotypes * variants$size[col(genotypes)]
-    region_effects <- t(apply(effects, 1, tapply, variants$region, sum))
-    nodes$phenotype <- rowSums(region_effects)
-
-    frequencies <- do.call(rbind, 
-                           tapply(1:nrow(genotypes), nodes$pop, function (k) colMeans(genotypes[k,])))
[…]

### 2019-05-08 Scripts/freshwater_allele_distributions.py
added all scripts needed to run simulations and plot

@@ -1,0 +1,148 @@
+import matplotlib as mpl
+mpl.use('pdf')
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import numpy as np
+import sys
+
+plt.rc('font', family='serif', serif='Times')
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=10)
+plt.rc('ytick', labelsize=10)
+plt.rc('axes', labelsize=10)
+
+num = '1'
+
+Traces = []
+
+recipes = ['9_'+num+'_0','9_'+num+'_1','9_'+num+'_2','9_'+num+'_3','9_'+num+'_4']
+mig_rates = ["$0.01$","$0.1$","$1.0$","$10.0$","$100.0$"]
+
+def getData(percentage=False):
+
+    filename = "/AvgFWAA"
+    if(percentage):
+        filename += "_divTotal"
+    filename += ".txt"
+
+    AvgFreshAllelesPerMarineInd = []
+    AvgFreshAllelesPerOrigInd = []
+    AvgFreshAllelesPerIntroInd = []	
+
+    for i in range(5):
+
+        M = []
+        O = []
+        I = []
+
+        File = open("./Output1/MyRecipe"+recipes[i]+filename,"r")
+        first = File.readline()
+        count = 0
+
+        for line in File:
+            n1,n2,n3 = (float(s) for s in line.split())
+            M.append(n1)
+            O.append(n2)
+            I.append(n3)
+
+        AvgFreshAllelesPerMarineInd.append(M)
+        AvgFreshAllelesPerOrigInd.append(O)
+        AvgFreshAllelesPerIntroInd.append(I)
+
+    return AvgFreshAllelesPerMarineInd,AvgFreshAllelesPerOrigInd,AvgFreshAllelesPerIntroInd
+
+def getDataTotal():
+
+    totalNumFreshwaterAdaptedAlleles = []
+
+    for i in range(5):
+        T = []
+
+        total_File = open("./Output1/MyRecipe"+recipes[i]+"/numfreshAlleles.txt","r")
+        first = total_File.readline()
+        header = first.split()
+        numSamples = int(header[0])
+        half = numSamples / 2
+        count = 0
+        for line in total_File:
+            if count > half:
[…]

### 2019-05-08 Scripts/fst_faa.py

@@ -1,0 +1,270 @@
+
+import matplotlib as mpl
+mpl.use('pdf')
+import msprime
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+
+plt.rc('font', family='serif', serif='Times')
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=12)
+plt.rc('ytick', labelsize=12)
+plt.rc('axes', labelsize=12)
+
+mig_rate = ["$0.01$","$0.1$","$1.0$","$10.0$","$100.0$"]
+combinations = ["$F_{st}$"+"\nMarine\nOld Lakes","$F_{st}$"+"\nMarine\nNew Lakes","$F_{st}$"+"\nOld Lakes\nNew Lakes"]
+x = [4949999, 5049999, 9999999, 14949999, 15049999, 19999999, 24949999, 25049999, 29999999, 34949999, 35049999, 39999999, 44949999, 45049999, 49999999, 54949999, 55049999, 59999999, 64949999, 65049999, 69999999, 74949999, 75049999, 79999999, 84949999, 85049999, 89999999, 94949999, 95049999, 99999999]
+
+effectIntervals = []
+
+for i in range(len(x)-1):
+    if(i % 3 == 0):
+        effectIntervals.append([x[i],x[i+1]])    
+
+meanEffect = []
+for e in effectIntervals:
+    meanEffect.append(sum(e)/len(e))
+
+treeSequences = []
+
+def populateTreeSequences():
+
+    for i in range(5):
+        ts = msprime.load("./Output1/MyRecipe9_1_"+str(i)+"/EndTrees.trees") 
+        mut_ts = msprime.mutate(ts,rate=1e-8,keep=True,random_seed=25)
+        treeSequences.append(mut_ts)
+
+
+def mutations_fst(tree_sequence, pop0, pop1, windows):
+    '''
+    Compute mean Fst across all variants in each window.
+    Assumes an infinite-sites model.
+    '''
+    n0 = len(pop0)
+    n1 = len(pop1)
+    num_windows = len(windows) - 1
+    if windows[0] != 0.0:
+        raise ValueError(
+            "Windows must start at the start of the sequence (at 0.0).")
+    if windows[-1] != tree_sequence.sequence_length:
+        raise ValueError("Windows must extend to the end of the sequence.")
+    for k in range(num_windows):
+        if windows[k + 1] <= windows[k]:
+            raise ValueError("Windows must be increasing.")
[…]

### 2019-05-08 Scripts/haplotypes.py

@@ -1,0 +1,105 @@
+import matplotlib as mpl
+mpl.use('pdf')
+import msprime
+import pyslim
+import io
+import sys
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import numpy as np
+
+plt.rc('font', family='serif', serif='Times')
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=12)
+plt.rc('ytick', labelsize=12)
+plt.rc('axes', labelsize=12)
+
+ts = pyslim.load("./Output1/MyRecipe9_1_2/EndTrees.trees")
+
+inds = [ind.id for ind in ts.individuals()]
+sample_subset = np.unique(np.random.choice(inds,6000))
+sample_nodes = []
+for i in sample_subset:
+    ind = ts.individual(i)
+    sample_nodes.append(ind.nodes[0])
+    sample_nodes.append(ind.nodes[1])
+    
+ts = ts.simplify(sample_nodes,filter_populations=False)
+
+extant_samples = []
+for n in ts.nodes():
+    if (n.time == 0.0):
+        extant_samples.append(n.id)
+
+tables = ts.dump_tables()
+x = [[],[],[]]
+y = [[],[],[]]
+c = [[],[],[]]
+varit = ts.variants()
+colors = ['m','c','y']
+
+ind_loc = np.array([i.location[0] + np.random.uniform(-0.5,0.5) for i in ts.individuals()])
+phenotypes = np.array([0.0 for _ in ts.individuals()])
+populations = np.array([ts.node(ind.nodes[0]).population for ind in ts.individuals()])
+
+for v in varit:
+    mut_meta = pyslim.decode_mutation(v.site.mutations[0].metadata)[0]
+    selectionCoeff = mut_meta.selection_coeff
+    mutType = mut_meta.mutation_type
+    color = None
+    dominance = None
+    if(mutType == 1 or mutType == 2):
+        color = 'm'
+        dominance = 'dom'
+    elif(mutType == 4 or mutType == 5):
+        color = 'c'
+        dominance = 'rec'
+    else:
+        color = 'y'
+        dominance = 'add'
+
+    for ind in ts.individuals():
+        num_muts = sum([v.genotypes[n] for n in ind.nodes])
+        if dominance == 'dom' and num_muts > 0:
+            phenotypes[ind.id] += 2 * selectionCoeff;
+        if dominance == 'rec' and num_muts == 1:
[…]

### 2019-05-08 Scripts/origin_adaptive_varients.py

@@ -1,0 +1,363 @@
+import matplotlib as mpl
+mpl.use('pdf')
+import msprime
+import numpy as np
+import pyslim
+import matplotlib.pyplot as plt
+
+plt.rc('font', family='serif', serif='Times')
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=12)
+plt.rc('ytick', labelsize=12)
+plt.rc('axes', labelsize=12)
+
+samples_prop_de_novo = []
+samples_prop_migrant = []
+samples_prop_standing = []
+
+
+def mutationsLocation(mutationSlimIds,treeSequence):
+    '''
+    take in SLiM mutation Id's and return their respective location using pyslim.
+    '''
+
+    positions = np.empty(len(mutationSlimIds),dtype=int)
+    mutation_msp_ids = np.empty(len(mutationSlimIds),dtype=int)
+    mutations = treeSequence.mutations()
+    found_count = 0
+    for m in mutations:
+        try:
+            idx = mutationSlimIds.index(int(m.derived_state))
+            positions[idx] = int(m.position)
+            mutation_msp_ids[idx] = int(m.id)
+            found_count += 1
+        except:
+            continue
+    #print(found_count," ",len(mutationSlimIds))
+    #print(positions)
+
+    assert(found_count == len(mutationSlimIds))
+    return positions, mutation_msp_ids
+
+def countFitnessOfAncestors(mutationSlimIds,treeSequence):
+
+    positions,mutationMspIds = mutationsLocation(mutationSlimIds,treeSequence)
+
+    #Get the extant samples from p3 and the initial population from p3. 
+    #They are all marked as samples in the treeSequence
+    extantSampleIds = []
+    IntroducedPopulationsIds = [] 
+    for sam in treeSequence.samples():
+        node = treeSequence.node(sam)
+        time = node.time
+        pop = node.population
+        if time == 0.0 and pop == 3:
+            extantSampleIds.append(sam)
+        elif pop == 3:
+            IntroducedPopulationsIds.append(sam)
+            T = time
+        else:
+            continue
+
+    fitnesses = np.array([0 for _ in IntroducedPopulationsIds])
+    #print(len(extantSampleIds))
+    #return
+    tt = ts.trees(tracked_samples=extantSampleIds)
+    t = next(tt)
+    for var in ts.variants():
[…]

### 2019-05-08 Scripts/phenotype_distributions.py

@@ -1,0 +1,102 @@
+import sys
+import matplotlib as mpl
+mpl.use('pdf')
+import matplotlib.pyplot as plt
+import numpy as np
+
+plt.rc('font', family='serif', serif='Times')
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=12)
+plt.rc('ytick', labelsize=12)
+plt.rc('axes', labelsize=12)
+
+num = "1"
+
+Traces = []
+
+recipes = ['9_'+num+'_0','9_'+num+'_1','9_'+num+'_2','9_'+num+'_3','9_'+num+'_4']
+
+xAxis = []
+xAxis2 = []
+col1 = []
+col2 = []
+col3 = []
+
+numSamplingPoints = 0
+interval = 0 
+intro = 0 
+end = 0
+newLakesIntroduced = 0 
+
+for i in range(5):
+
+	c1 = []
+	c2 = []
+	c3 = []
+
+	File = open("./Output1/MyRecipe"+recipes[i]+"/AveragePhenotypeThroughout.txt","r")
+	first = File.readline()
+	numSamplingPoints,interval,intro,end = (int(s) for s in first.split()) 
+
+	newLakesIntroduced = ( intro / interval ) -1
+
+	count = 0
+	for line in File:
+		if(count < newLakesIntroduced):
+			li = line.split()
+			c1.append(float(li[0]))
+			c2.append(float(li[1]))
+		else:
+			li = line.split()
+			c1.append(float(li[0]))
+			c2.append(float(li[1]))
+			c3.append(float(li[27]))
+
+		count += 1;		
+
+	col1.append(c1)
+	col2.append(c2)
+	col3.append(c3)
+
+data_a = col1
+data_b = col2
+data_c = col3
+
+ticks = ['0.01','0.1','1.0','10.0','100.0']
+
+def set_box_color(bp, color):
+    plt.setp(bp['boxes'], color=color)
+    plt.setp(bp['whiskers'], color=color)
+    plt.setp(bp['caps'], color=color)
+    plt.setp(bp['medians'], color=color)
+
+fig1,ax1 = plt.subplots()
+print(len(data_a))
+print(np.array(range(len(data_a)))*3.0-0.6)
+
+bpl = ax1.boxplot(data_a, positions=np.array(range(len(data_a)))*3.0-0.6, sym='', widths=0.4)
[…]

### 2019-05-08 Scripts/plot_phenotype.py

@@ -1,0 +1,103 @@
+import matplotlib as mpl
+mpl.use('pdf')
+import matplotlib.pyplot as plt
+import sys
+import numpy as np
+
+plt.rc('font', family='serif', serif='Times')
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=12)
+plt.rc('ytick', labelsize=12)
+plt.rc('axes', labelsize=12)
+
+def readHeader(firstLine):
+
+	numberOfSamplingPoints,samplingInterval,introduction,endofSim = (int(i) for i in firstLine.split())
+	samplesUntilIntroduction = introduction//samplingInterval - 1
+	return numberOfSamplingPoints,samplingInterval,introduction,endofSim,samplesUntilIntroduction
+
+def phenoData(recipeNum):
+
+    data = []
+    numLakes = 25
+
+    for i in range(0,((2*numLakes)+3)):
+        data.append([])
+
+    x_axis = []
+    x_axis2 = []
+    lake_traces = []
+    File = open("./Output1/MyRecipe"+recipeNum +"/AveragePhenotypeThroughout.txt","r")
+    first = File.readline()
+    numberOfSamplingPoints,samplingInterval,introduction,end,samplesUntilIntroduction = readHeader(first)
+    print(numberOfSamplingPoints,samplingInterval,introduction,end,samplesUntilIntroduction)
+
+    for i,line in enumerate(File):
+        for j,s in enumerate(line.split()):
+            data[j].append(float(s))
+
+    #Give all data a zero point because that wasn't sampled
+    for i in range(0,(numLakes+2)):
+        data[i] = [0] + data[i]
+
+    for i in range(0,numberOfSamplingPoints + 1):
+        x_axis.append(i*samplingInterval)
+
+    for i in range(samplesUntilIntroduction,numberOfSamplingPoints):
+        x_axis2.append(i*samplingInterval)
+    
+    return data, x_axis, x_axis2
+
+firstdata,x1,x2 = phenoData("9_1_0")
+seconddata,x11,x22 = phenoData("9_1_1")
+
+fig,axes = plt.subplots(nrows=2,ncols=1,sharex = True,sharey = True)
+
+lw1 = 0.7
+lw2 = 0.03
+
+
[…]

### 2019-05-08 Scripts/plot_phenotype_supp.py

@@ -1,0 +1,114 @@
+import matplotlib as mpl
+mpl.use('pdf')
+import matplotlib.pyplot as plt
+import sys
+import numpy as np
+
+plt.rc('font', family='serif', serif='Times')
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=12)
+plt.rc('ytick', labelsize=12)
+plt.rc('axes', labelsize=12)
+
+def readHeader(firstLine):
+
+	numberOfSamplingPoints,samplingInterval,introduction,endofSim = (int(i) for i in firstLine.split())
+	samplesUntilIntroduction = introduction//samplingInterval - 1
+	return numberOfSamplingPoints,samplingInterval,introduction,endofSim,samplesUntilIntroduction
+
+def phenoData(recipeNum):
+
+    data = []
+    numLakes = 25
+
+    for i in range(0,((2*numLakes)+3)):
+        data.append([])
+
+    x_axis = []
+    x_axis2 = []
+    lake_traces = []
+    File = open("./Output1/MyRecipe"+recipeNum +"/AveragePhenotypeThroughout.txt","r")
+    first = File.readline()
+    numberOfSamplingPoints,samplingInterval,introduction,end,samplesUntilIntroduction = readHeader(first)
+    print(numberOfSamplingPoints,samplingInterval,introduction,end,samplesUntilIntroduction)
+
+    for i,line in enumerate(File):
+        for j,s in enumerate(line.split()):
+            data[j].append(float(s))
+
+    #Give all data a zero point because that wasn't sampled
+    for i in range(0,(numLakes+2)):
+        data[i] = [0] + data[i]
+
+    for i in range(0,numberOfSamplingPoints + 1):
+        x_axis.append(i*samplingInterval)
+
+    for i in range(samplesUntilIntroduction,numberOfSamplingPoints):
+        x_axis2.append(i*samplingInterval)
+    
+    return data, x_axis, x_axis2
+
+firstdata,x1,x2 = phenoData("9_1_2")
+seconddata,x11,x22 = phenoData("9_1_3")
+thirddata,x111,x222 = phenoData("9_1_4")
+
[…]

### 2019-05-08 Scripts/time_to_adaptation.py

@@ -1,0 +1,73 @@
+import matplotlib as mpl
+mpl.use('pdf')
+import sys
+#import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+    
+plt.rc('font', family='serif', serif='Times')
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=12)
+plt.rc('ytick', labelsize=12)
+plt.rc('axes', labelsize=12)
+
+num = 1
+
+x = [0.00005,0.0005,0.005,0.05,0.5]
+x2 = ['0.01','0.1','1.0','10.0','100.0']
+color_shape = ['go','b^','r*']
+
+timeToAdaptation = []
+numShared = []
+totalNumFWAA = []
+numShared_DivTotal = []
+corFWAA = []
+corEffect = []
+
+for j in [1]:
+
+	TTA = []
+	num = str(j)
+	recipes = ['9_'+num+'_0','9_'+num+'_1','9_'+num+'_2','9_'+num+'_3','9_'+num+'_4']
+	
+	for i in recipes:
+
+		File = open("./Output1/MyRecipe"+i+"/Adaptation.txt","r")
+		header = File.readline()
+		FWAA_P2 = File.readline()
+		FWAA_P3 = File.readline()
+
+		TTA.append(int(File.readline().split()[0]))
+		numShared_ = File.readline().split()[0]
+		totalNumFWAA_ = File.readline().split()[0]
+
+		numShared_DivTotal.append(float(numShared_ )/float(totalNumFWAA_))
+
+		corFWAA.append(File.readline().split()[0])
+		corEffect.append(File.readline().split()[0])
+
+	timeToAdaptation.append(TTA)
+
+timeToAdaptation = np.array(timeToAdaptation)	
+means = np.mean(timeToAdaptation,axis=0)
+
+f1, ax1 = plt.subplots()
+for k in range(len(timeToAdaptation)):
+	ax1.plot(x2[:-1],timeToAdaptation[k][:-1],color_shape[k])
+ax1.plot(x2[:-1],means[:-1],"y--")
+ax1.plot(x2[-1],10,"")
+
+ax1.grid(axis = 'y')
+ax1.set_ylabel('Time (Generations)')
+ax1.set_xlabel('$M$')
+ax1.set_title('Time to Adaptation')
+
+ax1.set_yscale('log')
[…]

### 2019-05-08 Stickleback_Paper.tex
updated readme to include acstract

@@
 
-%\singlespacing
 \onehalfspacing
-%\doublespacing
 
@@
 \newcommand{\jgg}[1]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=black]{#1}}
-\newcommand{\bill}[1]{\todo[linecolor=green,backgroundcolor=yellow!25,bordercolor=black]{#1}}
+\newcommand{\wac}[1]{\todo[linecolor=green,backgroundcolor=yellow!25,bordercolor=black]{#1}}
 
@@
 
-% Jared's general notes for formatting
-% - ``citep'' for ``cite in parentheses''. (and, citations come before the punctuation)
-% - in latex, use ` `words' ' to get the open- and closed-double quotes (not the double-quote character)
-% - "citet" for "cite, as part of the text"
-
 The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations (\textbf{cite}).
@@
 
-%\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
-%\wac{I did a little editing but Peter can still totally rewrite. Regardless it needs to be streamlined a fair bit}
+\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
+\wac{I did a little editing but Peter can still totally rewrite. Regardless, it needs to be streamlined a fair bit}
 
@@
 Together these simulation findings enhance and extend the existing rich empirical work on stickleback evolutionary genomics.
-
-%%%%%%%%%%%%%%%%
 
@@
 
-We also used neutral mutations to calculate $F_{ST}$, calculated on a per-locus basis and then averaged in windows as a measure of between-population relative differentiation.
+We used neutral mutations to calculate $F_{ST}$, calculated on a per-locus basis and then averaged in windows as a measure of between-population relative differentiation.
 Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively, and $\bar p = (p_f + p_m)/2$, then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
 
-%%%%%%%%%%%%%%%%%%
-\section*{Results}
+\section*{Results \& Discussion}
 
@@
 
-%%%%%%%%%%%%
 \subsection*{Local Adaptation: differentiation with gene flow}
@@
         $F_{st}$ is calculated between all marine individuals, and all lakes pooled together.
-        %Ten vertical dotted pink lines in each subplot 
-                %show regions which have the potential to affect phenotype.
         All locations of pre-existing freshwater adapted alleles have been highlighted by 
@@
 
-%\begin{comment}
-%\paragraph{Standing genetic variation} 
-%Despite a dramatic difference in the amount of allelic sharing between lakes, 
-%standing genetic variance (Figure \ref{fig:SGV}) 
-%was around 0.05 in the freshwater habitat across all but the highest migration rates. 
-%Concurrently, genetic variation in the marine population 
-%steadily increased with migration to a similar value. 
-%On the one hand, it is not surprising that lakes, as a group, 
-%carry more genetic variation than the marine habitat, 
-%since population subdivision allows different alleles to become common in each. 
-%However, it is striking that at $m=.005$, 
-%the marine habitat carries as much genetic variation, 
-%despite a lack of any substantial migration load.
-%\plr{perhaps we want to remove this bit? revisit if it is telling us something important.}
-%\end{comment}
-
-%%%%%%%%%%%%
 \subsection*{Scans for selection}
@@
         across rates of gene flow,
-        % \textbf{``Best''} 
+        \textbf{``Best''} 
         quantified as the most negative trait value achievable
@@
         (see text for details).
-        % \textbf{``Expected''} shows the trait value
-        %     achieved by averaging haplotypic values weighted by fitness
-        %     (see text for details).
     } \label{tab:linkage}
@@
 
-%%%%%%%%%%%%
 \subsection*{Theoretical expectations}
@@
 %%%%%%%%%%%%%%%%%%%%
-\section*{Discussion}
+\section*{Summary}
 

### 2019-05-10 Stickleback_Paper.tex
Draft 5 before citations

@@
 
-\title{A few stickleback can transport adaptive alleles to new lakes}
+\title{A few stickleback suffice to transport adaptive alleles to new lakes}
 \author{Jared Galloway, William A. Cresko, and Peter Ralph}
 \maketitle
+\jgg{To me, this title makes it sound like only a "select few" stickleback have the ability to transport, no?}
 
@@
 Threespine stickleback provide a striking example of local adaptation to divergent habitats while still being connected by recurrent gene flow. 
-These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases smaller freshwater populations have been established ``de novo'' from marine colonists.
-Previous results show that a similar freshwater phenotype often evolves in independent populations soon after colonization using largely the same standing genetic variation, even in very distant geographic regions.
-But how many stickleback migrants are needed to capture a sufficient amount of standing variation to fuel adaptation?
+These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases the smaller freshwater populations have been established ``de novo'' from marine colonists.
+Previous results have found similar phenotypes, between independently evolved freshwater populations, are derived largely from the same standing genetic variants found to be Identical-by-decent
+Geographic isolation preventing direct migration between the freshwater populations strongly suggests that these shared adaptive alleles are transported through hybridization events with the wide-spanning marine population which acts as a medium for most freshwater populations.
+However, it is still largely unknown how gene flow, recombination, and selection jointly impact standing variation to fuel adaptation.
 Here we use genealogical simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback, and to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and re-use of standing genetic variation.
-In our simulations we find that repeated adaptation uses a shared set of alleles that are maintained at low frequency by migration-selection balance in ocean populations.
+In our simulations we find that repeated adaptation uses a shared set of alleles that are maintained at low frequency by migration-selection balance in oceanic populations. 
 This process occurs over a realistic range of intermediate levels of gene flow that match previous empirical population genomic studies in stickleback. 
-Examining these simulations more deeply reveals that low gene flow leads to slow, independent adaption to district habitats, whereas high gene flow leads to significant mutation load but an increased probability of successful population genomic scans for locally adapted alleles.
-We also find that the original generation of marine individuals that colonized the lakes is most often the sole source of most freshwater adapted alleles.
+Examining these simulations more deeply reveals that lower levels of gene flow leads to slow, independent adaptation to district habitats, whereas higher levels gene flow leads to significant mutation load, but an increased probability of successful population genomic scans for locally adapted alleles.
+Surprisingly, we find that the genealogical origins of most freshwater adapted alleles can be traced back to the original generation of marine individuals that colonized the lakes, as opposed to subsequent migrants.
 These simulations provide deeper context for existing studies of stickleback evolutionary genomics, and provide guidance for future empirical studies in this model.
 More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.  
+\jgg{Is this intro too stickleback-y?}
 
@@
 The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations (\textbf{cite}). 
-While it has been very useful heuristically for guiding studies in the field and lab (\textbf{cite}), this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures (\textbf{cite}).
-In particular, empirical studies have increasingly pointed to the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats and experiencing diverse selective pressures while still exhibiting significant gene flow (\textbf{cite}).
+While it has proved to be a useful baseline for understanding the genetic variation we see in species today (\textbf{cite}), this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures (\textbf{cite}).
+In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats.
+Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood (\textbf{cite}).
 A growing number of empirical studies have identified instances of convergent evolution using SGV \citep{nelson2017ancient,Schrider2017,Barrett2008}. 
-However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during invasion and adaptation to new environments (\textbf{cite}).
+However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during colonization and adaptation to new environments (\textbf{cite}).
 It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via SGV. 
-While emprical studies in the field and lab are essential, the use of simulation models of specific systems can greatly advanced our understanding of the adaptation genetic dynamics using SGV.
+While empirical studies in the field and lab are essential, modeling specific systems can help us understand the nature of standing genetic variants to fuel local adaptation as well as the genomic signals we can expect to see.
+% of the adaptation genetic dynamics using SGV.
+%\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
+%\wac{I did a little editing but Peter can still totally rewrite. Regardless, it needs to be streamlined a fair bit}
 
-\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
-\wac{I did a little editing but Peter can still totally rewrite. Regardless, it needs to be streamlined a fair bit}
-
-The threespine stickleback fish has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes (\textbf{cite}) and the much more extensive use of SGV than once thought (\textbf{cite}).
-The ancestral marine form of stickleback has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere (\textbf{cite}).
-While geographic isolation often prevents direct migration among freshwater populations, they frequently evolve similar (but not exactly the same) phenotypes (\textbf{cite}).
-The most recent evolutionary genomic studies on stickleback document that while adaptation to freshwater can occur over decades (\textbf{cite}), the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years (\textbf{cite}).
-Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations (\textbf{cite}).
+The ancestral marine form of threespine stickleback fish has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere (\textbf{cite}).
+This model organism has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes (\textbf{cite}) and the much more extensive use of SGV than once thought (\textbf{cite}).
+%The ancestral marine form of threespine stickleback has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere (\textbf{cite}).
+While geographic isolation often prevents direct migration between freshwater populations, they frequently evolve similar phenotypes (\textbf{cite}). 
+The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years (\textbf{cite}), independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations (\textbf{cite}).
 For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska (\textbf{cite}). 
 Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years (\textbf{cite}).
-Amazingly, nearly a quarter of the genome diverged significantly in this same time but - with very little change in the rest of the genome - in a pattern that mirrored previous studies of freshwater stickleback populations thousands of years old (\textbf{cite}).
+Amazingly, nearly a quarter of the genome diverged significantly in this same time but - with very little change in the rest of the genome - in a pattern that mirrored previous genomic studies of freshwater stickleback populations that have been geographically separated for thousands of years (\textbf{cite}).
 
-But how can evolution occur at such a rapid pace in this and other species?
+But how can evolution occur at such a rapid pace?
 Waiting for new mutations to arise in each lake or pond would take much longer than the decades since the Alaskan earthquake (\textbf{cite}). 
 Even more improbable is having divergence cluster in such similar genomic regions across independent populations. 
-An alternative hypothesis, supported by the most recent population genomic analyses, is that the marjority of alleles important for freshwater adaptation are maintained in the marine individuals due to recurrent gene flow from freshwater back in to marine populations.
-In 2009 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter hypothesis'' to describe the process of alleles benefical in freshwater populations are maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation into new freshwater habitats.
+An alternative hypothesis, supported by the most recent population genomic analyses, is that the majority of alleles important for freshwater adaptation are maintained in the marine individuals due to recurrent gene flow from freshwater back in to marine populations.
+In 2009 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter''- hypothesis to describe the process of alleles beneficial in freshwater environment being maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation to new freshwater habitats.
 The first clear example of the global reuse of SGV in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates (\textbf{cite}). 
@@
 A curious natural history observation undergirding many of these questions is the fact that some newly formed freshwater habitats, such as the ponds on Middleton Island, are quite small and presumably the number of initial marine migrants is few.
-Here we use forward simulations implementing in SLiM and incorporating linkage to model the stickleback metapopulation and address these questions.
+Here we use forward simulations implemented in SLiM incorporating linkage and realistic length genomes to model the stickleback metapopulation and address these questions \citep{haller2017slim,haller2018slim3}.
 We ask how variation in amount of gene flow affects the genetic architecture of local adaptation to newly created freshwater ponds.
-Because we are documenting the entire genealogy of all alleles, we can determine the distribution and abundance of haplotypes across all in silico marine and freshwater populations to know the timing and proportion of potentially adaptive alleles that are actually utilized in each population.
+Because we are documenting the entire genealogy of all alleles, we can determine the distribution and abundance of haplotypes across all in marine and freshwater populations to know the timing and proportion of potentially adaptive alleles that are actually utilized in each population.
 In addition, we can document how these adaptive alleles are distributed across the genomes, and as a consequence determine how this may affect the efficacy of genome scans for between-habitat differentiation.
@@
 \section*{Methods}
+\jgg{Any thoughts on putting Methods after results?}
 
-To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait.
+To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement complex forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait.
 The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography, and included divergent selection in the two habitats which had substantial spatial structure.
@@
 Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback populations observed in nature (see the Discussion for more on this).
+\jgg{I don't know where we should talk about our small population sizes ... Is this even worth mentioning here?}
 
@@
 
-Simulations have two habitat types -- Marine and Freshwater -- defined by their selective pressures, each with 5,000 diploid individuals.
+Our simulations have two habitat types -- Marine and Freshwater -- defined by their selective pressures, each with 5,000 diploid individuals. 
 The arrangement of these habitats, depicted in Figure~\ref{fig:Geo} roughly models a set of freshwater habitats along a stretch of coastline. 
@@
 Subsequent mutations at the same locus replace the previous allele.
-We also used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence} to add neutral mutations across the entire genome at a rate of $10^{-7}$ per locus per generation,  as described in \citet{kelleher2018efficient}.
+%We also used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence} to add neutral mutations across the entire genome at a rate of $10^{-8}$ per locus per generation,  as described in \citet{kelleher2018efficient}.
 
@@
 We use SLiM to simulate a Wright--Fisher population with non-overlapping generations and a fixed population size of 5,000 diploid individuals in each habitat. 
-Each generation, the two parents of each new offspring are chosen proportional to their fitness (all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-7}$ per locus per generation).
+Each generation, the two parents of each new offspring are chosen proportional to their fitness (all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-8}$ per locus per generation).
 Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation.
@@
 This creates two \emph{sets} of lakes, which along with the marine population have a total of 15,000 individuals.
-Since this introduction of lakes doubles the number of lake-to-marine immigrants,the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.
+Since this introduction of lakes doubles the number of lake-to-marine immigrants, the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.
 
@@
 This allowed us to directly query the true origins of adaptive alleles.
-In addition, it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations which were retroactively added to the gene trees at a rate of $10^{-7}$ per locus per generation, as described in \citet{kelleher2018efficient}.
-
+In addition, it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations which were retroactively added to the gene trees at a rate of $10^{-8}$ per locus per generation, as described in \citet{kelleher2018efficient}.
 The output tree sequence from each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
@@
 
-The proportion of trait-affecting alleles in new lakes that fall in these categories measures the degree to which selection in the new environments made use of
-(1) new mutation, 
-(2) post-colonization migration, 
-(3) standing variation at migration--selection balance, and
-(4) standing variation at mutation--selection balance.
+The proportion of trait-affecting alleles in new lakes that fall in these categories measures the degree to which selection in the new environments made use of (1) new mutation, (2) post-colonization migration, (3) standing variation at migration--selection balance, and (4) standing variation at mutation--selection balance.
 
@@
 To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$. 
-Below, we often refer to these migration rates in terms of the number of migrants per lake, per generation, which we denote $M$.
-Since each lake contains 200 individuals, $M=200m$.
+Below, we often refer to these migration rates in terms of the number of migrants per lake, per generation, which we denote $M$. Since each lake contains 200 individuals, $M=200m$. 
 Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
-At very low rates of gene flow, each new lake's population adapted almost completely independently throught \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations).
+At very low rates of gene flow, each new lake's population adapted almost completely independently through \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations). 
 At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles \citet{bolnick2007natural}. 
 Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes. 
-Perhaps surprisingly, only a few migrants per generation from the lakes to the ocean were needed to maintain sufficient genetic variation in the ocean to accelerate adaptation in new lakes.
+Perhaps surprisingly, only a few migrants per generation from the lakes to the ocean were needed to maintain sufficient genetic variation in the ocean to dramatically accelerate adaptation in new lakes.
 
@@
           \caption{
-            Distribution of mean trait values across generations of the simulation,
+            Distribution of mean individual trait values across generations of the simulation,
             for different migration rates.
@@
 Establishment of new alleles in the lakes is visible in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations.
-%Trait variation within each population was small compared to the difference between populations.
 At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites(about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats.
@@
         with columns corresponding to increasing gene flow from left to right.
-        $F_{st}$ is calculated between all marine individuals, and all lakes pooled together.
+        $F_{ST}$ is calculated between all marine individuals, and all lakes pooled together.
         All locations of pre-existing freshwater adapted alleles have been highlighted by 
@@
 The numbers in Figure~\ref{fig:MPFAI} suggest that the dramatic increase in speed of local adaptation we observed above occurs because higher gene flow between populations allows sharing of freshwater alleles between populations.
-
-%however, it is not enough for the marine population to hold bits and peices from independently formed haplotypes,
-%For selection to act on SGV on the scale we observe in nature, the gene flow 
-%initial generation to be able to rebuild
-
 We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele in each individual in the new lakes, as defined in the Methods. 
@@
 The rate at which migration load becomes substantial, 10 migrants per lake per generation, only replaces 5\% of each population each generation with migrants from the other habitat, but this is sufficient to shift the mean trait values to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
+These results present the sensitivity of a population to the amount of gene flow it experiences from neighboring species.
+In our simulations, the continued gene flow from freshwater populations meant that the alleles were maintained at a constant rate with fixed demographics. 
+However, future modeling efforts must be made to understand what population sizes, demographics, and evolutionary histories impact an ancestral population's capacity for carrying maladaptive alleles in its opposing environment, as well as what role a haplotype's genomic architecture play it to be rebuilt in a novel environment.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%
-\section*{Summary}
+\section*{Conclution}
 
-In this paper, we have described a realistic simulation model of a coastal meta-population to observe the effects of selection of genetic variation (SGV) across a wide range of gene flow rates.
+In this paper, we have described a realistic simulation model of a coastal meta-population to observe the effects of selection on genetic variation across a wide range of gene flow rates.
 We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
@@
 Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.
+We find that, after local adaptation in new lakes, the majority of adaptive variant's genealogy can be traced back to the original generation as opposed to post-colonization migrant alleles.
+Additionally, quantifying the inference power of $F_{ST}$ genomic scans has revealed that higher rates of gene flow are generally more useful for identifying 
+Perhaps the most notable finding is that the large marine population was able to harbor and distribute deleterious alleles to the surrounding freshwater populations with only a few migrants per population, per generation.
 
-While our findings prove useful to the understanding of stickleback, they are not autonomous to this species. 
-Now, we will discuss the broader implications of our findings on the more general framework of similar systems, we also explore what other questions remain. 
-Small patches of similar selective pressures can be found in many places such as high altitudes, underwater volcanic vents, or post forest fire environments. 
-Recently,  \citep{Ferretti2018}, discovered seven new species of Hapalotremus (Theraphosinae), living at the extremely high altitudes distributed along the Andes and Yungas in western South America.
-With the ancestral species at much lower altitudes, one could image high altitude alleles that could potentially be carried to help in the colonization of neighboring mountain environments. 
-Systems of meta-populations like this could potentially arise in many places
-where the ancestral species carries around with it beneficial alleles for selective pressures it has encountered in the past and remains connected to through hybridization events.
-
- \paragraph{Atmosphere's ability to carry whoever scotty is beaming up ... }
-Perhaps the most notable finding in this study is that large, ancestral populations have the capacity to maintain alleles which are potentially deleterious for an extended amount of time.  
-This principle, deriving solely from the properties of sex and natural selection, could be applicable to almost any species experiencing the influence of divergent selection with introgression.
-And to explore further how this process is carried out exactly, many of the questions lie in understanding which properties of genomic architecture and demographics allow this transportation to be possible.
-In our simulations, the continued gene flow from freshwater populations meant that the alleles were maintained at a constant rate with fixed demographics. 
-Two possible questions that might arise from this are:
-(1) How do population sizes, demographics, and evolutionary histories impact an ancestral population's capacity for carrying maladaptive alleles in its opposing environment, and 
-(2) What role does genomic architecture play in the ability for a haplotype to be rebuilt.
-
-\paragraph{Realized genomic architecture}
-Additionally have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
-This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
-differences between populations that have little gene flow between them.
-It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
-mistaken for being causative as they often display the same amount of divergence.
-In general, our study may suggest that diversity scans may not prove as useful in causal loci inference as previously thought.
-It is important to note that these summary statics could prove to be very useful if used beyond just the peaks seen when looked at as a function of genomic position. 
-Many methods, such as advanced machine learning, have proved to be very good at taking in noisy data and parsing out important information \citet{Schrider2018}.
-
-\paragraph{Implications for the ``Transporter"-hypothesis}
-This study was in large, supportive of the ``Transporter"-hypothesis suggested by \citep{schluter2009genetics}.
-Provided there is enough dispersal, we found that repeated colonization and adaptation of freshwater stickleback was able to produce beneficial alleles which could then be 
-supported by the marine stickleback population through the offspring of hybridization events. 
-When new freshwater environments were colonized by the marine stickleback, selection was able to rebuild the freshwater haplotype. 
-Shown in Table~\ref{tab:linkage}, we found that a populations capacity to cary maladaptive alleles remains pretty constant over realistic ranges of dispersal, but low rates of gene flow may 
-"mask" the alleles through linkage which counteracts the fitness disadvantage. 
-This linkage prevents the ability for a sub-population to rapidly select on the beneficial SGV as recombination acts as a bottleneck to evolution in this situation. 
-Additionally, In Figure~\ref{fig:Origin} we found that the genetic basis for local adaptation, when selecting upon SGV, is propagated largely from the initial generation of inhabitants. 
-This contrasts the belief that subsequent migration is a necessity in newly established populations ability to rebuild a haplotype.
-In other words, this finding suggests that with high probability, a randomly chosen subset from ancestral species in the meta-populations like marine stickleback, 
-carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%DO WE NEED THIS?
-\begin{comment}
-\subsection*{thresholds}
-
-We have found that too little migration leads to selection upon new mutations in all subpopulations and lakes alike.
-In contrast, at high migration rates we have seen that migration load limits the ability for species to locally adapt to the selective pressure of their environment.
-This leads us to consider a window of introgression which allows for the transportation of FAA's without migration load.
-
-\subsection*{connect results back to real data?}
-
-\paragraph{The adaptive filter?}
-RAMBLING THOUGHTS HERE
-Since larger effect alleles are more likely to establish,
-be it by mutation or migration,
-repeated colonization of new freshwater habitats will select for larger alleles,
-be it single alleles or haplotypes bound together by an inversion.
-However, these are more strongly selected against in the interstitial time.
-Being recessive would help with this,
-but would also make it more difficult to establish.
-
-\jgg{TODO: effect of smaller than realistic population sizes?}
-
-\jgg{TODO: Talk about the role recombination plays}
-This also brings to light the role of recombination in a system of migration selection balance. 
-On one hand, a lower recombination rate would allow for the freshwater haplotype to remain "intact" 
-within the marine environment, however, without significant introgression the haplotype would quickly get selected against in marine populations. 
-On the other hand a higher recombination rate would, in theory, allow freshwater adapted alleles more longevity 
-as they would hitchhike with marine adapted haplotypes.
-
-\paragraph{Modeling assumptions}
-Our simulations had much smaller population sizes than real populations.
-How might this affect things?
-\end{comment}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\jgg{This feels awkward, help please.}
+%This is evidence that standing genetic variation from gene flow likely plays a key role in local adaptation for many sub-species.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
 \clearpage
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Below is some content that may be useful, but we can delete once the manuscript is finalized.
+
+% Discussion
+
+While our findings prove useful to the understanding of stickleback, they are not autonomous to this species. 
+Now, we will discuss the broader implications of our findings on the more general framework of similar systems, we also explore what other questions remain. 
+Small patches of similar selective pressures can be found in many places such as high altitudes, underwater volcanic vents, or post forest fire environments. 
+Recently,  \citep{Ferretti2018}, discovered seven new species of Hapalotremus (Theraphosinae), living at the extremely high altitudes distributed along the Andes and Yungas in western South America.
+With the ancestral species at much lower altitudes, one could image high altitude alleles that could potentially be carried to help in the colonization of neighboring mountain environments. 
+Systems of meta-populations like this could potentially arise in many places
+where the ancestral species carries around with it beneficial alleles for selective pressures it has encountered in the past and remains connected to through hybridization events.
+
+ \paragraph{Atmosphere's ability to carry whoever scotty is beaming up ... }
+Perhaps the most notable finding in this study is that large, ancestral populations have the capacity to maintain alleles which are potentially deleterious for an extended amount of time.  
+This principle, deriving solely from the properties of sex and mutation, could be applicable to almost any species experiencing the influence of divergent selection with introgression.
+And to explore further how this process is carried out exactly, many of the questions lie in understanding which properties of genomic architecture and demographics allow this transportation to be possible.
+In our simulations, the continued gene flow from freshwater populations meant that the alleles were maintained at a constant rate with fixed demographics. 
+Two possible questions that might arise from this are:
+(1) How do population sizes, demographics, and evolutionary histories impact an ancestral population's capacity for carrying maladaptive alleles in its opposing environment, and 
+(2) What role does genomic architecture play in the ability for a haplotype to be rebuilt.
+
+\paragraph{Realized genomic architecture}
+Additionally have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
+This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
+differences between populations that have little gene flow between them.
+It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
+mistaken for being causative as they often display the same amount of divergence.
+In general, our study may suggest that diversity scans may not prove as useful in causal loci inference as previously thought.
+It is important to note that these summary statics could prove to be very useful if used beyond just the peaks seen when looked at as a function of genomic position. 
+Many methods, such as advanced machine learning, have proved to be very good at taking in noisy data and parsing out important information \citet{Schrider2018}.
+
+\paragraph{Implications for the ``Transporter"-hypothesis}
+This study was in large, supportive of the ``Transporter"-hypothesis suggested by \citep{schluter2009genetics}.
+Provided there is enough dispersal, we found that repeated colonization and adaptation of freshwater stickleback was able to produce beneficial alleles which could then be 
+supported by the marine stickleback population through the offspring of hybridization events. 
+When new freshwater environments were colonized by the marine stickleback, selection was able to rebuild the freshwater haplotype. 
+Shown in Table~\ref{tab:linkage}, we found that a populations capacity to cary maladaptive alleles remains pretty constant over realistic ranges of dispersal, but low rates of gene flow may 
+"mask" the alleles through linkage which counteracts the fitness disadvantage. 
+This linkage prevents the ability for a sub-population to rapidly select on the beneficial SGV as recombination acts as a bottleneck to evolution in this situation. 
+Additionally, In Figure~\ref{fig:Origin} we found that the genetic basis for local adaptation, when selecting upon SGV, is propagated largely from the initial generation of inhabitants. 
+This contrasts the belief that subsequent migration is a necessity in newly established populations ability to rebuild a haplotype.
+In other words, this finding suggests that with high probability, a randomly chosen subset from ancestral species in the meta-populations like marine stickleback, 
+carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%DO WE NEED THIS?
+\begin{comment}
+\subsection*{thresholds}
+
+We have found that too little migration leads to selection upon new mutations in all subpopulations and lakes alike.
+In contrast, at high migration rates we have seen that migration load limits the ability for species to locally adapt to the selective pressure of their environment.
+This leads us to consider a window of introgression which allows for the transportation of FAA's without migration load.
+
+\subsection*{connect results back to real data?}
+
+\paragraph{The adaptive filter?}
+RAMBLING THOUGHTS HERE
+Since larger effect alleles are more likely to establish,
+be it by mutation or migration,
+repeated colonization of new freshwater habitats will select for larger alleles,
+be it single alleles or haplotypes bound together by an inversion.
+However, these are more strongly selected against in the interstitial time.
+Being recessive would help with this,
+but would also make it more difficult to establish.
+
+\jgg{TODO: effect of smaller than realistic population sizes?}
+
+\jgg{TODO: Talk about the role recombination plays}
+This also brings to light the role of recombination in a system of migration selection balance. 
+On one hand, a lower recombination rate would allow for the freshwater haplotype to remain "intact" 
+within the marine environment, however, without significant introgression the haplotype would quickly get selected against in marine populations. 
+On the other hand a higher recombination rate would, in theory, allow freshwater adapted alleles more longevity 
+as they would hitchhike with marine adapted haplotypes.
+
+\paragraph{Modeling assumptions}
+Our simulations had much smaller population sizes than real populations.
+How might this affect things?
+\end{comment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Theoretical expectations.
+
 The calculation above ignores allele frequencies -- 

### 2019-05-13 Stickleback_Paper.tex
Latest edits & new colored plots

@@
 These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases the smaller freshwater populations have been established ``de novo'' from marine colonists.
-Previous results have found similar phenotypes, between independently evolved freshwater populations, are derived largely from the same standing genetic variants found to be Identical-by-decent
-Geographic isolation preventing direct migration between the freshwater populations strongly suggests that these shared adaptive alleles are transported through hybridization events with the wide-spanning marine population which acts as a medium for most freshwater populations.
+Previous results have found similar phenotypes, between independently evolved freshwater populations, are derived largely from the same standing genetic variants found to be Identical-by-decent.
+Geographic isolation preventing direct migration between the freshwater populations strongly suggests that these shared adaptive alleles are transported through the wide-spanning marine population.
 However, it is still largely unknown how gene flow, recombination, and selection jointly impact standing variation to fuel adaptation.
@@
 More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.
-\jgg{Is this intro too stickleback-y?}
 
@@
 
-The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations (\textbf{cite}). 
-While it has proved to be a useful baseline for understanding the genetic variation we see in species today (\textbf{cite}), this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures (\textbf{cite}).
+The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations \citep{smith_haigh_1974}. 
+While it has proved to be a useful baseline for understanding the genetic variation we see in species today, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures \citep{Lai2152, Schrider2017}.
 In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats.
-Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood (\textbf{cite}).
+Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood.
 A growing number of empirical studies have identified instances of convergent evolution using SGV \citep{nelson2017ancient,Schrider2017,Barrett2008}. 
-However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during colonization and adaptation to new environments (\textbf{cite}).
+However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during colonization and adaptation to new environments \citep{nelson2017ancient}.
 It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via SGV. 
 While empirical studies in the field and lab are essential, modeling specific systems can help us understand the nature of standing genetic variants to fuel local adaptation as well as the genomic signals we can expect to see.
-% of the adaptation genetic dynamics using SGV.
-%\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
-%\wac{I did a little editing but Peter can still totally rewrite. Regardless, it needs to be streamlined a fair bit}
+\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
+\wac{I did a little editing but Peter can still totally rewrite. Regardless, it needs to be streamlined a fair bit}
+\jgg{I think this encapsulates the feel of the paper, but happy for it to be streamlined}
 
@@
 This model organism has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes (\textbf{cite}) and the much more extensive use of SGV than once thought (\textbf{cite}).
-%The ancestral marine form of threespine stickleback has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere (\textbf{cite}).
 While geographic isolation often prevents direct migration between freshwater populations, they frequently evolve similar phenotypes (\textbf{cite}). 
@@
 \section*{Methods}
-\jgg{Any thoughts on putting Methods after results?}
 
@@
 Subsequent mutations at the same locus replace the previous allele.
-%We also used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence} to add neutral mutations across the entire genome at a rate of $10^{-8}$ per locus per generation,  as described in \citet{kelleher2018efficient}.
 
@@
 Each new individual in each habitat has parents from the other habitat with probability $m$ (in which case we call it a ``migrant''),and parents from the same habitat with probability $1-m$.
-The first parent of each nonmigrant individual in the freshwater habitat is chosen from the freshwater habitat proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness.
+The first parent of each non-migrant individual in the freshwater habitat is chosen from the freshwater habitat proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness.
 The offspring then lives in the same lake as the parents.
-Parents for each nonmigrant marine individual are chosen similarly:
+Parents for each non-migrant marine individual are chosen similarly:
 first, a single parent is chosen proportionally to fitness in the marine habitat, and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$.
@@
 Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, and reflected to stay within the habitat range.
-Parents for each freshwater migrant are chosen in the same way as for nonmigrant marine individuals, and are assigned to the lake nearest to the position of the first marine parent.
+Parents for each freshwater migrant are chosen in the same way as for non-migrant marine individuals, and are assigned to the lake nearest to the position of the first marine parent.
 Similarly, parents for each marine migrant are both chosen from the same lake as before, and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.
@@
 
-%%%%%%%%%%%%
 \subsection*{Sharing of freshwater adapted alleles}
@@
 Figure \ref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above, the average individual across the new lakes has nearly the same amount of pre-existing freshwater adapted alleles as individuals across the old lakes.
-As expected, the genetic basis of the freshwater phenotype seems to simplify as migration increases --
-higher rates of migration allow adaptive alleles of larger effect to travel more efficiently through the population, even though they are deleterious in the ocean.
+As expected, the genetic basis of the freshwater phenotype seems to simplify as migration increases -- higher rates of migration allow adaptive alleles of larger effect to travel more efficiently through the population, even though they are deleterious in the ocean.
 
-The numbers in Figure~\ref{fig:MPFAI} suggest that the dramatic increase in speed of local adaptation we observed above occurs because higher gene flow between populations allows sharing of freshwater alleles between populations.
+The numbers in Figure~\ref{fig:MPFAI} strongly suggest that the dramatic increase in speed of local adaptation we observed above occurs because higher gene flow between populations allows sharing of freshwater alleles between populations.
 We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele in each individual in the new lakes, as defined in the Methods. 
@@
 Higher rates of dispersal showed more distinct $F_{ST}$ peaks over polymorphic loci, while ``background'' levels of $F_{ST}$ increase as gene flow decreases, swamping out this signal until the regions under selection are indistinguishable.
-This is likely the contribution of two separate forces of natural selection: first, stronger genetic drift with less migration leads to higher background $F_{ST}$, and second, greater sharing of adaptive alleles providing a shared signal across populations.
+This is likely the contribution of two separate forces of natural selection:
+first, stronger genetic drift with less migration leads to higher background $F_{ST}$, and second, greater sharing of adaptive alleles providing a shared signal across populations.
 
@@
 But how many of these peaks are actually underlying trait differences that form the basis for local adaptation? 
-To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates
-that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.
+To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.
 
 \plr{need to say what true and false positive windows are} 
+\jgg{? like the pink lines in Figure \ref{fig:Fst}}
 
@@
 This may imply that genome scans for local adaptation based purely on measures of differentiation may not be as reliable as previously thought.
-To summarize, our study find that with high rates of dispersal between population, $F_{st}$ peaks may be a good indication of causal loci.
-However, this is a one way implication, meaning even with high dispersal, not all causal loci will appear as $F_{st}$ peaks.
+To summarize, our study find that with high rates of dispersal between population, $F_{ST}$ peaks may be a good indication of causal loci.
+However, this is a one way implication, meaning even with high dispersal, not all causal loci will appear as $F_{ST}$ peaks.
 
@@
 
-%In regions of the genome underlying individual trait value, we observed 
-%Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$. 
-%Knowing exactly which mutations effect phenotype in our simulations, we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome.
-%In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power. 
-%This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.
+This study shows introgression is beneficial for inferring causative loci from divergence ($F_{ST}$) along the genome. 
+This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
+differences between populations that have little gene flow between them.
+It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
+mistaken for being causative as they often display the same amount of divergence.
+In general, our study may suggest that diversity scans may not prove as useful in causal loci inference as previously thought.
+It is important to note that these summary statics could prove to be very useful if used beyond just the peaks seen when looked at as a function of genomic position.
 
@@
 We have thus far found that the speed of adaptation depends strongly on the degree to which alleles can be shared between populations. 
-However, the large increase in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$, seen in Figure \ref{fig:TimeToAdaptation}, remains mysterious.
-Figure~\ref{fig:MPFAI}A shows that with $M = 0.1$ migrants per lake per generation, there is substantial sharing of alleles between populations, and yet at this rate new lakes take around 2,000 generations to adapt.
+However, the \textit{origin} of the alleles underlying the phenotype is still unknown. 
+What is the nature of selection on standing genetic variants? 
+In our simulations migration from marine individuals into the new lakes continues throughout introduction and adaptation: but is consistent influx of SGV a necessity for rapid adaptation of the population?
+To answer this, we traced the genealogy of \textit{all} adaptive alleles in \textit{all} individuals from the introduced lakes, after local adaptation.
+Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back the original generation of inhabitants in the lake as seen in Figure \ref{fig:Origin}.
+This suggests that given enough SGV, any reasonable subset of the ancestral population has the potential to for rapid adaptation without the need post-colonization hybridization events.
+
+If effectively capturing standing genetics variants is the key to rapid adaptation, as we have presented thus far, why is there a large difference in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$ seen in Figure \ref{fig:TimeToAdaptation}.
+%Figure~\ref{fig:Origin} shows both migration levels being able to effectively capture and propagate the standing genetic variants, and
+%Figure~\ref{fig:MPFAI}A shows that with $M = 0.1$ migrants per lake per generation, there is substantial sharing of alleles between populations.
+%So why does it take $M = 0.1$ scenario nearly 2,000 generations to adapt. 
 This is surprising because we see a similar quantity of allele sharing at  $M = 1.0$, but much more rapid adaptation ($\approx 60$ generations).
-A reasonable explanation for this 33-fold discrepancy is that new lakes must wait for an additional influx of freshwater alleles through migration beyond what is present in the initial generation.
+%A reasonable explanation for this 33-fold discrepancy is that new lakes must wait for an additional influx of freshwater alleles through migration beyond what is present in the initial generation.
 However, Figure~\ref{fig:Origin} shows this is not the case: at both $m = 0.1$ and $m = 1.0$, the vast majority of alleles responsible for local adaptation came from individuals in the original generation (``captured'').
@@
 
-An alternative explanation is that at lower rates of gene flow, freshwater alleles present as standing variation in the marine habitat are more tightly linked to marine alleles.
+A possible explanation is that at lower rates of gene flow, freshwater alleles present as standing variation in the marine habitat are more tightly linked to marine alleles.
 This implies the adaptation in the new lakes must wait for recombination to separate freshwater adapted alleles from nearby marine adapted alleles.
@@
 Note that some of these haplotypes will likely be lost to drift -- indeed, if they did all fix, then populations at $M=0.1$ would adapt much more quickly.
-However, the comparison still provides a good comparison.
+However, this calculation still provides contrast and a reasonable explanation for the difference in time to adaptation.
 
@@
 
-% These calculations depend on there being no bottleneck in colonization of the lake; if there is a bottleneck, then an additional factor must be added.
-At what point do we expect new mutation to be more important than migration for adaptation?
-By the calculations above, if $M \ge 10$, we expect initial diversity in a lake to be sufficient  for adaptation, corresponding to our third-highest migration rate.
-If this does not happen, then we expect adaptation to take a multiple of $1/m$ generations;
-% with our values, $1/m$ ranges from 20,000 to 20 generations.
-Above we estimated that adaptive alleles due to new mutation fix locally about every 2,000 generations, suggesting that at our second-lowest migration rate (where $1/m = 2,000$), the two contributions of migration and new mutation are roughly equal.
-% This is in fact what we see: in Figure \ref{fig:MPFAI}, we see that at the second migration rate, alleles start to be shared between lakes, while by the third migration rate, they are almost entirely shared.
-
-\section*{Conclution}
+\section*{Conclusion}
 
@@
 Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.
-We find that, after local adaptation in new lakes, the majority of adaptive variant's genealogy can be traced back to the original generation as opposed to post-colonization migrant alleles.
-Additionally, quantifying the inference power of $F_{ST}$ genomic scans has revealed that higher rates of gene flow are generally more useful for identifying 
+We find that, after local adaptation in new lakes, the majority of adaptive variant's genealogy can be traced back to the original generation of inhabitants as opposed to post-colonization migrant alleles.
+In other words, this finding suggests that with high probability, a randomly chosen subset from ancestral species in the meta-populations like marine stickleback, 
+carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
+Additionally, quantifying the inference power of $F_{ST}$ genomic scans has revealed that higher rates of gene flow are generally more useful for identifying regions of the genome with locally adaptive alleles.
 Perhaps the most notable finding is that the large marine population was able to harbor and distribute deleterious alleles to the surrounding freshwater populations with only a few migrants per population, per generation.
+
+This study was in large, supportive of the ``Transporter"-hypothesis suggested by \citep{schluter2009genetics}.
+Provided there is enough dispersal, we found that repeated colonization and adaptation of freshwater stickleback was able to produce beneficial alleles which could then be 
+supported by the marine stickleback population through the offspring of hybridization events. 
+Shown in Table~\ref{tab:linkage}, we found that a populations capacity to cary maladaptive alleles remains pretty constant over realistic ranges of dispersal, but low rates of gene flow may 
+"mask" the alleles through linkage which counteracts the fitness disadvantage. 
+This linkage prevents the ability for a sub-population to rapidly select on the beneficial SGV as recombination acts as a bottleneck to evolution in this situation.
 
@@
           \caption{ 
-        Statistical Power and False Positives as a function of $F_{st}$ threshold. 
-        Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
-        False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
-            \textbf{
-                TODO: fix this caption; x-axis label ($F_{ST}$).}
+        Statistical Power and False Positives as a function of $F_{ST}$ threshold. 
+        Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected.
+        False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{ST}$ threshold.
+          \label{fig:Power_FP}
           }
-          \label{fig:Power_FP}
     \end{center}
@@
           \caption{ 
-        Statistical Power and False Positives as a function of $F_{st}$ threshold. 
-        Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
-        False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
-            \textbf{
-                TODO: fix this caption; x-axis label ($F_{ST}$).}
+        Statistical Power and False Positives as a function of $F_{ST}$ threshold. 
+        Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected.
+        False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{ST}$ threshold.
          	}

### 2019-05-28 Stickleback_Paper.tex
som minor edits

@@
 
-\title{A few stickleback can transport adaptive alleles to new lakes}
+\title{A few stickleback suffice for the transport of alleles to new lakes}
 \author{Jared Galloway, William A. Cresko, and Peter Ralph}
 \maketitle
-\jgg{To me, this title makes it sound like only a "select few" stickleback have the ability to transport, no?}
+%\jgg{To me, this title makes it sound like only a "select few" stickleback have the ability to transport, no?}
 
@@
 Geographic isolation preventing direct migration between the freshwater populations strongly suggests that these shared adaptive alleles are transported through the wide-spanning marine population.
-However, it is still largely unknown how gene flow, recombination, and selection jointly impact standing variation to fuel adaptation.
-Here we use genealogical simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback, and to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and re-use of standing genetic variation.
+However it is still largely unknown how gene flow, recombination, and selection jointly impact standing variation to fuel adaptation.
+Here we use genealogical simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback.
+We aim to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and re-use of standing genetic variation. 
 In our simulations we find that repeated adaptation uses a shared set of alleles that are maintained at low frequency by migration-selection balance in oceanic populations. 
 This process occurs over a realistic range of intermediate levels of gene flow that match previous empirical population genomic studies in stickleback. 
-Examining these simulations more deeply reveals that lower levels of gene flow leads to slow, independent adaptation to district habitats, whereas higher levels gene flow leads to significant mutation load, but an increased probability of successful population genomic scans for locally adapted alleles.
+Examining these simulations more deeply reveals that lower levels of gene flow leads to slow, independent adaptation to district habitats, whereas higher levels gene flow leads to significant mutation load -- but an increased probability of successful population genomic scans for locally adapted alleles.
 Surprisingly, we find that the genealogical origins of most freshwater adapted alleles can be traced back to the original generation of marine individuals that colonized the lakes, as opposed to subsequent migrants.
-These simulations provide deeper context for existing studies of stickleback evolutionary genomics, and provide guidance for future empirical studies in this model.
+These simulations provide deeper context for existing studies of stickleback evolutionary genomics, and guidance for future empirical studies in this model.
 More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.
@@
 While empirical studies in the field and lab are essential, modeling specific systems can help us understand the nature of standing genetic variants to fuel local adaptation as well as the genomic signals we can expect to see.
-\plr{I like the spirit of the first paragraph, but want to totally rewrite it}
-\wac{I did a little editing but Peter can still totally rewrite. Regardless, it needs to be streamlined a fair bit}
-\jgg{I think this encapsulates the feel of the paper, but happy for it to be streamlined}
 
@@
 How many differentially selected alleles exist, how many are used at any one time, and how are they arranged within the genome? 
-A curious natural history observation undergirding many of these questions is the fact that some newly formed freshwater habitats, such as the ponds on Middleton Island, are quite small and presumably the number of initial marine migrants is few.
-Here we use forward simulations implemented in SLiM incorporating linkage and realistic length genomes to model the stickleback metapopulation and address these questions \citep{haller2017slim,haller2018slim3}.
+A curious natural history observation underlying many of these questions is the fact that some newly formed freshwater habitats, such as the ponds on Middleton Island, are quite small and presumably the number of initial marine migrants is few.
+
+Here, we use forward simulations implemented in SLiM incorporating linkage and realistic length genomes to model the stickleback metapopulation and address these questions \citep{haller2017slim,haller2018slim3}.
 We ask how variation in amount of gene flow affects the genetic architecture of local adaptation to newly created freshwater ponds.
@@
 
-\plr{need to say what true and false positive windows are} 
-\jgg{? like the pink lines in Figure \ref{fig:Fst}}
-
 What we observe most plainly in this graph is that $M = 1.0$ and $M = 10.0$ migrants per lake, per generation are the only rates of dispersal which provide reliable true negatives when observing peaks.
@@
 
-\plr{cite \citet{yeaman2011genetic}?}
+%\plr{cite \citet{yeaman2011genetic}?}
 
@@
 \jgg{This feels awkward, help please.}
-%This is evidence that standing genetic variation from gene flow likely plays a key role in local adaptation for many sub-species.
 

### 2019-05-29 Stickleback_Paper.tex
Bill added citations and some edits to the main text file

@@
 %\jgg{To me, this title makes it sound like only a "select few" stickleback have the ability to transport, no?}
+%\wac{How about 'A small number of stickleback ... }
 
@@
 
-Threespine stickleback provide a striking example of local adaptation to divergent habitats while still being connected by recurrent gene flow. 
+Threespine stickleback populations provide a striking example of local adaptation to divergent habitats while still being connected by recurrent gene flow. 
 These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases the smaller freshwater populations have been established ``de novo'' from marine colonists.
 Previous results have found similar phenotypes, between independently evolved freshwater populations, are derived largely from the same standing genetic variants found to be Identical-by-decent.
-Geographic isolation preventing direct migration between the freshwater populations strongly suggests that these shared adaptive alleles are transported through the wide-spanning marine population.
+Geographic isolation preventing direct migration between the freshwater populations strongly suggests that these shared adaptive alleles are transported through the widely distributed marine population.
 However it is still largely unknown how gene flow, recombination, and selection jointly impact standing variation to fuel adaptation.
@@
 
-The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations \citep{smith_haigh_1974}. 
-While it has proved to be a useful baseline for understanding the genetic variation we see in species today, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures \citep{Lai2152, Schrider2017}.
-In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats.
-Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood.
-A growing number of empirical studies have identified instances of convergent evolution using SGV \citep{nelson2017ancient,Schrider2017,Barrett2008}. 
-However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during colonization and adaptation to new environments \citep{nelson2017ancient}.
+The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations \textbf{smith_haigh_1974 Endler_1977 Orr_2005}.
+While it has proved to be a useful baseline for understanding the genetic variation we see in species today, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures \textbf{Lai2152, Schrider2017}.
+In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats \textbf{Hermisson_2005, Barrett_2008, Peter_2012, Burke_2014}.
+Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood \textbf{Charlesworth_1997, Charlesworth_2003, Leonormand_2002, Nosil_2009, Flaxman_2013, Samuk 2017}.
+A growing number of empirical studies have identified instances of convergent evolution using SGV \citep{Schrider2017,Barrett2008}, \textbf{Burke 2014, Nelson_2018, Nelson_2019, Bassham_2018}. 
+However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during colonization and adaptation to new environments \citep{nelson2017ancient}\textbf{Pritchard_2010,Yeaman_2011, Messer_2013, Shrider2017}.
 It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via SGV. 
@@
 
-The ancestral marine form of threespine stickleback fish has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere (\textbf{cite}).
-This model organism has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes (\textbf{cite}) and the much more extensive use of SGV than once thought (\textbf{cite}).
-While geographic isolation often prevents direct migration between freshwater populations, they frequently evolve similar phenotypes (\textbf{cite}). 
-The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years (\textbf{cite}), independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations (\textbf{cite}).
-For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska (\textbf{cite}). 
-Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years (\textbf{cite}).
-Amazingly, nearly a quarter of the genome diverged significantly in this same time but - with very little change in the rest of the genome - in a pattern that mirrored previous genomic studies of freshwater stickleback populations that have been geographically separated for thousands of years (\textbf{cite}).
+The ancestral marine form of threespine stickleback fish has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere \textbf{Bell_1994, Bell_1994, Thompson_1997, Cresko_2007}.
+This model organism has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes \textbf{Hohenlohe_2010, Jones_2012, Terekhanova_2014, Marques_2016} and the much more extensive use of SGV than once thought \textbf{Schluter_2009, Roesti_2014,Nelson_2018, Bassham_2018, Nelson_2019}.
+While geographic isolation often prevents direct migration between freshwater populations, they frequently evolve similar phenotypes \textbf{Cresko_2004, Colosimo_2004, Shapiro_2005}. 
+The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years \textbf{Jones_2012, Nelson_2018}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \textbf{Terekhanova_2014, Lescak_2015, Bassham_2018}.
+For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska \textbf{Pflaker_1969, Pflaker_1978}. 
+Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years (\textbf{Lescak_2015, Bassham_2018}).
+Amazingly, nearly a quarter of the genome diverged significantly in this same time but - with very little change in the rest of the genome - in a pattern that mirrored previous genomic studies of freshwater stickleback populations that have been geographically separated for thousands of years \textbf{Hohenlohe_2010, Bassham_2018}.
 
 But how can evolution occur at such a rapid pace?
-Waiting for new mutations to arise in each lake or pond would take much longer than the decades since the Alaskan earthquake (\textbf{cite}). 
+Waiting for new mutations to arise in each lake or pond would take much longer than the decades since the Alaskan earthquake. 
 Even more improbable is having divergence cluster in such similar genomic regions across independent populations. 
@@
 In 2009 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter''- hypothesis to describe the process of alleles beneficial in freshwater environment being maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation to new freshwater habitats.
-The first clear example of the global reuse of SGV in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates (\textbf{cite}). 
-While the low lateral plate version of this gene arose millions of years ago, it is found in much younger freshwater ponds around the Northern Hemisphere (\textbf{cite}). 
+The first clear example of the global reuse of SGV in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates \textbf{Colosimo_2004, Cresko_2004, Colosimo_2005}.
+While the low lateral plate version of this gene arose millions of years ago, it is found in much younger freshwater ponds around the Northern Hemisphere \textbf{O'Brown_2015}. 
 More recently, genome-wide haplotype analyses have provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.

### 2019-06-06 Stickleback_Paper.tex
citations finished

@@
 \maketitle
-%\jgg{To me, this title makes it sound like only a "select few" stickleback have the ability to transport, no?}
-%\wac{How about 'A small number of stickleback ... }
 
@@
 
-The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations \textbf{smith_haigh_1974 Endler_1977 Orr_2005}.
-While it has proved to be a useful baseline for understanding the genetic variation we see in species today, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures \textbf{Lai2152, Schrider2017}.
-In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats \textbf{Hermisson_2005, Barrett_2008, Peter_2012, Burke_2014}.
-Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood \textbf{Charlesworth_1997, Charlesworth_2003, Leonormand_2002, Nosil_2009, Flaxman_2013, Samuk 2017}.
-A growing number of empirical studies have identified instances of convergent evolution using SGV \citep{Schrider2017,Barrett2008}, \textbf{Burke 2014, Nelson_2018, Nelson_2019, Bassham_2018}. 
-However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during colonization and adaptation to new environments \citep{nelson2017ancient}\textbf{Pritchard_2010,Yeaman_2011, Messer_2013, Shrider2017}.
+The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations 
+\citep{smith_haigh_1974, Endler_1977, Orr_2005}.
+While it has proved to be a useful baseline for understanding the genetic variation we see in species today, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures
+\citep{Lai2152, Schrider2017}. 
+In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats
+\citep{Hermisson_2008, Barrett2008}. 
+Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood 
+\citep{Charlesworth_1997, Charlesworth_2003, Nosil_2009, Flaxman_2013, Samuk_2017}. 
+A growing number of empirical studies have identified instances of convergent evolution using SGV 
+\citep{Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. 
+However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during colonization and adaptation to new environments
+\citep{nelson2017ancient, Pritchard_2010, yeaman2011genetic, Schrider2017}. 
 It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via SGV. 
@@
 
-The ancestral marine form of threespine stickleback fish has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere \textbf{Bell_1994, Bell_1994, Thompson_1997, Cresko_2007}.
-This model organism has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes \textbf{Hohenlohe_2010, Jones_2012, Terekhanova_2014, Marques_2016} and the much more extensive use of SGV than once thought \textbf{Schluter_2009, Roesti_2014,Nelson_2018, Bassham_2018, Nelson_2019}.
-While geographic isolation often prevents direct migration between freshwater populations, they frequently evolve similar phenotypes \textbf{Cresko_2004, Colosimo_2004, Shapiro_2005}. 
-The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years \textbf{Jones_2012, Nelson_2018}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \textbf{Terekhanova_2014, Lescak_2015, Bassham_2018}.
-For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska \textbf{Pflaker_1969, Pflaker_1978}. 
-Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years (\textbf{Lescak_2015, Bassham_2018}).
-Amazingly, nearly a quarter of the genome diverged significantly in this same time but - with very little change in the rest of the genome - in a pattern that mirrored previous genomic studies of freshwater stickleback populations that have been geographically separated for thousands of years \textbf{Hohenlohe_2010, Bassham_2018}.
+The ancestral marine form of threespine stickleback fish has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere 
+\citep{Bell_1994, Thompson_1997, Cresko_2007}. 
+This model organism has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes 
+and the much more extensive use of SGV than once thought 
+\citep{schluter2009genetics, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
+While geographic isolation often prevents direct migration between freshwater populations, they frequently evolve similar phenotypes 
+\citep{Cresko_2004, Colosimo_2004}. 
+The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years 
+\citep{Nelson_2018}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \citep{Terekhanova_2014, Lescak2015evolution, Bassham_2018}.
+For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska. 
+Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years
+\citep{Lescak2015evolution, Bassham_2018}. 
+Amazingly, nearly a quarter of the genome diverged significantly in this same time but - with very little change in the rest of the genome - in a pattern that mirrored previous genomic studies of freshwater stickleback populations that have been geographically separated for thousands of years
+\citep{Hohenlohe_2010, Bassham_2018}.
 
@@
 In 2009 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter''- hypothesis to describe the process of alleles beneficial in freshwater environment being maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation to new freshwater habitats.
-The first clear example of the global reuse of SGV in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates \textbf{Colosimo_2004, Cresko_2004, Colosimo_2005}.
-While the low lateral plate version of this gene arose millions of years ago, it is found in much younger freshwater ponds around the Northern Hemisphere \textbf{O'Brown_2015}. 
+The first clear example of the global reuse of SGV in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates 
+\citep{Colosimo_2004, Cresko_2004, Colosimo_2005}. 
+While the low lateral plate version of this gene arose millions of years ago, it is found in much younger freshwater ponds around the Northern Hemisphere 
+\citep{O'Brian_2015}. 
 More recently, genome-wide haplotype analyses have provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
@@
 \paragraph{Recording genealogical history}
+
 We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence,Kelleher2016} to output the genealogical history of all individuals at the time of introduction of new lakes, at the time of adaptation, and at the end of the simulation.

### 2019-06-16 Stickleback_Paper.tex
I went through and did a good edit on the paper

@@
 These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases the smaller freshwater populations have been established ``de novo'' from marine colonists.
-Previous results have found similar phenotypes, between independently evolved freshwater populations, are derived largely from the same standing genetic variants found to be Identical-by-decent.
+Previous results have found similar phenotypes between independently evolved freshwater populations are derived largely from the same standing genetic variants found to be Identical-by-decent.
 Geographic isolation preventing direct migration between the freshwater populations strongly suggests that these shared adaptive alleles are transported through the widely distributed marine population.
@@
 This process occurs over a realistic range of intermediate levels of gene flow that match previous empirical population genomic studies in stickleback. 
-Examining these simulations more deeply reveals that lower levels of gene flow leads to slow, independent adaptation to district habitats, whereas higher levels gene flow leads to significant mutation load -- but an increased probability of successful population genomic scans for locally adapted alleles.
+Examining these simulations more deeply reveals that lower levels of gene flow leads to slow, independent adaptation to different habitats, whereas higher levels gene flow leads to significant mutation load -- but an increased probability of successful population genomic scans for locally adapted alleles.
 Surprisingly, we find that the genealogical origins of most freshwater adapted alleles can be traced back to the original generation of marine individuals that colonized the lakes, as opposed to subsequent migrants.
@@
 \citep{Charlesworth_1997, Charlesworth_2003, Nosil_2009, Flaxman_2013, Samuk_2017}. 
-A growing number of empirical studies have identified instances of convergent evolution using SGV 
+Concurrently, a growing number of empirical studies have identified instances of convergent evolution using SGV 
 \citep{Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. 
@@
 \citep{schluter2009genetics, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
-While geographic isolation often prevents direct migration between freshwater populations, they frequently evolve similar phenotypes 
+While geographic isolation often prevents direct migration between freshwater populations, stickleback in them frequently evolve similar phenotypes 
 \citep{Cresko_2004, Colosimo_2004}. 
@@
 \citep{Lescak2015evolution, Bassham_2018}. 
-Amazingly, nearly a quarter of the genome diverged significantly in this same time but - with very little change in the rest of the genome - in a pattern that mirrored previous genomic studies of freshwater stickleback populations that have been geographically separated for thousands of years
+Amazingly, despite very little change in a majority of the genome, nearly a quarter of it diverged significantly in this same time in a pattern that mirrored previous genomic studies of freshwater stickleback populations that have been geographically separated for thousands of years
 \citep{Hohenlohe_2010, Bassham_2018}.
@@
 
-Here, we use forward simulations implemented in SLiM incorporating linkage and realistic length genomes to model the stickleback metapopulation and address these questions \citep{haller2017slim,haller2018slim3}.
+Here, we use forward simulations implemented in SLiM that incorporate linkage and realistic length genomes to model the stickleback metapopulation and address these questions \citep{haller2017slim,haller2018slim3}.
 We ask how variation in amount of gene flow affects the genetic architecture of local adaptation to newly created freshwater ponds.
@@
 In addition, we can document how these adaptive alleles are distributed across the genomes, and as a consequence determine how this may affect the efficacy of genome scans for between-habitat differentiation.
-We find that only a few stickleback migrants per lake per generation are sufficient to maintain SGV in the ocean, allowing new freshwater populations can adapt in only tens of generations. 
+We find that only a few stickleback migrants per lake per generation are sufficient to maintain SGV in the ocean, allowing new freshwater populations to adapt in only tens of generations. 
 We also find that continued gene flow of marine individuals into recently established lakes is not necessary for adaptation as the majority of alleles are present in the initial marine colonizers. 
@@
 \jgg{I don't know where we should talk about our small population sizes ... Is this even worth mentioning here?}
+\wac{I don't think it's necessary because the census population sizes of stickleback actually due vary widely, so we're not that far off.
+If anything you could say that we focused on smaller populations because of our questions about how many migrants are needed}
 
@@
 
-Our simulations have two habitat types -- Marine and Freshwater -- defined by their selective pressures, each with 5,000 diploid individuals. 
-The arrangement of these habitats, depicted in Figure~\ref{fig:Geo} roughly models a set of freshwater habitats along a stretch of coastline. 
+Our simulations include two habitat types -- Marine and Freshwater -- defined by the nature of their selective pressures, each with 5,000 diploid individuals. 
+The arrangement of these habitats, depicted in Figure~\ref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. 
 The marine habitat is a continuous, one-dimensional range of length 25 units, while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''), each connected to the marine habitat at regularly spaced intervals (positions $i - 1/2$ for $1 \le i \le 25$).
@@
 Divergent selection is mediated by a single quantitative trait with different optima in marine and freshwater habitats. 
-This situation roughly models the cumulative effect of the various phenotypes thought to be under divergent selection between the habitats, such as armor morphology, body size, craniofacial variation and opercle shape.
+This situation roughly models the cumulative effect of the various phenotypes thought to be under divergent selection between the habitats, such as armor morphology, body size, craniofacial variation and opercle shape (among many other traits).
 The optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively, and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$ is determined by a Gaussian kernel with standard deviation 15, i.e.,
@@
 
-In short, the difference between an individual's trait value and the optimum determines that individual's fitness. 
+Thus, the difference between an individual's trait value and the optimum determines that individual's fitness. 
 We chose the difference between optima and strength of stabilizing selection in each habitat so that (a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other, and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.
@@
             a single, continuous, one-dimensional marine habitat (blue)
-            is coupled to randomly mating ``lakes'' at discrete locations with arrows representing migration patterns.
+            is coupled to ``lakes'', in which stickleback randomly mate, at discrete locations with arrows representing migration patterns.
             After an initial period of 100K generations with 25 lakes,
@@
 As depicted in Figure~\ref{fig:Geo}, dispersal occurs both locally along the coastline in the marine habitat as well as between the marine habitat and the lakes,
-with a lake--ocean migration rate denoted $m$.All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
-Each new individual in each habitat has parents from the other habitat with probability $m$ (in which case we call it a ``migrant''),and parents from the same habitat with probability $1-m$.
+with a lake--ocean migration rate denoted $m$.
+No migration occurs among the freshwater populations.
+All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation.
+Each new individual in each habitat has parents from the other habitat with probability $m$ (in which case we call it a ``migrant''), and parents from the same habitat with probability $1-m$.
 The first parent of each non-migrant individual in the freshwater habitat is chosen from the freshwater habitat proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness.
@@
 
-\paragraph{New lakes}
+\paragraph{Colonization of newly formed lakes}
 
 To study how marine-derived populations adapt after colonizing newly appearing freshwater habitats, we introduce a new set of 25 lakes after 100,000 generations. 
-These new lakes are populated with marine individuals to emulate a freshwater lake being colonized by oceanic stickleback. 
+These new lakes are populated with marine individuals to emulate a freshwater lake being colonized by oceanic stickleback that had previously been exchanging alleles with older freshwater populations.
 This creates two \emph{sets} of lakes, which along with the marine population have a total of 15,000 individuals.
@@
 
+\wac{In general I suggest having the finding or statement at the beginning of the sentence with a reference to the figure or statistic at the end of the sentence.
+Notice that in many cases below (e.g.
+first sentence of the first paragraph in the 'Local Adaptation' section) the sentence starts with the Figure (e.g.
+'As shown in Figure 2').
+You can just start with the declarative statement (e.g.
+"We found that local adaptation occurred in all simulations ....
+was composed of migrants (Fig.
+2)")}
+
 To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$. 
@@
 At very low rates of gene flow, each new lake's population adapted almost completely independently through \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations). 
-At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles \citet{bolnick2007natural}. 
+At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles \citep{bolnick2007natural}. 
 Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes. 
@@
 \subsection*{Local Adaptation: differentiation with gene flow}
+\wac{Can we change the header to be a statement of finding? e.g. Local adaptation occurs at almost all levels of gene flow}
 
@@
 Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat.
-Establishment of new alleles in the lakes is visible in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations.
-At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites(about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats.
+Establishment of new alleles in the lakes is visible in the first few thousand generations as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations (Fig.
+\ref{fig:phenotype_time}).
+At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites (about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats.
 As expected, increasing migration rate decreased differentiation between habitats: as seen in Figure~\ref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.
@@
 \subsection*{Sharing of freshwater adapted alleles}
+\wac{same comment as above, is there a declarative statement of finding? e.g. Freshwater adapted alleles are shared extensively and primarily via initial migrants}
 
@@
 
-The numbers in Figure~\ref{fig:MPFAI} strongly suggest that the dramatic increase in speed of local adaptation we observed above occurs because higher gene flow between populations allows sharing of freshwater alleles between populations.
+The numbers in Figure~\ref{fig:MPFAI} strongly suggest that the dramatic increase in speed of local adaptation we observed above occurs because higher gene flow allows sharing of freshwater alleles among populations.
 We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele in each individual in the new lakes, as defined in the Methods. 
@@
 
-While increased migration allows sharing of adaptive alleles between lakes, at $m = 10.0$ the constant influx of alleles between the habitats creates substantial migration load.
+While increased migration allows sharing of adaptive alleles among lakes, at $m = 10.0$ the constant influx of alleles between the habitats creates substantial migration load.
 The rate at which migration load becomes substantial, 10 migrants per lake per generation, only replaces 5\% of each population each generation with migrants from the other habitat, but this is sufficient to shift the mean trait values to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
@@
 
-\subsection*{Scans for selection}
+\subsection*{The efficacy of genomic scans of selection depends on gene flow}
+\wac{just made a suggested change}
 
@@
 This may imply that genome scans for local adaptation based purely on measures of differentiation may not be as reliable as previously thought.
-To summarize, our study find that with high rates of dispersal between population, $F_{ST}$ peaks may be a good indication of causal loci.
-However, this is a one way implication, meaning even with high dispersal, not all causal loci will appear as $F_{ST}$ peaks.
-
+We therefore find that with high rates of dispersal between population, $F_{ST}$ peaks may be a good indication of causal loci that are identified, but that not all causal loci will appear as $F_{ST}$ peaks.
 It's important to note that our study in Figures~\ref{fig:Fst} and \ref{fig:Power_FP} pooled together all lakes and all marine individuals for the calculations. 
-For comparison, Figures~\ref{fig:fst_5lakes} \& \ref{fig:Power_FP_5lakes} shows the same scans but calculated only on between individuals in the marine with spacial location $\le 5.0$ and the individuals in the respective 5 lakes.
+For comparison, Figures~\ref{fig:fst_5lakes} \& \ref{fig:Power_FP_5lakes} shows the same scans but calculated only between individuals in the marine with spacial location $\le 5.0$ and the individuals in the respective 5 lakes.
 These graphs show even less power and True Positives only at $ M = 10.0$ migrants per lake, per generation
@@
 This study shows introgression is beneficial for inferring causative loci from divergence ($F_{ST}$) along the genome. 
-This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
+This is generally because noise of neutral divergence can appear causative when genetic drift causes more 
 differences between populations that have little gene flow between them.
@@
 \subsection*{Origin of introduced freshwater adaptive alleles}
+\wac{same thing here - a statement of the major finding of this section?}
 
@@
 Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back the original generation of inhabitants in the lake as seen in Figure \ref{fig:Origin}.
-This suggests that given enough SGV, any reasonable subset of the ancestral population has the potential to for rapid adaptation without the need post-colonization hybridization events.
+This suggests that given enough SGV, any reasonable subset of the ancestral population has the potential for rapid adaptation without the need post-colonization hybridization events.
 
@@
 A possible explanation is that at lower rates of gene flow, freshwater alleles present as standing variation in the marine habitat are more tightly linked to marine alleles.
-This implies the adaptation in the new lakes must wait for recombination to separate freshwater adapted alleles from nearby marine adapted alleles.
+This implies the adaptation in the new lakes must wait for recombination to separate freshwater and marine adapted alleles.
 This might be expected at low dispersal rates since freshwater alleles may need to be ``masked'' by nearby compensatory alleles to remain in the ocean for long periods of time.
@@
 
-\subsection*{Theoretical expectations}
+\subsection*{Alignment with theoretical expectations}
+\wac{This subsection is a little incongruous with the others sections with respect to how it is written.
+Seems that the bold headers for the paragraphs was residual from notes?
+Also, this section seems to be  in a different voice from the rest of the Results/Discussion section.
+I think it's worth a bit of time trying to hone it more.}
 
@@
 Suppose a new allele enters a lake, either by migration or mutation.
-If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ -- then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical} (assuming Poisson reproduction, as we roughly have here).
-If the current population all differ from the optimum trait by $z$, and the allele has effect size $-u$ in heterozygotes,  then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
+If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ it is then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical} (assuming Poisson reproduction, as we roughly have here).
+If the current populations all differ from the optimum trait by $z$, and the allele has effect size $-u$ in heterozygotes,  then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
 This tells us two things: (1) the rate of adaptation decreases as the population approaches the optimum, and (2) larger mutations (in the right direction) are more likely to fix.
@@
 % This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.
-However, this calculation treats each allele independently; in practice we found that standing freshwater variation in the ocean were masked by linkage to compensatory marine variants.
+However, this calculation treats each allele independently; in practice we found that standing freshwater variation in the ocean was masked by linkage to compensatory marine variants.
 
@@
 In this paper, we have described a realistic simulation model of a coastal meta-population to observe the effects of selection on genetic variation across a wide range of gene flow rates.
-We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
+We have shown that over a realistic range of parameter values historical introgression is able to reproduce rapid and parallel adaptation similar to that documented in natural stickleback populations, such as Middleton island (cites), and supporting the ``Transporter"-hypothesis suggested by \citet{schluter2009genetics}.
 Selection is able to rebuild the freshwater haplotype at a rapid pace (in tens of generations) from marine populations as a medium between all freshwater populations. 
-Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.
-We find that, after local adaptation in new lakes, the majority of adaptive variant's genealogy can be traced back to the original generation of inhabitants as opposed to post-colonization migrant alleles.
-In other words, this finding suggests that with high probability, a randomly chosen subset from ancestral species in the meta-populations like marine stickleback, 
-carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
-Additionally, quantifying the inference power of $F_{ST}$ genomic scans has revealed that higher rates of gene flow are generally more useful for identifying regions of the genome with locally adaptive alleles.
-Perhaps the most notable finding is that the large marine population was able to harbor and distribute deleterious alleles to the surrounding freshwater populations with only a few migrants per population, per generation.
-
-This study was in large, supportive of the ``Transporter"-hypothesis suggested by \citep{schluter2009genetics}.
-Provided there is enough dispersal, we found that repeated colonization and adaptation of freshwater stickleback was able to produce beneficial alleles which could then be 
-supported by the marine stickleback population through the offspring of hybridization events. 
-Shown in Table~\ref{tab:linkage}, we found that a populations capacity to cary maladaptive alleles remains pretty constant over realistic ranges of dispersal, but low rates of gene flow may 
-"mask" the alleles through linkage which counteracts the fitness disadvantage. 
-This linkage prevents the ability for a sub-population to rapidly select on the beneficial SGV as recombination acts as a bottleneck to evolution in this situation.
-
-\jgg{This feels awkward, help please.}
+Almost all rates of migration contributed to the efficiency of local adaptation except the highest due to migration load.
+A finding of practical consequence is that the efficacy of $F_{ST}$ based genomic scans of selection varies significantly with migration rate.
+A surprising new finding of our simulations is that the majority of adaptive variant's genealogy can be traced back to the original generation of inhabitants as opposed to post-colonization migrant alleles, and thus a randomly chosen subset of marine stickleback carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
+Thus, the large marine population was able to harbor and distribute deleterious alleles to the surrounding freshwater populations with only a few migrants per population, per generation.
+Our simulation findings significantly extend previous empirical and theoretical population genomic studies of stickleback to make testable predictions, and provide a framework for similar studies in other organisms to determine the generality of our findings.
 

### 2019-07-04 Stickleback_Paper.tex
peter edits

@@
 
-Threespine stickleback populations provide a striking example of local adaptation to divergent habitats while still being connected by recurrent gene flow. 
+Threespine stickleback populations provide a striking example of local adaptation to divergent habitats 
+in populations that are connected by recurrent gene flow. 
 These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases the smaller freshwater populations have been established ``de novo'' from marine colonists.
-Previous results have found similar phenotypes between independently evolved freshwater populations are derived largely from the same standing genetic variants found to be Identical-by-decent.
-Geographic isolation preventing direct migration between the freshwater populations strongly suggests that these shared adaptive alleles are transported through the widely distributed marine population.
-However it is still largely unknown how gene flow, recombination, and selection jointly impact standing variation to fuel adaptation.
-Here we use genealogical simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback.
+Independently evolved freshwater populations show similar phenotypes that have been shown to derive largely from the same standing genetic variants.
+Geographic isolation prevents direct migration between the freshwater populations, strongly suggesting that these shared locally adaptive alleles are transported through the marine population.
+However it is still largely unknown how gene flow, recombination, and selection jointly impact the standing variation that might fuel this adaptation.
+Here we use individual-based, spatially explicit simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback.
 We aim to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and re-use of standing genetic variation. 
@@
 This process occurs over a realistic range of intermediate levels of gene flow that match previous empirical population genomic studies in stickleback. 
-Examining these simulations more deeply reveals that lower levels of gene flow leads to slow, independent adaptation to different habitats, whereas higher levels gene flow leads to significant mutation load -- but an increased probability of successful population genomic scans for locally adapted alleles.
+Examining these simulations more deeply reveals how lower levels of gene flow leads to slow, independent adaptation to different habitats, whereas higher levels gene flow leads to significant mutation load -- but an increased probability of successful population genomic scans for locally adapted alleles.
 Surprisingly, we find that the genealogical origins of most freshwater adapted alleles can be traced back to the original generation of marine individuals that colonized the lakes, as opposed to subsequent migrants.
@@
 \citep{Lai2152, Schrider2017}. 
-In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats
+In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation into adaptation dynamics for metapopulations inhabiting an array of habitats
 \citep{Hermisson_2008, Barrett2008}. 
@@
 \citep{Charlesworth_1997, Charlesworth_2003, Nosil_2009, Flaxman_2013, Samuk_2017}. 
-Concurrently, a growing number of empirical studies have identified instances of convergent evolution using SGV 
+Concurrently, a growing number of empirical studies have identified instances of convergent evolution using standing genetic variation 
 \citep{Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. 
-However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during colonization and adaptation to new environments
+However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of standing genetic variation, particularly during colonization and adaptation to new environments
 \citep{nelson2017ancient, Pritchard_2010, yeaman2011genetic, Schrider2017}. 
-It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via SGV. 
-While empirical studies in the field and lab are essential, modeling specific systems can help us understand the nature of standing genetic variants to fuel local adaptation as well as the genomic signals we can expect to see.
+It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via standing genetic variation. 
+Theoretical work is mostly limited to one or two loci \plr{CITE}, or relies on approximations of uncertain validity.
+Therefore, detailed simulation models of specific systems can 
+provide an important complement to empirical studies in the lab and field,
+by helping us understand precisely how standing genetic variation might fuel local adaptation,
+and what genomic signals we can expect to see.
 
@@
 \citep{Bell_1994, Thompson_1997, Cresko_2007}. 
-This model organism has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes 
-and the much more extensive use of SGV than once thought 
+This model organism has provided some of the earliest data showing the heterogeneous nature of divergence across genomes 
+and the much more extensive use of standing genetic variation than once thought 
 \citep{schluter2009genetics, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
@@
 \citep{Lescak2015evolution, Bassham_2018}. 
-Amazingly, despite very little change in a majority of the genome, nearly a quarter of it diverged significantly in this same time in a pattern that mirrored previous genomic studies of freshwater stickleback populations that have been geographically separated for thousands of years
+Amazingly, the portions of the genomes of these populations that showed increased divergence from the oceanic population
+mirrored those previously found to differ between ocean and freshwater populations that
+that have been geographically separated for thousands of years 
 \citep{Hohenlohe_2010, Bassham_2018}.
@@
 But how can evolution occur at such a rapid pace?
-Waiting for new mutations to arise in each lake or pond would take much longer than the decades since the Alaskan earthquake. 
+Waiting for new mutations to arise in each lake or pond would take much longer than the decades since the Alaskan earthquake
+with any plausible target size.
 Even more improbable is having divergence cluster in such similar genomic regions across independent populations. 
-An alternative hypothesis, supported by the most recent population genomic analyses, is that the majority of alleles important for freshwater adaptation are maintained in the marine individuals due to recurrent gene flow from freshwater back in to marine populations.
-In 2009 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter''- hypothesis to describe the process of alleles beneficial in freshwater environment being maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation to new freshwater habitats.
-The first clear example of the global reuse of SGV in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates 
+An alternative hypothesis
+is that the majority of alleles important for freshwater adaptation are maintained in the marine individuals due to recurrent gene flow from freshwater back in to marine populations. 
+\citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter'' hypothesis to describe the process by which 
+alleles beneficial in freshwater environments are maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation to new freshwater habitats.
+(The alleles conferring adaptation to freshwater environments are thereby ``transported'' to and reassembled in new lakes.)
+The first clear example of the global reuse of such alleles in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates 
 \citep{Colosimo_2004, Cresko_2004, Colosimo_2005}. 
@@
 
-While this growing body of population genomic data on stickleback evolution support the transporter hypothesis, a number of questions remain. 
+While the growing body of population genomic data on stickleback evolution supports the transporter hypothesis, a number of questions remain. 
 What are the actual population sizes, migration rates, and fitness differentials consistent with this hypothesis? 
 How many differentially selected alleles exist, how many are used at any one time, and how are they arranged within the genome? 
-A curious natural history observation underlying many of these questions is the fact that some newly formed freshwater habitats, such as the ponds on Middleton Island, are quite small and presumably the number of initial marine migrants is few.
+A curious natural history observation underlying many of these questions is the fact that some newly formed freshwater habitats, such as the ponds on Middleton Island, 
+are quite small and presumably the number of initial marine migrants is few,
+thus presumably representing a small portion of available variation.
 
-Here, we use forward simulations implemented in SLiM that incorporate linkage and realistic length genomes to model the stickleback metapopulation and address these questions \citep{haller2017slim,haller2018slim3}.
+Here, we use individual-based forward simulations implemented in SLiM that incorporates selection on a quantitative trait
+explicitly determined additively from a realistically long genome to model the stickleback metapopulation and address these questions \citep{haller2017slim,haller2018slim3}. 
 We ask how variation in amount of gene flow affects the genetic architecture of local adaptation to newly created freshwater ponds.
-Because we are documenting the entire genealogy of all alleles, we can determine the distribution and abundance of haplotypes across all in marine and freshwater populations to know the timing and proportion of potentially adaptive alleles that are actually utilized in each population.
+Because we record the entire genealogy of all alleles \citep{kelleher2018efficient}, 
+we can determine the distribution and abundance of haplotypes across all marine and freshwater populations to know the timing and proportion of potentially adaptive alleles that are actually utilized in each population.
 In addition, we can document how these adaptive alleles are distributed across the genomes, and as a consequence determine how this may affect the efficacy of genome scans for between-habitat differentiation.
-We find that only a few stickleback migrants per lake per generation are sufficient to maintain SGV in the ocean, allowing new freshwater populations to adapt in only tens of generations. 
+We find that only a few stickleback migrants per lake per generation are sufficient to maintain standing genetic variation in the ocean, allowing new freshwater populations to adapt in only tens of generations.
 We also find that continued gene flow of marine individuals into recently established lakes is not necessary for adaptation as the majority of alleles are present in the initial marine colonizers. 
 Finally, we find that this rapid, local adaptation can only occur over an intermediate range of levels of gene flow. 
-Low gene flow prevents variation from being transported, while high gene flow prevents local adaptation.
+Low gene flow prevents variation from being 
+maintained in the ocean if it is maladaptive there, while high gene flow prevents local adaptation. 
 Together these simulation findings enhance and extend the existing rich empirical work on stickleback evolutionary genomics.
@@
 
-To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement complex forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait.
+To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait.
 The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography, and included divergent selection in the two habitats which had substantial spatial structure.
-Certain aspects of the model remain simplistic due to computational constraints.
-Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback populations observed in nature (see the Discussion for more on this).
-\jgg{I don't know where we should talk about our small population sizes ... Is this even worth mentioning here?}
-\wac{I don't think it's necessary because the census population sizes of stickleback actually due vary widely, so we're not that far off.
-If anything you could say that we focused on smaller populations because of our questions about how many migrants are needed}
+Certain aspects of the model remain simplistic due to computational constraints;
+in particular, total population sizes are much smaller than in reality,
+although we may capture a good picture of local dynamics.
 
 \paragraph{Habitat and geography}
-
-Our simulations include two habitat types -- Marine and Freshwater -- defined by the nature of their selective pressures, each with 5,000 diploid individuals. 
+Our simulations include two habitat types -- marine and freshwater -- defined by the nature of their selective pressures, each with 5,000 diploid individuals. 
 The arrangement of these habitats, depicted in Figure~\ref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. 
@@
 Divergent selection is mediated by a single quantitative trait with different optima in marine and freshwater habitats. 
-This situation roughly models the cumulative effect of the various phenotypes thought to be under divergent selection between the habitats, such as armor morphology, body size, craniofacial variation and opercle shape (among many other traits).
+This situation roughly models the cumulative effect of the various phenotypes thought to be under divergent selection between the habitats, such as armor morphology, body size, craniofacial variation and opercle shape.
 The optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively, and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$ is determined by a Gaussian kernel with standard deviation 15, i.e.,
@@
 We use SLiM to simulate a Wright--Fisher population with non-overlapping generations and a fixed population size of 5,000 diploid individuals in each habitat. 
-Each generation, the two parents of each new offspring are chosen proportional to their fitness (all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-8}$ per locus per generation).
-Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation.
+Each generation, the two parents of each new offspring are chosen proportional to their fitness 
+(unlike actual stickleback, all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-8}$ per locus per generation).
+Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual so that approximately 200 offspring are generated in each lake, each generation.
+(A better implementation would use SLiM's ``non-Wright-Fisher'' model type,
+which was not available when we wrote the simulations; but we do not expect the difference to affect results.)
 To do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, so that the mean fitnesses of all lakes are equal before selection happens.
@@
 In addition, it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations which were retroactively added to the gene trees at a rate of $10^{-8}$ per locus per generation, as described in \citet{kelleher2018efficient}.
-The output tree sequence from each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
+The tree sequence output by each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
 To do this, we constructed the genealogical tree relating all extant chromosomes at each locus along the genome. 
@@
 
-We used neutral mutations to calculate $F_{ST}$, calculated on a per-locus basis and then averaged in windows as a measure of between-population relative differentiation. 
-Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively, and $\bar p = (p_f + p_m)/2$, then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.
+We used neutral mutations to calculate $F_{ST}$ for each locus, which we then averaged in windows as a measure of between-population relative differentiation. 
+Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively, and $\bar p = (p_f + p_m)/2$, 
+then we compute $F_{ST}$ for that mutation as $1 - (p_f (1-p_f) + p_m (1-p_m))/ (2 \bar p (1-\bar p))$.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Results \& Discussion}
@@
 At very low rates of gene flow, each new lake's population adapted almost completely independently through \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations). 
-At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles \citep{bolnick2007natural}. 
+At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles. % \citep{bolnick2007natural}. 
 Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes. 
@@
           \caption{ 
-                Mean individual trait values in the marine habitat (blue line),
-                the original lakes (yellow lines; average in orange),
-                and the new lakes (light green lines; average in dark green),
+        Mean individual trait values in the marine habitat (red line),
+                the original lakes (tan lines; average thicker),
+                and the new lakes (black lines; average thicker),
                 across the course of two simulations, with migration rates of
-                \textbf{(top)} $M=0.01$ and
-                \textbf{(bottom)} $M=0.1$ 
+                \textbf{(A)} $M=0.01$ and
+                \textbf{(B)} $M=0.1$ 
                 migrants per lake per generation, respectively.
@@
 Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat.
-Establishment of new alleles in the lakes is visible in the first few thousand generations as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations (Fig.
+Establishment of new alleles in the lakes during the first few thousand generations manifests as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations (Fig.
 \ref{fig:phenotype_time}).
@@
         All locations of pre-existing freshwater adapted alleles have been highlighted by 
-        pink dashed vertical lines with an alpha value of 0.3, so darker shades of pink imply more 
+        partially transparent pink dashed vertical lines, so darker shades of pink show more 
         freshwater adapted alleles at that location.
@@
 We measured this ``time to adaptation'' as the number of generations until average trait values in old and new lakes were within 0.5 of each other, shown in Figure~\ref{fig:TimeToAdaptation} for different rates of gene flow.
-Adaptation of new lakes took over 18,000 generations at the lowest rate of $m = 0.01$, while at $m = 1$ migrant per, lake per generation, new lakes managed to adapt in just under 60 generations.
+Adaptation of new lakes took over 18,000 generations at the lowest rate of $M = 0.01$, while at $M = 1$ migrant per lake per generation, new lakes adapted in just under 60 generations.
 
@@
 At low migration rates, the \emph{initial} period of adaptation takes roughly 25 times longer for lakes than it does for the ocean. 
-This difference occurs because at low migration rates, each lake must wait for its own novel mutations to arise in order to adapt. 
+This difference occurs because in the absence of gene flow, each lake must wait for its own novel mutations to arise in order to adapt. 
 Because the marine habitat is continuous, with 25 times more individuals than any one lake, there is a much larger influx of new mutations to be selected upon. 
@@
 Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lake and below 50\% in the marine habitat. 
+(So, ``captured'' alleles as defined above are pre-existing, but not all pre-existing alleles are captured.)
 Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles across generations.
@@
 If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%.
-At the lowest migration rate $m = 0.01$, each genome in the original lakes have almost exactly~$1/25^{th}$ of the total number of pre-existing freshwater adapted alleles -- this is because each one of the 25 lakes has adapted with a unique set of alleles.
+At the lowest migration rate ($M = 0.01$), each genome in the original lakes have almost exactly~$1/25^{th}$ of the total number of pre-existing freshwater adapted alleles -- this is because each one of the 25 lakes has adapted with a unique set of alleles.
 Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes. 
@@
 
-While increased migration allows sharing of adaptive alleles among lakes, at $m = 10.0$ the constant influx of alleles between the habitats creates substantial migration load.
-The rate at which migration load becomes substantial, 10 migrants per lake per generation, only replaces 5\% of each population each generation with migrants from the other habitat, but this is sufficient to shift the mean trait values to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
-These results present the sensitivity of a population to the amount of gene flow it experiences from neighboring species.
-In our simulations, the continued gene flow from freshwater populations meant that the alleles were maintained at a constant rate with fixed demographics. 
-However, future modeling efforts must be made to understand what population sizes, demographics, and evolutionary histories impact an ancestral population's capacity for carrying maladaptive alleles in its opposing environment, as well as what role a haplotype's genomic architecture play it to be rebuilt in a novel environment.
+While increased migration allows sharing of adaptive alleles among lakes, at $M = 10.0$ migrants per lake per generation
+the constant influx of alleles between the habitats creates substantial migration load. 
+This level of gene flow only replaces 5\% of each population each generation with migrants from the other habitat, 
+but is sufficient to shift the mean trait values to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{The efficacy of genomic scans of selection depends on gene flow}
@@
 
-What we observe most plainly in this graph is that $M = 1.0$ and $M = 10.0$ migrants per lake, per generation are the only rates of dispersal which provide reliable true negatives when observing peaks.
-This means that a large percentage of peaks past an $F_{ST}$ cutoff of $0.05$ actually lie on top of regions which could impact trait differences.
-Unfortunately, the low statistical power for all rates of $M$ would seem to suggest that many regions which impact trait differences do not appear as $F_{st}$ peaks. 
-This may imply that genome scans for local adaptation based purely on measures of differentiation may not be as reliable as previously thought.
-We therefore find that with high rates of dispersal between population, $F_{ST}$ peaks may be a good indication of causal loci that are identified, but that not all causal loci will appear as $F_{ST}$ peaks.
-It's important to note that our study in Figures~\ref{fig:Fst} and \ref{fig:Power_FP} pooled together all lakes and all marine individuals for the calculations. 
-For comparison, Figures~\ref{fig:fst_5lakes} \& \ref{fig:Power_FP_5lakes} shows the same scans but calculated only between individuals in the marine with spacial location $\le 5.0$ and the individuals in the respective 5 lakes.
-These graphs show even less power and True Positives only at $ M = 10.0$ migrants per lake, per generation
+What we observe most plainly in this graph is that $M = 1.0$ and $M = 10.0$ migrants per lake per generation 
+are the only rates of dispersal at which a large percentage of peaks (above $F_{ST} = 0.05$) 
+actually lie on top of regions which impact trait differences.
+Unfortunately, the low statistical power for all rates of $M$ would seem to suggest that many regions which impact trait differences do not appear as $F_{ST}$ peaks. 
+In other words, with high rates of dispersal between population, $F_{ST}$ peaks may reliably identify causal loci, 
+but not all causal loci will appear as $F_{ST}$ peaks.
+It's important to note that $F_{ST}$ in Figures~\ref{fig:Fst} and \ref{fig:Power_FP} was calculated after pooling together all lakes and all marine individuals.
+For comparison, Figures~\ref{fig:fst_5lakes} \& \ref{fig:Power_FP_5lakes} shows $F_{ST}$ calculated with individuals only in the first fifth of the range (5 lakes and corresponding marine habitat).
+These show a substantially lower true positive rate, as population-specific noise swamps any signal between the lakes.
 
-This study shows introgression is beneficial for inferring causative loci from divergence ($F_{ST}$) along the genome. 
-This is generally because noise of neutral divergence can appear causative when genetic drift causes more 
-differences between populations that have little gene flow between them.
-It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
-mistaken for being causative as they often display the same amount of divergence.
-In general, our study may suggest that diversity scans may not prove as useful in causal loci inference as previously thought.
-It is important to note that these summary statics could prove to be very useful if used beyond just the peaks seen when looked at as a function of genomic position.
+As has been suggested elsewhere, introgression aids identification of locally adaptive loci from scans for differentiation.
+\plr{CITE}
+This is generally because shared adaptive alleles provide something akin to independent replicates,
+allowing signal to emerge above the noise created by genetic drift.
+However, the required sample sizes and conditions on levels of admixture may be restrictive in practice.
+More generally, the difficulty of correctly identifying causal loci in even our most ideal situation
+calls for caution (and proof-of-concept simulations) in designing and carrying out scans for alleles underlying locally adaptive, polygenic traits.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Origin of introduced freshwater adaptive alleles}
@@
 What is the nature of selection on standing genetic variants? 
-In our simulations migration from marine individuals into the new lakes continues throughout introduction and adaptation: but is consistent influx of SGV a necessity for rapid adaptation of the population?
+In our simulations migration from marine individuals into the new lakes continues throughout introduction and adaptation: but is consistent influx of standing genetic variation a necessity for rapid adaptation of the population?
 To answer this, we traced the genealogy of \textit{all} adaptive alleles in \textit{all} individuals from the introduced lakes, after local adaptation.
-Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back the original generation of inhabitants in the lake as seen in Figure \ref{fig:Origin}.
-This suggests that given enough SGV, any reasonable subset of the ancestral population has the potential for rapid adaptation without the need post-colonization hybridization events.
+Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back the original generation of inhabitants in the lake (Figure \ref{fig:Origin}).
+In our simulations,
+any reasonable subset of the ancestral population has the potential for rapid adaptation without the need post-colonization hybridization events.
 
-If effectively capturing standing genetics variants is the key to rapid adaptation, as we have presented thus far, why is there a large difference in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$ seen in Figure \ref{fig:TimeToAdaptation}.
+If effectively capturing standing genetics variants is the key to rapid adaptation, as we have presented thus far, why is there a large difference in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$ seen in Figure \ref{fig:TimeToAdaptation}?
 %Figure~\ref{fig:Origin} shows both migration levels being able to effectively capture and propagate the standing genetic variants, and
@@
 %So why does it take $M = 0.1$ scenario nearly 2,000 generations to adapt. 
-This is surprising because we see a similar quantity of allele sharing at  $M = 1.0$, but much more rapid adaptation ($\approx 60$ generations).
+This is surprising because we see a similar quantity of allele sharing at  $M = 1.0$, but much more rapid adaptation ($\approx 60$ versus 2000 generations).
 %A reasonable explanation for this 33-fold discrepancy is that new lakes must wait for an additional influx of freshwater alleles through migration beyond what is present in the initial generation.
@@
 A possible explanation is that at lower rates of gene flow, freshwater alleles present as standing variation in the marine habitat are more tightly linked to marine alleles. 
-This implies the adaptation in the new lakes must wait for recombination to separate freshwater and marine adapted alleles.
+This implies that adaptation in the new lakes must wait for recombination to separate freshwater and marine adapted alleles. 
 This might be expected at low dispersal rates since freshwater alleles may need to be ``masked'' by nearby compensatory alleles to remain in the ocean for long periods of time. 
@@
 Each local haplotype is therefore neutral.
-This might also explain why at $M=0.1$, marine individuals still hold on average several alleles that shift the trait in the freshwater direction.
+This might also explain why at $M=0.1$, marine individuals still hold on average several alleles that shift the trait in the freshwater direction
 (Figure \ref{fig:MPFAI}A).
@@
 Note that some of these haplotypes will likely be lost to drift -- indeed, if they did all fix, then populations at $M=0.1$ would adapt much more quickly. 
-However, this calculation still provides contrast and a reasonable explanation for the difference in time to adaptation.
+However, this calculation supports our explanation above:
+it appears that adaptation at $M=0.1$ is slower because it must wait for intralocus recombination to free up genetic variation present but masked in the marine population.
 
@@
 Suppose a new allele enters a lake, either by migration or mutation.
-If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ it is then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical} (assuming Poisson reproduction, as we roughly have here).
+If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical} (assuming Poisson reproduction, as we roughly have here).
 If the current populations all differ from the optimum trait by $z$, and the allele has effect size $-u$ in heterozygotes,  then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
@@
 and so the rate of establishment of dominant mutations is $\beta z / 75$, i.e., about one such mutation every $33750/z$ generations. 
-At the beginning of the simulation where $z = 10$, we would then expect the fixation of alleles to take around 3,000 generations.
+% At the beginning of the simulation where $z = 10$, we would then expect the fixation of alleles to take around 3,000 generations. 
 The distribution of the effect sizes of these successfully established mutations 
-has density proportional to $u \exp(-2u)$, i.e., is Gamma with mean 1 and shape parameter 2.
+has density proportional to $u \exp(-2u)$, i.e., a Gamma distribution with mean 1 and shape parameter 2. 
 Since additive alleles have half the effect in heterozygotes, they have half the probability of establishment. 
@@
 
+\paragraph{Population size}
+In our simulations,
+the marine and freshwater populations are of equal size.
+More generally, we expect the required downstream migration rate to scale with the size of the marine population,
+since the same number of migrants make up a smaller proportion of a larger population,
+and so are less likely to have left genetic variation present in a new, founding population.
+So, if the marine population was well-mixed and tenfold larger than all freshwater populations combined,
+the number of migrants per lake per generation would need to be ten times larger than in our simulations to maintain sufficient standing genetic variation.
+If the marine population was a thousandfold larger,
+adaptation from standing genetic variation that is deleterious in the ocean might be impossible.
+
 \section*{Conclusion}
 
-In this paper, we have described a realistic simulation model of a coastal meta-population to observe the effects of selection on genetic variation across a wide range of gene flow rates.
-We have shown that over a realistic range of parameter values historical introgression is able to reproduce rapid and parallel adaptation similar to that documented in natural stickleback populations, such as Middleton island (cites), and supporting the ``Transporter"-hypothesis suggested by \citet{schluter2009genetics}.
-Selection is able to rebuild the freshwater haplotype at a rapid pace (in tens of generations) from marine populations as a medium between all freshwater populations. 
-Almost all rates of migration contributed to the efficiency of local adaptation except the highest due to migration load.
-A finding of practical consequence is that the efficacy of $F_{ST}$ based genomic scans of selection varies significantly with migration rate.
-A surprising new finding of our simulations is that the majority of adaptive variant's genealogy can be traced back to the original generation of inhabitants as opposed to post-colonization migrant alleles, and thus a randomly chosen subset of marine stickleback carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
+In this paper, we have analyzed a relatively realistic simulation study of a coastal meta-population to 
+understand how repeated local adaptation operates on complex standing variation across a wide range of gene flow rates.
+We have shown that rapid and parallel adaptation similar to that documented in natural stickleback populations
+occurs over a realistic range of parameter values.
+These simulations therefore provide an in-depth look at (and quantitative proof-of-concept of)
+the ``transporter'' hypothesis suggested by \citet{schluter2009genetics}.
+Selection is able to rebuild the freshwater haplotype at a rapid pace (in tens of generations) from variation present at migration-selection balance in marine populations 
+A finding of practical consequence is that the efficacy of $F_{ST}$-based genomic scans varies significantly with migration rate.
+Perhaps surprisingly, we found
+that the majority of adaptive variant can be traced back to the original generation of new habitats,
+with little role for post-colonization migrant alleles. 
+In other words, a randomly chosen subset of marine stickleback carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
 Thus, the large marine population was able to harbor and distribute deleterious alleles to the surrounding freshwater populations with only a few migrants per population, per generation. 
-Our simulation findings significantly extend previous empirical and theoretical population genomic studies of stickleback to make testable predictions, and provide a framework for similar studies in other organisms to determine the generality of our findings.
+% Our simulation findings significantly extend previous empirical and theoretical population genomic studies of stickleback to make testable predictions, and provide a framework for similar studies in other organisms to determine the generality of our findings.
+
+\plr{TODO: discuss total population size caveats}
 
@@
           \caption{
-        Mean individual trait values in the marine habitat (blue line),
-                the original lakes (yellow lines; average in orange),
-                and the new lakes (light green lines; average in dark green),
+        Mean individual trait values in the marine habitat (red line),
+                the original lakes (tan lines; average thicker),
+                and the new lakes (black lines; average thicker),
                 across the course of two simulations, with migration rates of
@@
         \textbf{(B)} $m=5 \times 10^{-1}$.
-                (i.e., 1, 10, and 100 migrants per lake per generation, respectively).
+                (i.e., $M=1$, 10, and 100 migrants per lake per generation, respectively).
                 Optimal trait values in the two habitats are at $\pm 10$.
@@
 "mask" the alleles through linkage which counteracts the fitness disadvantage. 
-This linkage prevents the ability for a sub-population to rapidly select on the beneficial SGV as recombination acts as a bottleneck to evolution in this situation. 
-Additionally, In Figure~\ref{fig:Origin} we found that the genetic basis for local adaptation, when selecting upon SGV, is propagated largely from the initial generation of inhabitants. 
+This linkage prevents the ability for a sub-population to rapidly select on the beneficial standing genetic variation as recombination acts as a bottleneck to evolution in this situation. 
+Additionally, In Figure~\ref{fig:Origin} we found that the genetic basis for local adaptation, when selecting on standing genetic variation, is propagated largely from the initial generation of inhabitants.
 This contrasts the belief that subsequent migration is a necessity in newly established populations ability to rebuild a haplotype.

### 2019-07-11 Stickleback_Paper.tex
.

@@
 But how many of these peaks are actually underlying trait differences that form the basis for local adaptation? 
-To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.
+To quantify this, Figure \ref{fig:Power_FP} shows the power and true positive rates that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.
 
@@
 %A reasonable explanation for this 33-fold discrepancy is that new lakes must wait for an additional influx of freshwater alleles through migration beyond what is present in the initial generation.
-However, Figure~\ref{fig:Origin} shows this is not the case: at both $m = 0.1$ and $m = 1.0$, the vast majority of alleles responsible for local adaptation came from individuals in the original generation (``captured'').
+However, Figure~\ref{fig:Origin} shows this is not the case: at both $M = 0.1$ and $M = 1.0$, the vast majority of alleles responsible for local adaptation came from individuals in the original generation (``captured'').
 
@@
       \hline
-     & m=0.01 & m=0.1 & m=1 & m=10 \\ 
+     & M=0.01 & M=0.1 & M=1 & M=10 \\ 
       \hline
@@
           \caption{ 
-        Statistical Power and False Positives as a function of $F_{ST}$ threshold. 
-        Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected.
-        False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{ST}$ threshold.
+        Statistical power and true positive rate as a function of $F_{ST}$ threshold. 
+        Statistical power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected.
+        True positive rate is the proportion of SNPs above the $F_{ST}$ threshold that affect phenotype.
           \label{fig:Power_FP}
@@
           \caption{ 
-        Statistical Power and False Positives as a function of $F_{ST}$ threshold. 
-        Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected.
-        False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{ST}$ threshold.
+        Statistical power and true positive rate as a function of $F_{ST}$ threshold. As in Figure \ref{fig:Power_FP}
+        but computed only using the first one-fifth of the total range -- i.e., fish in the first five lakes and those in the ocean with coordinate less than 5.0.
          	}
@@
           \caption{
-            Phenotypes (\textbf{Top}) (effect size and mutation type) and haplotypes \textbf{(Botton)} as a function of spatial position of the individual 
+            Phenotypes (\textbf{Top}) (effect size and mutation type) and haplotypes \textbf{(Bottom)} as a function of spatial position of the individual 
             possessing the genome. The phenotype of each haplotype is, by definition, the sum of the 
@@
         All locations of pre-existing freshwater adapted alleles have been highlighted by 
-        pink dashed vertical lines with an alpha value of 0.3, so darker shades of blue imply more 
+        partially transparent pink dashed vertical lines, so darker shades of blue imply more 
         freshwater adapted alleles at that location.
@@
 When new freshwater environments were colonized by the marine stickleback, selection was able to rebuild the freshwater haplotype. 
-Shown in Table~\ref{tab:linkage}, we found that a populations capacity to cary maladaptive alleles remains pretty constant over realistic ranges of dispersal, but low rates of gene flow may 
+Shown in Table~\ref{tab:linkage}, we found that a populations capacity to carry maladaptive alleles remains pretty constant over realistic ranges of dispersal, but low rates of gene flow may 
 "mask" the alleles through linkage which counteracts the fitness disadvantage. 
@@
 
-Between $m = 0.1$ and $m = 1.0$ we see a large difference in the total effect sizes of the haplotypes that are expected to fix.
+Between $M = 0.1$ and $M = 1.0$ we see a large difference in the total effect sizes of the haplotypes that are expected to fix.
 Given these numbers were computed assuming additivity (slight overestimate) and individuals are diploid,
-at $m = 0.1$ the mean total genotype expected to fix in the populations puts individuals at a trait value of $\approx -8$ 
+at $M = 0.1$ the mean total genotype expected to fix in the populations puts individuals at a trait value of $\approx -8$ 
 after selection on alleles haplotypes in the initial population. 

### 2019-07-11 Stickleback_Paper.tex
reword theory

@@
 It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via standing genetic variation. 
-Theoretical work is mostly limited to one or two loci \plr{CITE}, or relies on approximations of uncertain validity.
+Theoretical work is mostly limited to one or two loci, or relies on approximations of uncertain validity
+\citep[e.g.,][]{slatkin1975selection,petry1983effect,barton1986barrier,yeaman2011genetic,ralph2015standing}.
 Therefore, detailed simulation models of specific systems can 
@@
 
-\wac{In general I suggest having the finding or statement at the beginning of the sentence with a reference to the figure or statistic at the end of the sentence.
-Notice that in many cases below (e.g.
-first sentence of the first paragraph in the 'Local Adaptation' section) the sentence starts with the Figure (e.g.
-'As shown in Figure 2').
-You can just start with the declarative statement (e.g.
-"We found that local adaptation occurred in all simulations ....
-was composed of migrants (Fig.
-2)")}
-
 To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$. 
@@
 
-\subsection*{Local Adaptation: differentiation with gene flow}
-\wac{Can we change the header to be a statement of finding? e.g. Local adaptation occurs at almost all levels of gene flow}
+\subsection*{Rapid local adaptation at intermediate gene flow rates}
 
@@
 
-As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred in all simulations with the exception of the highest gene flow at which half of each population was composed of migrants. 
-Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat.
+Local adaptation occurred in all simulations with the exception of the highest gene flow at which half of each population was composed of migrants, 
+as can be seen by the mean trait values of Figure~\ref{fig:MeanPhenotype}.
+Figures~\ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat.
 Establishment of new alleles in the lakes during the first few thousand generations manifests as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations (Fig.
@@
 
-\subsection*{Sharing of freshwater adapted alleles}
-\wac{same comment as above, is there a declarative statement of finding? e.g. Freshwater adapted alleles are shared extensively and primarily via initial migrants}
+\subsection*{Widespread allele sharing between lakes}
 
+Adaptive alleles were shared between lakes at many rates of gene flow, but not at the lowest.
 At low migration rates, the \emph{initial} period of adaptation takes roughly 25 times longer for lakes than it does for the ocean. 
@@
 \subsection*{The efficacy of genomic scans of selection depends on gene flow}
-\wac{just made a suggested change}
 
@@
 As has been suggested elsewhere, introgression aids identification of locally adaptive loci from scans for differentiation.
-\plr{CITE}
+% \plr{CITE}
 This is generally because shared adaptive alleles provide something akin to independent replicates,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Origin of introduced freshwater adaptive alleles}
-\wac{same thing here - a statement of the major finding of this section?}
+\subsection*{Freshwater adaptation is mostly present in the initial generation}
 
@@
 
-\subsection*{Alignment with theoretical expectations}
-\wac{This subsection is a little incongruous with the others sections with respect to how it is written.
-Seems that the bold headers for the paragraphs was residual from notes?
-Also, this section seems to be  in a different voice from the rest of the Results/Discussion section.
-I think it's worth a bit of time trying to hone it more.}
+\subsection*{Theoretical expectations}
 
-We now revisit our observations above in the light of population genetics theory.
-The discussion will be roughly based on estimates, with a focus on intuition rather than precise calculations.
-Because we model stabilizing selection on an additive trait controlled by a moderately large number of loci within each population, more precise expectations might be obtained through quantitative genetics \citep{svardal2014general} or even Fisher's geometric model \citep{barton2001hybridization,chevin2014niche}.
+How do our results compare to what is expected from population genetics theory?
+Our simulations included many loci under selection and in linkage to many other selected alleles,
+which makes precise calculation impossible.
+Nonetheless, rough calculations based on simple population genetics theory
+-- with the benefit of hindsight --
+turn out to describe qualitatively most of the aspects of adaptation we observed above.
+Because we model stabilizing selection on an additive trait controlled by a moderately large number of loci within each population, more precise expectations might be obtained through quantitative genetics \citep{svardal2014general} or even Fisher's geometric model \citep{barton2001hybridization,chevin2014niche},
+but doing so is beyond the scope of this paper
 
-Suppose a new allele enters a lake, either by migration or mutation.
-If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical} (assuming Poisson reproduction, as we roughly have here).
-If the current populations all differ from the optimum trait by $z$, and the allele has effect size $-u$ in heterozygotes,  then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$.
+One of the most useful things that theory tells us is about
+the fate of a new allele in a lake, that has appeared by either migration or mutation. 
+If the allele has fitness advantage $s$ --
+-- i.e., when it is rare but present in $n$ copies,
+the expected number of copies in the next generation is $(1+s)n$ --
+then the probability that it escapes demographic stochasticity to become common in the population 
+is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical} 
+(assuming Poisson reproduction, as we roughly have here). 
+Since we study a quantitative trait, the fitness effect of each allele
+depends on the population context:
+if the individuals in the population all have trait values $z$ units above the optimum,
+and the allele has effect size $-u$ in heterozygotes,
+then the fitness advantage of the allele is the ratio of fitnesses with and without the allele.
+As we calculate fitness here, this is
+$s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$,
+where $\beta = 1 / 450$. 
 This tells us two things: (1) the rate of adaptation decreases as the population approaches the optimum, and (2) larger mutations (in the right direction) are more likely to fix.
@@
 \paragraph{New mutations}
-
-The total rate of appearance of new mutations per lake is $\mu_L = 0.04$, which are divided evenly in six categories: additive, dominant, and recessive, in either direction.
-This implies that a new additive or dominant effect mutation appears once every 75 generations, on average.
-Effect sizes are randomly drawn from an Exponential distribution with mean $1/2$,
-and so the probability that a dominant mutation manages to establish in a population differing from the optimum by $z$ is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$, 
-and so the rate of establishment of dominant mutations is $\beta z / 75$, i.e., about one such mutation every $33750/z$ generations. 
+The total rate of appearance of new mutations per lake is $\mu_L = 0.04$ per generation, 
+and these are divided evenly in six categories: additive, dominant, and recessive, in either direction. 
+Therefore, a new additive or dominant effect mutation appears once every 75 generations, on average. 
+The effect size of each new mutation is randomly drawn from an Exponential distribution with mean $1/2$, 
+and so averaging the probability of establishment over this distribution,
+we get that the probability that a dominant mutation manages to establish in a population differing from the optimum by $z$ 
+is roughly $\int_0^\infty 4 \beta z u \exp(-2u) du = \beta z$.
+Multiplying the rate of appearance of these mutations with the probability they establish
+predicts that the rate of establishment of dominant mutations is $\beta z / 75$, 
+i.e., about one such mutation every $33750/z$ generations. 
 % At the beginning of the simulation where $z = 10$, we would then expect the fixation of alleles to take around 3,000 generations. 
-The distribution of the effect sizes of these successfully established mutations 
-has density proportional to $u \exp(-2u)$, i.e., a Gamma distribution with mean 1 and shape parameter 2. 
-Since additive alleles have half the effect in heterozygotes, they have half the probability of establishment. 
-During the initial phase of adaptation, the populations begin at around distance $z=10$ from the optimum. 
-Combining these facts, we expect adaptive alleles to appear through mutation within lakes at first on a time scale of 3,000 generations, with the time between local fixation of new alleles increasing as adaptation progresses, and each to move the trait by a distance of order 1.
-This agrees roughly with what we see in Figures~\ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp}.
+During the initial phase of adaptation, the populations begin at around distance $z=10$ from the optimum,
+and $z$ decreases as adaptation progresses.
+Mutations that successfully established are more likely to be strong:
+the distribution of the effect sizes of these successfully established mutations 
+has density proportional to $u \exp(-2u)$,
+which is a Gamma distribution with mean 1 and shape parameter 2. 
+There are also additive alleles: these have half the effect in heterozygotes, 
+and so roughly half the probability of establishment. 
+Combining these facts,
+we expect adaptive alleles to appear through mutation within lakes 
+at first on a time scale of 3,000 generations, 
+with the time between local fixation of new alleles increasing as adaptation progresses, 
+and each to move the trait by a distance of order 1. 
+This agrees roughly with what we see in Figures~\ref{fig:phenotype_time} 
+and \ref{fig:phenotype_time_supp}.
 
 \paragraph{Standing variation}
-An allele that moves the trait $u$ units in the freshwater direction, in heterozygotes, has fitness roughly $\exp(-\beta u^2) \approx 1 - \beta u^2$ in the marine environment, i.e, a fitness differential of $s = \beta u^2$.
-The product of population size and fitness differential in the marine environment for a mutation with $u=1$ is therefore $2Ns \approx 22$, implying that these alleles are strongly selected against but might drift to moderate frequency if recessive.
-The average frequency of such an allele in the marine environment at migration-selection equilibrium is equal to the proportion of individuals in the ocean replaced by migrants per generation divided by the selective disadvantage, i.e., around $m / \beta u^2$.
+How are freshwater alleles maintained in the ocean, where they are deleterious?
+Assuming that fish in the marine environment are close to their phenotypic optimum,
+an allele that when heterozygous moves the trait $u$ units in the freshwater direction,
+has fitness roughly $\exp(-\beta u^2) \approx 1 - \beta u^2$,
+i.e, a fitness differential of $s = \beta u^2$. 
+The product of population size and fitness differential in the marine environment for a mutation with $u=1$ is therefore $2Ns \approx 22$, 
+implying that these alleles are strongly selected against but might drift to moderate frequency if recessive. 
+The average frequency of such an allele in the marine environment at migration-selection equilibrium 
+is equal to the proportion of individuals in the ocean replaced by migrants per generation divided by the selective disadvantage, 
+i.e., around $m / \beta u^2$. 
 Each new lake is likely to contain a few copies of alleles at frequency above 1/400 (since each new lake is initialized with 400 randomly selected genomes). 
-With $u=1$, this occurs if $m/\beta > 1/400$,
-and so since $1/\beta = 450$, if $m \ge 5 \times 10^{-6}$ the chances are good that any particular lake-adapted allele that is present in all pre-existing lakes will appear at least once in the fish that colonize a new lake.
-However, an allele with $u=1$ only has probability of around 1/20 of establishing locally, so an allele of this size must be present in about 20 copies to ensure adaptation.
-Putting these together, we expect migration-selection balance to maintain sufficient genetic variation for new lakes to adapt if $m \ge 10^{-4}$, which corresponds to $M \ge 0.4$,
-agreeing with our results.
+For a typical allele with effect size $u=1$, 
+the equilibrium migration-selection frequency is greater than this threshold if $m/\beta > 1/400$.
+Since $1/\beta = 450$, if $m \ge 5 \times 10^{-6}$ this suggests that there is a good chance that 
+any particular lake-adapted allele that is present in all pre-existing lakes 
+will appear at least once in the fish that colonize a new lake. 
+However, an allele with effect size $u=1$ only has probability of around 1/20 of establishing locally, 
+and so must be present in about 20 copies to ensure establishment.
+Putting these calculations together, 
+we expect migration-selection balance to maintain sufficient genetic variation for new lakes to adapt if $m \ge 10^{-4}$, which corresponds to $M \ge 0.4$.
+This is in good agreement with our observations.
 % This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.
@@
 \paragraph{Migration}
-The key quantity regulating the amount of standing variation in the ocean was the \emph{downstream} migration rate, from lakes to the ocean. How important is the upstream migration rate?
+The key quantity regulating the amount of standing variation in the ocean is the \emph{downstream} migration rate, from lakes to the ocean. 
+How important is the upstream migration rate?
 If sufficient genetic variation is not present in a new lake initially, it must appear either by new mutation or by migration. 
@@
 it takes $1/m$ generations until the genetic variation introduced by migrants equals the amount initially present at colonization. 
-This implies a dichotomy: either (a) migration is high, and adaptation is possible using variants present at colonization or arriving shortly thereafter, or (b) migration is low, so adaptation takes many multiples of $1/m$ generations.
-Since in our model lower migration also reduces the amount of variation available in the ocean, we expect very little contribution of subsequent migration across any value of $m$, as seen in Figure~\ref{fig:Origin}.
+This implies a dichotomy: either 
+(a) migration is high, and adaptation is possible using variants present at colonization or arriving shortly thereafter, or 
+(b) migration is low, so adaptation takes many multiples of $1/m$ generations. 
+Since in our model lower migration also reduces the amount of variation available in the ocean, 
+we expect very little contribution of subsequent migration across any value of $m$, as seen in Figure~\ref{fig:Origin}.
 
@@
 the marine and freshwater populations are of equal size.
-More generally, we expect the required downstream migration rate to scale with the size of the marine population,
-since the same number of migrants make up a smaller proportion of a larger population,
-and so are less likely to have left genetic variation present in a new, founding population.
+More generally, we expect the downstream migration rate that is required to maintain standing variation there
+to scale with the size of the marine population,
+since the same number of migrants make up a smaller proportion of a larger population.
 So, if the marine population was well-mixed and tenfold larger than all freshwater populations combined,

### 2019-07-12 Stickleback_Paper.tex
final updates from Bill before uploading to the arXiv

@@
 However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of standing genetic variation, particularly during colonization and adaptation to new environments
-\citep{nelson2017ancient, Pritchard_2010, yeaman2011genetic, Schrider2017}. 
+\citep{Nelson_2018, Pritchard_2010, yeaman2011genetic, Schrider2017}. 
 It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via standing genetic variation. 
@@
 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter'' hypothesis to describe the process by which 
-alleles beneficial in freshwater environments are maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation to new freshwater habitats.
+alleles beneficial in freshwater environments are maintained at migration-selection balance in the larger oceanic population and therefore available to be utilized during subsequent adaptation to new freshwater habitats.
 (The alleles conferring adaptation to freshwater environments are thereby ``transported'' to and reassembled in new lakes.)
@@
 A curious natural history observation underlying many of these questions is the fact that some newly formed freshwater habitats, such as the ponds on Middleton Island, 
-are quite small and presumably the number of initial marine migrants is few,
-thus presumably representing a small portion of available variation.
+are quite small and presumably the number of initial marine migrants is few.
+The variation carried by these few initial migrants might therefore be a small subset of the total variation, and thus be insufficient to fuel adaptation without subsequent influx of alleles from additional generations of marine migrants carrying the remaining freshwater adaptive alleles.
 
@@
 In addition, we can document how these adaptive alleles are distributed across the genomes, and as a consequence determine how this may affect the efficacy of genome scans for between-habitat differentiation.
-We find that only a few stickleback migrants per lake per generation are sufficient to maintain standing genetic variation in the ocean, allowing new freshwater populations to adapt in only tens of generations.
+
+-- I'm wondering if we can drop the following sentences that describe the findings from the Introduction.
+We state similar things in the abstract already, and the Introduction ends nicely with the previous sentence --
+
+**We find that only a few stickleback migrants per lake per generation are sufficient to maintain standing genetic variation in the ocean, allowing new freshwater populations to adapt in only tens of generations.
 We also find that continued gene flow of marine individuals into recently established lakes is not necessary for adaptation as the majority of alleles are present in the initial marine colonizers. 
@@
 maintained in the ocean if it is maladaptive there, while high gene flow prevents local adaptation. 
-Together these simulation findings enhance and extend the existing rich empirical work on stickleback evolutionary genomics.
+Together these simulation findings enhance and extend the existing rich empirical work on stickleback evolutionary genomics.**
 
@@
 
-To investigate in more depth how locally adaptive alleles found in the original lakes are shared between lakes, as well as how they spread to the new lakes, we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation.
+To investigate in more depth how locally adaptive alleles found in the original lakes are shared among newly derived lakes, as well as how they spread to the new lakes, we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation.
 To be considered in this category, an allele must participate in the genetic basis of local adaptation for at least one of the original lakes. 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Freshwater adaptation is mostly present in the initial generation}
+\subsection*{Alleles for freshwater adaptation are mostly present in the initial generation}
 
@@
 
-If effectively capturing standing genetics variants is the key to rapid adaptation, as we have presented thus far, why is there a large difference in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$ seen in Figure \ref{fig:TimeToAdaptation}?
+If effectively capturing standing genetic variants is the key to rapid adaptation, as we have presented thus far, why is there a large difference in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$ seen in Figure \ref{fig:TimeToAdaptation}?
 %Figure~\ref{fig:Origin} shows both migration levels being able to effectively capture and propagate the standing genetic variants, and
@@
 
-\subsection*{Theoretical expectations}
+\subsection*{Simulation results align with theoretical expectations}
 
@@
 is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical} 
-(assuming Poisson reproduction, as we roughly have here). 
-Since we study a quantitative trait, the fitness effect of each allele
+, assuming Poisson reproduction, as we roughly have here. 
+Since we are studying a quantitative trait, the fitness effect of each allele
 depends on the population context:
@@
 adaptation from standing genetic variation that is deleterious in the ocean might be impossible. 
+The fact that repeated adaptation from standing genetic variation has been shown to occur in real world threespine stickleback populations argues that the population sizes of marine and combined freshwater stickleback are not that different.
+While this might at first seem implausible given the ocean is large and lakes are small, in reality there are hundreds of thousands if not millions of lakes inhabited by threespine stickleback in coastal regions of the Northern Hemisphere, and the limiting habitat in the ocean constraining marine stickleback population sizes is likely to be coastal nesting areas.
+Together, these facts make the equivalency of marine and freshwater stickleback population sizes much more plausible.
 
@@
 A finding of practical consequence is that the efficacy of $F_{ST}$-based genomic scans varies significantly with migration rate.
-Perhaps surprisingly, we found
-that the majority of adaptive variant can be traced back to the original generation of new habitats,
+Perhaps surprisingly, we found that the majority of adaptive variant can be traced back to the original generation of new habitats,
 with little role for post-colonization migrant alleles. 
-In other words, a randomly chosen subset of marine stickleback carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.
-Thus, the large marine population was able to harbor and distribute deleterious alleles to the surrounding freshwater populations with only a few migrants per population, per generation. 
+In other words, a randomly chosen subset of marine stickleback carry the capacity to rapidly adapt in freshwater habitats without \emph{any} continued connection to the rest of of the species.
+Thus, the large marine population was able to harbor and distribute alleles that were deleterious in the ocean habitat, but adaptive in freshwater habitats, to surrounding freshwater populations with only a few migrants per population, per generation.
 % Our simulation findings significantly extend previous empirical and theoretical population genomic studies of stickleback to make testable predictions, and provide a framework for similar studies in other organisms to determine the generality of our findings.

### 2019-07-18 Stickleback_Paper.tex
removed bit of intro as suggested by bill

@@
 In addition, we can document how these adaptive alleles are distributed across the genomes, and as a consequence determine how this may affect the efficacy of genome scans for between-habitat differentiation.
-
--- I'm wondering if we can drop the following sentences that describe the findings from the Introduction.
-We state similar things in the abstract already, and the Introduction ends nicely with the previous sentence --
-
-**We find that only a few stickleback migrants per lake per generation are sufficient to maintain standing genetic variation in the ocean, allowing new freshwater populations to adapt in only tens of generations.
-We also find that continued gene flow of marine individuals into recently established lakes is not necessary for adaptation as the majority of alleles are present in the initial marine colonizers. 
-Finally, we find that this rapid, local adaptation can only occur over an intermediate range of levels of gene flow. 
-Low gene flow prevents variation from being 
-maintained in the ocean if it is maladaptive there, while high gene flow prevents local adaptation. 
-Together these simulation findings enhance and extend the existing rich empirical work on stickleback evolutionary genomics.**
 
@@
 
-\plr{TODO: discuss total population size caveats}
-
 \bibliographystyle{plainnat}

### 2019-09-17 review-responses.tex
reviews

%%%%%%
%%
%%  Don't reorder the reviewer points; that'll mess up the automatic referencing!
%%
%%%%%

\begin{minipage}[b]{2.5in}
  Resubmission Cover Letter \\
    {\it G3}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
    Jared Galloway, \\
    William Cresko, \\
    \emph{and} Peter Ralph \\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revised version of our manuscript, 
``A few stickleback suffice for the transport of alleles to new lakes'',
XXX

\vspace{5em}

\noindent \hspace{4em}
\begin{minipage}{3in}
\noindent
{\bf Sincerely,}

\vskip 2em

{\bf 
Jared Galloway, Bill Cresko, and Peter Ralph.
}\\
\end{minipage}

\vskip 4em

\pagebreak
\setcounter{page}{1}

%%%%%%%%%%%%%%
\reviewersection{AE}

\begin{quote}
    Please submit your revision along with a response to the reviewers' concerns
    and suggestions, which can be viewed at the bottom of this email. I agree
    particularly with questions from two of the reviewers asking for
    explanation or additional simulations on the founder population size, with
    reviewer 3's point about clarifying old vs new, and reviewer 1's request
    for more simulations. I do understand that simulations can be time
    consuming, however, and want to make clear additional simulations are not a
    requirement for acceptance (but would be rad if feasible). My lab and I
    also spent some time debating what was meant by the explanation at the
    bottom of page 13 -- I think more effort to clarify here would be helpful.
    Finally, we also wondered why the simulations started at a trait value of 0
    instead of 10 -- it makes comparison of old vs new a bit harder than seems
    necessary. I expect these revisions can be done within 30 days, but don't
    hesitate to ask if you need more time. 
\end{quote}

\plr{IMPORTANT: don't reorder or delete ``points'' below - it messes up the automatic numbering!}

%%%%%%%%%%%%%%
\reviewersection{1}

\begin{quote}
This is a very well done work addressing the fundamental problem of local
adaptation in metapopulations; stickleback are a nice study system but the
findings are relevant for a wide variety of systems. I enjoyed it very much,
and my only wish is that the simulation has been run more than once. This is
not necessary for speed of adaptation - within-simulation replication of lakes
is totally sufficient for this - but for aspects related to genome scanning,
numbers of adaptive alleles in a metapopulation, and haplotype-wise effect
sizes (table 1), since (especially at high M) the number of adaptive alleles is
very few across the whole metapopulation. That said, I do not insist on
rerunning the simulation, since it must be very computationally intensive.
\end{quote}

\begin{point}{}
    \textit{``As expected, the genetic basis of the freshwater phenotype seems to simplify
    as migration increases - higher rates of migration allow adaptive alleles of
    larger effect to travel more efficiently through the population, even though
    they are deleterious in the ocean.''}
    Very cool. It gives a very intuitive guideline as to which genetic
    architecture of local adaptation to expect given a certain migration rate.
    There is one neat aspect of this result that is perhaps worth better
    emphasizing (please correct me if I an : detectable Fst peaks at truly adaptive
    alleles can only be very few. If we had replicate runs we could say how many
    true Fst peaks, on average, we see at M=1 and M=10, Many published studies
    (particularly in marine animals) report dozens if not hundreds Fst outliers
    throughout the genome; your simulation here suggests that most of them have to
    be false positives of some kind, especially since migration in the sea is
    likely to be high.
\end{point}

\reply{
}

\begin{point}{}
Also, about Fig. S1: I am confused what is ``statistical power'' here :
``likelihood that a SNP will be predicted to have an effect on phenotype when
there is an effect to be detected''. In this model every SNP has at least *some*
effect on phenotype, so what defines ``effect to be detected'' then?
\end{point}

\reply{
}

\begin{point}{}
    \textit{``Surprisingly, for all cases with the exception of M = 0.01, we found that the
    majority of adaptive allele origins traced back the original generation of
    inhabitants in the lake (Figure 7)''}
    This is really cool and counter-intuitive,
    but maybe it is not too surprising since first-arriving (``captured'') alleles
    are simply much more numerous relative to the alleles that trickle in later,
    *before the adaptation is complete*? Indeed, initial colonization involves
    arrival of 200 individuals in a single generation, which is equivalent to 200/M
    generations of subsequent immigration. Considering that at any appreciable M
    (>=0.1) adaptation happens very rapidly (fig 3 and S3), there might simply be
    not enough additional immigrant alleles to make a contribution. I wonder how
    this would work if colonization of a lake was a more gradual process, say 2
    individuals per generation (not all 200 at once)? Would the early-arriving
    alleles still have advantage? (I would bet yes)
\end{point}

\reply{
}

%%%%%%%%%%%%%%
\reviewersection{2}

\begin{quote}
    This is an interesting, and as far as I can judge, a sound simulation study
    focused on population genetics of adaptation in the threespine stickleback.
    While the problem(s) approached in the manuscript are interesting and relevant,
    the way the ms has been pitched makes is mostly interesting for community of
    researchers interested on sticklebacks. By angling the story in more general
    terms and bringing stickleback in as model system, broader general appeal might
    have been achieved.
\end{quote}

\begin{point}{}
Latin binomial of the focal model species appears only in the references. It would make sense to incorporate it in the abstract and introduction given that the entire ms is focused on this species.
\end{point}

\reply{
    Good point; we've added it to the first use of the name after the abstract \revref.
}

\begin{point}{Title:}
I would make it clear that this is about the ``threespine stickleback'' - there are many different stickleback species.
\end{point}

\reply{
}

\begin{point}{\revref}
``to millions of independently derived''. Do the references you cite say so? Can you really back up your claim of millions of freshwater populations?
\end{point}

\reply{
}

\begin{point}{\revref}
``heterogenous nature of divergence''. I think you should cite here the first (paper using global sampling) showing evidence for heterogenous nature of this genetic divergence: DeFaveri et al. (2011). Global analysis of genes involved in freshwater adaptation in threespine sticklebacks (Gasterosteus aculeatus). Evolution, 65(6), 1800-1807.
\end{point}

\reply{
}

\begin{point}{\revref}
P2. ``stickleback in them'' - threespine stickleback in them
\end{point}

\reply{
}

\begin{point}{p.3 Methods}
From here on, the ms is generally very thinly referenced. In particular, it remains unclear to reader where the many equations come from: are they something authors have derived themselves (I guess not?), or who did that?
\end{point}

\reply{
}

\begin{point}{p.5}
One is left to wonder if it realistic to think that the marine population is as big as 25 freshwater populations. Admittedly, this is discussed to some extent in the discussion, but given the abundance of the threespine stickleback in many seas (there is actually quantitative estimates of this in literature), it seems plausible that the marine population(s) is orders of magnitudes larger than that of typical freshwater population, even if combined. I guess a more quantitative view on this in light of literature would be something worth attempting.
\end{point}

\reply{
}

\begin{point}{\revref}
    Where does the mutation rate $10^{-10}$ come from? Why this particular value, and how are the results influenced by choice of this?
\end{point}

\reply{
}

\begin{point}{}
P4. It is bit unclear here how LD as was dealt with in the simulations. Are all the loci assumed to be unlinked?
\end{point}

\reply{
}

\begin{point}{\revref}
What happens if generations are overlapping, as is the case in sticklebacks?
\end{point}

\reply{
}

\begin{point}{\revref}
Here the authors say that a different model in SLiM would have been more appropriate to be used, but that they ``do not expect this to influence the results''. This does not sound very good - why not to check this? Why not to rerun the analyses if a more appropriate model is available?
\end{point}

\reply{
}

\begin{point}{Fig 1}
The modeled scenario is probably correct over shorter geographic distances. But there is plenty of empirical evidence to show that marine threespine stickleback populations over larger geographic distances are not genetically homogenous. I think it would be appropriate to acknowledge this somewhere in manuscript.
\end{point}

\reply{
}

\begin{point}{\revref}
``The efficacy of genomic scans of selection depends on gene flow''. I found this section rather uninformative: all this has been studied in quite detail already in many theoretical (and empirical) papers. This could be shortened. And references to earlier findings would be in place here too - there is no single reference to earlier works here at all.
\end{point}

\reply{
}

\begin{point}{\revref}
``Simulation results align with..'' It is nice that the results are referenced back to theory. I was left to think that it might have been better to do this by combining this section with the actual results. Now the structure of the ms appears to be: 1) Results \& Discussion, 2) Discussion. Furthermore, references to theory are given in the first paragraph, but not after this. Readers do not know where all the equations are derived from.
\end{point}

\reply{
}

\begin{point}{\revref}
``we found that standing..'' Is ``found'' right word here? You made an inference, even a guess, that this was so. But you did not prove this?
\end{point}

\reply{
}

\begin{point}{\revref}
``Population size''. Here it is argued that population sizes of marine and freshwater populations are not so different. I found it hard to believe this in the light of quantitative data there is. Could it simply be so that marine population is actually very much larger, but structured (as there is genetic evidence for this in many papers) so that the fraction feeding the freshwater colonizations is to some degree isolated from ones feeding the freshwaters in a particular area? This would in effect make relevant portion of the marine population smaller, without need to postulate that the total marine population equals in size to freshwater populations, and that there are millions of lakes inhabited by three spine sticklebacks. Note also that in introduction it was claimed millions of freshwater populations, here you say ``if not millions'' - inconsistent.
\end{point}

\reply{
}

%%%%%%%%%%%%%%
\reviewersection{3}

\begin{quote}
    \ldots
    Overall, the paper finds clear support for the transporter hypothesis, and
    uncovers some key conditions under which the transporter mechanism is most
    likely to occur. An interesting insight is that little gene flow can suffice to
    provide enough standing variation in the marine gene pool for parallel
    freshwater adaptation to mostly involve pre-existing (instead of de novo). I
    also found interesting that freshwater-adaptive alleles can be maintained for a
    relatively long time in the marine population, thus indicating that it's not a
    necessity for gene flow to occurs at all times between freshwater and the sea
    for genetic variation to be shared by descent among freshwater populations. I
    also liked the paper's insights on the importance of considering entire
    adaptive haplotypes - and not only the number of adaptive alleles - for
    repeated adaptation.  
\end{quote}

\begin{quote}
    The paper is well written and I like this work and don't see any
    fundamental flaws in the way the simulations were set up or in how the results
    are interpreted. Of course, one always needs to make certain assumptions when
    setting up simulations (such as on effect sizes of loci etc.), but I think that
    the authors explore the parameter space appropriately. 
\end{quote}

\begin{quote}
    In what follows, I provide suggestions on how to further improve the work. Most
    of these points should be understood as recommendations. A side-note:
    Please make sure to add line numbers to your next submission. It is
    difficult for reviewers to refer to specific text passages without line
    numbers being present.
\end{quote}

\begin{point}{}
Founder effects followed by population growth: I wonder how founder effects during the colonization of the 'new lakes', followed by population growth, would influence the results? It seems that this was never explored, as each new lake had a constant population size of 200 individuals (rather low). I don't expect there to be a qualitative change in the overall results, but the relative contribution of de novo vs. standing variants to adaptation might well be influenced by a more colonization scenario. If the authors decided not to run additional simulations on this issue, it might at least be worth speculating on how founder effects followed by population growth might influence some of their results (such as the results shown in Fig. 7).
\end{point}

\reply{
}

\begin{point}{}
If possible, it would be interesting to check what type of mutations - in terms of effect size and in terms of dominance - fall under the 'captured', 'marine' and 'de novo' categories. I could well imagine, for example, that mutations that can stay in the marine populations relatively long are more likely to be recessive. Is this true?
\end{point}

\reply{
}

\begin{point}{}
I'm a bit confused by how the adaptive loci along the chromosome were modeled. I don't fully understand what you mean by mutations occurring at a rate $10^{-10}$ per locus per generations in ten regions of $10^5$ loci each (Page 4, 'Genetic architecture of the trait'). Why are the adaptive loci themselves organized in distinct regions? Why isn't there simply a given number of adaptive loci that are evenly spread across the chromosome?
\end{point}

\reply{
}

\begin{point}{Results \& Discussion section}
    Sometimes, it is not fully clear whether you talk about the 'old' lakes or the 'new' lakes. This is often unclear, for example, in the section ``Rapid local adaptation at intermediate gene flow rates''. Could you make this clear throughout the manuscript?
\end{point}

\reply{
}

\begin{point}{Figure 2}
    Are these the results after 50k generations?
\end{point}

\reply{
}

\begin{point}{}
    In the legend of Figure 4, it might be good to re-iterate what the number on top of the panels means exactly. Say that these are actual numbers of migrants per population and what the population size is. What is meant by 'M' is said more clearly in the legend to Figure 6, yet, it would be good to here also state (again) how large the populations are. M can only be understood in context of the actual population size. Otherwise, you would need to provide 'm'.
\end{point}

\reply{
}

\begin{point}{}
Figure 5: delete '100' along the x-axis.
\end{point}

\reply{
}

\begin{point}{}
Typo on page 14: two '- -' instead of only one.
\end{point}

\reply{
    Whoops! Fixed; thanks. \revref
}

\begin{point}{}
Page 2, 2nd paragraph (beginning): When you talk about the evolutionary history of stickleback, an appropriate citation would also be Bell \& Foster 1994.
\end{point}

\reply{
}

\begin{point}{}
Page 2, 2nd paragraph: You provide several empirical papers that demonstrate the re-use of standing variation in stickleback. Your list of papers is fair, although I don't think you chose the best papers of all the ones available. There are papers that are not cited that directly show evidence of standing variation in the sea that is being re-used for freshwater adaptation, such as: Jones et al. 2012 (Curr Biol); Haenel et al. 2019 (EvolLet), Feulner et al. 2014 (MolEcol) etc. Evidence for the reuse of the exact same genetic variation at different locations on the northern hemisphere also comes from inversions in the stickleback genome (Jones et al. 2012 (Nature); Roesti et al. 2015 (NatCom)).
\end{point}

\reply{
}

\begin{point}{}
Page 2, 2nd paragraph: you here talk about 'similar phenotypes' that have evolved repeatedly in freshwater, but you provide only two examples. The cited papers are mostly concerned with the genetic basis of a specific phenotype that has repeatedly evolved (plate loss). However, at the phenotypic level, there are many more examples for marine-freshwater adaptations in stickleback.
\end{point}

\reply{
}

\begin{point}{}
Page 3, 2nd sentence ('An alternative hypothesis ... maintained in the marine .... due to recurrent gene flow'). I would replace 'maintain' with 'present' because it is unclear at this point whether alleles are actually maintained in the marine population (i.e., stay there for many generations), or whether they are just 'washed into' the marine population over and over again from freshwater without staying long in the marine population at all.
\end{point}

\reply{
}

\begin{point}{}
It might be good to cite Hohenlohe et al. 2010 and/or Roesti et al. 2014 to justify that the marine population is modeled to be panmictic (page 4, top). Both these studies have shown very little genome-wide differentiation between samples taken at different marine locations.
\end{point}

\reply{
}

### 2019-09-17 Stickleback_Paper.tex
reviews

@@
 \newcommand{\wac}[1]{\todo[linecolor=green,backgroundcolor=yellow!25,bordercolor=black]{#1}}
+
+%%%%%%%%%%%
+% for responses to reviewers
+% set this to show line numbers and include responses to reviews or not
+\newif\ifreviewresponses
+\reviewresponsestrue  % include them
+% \reviewresponsesfalse  % don't include them
+\input{review-response-commands}
+\newcommand{\responsefile}{review-responses.tex}  % name of the review reponses file
 
@@
 
-The ancestral marine form of threespine stickleback fish has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere 
+The ancestral marine form of threespine stickleback fish (\textit{Gasterosteus aculeatus}) \revpoint{2}{1}
+has given rise to millions of independently derived freshwater populations \revpoint{2}{3}
+in recently de-glaciated regions around the Northern Hemisphere 
 \citep{Bell_1994, Thompson_1997, Cresko_2007}. 
-This model organism has provided some of the earliest data showing the heterogeneous nature of divergence across genomes 
+This model organism has provided some of the earliest data showing the heterogeneous nature of divergence across genomes \revpoint{2}{4}
 and the much more extensive use of standing genetic variation than once thought 
 \citep{schluter2009genetics, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
-While geographic isolation often prevents direct migration between freshwater populations, stickleback in them frequently evolve similar phenotypes 
-\citep{Cresko_2004, Colosimo_2004}. 
+While geographic isolation often prevents direct migration between freshwater populations, stickleback in them \revpoint{2}{5}
+frequently evolve similar phenotypes \citep{Cresko_2004, Colosimo_2004}. 
 The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years 
@@
 Each individual carries two linear chromosomes, each of size $10^8$ loci. 
-Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation in ten regions of $10^5$ loci each, spread evenly along the chromosome. 
+Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation \revpoint{2}{8}
+in ten regions of $10^5$ loci each, spread evenly along the chromosome. 
 Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
@@
 
-We use SLiM to simulate a Wright--Fisher population with non-overlapping generations and a fixed population size of 5,000 diploid individuals in each habitat. 
+We use SLiM to simulate a Wright--Fisher population with non-overlapping generations \revpoint{2}{10}
+and a fixed population size of 5,000 diploid individuals in each habitat. 
 Each generation, the two parents of each new offspring are chosen proportional to their fitness 
@@
 Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual so that approximately 200 offspring are generated in each lake, each generation.
-(A better implementation would use SLiM's ``non-Wright-Fisher'' model type,
+(A better implementation would use SLiM's ``non-Wright-Fisher'' model type, \revpoint{2}{11}
 which was not available when we wrote the simulations; but we do not expect the difference to affect results.)
@@
 
+\subsection*{Data availability}
+
+The source code used to run and analyze the simulations in this paper
+are available at \url{https://github.com/jgallowa07/SticklebackPaper}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 As expected, increasing migration rate decreased differentiation between habitats: as seen in Figure~\ref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.
+\revpoint{1}{1}
 
@@
 
+\revpoint{2}{13}
 Here, we take a closer look at the genomic architecture of local adaptation between the two habitats. 
@@
 
+\revpoint{2}{14}
 How do our results compare to what is expected from population genetics theory?
@@
 the fate of a new allele in a lake, that has appeared by either migration or mutation. 
-If the allele has fitness advantage $s$ --
--- i.e., when it is rare but present in $n$ copies,
+If the allele has fitness advantage $s$ -- \revpoint{3}{8}
+i.e., when it is rare but present in $n$ copies,
 the expected number of copies in the next generation is $(1+s)n$ --
@@
 % This corresponds to our highest two migration rates, as in e.g., Figure \ref{fig:phenotype_ts2}B.
-However, this calculation treats each allele independently; in practice we found that standing freshwater variation in the ocean was masked by linkage to compensatory marine variants.
+However, this calculation treats each allele independently;
+in practice we found that standing freshwater variation in the ocean
+was masked by linkage to compensatory marine variants. \revpoint{2}{15}
 
@@
 \paragraph{Population size}
-In our simulations,
+In our simulations, \revpoint{2}{16}
 the marine and freshwater populations are of equal size.
@@
 \end{figure}
+
+%%%%%%% reviews
+\newpage
+
+\includereviews
 

### 2019-09-18 review-responses.tex
finished reviwer comments for 1 & 2

@@
 
-\begin{point}{}
+\begin{point}{\revref}
     \textit{``As expected, the genetic basis of the freshwater phenotype seems to simplify
@@
 \reply{
+We agree with the reviewer that this is an important point, 
+and that our results are consistant with the above hypothesis. 
+However, to quantify and assert that previous empiracal studies 
+are reporting false-positives would require many 
+more simulations (and \textit{careful} analysis), which we believe is outside the rhelm of this paper.
 }
@@
 \reply{
+TODO: Peter, you might be able to best describe this most clearly.
 }
 
-\begin{point}{}
+\begin{point}{\revref}
     \textit{``Surprisingly, for all cases with the exception of M = 0.01, we found that the
@@
     generations of subsequent immigration. Considering that at any appreciable M
-    (>=0.1) adaptation happens very rapidly (fig 3 and S3), there might simply be
+    ($>=0.1$) adaptation happens very rapidly (fig 3 and S3), there might simply be
     not enough additional immigrant alleles to make a contribution. I wonder how
@@
 \reply{
+We agree that this is an important point to explore, and a good observation.
+We have appended to the section on migration to more clearly address this \llname{disc_colon}.  
+However, the available code could certainly be used in further studies to investigate this question
+more in-depth.
 }
@@
 \reply{
+    We believe the first sentence in the abstract makes this clear to the reader.
 }
@@
 \reply{
+    This order of magnitute estimate is roughly based upon the length of the coastal range and coastal lake density.
+Because there is no concrete evidence, we have re-worded this \llname{hundreds-thousands-or-more}.
 }
@@
 \reply{
+    Thank you for pointing this out. The citations has been added.
 }
@@
 \reply{
+    Fixed.
 }
@@
 \reply{
+    All equations explaining dynamics of the simulation model (outside of \llname{Fst}) were chosen explicitely by the authors as input to SLiM.
+We have added citations for the population genetics theoretical expectations (\llname{Gil_cite} \llname{Fst}) as well as clarified the derivation of equations in the Methods section \llname{clarify_equations}.
 }
@@
 \reply{
+    TODO
 }
@@
 \reply{
+    This was chosen as an arbitrary, yet realistic, rate at which you could expect incoming mutations to affect phenotype. 
+At $10^{6}$ loci where these mutations can arise at a rate of $10^{-10}$, and $5000$ diploid individuals in each population, you can expect $\approx 1$ mutations to arise in each population, each generation, which impact the trait value.
+Another way to look at this is that it is a per-nucleotide mutation rate of $10^{-8}$ with a total target size of $10^{4}$. 
+Because we hold selective pressure (optimum trait value) constant and stack mutations, mutation rate will only affect the speed of adaptation through the use of ``de-novo'' mutation. 
+While this comparison is an important point in our ms,
+the authors believe this rate is, if anything, a generous assumption to make.
+We have re-worded and appended to the ms to hopefully make this more clear. \llname{mu_exp}
 }
@@
 \reply{
+    All chromsomes are linear with a per-nucleotide recombination rate of $10^{-8}$. This has been clarified here \revref.
 }
@@
 \reply{
+The authors expect no fundemental difference between results yeilded from WF model vs. non-WF model.
 }
@@
 \reply{
+In a perfect world, we would re-run and compare. 
+However, to re-write this model as non-WF would require a relatively significant amount of change to the code as well as much more computational power to re-run the analysis.
+With no fundemental differences expected between the two models other than one being slightly more realistic, the authors opted to stick with WF model. 
+We agree the ms makes this slightly confusing, so we have altered the wording a bit here \llname{realistic}.
 }
@@
 \reply{
+The authors agree, it should be made clear that the oceanic population is not panmictic but rather has one-dimential continuous spatial structure.
+We have clarified this here \llname{continuous}.
 }
@@
 \reply{
+The authors believe this finding supports existing studies, specifically, how stronger genetic drift with less migration leads to higher background $F_{ST}$.
+However, for our study, showing how greater sharing of adaptive alleles provides a shared signal across populations seemed necessary for a study of selection on standing variants.
+Thus, the authors have decided to leave this section as is.
+TODO: Looks like we could use some citations, I was unable to find any good ones :(.
 }
@@
 \reply{
+The authors thought about connected it back to the results and decided it would be more confusing to the reader. 
+As above, we have decided to reference Gillespie throughout this section.
 }
@@
 \reply{
+TODO
+\jgg{Justify that this is actually a finding, by adding a sentence to the bit on masking to make this more concrete.} 
 }
@@
 \reply{
+TODO
+\jgg{This is clearly tied to the above comments about pop size, not sure how to answer these effectively}
 }
@@
 \reply{
+Good point: Fixed.
 }

### 2019-09-18 Stickleback_Paper.tex
finished reviwer comments for 1 & 2

@@
 The ancestral marine form of threespine stickleback fish (\textit{Gasterosteus aculeatus}) \revpoint{2}{1}
-has given rise to millions of independently derived freshwater populations \revpoint{2}{3}
+has given rise to hundreds of thousands if not millions of independently derived freshwater populations \revpoint{2}{3} \llabel{hundreds-thousands-or-more}
 in recently de-glaciated regions around the Northern Hemisphere 
@@
 and the much more extensive use of standing genetic variation than once thought 
-\citep{schluter2009genetics, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
-While geographic isolation often prevents direct migration between freshwater populations, stickleback in them \revpoint{2}{5}
+\citep{schluter2009genetics, DeFaveri_2011, DeFaveri_2013, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
+While geographic isolation often prevents direct migration between freshwater populations, threespine stickleback in them \revpoint{2}{5}
 frequently evolve similar phenotypes \citep{Cresko_2004, Colosimo_2004}. 
@@
 To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait.
-The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography, and included divergent selection in the two habitats which had substantial spatial structure.
+The details and equations of the model were motivated by current understanding of threespine stickleback evolutionary, history and demography, as well as population genetics theory \citep{Gillespie_1998}.
+\llabel{clarify_equations}
+The following model included divergent selection in the two habitats which had substantial spatial structure.
 Certain aspects of the model remain simplistic due to computational constraints;
@@
 
-Each individual carries two linear chromosomes, each of size $10^8$ loci. 
-Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation \revpoint{2}{8}
-in ten regions of $10^5$ loci each, spread evenly along the chromosome. 
+Each individual carries two linear chromosomes, each of size $10^8$ loci,
+with recombination events occurring at a intralocus rate of $10^{-8}$ every generation. \revpoint{2}{9}
+%Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation \revpoint{2}{8}
+%in ten regions of $10^5$ loci each, spread evenly along the chromosome. 
+Mutations that can affect the trait under selection can occur at rate $10^{-8}$ with a target size of $10$ Kb, (spread evenly among $1$ Mb) across each individual genome. \revpoint{2}{8}
+With 5000 (diploid) individuals per population, this equates to $\approx 1$ mutation, per population, per generation, which impact individual trait value. \llabel{mu_exp}
 Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
@@
 Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual so that approximately 200 offspring are generated in each lake, each generation.
-(A better implementation would use SLiM's ``non-Wright-Fisher'' model type, \revpoint{2}{11}
-which was not available when we wrote the simulations; but we do not expect the difference to affect results.)
+(A more realistic implementation would use SLiM's ``non-Wright-Fisher'' model type, \revpoint{2}{11}
+which was not available when we wrote the simulations; but we do not expect the difference to affect results.) \llabel{realistic}
 To do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, so that the mean fitnesses of all lakes are equal before selection happens.
 
-As depicted in Figure~\ref{fig:Geo}, dispersal occurs both locally along the coastline in the marine habitat as well as between the marine habitat and the lakes,
+As depicted in Figure~\ref{fig:Geo} the marine population is continuous (unidimensional) while lakes are discrete. \llabel{continuous}
+Dispersal occurs both locally along the coastline within the marine habitat, as well as between the marine habitat and the lakes,
 with a lake--ocean migration rate denoted $m$. 
-No migration occurs among the freshwater populations.
+No migration occurs directly between the freshwater populations. 
 All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
@@
 Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively, and $\bar p = (p_f + p_m)/2$, 
-then we compute $F_{ST}$ for that mutation as $1 - (p_f (1-p_f) + p_m (1-p_m))/ (2 \bar p (1-\bar p))$.
+then we compute $F_{ST}$ for that mutation as $1 - (p_f (1-p_f) + p_m (1-p_m))/ (2 \bar p (1-\bar p))$ \citep{Gillespie_1998}. \llabel{Fst}
 
@@
 As expected, the genetic basis of the freshwater phenotype seems to simplify as migration increases -- higher rates of migration allow adaptive alleles of larger effect to travel more efficiently through the population, even though they are deleterious in the ocean.
+\revpoint{1}{1}
 
@@
 Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back the original generation of inhabitants in the lake (Figure \ref{fig:Origin}).
+\revpoint{1}{3}
 In our simulations,
@@
 -- with the benefit of hindsight --
-turn out to describe qualitatively most of the aspects of adaptation we observed above.
+turn out to describe qualitatively most of the aspects of adaptation we observed above \citep{Gillespie_1998}. \llabel{Gil_cite}
 Because we model stabilizing selection on an additive trait controlled by a moderately large number of loci within each population, more precise expectations might be obtained through quantitative genetics \citep{svardal2014general} or even Fisher's geometric model \citep{barton2001hybridization,chevin2014niche},
@@
 we expect very little contribution of subsequent migration across any value of $m$, as seen in Figure~\ref{fig:Origin}.
+It remains unknown how different colonization schemes would impact the origin of adapted alleles, 
+but it stands to reason based upon our results that early arriving alleles are favored as adaptive alleles in any scenario. \llabel{disc_colon}
 

### 2019-09-18 review-responses.tex
finished reviewer comments

@@
 are reporting false-positives would require many 
-more simulations (and \textit{careful} analysis), which we believe is outside the rhelm of this paper.
+more simulations (and \textit{careful} analysis), which we believe is outside the realm of this paper.
 }
@@
 We have appended to the section on migration to more clearly address this \llname{disc_colon}.  
-However, the available code could certainly be used in further studies to investigate this question
-more in-depth.
+The available code could certainly be used in further studies to investigate this question more in-depth, but again feel that is outside the scope of this paper and have decidecided not to conduct more simulation and analysis.
 }
@@
 \reply{
+This is a good point and was thought about by the authors. 
+We believe this is worth noting \llname{disc_colon} in the manuscript but further exploration is outside the scope of this paper. 
 }
@@
 \reply{
+The authors agree that this is an interesting point of study. While this was explored in previous simulations, we thought it was outside the scope of this paper to make any statements about dominance.
 }
@@
 \reply{
+We agree this was confusing and have edited the ms to hopefully make it more clear \llname{mu_exp}. The motivation for 10 regions of effect was to model 'islands' of genetic divergence.
 }
@@
 \reply{
+Good point, I have found a few ambiguous spots and made sure to make this more clear
+\revref
 }
@@
 \reply{
+    Actually, at the time of introduction (so 100,000 generations), thank you for the clarification, this has been noted in the figure caption. 
 }
@@
 \reply{
+Makes sense, the figure caption has been revised.
 }
@@
 \reply{
+Good catch - added. \revref
 }
@@
 \reply{
+Great catch, citations added. \revref
 }
@@
 \reply{
+TODO: 
+\jgg{I feel like this may be brought up in a lot a papers, but is just an observation and not a finding worth citing super extensively? I could be wrong here}
 }
@@
 \reply{
+Good point. Changed. \revref
 }
@@
 \reply{
-}
+This was clarified \llname{continuous}. The marine population was not modeled as panmictic but rather a unidimensional continuous space with dispersal.

### 2019-09-18 Stickleback_Paper.tex
finished reviewer comments

@@
 Concurrently, a growing number of empirical studies have identified instances of convergent evolution using standing genetic variation 
-\citep{Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. 
+\citep{jones_genomic_2012, roesti_genome_2012, feulner_genomics_2015, Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. \revpoint{3}{10}
 However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of standing genetic variation, particularly during colonization and adaptation to new environments
@@
 in recently de-glaciated regions around the Northern Hemisphere 
-\citep{Bell_1994, Thompson_1997, Cresko_2007}. 
+\citep{Bell_1994, Thompson_1997, Cresko_2007, Hunt_2008} \revpoint{3}{9}
 This model organism has provided some of the earliest data showing the heterogeneous nature of divergence across genomes \revpoint{2}{4}
@@
 An alternative hypothesis
-is that the majority of alleles important for freshwater adaptation are maintained in the marine individuals due to recurrent gene flow from freshwater back in to marine populations. 
+is that the majority of alleles important for freshwater adaptation are present in the marine individuals due to recurrent gene flow from freshwater back in to marine populations. \revpoint{3}{12}
 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter'' hypothesis to describe the process by which 
@@
           \caption{
-            Distribution of mean individual trait values across generations of the simulation,
+            Distribution of mean individual trait values across generations of the simulation 
+            (recording started after introduction of the new lakes at 100,000 generations),
             for different migration rates.
-            The dashed pink and purple lines at $\pm 10$ give the optimum phenotypes
+            The dashed yellow and grey lines at $\pm 10$ give the optimum phenotypes
             in the marine and freshwater environments, respectively.
@@
 Figures~\ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat.
-Establishment of new alleles in the lakes during the first few thousand generations manifests as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations (Fig.
+Establishment of adaptive alleles in the old lakes during the first few thousand generations manifests as jumps in the mean trait value - which move the trait by an amount of order 1 every few hundred generations (Fig.
 \ref{fig:phenotype_time}).
+\revpoint{3}{4}
 At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites (about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats.
@@
         \textbf{(bottom)} old and new lakes.
-        Each plot shows $F_{ST}$ values for a separate simulation,
-        with columns corresponding to increasing gene flow from left to right.
+        The columns are ordered by the $M$ (number of migrants per lake, per generation) 
+        and show $F_{ST}$ values for a separate simulation,
         $F_{ST}$ is calculated between all marine individuals, and all lakes pooled together.
@@
 we expect very little contribution of subsequent migration across any value of $m$, as seen in Figure~\ref{fig:Origin}.
-It remains unknown how different colonization schemes would impact the origin of adapted alleles, 
-but it stands to reason based upon our results that early arriving alleles are favored as adaptive alleles in any scenario. \llabel{disc_colon}
+Outside of our specific colonization scheme two other possible scenarios, and their respective impact on our results, are relevant to discuss. \llabel{disc_colon}
+In the case of a bottleneck and expanding population size post-colonization \emph{with} continued influx of migrant alleles,
+we would expect time to adaptation for $M = \{0.1, 1.0, 10.0\}$ to be greater, and the origin of adaptive alleles to 
+be weighted more towards ``migrant'' alleles. 
+In the case of a bottleneck and expanding population size post-colonization \emph{without} continued influx of migrant alleles,
+We would expect much less chance at selection on standing variants and most migration rates ($M = \{0.1, 1.0, 10.0\}$) 
+to adapt at the same pace as selection on ``de-novo'' alleles ($M=0.01$).
+
+%It remains unknown how different colonization schemes would impact the origin of adapted alleles, 
+%but it stands to reason based upon our results that early arriving alleles are favored as adaptive alleles in any scenario. \llabel{disc_colon}
 

### 2019-09-18 Scripts/time_to_adaptation.py
fixed plot

@@ -14,7 +14,8 @@ plt.rc('axes', labelsize=12)
 num = 1
 
 x = [0.00005,0.0005,0.005,0.05,0.5]
-x2 = ['0.01','0.1','1.0','10.0','100.0']
+#x2 = ['0.01','0.1','1.0','10.0','100.0']
+x2 = ['0.01','0.1','1.0','10.0']
 color_shape = ['go','b^','r*']
 
 timeToAdaptation = []
@@ -28,10 +29,10 @@ for j in [1]:
 
 	TTA = []
 	num = str(j)
-	recipes = ['9_'+num+'_0','9_'+num+'_1','9_'+num+'_2','9_'+num+'_3','9_'+num+'_4']
+	#recipes = ['9_'+num+'_0','9_'+num+'_1','9_'+num+'_2','9_'+num+'_3','9_'+num+'_4']
+	recipes = ['9_'+num+'_0','9_'+num+'_1','9_'+num+'_2','9_'+num+'_3']
 	
 	for i in recipes:
-
 		File = open("./Output1/MyRecipe"+i+"/Adaptation.txt","r")
 		header = File.readline()
 		FWAA_P2 = File.readline()
@@ -44,7 +45,7 @@ for j in [1]:
 		numShared_DivTotal.append(float(numShared_ )/float(totalNumFWAA_))
 
 		corFWAA.append(File.readline().split()[0])
-		corEffect.append(File.readline().split()[0])
+		#corEffect.append(File.readline().split()[0])
 
 	timeToAdaptation.append(TTA)
 
@@ -52,10 +53,11 @@ timeToAdaptation = np.array(timeToAdaptation)
 means = np.mean(timeToAdaptation,axis=0)
 
 f1, ax1 = plt.subplots()
-for k in range(len(timeToAdaptation)):
-	ax1.plot(x2[:-1],timeToAdaptation[k][:-1],color_shape[k])
-ax1.plot(x2[:-1],means[:-1],"y--")
-ax1.plot(x2[-1],10,"")
+#for k in range(len(timeToAdaptation)):
+#ax1.plot(x2[:-1],timeToAdaptation[k][:-1],color_shape[k])
+ax1.plot(x2,timeToAdaptation[0],color_shape[0])
+ax1.plot(x2,means,"y--")
+#ax1.plot(x2[-1],10,"")
 
 ax1.grid(axis = 'y')
 ax1.set_ylabel('Time (Generations)')

### 2019-09-20 review-responses.tex
peter edits

--- 2019-09-17 review-responses.tex
@@
 We are writing to submit a revised version of our manuscript, 
-``A few stickleback suffice for the transport of alleles to new lakes'',
-XXX
+``A few stickleback suffice for the transport of alleles to new lakes''.
+Many thanks to the reviewers for thoughtful and helpful comments,
+to which we have responded in detail.
+We sympathize that additional simulations would be informative and interesting,
+but the amount of work that these would entail is beyond our capability at the moment.
+We appreciate being explicitly given the option to omit them.
+Otherwise, we think that the revisions have helped to clarify and generally improve the manuscript.
+We hope you agree.
 
@@
 \begin{quote}
-    Please submit your revision along with a response to the reviewers' concerns
-    and suggestions, which can be viewed at the bottom of this email. I agree
-    particularly with questions from two of the reviewers asking for
+    I agree particularly with questions from two of the reviewers asking for
     explanation or additional simulations on the founder population size, with
@@
     consuming, however, and want to make clear additional simulations are not a
-    requirement for acceptance (but would be rad if feasible). My lab and I
-    also spent some time debating what was meant by the explanation at the
+    requirement for acceptance (but would be rad if feasible). 
+\end{quote}
+
+Additional simulations exploring more of parameter space would be totally awesome, we agree.
+However, besides the constraints on our time noted above,
+we also note that it's not clear how much additional value it would have to delimit
+the region of paramter space on which our results hold:
+it would be more satisfying to develop theory that would apply to
+larger populations of varying sizes with fluctuating migration rates, for instance.
+
+\begin{point}{\revref}
+    My lab and I also spent some time debating what was meant by the explanation at the
     bottom of page 13 -- I think more effort to clarify here would be helpful.
+\end{point}
+
+\reply{
+    Thanks! We have rewritten this to hopefully be more precise. \revref
+}
+
+\begin{point}{}
     Finally, we also wondered why the simulations started at a trait value of 0
     instead of 10 -- it makes comparison of old vs new a bit harder than seems
-    necessary. I expect these revisions can be done within 30 days, but don't
-    hesitate to ask if you need more time. 
-\end{quote}
+    necessary. 
+\end{point}
+
+\reply{
+    We don't have a good reason. Apologies.
+}
 
@@
 
-\begin{point}{}
+Thanks for the nice words!
+We sympathize, but think that the number of simulations is sufficient
+for at least estimating the rough size of these effects -- 
+these effects are mostly monotonic in migration rate,
+so that for instance we have an estimate of the number of adaptive alleles at $M=1.0$
+but also know it is bracketed by the values at $M=0.1$ and $M=10$.
+Further simulations would only give us more precise estimates of the mean
+and an idea of the standard deviation.
+
+\begin{point}{\revref}
     \textit{``As expected, the genetic basis of the freshwater phenotype seems to simplify
@@
     There is one neat aspect of this result that is perhaps worth better
-    emphasizing (please correct me if I an : detectable Fst peaks at truly adaptive
+    emphasizing: detectable Fst peaks at truly adaptive
     alleles can only be very few. If we had replicate runs we could say how many
     true Fst peaks, on average, we see at M=1 and M=10, Many published studies
-    (particularly in marine animals) report dozens if not hundreds Fst outliers
+    (particularly in marine animals) report dozens if not hundreds of Fst outliers
     throughout the genome; your simulation here suggests that most of them have to
@@
 \reply{
+    We agree with the reviewer that this is an important point,
+    and that our results are consistant with the above hypothesis. 
+    However, to quantify and assert that previous empirical studies 
+    are reporting false-positives would require many 
+    more simulations and analysis in many more situations,
+    which we believe is outside the realm of this paper.
 }
@@
 \reply{
+    Actually, there are many SNPs that do not affect the phenotype,
+    but these were added as a secondary step after simulation,
+    so this is an easy mistake to make.
+    We've added a reminder about this. \revref
 }
 
-\begin{point}{}
+\begin{point}{\revref}
     \textit{``Surprisingly, for all cases with the exception of M = 0.01, we found that the
@@
 \reply{
+We agree that this is an important point to explore, and a good observation.
+We have appended to the section on migration to more clearly address this \llname{disc_colon}.  
+The available code could certainly be used in further studies to investigate this question more in-depth, but again feel that is outside the scope of this paper and have decidecided not to conduct more simulation and analysis.
 }
@@
 \reply{
+    This order of magnitude estimate is roughly based on the length of the coastline
+    (roughly 10,000 miles) and the high density of freshwater lakes in many flat, coastal areas.
+    We have reworded this to be more conservative, however. \revref
 }
@@
 \reply{
+    We're not sure what is meant here -- the equations in the Methods are almost entirely choices
+    we made in setting up the model:
+    for instance, the choice of a Gaussian fitness function is almost ubiquitous,
+    and the equation for the phenotype in terms of the selection and dominance coefficients
+    is just writing down the definition.
+    Most of these are options available in SLiM, which we cite.
+    We have included a reference for the (simple) defintion of $F_{ST}$ that we use \llname{Fst},
+    and added a clarifying comment to the section of the Results and Discussion
+    that is most equation-heavy \llname{Gil_cite}.
 }
@@
 \reply{
+    We agree, this would be very interesting, and an informative exercise.
+    However, we think it's outside the scope of this paper,
+    and hope the reviewer agrees with our discussion of this point at \revref.
 }
@@
 \reply{
+    We chose this to reflect our guess at what might be a general consensus for the mutational target size
+    of a polygenic trait (although we recognize that opinions differ).
+    We've added some more discussion to contextualize our choice to the paper \revref.
+    Some more on this: since per-bp mutation rates are likely closer to $10^{-8}$,
+    a rate of $10^{-10}$ represents roughly one adaptive target per 100 bases
+    across the $10^6$ bases where mutations that affect the phenotype might occur.
+    This translates to a total \emph{potential} target size of $10^4$ bp,
+    which seems generous but not impossible for a polygenic trait.
 }
@@
 \reply{
+    We know of no reason why in this case overlapping generation would affect the results.
 }
@@
 \reply{
+In a perfect world, we would re-run and compare. 
+However, to re-write this model as non-WF would require a relatively significant amount of change to the code as well as much more computational power to re-run the analysis.
+We know of no reason to expect differences between the two implementations --
+the main difference being in elegance --
+and so have opted to stick with the current model. 
+We have altered the wording a bit to better convey this \llname{realistic}.
 }
@@
 \reply{
+We agree, it should be made clear that the oceanic population is not panmictic but rather has one-dimential continuous spatial structure.
+We have clarified this here \llname{continuous}.
 }
@@
 \reply{
+We believe this finding supports existing studies, specifically, how stronger genetic drift with less migration leads to higher background $F_{ST}$.
+However, for our study, showing how greater sharing of adaptive alleles provides a shared signal across populations seemed necessary for a study of selection on standing variants.
+Thus, we have decided to leave this section as is.
+TODO: Looks like we could use some citations, I was unable to find any good ones :(.
 }
@@
 \reply{
+    We considered this sort of re-ordering when writing the paper,
+    but decided that it would be more confusing to jump back and forth
+    to the population genetics heuristics.
+    The equations are standard, and we've tried better to flag this at the start of the section.
+    \llname{Gil_cite}
 }
@@
 \reply{
+    It is a good point that although we found linkage between opposite-effect alleles,
+    this masking might not have a substantial effect.
+    We've hedged a bit more in this statement.
 }
@@
 \begin{point}{\revref}
-``Population size''.
-Here it is argued that population sizes of marine and freshwater populations are not so different.
-I found it hard to believe this in the light of quantitative data there is.
-Could it simply be so that marine population is actually very much larger, but structured (as there is genetic evidence for this in many papers) so that the fraction feeding the freshwater colonizations is to some degree isolated from ones feeding the freshwaters in a particular area?
-This would in effect make relevant portion of the marine population smaller, without need to postulate that the total marine population equals in size to freshwater populations, and that there are millions of lakes inhabited by three spine sticklebacks.
-Note also that in introduction it was claimed millions of freshwater populations, here you say ``if not millions'' - inconsistent.
+    ``Population size''. Here it is argued that population sizes of marine and
+    freshwater populations are not so different. I found it hard to believe this in
+    the light of quantitative data there is. Could it simply be so that marine
+    population is actually very much larger, but structured (as there is genetic
+    evidence for this in many papers) so that the fraction feeding the freshwater
+    colonizations is to some degree isolated from ones feeding the freshwaters in a
+    particular area? This would in effect make relevant portion of the marine
+    population smaller, without need to postulate that the total marine population
+    equals in size to freshwater populations, and that there are millions of lakes
+    inhabited by three spine sticklebacks. Note also that in introduction it was
+    claimed millions of freshwater populations, here you say ``if not millions'' -
+    inconsistent. 
 \end{point}
@@
 \reply{
+    We've made the two sections consistent,
+    and still heartily agree that quantifying this would be an interesting and useful endeavor,
+    but beyond the scope of the current paper.
+    The proposed substructure in the ocean doesn't seem very plausible to us,
+    although there is little enough data on oceanic demography that this may well be correct.
+    Still, is our argument that they face similar constraints on breeding ground not convincing?
 }
@@
 
+Thanks! We're glad you liked it.
+
 \begin{quote}
@@
 \end{quote}
+
+Sorry about that; we're submitting with line numbers this time.
 
@@
 \reply{
+Good point: we agree this would be one of the first additional parameters to vary.
+We have discussed this further \llname{disc_colon},
+but as we explained above, haven't done more simulations to expand coverage of parameter space.
 }
@@
 \reply{
+    We thought about this also, and did this comparison in some earlier simulations,
+    but any differential representation was not obvious
+    -- there was no large effect --
+    and that we'd need many more simulations to detect anything.
+    So: this seems plausible, but speculatively,
+    it doesn't seem to be a major driver of genetic architecture,
+    but rather a minor adjustment.
 }
@@
 \reply{
+    We agree this was confusing and have edited this section to hopefully make it more clear \revreffull{2}{8}. 
+    The motivation for 10 regions of effect was to roughly model the distribution of QTL seen in experiments,
+    and to allow us to contrast closely linked variants (within the same region)
+    with more distant ones.
 }
@@
 \reply{
+Good point, we have clarified a few ambiguous spots, e.g., \revref.
 }
@@
 \reply{
+    We have revised the figure caption and added the label $M$ to the header.
 }

### 2019-09-20 Stickleback_Paper.tex
fixed plot
peter edits

@@
 \citep{jones_genomic_2012, roesti_genome_2012, feulner_genomics_2015, Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. \revpoint{3}{10}
-However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of standing genetic variation, particularly during colonization and adaptation to new environments
+However, it is still not clear how variation in evolutionary processes -- such as gene flow, recombination, selection and mutation --  can promote the maintenance and re-use of standing genetic variation, particularly during colonization and adaptation to new environments
 \citep{Nelson_2018, Pritchard_2010, yeaman2011genetic, Schrider2017}. 
@@
 The ancestral marine form of threespine stickleback fish (\textit{Gasterosteus aculeatus}) \revpoint{2}{1}
-has given rise to hundreds of thousands if not millions of independently derived freshwater populations \revpoint{2}{3} \llabel{hundreds-thousands-or-more}
+has given rise to hundreds of thousands to millions of independently derived freshwater populations \revpoint{2}{3}
 in recently de-glaciated regions around the Northern Hemisphere 
@@
 To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait.
-The details and equations of the model were motivated by current understanding of threespine stickleback evolutionary, history and demography, as well as population genetics theory \citep{Gillespie_1998}.
-\llabel{clarify_equations}
+The details of the model were motivated by current understanding of threespine stickleback evolutionary, history and demography.
 The following model included divergent selection in the two habitats which had substantial spatial structure.
@@
 Our simulations include two habitat types -- marine and freshwater -- defined by the nature of their selective pressures, each with 5,000 diploid individuals. 
-The arrangement of these habitats, depicted in Figure~\ref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. 
+The arrangement of these habitats, depicted in \autoref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. 
 The marine habitat is a continuous, one-dimensional range of length 25 units, while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''), each connected to the marine habitat at regularly spaced intervals (positions $i - 1/2$ for $1 \le i \le 25$).
@@
 
-Each individual carries two linear chromosomes, each of size $10^8$ loci,
-with recombination events occurring at a intralocus rate of $10^{-8}$ every generation. \revpoint{2}{9}
-%Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation \revpoint{2}{8}
-%in ten regions of $10^5$ loci each, spread evenly along the chromosome. 
-Mutations that can affect the trait under selection can occur at rate $10^{-8}$ with a target size of $10$ Kb, (spread evenly among $1$ Mb) across each individual genome. \revpoint{2}{8}
-With 5000 (diploid) individuals per population, this equates to $\approx 1$ mutation, per population, per generation, which impact individual trait value. \llabel{mu_exp}
+Individuals are diploid, with a linear chromosome $10^8$ bp long,
+with recombination events occurring at a rate of $10^{-8}$ per bp per generation. \revpoint{2}{9}
+Mutations that can affect the trait under selection occur at rate $10^{-10}$ per locus per generation
+in ten regions of $10^5$ loci each, spread evenly along the chromosome.
+This translates to a total adaptive target size of $10^{-4}$ per individual per generation, \revpoint{2}{8}
+or $\approx 1$ trait-affecting mutation, per population of 5,000 diploids, per generation.
+This architecture was chosen to model a moderate number of clusters of trait-affecting genes,
+each of which having about 1Kb of potentially adaptive target
+(since mutation rates are closer to $10^{-8}$, a mutation rate of $10^{-10}$
+is roughly equivalent to having potentially adaptive sites at one of every 100 bases).
 Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
@@
 Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual so that approximately 200 offspring are generated in each lake, each generation.
-(A more realistic implementation would use SLiM's ``non-Wright-Fisher'' model type, \revpoint{2}{11}
+(A simpler implementation would use SLiM's ``non-Wright-Fisher'' model type, \revpoint{2}{11}
 which was not available when we wrote the simulations; but we do not expect the difference to affect results.) \llabel{realistic}
@@
 
-As depicted in Figure~\ref{fig:Geo} the marine population is continuous (unidimensional) while lakes are discrete. \llabel{continuous}
+As depicted in \autoref{fig:Geo} the marine population is continuous (one-dimensional) while lakes are discrete. \llabel{continuous}
 Dispersal occurs both locally along the coastline within the marine habitat, as well as between the marine habitat and the lakes,
@@
                     Analogous plots for other migration rates
-                    are shown in Figure \ref{fig:phenotype_time_supp}.
+                    are shown in \autoref{fig:phenotype_time_supp}.
         }
@@
 Local adaptation occurred in all simulations with the exception of the highest gene flow at which half of each population was composed of migrants, 
-as can be seen by the mean trait values of Figure~\ref{fig:MeanPhenotype}.
+as can be seen by the mean trait values of \autoref{fig:MeanPhenotype}.
 Figures~\ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat.
-Establishment of adaptive alleles in the old lakes during the first few thousand generations manifests as jumps in the mean trait value - which move the trait by an amount of order 1 every few hundred generations (Fig.
-\ref{fig:phenotype_time}).
+Establishment of adaptive alleles in the old lakes during the first few thousand generations manifests as jumps in the mean trait value -- which move the trait by an amount of order 1 every few hundred generations (\autoref{fig:phenotype_time}).
 \revpoint{3}{4}
 At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites (about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats.
-As expected, increasing migration rate decreased differentiation between habitats: as seen in Figure~\ref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.
+As expected, increasing migration rate decreased differentiation between habitats: as seen in \autoref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.
 \revpoint{1}{1}
@@
         \textbf{(bottom)} old and new lakes.
-        The columns are ordered by the $M$ (number of migrants per lake, per generation) 
+        The columns are ordered by $M$ (the number of migrants per lake, per generation) 
         and show $F_{ST}$ values for a separate simulation,
@@
 Adaptation occurred much more quickly at higher migration rates, both in the old and new sets of lakes.
-We measured this ``time to adaptation'' as the number of generations until average trait values in old and new lakes were within 0.5 of each other, shown in Figure~\ref{fig:TimeToAdaptation} for different rates of gene flow.
+We measured this ``time to adaptation'' as the number of generations until average trait values in old and new lakes were within 0.5 of each other, shown in \autoref{fig:TimeToAdaptation} for different rates of gene flow.
 Adaptation of new lakes took over 18,000 generations at the lowest rate of $M = 0.01$, while at $M = 1$ migrant per lake per generation, new lakes adapted in just under 60 generations.
@@
     \begin{center}
-          \includegraphics{Final_Plots/Time_Adapt.pdf}
+          \includegraphics{Final_Plots/Time_Adapt2.pdf}
           \caption{
@@
 (So, ``captured'' alleles as defined above are pre-existing, but not all pre-existing alleles are captured.)
-Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles across generations.
+\autoref{fig:MPFAI}A shows the distribution of the number of these alleles across generations.
 At $M = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere: new lakes independently acquire new adaptive alleles rather than pre-existing ones.
@@
 Concurrently, the average marine individual carries $\approx 2$ pre-existing freshwater adapted alleles, standing variation which was nearly absent at $M=0.01$.
-As migration rate increases past this, the total number of pre-existing freshwater adapted alleles declines (Figure \ref{fig:MPFAI}C).
+As migration rate increases past this, the total number of pre-existing freshwater adapted alleles declines (\autoref{fig:MPFAI}C).
 Interestingly, the frequency of these alleles in the ocean
@@
 
-Figure~\ref{fig:MPFAI}B shows the distribution, through time, of the mean \emph{percentage} of currently-defined freshwater adapted alleles that each genome in each of the populations carries. 
+\autoref{fig:MPFAI}B shows the distribution, through time, of the mean \emph{percentage} of currently-defined freshwater adapted alleles that each genome in each of the populations carries. 
 If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%.
@@
 Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes. 
-Figure \ref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above, the average individual across the new lakes has nearly the same amount of pre-existing freshwater adapted alleles as individuals across the old lakes.
+\autoref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above, the average individual across the new lakes has nearly the same amount of pre-existing freshwater adapted alleles as individuals across the old lakes.
 As expected, the genetic basis of the freshwater phenotype seems to simplify as migration increases -- higher rates of migration allow adaptive alleles of larger effect to travel more efficiently through the population, even though they are deleterious in the ocean.
@@
 
-The numbers in Figure~\ref{fig:MPFAI} strongly suggest that the dramatic increase in speed of local adaptation we observed above occurs because higher gene flow allows sharing of freshwater alleles among populations.
+The numbers in \autoref{fig:MPFAI} strongly suggest that the dramatic increase in speed of local adaptation we observed above occurs because higher gene flow allows sharing of freshwater alleles among populations.
 We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele in each individual in the new lakes, as defined in the Methods. 
-Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo''} mutation. 
+\autoref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo''} mutation. 
 As gene flow increases, a larger fraction of adaptive alleles derive from pre-existing variation in the marine population at the time of introduction. 
@@
 This level of gene flow only replaces 5\% of each population each generation with migrants from the other habitat, 
-but is sufficient to shift the mean trait values to nearly half their optimal values, as seen in Figure \ref{fig:MeanPhenotype}.
+but is sufficient to shift the mean trait values to nearly half their optimal values, as seen in \autoref{fig:MeanPhenotype}.
 
@@
 Can measures of local differentiation such as $F_{ST}$ be used to identify the causal loci? 
-Figure \ref{fig:Fst} shows plots along the genome of average per-locus $F_{ST}$ values in 500bp windows between the marine habitat and all freshwater habitats pooled together.
+\autoref{fig:Fst} shows plots along the genome of average per-locus $F_{ST}$ values in 500bp windows between the marine habitat and all freshwater habitats pooled together. 
+These $F_{ST}$ values are calculated using all polymorphic loci,
+including both those that affect phenotype as well as neutral variants. \revpoint{1}{2}
 Higher rates of dispersal showed more distinct $F_{ST}$ peaks over polymorphic loci, while ``background'' levels of $F_{ST}$ increase as gene flow decreases, swamping out this signal until the regions under selection are indistinguishable.
@@
 But how many of these peaks are actually underlying trait differences that form the basis for local adaptation? 
-To quantify this, Figure \ref{fig:Power_FP} shows the power and true positive rates that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.
+To quantify this, \autoref{fig:Power_FP} shows the power and true positive rates that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.
 
@@
 To answer this, we traced the genealogy of \textit{all} adaptive alleles in \textit{all} individuals from the introduced lakes, after local adaptation.
-Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back the original generation of inhabitants in the lake (Figure \ref{fig:Origin}).
+Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back the original generation of inhabitants in the lake (\autoref{fig:Origin}).
 \revpoint{1}{3}
@@
 
-If effectively capturing standing genetic variants is the key to rapid adaptation, as we have presented thus far, why is there a large difference in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$ seen in Figure \ref{fig:TimeToAdaptation}?
+If effectively capturing standing genetic variants is the key to rapid adaptation, as we have presented thus far, why is there a large difference in speed of adaptation between dispersal rates of $M = 0.1$ and $M = 1.0$ seen in \autoref{fig:TimeToAdaptation}?
 %Figure~\ref{fig:Origin} shows both migration levels being able to effectively capture and propagate the standing genetic variants, and
@@
 
-A possible explanation is that at lower rates of gene flow, freshwater alleles present as standing variation in the marine habitat are more tightly linked to marine alleles. 
+A possible explanation is that at lower rates of gene flow, freshwater alleles present as standing variation in the marine habitat are more tightly linked to marine alleles.  \revpoint{AE}{1}
 This implies that adaptation in the new lakes must wait for recombination to separate freshwater and marine adapted alleles.
-This might be expected at low dispersal rates since freshwater alleles may need to be ``masked'' by nearby compensatory alleles to remain in the ocean for long periods of time. 
-Reversing this ``masking'' then slows the process of adaptation that uses this genetic variation.
+A freshwater and a marine allele in close proximity
+could form an essentially neutral locus,
+which would persist in the marine population much longer than a freshwater allele
+not ``masked'' in this way.
+This suggests that a reasonable portion of freshwater alleles in the marine population
+would be linked to marine alleles.
+This genetic variation would be more difficult to access, however,
+so if there is sufficient additive genetic variation available to a new population in unlinked alleles,
+this will be used first.
+However, if the genetic variation initially available is not sufficient for a new population to adapt,
+then adaptation might need to wait for recombination to separate these haplotypes.
+Similar effects have been noted before in simulation \citep[e.g.,][]{yeaman2011genetic}.
 
-%\plr{cite \citet{yeaman2011genetic}?}
-
-In other words, higher dispersal from freshwater to the ocean maintains relatively intact freshwater haplotypes that can be more easily rebuilt in the marine environments.
+In other words, perhaps higher dispersal from freshwater to the ocean maintains relatively intact freshwater haplotypes that can be more easily rebuilt in the marine environments.
 Recall that trait-affecting mutations only occur in relatively small regions of $10^5$ loci in which recombination occurs only once in every one thousand meioses.
@@
 This might also explain why at $M=0.1$, marine individuals still hold on average several alleles that shift the trait in the freshwater direction
-(Figure \ref{fig:MPFAI}A).
+(\autoref{fig:MPFAI}A).
 %(and so likely to be at higher frequency in the marine population).
@@
 -- with the benefit of hindsight --
-turn out to describe qualitatively most of the aspects of adaptation we observed above \citep{Gillespie_1998}. \llabel{Gil_cite}
+turn out to describe qualitatively most of the aspects of adaptation we observed above.
 Because we model stabilizing selection on an additive trait controlled by a moderately large number of loci within each population, more precise expectations might be obtained through quantitative genetics \citep{svardal2014general} or even Fisher's geometric model \citep{barton2001hybridization,chevin2014niche},
-but doing so is beyond the scope of this paper
+but doing so is beyond the scope of this paper.
+The arguments we use here are standard in population genetics
+\citep[see e.g.,][]{Gillespie_1998}. \llabel{Gil_cite}
 
@@
 in practice we found that standing freshwater variation in the ocean
-was masked by linkage to compensatory marine variants. \revpoint{2}{15}
+might be masked by linkage to compensatory marine variants. \revpoint{2}{15}
 
@@
 Since in our model lower migration also reduces the amount of variation available in the ocean, 
-we expect very little contribution of subsequent migration across any value of $m$, as seen in Figure~\ref{fig:Origin}.
-Outside of our specific colonization scheme two other possible scenarios, and their respective impact on our results, are relevant to discuss. \llabel{disc_colon}
-In the case of a bottleneck and expanding population size post-colonization \emph{with} continued influx of migrant alleles,
-we would expect time to adaptation for $M = \{0.1, 1.0, 10.0\}$ to be greater, and the origin of adaptive alleles to 
-be weighted more towards ``migrant'' alleles. 
-In the case of a bottleneck and expanding population size post-colonization \emph{without} continued influx of migrant alleles,
-We would expect much less chance at selection on standing variants and most migration rates ($M = \{0.1, 1.0, 10.0\}$) 
-to adapt at the same pace as selection on ``de-novo'' alleles ($M=0.01$).
+we expect very little contribution of subsequent migration across any value of $m$, as seen in \autoref{fig:Origin}.
 
-%It remains unknown how different colonization schemes would impact the origin of adapted alleles, 
-%but it stands to reason based upon our results that early arriving alleles are favored as adaptive alleles in any scenario. \llabel{disc_colon}
+We expect at least two other aspects of colonization and migration to be relevant.
+First, the size of the initial population of marine individuals colonizing each new lake
+affects how much of the marine genetic diversity is present. \llabel{disc_colon}
+In our simulations, new lakes began with a full set of individuals,
+as may happen in uplifted tidal habitats.
+If instead each new lake began with only a few individuals (inducing an initial bottleneck),
+subsequent migration of alleles would likely become more relevant.
+If in addition, there was no subsequent ``upstream'' migration,
+adaptation might need to rely on new mutations, as at our lowest migration rates.
+However, exactly where these boundaries lie will require additional theoretical or simulation work.
 
@@
 The fact that repeated adaptation from standing genetic variation has been shown to occur in real world threespine stickleback populations argues that the population sizes of marine and combined freshwater stickleback are not that different.
-While this might at first seem implausible given the ocean is large and lakes are small, in reality there are hundreds of thousands if not millions of lakes inhabited by threespine stickleback in coastal regions of the Northern Hemisphere, and the limiting habitat in the ocean constraining marine stickleback population sizes is likely to be coastal nesting areas.
+While this might at first seem implausible given the ocean is large and lakes are small, \revpoint{2}{7}
+there are hundreds of thousands to millions of lakes inhabited by threespine stickleback in coastal regions of the Northern Hemisphere, and the limiting habitat in the ocean constraining marine stickleback population sizes is likely to be coastal nesting areas.
 Together, these facts make the equivalency of marine and freshwater stickleback population sizes much more plausible.
@@
         True positive rate is the proportion of SNPs above the $F_{ST}$ threshold that affect phenotype.
+        Line labels give the respective values of $M$.
           \label{fig:Power_FP}
@@
           \caption{ 
-        Statistical power and true positive rate as a function of $F_{ST}$ threshold. As in Figure \ref{fig:Power_FP}
+        Statistical power and true positive rate as a function of $F_{ST}$ threshold. As in \autoref{fig:Power_FP}
         but computed only using the first one-fifth of the total range -- i.e., fish in the first five lakes and those in the ocean with coordinate less than 5.0.
+        Line labels give the respective values of $M$.
          	}
@@
 This linkage prevents the ability for a sub-population to rapidly select on the beneficial standing genetic variation as recombination acts as a bottleneck to evolution in this situation. 
-Additionally, In Figure~\ref{fig:Origin} we found that the genetic basis for local adaptation, when selecting on standing genetic variation, is propagated largely from the initial generation of inhabitants.
+Additionally, In \autoref{fig:Origin} we found that the genetic basis for local adaptation, when selecting on standing genetic variation, is propagated largely from the initial generation of inhabitants.
 This contrasts the belief that subsequent migration is a necessity in newly established populations ability to rebuild a haplotype.
@@
 attaining a mean average trait value at the optimum (-10).
-It follows in Figure \ref{fig:phenotype_ts2}B that the majority time to adaptation is spent within 2 units of mean trait value.
+It follows in \autoref{fig:phenotype_ts2}B that the majority time to adaptation is spent within 2 units of mean trait value.
 The initial adaptation on selection of standing variation up to that point happens quite rapidly.

### 2019-09-26 Stickleback_Paper.tex
Bill's last edits

@@
 These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases the smaller freshwater populations have been established ``de novo'' from marine colonists.
-Independently evolved freshwater populations show similar phenotypes that have been shown to derive largely from the same standing genetic variants.
+Independently evolved freshwater populations exhibit similar phenotypes that have been shown to derive largely from the same standing genetic variants.
 Geographic isolation prevents direct migration between the freshwater populations, strongly suggesting that these shared locally adaptive alleles are transported through the marine population.
@@
 This process occurs over a realistic range of intermediate levels of gene flow that match previous empirical population genomic studies in stickleback. 
-Examining these simulations more deeply reveals how lower levels of gene flow leads to slow, independent adaptation to different habitats, whereas higher levels gene flow leads to significant mutation load -- but an increased probability of successful population genomic scans for locally adapted alleles.
+Examining these simulations more deeply reveals how lower levels of gene flow leads to slow, independent adaptation to different habitats, whereas higher levels of gene flow leads to significant mutation load -- but an increased probability of successful population genomic scans for locally adapted alleles.
 Surprisingly, we find that the genealogical origins of most freshwater adapted alleles can be traced back to the original generation of marine individuals that colonized the lakes, as opposed to subsequent migrants.
@@
 Concurrently, a growing number of empirical studies have identified instances of convergent evolution using standing genetic variation 
-\citep{jones_genomic_2012, roesti_genome_2012, feulner_genomics_2015, Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. \revpoint{3}{10}
-However, it is still not clear how variation in evolutionary processes -- such as gene flow, recombination, selection and mutation --  can promote the maintenance and re-use of standing genetic variation, particularly during colonization and adaptation to new environments
+\citep{jones_genomic_2012, roesti_genome_2012, feulner_genomics_2015, Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}.
+\revpoint{3}{10} However, it is still not clear how variation in evolutionary processes -- such as gene flow, recombination, selection and mutation --  can promote the maintenance and re-use of standing genetic variation, particularly during colonization and adaptation to new environments
 \citep{Nelson_2018, Pritchard_2010, yeaman2011genetic, Schrider2017}. 
@@
 The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years 
-\citep{Nelson_2018}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \citep{Terekhanova_2014, Lescak2015evolution, Bassham_2018}.
+\citep{Nelson_2018, Nelson_2019}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \citep{Terekhanova_2014, Lescak2015evolution, Bassham_2018}.
 For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska. 
@@
 Below, we often refer to these migration rates in terms of the number of migrants per lake, per generation, which we denote $M$. Since each lake contains 200 individuals, $M=200m$. 
-Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
+Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles among lakes, and the population genetic signals left behind.
 At very low rates of gene flow, each new lake's population adapted almost completely independently through \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations). 
@@
 
-\subsection*{Widespread allele sharing between lakes}
+\subsection*{Widespread allele sharing among lakes}
 
@@
 The numbers in \autoref{fig:MPFAI} strongly suggest that the dramatic increase in speed of local adaptation we observed above occurs because higher gene flow allows sharing of freshwater alleles among populations.
-We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele in each individual in the new lakes, as defined in the Methods. 
+We confirmed this finding by using recorded tree sequences to identify the origin of each trait-affecting allele in each individual in the new lakes, as defined in the Methods. 
 \autoref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo''} mutation. 
@@
 However, the \textit{origin} of the alleles underlying the phenotype is still unknown. 
-What is the nature of selection on standing genetic variants? 
 In our simulations migration from marine individuals into the new lakes continues throughout introduction and adaptation: but is consistent influx of standing genetic variation a necessity for rapid adaptation of the population?
 To answer this, we traced the genealogy of \textit{all} adaptive alleles in \textit{all} individuals from the introduced lakes, after local adaptation.
-Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back the original generation of inhabitants in the lake (\autoref{fig:Origin}).
+Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back to the original generation of inhabitants in the lake (\autoref{fig:Origin}).
 \revpoint{1}{3}
@@
 This implies that adaptation in the new lakes must wait for recombination to separate freshwater and marine adapted alleles.
-A freshwater and a marine allele in close proximity
+A freshwater and a marine allele (of roughly equal and opposite effects) in close proximity
 could form an essentially neutral locus,
@@
 Thus, the large marine population was able to harbor and distribute alleles that were deleterious in the ocean habitat, but adaptive in freshwater habitats, to surrounding freshwater populations with only a few migrants per population, per generation.
+A few good stickleback suffice.
 % Our simulation findings significantly extend previous empirical and theoretical population genomic studies of stickleback to make testable predictions, and provide a framework for similar studies in other organisms to determine the generality of our findings.

### 2019-09-27 review-responses.tex
added more refs

@@
 However, for our study, showing how greater sharing of adaptive alleles provides a shared signal across populations seemed necessary for a study of selection on standing variants.
-Thus, we have decided to leave this section as is.
-TODO: Looks like we could use some citations, I was unable to find any good ones :(.
+Thus, we have added some more citations to previous work,
+but otherwise have left this section as is.
 }
@@
 
-\begin{point}{}
+\begin{point}{\revref}
 Page 2, 2nd paragraph: you here talk about 'similar phenotypes' that have evolved repeatedly in freshwater, but you provide only two examples.
@@
 \reply{
-TODO: 
-\jgg{I feel like this may be brought up in a lot a papers, but is just an observation and not a finding worth citing super extensively? I could be wrong here}
+Good point; we have added citations to more examples. \revref
 }

### 2019-09-27 Stickleback_Paper.tex
added more refs

@@
 While geographic isolation often prevents direct migration between freshwater populations, threespine stickleback in them \revpoint{2}{5}
-frequently evolve similar phenotypes \citep{Cresko_2004, Colosimo_2004}. 
+frequently evolve similar phenotypes \citep{Cresko_2004, Colosimo_2004, stuart2017contrasting, hanson2016testing, hanson2017heritable, hirase2014parallel}. \revpoint{3}{11}
 The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years 
@@
 Can measures of local differentiation such as $F_{ST}$ be used to identify the causal loci? 
+This question has been extensively studied,
+both with simulation \citep[e.g.,][]{southcott2017neutral,lotterhos2015relative}
+and in the wild \citep[reviewed in][]{harrison2016heterogeneous,hoban2016finding};
+our simulations provide a detailed look where the truth is known.
 \autoref{fig:Fst} shows plots along the genome of average per-locus $F_{ST}$ values in 500bp windows between the marine habitat and all freshwater habitats pooled together. 

### 2019-11-15 Stickleback_Paper.tex
reviewer edits

@@
 stays relatively constant across the reasonable rates of gene flow.
+We explored, the possibility of dominance being a good predictor of how long an adaptive allele can be maintained in the 
+marine environment with no conclusive evidence, further modeling efforts are needed to say anything conclusive about dominance and allele maintenance in the marine.
 
@@
 \revpoint{1}{1}
+Because the detectable traits at high dispersal were very few, empirical studies in situation of high dispersal which report hundreds or thousands 
+of causal loci may benefit from exploring false positive rates in similar simulations runs.
 

### 2019-11-15 Stickleback_Paper.tex
revised edits

--- 2019-09-27 Stickleback_Paper.tex
@@
 stays relatively constant across the reasonable rates of gene flow.
+We explored the possibility of dominance being a good predictor of how long an adaptive allele can be maintained in the 
+marine environment with no conclusive evidence. 
+Further modeling efforts are needed to say anything conclusive about dominance and allele maintenance in the marine.
 
@@
 \revpoint{1}{1}
+Because the detectable traits at high dispersal were very few, empirical studies in the situation of high dispersal which report hundreds or thousands of causal loci may benefit from exploring false-positive rates in similar modeling studies.
 

### 2019-11-17 review-responses.tex
tweaks to letter
older changes

@@
 to which we have responded in detail.
-We sympathize that additional simulations would be informative and interesting,
-but the amount of work that these would entail is beyond our capability at the moment.
+We agree that additional simulations may well be informative and interesting,
+but the amount of work that these would entail is beyond our capability at the moment,
+and we think that the additional information added would probably be marginal.
 We appreciate being explicitly given the option to omit them.
@@
 We hope you agree.
+
+The responses to reviews includes page and line numbers that refer to the resubmitted manuscript;
+we are also uploading a color-coded pdf that shows differences from the previously submitted version.
 
@@
 }
-
-\plr{IMPORTANT: don't reorder or delete ``points'' below - it messes up the automatic numbering!}
 

### 2019-11-18 second-review-responses.tex
revew to responses again

%%%%%%
%%
%%  Don't reorder the reviewer points; that'll mess up the automatic referencing!
%%
%%%%%

\begin{minipage}[b]{2.5in}
  Resubmission Cover Letter \\
    {\it G3}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
    Jared Galloway, \\
    William Cresko, \\
    \emph{and} Peter Ralph \\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

Thanks very much for the prompt and positive decision.
We have added some additional discussion about the two reviewer points
(that previously we had only responded to in the Response to Reviewers),
as requested by the editor.
This was a good idea, as other readers will likely have the same thoughts.

We hope that you agree with our assessment.

\vspace{5em}

\noindent \hspace{4em}
\begin{minipage}{3in}
\noindent
{\bf Sincerely,}

\vskip 2em

{\bf 
Jared Galloway, Bill Cresko, and Peter Ralph.
}\\
\end{minipage}

\vskip 4em

\pagebreak
\setcounter{page}{1}

%%%%%%%%%%%%%%
\reviewersection{AE}

\begin{quote}
I would ask that you add some text to the discussion addressing reviewer points 1.1 and 3.2.
\end{quote}

Here is Reviewer 1 point 1:
\begin{quote}
    \textit{``As expected, the genetic basis of the freshwater phenotype seems to simplify
    as migration increases -- higher rates of migration allow adaptive alleles of
    larger effect to travel more efficiently through the population, even though
    they are deleterious in the ocean.''}
    Very cool. It gives a very intuitive guideline as to which genetic
    architecture of local adaptation to expect given a certain migration rate.
    There is one neat aspect of this result that is perhaps worth better
    emphasizing: detectable $F_{ST}$ peaks at truly adaptive
    alleles can only be very few. If we had replicate runs we could say how many
    true $F_{ST}$ peaks, on average, we see at $M=1$ and $M=10$. Many published studies
    (particularly in marine animals) report dozens if not hundreds of $F_{ST}$ outliers
    throughout the genome; your simulation here suggests that most of them have to
    be false positives of some kind, especially since migration in the sea is
    likely to be high.
\end{quote}

% our previous \reply{
%     We agree with the reviewer that this is an important point,
%     and that our results are consistant with the above hypothesis. 
%     However, to quantify and assert that previous empirical studies 
%     are reporting false-positives would require many 
%     more simulations and analysis in many more situations,
%     which we believe is outside the realm of this paper.
% }

We have added, at \llname{disco11}:

\textit{
We find that, unsurprisingly, sharing of alleles across lakes leads to
a very simple genetic basis -- only tens of loci, at $M \ge 1$.
This suggests that all else being equal, systems with higher migration
should have fewer loci responsible for local adaptation.
However, ``all else is equal'' is rarely plausible:
the actual number of locally adaptive alleles in a particular system
depends on many other factors, including
the strength of selection, degree of pleiotropy, and the genetic basis of the trait.
A different genetic basis could well lead to a much larger number of alleles, even at high migration rates:
but, could these be reliably identified?
Next, we look at how well adaptive alleles are identified with genome scans.
}

% \textit{
% Because the detectable traits at high dispersal were very few,
% empirical studies in the situation of high dispersal which report
% hundreds or thousands of causal loci may benefit from exploring
% false-positive rates in similar modeling studies.
% }

\bigskip

Here is Reviewer point 3.2:
\begin{quote}
If possible, it would be interesting to check what type of mutations - in terms of effect size and in terms of dominance - fall under the 'captured', 'marine' and 'de novo' categories. I could well imagine, for example, that mutations that can stay in the marine populations relatively long are more likely to be recessive. Is this true?
\end{quote}

% our previous \reply{
%     We thought about this also, and did this comparison in some earlier simulations,
%     but any differential representation was not obvious
%     -- there was no large effect --
%     and that we'd need many more simulations to detect anything.
%     So: this seems plausible, but speculatively,
%     it doesn't seem to be a major driver of genetic architecture,
%     but rather a minor adjustment.
% }

We have added, at \llname{disco32}:

\textit{
We explored the possibility of dominance being a good predictor of
how long an adaptive allele can be maintained in the marine environment,
but found no effect strong enough to be seen with our relatively small sample size of adaptive alleles.
}

### 2019-11-18 Stickleback_Paper.tex
revew to responses again

--- 2019-09-27 Stickleback_Paper.tex
@@
 % \reviewresponsesfalse  % don't include them
+\newcommand{\responsefile}{second-review-responses.tex}  % name of the review reponses file
 \input{review-response-commands}
-\newcommand{\responsefile}{review-responses.tex}  % name of the review reponses file
 
@@
 As expected, increasing migration rate decreased differentiation between habitats: as seen in \autoref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.
-\revpoint{1}{1}
 
@@
 stays relatively constant across the reasonable rates of gene flow.
+We explored the possibility of dominance being a good predictor of
+how long an adaptive allele can be maintained in the marine environment,
+but found no effect strong enough to be seen with our relatively small sample size of adaptive alleles.
+\llabel{disco32}
 
@@
 \revpoint{1}{1}
+% Because the detectable traits at high dispersal were very few, empirical studies in the situation of high dispersal which report hundreds or thousands of causal loci may benefit from exploring false-positive rates in similar modeling studies.
+\llabel{disco11}
 
@@
 
+We find that, unsurprisingly, sharing of alleles across lakes leads to
+a very simple genetic basis -- only tens of loci, at $M \ge 1$.
+This suggests that all else being equal, systems with higher migration
+should have fewer loci responsible for local adaptation.
+However, ``all else is equal'' is rarely plausible:
+the actual number of locally adaptive alleles in a particular system
+depends on many other factors, including
+the strength of selection, degree of pleiotropy, and the genetic basis of the trait. \llabel{disco11}
+A different genetic basis could well lead to a much larger number of alleles, even at high migration rates:
+but, could these be reliably identified?
+Next, we look at how well adaptive alleles are identified with genome scans.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 It's important to note that $F_{ST}$ in Figures~\ref{fig:Fst} and \ref{fig:Power_FP} was calculated after pooling together all lakes and all marine individuals.
-For comparison, Figures~\ref{fig:fst_5lakes} \& \ref{fig:Power_FP_5lakes} shows $F_{ST}$ calculated with individuals only in the first fifth of the range (5 lakes and corresponding marine habitat).
+For comparison, Figures~ \ref{fig:Power_FP_5lakes} and~\ref{fig:fst_5lakes} show $F_{ST}$ calculated with individuals only in the first fifth of the range (5 lakes and corresponding marine habitat).
 These show a substantially lower true positive rate, as population-specific noise swamps any signal between the lakes.

### 2019-11-20 second-review-responses.tex
resubmitted

@@
 This was a good idea, as other readers will likely have the same thoughts.
+One of the points turned out to be somewhat tricky to discuss in an accurate way,
+since it involved speculating beyond the bounds of our study,
+but we've got something in that we're satisified with.
 
-We hope that you agree with our assessment.
+We hope that you agree with us.
+
+In addition to the revised manuscript (Stickleback\_Paper\_text.pdf),
+we have uploaded a version with the changes we have made highlighted (Stickleback\_Paper\_diff.pdf).
 

### 2019-12-17 Stickleback_Paper.tex
added acknowledgements
ack fix-up
fixes

@@
 
-\onehalfspacing
-
 \newcommand{\E}{\mathbb{E}}
@@
 \newif\ifreviewresponses
-\reviewresponsestrue  % include them
-% \reviewresponsesfalse  % don't include them
+% \reviewresponsestrue  % include them
+\reviewresponsesfalse  % don't include them
 \newcommand{\responsefile}{second-review-responses.tex}  % name of the review reponses file
@@
 in recently de-glaciated regions around the Northern Hemisphere 
-\citep{Bell_1994, Thompson_1997, Cresko_2007, Hunt_2008} \revpoint{3}{9}
+\citep{Bell_1994, Thompson_1997, Cresko_2007, Hunt_2008}. \revpoint{3}{9}
 This model organism has provided some of the earliest data showing the heterogeneous nature of divergence across genomes \revpoint{2}{4}
@@
     \exp\left\{
-        \frac{1}{2}
+        - \frac{1}{2}
             \left(
@@
           \caption{
-        Average $F_{ST}$ in windows of 500Mb between:
+        Average $F_{ST}$ in windows of 500bp between:
         \textbf{(top)} marine habitat and old lakes;
@@
         \textbf{(red)} new mutations,
-        \textbf{(blue)} post-colonization migrants,
-        \textbf{(green)} ``captured'' from pre-existing lakes, or
-        \textbf{(orange)} standing marine variation.
+        \textbf{(black)} post-colonization migrants,
+        \textbf{(tan)} ``captured'' from pre-existing lakes, or
+        \textbf{(grey)} standing marine variation.
         See Methods for precise definitions of these categories.
@@
 As we calculate fitness here, this is
-$s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$,
+$s(u) = \exp(-\beta(z - u)^2) / \exp( - \beta z^2) \approx 2 \beta z u$,
 where $\beta = 1 / 450$. 
@@
 % Our simulation findings significantly extend previous empirical and theoretical population genomic studies of stickleback to make testable predictions, and provide a framework for similar studies in other organisms to determine the generality of our findings.
+
+\section *{Acknowledgements}
+
+We'd like to thank the entire Cresko lab, as well as Jeff Ross-Ibarra and several anonymous reviewers for useful input.
+Thank you to Ben Haller for suggesting SLiM model improvements and interpretation. 
+This work benefited from access to the University of Oregon high performance computer, Talapas.
 
